### 2014-10-01 stat304/hw1.tex
First homework set

\documentclass[11pt]{article}
\usepackage[paper=letterpaper]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2014-10-11 stat343/hw1/hw1.tex
Adding folder for stats 343

\documentclass[11pt]{article}
\usepackage[paper=letterpaper]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx}
\DeclareGraphicsExtensions{.jpg}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
Homework 1 \hfill Aaron Maurer
\hrule
\begin{enumerate}
    \item[1.]
        \begin{enumerate}
            \item[a)]
                \textbf{Task:} Prove that $RSS = SYY - \frac{(SXY)^2}{SXX}$
                \begin{align*}
                    RSS &= \sum_{i=1}^n \left [y_i - (\hat{\beta}_0 + \hat{\beta}_1 x_i) \right ]^2
                    \\ &= \sum_{i=1}^n \left [y_i - (\bar{y} - \hat{\beta}_1 \bar{x} + \hat{\beta}_1 x_i) \right ]^2 
                    \\ &= \sum_{i=1}^n \left [y_i - \bar{y} + \frac{SXY}{SXX} \bar{x} - \frac{SXY}{SXX} x_i \right ]^2
                    \\ &= \sum_{i=1}^n \left [(y_i - \bar{y}) - \frac{SXY}{SXX} (x_i - \bar{x}) \right ]^2
                    \\ &= \sum_{i=1}^n \left [(y_i - \bar{y})^2 - 2 \frac{SXY}{SXX}(y_i - \bar{y})(x_i - \bar{x}) + \left(\frac{SXY}{SXX}\right)^2 (x_i - \bar{x})^2 \right ]
                    \\ &= \sum_{i=1}^n (y_i - \bar{y})^2 - 2 \frac{SXY}{SXX}\sum_{i=1}^n (y_i - \bar{y})(x_i - \bar{x}) + \left(\frac{SXY}{SXX}\right)^2 \sum_{i=1}^n (x_i - \bar{x})^2
                    \\ &= SYY - 2\frac{SXY}{SXX}SXY + \left(\frac{SXY}{SXX}\right)^2SXX
                    \\ &= SYY - \frac{SXY^2}{SXX}
                \end{align*}
            \item[b)]
                \textbf{Task:} Use Cauchy-Schwarz to prove that RSS is nonnegative.
                \begin{align*}
                    RSS &= SYY - \frac{SXY^2}{SXX}
                    \\ &=  \|y-\bar{y}\|^2 - \frac{\langle x - \bar{x}, y - \bar{y} \rangle^2}{\|x-\bar{x}\|^2}
                    \\ &\geq \|y-\bar{y}\|^2 - \frac{\|x-\bar{x}\|^2 \|y-\bar{y}\|^2}{\|x-\bar{x}\|^2}
                    \\ &\geq \|y-\bar{y}\|^2 - \|y-\bar{y}\|^2
                    \\ &\geq 0
                \end{align*}
        \end{enumerate}
    \item[2.]
        \textbf{Task:} Show that linear regression and k-nearest neighbor regression are both of the class $\hat{f}(x_0)=\sum_{i=1}^N l_i(x_0;\mathcal{X})y_i$. \\
        Linear regression is a model of the form: \\
        \begin{align*}
            \hat{f}(x_0) &= \hat{\beta}_0 + \hat{\beta}_1 x_0
            \\ &= \bar{y} - \hat{\beta}_1 \bar{x} + \hat{\beta}_1 x_0
            \\ &= \bar{y} + (x_0 - \bar{x})\frac{SXY}{SXX}
            \\ &= \bar{y} + (x_0 - \bar{x})\frac{\sum_{i=1}^N (x_i - \bar{x})y_i}{SXX}
            \\ &= \bar{y} + \sum_{i=1}^N \frac{(x_0 - \bar{x})(x_i - \bar{x})}{SXX}y_i
            \\ &= \sum_{i=1}^N \left [ \frac{(x_0 - \bar{x})(x_i - \bar{x})}{SXX}y_i + \frac{\bar{y}}{N} \right ]
            \\ &= \sum_{i=1}^N \left [ \frac{(x_0 - \bar{x})(x_i - \bar{x})}{SXX} + \frac{1}{N} \right ]y_i
        \end{align*}
        Accordingly, it is of the desired form for 
        \[ l_i(x_0;\mathcal{X})=\frac{(x_0 - \bar{x})(x_i - \bar{x})}{SXX} + \frac{1}{N} \]
        Nearest neighbors regression is trivially of that form, with $N_{k,i}$ being the set of $k$ $x_j$ closest to $x_0$:
        \[ l_i(x_0;\mathcal{X})=\frac{I_{N_{k,x_0}}(x_i)}{k} \]
    \item[3.]
        Please note that all work is in the R supplement
        \begin{enumerate}
            \item[2.1.1]
                \\ \includegraphics[width=8cm]{hw1_2_1_1} \\ 
                I would argue that this data, though seeming to indicate some sort of correlation bet ween height and weight befitting a linear model, is too small considering the amount of noise for a linear regression to make sense. The estimates would be far too noisy to read anything into it. That said, if there was more data it looks like a linear model would probably make sense.
            \item[2.1.2]
                See the R output for the calculation and confirmation of most of the values. 
                \[ \hat{\beta}_1=.582, \hat{\beta}_0=-36.876, \]
                \\ \includegraphics[width=8cm]{hw1_2_1_2} \\ 
            \item[2.1.3]
            \[\sigma^2=71.502, se(\hat{\beta}_1)=.389, se(\hat{\beta}_0)=64.473, Cov(\beta_0,\beta_1)=-25.070 \] 
            \[t_{\beta_1} = 1.496, t_{\beta_0}=-94.752, p_{\beta_1} = .173, p_{\beta_0} = 1.718e-13 \]
            \item[2.1.4]
                \begin{center}
                    \begin{tabular}{r | r r r r r}
                        Regression & df      & ss      & ms      & F     & pval \\
                        \hline
                        Regression & 1       & 159.947 & 159.947 & 2.237 & \approx 0 \\
                        Residual   & 8       & 572.014 & 71.502  &       &  \\
                        Total      & 731.961 &         &         &       & 
                    \end{tabular}
                \end{center}
                t_{\beta_1}^2 = 2.239 = F
        \end{enumerate}

    \item[4.]
        Please note hat all work is in the R supplement
        \begin{enumerate}
            \item[b)]
                While certain variables seem to be approximately normally distributed, such as pH, others seem to be very distinctly not normally distributed, such as alcohol. 
                \\ \includegraphics[width=8cm]{hw1_4b_pH} \\ 
                \\ \includegraphics[width=8cm]{hw1_4b_alcohol} \\
                We also see some variables that clearly have a strong correlation, such as density and fixed.acidity, while with others like alcohol and citric.acid, if there is a correlation is hardly obvious.
                \\ \includegraphics[width=8cm]{hw1_4den_to_acid} \\ 
                \\ \includegraphics[width=8cm]{hw1_4alc_to_cit} \\
                Also, quality is clearly a categorical, rather than truly continuous.
            \item[c)]
                \\ \includegraphics[width=8cm]{hw1_4c} \\
                Most of these pairwise relationships at least seem to indicate some kind of correlation based on eyeballing the plot. pH and density look the closest to being independent, though there may be a very weak negative correlation. citric.acid and density look like they at best have a weak positive correlation. Its hard to tell if the correlations are linear, but they look to be at worst a little bit curved, such as in the case of fixed.acidity and pH, which can be modeled with a linear model
                \\ \includegraphics[width=8cm]{hw1_4c_fixed_acidity} \\
                \\ \includegraphics[width=8cm]{hw1_4c_pH} \\
                \\ \includegraphics[width=8cm]{hw1_4c_citric_acid} \\
                \\ \includegraphics[width=8cm]{hw1_4c_density} \\
                However, from comparing the quantile plots with a vector of normals, it looks like possibly fixed.acidity and definitely citric.acid are not normally distributed. This means are standard error estimates will be off with a linear regression.
            \item[d)]
                Both models seem appropriate based on the usuals statistical tests. For the pH regression, we have an $R^2$ of $.2937$, with the F-test for the model and T-tests for the $\beta$s essentially yielding zero (smaller than the smallest non-zero number r can represent). Similarly, we get an $R^2$ of $.4463$ and p-values of essentially 0 for the density regression as well. However, looking at the residuals plots, it is a little less clear. 
                \\ \includegraphics[width=8cm]{hw1_4d_pHresiduals} \\
                I would say the residuals for the pH model look fine, they look symmetrically distributed around 0 throughout the range of fitted values, indicating there isn't bias and that the model is alright.
                \\ \includegraphics[width=8cm]{hw1_4d_denresiduals} \\
                However, for the density model this doesn't quite seem to be true. We seem to get a negative bias on the tails of the range of fitted values and a positive bias in the middle. This indicates that the linear model, as is, isn't appropriate.

            \item[e)]
                \\ \includegraphics[width=8cm]{hw1_4e_phfit} \\
                \\ \includegraphics[width=8cm]{hw1_4e_denfit} \\
            \item[f)]
                It is clear for the pH model that the linear model (diamonds) is a better fit than the nearest neighbor model (circles), since its global  assumption of linearity appears to hold. The nearest neighbor model is essentially bouncing around the linear estimate, indicating that it merely has higher variance without capturing the underlying relationship better (it doesn't seem to have less bias). \\

                On the other hand, the nearest neighbor model does seem to be doing a better job on the density model. Since it is a local model that introduces less bias, it has picked up on the not quite linear (without transforming fixed.acidity anyway) relationship. The linear regression, since its global and based on a strong assumption, has introduced bias since the assumption failed.
            \item[g)]
                Considering that citric.acid is an acid, and pH is a measure for acidity, with lower values indicating higher acidity, it would stand to reason that a decrease in citric.acid would cause lower pH, as we have seen. \\
                Fixed acids seem to, in general, have much higher density than alcohol and water, which are the two main ingredients in wine. Thus, fixed acids increase the density of wine, and all else equal, one would expect higher concentrations of them would correspond to higher density.
        \end{enumerate}

\end{enumerate}

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-10-13 stat304/hw1.tex
Homework 1 done

@@
 \documentclass[11pt]{article}
-\usepackage[paper=letterpaper]{geometry}
+\usepackage[paper=letterpaper, margin=.5in]{geometry}
 \pdfpagewidth 8.5in
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Homework 1 \hfill Aaron Maurer
+\vspace{2mm}
+\hrule
+\vspace{2mm}
 
+\begin{itemize}
+    \item[1.1.] 
+        \begin{itemize}
+            \item[DF 3]
+                Suppose, to the contrary, $\lim_{x \to -\infty} F(x) \neq 0$. \\
+                $\Rightarrow \exists \epsilon>0: \lim_{x \to -\infty} F(x) = \epsilon$, since $F$ is bounded below by 0 and nondecreasing. \\
+                $\Rightarrow$ For any sequence $x_n: \lim_{n \to \infty}x_n = -\infty$, $\lim_{n \to \infty} F(x_n)=\epsilon$. \\
+                Let $x_n$ be such a sequence, with the additional restriction that it is strictly decreasing. \\
+                $\Rightarrow \forall n, F(x_n)>\epsilon$, since $F$ is nondecreasing. \\
+                Let $A_n = {X\leq x_n}$ and $A = \inter{n}A_n$. Since $x_n$ is strictly decreasing, $\forall n>m, A_n \subset A_m$. \\
+                $\Rightarrow P[A_n] \downarrow P[A]$, since $A_n \downarrow A$. \\
+                $\Rightarrow \exists a \in A$, since $P[A]\downarrow \epsilon$ and $P[\emptyset]=0$. \\
+                $\Rightarrow \forall n, a\leq x_n$. \\
+                This is a contradiction, since $\lim_{n \to \infty}x_n = -\infty$. Thus, we must conclude that $\lim_{x \to -\infty} F(x) = 0$ 
+                \smallskip
+
+                Suppose, to the contrary, $\lim_{x \to \infty} F(x) \neq 1$. \\
+                $\Rightarrow \exists \epsilon<1: \lim_{x \to \infty} F(x) = \epsilon$, since $F$ is bounded above by 1 and nondecreasing. \\
+                $\Rightarrow$ For any sequence $x_n: \lim_{n \to \infty}x_n = \infty$, $\lim_{n \to \infty} F(x_n)=\epsilon$. \\
+                Let $x_n$ be such a sequence, with the additional restriction that it is strictly increasing. \\
+                $\Rightarrow \forall n, F(x_n)<\epsilon$, since $F$ is nondecreasing. \\
+                Let $A_n = {X\leq x_n}$ and $A = \un{n}A_n$. Since $x_n$ is strictly increasing, $\forall n>m, A_m \subset A_n$. \\
+                $\Rightarrow P[A_n] \uparrow P[A]$, since $A_n \uparrow A$. \\
+                $\Rightarrow P[A^c]>0$, since $P[A]\uparrow \epsilon$ and $\epsilon<1$. \\
+                $\Rightarrow \exists a \in A^c$, since $P[\emptyset]=0$. \\
+                $\Rightarrow \forall n, a> x_n$. \\
+                This is a contradiction, since $\lim_{n \to \infty}x_n = \infty$. Thus, we must conclude that $\lim_{x \to \infty} F(x) = 1$ \\
+            \item[DF 4]
+                $F$ is bounded above and nondecreasing, so we know that $F(x-)$ exists. Let $x_n$ be a strictly increasing sequence such that $\lim_{n \to \infty} x_n = x$. 
+                Let $A_n = {X\leq x_n}$ and $A = \un{n} A_n$. It is clearly the case that $A = {X<x}$, and since $A_n\uparrow A$, $P[A_n]\uparrow P[A]$. Since $P[A_n]=F(x_n)$, we may 
+                conclude that $\lim_{n \to \infty} F(x_n) = P[X<x]$. Thus, we can conclude $F(x-)=P[X<x]$ as well.               
+            \item[DF 5]
+                \begin{align*}
+                    F(x) - F(x-) &= P[X\leq x] - P[X<x] \\
+                                 &= P[X=x]
+                \end{align*}
+        \end{itemize}
+    \item[1.2.]
+        Let 
+        \[F(x)= 
+            \begin{cases} 
+                0,           &\mbox{if } x<0          \\
+                \frac{1}{2}, &\mbox{if } 0\leq x < 2  \\
+                1,           &\mbox{if } 2\leq x
+            \end{cases}
+        \]
+        Here, $F^*(\frac{1}{2})=0<1$, yet $\frac{1}{2}=F(1)$. As well, $\frac{1}{4}<\frac{1}{2}=F(0)$, yet $F^*(\frac{1}{4})=0$.
+    \item[1.3.]
+        \(\forall v\in (0,1): \lim_{u\downarrow 0} F^*(u) \leq F^*(v)\), since $F^*$ nondecreasing. \\
+        \(\Rightarrow \forall v\in (0,1): \forall x \in \{x\in\R: v\leq F(x)\}: \lim_{u\downarrow 0} F^*(u) \leq x$, by the definition of $F^*\) \\
+        \(\Rightarrow \forall x \in \{x\in \R: F(x)>0\}: \lim_{u\downarrow 0} F^*(u) \leq x\) \\
+        Also, \(\nexists w: \forall v\in (0,1): \lim_{u\downarrow 0} F^*(u) < w \leq F^*(v)\). \\
+        \(\Rightarrow \nexists w: \forall v\in (0,1): \forall x \in \{x\in \R: v\leq F(x)\}: \lim_{u\downarrow 0} F^*(u) < w \leq F(x) \) \\
+        \(\Rightarrow \nexists w: \forall x \in \{x\in \R: F(x)>0\}: \lim_{u\downarrow 0} F^*(u) < w \leq F(x) \) \\
+        Accordingly, since \(\lim_{u\downarrow 0} F^*(u)\) is the greatest upper bound of \(\{x\in \R: F(x)>0\}\), \\
+        \[\lim_{u\downarrow 0} F^*(u) = \inf\{x\in \R: F(x)>0\}\]
+        \smallskip
+
+        Let \(x\in\R: F(x)\in(0,1)\). \\
+        \(\Rightarrow \exists v\in(0,1): F(x)<v<1 \) \\
+        \(\Rightarrow F^*(v)>x \) by the switching formula. \\
+        \(\Rightarrow \lim_{u\uparrow 1} F^*(u) \geq F^*(v) > x\), since \(F^*\) is nondecreasing.\\
+        \(\Rightarrow \lim_{u\uparrow 1} F^*(u)\) is an upper bound of \(\{x\in\R:F(x)<1\}\) \\
+        Suppose, to the contrary, \(\exists w: \forall x \in \{x\in\R: F(x)<1\}: x\leq w < \lim_{u\uparrow 1} F^*(u)\). \\
+        \(\Rightarrow \forall v\in(0,1): F^*(v)\leq w < \lim_{u\uparrow 1} F^*(u)\). \\
+        \(\Rightarrow \lim_{u\uparrow 1} F^*(u) \leq w < \lim_{u\uparrow 1} F^*(u)\) \\
+        Since the above is a contradiction, we must conclude that \(\lim_{u\uparrow 1} F^*(u)\) is the least upper bound of \(\{x\in\R:F(x)<1\}\), and therefor \\
+        \[\lim_{u\uparrow 0} F^*(u) = \sup\{x\in \R: F(x)>0\}\]
+    \item[1.4.]
+        \(F(x)\leq F(x)\) \\
+        \(\Rightarrow F^*(F(x))\leq x \) by the switching formula. \\
+        Let \(\delta>0\). \\
+        \(\Rightarrow F(x) < F(x) + \delta \) \\
+        \(\Rightarrow x < F^*(F(x+\delta)) \) by the switching formula. \\
+        \(\Rightarrow x \leq F^*(F(x)+) \) \\
+        Accordingly,
+        \[F^*(F(x))\leq x \leq F^*(F(x)+) \]
+    \item[1.5.]
+        We can show that
+        \begin{align*}
+            F^*(u) &\leq F^*(u)         & \\
+                u  &\leq F(F^*(u))      & \mbox{by the switching formula }  \\
+            F^*(u) &\leq F^*(F(F^*(u))) & \mbox{since } F^* \mbox{ is nondecreasing}             
+        \end{align*}    
+        Now suppose, to the contrary, that \(F^*(u)<F^*(F(F^*(u)))\). \\
+        \(\Rightarrow F(F^*(u))<F(F^*(u)) \) by the switching formula. \\
+        Since this is an obvious contradiction, we may conclude that
+        \[F^*(u) = F^*(F(F^*(u))) \]
+        \smallskip
+
+        We can also show that
+        \begin{align*}
+            F(x) &\leq     F(x)     & \\
+              x  &\leq F^*(F(x))    & \mbox{by the switching formula} \\
+            F(x) &\leq F(F^*(F(x))) & \mbox{since } F \mbox{ is nondecreasing}
+        \end{align*}    
+        If \(F(x) < F(F^*(F(x)))\), this would imply by the switching formula that \(F^*(F(x))>F^*(F(x))\), which an obvious contradiction. Thus, we can conclude that
+        \[F(x) = F(F^*(F(x)))\]
+    \item[1.6.]
+        Let $F$ not be continious at $x$.  \\
+        \(\Rightarrow F(x-)\neq F(x)\), since \(F(x)=F(x)\) by DF2. \\
+        \(\Rightarrow F(x-)< F(x)\) since $F$ is nondecreasing. \\
+        \(\Rightarrow \exists \epsilon>0: \forall \delta>0: \exists x_1\in(x-\delta,x): F(x)-F(x_1)>\epsilon \) \\
+        \(\Rightarrow \forall w<x, F(x)-\epsilon>F(w)\), since F is nondecreasing, and \(\exists x_1\in(w,x): F(w)\leq F(x_1)< F(x)-\epsilon\) \\
+        \(\Rightarrow \forall u \in (F(x)-\epsilon,F(x)): F^*(u)\geq x \), since $x$ is a lower bound of \(\{x\in\R:u\leq F(x)\}\). \\
+        \(\Rightarrow F(F^*(u)) \geq F(x) > u \), since F is nondecreasing. \\
+        \(\Rightarrow\) It is not the case that \(\forall u \in(0,1): F(F^*(u))=u \)  \\
+        Now, going the other direction, let $F$ be continious, and, to the contrary, let \(F(F^*(u))>u\) for some \(u\in(0,1)\) \\
+        \(\Rightarrow \forall x\in\R: F(x-)= F(x) \). \\
+        \(\Rightarrow \forall \epsilon>0: \exists \delta>0: \forall x_1\in (x-\delta,x), F(x)-F(x_1)<\delta \) \\
+        \(\Rightarrow \exists \delta>0: \forall w \in (F^*(u)-\delta,F^*(u)): F(F^*(u))-F(w)<F(F^*(u))-u \) \\
+        \(\Rightarrow F(w)>u \) and \(F^*(u)>w\) \\
+        However, by definition, \(F^*(u)\) is less than or equal to all $x$ such that \(F(x)\geq u\), so we have reached a contradiction. Thus, we may conclude that if $F$ is continious,
+        it must also be the case that \(F(F^*(u))\leq u \).
+We can reject that \(F(F^*(u))< u \), since that would mean \(F^*(u)<F^*(u) \) by the switching formula, so we are left with \(F(F^*(u))= u \).
+\\
+        Taking both direction together, we get the result 
+        \[\forall u\in(0,1): F(F^*(u))=u \Leftrightarrow F \mbox{is continious} \]
+        \smallskip
+
+        Let $F$ be strictly increasing. \\
+        \( \Rightarrow\) if \(x_2<x_1\), then \(F(x_2)<F(x_1)\).\\
+        \( \Rightarrow x_2<F^*(F(x_1)) \) by the switching formula. \\
+        \( \Rightarrow x_2<F^*(F(x_1))\leq x_1 \) by IDF4. \\
+        \( \Rightarrow \forall \epsilon>0: x_1-\epsilon<F^*(F(x_1))\leq x_1 \). \\
+        \( \Rightarrow x_1 = F^*(F(x_1)) \). \\
+        Now, let $F$ not be strictly increasing. \\
+        \( \Rightarrow \exists x_1,x_2 \in \R: x_2<x_1 \mbox{ and } F(x_2)\geq F(x_1) \) \\
+        \( \Rightarrow F(x_2)= F(x_1) \), since $F$ is nondecreasing. \\
+        \( \Rightarrow F^*(F(x_1))\leq x_2 < x_1 \) \\
+        \( \Rightarrow \) It is not the case that \(\forall x\in\{x\in\R:0<F(x)<1\}: F^*(F(x))=x\). \\
+        Taking both directions together, we get the result that
+        \[ \forall x\in A:=\{x\in\R:0<F(x)<1\}: F^*(F(x))=x\ \Leftrightarrow F \mbox{ is strictly increasing over } A \]
+    \item[1.7.]
+        To acheive a desired $p$-value, it is sufficient that, for a test statistic $x$, \(F(x)\leq p\).
+    \item[1.8.]
+        \begin{itemize}
+            \item[a)]
+                Let \(u\geq F(x-)\) \\
+                \( \Leftrightarrow \forall \delta_1>0: u\geq F(x-\delta_1) \), since $F$ is nondecreasing. \\
+                \( \Leftrightarrow \forall \delta_1,\delta_2>0: u + \delta_2 > F(x-\delta_1) \) \\
+                \( \Leftrightarrow \forall \delta_1,\delta_2>0: F^*(u + \delta_2) > x-\delta_1 \) by the switching formula. \\
+                \( \Leftrightarrow \forall \delta_2>0: F^*(u + \delta_2) \geq x \) \\
+                \( \Leftrightarrow F^*(u+) \geq x \) \\
+            \item[b)]
+                Let \(A = \{w:u\geq F(w-)\}\) \\
+                \( \Rightarrow A = \{w:F^*(u+)\geq w\} \) by part a) \\
+                \( \Rightarrow \forall x \in A: F^*(u+)\geq x \) \\
+                \( \Rightarrow F^*(u+)\geq \sup(A) \) \\
+                Now suppose, to the contrary, that \(F^*(u+)> \sup(A)\) \\
+                \( \Rightarrow \exists x\in \R : F^*(u+)>x>\sup(A) \) \\
+                \( \Rightarrow x\in A \mbox{ and } x>\sup A \) \\
+                This is a contradiction, so we must conclude that it cannot be the case that \(\sup A < F^*(u+)\), and rather it must be that \(\sup A = F^*(u+)\).
+        \end{itemize}
+    \item[1.9.]
+        \begin{itemize}
+            \item[a)]
+                Let $A$ be a closed and bounded supbinterval of $B$.\\
+                \( \Rightarrow f(A)\subseteq [f(\min(A)),f(\max(A))] \), since $f$ nondecreasing. \\
+                \( \Rightarrow \forall \epsilon>0:\) Let \(J_{A,\epsilon}=\{x\in A:f(x+)-f(x-)>\epsilon\} \) \\
+                \( \Rightarrow \vert J_{A,\epsilon}\vert\in \N \), since, due to $f$ being nondecreasing, \\
+                \[ \sum_{j\in J_{A,\epsilon}} \epsilon \leq \sum_{j\in J_{A,\epsilon}} f(j+)-f(j-) \leq f(\max(A)) - f(\min(A)) \]
+                Now, let, where \(K=\{z\in\Z : B\cap [z-1,z] \neq \emptyset\}\)  \\
+                \[ J=\bigcup_{z\in K} \un{n} J_{[z-1,z],\frac{1}{n}} \]
+                \( \Rightarrow D_f\subseteq J\), since if \(j\in D_f\), \(\exists z\in \Z: j\in[z-1,z] \) and \(\exists n\in \N_0: f(j+)-f(j-)>\frac{1}{n} \). \\
+                Let \(h:\Z \times \N_0 \times \N_0 \to J\) be defined as such: The triple \((z,n_1,n_2)\) gets mapped to the $n_2$th largest $j$ in \(J_{[z-1,z],\frac{1}{n_1}}\).
+                We may choose the $n_2$th largest since we have already shown \(J_{[z-1,z],\frac{1}{n_1}}\) is finite. This is clearly a surjective function, where 
+                \(\forall j\in J: \exists (z,n_1,n_2)\in \Z \times \N_0 \times \N_0 : h(z,n_1,n_2) = j \). Therefor, since such a mapping exists, we may conclude that
+                \[\left\vert \Z \times \N_0 \times \N_0 \right\vert \geq \left\vert J \right\vert \geq \left\vert D_f \right\vert \]
+                Since \(\Z \times \N_0 \times \N_0\) is countable, this means that $D_f$ is also at most countable.
+            \item[b)]
+                Let \(x\in B \mbox{ and } \epsilon>0\). \( (x-\epsilon,x+\epsilon) \) is an uncountable set, so there can't possibly be a surjective function \(h:D_f \to (x-\epsilon,x_\epsilon)\), 
+                since $D_f$ is at most countable. Thus, it must be the case that \(\exists c\in(x-\epsilon,x_\epsilon): c\in D_f^c=C_f\). We can thus conclude $C_f$ is dense in $B$.
+        \end{itemize}
+    \item[1.10.]
+        \begin{itemize}
+            \item[a)]
+                Let $w$ be a continuity point of $F$ and \(u\in(0,1)\) with \(F^*(u)>w \) \\
+                \( \Rightarrow u>F(w) \) by the switching formula \\
+                Let \(\epsilon = u - F(w)\).
+                \( \Rightarrow \exists m : \forall m\geq n: \vert F_n(w) - F(w) \vert < \epsilon \), since \(\lim_{n\to\infty} F_n(w)=F(w) \) \\
+                \( \Rightarrow \forall n \geq m \)
+                \begin{align*}
+                    u &= F(w) + (u - F(w) \\
+                      &> F(w) + \vert F_n(w) - F(w) \vert \\
+                      &> F(w) + F_n(w) - F(w) \\
+                      &> F_n(w)
+                \end{align*}
+            \item[b)]
+                Let $y$ be a continuity point of $F$ and \(u\in(0,1)\) with \(F^*(u+)<y \) \\
+                \(\Rightarrow \exists \delta>0: \forall \epsilon\in (0,\delta): F^*(u+\epsilon)<y \) \\
+                \(\Rightarrow u+\epsilon \leq F(y) \) by the switching formula \\
+                \(\Rightarrow u \leq F(y) - \epsilon \) \\
+                Since \(\lim_{n\to\infty} F_n(y)=F(y), \exists m: \forall n>m: \vert F_n(y) - F(y) \vert < \epsilon \) \\
+                \(\Rightarrow \forall n>m \)
+                \begin{align*}
+                        u &\leq F(y) - \epsilon \\
+                          &<    F(y) - \vert F_n(y) - F(y) \vert \\
+                          &<    F(y) - (F(y) -  F_n(y)) \\
+                          &<    F_n(y)                  \\
+                    F^*(u)&<      y                            & \mbox{By the switching formula}
+                \end{align*}
+            \item[c)]
+                Suppose, to the contrary, \(F^*(u) > \lim \inf_n F_n^*(u) \) \\
+                \( \Rightarrow \exists c\in C_f: F^*(u) > c > \liminf_n F_n^*(u) \), since $C_f$ is dense in $\R$. \\
+                \( \Rightarrow \exists m \in \N: \forall n>m: F_n^*(u)>c \), by part a. \\
+                \( \Rightarrow \forall n>m, \inf_n F_n^*(u) \geq c \) \\
+                \( \Rightarrow \liminf_n F_n^*(u) \geq c \) \\
+                This is a contradiction, so we must conclude \(\liminf_n F_n^*(u)\geq F^*(u)\)
+                \smallskip
+
+                \( \inf_n F^*(u) \leq \sup_n F^*(u) \) by definition. \\
+                \( \Rightarrow \liminf_n F_n^*(u)\leq \limsup_n F_n^*(u) \)
+                \smallskip
+
+                Suppose, to the contrary, \(\limsup_n F_n^*(u) > F^*(u+) \) \\
+                \( \Rightarrow \exists c \in C_f: \limsup_n F_n^*(u) > c > F^*(u+) \), since $C_f$ is dense in $\R$a \\ 
+                \( \Rightarrow \exists m \in \N: \forall n>m: F_n^*(u)\leq c \), by part b. \\
+                \( \Rightarrow \forall n>m, \sup_n F_n^*(u) \leq c \) \\
+                \( \Rightarrow \limsup_n F_n^*(u) \leq c \) \\
+                This is a contradiction, so we must conclude \(\limsup_n F_n^*(u)\leq F^*(u+)\)
+                \smallskip
+
+                Combining the three above results, we have that 
+                \[  F^*(u) \leq \liminf_n F_n^*(u) \leq \limsup_n F_n^*(u) \leq F^*(u+) \]
+            \item[d)]
+                If \(u\in C_f\), by definition \(F^*(u) = F^*(u+)\) \\
+                \( \Rightarrow F^*(u) = \liminf_n F_n^*(u) = \limsup_n F_n^*(u) \) by part c. \\
+                \( \Rightarrow \forall \epsilon>0: \exists m\in\N: \forall n>m: \) \\
+                \[ F^*(u) - \epsilon < \inf_n F_n^*(u) \leq \sup_n F_n^*(u) < F^*(u) + \epsilon \]
+                \( \Rightarrow \forall n>m : F^*(u) - \epsilon < F_n^*(u) < F^*(u) + \epsilon \) \\
+                \( \Rightarrow \lim_{n\to\infty} F_n^*(u) = F^*(u) \)
+            \item[e)]
+                Let
+                \[F_n^*(u) = (2u-1)^{-n} \mbox{ and } F^*(u) = 
+                    \begin{cases}
+                        -1, &\mbox{if } 0<x\leq\frac{1}{2} \\
+                         1, &\mbox{if } \frac{1}{2}<x\leq1 
+                    \end{cases}
+                \]
+                arising from 
+                \[F_n(u) =
+                    \begin{cases}
+                        0,               &\mbox{ if } x<-1        \\
+                        \frac{x^n+1}{2}, &\mbox{ if } -1\leq x< 1 \\
+                        1,               &\mbox{ if } 1 \leq x
+                    \end{cases}
+                \mbox{ and } F(u) =
+                    \begin{cases}
+                        0,               &\mbox{ if } x<-1        \\
+                        \frac{1}{2},     &\mbox{ if } -1\leq x< 1 \\
+                        1,               &\mbox{ if } 1 \leq x
+                    \end{cases}
+                \]
+                \( \lim_{n\to\infty} F_n^*(u) = F^*(u) \) for all continuity points of $F^*$. However, where $F^*$ isn't continious at $\frac{1}{2}$, 
+                \(\lim_{n\to\infty} F_n^*(\frac{1}{2}) = 0 \neq -1 = F^*(\frac{1}{2})\).
+        \end{itemize}
+    \item[2.1.]
+        Let $F(x)$ not be strictly increasing over \(A=\{x\in\R: 0<F(x)<1\}\). \\
+        \(\Rightarrow \exists x_1,x_2\in A: x_1<x_2 \mbox{ and } F(x_1)=F(x_2)\) \\
+        \(\Rightarrow F^*(F(x_1))\leq x_1, \mbox{ since } \inf\{x:F(x_1)\leq F(x)\} \leq x_2 \) \\
+        Let \(B = \{x\in A: F(x)=F(x_1)\} \) \\
+        \(\Rightarrow \forall x\in \{ x\in A: F(x)>F(x_1)\}, x\geq \sup B \), since the fact that $F$ is nondecreasing makes $x$ and upper bound on $B$ \\
+        \(\Rightarrow \forall v> F(x_1), F^*(v)\geq \sup B \) \\
+        \(\Rightarrow \forall \delta > 0 \)
+        \begin{align*}
+            F^*(F(x)+\delta)             &\geq \sup B -F^*(F(x_1)+F^*(F(x_1) \\
+            F^*(F(x)+\delta) -F^*(F(x_1)) &\geq \sup B -F^*(F(x_1) \\
+                                         &\geq x_2 - x_1
+        \end{align*}
+        \(\Rightarrow F^*(u+)\neq F^*(u)\)
+        \smallskip
+
+        Now, let $F$ be strictly increasing over $A$ and $x\in A$ such that \(x=F^*(u)\) and \(F(x)<1\). Let \(y\in(x,1)\). \\
+        \( \Rightarrow \exists w\in(x,y) \) \\
+        \( \Rightarrow u\leq F(x) < F(w) < F(y) \) since $F$ is strictly increasing.\\
+        \( \Rightarrow u+F(w)-F(x) \leq F(w) \) \\
+        \( \Rightarrow F^*(u+F(w)-F(x)) \leq w < y \) by the switching formula. \\
+        \( \Rightarrow F^*(u+) < y \) \\
+        \( \Rightarrow F^*(u+) = x = F^*(u) \)
+        \smallskip
+
+        Between these two results, we get that $F^*$ is continious if and only if $F$ is strictly increasing.
+    \item[2.2.]
+        \begin{itemize}
+            \item[a)]
+                \begin{align*}
+                    P[T(Y)\in B] &= P[Y \in {x: T(x)\in B}] \\
+                                 &= P[Z \in {x: T(x)\in B}] \\
+                                 &= P[T(Z) \in B]
+                \end{align*}
+                \( \Rightarrow Y~Z \)
+            \item[b)]
+                Let \(P[Y=Z]=1\)
+                \begin{align*}
+                    P[Y\in B] &= P[Y=Z \cap Y\in B] \\
+                              &= P[Y=Z \cap Z\in B] \\
+                              &= P[Z\in B]
+                \end{align*}
+                \( \Rightarrow Y~Z \)
+        \end{itemize}
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2014-10-14 stat343/hw2.R
Done with homework 2

@@ -1,0 +1,28 @@
+wine<-read.csv(file="hw1/winequalityRed.csv",head=TRUE,sep=";")
+lm<-lm(pH ~ fixed.acidity, data=wine)
+
+# Generate confidence intervals
+confint(lm,level=.66)
+confint(lm,level=.9)
+confint(lm,level=.95)
+
+# Generate variance estimate
+c<-c(1,10)
+n<-nrow(wine)
+cov<-summary(lm)$cov.unscaled
+sig2 <- summary(lm)$sigma^2
+varpred<- sig2 * t(c) %*% cov %*% c 
+varpred
+
+# Make prediction
+predict(lm,newdata=data.frame(fixed.acidity=13.1), interval="confidence")
+predict(lm,newdata=data.frame(fixed.acidity=13.1), interval="prediction")
+
+# Make and plot 100 predictions
+points<-data.frame(fixed.acidity =seq(min(wine$fixed.acidity),max(wine$fixed.acidity),length.out=100))
+predictions<-predict(lm,newdata=points, interval="confidence")
+pdf('hw2_plot.pdf')
+plot(points$fixed.acidity,predictions[,1],cex=.5,main="CI of Mean pH Estimate", xlab="Fixed Acidity", ylab="pH")
+points(points$fixed.acidity,predictions[,2],cex=.5, pch=18)
+points(points$fixed.acidity,predictions[,3],cex=.5, pch=18)
+dev.off()

### 2014-10-16 stat374/hw1.tex
Done with homework 1 for 374

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
HW1 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[a)]
                I've only included the graphs here, the commands used to generate them can be found in the R supplement in the back. For all of these, $B$, the number of trials for a particular $N$, was set to $100$. \\
                \includegraphics[width=9cm]{hw1_1_a_reg} 
                \includegraphics[width=9cm]{hw1_1_a_log} 
            \item[b)]
                I chose $N$ to be $10$, $50$, and $100$. For each of these, the density plot is again off of $B=100$ trials.
                \includegraphics[width=9cm]{hw1_1_b_10} 
                \includegraphics[width=9cm]{hw1_1_b_50} \\
                \includegraphics[width=9cm]{hw1_1_b_100} 
        \end{itemize}
    \item[2.]
        \begin{itemize}
            \item[a)]
                For $\sigma=1$
                \smallskip

                The risk of the MLE estimator is just the variance. Thus,
                \[ R(Z,\theta) = \sum_{i=1}^{1000} \sigma^2 = \sum_{i=1}^{1000} 1 = 1000 \]

                The risk of a shrinkage estimator will be:
                \[ R(\hat\theta,\theta) = b^2n\sigma^2 + (1-b)^2 \| \theta \|^2 = b^2 1000 + (1-b)^2 \sum_{i=1}^{1000} \frac{1}{i^4} = b^2 1000 + (1-b)^2*1.082323 \]
                Which is plotted here: \\
                \includegraphics[width=9cm]{hw1_2_plot_1} \\
                As the graph suggests, the minimum is achieved at
                \[b_*=\frac{\|\theta \|^2}{n\sigma^2+\|\theta \|^2} = \frac{1.082323}{1+1.082323} = .0010812 \]
                The James-Stein estimator seems to be a good approximation of this; it appears to be centered at the optimal $b$ (the mean is higher because negative values are excluded), with some random variance: \\
                \includegraphics[width=9cm]{hw1_2_sim_1} \\

                Setting $c^2$ to $1.082323$, we get a Pinsker Bound of \(\frac{1*1.082323}{1+1.082323}=.52976\). The MLE risk blows right by this by a factor of $2000$, partly because \(\sigma_n^2\neq \frac{\sigma^2}{n} \). The simulated mean squared error of the JS estimator is, as it should be, always coming in much below this threshold" \\
                \includegraphics[width=9cm]{hw1_2_risk_1} \\

            \item[b)]
                For \(\sigma= \frac{1}{n}\)
                \smallskip

                The risk of the MLE estimator is just the variance. Thus,
                \[ R(Z,\theta) = \sum_{i=1}^{1000} \sigma^2 = \sum_{i=1}^{1000} \frac{1}{1000} = 1 \]

                The risk of a shrinkage estimator will be:
                \[ R(\hat\theta,\theta) = b^2n\sigma^2 + (1-b)^2 \| \theta \|^2 = b^2 + (1-b)^2 \sum_{i=1}^{1000} \frac{1}{i^4} = b^2 + (1-b)^2*1.082323 \]
                Which is plotted here: \\
                \includegraphics[width=9cm]{hw1_2_plot_n} \\
                As the graph suggests, the minimum is achieved at
                \[b_*=\frac{\|\theta \|^2}{n\sigma^2+\|\theta \|^2} = \frac{1.082323}{1+1.082323} = .5197671 \]
                The James-Stein estimator seems to be a good approximation of this; it appears to be centered at the optimal $b$, with some random variance: \\
                \includegraphics[width=9cm]{hw1_2_sim_n} \\
                With the same Pinkser Bound, with the variance assumption actually holding, the MLE risk is still about twice as high at 1. The simulated mean squared error of the JS estimator comes in far below, as one would expect. \\
                \includegraphics[width=9cm]{hw1_2_risk_n} \\

        \end{itemize}
    \item[3.]
        After some investigation, it is my belief that the actual $\theta$ vector was generated by a Pareto distribution, or a distribution that is quite similar to it. Choosing a shape parameter of 2 and a minimum of 2.4 for the Pareto distribution, I certainly seem to be very close: \\
        \includegraphics[width=9cm]{hw1_3_den} 
        \includegraphics[width=9cm]{hw1_3_qq}  \\
        This is about as good a fit as you will see with a parametric model. Its hard to judge the fit of the tail, but even if it is off, since the signal overpowers the noise there, it doesn't matter that much for this exercise.
        \smallskip

        Armed with the assumption that \(\theta \sim Pareto(2,2.4)\), I can calculate the density function of $\theta$ conditioned on the particular $Z$:
        \begin{align*}
            f_\theta (\theta \,|\, Z) &= f_\theta (\theta \,|\, \theta + N(0,1) = z) \\
                                  &= \dfrac{ 2 \cdot 2.4^2 \theta^{-3}}{\int_{2.4}^\infty 2 \cdot 2.4^2 x^{-3} \phi(z-x) dx}
        \end{align*}
        Using this, I can calculate the expected mean square error for a particular estimate $\hat\theta_i$ conditioned on $Z_i$:
        \[ \E[(\theta_i-\hat\theta_i)^2 \,|\, Z_i] = \int_{2.4}^\infty (\theta_i-\hat\theta_i)^2 f_\theta(\theta \,|\, Z_i) d\theta \]
        Combining all of this, I chose the $\hat\theta_i$ which minimized this expected mean squared error:
        \[\hat\theta_i = \arg\!\min_b \E[(\theta_i-b)^2 \,|\, Z_i] \]
        This was all implemented numerically in R. The one caveat is that the numerical integral ran into precision issues in calculation when $Z$ was large. I assume the reason is the integral was summing over values smaller than can be represented in double precision. Thus, for $Z$ larger than $30$, I instead used \(\hat\theta_i = \frac{Z_i-1}{Z_i} \), which is what the James Stein estimate would be for that observation by itself. The end result is this: \\
        \includegraphics[width=9cm]{hw1_3_prediction} \\
        Values of $Z$ below $2.4$ all got $\hat\theta_i$ higher than $2.4$ (since this was the assumed minimum), and there was a general pull towards $3$ or so, with larger values of $Z$ getting shrunk less than smaller values greater than $3$.

    \item[4.]
        We will begin by calculating the risk for a given \(\bar X_n\) \\
        \begin{align*}
            \E\left[\left(f_\theta(x) - f_{\bar X_n}(x)\right)^2\,|\, \bar X_n\right] 
            &= \int_{-\infty}^\infty (f_\theta(x) - f_{\bar X_n}(x))^2 dx \\
            &=\int_{-\infty}^\infty f_\theta(x)^2 dx + \int_{-\infty}^\infty f_{\bar X_n}(x)^2 dx - 2\int_{-\infty}^\infty f_\theta(x) f_{\bar X_n}(x) dx \\
            &=2\int_{-\infty}^\infty \left( \frac{1}{\sqrt{2\pi}} e^{\frac{-(x-\mu)^2}{2}} \right )^2 dx - 2\int_{-\infty}^\infty \left( \frac{1}{\sqrt{2\pi}} e^{\frac{-(x-\theta)^2}{2}} \right) \left( \frac{1}{\sqrt{2\pi}} e^{\frac{(-(x-\bar X_n)^2}{2}} \right) dx \\
            &=\frac{1}{\sqrt{\pi}}\int_{-\infty}^\infty \frac{1}{\sqrt{\pi}} e^{-(x-\mu)^2} dx -  \frac{1}{\sqrt{\pi}}\int_{-\infty}^\infty \frac{1}{\sqrt{\pi}} e^{\frac{-(x-\theta)^2-(x-\bar X_n)^2}{2}}  dx \\
            &=\frac{1}{\sqrt{\pi}} - \frac{e^{-\frac{1}{4}(\theta-\bar X_n)^2}}{\sqrt{\pi}}\int_{-\infty}^\infty \frac{1}{\sqrt{\pi}} e^{-x^2+x\theta+x\bar X_n-\frac{\theta^2+X_n^2}{2}+\frac{\theta^2-2\theta\bar X_n+X_n^2}{4}}  dx \\
            &=\frac{1}{\sqrt{\pi}} - \frac{e^{-\frac{1}{4}(\theta-\bar X_n)^2}}{\sqrt{\pi}}\int_{-\infty}^\infty \frac{1}{\sqrt{\pi}} e^{-x^2+x\theta+x\bar X_n-\frac{\theta^2+2\theta\bar X_n+\bar X_n^2}{4}}  dx \\
            &=\frac{1}{\sqrt{\pi}} - \frac{e^{-\frac{1}{4}(\theta-\bar X_n)^2}}{\sqrt{\pi}}\int_{-\infty}^\infty \frac{1}{\sqrt{\pi}} e^{-(x-\frac{\theta+X_n}{2})^2}  dx \\
            &=\frac{1}{\sqrt{\pi}} - \frac{e^{-\frac{1}{4}(\theta-\bar X_n)^2}}{\sqrt{\pi}} \\
        \end{align*}
        Now, taking the expectation of this over all \(\bar X_n\) will give us the unconditional expectation of the risk.
        \begin{align*}
            \E\left[\left(f_\theta(x) - f_{\bar X_n}(x)\right)^2 \right] 
            &=\E\left[\E\left[\left(f_\theta(x) - f_{\bar X_n}(x)\right)^2\,|\, \bar X_n\right]\right] \\
            &= \int_{-\infty}^\infty \left( \frac{1}{\sqrt{\pi}} - \frac{e^{-\frac{1}{4}(\theta-\bar X_n)^2}}{\sqrt{\pi}} \right)  \left(\frac{\sqrt{n}}{\sqrt{2\pi}} e^{\frac{-n(\theta-\bar X_n)^2}{2}}\right) d\bar X_n  \\
            &=\frac{1}{\sqrt{\pi}} - \frac{1}{\sqrt{\pi}}\int_{-\infty}^\infty \frac{\sqrt{n}}{\sqrt{2\pi}} e^{\frac{-(\frac{1}{2}+n)(\theta-\bar X_n)^2}{2}} d\bar X_n \\
            &=\frac{1}{\sqrt{\pi}} - \frac{1}{\sqrt{\pi}}\sqrt{\frac{n}{\frac{1}{2}+n}}\int_{-\infty}^\infty \frac{\sqrt{\frac{1}{2}+n}}{\sqrt{2\pi}} e^{\frac{-(\frac{1}{2}+n)(\theta-\bar X_n)^2}{2}} d\bar X_n \\
            &=\frac{1}{\sqrt{\pi}}\left(1 - \sqrt{\frac{2n}{1+2n}} \right)\\
        \end{align*}
    \item[5.]
        Doppler function with $\sigma=.1$: \\
        \includegraphics[width=9cm]{hw1_5_p1_fit} 
        \includegraphics[width=9cm]{hw1_5_p1_gcv} \\
        Doppler function with $\sigma=1$: \\
        \includegraphics[width=9cm]{hw1_5_1_fit} 
        \includegraphics[width=9cm]{hw1_5_1_gcv} \\

    \item[6.]
        Per the definition of leave-one-out cross-validation, we have that
        \begin{align*}
            \hat R(h) &= \frac{1}{n} \sum_{i=1}^n \left(Y_i-\hat r_{(-i)}(x) \right )^2 \\
            \hat R(h) &= \frac{1}{n} \sum_{i=1}^n \left(Y_i-\sum_{j\neq i} Y_i l_{j,(-i)} (x)  \right )^2 \\
            \hat R(h) &= \frac{1}{n} \sum_{i=1}^n \left(Y_i- \frac{1}{\sum_{j\neq i} l_{j}(x)} \sum_{j\neq i} Y_i l_{j} (x)  \right )^2 \\
            \hat R(h) &= \frac{1}{n} \sum_{i=1}^n \left(Y_i- \frac{\hat r_n (x) - Y_i l_{i}(x)}{1- l_{i}(x)} \right )^2 \\
            \hat R(h) &= \frac{1}{n} \sum_{i=1}^n \left(\frac{Y_i - Y_i l_{i}(x) - \hat r_n (x) + Y_i l_{i}(x)}{1- l_{i}(x)} \right )^2 \\
            \hat R(h) &= \frac{1}{n} \sum_{i=1}^n \left(\frac{Y_i -\hat r_n (x) }{1- l_{i}(x)} \right )^2 \\
            \hat R(h) &= \frac{1}{n} \sum_{i=1}^n \left(\frac{Y_i -\hat r_n (x) }{1- L_{ii}} \right )^2 \\
        \end{align*}

\end{itemize}
\hrule
\vspace{2mm}
This is the R code I used for this homework:
\lstinputlisting{hw1.R}

\end{document}

### 2014-10-16 stat374/hw1.R

@@ -1,0 +1,211 @@
+library('parallel','locfit')
+
+# Option on which section of code to run
+
+############################################
+# Problem 1
+############################################
+
+#Part a
+m<-1
+s<-3
+b<-100
+nseq<-1:200
+uhat<-function(n) mean(rnorm(n,m,s))
+meansqerr<-function(n) mean((m-sapply(rep(n,times=b),uhat))^2)
+empmsqerr<-sapply(nseq, meansqerr)
+anamsqerr<-sapply(nseq, function(n) (s^2)/n)
+
+pdf('hw1_1_a_reg.pdf')
+plot(1:200,empmsqerr,main="Empirical vs Analytic Mean Squared Error", pch=1, cex=.5, xlab="N", ylab="Mean Squarred Error")
+points(1:200,anamsqerr,cex=.5, pch=18)
+legend('topright', legend=c('Empirical','Analytic'), pch=c(1,18))
+dev.off()
+
+pdf('hw1_1_a_log.pdf')
+plot(1:200,empmsqerr,main="Empirical vs Analytic Mean Squared Error", pch=1, cex=.5, xlab="Log N", ylab="Log Mean Squarred Error", log='xy')
+points(1:200,anamsqerr,cex=.5, pch=18)
+legend('topright', legend=c('Empirical','Analytic'), pch=c(1,18))
+dev.off()
+
+#Part b 
+nseq<-c(10,50,100)
+genZ<-function(n) sqrt(n)*(m-sapply(rep(n,times=b),uhat))
+for (n in nseq) {
+    pdf(paste('hw1_1_b_',toString(n),'.pdf', sep=""))
+    plot(density(genZ(n)), main="Empirical vs Analytic Density of Z", sub=paste('N=',toString(n),sep=""), lty=2 , xlab="Z", ylab="Density", xlim=c(-10,10), ylim=c(0,.25))
+    curve(dnorm(x,0,s), add=TRUE, lty=1)
+    legend('topright', legend=c('Empirical','Analytic'), lty=c(2,1))
+    dev.off()
+}
+
+
+############################################
+# Problem 2
+############################################
+
+### Sigma^2 = 1
+
+#plot the risk in terms of b
+pdf('hw1_2_plot_1.pdf')
[…]

### 2014-10-28 stat343/hw4.tex
Started homework 4 and brought hw3 back

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Homework 4 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[a)]
                \begin{align*}
                    \|Q\beta - y\|^2 &= (Q\beta - y)^T(Q\beta - y) \\                
                                     &= (\beta^T Q^T - y^T)(Q\beta - y) \\
                                     &= \beta^TQ^TQ\beta - y^TQ\beta - \beta^T Q^T y + y^Ty  \\
                                     &= (\beta^T\beta - y^TQ\beta - \beta^T Q^T y + y^TQQ^Ty) + (y^Ty - y^TQQ^Ty) \\
                                     &= (\beta^T - y^TQ)(\beta + Q^Ty) + (y^Ty - y^TQQ^Ty- y^TQQ^Ty + y^TQQ^Ty) \\
                                     &= \|\beta - Q^Ty\|^2 + (y^Ty - y^TQQ^Ty- y^TQQ^Ty + y^TQQ^TQQ^Ty) \\
                                     &= \|\beta - Q^Ty\|^2 + (y^T - y^TQQ^T)(y - QQ^Ty)\\
                                     &= \|\beta - Q^Ty\|^2 + \|y - QQ^Ty\|^2\\
                \end{align*}
            \item[b)]
                Since $Q^T$ is orthogonal,
                \[RSS(\beta) = \|Y-X\beta\|^2 = \|Q^TY-Q^TX\beta\|^2 = \|Q^TY - RP^T\beta\|^2 \]
                Using the result from a, we have that
                \[RSS(\beta) = \|QRP^T\beta - Y\|^2 = \|RP^T\beta-Q^TY\|^2 + \|Y-QQ^TY\|^2 \]
                Since \(\|y-QQ^Ty\|^2\) is fixed, the $RSS$ will be at a minimum when \(\|RP^T\beta-Q^TY\|^2\) is minimized. If \(R=\mathbf{0}\), then trivially all \(\beta\) achieve the same $RSS$. Otherwise, we want to minimize the non-trivial part of the expression. Where $R_{11}$ has $m$ rows, let $Y_1$ be the first $m$ rows of $Y$ and $Y_2$ be the last $n-m$ rows, and $Q_1$ the first $m$ rows of $Q$ with $Q_2$ the remiander. Thus, 
                \begin{align*}
                \|RP^T\beta-Q^TY\|^2 &= \left\| \left[\begin{array}{cc} R_{11} & R_{12} \\ \mathbf{0} & \mathbf{0} \end{array} \right] P^T \beta - [Q_1^T Q_2^T]\left[\begin{array}{c} Y_1 \\ Y_2 \end{array} \right] \right \|^2 \\
                    &= \| \left[R_{11} R_{12}\right] P^T \beta - Q_1^T Y_1 \|^2 + \|Q_2^T Y_2\|^2
                \end{align*}
                Since $\|Q_2^TY_2\|^2$ is also fixed, we can reduce our minimization problem to minimizing \(\| [R_{11} R_{12}] P^T \beta - Q_1^T Y_1 \|^2\). Let the first $m$ rows of $P^T\beta$ be $\beta_1$ and the remainder be $\beta_2$. $P^T\beta$ is just a permutation of $\beta$, so it is the same coefficients, just out of order. Thus, we will get a minimum when,
                \begin{align*}
                    \left[R_{11} R_{12}\right] P^T \beta &= Q_1^T Y_1 \\
                    \left[R_{11} R_{12}\right] \left[ \begin{array}{c} \beta_1 \\ \beta_2 \end{array} \right]  &= Q_1^T Y_1 \\
                    R_{11}\beta_1 + R_{12}\beta_2 &= Q_1^T Y_1 \\
                    R_{11}\beta_1  &= Q_1^T Y_1 - R_{12}\beta_2 \\
                    \beta_1  &= R_{11}^{-1}(Q_1^T Y_1 - R_{12}\beta_2) 
                \end{align*}
                Thus, when $\beta_1$ is as defined, we will have minimized $RSS$.
            \item[c)]
                We want to solve \(\left[R_{11} R_{12}\right] P^T \beta = Q_1^T Y_1\). Let \(W=\left[R_{11} R_{12}\right] P^T\). If we have some solution $\beta$, and a vector $a$ in the null space of \(W\), then 
                \[ W(\beta+a) = W \beta + W a = Q_1^T Y_1 + 0 \]
                So $\beta+a$ is also a solution. Let $\beta^*$ be the vector that acheives the smallest norm and solves the equation. Assuming $R$ isn't all $0$, then there is some $\beta$ perpendicular to all $a$. If $\beta_1$ is and another $\beta_2$ isn't, by Cauchy-Schwarz
                \begin{align*}
                    \|\beta_2+a\|^2 &\geq \|\beta_1 + a\|^2 \\
                    \|\beta_2\|^2 + \|a\|^2 &> \|\beta_1\|^2 + \|a\|^2 \\
                    \|\beta_2\|^2 &> \|\beta_1\|^2 \\
                \end{align*}
                Thus, $\beta^*$ must be perpendicular to all $a$. If $b$ is a vector, 
                \[0= b^T W a = (W^Tb)^T a \]
                So the range of \(W^T\) is perpendicular to all $a$ as well. Thus, since $\beta^*$ is perpendicular to the null space of \(W\), it must be in the range of \(W^T\) (This range fills the complement of the null space because taking the transpose of a matrix doesn't reduce its rank). Accordingly, for some $v$,
                \begin{align*}
                    \beta^* &= W^Tv \\
                    W\beta^* &= WW^Tv \\
                    Q_1^T Y_1 &= WW^Tv  \\
                    W^T(WW^T)^{-1} Q_1^T Y_1 &= W^T v \\
                    W^T(WW^T)^{-1} Q_1^T Y_1 &= \beta^*  \\
                \end{align*}
                Thus we have our answer. We can invert \(WW^T\) since \(W\) is full row rank (since \(R_{11}\) is invertible) and similarly \(W^T\) is full column rank.
        \end{itemize}

    \item{2.}
        \begin{itemize}
            \item{a)}
                The trace of $H$ is its rank, since its idempotent. 
        \end{itemize}
    \item{4.}
        \begin{itemize}
            \item{2.13.}
                \begin{itemize}
                    \item{1.}

                \end{itemize}
        \end{itemize}

\end{itemize}

\end{document}

### 2014-10-28 stat343/hw3.R

@@ -1,0 +1,94 @@
+library('faraway')
+
+#############################
+# Problem 1
+#############################
+
+# Load the data and create the matrices 
+data('savings')
+attach(savings)
+n<-nrow(savings)
+x<-as.matrix(data.frame(rep(1,n),savings[-1]))
+y<-as.matrix(savings[1])
+
+### Part c)
+# Loop over epsilons and try methods until failure
+eps <- .1
+norm.diff <- NULL
+continue <- TRUE
+while (continue) {
+    eps <- eps *.1
+    # Add pop75 + noise
+    x.i<-cbind(x,as.matrix(x[,3]+ eps*rnorm(n))) 
+
+    #See if the QR method still works
+    QR.success <-tryCatch(
+        {
+            q<-qr.Q(qr(x.i))
+            r<-qr.R(qr(x.i))
+            beta.qr <- solve(r, t(q) %*% y)
+            TRUE
+        },
+        error=function(e) return(FALSE),
+        warning=function(e) return(FALSE)
+        )
+
+    # See if the normal method still works
+    Normal.success <-tryCatch(
+        {
+            beta.x.i  <- solve(t(x.i)%*%x.i) %*% t(x.i) %*% y
+            TRUE
+        },
+        error=function(e) return(FALSE),
+        warning=function(e) return(FALSE)
+        )
+
+    # Keep going and calculate norm of difference if both succeed 
+    continue <- QR.success & Normal.success
+    if (continue) norm.diff <- c(norm.diff,norm(beta.qr-beta.x.i))
+}
+if (!QR.success) print(paste('The QR method failed at an epsilon of ',toString(eps)))
+if (!Normal.success) print(paste('The normal equation failed at an epsilon of ',toString(eps)))
+
+### Part d)
+model<-lm(y~x.i)
+summary(model)
+beta.qr
+
+#############################
+# Problem 5
+#############################
+
+data(prostate)
+
+### 3.1.a)
+m<-lm(lpsa~lcavol + lweight + age +lbph + svi + lcp + gleason + pgg45, data=prostate)
+coe<-summary(m)$coef
+ci.9 <- c(coe[4,1] + qt(.05,88)*coe[4,2],coe[4,1] - qt(.05,88)*coe[4,2])
[…]

### 2014-10-28 stat343/hw4.R

@@ -1,0 +1,18 @@
+library("alr3")
+
+###########################
+# Problem 4
+###########################
+data(wm1)
+attach(wm1)
+
+### 2.13
+# 1
+pdf("4.2.13.1.pdf")
+plot(RSpd,CSpd)
+dev.off
+# 2
+reg<-lm(CSpd~RSpd)
+summary(reg)
+# 3
+predict(reg,data.frame(CSpd=NA,RSpd=7,4285), interval="confidence")

### 2014-10-29 stat343/hw4.tex
Finnished homework 4

@@
 Otherwise, we want to minimize the non-trivial part of the expression.
-Where $R_{11}$ has $m$ rows, let $Y_1$ be the first $m$ rows of $Y$ and $Y_2$ be the last $n-m$ rows, and $Q_1$ the first $m$ rows of $Q$ with $Q_2$ the remiander.
+Where $R_{11}$ has $m$ rows, let $Q_1$ be the first $m$ rows of $Q$ with $Q_2$ the remainder.
 Thus,
                 \begin{align*}
-                \|RP^T\beta-Q^TY\|^2 &= \left\| \left[\begin{array}{cc} R_{11} & R_{12} \\ \mathbf{0} & \mathbf{0} \end{array} \right] P^T \beta - [Q_1^T Q_2^T]\left[\begin{array}{c} Y_1 \\ Y_2 \end{array} \right] \right \|^2 \\
-                    &= \| \left[R_{11} R_{12}\right] P^T \beta - Q_1^T Y_1 \|^2 + \|Q_2^T Y_2\|^2
+                \|RP^T\beta-Q^TY\|^2 &= \left\| \left[\begin{array}{cc} R_{11} & R_{12} \\ \mathbf{0} & \mathbf{0} \end{array} \right] P^T \beta - [Q_1^T Q_2^T]Y \right \|^2 \\
+                    &= \| \left[R_{11} R_{12}\right] P^T \beta - Q_1^T Y \|^2 + \|Q_2^T Y\|^2
                 \end{align*}
-                Since $\|Q_2^TY_2\|^2$ is also fixed, we can reduce our minimization problem to minimizing \(\| [R_{11} R_{12}] P^T \beta - Q_1^T Y_1 \|^2\).
-Let the first $m$ rows of $P^T\beta$ be $\beta_1$ and the remainder be $\beta_2$.
-$P^T\beta$ is just a permutation of $\beta$, so it is the same coefficients, just out of order.
+                Since $\|Q_2^TY_2\|^2$ is also fixed, we can reduce our minimization problem to minimizing 
+                \[\| [R_{11} R_{12}] P^T \beta - Q_1^T Y \|^2\]
                 Thus, we will get a minimum when,
-                \begin{align*}
-                    \left[R_{11} R_{12}\right] P^T \beta &= Q_1^T Y_1 \\
-                    \left[R_{11} R_{12}\right] \left[ \begin{array}{c} \beta_1 \\ \beta_2 \end{array} \right]  &= Q_1^T Y_1 \\
-                    R_{11}\beta_1 + R_{12}\beta_2 &= Q_1^T Y_1 \\
-                    R_{11}\beta_1  &= Q_1^T Y_1 - R_{12}\beta_2 \\
-                    \beta_1  &= R_{11}^{-1}(Q_1^T Y_1 - R_{12}\beta_2) 
-                \end{align*}
-                Thus, when $\beta_1$ is as defined, we will have minimized $RSS$.
+                \[ \left[R_{11} R_{12}\right] P^T \beta = Q_1^T Y \]
             \item[c)]
-                We want to solve \(\left[R_{11} R_{12}\right] P^T \beta = Q_1^T Y_1\).
-Let \(W=\left[R_{11} R_{12}\right] P^T\).
-If we have some solution $\beta$, and a vector $a$ in the null space of \(W\), then
-                \[ W(\beta+a) = W \beta + W a = Q_1^T Y_1 + 0 \]
+                We want to solve the expression above. Let \(W=\left[R_{11} R_{12}\right] P^T\). If we have some solution $\beta$, and a vector $a$ in the null space of \(W\), then 
+                \[ W(\beta+a) = W \beta + W a = Q_1^T Y + 0 \]
                 So $\beta+a$ is also a solution.
-Let $\beta^*$ be the vector that acheives the smallest norm and solves the equation.
+Let $\beta^*$ be the vector that achieves the smallest norm and solves the equation.
 Assuming $R$ isn't all $0$, then there is some $\beta$ perpendicular to all $a$.
-If $\beta_1$ is and another $\beta_2$ isn't, by Cauchy-Schwarz
+If $\beta_1$ is and another $\beta_2$ isn't, by Cauchy-Schwartz
                 \begin{align*}
@@
                     W\beta^* &= WW^Tv \\
-                    Q_1^T Y_1 &= WW^Tv  \\
-                    W^T(WW^T)^{-1} Q_1^T Y_1 &= W^T v \\
-                    W^T(WW^T)^{-1} Q_1^T Y_1 &= \beta^*  \\
+                    Q_1^T Y &= WW^Tv  \\
+                    W^T(WW^T)^{-1} Q_1^T Y &= W^T v \\
+                    W^T(WW^T)^{-1} Q_1^T Y &= \beta^*  \\
                 \end{align*}
                 Thus we have our answer. We can invert \(WW^T\) since \(W\) is full row rank (since \(R_{11}\) is invertible) and similarly \(W^T\) is full column rank.
+            \item[d)]
+                Here is the implemented R script:
+                \lstinputlisting[firstline=8,lastline=19]{hw4.R}
+            \item[e)]
+                The two sets of coefficients are identical, except for the coefficients for pop75.
+In the normal regression, the coefficient was -1.6915, while in the solution to the singular version, the coefficients are both -.8457.
+\(-.8457+-.8457=-1.6915\) because the total additive effect needs to be the same.
+For any $a$, \(-1.6915a\) and \(-1.6915(1-a)\) would have accomplished this, but since we insisted on the smallest norm, we got $a=.5$, which is the minimum of \((-1.6915a)^2+(-1.6915(1-a))^2\).
+            \item[f)]
+                It seems like a robust heuristic to compute the rank of $R$ would be to count the number of entries on the diagonal which have a magnitude above a very small threshold (maybe $1e^{-10}$.
+Here, the diagonal for the second occurrence of pop75 is $-1.7e^{-15}$, despite the fact that variable is linearly dependent on the first occurrence of pop75 and should be $0$.
+Thus, it seems like the $QR$ has small rounding errors which we could exclude by this heuristic.
+\\
+                If $R$ is not singular but extremely close, this suggests that a robust approach would be to exclude the variables which have an extremely small diagonal value in $R$, or, alternatively, treat them as singular and use the method developed in this problem to find $\beta$.
         \end{itemize}
 
-    \item{2.}
+    \item[2.]
         \begin{itemize}
-            \item{a)}
-                The trace of $H$ is its rank, since its idempotent. 
+            \item[a)]
+                We know that when $A=BC$, $a_{ij}=b_i c_j$, where $b_i$ and $c_j$ are respectively the $i$th row and $j$th column of $B$ and $C$.
+Further, if $C$ is itself the product of matrices $D$ and $E$, then $c_j=De_j$, where $e_j$ is the $j$th column of $D$.
+Combining this, $a_{ij}=b_i D e_j$.
+If we treat the $i$th and $j$th rows of $X$ as column vectors on their own, this leads to the conclusion that
+                \[h_{ij}=x_i^T(X^TX)^{-1}x_j\]
+                \smallskip
+                Since $H$ is symmetric and idempotent,
+                \[h_{ii}=\sum_{j=1}^n h_{ij}h_{ji} = \sum_{j=1}^n h_{ij}^2 \]
+                Also, if \(J=\{j:x_i=x_j\}\), then for all \(j\in J\),
+                \[h_{ij} = x_j^T (X^TX)^{-1} x_i = x_i^T (X^TX)^{-1} x_i = h_{ii} \]
+                So, combining these,
+                \begin{align*}
+                    h_{ii} &= \sum_{j=1}^n h_{ij}^2 \\
+                    h_{ii} &= r h_{ii}^2 + \sum_{j\not\in J} h_{ij}^2 \\
+                    h_{ii} &\geq r h_{ii}^2 \\
+                    \frac{h_{ii}}{r} &\geq h_{ii}^2  \\
+                    \frac{1}{r} &\geq h_{ii} \\
+                \end{align*}
+                \smallskip
+                If $X$ includes a column of $1$s at the front (for an intercept), then $\mathbf{1}_n$ is in the span of $H$ and $H\mathbf{1}_n=\mathbf{1}_n$. We may conclude that 
+                \[\sum_{j=1}^n h_{ij} = 1 \]
+                This would imply that, since $\sum_{j=1}^n h_{ij}^2$ is minimized when $h_{ij}=\frac{1}{n}$ for all $h_{ij}$,
+                \[\sum_{j=1}^n h_{ij}^2 \geq \sum_{j=1}^n \frac{1}{n^2} = \frac{1}{n} \]
+                Combining this with what we showed earlier,
+                \[h_{ii} = \sum_{j=1}^n h_{ij}^2 \geq \frac{1}{n}\]
+                Giving us the desired results that 
+                \[\frac{1}{r} \geq h_{ii} \geq \frac{1}{n} \]
+            \item[b)]
+                When $X$ is an invertible matrix, 
+                \[\sum_{i=1}^n h_{ii} = \mathrm{tr}(H) = \mathrm{rk}(H) = n\]
+                So $h_{ii}=1=\frac{1}{r}$, since $r=1$ for all $i$. If $X$ is just one row repeated $n$ times, then $\mathrm{rk}(X)=1$, and 
+                \[\sum_{i=1}^n h_{ii} = \mathrm{tr}(H) = \mathrm{rk}(H) = 1 \]
+                So it must be the case that $h_{ii}=1=\frac{1}{n}$
         \end{itemize}
-    \item{4.}
+    \item[3.]
         \begin{itemize}
-            \item{2.13.}
-                \begin{itemize}
-                    \item{1.}
+            \item[a)]
+                Under the normal assumption,
+                \[ \hat\beta \sim N_p(\beta,\sigma^2(X^TX)^{-1}) \]
+                So we can also assume 
+                \[ R\hat\beta \sim N_q(R\beta,\sigma^2R(X^TX)^{-1}R^T) \]
+                In turn since, $z$ is independent of $Y$, and thus independent of $\hat\beta$, 
+                \begin{align*}
+                     z &\sim N_q(R\beta, \sigma^2 \mathbb{I}_q) \\
+                    \Rightarrow  R\hat\beta - z &\sim N_q(R\beta - R\beta,\sigma^2R(X^TX)^{-1}R^T+\sigma^2 \mathbb{I}_q) \\
+                    \Rightarrow  \frac{R\hat\beta - z}{\sigma} &\sim N_q(\mathbf{0},(R(X^TX)^{-1}R^T+\mathbb{I}_q))
+                \end{align*}
+                Accordingly, we would expect that 
+                \[\frac{\frac{1}{q}(R\hat\beta - z)^T(R(X^TX)^{-1}R^T+\mathbb{I}_q)^{-1}(R\hat\beta - z)}{\hat\sigma^2} \sim F_{q,n-p} \]
+                So we would get a confidence set of 
+                \[ E_\alpha = \left\{\beta\in \R^q : (R\hat\beta - z)^T(R(X^TX)^{-1}R^T+\mathbb{I}_q)^{-1}(R\hat\beta - z) \leq q\hat\sigma^2 f_{q,n-p,\alpha} \right\} \]
+            \item[b)]
+                I have plotted the confidence intervals for where $z$ will land:
+                \begin{center}
+                    \includegraphics[width=12cm]{hw4_3_b_ci} 
+                \end{center}
+                You will notice that the confidence interval is very close to a circle.
+If we create just a confidence interval for \(R\hat\beta\), it is a very elongated ellipse, with the main axis not parallel to the $z1$ or $z2$ axis.
+However, the noise of $z$ around$R\beta$ has no covariance, and there is equal variance in the direction of $z1$ and $z2$, so if it we created a confidence interval for it conditioned on a $R\beta$, we would get a circle.
+Since the magnitude variance of $z$ around $X\beta$ is much higher than the variance of the estimate of $R\hat\beta$, this makes the confidence interval close to a circle.
+                Here is the R code I used:
+                \lstinputlisting[firstline=38,lastline=65]{hw4.R}
+            \item[c)]
+                Having repeated the simulation $1000$ times, $z$ fell within the $90\%$ confidence interval $91.3\%$ of the time, which indicates that my expression for the confidence interval was pretty accurate.
+I calculated whether or not the simulated data point fell in the confidence interval using exactly the formulation from part a.
+I've attached the code:
+
+                \lstinputlisting[firstline=68,lastline=79]{hw4.R}
 
         \end{itemize}
+    \item[4.]
+        \begin{itemize}
+            \item[2.13.]
+                \begin{itemize}
+                    \item[1.]
+                        Looking over this plot, simple linear regression looks extremely plausible for this model.
+We see what looks like a linear relationship between the two variables with a multivariate normal noise distribution, which is exactly what we would want.
+                        \begin{center}
+                            \includegraphics[width=12cm]{hw4_4_2_13_1_plot} 
+                        \end{center}
+                    \item[2.]
+                        For the coefficients, we get \\
+                        \begin{tabular}{c | c c c c}
+                            & Estimate & Std. Error & t value & Prob \\
+                            \hline
+                            Intercept & 3.141 & .120 & 18.52 & $<$2e-16 \\
+                            RSpd      & .756 & .020 & 38.50 & $<$2e-16
+                        \end{tabular} \\
+                        These coefficients are both quite meaningful, the t values indicating that there is a negligible probability of the estimates being positive by chance if each coefficient was actually 0.
+The most interesting is the coefficient on RSpd, which indicates that their is a strong correlation between the wind speed at the two sites.
+Additionally, we have a residual standard error on $2.466$ degrees of freedom, an R-squared of $.5709$, and a F-statistic of $1482$ on $1$ and $1114$ degrees of freedom, resulting in a p-value of $<2e-16$.
+This speaks very strongly that this regression is meaningful; it explains $57\%$ of the original variance, and the probability that we would see these results if we just had a mean model is negligible.
+                    \item[3.]
+                        From the model, we get a prediction that when the wind at the reference site was $7.4285$, we would see wind at the test site of $\hat y_0 = [1 \> 7.4285]\beta = 8.7552$.
+We predict with $95\%$ certainty, based on the normality assumption, that the actual test site wind speed would be in the range \(\hat y_0 \pm t_{1114}^{.025} \hat\sigma\sqrt{1 + [1\> 7.4285]^T(X^TX)[1\> 7.4285]} = (3.914,13.570) \).
+                    \item[4.]
+                        Starting with the mean of the prediction, 
+                        \begin{align*}
+                            \frac{1}{m} \sum_{i=1}^m \hat y_{*i} x &= \frac{1}{m} \sum_{i=1}^m (x_{*i}\beta_1 + \beta_0) \\ 
+                            \frac{1}{m} \sum_{i=1}^m \hat y_{*i} x &= \hat\beta_0 + \frac{1}{m} \sum_{i=1}^m (x_{*i}\beta_1) \\ 
+                            \frac{1}{m} \sum_{i=1}^m \hat y_{*i} x &= \hat\beta_0 + \frac{\hat\beta_1}{m} \sum_{i=1}^m (x_{*i})  \\ 
+                            \frac{1}{m} \sum_{i=1}^m \hat y_{*i} x &= \hat\beta_0 + \hat\beta_1 \bar x  
+                        \end{align*}
+                        So we conclude the prediction for the mean is equal to the mean of the predictions. If we call the prediction at $\bar x_*$ $\bar y_*$ and the expectation $\hat y_*$, then
+                        \begin{align*}
+                            \var(\bar y_*) &= \var(\bar y_* - \hat y_* + \hat y_m ) \\  
+                            \var(\bar y_*) &= \var(\bar y_* - \hat y_*) + \var(\hat y_*) -2\cov(\bar y_* - \hat y_*,\hat y_*) \\  
+                            \var(\bar y_*) &= \frac{\hat\sigma^2}{m} + \hat\sigma^2\left(\frac{1}{n} + \frac{(\bar x_* - \bar x)^2}{SXX} \right) \\  
+                            se \>\bar y_* &= \sqrt{\frac{\hat\sigma^2}{m} + \hat\sigma^2\left(\frac{1}{n} + \frac{(\bar x_* - \bar x)^2}{SXX} \right)}   
+                        \end{align*}
+                        We know the covariance term is equal to $0$ since the residuals and the predictions are orthogonal, and thus have $0$ covariance.
+                    \item[5.]
+                        We can plug the values given and from the data set into the above equation to get the standard error.
+We get a $\sigma^2=6.082$ from the regression, $\bar x=7.778$, and $SXX= 15785.56$ from the data.
+Thus, the standard error is:
+                        \[ se \>\bar y_* = \sqrt{\frac{6.082}{62039} + 6.082\left(\frac{1}{1116} + \frac{(7.429 - 7.778)^2}{15785.56} \right)} = .075 \]   
+                        Combining this with the model's prediction of the expectation of $\bar y_*=8.7552$ from earlier, we get a $95\%$ confidence interval of 
+                        \[\bar y_* \pm se(\bar y_*) t_{1114}^{.025} = (8.609,8.901) \]
+                \end{itemize}
+
+            \item[4.6.]
+                When I bootstrapped the mean wind speed at the candidate site, I got a $95\%$ of the means landing in $(8.800,9.22)$.
+This is surprising, because its in general higher, and only partially overlaps with the analytic predicted confidence interval.
+So, I checked the assumption of normality:
+                \begin{center}
+                    \includegraphics[width=12cm]{hw4_4_4pt6_qq} 
+                \end{center}
+                Here, the qq plot shows that the distribution of the residuals has a thicker tail than the normal distribution.
+This supports the notion that there is more variability than we assumed, and in particular the upper bound of the analytic confidence interval is too low, supporting the bootstrapped confidence interval.
         \end{itemize}

### 2014-10-29 stat343/hw4.R

@@ -1,4 +1,82 @@
-library("alr3")
+library('alr3')
+library('faraway')
+library('car')
+library('MASS')
+
+#############################
+# Problem 1
+#############################
+
+### Part d)
+# Load the data and create the matrices 
+data('savings')
+attach(savings)
+n<-nrow(savings)
+x<-as.matrix(data.frame(rep(1,n),savings[c(2,3,4,3,5)]))
+y<-as.matrix(savings[1])
+qr<-qr(x)
+
+R<-qr.R(qr)[1:qr$rank,]
+Q<-qr.Q(qr)[,1:qr$rank]
+W<-R[,qr$pivot]
+beta<-t(W) %*% solve(W%*%t(W)) %*% t(Q) %*% y 
+beta 
+
+### part e)
+
+model<-lm(y~x)
+summary(model)
+
+### part f)
+qr.R(qr)
+
+#############################
+# Problem 3 
+#############################
+
+### part b
+# Linear model and output
+model<-lm(sr ~ pop15 + pop75 + dpi + ddpi)
+X <- model.matrix(model)
+beta<-model$coef
+sigma2 <- summary(model)$sigma^2
+df <- summary(model)$df[2]
+
+# Set up for ellipse
+R<-matrix(c(0,0,1,-1,-1,0,0,0,0,1),nrow=2,ncol=5)
+variance.pred <- R %*% solve(t(X) %*% X) %*% t(R) + diag(2) 
+variance.mean <- R %*% solve(t(X) %*% X) %*% t(R)
+
+mycenter <- R%*%beta
+radius9<- sqrt(2* sigma2 * qf(.1,2,df, lower.tail=FALSE))
+radius95<- sqrt(2* sigma2 * qf(.05,2,df, lower.tail=FALSE))  
+
+# Draw the ellipses
+# Ellipse for 90% CI
+pdf("hw4_3_b_ci.pdf")
+plot.new()
+plot.window(xlim=c(-10,15),ylim=c(-10,15))
+title(main='Prediction Interval of z', xlab='z1', ylab='z2')
+car::ellipse(center=c(mycenter), shape=variance.pred, radius=radius9, lty=2, main='Confidence Interval for Prediction', col=palette()[1])
[…]

### 2014-10-29 stat374/hw2.tex
started homework 2

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Homework 2 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[a)]
                I will use SURE to estimate the risk of the generalized James-Stein estimator. Let
                \[\hat\theta^{JS,v}= Z - \frac{(Z-v)(n-2)\sigma^2}{\sum_{i=1}^n (Z_i - v)^2}= Z + g(Z) \]  
                Accordingly, 
                \begin{align*}
                    D_i &= \frac{\partial g_i}{\partial z_i} \\
                    D_i  &= -(n-2)\sigma^2\left(\frac{1}{\sum_{i=1}^n (z_i - v)^2} - \frac{(z_i-v)^2}{\left(\sum_{i=1}^n 2(z_i - v)^2\right)^2}\right) \\
                    \sum_{i=1}^n D_i &= -(n-2)\sigma^2\left(\frac{n}{\sum_{i=1}^n (z_i - v)^2} - \frac{2\sum_{i=1}^n(z_i-v)^2}{\left(\sum_{i=1}^n (z_i - v)^2\right)^2}\right) \\
                    \sum_{i=1}^n D_i &= -(n-2)\sigma^2\left(\frac{n-2}{\sum_{i=1}^n (z_i - v)^2}\right)  \\
                    \sum_{i=1}^n D_i &= -\frac{(n-2)^2\sigma^2}{\sum_{i=1}^n (z_i - v)^2} 
                \end{align*}
                Plugging this into the SURE formula, we get that
                \begin{align*}
                    \hat R (Z) &= n\sigma^2 + 2\sigma^2 \sum_{i=1}^n D_i + \sum_{i=1}^n g_i^2 \\
                    \hat R (Z) &= n\sigma^2 - 2\frac{(n-2)^2\sigma^4}{\sum_{i=1}^n (Z_i - v)^2} + \sum_{i=1}^n \left(\frac{-(Z_i-v)(n-2)\sigma^2}{\sum_{i=1}^n (Z_i - v)^2}\right)^2 \\
                    \hat R (Z) &= n\sigma^2 - 2 \frac{(n-2)^2\sigma^4}{\sum_{i=1}^n (Z_i - v)^2} + \frac{(n-2)^2\sigma^4\sum_{i=1}^n (Z_i - v)^2}{\left(\sum_{i=1}^n (Z_i - v)^2\right)^2} \\
                    \hat R (Z) &= n\sigma^2 - 2 \frac{(n-2)^2\sigma^4}{\sum_{i=1}^n (Z_i - v)^2} + \frac{(n-2)^2\sigma^4}{\sum_{i=1}^n (Z_i - v)^2} \\
                    \hat R (Z) &= n\sigma^2 - \frac{(n-2)^2\sigma^4}{\sum_{i=1}^n (Z_i - v)^2} \\
                \end{align*}
                Since the estimate is unbiased,
                \[ R(\hat\theta^{JS,v},\theta) = \E(\hat R(Z)) = n\sigma^2 - (n-2)^2\sigma^4\E\left(\frac{1}{\sum_{i=1}^n (Z_i - v)^2}\right) \]
                Since \(Z_i-v \sim N(\theta_i - v,\sigma^2)\), we can use a result from the book that 
                \[\E\left(\frac{1}{\sum_{i=1}^n (Z_i - v)^2}\right) \geq \frac{1}{(n-2)\sigma^2 + \|\theta-v\|^2} \]
                This gives us that
                \begin{align*}
                    R(\hat\theta^{JS,v},\theta) &\leq n\sigma^2 - \frac{(n-2)^2\sigma^4}{(n-2)\sigma^2 + \|\theta-v\|^2} \\
                    R(\hat\theta^{JS,v},\theta) &\leq \frac{n\sigma^2((n-2)\sigma^2 + \|\theta-v\|^2) - (n-2)^2\sigma^4}{(n-2)\sigma^2 + \|\theta-v\|^2} \\
                    R(\hat\theta^{JS,v},\theta) &\leq \frac{n(n-2)\sigma^4 + n\sigma^2\|\theta-v\|^2 - (n-2)^2\sigma^4}{(n-2)\sigma^2 + \|\theta-v\|^2} \\
                    R(\hat\theta^{JS,v},\theta) &\leq \frac{2\sigma^2(n-2)\sigma^2 + (2 + n -2)\sigma^2\|\theta-v\|^2}{(n-2)\sigma^2 + \|\theta-v\|^2} \\
                    R(\hat\theta^{JS,v},\theta) &\leq 2\sigma^2 + \frac{(n -2)\sigma^2\|\theta-v\|^2}{(n-2)\sigma^2 + \|\theta-v\|^2} \\
                \end{align*}
                Because 
                \[1 < \frac{\|\theta-v\|^2}{(n-2)\sigma^2 + \|\theta-v\|^2}\]
                We get that, for $n\geq3$,
                \[R(\hat\theta^{JS,v},\theta) \leq 2\sigma^2 + \frac{(n -2)\sigma^2\|\theta-v\|^2}{(n-2)\sigma^2 + \|\theta-v\|^2} < 2\sigma^2 + (n-2)\sigma^2 = n\sigma^2 \]
                \(R(\hat\theta^{JS,v},\theta)\) will be smallest when \(\frac{\|\theta-v\|^2}{(n-2)\sigma^2 + \|\theta-v\|^2}\) is smallest, which in turn is when \(\|\theta-v\|^2\) is smallest. Thus, $v=\bar\theta$, the mean of $\theta$, will result in the smallest risk.
            \item[b)]
                In the case where $n=1$, since $\hat\theta = Z$ is admissiable, there is no estimator $\tilde\theta$ such that \(R(\tilde\theta,\theta)<R(\hat\theta,\theta)\), so $\hat\theta^{a,b}$ is admissible only if 
                \[R(\hat\theta,\theta)=\E(\hat\theta-\theta)^2=\E(\hat\theta^{a,b}-\theta)^2=R(\hat\theta^{a,b},\theta)\]
                This implies that, to be admissiable, it must be the case that $Z=aZ+b$ for all potential $Z$. Thus, since this is not the case for i, ii, iii, and iv, we must conclude the estimator is not admissable in these cases. Only in case iv can $Z=aZ+b$ for all potential $Z$, so there the estimator is admissable. 
                \smallskip
                As we showed in part a, $R(Z,\theta)=n\sigma^2>R(\hat\theta^{JS,v},\theta)$ when $n\geq3, so $\hat\theta = Z$ is not admissable. However, there are certainly cases where \(R(\hat\theta^{JS,v},\theta)\) is not admissable. Take for instance the example of $\theta_i=c$ for some constant $c$. The mean of $Z$ of has risk \(R(\bar Z,\theta) = \frac{\sigma^2}{n} < \)

        \end{itemize}
\end{itemize}
\end{document}

### 2014-10-30 stat374/hw2.tex
Mostly done with homework 2

@@
             \item[b)]
-                In the case where $n=1$, since $\hat\theta = Z$ is admissiable, there is no estimator $\tilde\theta$ such that \(R(\tilde\theta,\theta)<R(\hat\theta,\theta)\), so $\hat\theta^{a,b}$ is admissible only if
-                \[R(\hat\theta,\theta)=\E(\hat\theta-\theta)^2=\E(\hat\theta^{a,b}-\theta)^2=R(\hat\theta^{a,b},\theta)\]
-                This implies that, to be admissiable, it must be the case that $Z=aZ+b$ for all potential $Z$.
-Thus, since this is not the case for i, ii, iii, and iv, we must conclude the estimator is not admissable in these cases.
-Only in case iv can $Z=aZ+b$ for all potential $Z$, so there the estimator is admissable.
+                Let \(e\sim N(0,\sigma^2) \). For $n=1$, we can calculate the risk of \(\hat\theta^{a,b}\): 
+                \begin{align*}
+                    R(\hat\theta^{a,b},\theta) &= \E((aZ+b-\theta)^2) \\
+                                               &= \E((ae + a\theta -\theta + b)^2) \\
+                                               &= \E(a^2e^2+a^2\theta^2 +\theta^2 +b^2 + 2a^2e\theta - 2ae\theta + 2aeb - 2a\theta^2 + 2ba\theta - 2b\theta) \\
+                                               &= a^2\sigma^2+a^2\theta^2 +\theta^2 +b^2 - 2a\theta^2 + 2ba\theta - 2b\theta \\
+                                               &= a^2\sigma^2+(a-1)^2\theta^2 +b^2 + 2(a-1)b\theta  \\
+                                               &= a^2\sigma^2+((a-1)\theta +b)^2  \\
+                \end{align*}
+                \begin{itemize}
+                    \item[(i)]
+                        \(a=0\) is admissible, since this is the best predictor when \(b=\theta\), as we can see from the risk:
+                        \[R(\hat\theta^{0,b},\theta) = \theta^2 +b^2 + 2(-1)b\theta = 2\theta^2 -2\theta^2 = 0 \] 
+                        If a estimator $\tilde\theta$ is not uniformly $\theta=b$, then it will have positive risk, so \(R(\hat\theta^{0,b},\theta)< R(\tilde\theta,\theta)\), meaning $\hat\theta^{0,b}$ must be admissible.
+                    \item[(ii)]
+                        \(\hat\theta^{a,b}\) for \(a<0\) is not admissible, by this line of reasoning. If \(a<0\), the \(0<\left(\frac{1}{a-1}\right)^2<1\), so                      
+                        \begin{align*}
+                            R(\hat\theta^{a,b},\theta) &= a^2\sigma^2+((a-1)\theta +b)^2  \\
+                            R(\hat\theta^{a,b},\theta) &>\left(\frac{1}{a-1}\right)^2\left(a^2\sigma^2+\left((a-1)\theta +b\right)^2\right)  \\
+                            R(\hat\theta^{a,b},\theta) &>\left(\frac{a}{a-1}\right)^2\sigma^2+\left(\theta +\frac{b}{a-1}\right)^2  \\
+                            R(\hat\theta^{a,b},\theta) &>\left(-\theta -\frac{b}{a-1}\right)^2  \\
+                            R(\hat\theta^{a,b},\theta) &>R(\hat\theta^{0,\frac{-b}{a-1}},\theta)  \\
+                        \end{align*}
+                    \item[(iii)]
+                        \(\hat\theta^{a,b}\) for \(a>1\) is not admissible since, for all $\theta$,
+                        \[R(\hat\theta^{a,b},\theta)=a^2\sigma^2+((a-1)\theta +b)^2>\sigma^2+b^2 = R(\hat\theta^{1,b},\theta) \]
+                    \item[(iv)]+
+                        If \(a=1\) and \(b\neq0\), then \(\hat\theta^{a,b}\) is not admissible, since, for all $\theta$ 
+                        \[R(\hat\theta^{1,b},\theta)=\sigma^2+b^2>\sigma^2 = R(Z,\theta)\]
+                \end{itemize}
                 \smallskip
-                As we showed in part a, $R(Z,\theta)=n\sigma^2>R(\hat\theta^{JS,v},\theta)$ when $n\geq3, so $\hat\theta = Z$ is not admissable.
-However, there are certainly cases where \(R(\hat\theta^{JS,v},\theta)\) is not admissable.
-Take for instance the example of $\theta_i=c$ for some constant $c$.
-The mean of $Z$ of has risk \(R(\bar Z,\theta) = \frac{\sigma^2}{n} < \)
+                When $n\geq 3$, part a) proves that, for all $\theta$,
+                \[R(Z,\theta)=n\sigma^2>R(\hat\theta^{JS,v},\theta)\]
+                So $Z=\hat\theta$ is not admissible.
+                \smallskip
+                As (i) showed, for any particular $\theta$, A linear estimate like $\hat\theta^{a,b}$ is not admissible if $a<0$.
+Since the generalized James-Stein estimator is of this form, and can have a negative $a$, we can make a strictly improved version of it.
+This estimator would be equal to the generalize James-Stein estimator when it has a non-negative $b$, and \(\hat\theta^{0,\frac{-b}{a-1}}\), for the $b$ and $a$ from the generalized James-Stein estimator when it wasn't.As proved in (ii), this estimator would have the same risk when $a\geq0$, and would be strictly less risk when it wasn't.
+Thus, we can conclude the generalized James-Stein estimator is not admissible.
+            \item[c)]
+                Running the three estimates (all results in radians), I got a total risk of $17.867$ with the MLE, $16.636$ with the normal James-Stein estimator ($v=0$), and $5.280$ with the generalized James-Stein estimator with $v=\bar Z= -3.317$.
+Unsurprisingly with a data set that is not centered around $0$, we see a huge improvement with the generalized James-Stein method set to the population mean over the MLE and original James-Stein estimators.
+            \item[d)]
+                $\frac{9}{45}$ is almost certainly a less risky estimate.
+This is because there is no reason to expect that the portion of imported cars seen in Chicago and $18$ players batting averages have the same variance.
+Since this violates an assumption of the James-Stein estimator, it would be inappropriate to use it.
+        \end{itemize}
+    \item[2.]
+        \begin{itemize}
+            \item[a)]
+                For both log counts, I used a linear model that included the season as factor, working day as a factor, the interaction of season and working day as a factor, weather as factor, day label as a continuous variable, and hours separation from 5pm as a continuous variable.
+In total, I got a RMSLE of $1872.738$ for the registered users and $2163.457$ for the casual non-registered users.
+            \item[b)]
+                These are the fits of the two local linear regressions on the mean hourly transformed counts by day: \\
+                \includegraphics[width=9cm]{hw2_2b_reg_plot}
+                \includegraphics[width=9cm]{hw2_2b_cas_plot}
+                With the new local linear regression and the linear model predicted on its residuals, I got RMSLE of $1828.168$ for the registered users and $1875.078$ for the casual non-registered users.
+            \item[b)]
+
+        \end{itemize}
+    \item[3.]
+        Starting with the bias:
+        \begin{align*}
+            bias(\hat p_n(x_0,y_0)) &= p(x_0,y_0) - \E(\hat p_n(x_0,y_0)) \\
+                                    &= p(x_0,y_0) - \frac{1}{h^2}\iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)p(q,t)\,dt\,dq \\ 
+                                    &= \frac{1}{h^2}\iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)(p(q,t)-p(x_0,y_0))\,dt\,dq \\ 
+                                    &\leq \frac{1}{h^2}\iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)L(\lvert t - x_0 \rvert^\beta +\lvert q - y_0 \rvert^\beta )\,dt\,dq \\ 
+                                    &\leq \iint K\left(t\right)K\left(q\right)L(\lvert ht \rvert^\beta +\lvert hq \rvert^\beta )\,dt\,dq \\ 
+                                    &\leq 4Lh^\beta\iint_{\R^+\times\R^+} K\left(t\right)K\left(q\right)(t^\beta + q^\beta )\,dt\,dq \\ 
+                                    &\leq h^\beta c_1\\ 
+        \end{align*}
+        Since each $X_i$ is iid and the kernel estimator is linear, we can decompose the variance:
+        \begin{align*}
+            \var\left(\hat p_n(x_0,y_0)\right) &= \frac{1}{nh^4}\var\left(K\left(\frac{X_i-x_0}{h}\right)K\left(\frac{Y_i-y_0}{h}\right)\right) \\
+                                               &= \frac{1}{nh^4}\iint \left( K\left(\frac{x-x_0}{h}\right)K\left(\frac{y-y_0}{h}\right) - \iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)p(q,t)\,dt\,dq\right)^2 p(x,y)\,dx\,dy \\
+        \end{align*}
+    \item[4.]
+        \begin{itemize}
+            \item[a)]
+                Here are the histograms of the variable densities. The number of bins was chosen between 1 and 200 based on cross validation score.\\
+                \includegraphics[width=9cm]{hw2_4a_Ba_hist}
+                \includegraphics[width=9cm]{hw2_4a_Ca_hist} \\ 
+                \includegraphics[width=9cm]{hw2_4a_Fe_hist} 
+                \includegraphics[width=9cm]{hw2_4a_K_hist}   \\
+                \includegraphics[width=9cm]{hw2_4a_Mg_hist}
+                \includegraphics[width=9cm]{hw2_4a_Na_hist}  \\
+                \includegraphics[width=9cm]{hw2_4a_RI_hist} 
+                \includegraphics[width=9cm]{hw2_4a_Si_hist}  \\
+            \item[b)]
+                These are the density estimates with bandwidth picked by cross validation.
+The Gaussian Kernel's bandwidth was picked with R's built in unbiased cross validation tool, while the other two I calculated manually on a series of points around the optimal Gaussian bandwidth, picking the smallest estimated risk.
+In general, the Gaussian seemed to go with the higher bandwidth, producing the smoother estimate, but that may be in part due to a failure in my implementation of cross validation.
+Certainly it is smoother than the other two kernels by construction, since they are $0$ except over a closed interval.
+                \includegraphics[width=9cm]{hw2_4b_Ba_den}
+                \includegraphics[width=9cm]{hw2_4b_Ca_den} \\ 
+                \includegraphics[width=9cm]{hw2_4b_Fe_den} 
+                \includegraphics[width=9cm]{hw2_4b_K_den}   \\
+                \includegraphics[width=9cm]{hw2_4b_Mg_den}
+                \includegraphics[width=9cm]{hw2_4b_Na_den}  \\
+                \includegraphics[width=9cm]{hw2_4b_RI_den} 
+                \includegraphics[width=9cm]{hw2_4b_Si_den}  \\
+            \item[c)]
+                Cross validation seems to do poorly with variables with one or a few large point masses.
+We see this several times with a large number of points being at $0$.
+This is because cross validation rewards the estimator for having an extremely sharp peak around these point masses, even though the rest of the distribution suffers.
+One might view these as almost two separate distributions.
+A better way to estimate these sort of variables would be to predict the point mass separately, remove those points, and then run the density estimate on the rest of the distribution.
 

### 2014-11-10 stat343/hw5.tex
Done homework 5

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 343, Homework 5 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[a)] 
                The constant variance assumption apiears to hold; we do not seem to have the variance change over the range of predicted values. The plot on the left, which is the residuals versus the fitted values, indicates by the eyeball test that the distribution of residuals is constant across the range of predicted values. This is shown more formally by the plot on the right, where I have fit a linear model to the square root of the absolute value of each residual in terms of the fitted value. This yields a coefficient of $-.00016$ for the fitted values, with a monstrous p-value of $.969$, dictating that the fitted values of the original regression really don't have any correlation with the deviation of the residuals. \\
                \includegraphics[width=9cm]{hw5_1_a_res} 
                \includegraphics[width=9cm]{hw5_1_a_sqrtres} 
            \item[b)]
                The normality assumption also seems to hold; it appears the residuals are normally distributed around the predicted values. This is indicated by the eyeball test, where we can see from the quantile-quantile plot that the quantiles of the residuals match the quantiles of a normal distribution fairly closely. Additionally, testing that they are normal formally, the Shapiro-Wilkes test yields a p-value of $.430$, which is far to high to reject normality.
                \begin{center}
                    \includegraphics[width=9cm]{hw5_1_b_qq} 
                \end{center}
            \item[c)]
                We do see some points with large leverage, but it doesn't seem that any stand out as being unusually large. I have made a quantile-quantile plot of the leverages against the half normal distribution. We see the states with the highest leverage are Utah followed by California. However, they don't appear to be particular extreme or out of line with the other leverages.
                \begin{center}
                    \includegraphics[width=9cm]{hw5_1_c_halfnorm} 
                \end{center}
            \item[d)]
                It doesn't seem that we he have any outliers. The studentized residuals should have a $t_{n-p-1}$ distribution should the assumptions hold, and as we can see from the quantile-quantile plot, that seems to be true. Further, if we look at the most extreme studentized residual, which is the best candidate to be an outlier, we find it is only $-3.124$, which is less than the Bonferroni critical value at a $.1$ alpha level, so we can conclude neither it, nor any other point, is an outlier. 
                \smallskip
                \begin{center}
                    \includegraphics[width=9cm]{hw5_1_d_tqq} 
                \end{center}
                The Bonferroni critical value at an alpha level of $.1$ is the \(\frac{\alpha}{2p}\) quantile of a $t_{n-p-1}$ distribution, giving a value of $-3.281$ in this case. Since the Bonferroni critical value is conservative by construction, I tested at a higher alpha level just to see if their were any outlier candidates.
            \item[e)]
                We seem to have one point that is extremely influential, Utah, and a few additional observations in West Virginia, Florida, New Hampshire and Connecticut which are fairly influential. By looking at the first plot, the Cook Distance, one sees that the inclusion of Utah has far bigger effect on the predictions than any other point, though West Virginia and New Hampshire have pretty large effects in their own right. Looking at the effect of Utah's inclusion on each coefficient individually (the last five graphs), it is the most influential observation on four of the five coefficients, supporting this. However, despite the fact the cook distance of Connecticut and Florida wasn't that high, they are each fairly influential on a few variables. New Hampshire and West Virginia are fairly influential on a variable or two a piece as well, as the cook distance pointed to.
                \begin{center}
                    \includegraphics[width=9cm]{hw5_1_e_halfnorm} 
                \end{center}
                \includegraphics[width=9cm]{hw5_1_e_Intercept}
                \includegraphics[width=9cm]{hw5_1_e_takers} \\
                \includegraphics[width=9cm]{hw5_1_e_salary} 
                \includegraphics[width=9cm]{hw5_1_e_ratio}  \\
                \includegraphics[width=9cm]{hw5_1_e_expend} 
            \item[f)]
                The assumption that there is a linear relationship between the predictors and the outcome does not quite seem to hold. We can see this most clearly when we try to predict the residuals based on the predicted values with a more flexible local linear regression (the first plot); when we do, we can see that the mean of the residuals is lower than $0$ in the center of the predicted values, and higher on the edges, indicating a non-linear, possibly quadratic relationship. Digging into this further, repeating this process on the four partial residual plots (the last four plots), it seems like this is because there is a non-linear fit between takers and the outcome. For the other three variables, while the local linear regressions may be picking up some meaningful non-linearity, they generally appear to just be fluctuating around the linear fit with noise.
                \begin{center}
                    \includegraphics[width=9cm]{hw5_1_f_locfitres} 
                \end{center}
                \includegraphics[width=9cm]{hw5_1_f_takers}
                \includegraphics[width=9cm]{hw5_1_f_salary} \\
                \includegraphics[width=9cm]{hw5_1_f_ratio}  
                \includegraphics[width=9cm]{hw5_1_f_expend} \\ 
        \end{itemize}
    \item[2.]
        \begin{itemize}
            \item[a)]
                There are only discrete values for waiting, which is in minutes. This indicates that there is some kind of rounding error, where the actual waiting time was rounded to the minute. Since there is no reason to expect that an eruption was more or less likely based on how close the actual time was to a full minute, I am going to assume that the error is uniformly distributed from $-.5$ to $.5$. In other words, I am assuming the time was rounded to the nearest minute. 
                \begin{center}
                    \includegraphics[width=9cm]{hw5_2_a} 
                \end{center}
                The regression I fit can be seen above. There is an intercept of $-1.8740$, and a coefficient on waiting of $.07563$.
            \item[b)]
                If we have a continuous, uniformly distributed error, we can simulate larger continuous uniform errors by summing it with the right discrete uniform random error. Thus, we can simulate a random error that is uniformly distributed on $(-1.5,1.5)$ by randomly choosing from ${-1,0,1}$ and adding it to the original error, an error on $(-2.5,2.5)$ by randomly choosing from ${-2,-1,0,1,2}$ and so on. Taking the mean beta over $1000$ simulations where a particular discrete error was added to the waiting time and the model was fit, we get a distribution that seems to have a linear relationship between variance and $\beta$. Fitting a linear model to it, we can extrapolate to a measurement error level of $0$, giving us a predicted $\beta$ of $.07565$.
                \begin{center}
                    \includegraphics[width=9cm]{hw5_2_b}
                \end{center}
                The above shows the distribution of $\beta$ at different variance levels corresponding to errors uniformly distributed over $(-.5-z,.5+z)$, for $z\in\Z$. These were chosen since, to the best of my knowledge, there is no random variable $X$ such that \(u_{-.5,.5}+X \sim u_{-.5-q,.5+q}\) when $q\not\in\Z$, making these errors hard to simulate.
            \item[c)] If we tried to apply the SIMEX method to variance in the response variable, by running the regression with additional variance added, the expectation of $\beta$ would still be the same irrespective of additional noise, and any fluctuation would be distributed around the original $\beta$. Thus, our extrapolation to $0$ variance would give the same $\beta$. \\
                {\bf Proof:} Let $Y$ be the original set of responses, and $Y^*=Y+e$ for some random variable $e$ with expectation $0$. Thus, \(\E(y^*)=y,\E(\bar y^*) = \bar y\), and \(\E(y^*-\bar y^*) = y -\bar y\). Now, starting with the coefficient on the predictor for a model fit on $Y^*$, $\beta^*$,
                \begin{align*}
                    \beta^* &= r_{xy}\frac{SD_{y^*}}{SD_x} \\
                    \beta^* &= \frac{s_{xy}}{SD_x SD_{y^*}}\frac{SD_{y^*}}{SD_x} \\
                    \beta^* &= \frac{\sum(x_i-\bar x)(y_i^*-\bar y^*)}{(n-1)SD_x^2}\\
                    \E[\beta^*] &= \E\left[\frac{\sum(x_i-\bar x)(y_i^*-\bar y^*)}{(n-1)SD_x^2}\right] \\
                    \E[\beta^*] &= \frac{\sum(x_i-\bar x)(y_i-\bar y)}{(n-1)SD_x^2} \\
                    \E[\beta^*] &= \beta \\
                \end{align*}
                So in the expectation, you would get the same $\beta$ irrespective of the variance level.
            \item[d)]
                \begin{center}
                    \includegraphics[width=9cm]{hw5_2_d}
                \end{center}
                The residual plot above seems to show that the data really represents two separate regions, for each of which we are systematically underestimating the lower end and overestimating the higher end. The split here corresponds to approximately a waiting time of $68$. This would suggest we really need two different models for waiting times above and below $68$. To test this, I fit a new model which had three new variables.
                \begin{itemize}
                    \item[1-] A binary for whether the waiting time was higher than $68$
                    \item[2-] A variable which was the waiting time if higher than $68$ and otherwise $0$.
                    \item[2-] A variable which was the waiting time if lower than $68$ and otherwise $0$.
                \end{itemize}
                This is essentially two separate models. Since the original model corresponds to a subspace of this model (where the coefficient on the first variable is $0$ and the second two hav equal coefficients), we can compare the two models with a $F$ test: \\
                \begin{table}[ht]
                \centering
                \begin{tabular}{lrrrrrr}
                    \hline
                       & Res.Df & RSS & Df & Sum of Sq & F & Pr($>$F) \\ 
                        \hline
                        1 & 270 & 66.56 &  &  &  &  \\ 
                          2 & 268 & 40.21 & 2 & 26.35 & 87.83 & 0.0000 \\ 
                           \hline
                   \end{tabular}
               \end{table} \\
               Since the p-value is far below $.05$, we can reject the null hypothesis that the original linear model is sufficient, and accept that really two separate models are needed.
        \end{itemize}
    \item[3.]
        \begin{itemize}
            \item[a)]
                Simulating the largest studentized residual on the star data set, I found that there was a larger studetnized residual in the simulation $.974$ of the time. On the savings data set, there was a larger studentized residual $.312$ of the time. I would reject the corresponding point as an outlier at $.05$, since that's the accepted threshold, and in this case would indicate that there is only a $.05$ chance of the point fitting the model so poorly if it indeed came from it.
            \item[b)]
                I found that $.219$ of the simulations of the star data set had a point with as big a Cook Distance. $.408$ of the simulations of the savings data set had a point with as big a Cook Distance. I would not exclude a point automatically at any confidence level of the Cook Statistic, since having a high value is indicative of being influential, but not necessarily a poor fit for the model.
            \item[c)] Both of these conclusions are similar to the analysis from class. With studentized residuals, we checked against the Bonferroni critical value to decide to exclude a point, while with Cook Distance we merely noted high values for further evaluation.
        \end{itemize}
    \item[4.]
        \[ M^{-1} = \left[\begin{array}{cc} H^{-1}+H^{-1}e(f^T-e^TH^{-1}e)^{-1}e^TH^{-1} & -H^{-1}e(f-e^TH^{-1}e)^{-1} \\ -(f^T-e^TH^{-1}e)^{-1}e^TH^{-1} & (f-e^TH^{-1}e)^{-1} \end{array} \right] \]
        {\bf How I Got Here:} First, I assumed that, since $M$ has nice symmetry,
    \[M^{-1} = \left[\begin{array}{cc} A & B \\ B^T & C \end{array}\right] \]
        Since $H$ is positive definite, it has an inverse. Since it is symmetric, the inverse is as well. Since \(MM^{-1}=I\), we get that
        \[HA+eB=I, \, HB+eC=0, \, e^TA+fB^T=0, \, e^TB+fC=I\]
        From the second equation, using $H^{-1}$, we get \(B=-H^{-1}eC\). Plugging this into the fourth equation, we get that \((f-e^TH^{-1}e)C=I\). Since we can repeat this process for \(M^{-1}M=I\) and get \(C(f-e^TH^{-1}e)=I\), we can conclude that $C$ is invertible with \(C=(f-e^TH^{-1}e)^{-1}\). Plugging the first two results into the first equation, we get
        \[HA-e(f-e^TH^{-1}e)^{-1}e^TH^{-1}=I\]
        so
        \[A=H^{-1} + H^{-1}e(f-e^TH^{-1}e)^{-1}e^TH^{-1}\]
        Plugging everything in and multiplying $MM^{-1}$, I get the desired result, so my initial assumption appears to be correct.
    \item[5)]
        Since 
        \[r_{u,v}=\frac{\langle u - \bar u, v -\bar v\rangle}{\|u-\bar u\|\|v-\bar v\|} \]
        if \(r_{x,y}=\pm 1\) and \(\bar v = \bar u = 0\), we get that
        \[\vert\langle u, v\rangle\vert = \|u\|\|v\| \]
        By Cauchy-Schwarz, this implies that \(u\) and \(v\) are linearly dependent. Accordingly, the design matrix, which is \( [\begin{array}{cc}u & v\end{array}]\), will only have one linearly independent column, and will thus have a rank of 1.
        \smallskip
        To get the variance of the estimators, we need to get the covariance matrix of the coefficient vector:
        \begin{align*}
            \cov\left[\begin{array}{c}\hat a \\ \hat b \end{array}\right] &= \sigma^2\left([\begin{array}{cc}u & v\end{array}]^T[\begin{array}{cc}u & v\end{array}]\right)^{-1} \\
            \left[\begin{array}{cc} \var(\hat a) & \cov(\hat a,\hat b) \\ \cov(\hat a,\hat b) & \var(\hat b) \end{array} \right] &= \left[\begin{array}{cc} \|u\|^2 & \langle u, v \rangle \\ \langle u, v \rangle & \|v\|^2 \end{array} \right]^{-1} \\
            \left[\begin{array}{cc} \var(\hat a) & \cov(\hat a,\hat b) \\ \cov(\hat a,\hat b) & \var(\hat b) \end{array} \right] &= \frac{1}{\|u\|^2\|v\|^2-\langle u, v \rangle^2} \left[\begin{array}{cc} \|u\|^2 & -\langle u, v \rangle \\ -\langle u, v \rangle & \|v\|^2 \end{array} \right] \\
            \left[\begin{array}{cc} \var(\hat a) & \cov(\hat a,\hat b) \\ \cov(\hat a,\hat b) & \var(\hat b) \end{array} \right] &= \frac{1}{\|u\|^2\|v\|^2(1-r^2)} \left[\begin{array}{cc} \|u\|^2 & - r\|u\|\|v\| \\ -r\|u\|\|v\| & \|v\|^2 \end{array} \right]
        \end{align*}
        Pulling what we need out of this, we get that \(\var(\hat a) = \frac{\|u\|^2}{\|u\|^2\|v\|^2(1-r^2)}\), \(\var(\hat b) = \frac{\|v\|^2}{\|u\|^2\|v\|^2(1-r^2)}\),
        \[\var(\hat a + \hat b) = \var(\hat a) + \var(\hat b) + 2\cov(\hat a,\hat b) = \frac{\|u\|^2 + \|v\|^2 - 2r\|u\|\|v\| }{\|u\|^2\|v\|^2(1-r^2)}\]
        and
        \[\var(\hat a - \hat b) = \var(\hat a) + \var(\hat b) - 2\cov(\hat a,\hat b) = \frac{\|u\|^2 + \|v\|^2 + 2r\|u\|\|v\| }{\|u\|^2\|v\|^2(1-r^2)}\]
        So, when $r=-.99$, we are going to see $\var(\hat a + \hat b)$ be large, since its numerator is large and its denominator is small (compared to a less negative $r$). $\var(\hat a)$ and $\var(\hat b)$ are going to have a small denominator, so they will also be large. Only $\var(\hat a - \hat b)$, with a numerator that shrinks as $r$ decreases will be relatively small. This means that \(\hat a - \hat b\) will be the easiest to estimate, while the others will be relatively harder.  
    \item[6)]
        \begin{itemize}
            \item[a)]
                I've calculated the condition number, which is, for each eigenvalue $\lambda$, \(\sqrt{\frac{\lambda}{\lambda_{min>0}}}\), below. The largest condition number is an astronomical $5751.22$, which is way higher than the $30$ which is considered large. That we have several other that are also large indicates we have several variables which are independently extremely close to linear combinations of other variables. All in all this dictates a lot of multicollinearity. \\
                \begin{table}[ht]
                    \centering
                    \begin{tabular}{rrrrrrr}
                          \hline
                           & 1 & 2 & 3 & 4 & 5 & 6 \\ 
                            \hline
                            1 & 5751.22 & 322.11 & 228.65 & 94.62 & 3.49 & 1.00 \\ 
                               \hline
                           \end{tabular}
                       \end{table}
            \item[b)] 
                I have calculated the pairwise collinearity below. We see a number of variables which have pairwise collinearities very close to one, such as between Year and GNP or Year and Population. This supports the notion from part a) that their is a ton of collinearity, showing several variables are pairwise collinear, though not addressing any multicollinearity. \\
                \begin{table}[ht]
                    \centering
                    \begin{tabular}{rrrrrrr}
                          \hline
                           & GNP.deflator & GNP & Unemployed & Armed.Forces & Population & Year \\ 
                            \hline
                            GNP.deflator & 1.00 & 0.99 & 0.62 & 0.47 & 0.98 & 0.99 \\ 
                              GNP & 0.99 & 1.00 & 0.60 & 0.45 & 0.99 & 0.99 \\ 
                              Unemployed & 0.62 & 0.60 & 1.00 & -0.18 & 0.69 & 0.67 \\ 
                              Armed.Forces & 0.47 & 0.45 & -0.18 & 1.00 & 0.36 & 0.42 \\ 
                              Population & 0.98 & 0.99 & 0.69 & 0.36 & 1.00 & 0.99 \\ 
                              Year & 0.99 & 0.99 & 0.67 & 0.42 & 0.99 & 1.00 \\ 
                               \hline
                           \end{tabular}
                       \end{table}
            \item[c)]
                The variance inflation factors below are calculated as \(\frac{1}{1-R_j^2}\), where \(R_j^2\) is the $R^2$ of a regression predicting one predictor variable using all the other predictor variables. That these are so high speak to the collinearity we've already observed. More specifically though, high variance inflation factors dictate that the estimate of the coefficient will have high variance. Thus, our estimates of the coefficients for GNP, Year, Population, among others, are going to be very noisy. \\
                \begin{table}[ht]
                    \centering
                    \begin{tabular}{rrrrrrr}
                          \hline
                           & GNP.deflator & GNP & Unemployed & Armed.Forces & Population & Year \\ 
                            \hline
                            1 & 135.53 & 1788.51 & 33.62 & 3.59 & 399.15 & 758.98 \\ 
                               \hline
                           \end{tabular}
                       \end{table}
        \end{itemize}
\end{itemize}
\end{document}

### 2014-11-10 stat374/hw2.tex
Done homework 5

@@
                 With the new local linear regression and the linear model predicted on its residuals, I got RMSLE of $1828.168$ for the registered users and $1875.078$ for the casual non-registered users.
-            \item[b)]
+            \item[c)]
+                I ran a generalized additive model with just smoothed functions for daylabel, humidity, temp, windspeed, and hour.
+The end result was that I got a RMSLE of $1352.043$ for registered customers and $1326.217$ for casual customers.
+            \item[d)]
+                For the solution data set I turned in, I predicted the total counts with an additive linear model predicting the log of the total counts with smoothed daylabel, humidity, temp, wind speed, hour, and categorical of working day, and weather interacted with season.
 
@@
             \var\left(\hat p_n(x_0,y_0)\right) &= \frac{1}{nh^4}\var\left(K\left(\frac{X_i-x_0}{h}\right)K\left(\frac{Y_i-y_0}{h}\right)\right) \\
-                                               &= \frac{1}{nh^4}\iint \left( K\left(\frac{x-x_0}{h}\right)K\left(\frac{y-y_0}{h}\right) - \iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)p(q,t)\,dt\,dq\right)^2 p(x,y)\,dx\,dy \\
+                                               &= \frac{1}{nh^4} \E\left(\left(\iint K\left(\frac{X-x_0}{h}\right)K\left(\frac{Y-y_0}{h}\right)p(t,q) - K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)p(t,q)\,dt\,dq\right)\right)^2 \\
+                                               &\leq \frac{1}{nh^4} \E\left(\left(\iint K\left(\frac{X-t}{h}\right)K\left(\frac{Y-q}{h}\right)p(t,q)\,dt\,dq\right)\right)^2 \\
         \end{align*}
+        And here is where I am stuck. However, if we had an expression for variance, then the mean squared error would be 
+        \[bias(\hat p_n(x_0,y_0))^2 + \var\left(\hat p_n(x_0,y_0)\right) \leq h^{2\beta}c_1 + \var\left(\hat p_n(x_0,y_0)\right)\]
+        Differentiating the right side of that with respect to $h$ and setting it to $0$, we would find the optimal $h$.
+That would have a leading term of the order $h^{2\beta-1}$, at least assuming the bias has the larger $h$ power.
+Then, we would expect to see $h$ shrink with $cn^\frac{-1}{2\beta+1}$., dictating $R$ converge at a rate of $O(n^\frac{2\beta}{2\beta+1})$.
     \item[4.]

### 2014-11-11 stat374/hw3.tex
Started homework 3

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 374 HW3 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[a)]
                \begin{align*}
                    \E[\hat p_{n,s}(x)] &= \E\left[\frac{1}{nh^{s+1}}\sum_{i=1}^n K\left(\frac{X_i-x}{h}\right)\right] \\
                                        &= \frac{1}{h^{s+1}}\int K\left(\frac{u-x}{h}\right)p(u) du\\
                                        &= \frac{1}{h^{s}}\int K\left(u\right)p(x+hu) du\\
                                        &= \frac{1}{h^{s}}\int K(u)\left(\sumj \frac{(hu)^j}{j!}p^{(j)}(x)\right) du\\
                                        &= \sum_{j=1}^{l-1}\frac{h^{j-s}}{j!}p^{(j)}(x)\int K(u)u^j du + \frac{1}{h^{s}}\int K(u)\left(\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x)\right) du\\
                                        &= p^{(s)}(x)+ \frac{1}{h^{s}}\int K(u)\left(\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x)\right) du
                \end{align*}
                Using the Lagrange form of the remainder of a taylor series, for some $\xi\in(x,x+hu)$,
                \[\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) = \frac{(hu)^l}{l!}p^{(l)}(\xi)\]
                By the conditionts of the H{\"o}lder class,
                \begin{align*}
                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\leq \frac{(hu)^l}{l!}(L\vert \xi - x\vert^{\beta-l} + \vert p(x)\vert) \\
                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\leq \frac{(hu)^l}{l!}(L\vert hu\vert^{\beta-l}  + \vert p(x)\vert) \\
                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\leq h^\beta\frac{u^l\vert u \vert^{\beta-l}L}{l!}+ u^l\frac{h^l}{l!}\vert p(x)\vert \\
                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\leq h^\beta f_1(u)+ u^lf_2(x) \\
                \end{align*}
                For \(f_1(u):=\frac{u^l\vert u \vert^{\beta-l}L}{l!}\) and \(f_2(x):=\frac{h^l}{l!}\vert p(x)\vert\). By similar logic, we also get
                \begin{align*}
                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\geq -\frac{(hu)^l}{l!}(L\vert \xi - x\vert^{\beta-l} + \vert p(x)\vert) \\
                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\geq -h^\beta f_1(u)- u^lf_2(x) 
                \end{align*}

                Plugging these into our expression for \(\E[\hat p_{n,s}(x)]\), we get that 
                \begin{align*}
                    \E[\hat p_{n,s}(x)] &= p^{(s)}(x)+ \frac{1}{h^{s}}\int K(u)\left(\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x)\right) du \\
                    \vert \E[\hat p_{n,s}(x)]-p^{(s)}(x) \vert &\leq \frac{1}{h^{s}}\int K(u) (h^\beta f_1(u)+ u^lf_2(x)) du \\
                                                                &\leq h^{\beta-s}\int K(u) f_1(u) du + \frac{f_2(x)}{h^{s}}\int K(u) u^l du\\
                                                                &\leq h^{\beta-s}c
                \end{align*}

                For \(c:=\int K(u) f_1(u) du\), showing the bias is bounded as desired.
        \end{itemize}
    \item[2.]
        \begin{itemize}
            \item[a)]
                \begin{align*}
                    L_F(x)&=\lim_{\epsilon\to 0} \frac{T((1-\epsilon)F+\epsilon\delta_x)-T(F)}{\epsilon} \\
                    L_F(x)&=\lim_{\epsilon\to 0} \frac{((1-\epsilon)F(b)+\epsilon\delta_x(b))-((1-\epsilon)F(a)+\epsilon\delta_x(a))-(F(b)-F(a))}{\epsilon} \\
                    L_F(x)&=\lim_{\epsilon\to 0} \frac{\epsilon(I_{[x,\infty)}(b)-I_{[x,\infty)}(a))-\epsilon (F(b)-F(a))}{\epsilon} \\
                    L_F(x)&= I_{[a,b)}(x)-T(F) \\
                \end{align*}
            \item[b)]
                \begin{align*}
                    \hat{se}(\hat \theta) &= \frac{\hat \tau}{\sqrt{n}} \\
                    \hat{se}(\hat \theta) &= \frac{1}{n^\frac{3}{2}}\sum_{i=1}^n(I_{[a,b)}(X_i)-T(\hat F_n))^2 \\
                    \hat{se}(\hat \theta) &= \frac{1}{n^\frac{3}{2}}\sum_{i=1}^n(I_{[a,b)}(X_i)-\hat F_n(b)+\hat F_n(a))^2 \\
                    \hat{se}(\hat \theta) &= \frac{1}{\sqrt{n}}\left((\hat F_n(b)-\hat F_n(a))(1-\hat F_n(b)+\hat F_n(a))^2+(1-\hat F_n(b)+\hat F_n(a))(\hat F_n(b)-\hat F_n(a))^2\right) \\
                    \hat{se}(\hat \theta) &= \frac{1}{\sqrt{n}}\left((\hat F_n(b)-\hat F_n(a)+1- \hat F_n(b)+\hat F_n(a))(\hat F_n(b)-\hat F_n(a))(1-\hat F_n(b)+\hat F_n(a))\right) \\
                    \hat{se}(\hat \theta) &= \frac{1}{\sqrt{n}}\left((\hat F_n(b)-\hat F_n(a))(1-\hat F_n(b)+\hat F_n(a))\right) \\
                \end{align*}
            \item[c)]
                An approximate \(1-\alpha\) confidence interval is given by 
                \[\hat \theta \pm z_{\frac{\alpha}{2}}\hat{se}(\hat \theta) = F_n(b)-\hat F_n(a) \pm z_{\frac{\alpha}{2}}\frac{1}{\sqrt{n}}\left((\hat F_n(b)-\hat F_n(a))(1-\hat F_n(b)+\hat F_n(a))\right)\]
            \item[d)]
                To estimate \(se(\hat \theta)\), one would draw a sample, with replacement, of $n$ from the $X_i$s. For each of these samples, $\hat\theta$ would be calculated (this is the portion of the sample in \([a,b)\). Then, \(se(\hat \theta)\) is the standard deviation of $\hat\theta$ among all the samples.

        \end{itemize}
\end{itemize}
\end{document}

### 2014-11-12 stat374/hw3.tex
Still working on HW3
Minor update hw5

@@
                 \end{align*}
-                Using the Lagrange form of the remainder of a taylor series, for some $\xi\in(x,x+hu)$,
+                Using the Lagrange form of the remainder of a Taylor series, for some $\xi\in(x,x+hu)$,
                 \[\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) = \frac{(hu)^l}{l!}p^{(l)}(\xi)\]
-                By the conditionts of the H{\"o}lder class,
+                By the conditions of the H{\"o}lder class,
                 \begin{align*}
@@
                 For \(c:=\int K(u) f_1(u) du\), showing the bias is bounded as desired.
+            \item[b)]
+                Combining the bounds on bias squared and variance from part a) to get the risk, we get
+                \begin{align*}
+                    \mbox{max}(\L^2(\hat P_{n,s}(x))) &= c^2h^{2(\beta-s)}+\frac{c'}{nh^{2s+1}} \\
+                    \frac{d}{dh}\mbox{max}((\L^2(\hat P_{n,s}(x))) &= 2(\beta-s)c^2h^{2(\beta-s)-1}-\frac{(2s+1)c'}{nh^{2s+2}} 
+                \end{align*}
+                Then, this risk bound is minimized when its derivative is at $0$, so for the optimal $h$,
+                \begin{align*}
+                    0 &= 2(\beta-s)c^2h^{2(\beta-s)-1}-\frac{(2s+1)c'}{nh^{2s+2}} \\
+                    \frac{(2s+1)c'}{2(\beta-s)c^2}\frac{1}{n} &= h^{2\beta+1} \\
+                    \left(\frac{(2s+1)c'}{2(\beta-s)c^2}\right)^\frac{1}{2\beta+1}n^{-\frac{1}{2\beta+1}} &= h 
+                \end{align*}
+                Plugging this into the original risk function, and defining a new constant \(c_1=\left(\frac{(2s+1)c'}{2(\beta-s)c^2}\right)^\frac{1}{2\beta+1}\), we get
+                \begin{align*}
+                    \mbox{max}(\L^2(\hat P_{n,s}(x)))  &= c^2c_1^{2(\beta-s)}n^{\frac{2(\beta-s)}{2\beta+1}}-\frac{c'}{c^{2s+1}}n^{\frac{2s+1-2\beta+1}{2\beta+1}} \\
+                    \mbox{max}(\L^2(\hat P_{n,s}(x))) &= \left(c^2c_1^{2(\beta-s)}+\frac{c'}{c^{2s+1}}\right)n^{\frac{2(\beta-s)}{2\beta+1}}
+                \end{align*}
+                So the maximum risk is of order \(O\left(n^{\frac{2(\beta-s)}{2\beta+1}}\right)\).
         \end{itemize}
@@
         \end{itemize}
+    \item[3.]
+        \begin{itemize}
+            \item[a)]
+                I fit a kernel density estimate with a Gaussian kernel and bandwidth chosen by the Normal Reference Rule.
+I didn't use cross validation, since the rounding in the magnitudes made it so that cross validation would default to an arbitrarily small bandwidth (as we discovered in our last homework).
+                \begin{center}
+                    \includegraphics[width=9cm]{hw3_3_a}
+                \end{center}
+            \item[b)]
+                Below I have plotted the estimated CDF along with a $95\%$ confidence interval.
+This band is generated from \(\hat F(x) \pm \sqrt{\frac{1}{2n}\log\left(\frac{2}{.05}\right)}\), winsorized above at $1$ and below at $0$.
+                \begin{center}
+                    \includegraphics[width=9cm]{hw3_3_b}
+                \end{center}
+            \item[b)]
+                Using the density from problem a, we get estimate for \(F(4.6)-F(4.3)\) of 
+                \[\int_{4.3}^{4.6} f(x)dx = .533 \]
+                Using the plug-in estimator, we get \(\hat F(4.6) - \hat F(4.3)=.526\).
+Then, using the result from problem 2, we get an approximate confidence interval of \(.526\pm z_{.025}\frac{.526(1-.526)}{\sqrt{1000}} = (.522,.530)\).
+
+        \end{itemize}
+    \item[4.]
+        \begin{itemize}
+            \item[a)]
+                To get the analytical distribution of \(\hat \theta\), its easiest to start by calculating the distribution function. For any $x\in[0,\theta]$
+                \begin{align*}
+                    P(\hat \theta \leq x) &= P(X_1\leq x \cap ... \cap X_n\leq x) \\
+                    P(\hat \theta \leq x) &= \prod_{i=1}^n P(X_i\leq x) \\
+                    P(\hat \theta \leq x) &= \prod_{i=1}^n \left(\frac{x}{\theta}\right)^n
+                \end{align*}
+                Differentiating both sides, we get the distribution function \(f(x)=\frac{nx^{n-1}}{\theta^n}\).
+Below, I drew a sample of $50$ from a uniform with $\theta=1$, and then preformed a parametric and nonparametric bootstrap of $\hat\theta$: \\
+                \includegraphics[width=9cm]{hw3_4_a_np}
+                \includegraphics[width=9cm]{hw3_4_a_p} \\
+                As you can see, the parametric bootstrap gives a distribution fairly close to the analytic one (just shifted a little to the left), while the nonparametric bootstrap fails to do so, having the majority of the point mass at the max from the original sample.
+            \item[b]
+                Since in the parametric case \(X_i^*\sim \mbox{Uniform}(0,\hat\theta)\),
+                \begin{align*}
+                    P(\hat\theta^*=\hat\theta) &= P(\exists i, X_i^*=\hat\theta\cap \forall j, X_j^*\leq\hat\theta) \\
+                    &\leq P(\exists i, X_i^*=\hat\theta) \\
+                    &\leq \prod_{i=1}^n (1-P(X_i^*=\hat\theta)) \\
+                    &\leq 0
+                \end{align*}
+                On the other hand, in the nonparametric case,
+                \begin{align*}
+                    P(\hat\theta^*=\hat\theta) &= P(\exists i, X_i^*=\hat\theta\cap \forall j, X_j^*\leq\hat\theta) \\
+                    &= P(\exists i, X_i^*=\hat\theta) \\
+                    &= 1-P(\forall i, X_i^*\neq\hat\theta) \\
+                    &= 1-\left(1-\frac{1}{n}\right)^{n} \\
+                \end{align*}
+                For $n=50$, this gives us $.636$, but in the limiting case, we get 
+                \begin{align*}
+                    \lim_{n\to\infty}P(\hat\theta^*=\hat\theta)&= 1-\lim_{n\to\infty}\left(1-\frac{1}{n}\right)^{n}  \\
+                                                               &= 1-\lim_{n\to\infty}\left(\frac{n}{n-1}\right)^{-n}  \\
+                                                               &= 1-\lim_{n\to\infty}\left(1+\frac{1}{n-1}\right)^{-n}  \\
+                                                               &= 1-\lim_{n\to\infty}\left(1+\frac{1}{n-1}\right)\left(1+\frac{1}{n-1}\right)^{-n+1}  \\
+                                                               &= 1-\frac{1}{e}  \\
+                                                               &\approx .632  \\
+                \end{align*}
+        \end{itemize}
+    \item[5.]
+        \begin{itemize}
+            \item[a)]
+                For this model, the bootstrapped $95\%$ confidence interval for \(\theta\) I got was \((.240,.282)\), and the "true" $\theta$, which was the result of running the experiment for $n=500$, was $.264$.
+Thus, the confidence interval included the "true" parameter, and also the real \(\theta=.25\).
+            \item[b)]
+                For this model, my confidence interval was \(-.280,.120)\) and the "true" $\theta$ was \(.023\).
+Thus, the confidence interval covered the "true" parameter and also the real parameter \(\theta=0\).
+            \item[c)]
+                For this model, my confidence interval was \((.282,.482)\) with the "true" $\theta=.865\).
+Thus, my confidence interval was far below the "true" $\theta$, and even farther below the actual $\theta$ of $0$.
+        \end{itemize}
+
 \end{itemize}

### 2014-11-13 stat374/hw3.tex
done with hw3

@@
                 \begin{align*}
-                    \E[\hat p_{n,s}(x)] &= \E\left[\frac{1}{nh^{s+1}}\sum_{i=1}^n K\left(\frac{X_i-x}{h}\right)\right] \\
-                                        &= \frac{1}{h^{s+1}}\int K\left(\frac{u-x}{h}\right)p(u) du\\
-                                        &= \frac{1}{h^{s}}\int K\left(u\right)p(x+hu) du\\
-                                        &= \frac{1}{h^{s}}\int K(u)\left(\sumj \frac{(hu)^j}{j!}p^{(j)}(x)\right) du\\
-                                        &= \sum_{j=1}^{l-1}\frac{h^{j-s}}{j!}p^{(j)}(x)\int K(u)u^j du + \frac{1}{h^{s}}\int K(u)\left(\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x)\right) du\\
-                                        &= p^{(s)}(x)+ \frac{1}{h^{s}}\int K(u)\left(\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x)\right) du
+                    \E[\hat p_{n,s}(x_0)] &= \E\left[\frac{1}{nh^{s+1}}\sum_{i=1}^n K\left(\frac{X_i-x_0}{h}\right)\right] \\
+                                        &= \frac{1}{h^{s+1}}\int K\left(\frac{u-x_0}{h}\right)p(u) du\\
+                                        &= \frac{1}{h^{s}}\int K\left(u\right)p(x_0+hu) du\\
+                                        &= \frac{1}{h^{s}}\int K(u)\left(\sum_{j=0}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0)\right) du\\
+                                        &= \sum_{j=0}^{l-1}\frac{h^{j-s}}{j!}p^{(j)}(x_0)\int K(u)u^j du + \frac{1}{h^{s}}\int K(u)\left(\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0)\right) du\\
+                                        &= p^{(s)}(x_0)+ \frac{1}{h^{s}}\int K(u)\left(\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0)\right) du
                 \end{align*}
-                Using the Lagrange form of the remainder of a Taylor series, for some $\xi\in(x,x+hu)$,
-                \[\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) = \frac{(hu)^l}{l!}p^{(l)}(\xi)\]
+                Using the Lagrange form of the remainder of a Taylor series, for some $\xi\in(x_0,x_0+hu)$,
+                \[\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0) = \frac{(hu)^l}{l!}p^{(l)}(\xi)\]
                 By the conditions of the H{\"o}lder class,
                 \begin{align*}
-                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\leq \frac{(hu)^l}{l!}(L\vert \xi - x\vert^{\beta-l} + \vert p(x)\vert) \\
-                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\leq \frac{(hu)^l}{l!}(L\vert hu\vert^{\beta-l}  + \vert p(x)\vert) \\
-                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\leq h^\beta\frac{u^l\vert u \vert^{\beta-l}L}{l!}+ u^l\frac{h^l}{l!}\vert p(x)\vert \\
-                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\leq h^\beta f_1(u)+ u^lf_2(x) \\
+                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0) &\leq \frac{(hu)^l}{l!}(L\vert \xi - x_0\vert^{\beta-l} + \vert p(x_0)\vert) \\
+                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0) &\leq \frac{(hu)^l}{l!}(L\vert hu\vert^{\beta-l}  + \vert p(x_0)\vert) \\
+                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0) &\leq h^\beta\frac{u^l\vert u \vert^{\beta-l}L}{l!}+ u^l\frac{h^l}{l!}\vert p(x_0)\vert \\
+                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0) &\leq h^\beta f_1(u)+ u^lf_2(x_0) \\
                 \end{align*}
-                For \(f_1(u):=\frac{u^l\vert u \vert^{\beta-l}L}{l!}\) and \(f_2(x):=\frac{h^l}{l!}\vert p(x)\vert\). By similar logic, we also get
+                For \(f_1(u):=\frac{u^l\vert u \vert^{\beta-l}L}{l!}\) and \(f_2(x_0):=\frac{h^l}{l!}\vert p(x_0)\vert\). By similar logic, we also get
                 \begin{align*}
-                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\geq -\frac{(hu)^l}{l!}(L\vert \xi - x\vert^{\beta-l} + \vert p(x)\vert) \\
-                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x) &\geq -h^\beta f_1(u)- u^lf_2(x) 
+                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0) &\geq -\frac{(hu)^l}{l!}(L\vert \xi - x_0\vert^{\beta-l} + \vert p(x_0)\vert) \\
+                    \sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0) &\geq -h^\beta f_1(u)- u^lf_2(x_0) 
                 \end{align*}
 
-                Plugging these into our expression for \(\E[\hat p_{n,s}(x)]\), we get that 
+                Plugging these into our expression for \(\E[\hat p_{n,s}(x_0)]\), we get that 
                 \begin{align*}
-                    \E[\hat p_{n,s}(x)] &= p^{(s)}(x)+ \frac{1}{h^{s}}\int K(u)\left(\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x)\right) du \\
-                    \vert \E[\hat p_{n,s}(x)]-p^{(s)}(x) \vert &\leq \frac{1}{h^{s}}\int K(u) (h^\beta f_1(u)+ u^lf_2(x)) du \\
-                                                                &\leq h^{\beta-s}\int K(u) f_1(u) du + \frac{f_2(x)}{h^{s}}\int K(u) u^l du\\
+                    \E[\hat p_{n,s}(x_0)] &= p^{(s)}(x_0)+ \frac{1}{h^{s}}\int K(u)\left(\sum_{j=l}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0)\right) du \\
+                    \vert \E[\hat p_{n,s}(x_0)]-p^{(s)}(x_0) \vert &\leq \frac{1}{h^{s}}\int K(u) (h^\beta f_1(u)+ u^lf_2(x_0)) du \\
+                                                                &\leq h^{\beta-s}\int K(u) f_1(u) du + \frac{f_2(x_0)}{h^{s}}\int K(u) u^l du\\
                                                                 &\leq h^{\beta-s}c
@@
 
-                For \(c:=\int K(u) f_1(u) du\), showing the bias is bounded as desired.
+                For \(c:=\int K(u) f_1(u) du\), showing the bias is bounded as desired. To bound the variance term, we have that
+                \begin{align*}
+                    \var\left(\hat p_{n,s}(x_0)\right) &= \frac{1}{nh^{2s+2}}\var\left(K\left(\frac{X_i-x_0}{h}\right)\right) \\
+                                                     &= \frac{1}{nh^{2s+2}}\left(\E\left[K\left(\frac{X_i-x_0}{h}\right)^2\right]-\E\left[K\left(\frac{X_i-x_0}{h}\right)\right]^2\right) \\
+                                                     &\leq \frac{1}{nh^{2s+2}}\left(h\int K^2(u)\left(\sum_{j=0}^\infty \frac{(hu)^j}{j!}p^{(j)}(x_0)\right)du-\left(h^{s+1}p^{(s)}(x_0)+ h^{\beta+1}c\right)^2\right) \\
+                                                     &\leq \frac{p(x_0)}{nh^{2s+1}}\int K^2(u)du + O\left(\frac{1}{n}\right)
+                \end{align*}
+                So \(\var\left(\hat p_{n,s}(x_0)\right)\) is bounded by \(\frac{c'}{nh^{2s+1}}\) for \(c'>0\).
             \item[b)]

### 2014-11-19 stat343/hw6.tex
4/5 done with homework 6

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 343, Homework 6 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
\begin{itemize}
    \item[1.] Problem 5.3 in Weisberg \\
        I got an estimate of \(\hat\sigma^2=1860.276\), so for the test statistic I got that \(G=\frac{\sum_{i=1}^n(\hat y_i - \tilde y_i)^2}{\hat\sigma^2}=15.296\). Then, running the bootstrap algorithm with $1000$ bootstraps, I got a significance level of  $.002$, since only $\frac{1}{1000}$ of the bootstrap samples resulted in a higher test statistic.
    \item[2.]
        \begin{itemize}
            \item[a)]
                \begin{align*}
                    X_{(i)}^TX_{(i)}&=\sum_{j\neq i} x_j^Tx_j \\
                    X_{(i)}^TX_{(i)}&= X^TX - x_i^Tx_i \\
                    (X_{(i)}^TX_{(i)})^{-1}&= (X^TX - x_i^Tx_i)^{-1} \\
                \end{align*}
                Applying the Sherman-Morrison formula to the right hand side, we get
                \begin{align*}
                    (X_{(i)}^TX_{(i)})^{-1}&= (X^TX)^{-1} + \frac{(X^TX)^{-1}x_i^Tx_i(X^TX)^{-1}}{1-x_i(X^TX)^{-1}x_i^T} \\
                    (X_{(i)}^TX_{(i)})^{-1}&= (X^TX)^{-1} + \frac{(X^TX)^{-1}x_i^Tx_i(X^TX)^{-1}}{1-h_{ii}} \\
                \end{align*}
            \item[b)]
                \begin{align*}
                    \hat \beta_{(i)} &= (X_{(i)}^TX_{(i)})^{-1}X_{(i)}^T Y_{(i)}   \\
                    \hat \beta_{(i)} &= \left((X^TX)^{-1} + \frac{(X^TX)^{-1}x_i^Tx_i(X^TX)^{-1}}{1-h_{ii}} \right)\left(X^TY - x_i^Ty_i\right)  \\
                    \hat \beta_{(i)} &= (X^TX)^{-1}\left(X^TY - x_i^Ty_i\right) + \frac{(X^TX)^{-1}x_i^Tx_i(X^TX)^{-1}}{1-h_{ii}}\left(X^TY - x_i^T y_i\right)   \\
                    \hat \beta_{(i)} &= \hat \beta - (X^TX)^{-1}x_i^Ty_i + \frac{(X^TX)^{-1}x_i^Tx_i(X^TX)^{-1}\left(X^TY - x_i^T y_i\right)}{1-h_{ii}}   \\
                    \hat \beta_{(i)} &= \hat \beta + \frac{(X^TX)^{-1}x_i^Ty_i(h_{ii}-1) +(X^TX)^{-1}x_i^T\left(x_i\hat\beta - h_{ii}y_i\right)}{1-h_{ii}}   \\
                    \hat \beta_{(i)} &= \hat \beta + \frac{(X^TX)^{-1}x_i^T\left(x_i\hat\beta - h_{ii}y_i + (1-h_{ii})y_i\right)}{1-h_{ii}}   \\
                    \hat \beta_{(i)} &= \hat \beta + \frac{(X^TX)^{-1}x_i^T\left(x_i\hat\beta - y_i\right)}{1-h_{ii}} \\
                    \hat \beta_{(i)} &= \hat \beta - \frac{(X^TX)^{-1}x_i^T\hat e_i}{1-h_{ii}}   
                \end{align*}
            \item[c)]
                \begin{align*}
                    y_i - x_i\hat\beta_{(i)} &= y_i- x_i\left(\hat \beta - \frac{(X^TX)^{-1}x_i^T\hat e_i}{1-h_{ii}}\right) \\
                    y_i - x_i\hat\beta_{(i)} &= y_i  - \hat y_i + \frac{h_{ii}\hat e_i}{1-h_{ii}} \\
                    y_i - x_i\hat\beta_{(i)} &= \hat e_i \left(1+ \frac{h_{ii}}{1-h_{ii}}\right) \\
                    y_i - x_i\hat\beta_{(i)} &= \hat e_i \frac{1}{1-h_{ii}} \\
                \end{align*}
            \item[d)]
                \begin{align*}
                    D_i &= \frac{\|\hat Y_{(i)} - \hat Y\|^2}{p\hat\sigma^2} \\
                    D_i &= \frac{\|X\beta_{(i)} - X\beta\|^2}{p\hat\sigma^2} \\
                    D_i &= \frac{\|X\left(\hat \beta - \frac{(X^TX)^{-1}x_i^T\hat e_i}{1-h_{ii}}\right) - X\beta\|^2}{p\hat\sigma^2} \\
                    D_i &= \frac{\|- X\frac{(X^TX)^{-1}x_i^T\hat e_i}{1-h_{ii}}\|^2}{p\hat\sigma^2} \\
                    D_i &= \frac{\frac{\hat e_ix_i(X^TX)^{-1}}{1-h_{ii}}X^TX\frac{(X^TX)^{-1}x_i^T\hat e_i}{1-h_{ii}}}{p\hat\sigma^2} \\
                    D_i &= \frac{\frac{\hat e_ix_i(X^TX)^{-1}}{1-h_{ii}}X^TX\frac{(X^TX)^{-1}x_i^T\hat e_i}{1-h_{ii}}}{p\hat\sigma^2} \\
                    D_i &= \frac{\frac{\hat e_ix_i(X^TX)^{-1}}{1-h_{ii}}\frac{x_i^T\hat e_i}{1-h_{ii}}}{p\hat\sigma^2} \\
                    D_i &= \frac{1}{p}\frac{\hat e_i^2}{\hat\sigma^2(1-h_{ii})}\frac{x_i(X^TX)^{-1}x_i^T}{1-h_{ii}} \\
                    D_i &= \frac{1}{p} r_i^2 \frac{h_ii}{1-h_{ii}} 
                \end{align*}
        \end{itemize}
    \item[3)]
        \begin{itemize}
            \item[a)]
                The two graphs make it pretty clear there is autocorrelated errors. The first, "Residual by Year", seems to show the errors moving up and down together, rather than being the cloud centered around $0$ if there was no autocorrelation. The second, "Residual vs Residual With Lag of 1" shows each residual plotted against the next residual. The evidence of autocorrelation is overwhelming here, with there being a clear positive linear relationship between one error and the next. \\
                \includegraphics[width=9cm]{hw6_3_a_res} 
                \includegraphics[width=9cm]{hw6_3_a_auto}  \\
            \item[b)]
                Running the AR1 model, we see an estimated autocorrelation of $.972$, which is extremely significant. We also see unemployed switch to being significant with a p-value of $.02$ in the AR1 model while just outside of of significance at $.051$ in the ordinary linear model. Much more importantly though, the coefficient switches sign, speaking to the autocorrelation having a meaningful effect.\\
                Confidence interval and estimate for autocorrelation: \\

                \input{hw6_3_alm_acor}

                Coefficients for autocorrelated model: \\

                \input{hw6_3_alm_coef}

                Coefficients for regular linear model: \\
                \input{hw6_3_lm_coef}

                Only one coefficient, that of unemployed, switched to the other side of the $.05$ threshold with the auto correlated term, but the estimate and p-value of all the other variables changes noticeably, speaking to how not dealing with autocorrelation can invalidate a model.

            \item[c)]
                My suspicion is that the autocorrelation largely has to do with the relative size of different generations. The portion of the US population in different age groups varies a lot over time, and divorce rates are likely quite different between different age groups. Since the relative portion of people in each age group is highly auto-correlated (since everyone either dies or gets just one year older), our divorce rate is also auto-correlated.
        \end{itemize}
    \item[4)]
        It looks like we have a lack of fit since the variance isn't constant. Looking at the first plot, the residual plot, nothing is glaringly wrong, but it looks like the points farther to the right are somewhat more spread out. Plotting the square root transformed residuals against the predicted values, and adding a trend line, we see a much more obvious increase in variance over the range, with a clear upwards trend. Testing this formally with the variance test, where we split the residuals into those with speed above and below 14, we get a ratio of $.206$, with an associated p-value of $.0009$, which seals it. We also appear to have a mildly non-linear response. \\
        \includegraphics[width=9cm]{hw6_4_res} 
        \includegraphics[width=9cm]{hw6_4_sqrtres}  \\
             As far as other issues, none jump out. From the residual plot, no residual looks extreme enough to be an outlier. Running a QQ plot, the residuals on the whole look roughly normal. \\
        \begin{center}
            \includegraphics[width=9cm]{hw6_4_qq} 
        \end{center}
        The one potential issue is that the true relationship between the predictor and outcome may not quite be linear. Looking at the linear fit versus a kernel regression fit, it seems like the function may actual be somewhat convex. 
        \begin{center}
            \includegraphics[width=9cm]{hw6_4_ker} 
        \end{center}
        Attempting to approximate this with a linear spline, we allow for a different slope above and below a speed of $20$, where we seem to have a mild kink. When we do this, and run we ANOVA, we just get significant decrease in variance, with a p-value of $.0378$. \\
        \input{hw6_4}
    \item[5)]

\end{itemize}

\end{document}

### 2014-11-19 stat343/hw6.R

@@ -1,0 +1,112 @@
+library('parallel')
+library('alr3')
+library('faraway')
+library('car')
+library('MASS')
+library('locfit')
+library('xtable')
+library('nlme')
+
+#############################
+# Problem 1
+#############################
+
+data(ufcgf)
+attach(ufcgf)
+
+### 5.3.1.1
+fir.lm<-lm(Height~Dbh)
+fir.loess<-loess(Height~Dbh)
+
+G<-sum((predict(fir.lm)-predict(fir.loess))^2)/(summary(fir.lm)$sigma^2)
+
+### 5.3.1.2-4
+B<-1000
+Yhat<-predict(fir.lm)
+ehat<-residuals(fir.lm)
+
+res.boot<- function(c,x,y,e) {
+    ystar<-y+sample(e,length(y),replace=TRUE)
+    lm.m    <-lm(ystar~x)
+    loess.m <-loess(ystar~x)
+    Gstar<-sum((predict(lm.m)-predict(loess.m))^2)/(summary(lm.m)$sigma^2)
+    return(Gstar)
+}
+res.boots<-mcmapply(res.boot,c=1:B,MoreArgs=list(x=Dbh,y=Yhat,e=ehat),mc.cores=4)
+fir.p<-sum(res.boots>G)/B
+
+#############################
+# Problem 3 
+#############################
+
+data(divusa)
+attach(divusa)
+n<-nrow(divusa)
+div.lm<-lm(divorce~unemployed+femlab+marriage+birth+military)
+
+### a)
+pdf("hw6_3_a_res.pdf")
+plot(year,residuals(div.lm),ylab="Residual",xlab="Year",main="Residual by Year")
+dev.off()
+
+pdf("hw6_3_a_auto.pdf")
+plot(tail(residuals(div.lm),n-1)~head(residuals(div.lm),n-1),xlab=expression(hat(epsilon)[i]),ylab=expression(hat(epsilon)[i+1]),main="Residual vs Residual With Lag of 1")
+dev.off()
+
+### b)
+div.alm<-gls(divorce~unemployed+femlab+marriage+birth+military,correlation=corAR1(form=~year),method="ML")
+
+d<- 3
+print(xtable(summary(div.alm)$tTable,digits=d),type="latex",file="hw6_3_alm_coef.tex")
+print(xtable(intervals(div.alm)$corStruct,digits=d),type="latex",file="hw6_3_alm_acor.tex")
[…]

### 2014-11-19 stat343/hw6.tex
Finnished hw6

@@
 %% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
-\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
+\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings, placeins}
 \newcommand{\Z}{\mathbb{Z}}
@@
                     \hat \beta_{(i)} &= \hat \beta + \frac{(X^TX)^{-1}x_i^Ty_i(h_{ii}-1) +(X^TX)^{-1}x_i^T\left(x_i\hat\beta - h_{ii}y_i\right)}{1-h_{ii}}   \\
-                    \hat \beta_{(i)} &= \hat \beta + \frac{(X^TX)^{-1}x_i^T\left(x_i\hat\beta - h_{ii}y_i + (1-h_{ii})y_i\right)}{1-h_{ii}}   \\
+                    \hat \beta_{(i)} &= \hat \beta + \frac{(X^TX)^{-1}x_i^T\left(x_i\hat\beta - h_{ii}y_i + (h_{ii}-1)y_i\right)}{1-h_{ii}}   \\
                     \hat \beta_{(i)} &= \hat \beta + \frac{(X^TX)^{-1}x_i^T\left(x_i\hat\beta - y_i\right)}{1-h_{ii}} \\
@@
                 The two graphs make it pretty clear there is autocorrelated errors.
-The first, "Residual by Year", seems to show the errors moving up and down together, rather than being the cloud centered around $0$ if there was no autocorrelation.
-The second, "Residual vs Residual With Lag of 1" shows each residual plotted against the next residual.
+The first, ``Residual by Year", seems to show the errors moving up and down together, rather than being the cloud centered around $0$ if there was no autocorrelation.
+The second, ``Residual vs Residual With Lag of 1" shows each residual plotted against the next residual.
 The evidence of autocorrelation is overwhelming here, with there being a clear positive linear relationship between one error and the next.
@@
                 Confidence interval and estimate for autocorrelation: \\
-
+                \FloatBarrier
                 \input{hw6_3_alm_acor} 
+                \FloatBarrier
 
                 Coefficients for autocorrelated model: \\
-
+                \FloatBarrier
                 \input{hw6_3_alm_coef}
+                \FloatBarrier
 
                 Coefficients for regular linear model: \\
+                \FloatBarrier
                 \input{hw6_3_lm_coef}
+                \FloatBarrier
 
@@
     \item[5)]
+        In general, it seems to be the case that without dealing with outliers, one can get meaningfully different results between robust regression and non-robust regression.
+However, when outliers are taken out, they give much more similar results.
+We see this most strongly with the coefficient for air flow.
+With the outlier in, we get coefficients of about $.8$ with robust regression and a smaller, though still significant, coefficient of $.7$ with regular linear regression.
+When an outlier is removed, all the methods get about $.8$.\\
+
+        \smallskip
+        With the outliers in, we got the results below.
+In addition to the difference with air flow noted above, the coefficient with water temp seems to vary a lot between methods, though is always significant, while  acid concentration is significant in every one except the least absolute deviation, where its just barely significant.
+        \begin{itemize}
+            \item[a)] Regular least squares: \\
+                \FloatBarrier
+                \input{hw6_5_lm}
+                \FloatBarrier
+            \item[b)] Least absolute deviation: \\
+                \FloatBarrier
+                \input{hw6_5_lad}
+                \FloatBarrier
+            \item[c)] Hubber method:\\
+                \FloatBarrier
+                \input{hw6_5_hub}
+                \FloatBarrier
+            \item[d)] Least trimmed squares: \\
+                \FloatBarrier
+                \input{hw6_5_lts}
+                \FloatBarrier
+        \end{itemize}
+        Looking at the Cook's Distances, its clear that the $21$st observation is immensely influential.
+$3$, $4$, and $1$ also have relatively high Cook's Distances, but they don't seem that out of line with the rest of them.
+Double checking the studentized residuals, $21$ a value of $3.330$, while the next highest is $2.052$.
+Since the Bonferroni critical value at a p-value of $.05$ is $3.256$, I concluded that $21$ is worth removing, while the rest can remain.
+        \begin{center}
+            \includegraphics[width=9cm]{hw6_5_halfnorm} 
+        \end{center}
+        Looking over the effects of the inclusion of each observation on the coefficients, we see that $21$ has a much bigger effect over air flow and water temp than other variables, further supporting that it should be removed as an outlier.
+\\
+            \includegraphics[width=9cm]{hw6_5_af} 
+            \includegraphics[width=9cm]{hw6_5_wt} \\
+
+        Rerunning each of the regressions with $21$ removed, we get the results below.
+The air flow coefficient is now fairly consistent between the robust and non-robust regressions.
+The water temp coefficient is still meaningfully different between the robust and non-robust regressions, though much smaller than before.
+It may be that observation $4$, as seen above, is still having undue influence.
+Finally, acid concentration continues to be insignificant, or just barely significant.
+        \begin{itemize}
+            \item[a)] Regular least squares: \\
+                \FloatBarrier
+                \input{hw6_5_lm_nout}
+                \FloatBarrier
+            \item[b)] Least absolute deviation: \\
+                \FloatBarrier
+                \input{hw6_5_lad_nout}
+                \FloatBarrier
+            \item[c)] Hubber method:\\
+                \FloatBarrier
+                \input{hw6_5_hub_nout}
+                \FloatBarrier
+            \item[d)] Least trimmed squares: \\
+                \FloatBarrier
+                \input{hw6_5_lts_nout}
+                \FloatBarrier
+        \end{itemize}
 

### 2014-11-19 stat343/hw6.R

@@ -6,6 +6,10 @@ library('MASS')
 library('locfit')
 library('xtable')
 library('nlme')
+library('quantreg')
+
+#digits of precision
+d<- 3
 
 #############################
 # Problem 1
@@ -56,7 +60,6 @@ dev.off()
 ### b)
 div.alm<-gls(divorce~unemployed+femlab+marriage+birth+military,correlation=corAR1(form=~year),method="ML")
 
-d<- 3
 print(xtable(summary(div.alm)$tTable,digits=d),type="latex",file="hw6_3_alm_coef.tex")
 print(xtable(intervals(div.alm)$corStruct,digits=d),type="latex",file="hw6_3_alm_acor.tex")
 print(xtable(summary(div.lm),digits=d),type="latex",file="hw6_3_lm_coef.tex")
@@ -84,7 +87,6 @@ var.test(residuals(car.lm)[speed<14],residuals(car.lm)[speed>=14])
 pdf("hw6_4_qq.pdf")
 qqnorm(residuals(car.lm),ylab="Residuals",main="Residual vs Normal Qunatiles")
 qqline(residuals(car.lm))
-line(x)
 dev.off()
 
 pdf("hw6_4_ker.pdf")
@@ -100,6 +102,61 @@ speed.high<-ifelse(higher,speed-s,0)
 car.lm2<-lm(dist~speed.low + speed.high)
 ana<-anova(car.lm,car.lm2)
 print(xtable(ana,digits=3),type="latex",file="hw6_4.tex")
+
+#############################
+# Problem 5 
+#############################
+data(stackloss)
+
+sl.lm<-lm(stack.loss~Acid.Conc.+Water.Temp+Air.Flow,data=stackloss)
+print(xtable(summary(sl.lm)$coefficients,digits=d),type="latex",file="hw6_5_lm.tex")
+
+sl.hub<-rlm(stack.loss~Acid.Conc.+Water.Temp+Air.Flow,data=stackloss)
+print(xtable(summary(sl.hub)$coefficients,digits=d),type="latex",file="hw6_5_hub.tex")
+
+sl.lts<-ltsreg(stack.loss~Acid.Conc.+Water.Temp+Air.Flow,data=stackloss,nsamp="exact")
[…]

### 2014-11-24 stat374/final_project.R
Added nonwork and test data set factors

@@ -8,8 +8,8 @@ library('reshape2')
 library('chron')
 
 # Turns on different parts of the code
-clean.data <- FALSE
-data.sum   <- TRUE 
+clean.data <- TRUE 
+data.sum   <- FALSE
 par.reg    <- FALSE
 nonpar.reg <- FALSE
 
@@ -36,6 +36,11 @@ if (clean.data) {
     ### Merge the two data sets. Weather is only missing for a few hours, so we just drop them
     rob.data <- merge(hourly.robberies,hourly.weather,by=c("date","hour"),all.y=TRUE,all.x=FALSE)
     rob.data$robberies[is.na(rob.data$robberies)] <- 0
+    # Mark weekend/holidays
+    rob.data$nonworkday <- is.weekend(rob.data$date) | is.holiday(rob.data$date)
+    # Pick a 20% subset as our test data set
+    rob.data$test.set <- FALSE
+    rob.data$test.set[sample.int(nrow(rob.data),size=floor(nrow(rob.data)*.2))] <- TRUE 
     rm(hourly.weather,hourly.robberies)
     save(rob.data,file="final_project/hourly_robberies.RData")
     

### 2014-11-25 stat343/hw7.tex
Started homework 7

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 343 Homework 6 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
\begin{itemize} 
    \item[1.]
        \begin{itemize} 
            \item[a)]
                When we run a linear regression, we see get a statistically significant positive coefficient on year. This at least is a pretty good indicator for some kind of increasing trend. However, looking at the plot, it looks like there is some kind of concavity at least during much of the 20th century, which doesn't support their being a simple linear trend. So, I would conclude there is an upwards trend that probably isn't quite linear.  
                \smallskip
                The estimate of coefficients:
                \FloatBarrier
                \input{hw7/hw7_1_a} 
                \FloatBarrier
                \begin{center}
                    \includegraphics[width=9cm]{hw7/hw7_1_a} 
                \end{center}
            \item[b)]
                When we run a linear model that includes an AR(1) term, we find a statistically significant autoregressive term, with an estimate autocorrelation of $.208$. However, even with this, the estimates for the normal coefficients are pretty much the same as with the simple linear regression, with the coefficient on year still at $.012$ and only a small change in the intercept. The plot also shows how the plot barely changed. So, the autocorrelation  doesn't change my opinion about the trend. 
                \smallskip
                The estimate of autocorrelation:
                \FloatBarrier
                \input{hw7/hw7_1_b_acor} 
                \FloatBarrier
                The estimate of coefficients:
                \FloatBarrier
                \input{hw7/hw7_1_b_ttable} 
                \FloatBarrier
                \begin{center}
                    \includegraphics[width=9cm]{hw7/hw7_1_b} 
                \end{center}

            \item[c)]
                Starting with orthogonal polynomials of degree one through ten and backwards eliminating them with a critical value of $.2$, I we end up left with polynomials of degree one through five. The coefficients and plot are below. Looking at the plot, the model seems to fit quite well within the range of the data, but when we predict out to $2020$, we get a prediction of $60.078$, which is way outside of the range seen historically and thus not believable. This reflects the issues with polynomials only being a good representation of arbitrary functions where there is data to fit them.

                \smallskip
                Predicted value and range:
                \FloatBarrier
                \input{hw7/hw7_1_c_pred} 
                \FloatBarrier

                Estimated coefficients:
                \FloatBarrier
                \input{hw7/hw7_1_c} 
                \FloatBarrier
                \begin{center}
                    \includegraphics[width=9cm]{hw7/hw7_1_c} 
                \end{center}

            \item[d)]
                To test this trend, we fit a simple linear spline on the data, with an intercept and a continuous term for year which is $0$ before 1930. Looking at the results, we get a statistically significant coefficient on year after 1930. However, this coefficient is pretty close to the coefficient on year we had in the simple linear model, and the fit, when plotted, looks like a less convincing fit than the original simple linear model, with the actual values not symmetrically distributed around the prediction in the flat section. We don't have a formal statistical test to reject this model, but it certainly doesn't seem like an improvement over the simple linear model.

                \smallskip
                Estimated coefficients:
                \FloatBarrier
                \input{hw7/hw7_1_d} 
                \FloatBarrier
                \begin{center}
                    \includegraphics[width=9cm]{hw7/hw7_1_d} 
                \end{center}

            \item[e)]
                Fitting a cubic spline model, we certainly seem to see an improvement over the simple linear model. It picks up the apparent concavity over most of the 20th century, and also seems to capture the behavior towards the boundary reasonably, besides possibly turning up too sharply at the high end of the domain. Its fit is fairly similar to the polynomial fit though, with the exception of not having an inflection point on the lower end of the domain. There isn't much data in that section though, so it looks more like the polynomial is over fitting than the cubic splines.

                \smallskip
                Estimated coefficients:
                \FloatBarrier
                \input{hw7/hw7_1_e} 
                \FloatBarrier
                \begin{center}
                    \includegraphics[width=9cm]{hw7/hw7_1_e} 
                \end{center}
        \end{itemize}
    \item[2.]
    \item[4.]
        As I think we had to show on the midterm,
        \begin{align*}
            H&= X(X^TX)^{-1}X^T //
            H&= QR(R^TQ^TQR)^{-1}R^TQ^T //
            H&= QR(R^TR)^{-1}R^TQ^T //
            H&= QRR^{-1}(R^T)^{-1}R^TQ^T //
            H&= QQ^T
        \end{align*}

\end{itemize}
\end{document}

### 2014-11-25 stat343/hw7.R

@@ -1,0 +1,84 @@
+library('parallel')
+library('alr3')
+library('faraway')
+library('car')
+library('MASS')
+library('locfit')
+library('xtable')
+library('nlme')
+library('quantreg')
+library('splines')
+
+#digits of precision
+d<- 3
+
+#####################
+# Problem 1
+#####################
+
+data(aatemp)
+
+### a)
+temp.lm <- lm(temp~year,data=aatemp)
+print(xtable(summary(temp.lm),digits=d),type="latex",file="hw7/hw7_1_a.tex")
+
+pdf("hw7/hw7_1_a.pdf")
+plot(aatemp,xlab="Year",ylab="Temperature",main="Linear Fit",xlim=c(1850,2000),pch=20)
+abline(temp.lm,xlim=c(1850,2000))
+dev.off()
+
+### b)
+temp.alm<-gls(temp~year,correlation=corAR1(form=~year),method="ML",data=aatemp)
+print(xtable(summary(temp.alm)$tTable,digits=d),type="latex",file="hw7/hw7_1_b_ttable.tex")
+print(xtable(intervals(temp.alm)$corStruct,digits=d),type="latex",file="hw7/hw7_1_b_acor.tex")
+
+pdf("hw7/hw7_1_b.pdf")
+matplot(aatemp$year, cbind(aatemp$temp,temp.alm$fit),type="pl",pch=20,lty=1,col=1,xlim=c(1850,2000),ylab="Temperature",xlab="Year",main="Autoregressive Fit")
+dev.off()
+
+### c)
+crit<-.2
+p<-1
+deg<-11
+while (p>crit) {
+    deg <- deg-1
+    temp.poly<-poly(aatemp$year,deg)
+    poly.data<-data.frame(temp=aatemp$temp,temp.poly=temp.poly)
+    temp.poly.lm <- lm(temp ~ .,data=poly.data)
+    p<-coef(summary(temp.poly.lm))[deg+1,4]
+}
+print(xtable(summary(temp.poly.lm),digits=d),type="latex",file="hw7/hw7_1_c.tex")
+
+pdf("hw7/hw7_1_c.pdf")
+plot(aatemp,xlab="Year",ylab="Temperature",main="Polynomial Fit",xlim=c(1850,2000),pch=20)
[…]

### 2014-11-25 stat343/hw7.tex
hw7 question 4 almost done

@@
         \begin{align*}
-            H&= X(X^TX)^{-1}X^T //
-            H&= QR(R^TQ^TQR)^{-1}R^TQ^T //
-            H&= QR(R^TR)^{-1}R^TQ^T //
-            H&= QRR^{-1}(R^T)^{-1}R^TQ^T //
+            H&= X(X^TX)^{-1}X^T \\
+            H&= QR(R^TQ^TQR)^{-1}R^TQ^T \\
+            H&= QR(R^TR)^{-1}R^TQ^T \\
+            H&= QRR^{-1}(R^T)^{-1}R^TQ^T \\
             H&= QQ^T
         \end{align*}
+        As was discussed with added variable plots, for an additional variable $X_i$ added to a regression, if $e_{(-i)}$ are the residuals from the regression excluding $X_i$ and $e$ are the residuals from the regression including it,
+        \begin{align*}
+            e_{(-i)} &= \hat\beta_i (I-H) X_i + e \\ 
+            e_{(-i)} &= \hat\beta_i (I-QQ^T) X_i - e \\ 
+        \end{align*}
+        From simple linear regression, since \(e_{(-i)}\) has mean $0$, it must be the case that \(\hat\beta_i=\frac{\cov((I-QQ^T) X_i,e_{(-i)})}{\var((I-QQ^T) X_i)}\). Plugging this in, we get
+        \begin{align*}
+            e_{(-i)} &= \frac{\cov((I-QQ^T) X_i,e_{(-i)})}{\var((I-QQ^T) X_i)} (I-QQ^T) X_i - e \\ 
+            e &= e_{(-i)} - \frac{\cov\left((I-QQ^T) X_i,e_{(-i)}\right)}{\var\left((I-QQ^T) X_i\right)} (I-QQ^T) X_i \\ 
+            \|e\|^2 &= \left\|e_{(-i)} - \frac{\cov\left((I-QQ^T) X_i,e_{(-i)}\right)}{\var\left((I-QQ^T) X_i\right)} (I-QQ^T) X_i \right \|^2
+            RSS &= \left\|e_{(-i)} - \frac{\cov\left((I-QQ^T) X_i,e_{(-i)}\right)}{\var\left((I-QQ^T) X_i\right)} (I-QQ^T) X_i \right \|^2
+        \end{align*}
+        Thus,
 

### 2014-11-26 stat343/hw2.tex
added homework 2 and 5 tex files

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Homework 2 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[a)]
                \begin{align*}
                    \E[\hat{\beta} \vert X] &= \E[(X'X)^{-1}X'Y \vert X] \\
                                            &= \E[(X'X)^{-1}X'(X\beta+e) \vert X] \\
                                            &= \E[(X'X)^{-1}X'X\beta+(X'X)^{-1}X'e \vert X] \\
                                            &= \E[\beta \vert X] + (X'X)^{-1}X' \E[e \vert X] \\
                                            &= \beta + (X'X)^{-1}X' b(X)
                \end{align*}
                Thus, $\hat{\beta}$ is only unbiased when \((X'X)^{-1}X' b(X) = 0\).
                In other words, $\hat{\beta}$ is biased if  $b(X)$ is in the span of \((X'X)^{-1}X'\).
            \item[b)]
                \begin{align*}
                    \var[\hat{\beta} \vert X ] &= \var[(X'X)^{-1}X'Y \vert X ] \\
                                               &= \var[(X'X)^{-1}X'(X\beta+e) \vert X ] \\
                                               &= \var[(X'X)^{-1}X'X\beta+(X'X)^{-1}X'e \vert X] \\
                                               &= \var[\beta+(X'X)^{-1}X'e \vert X] \\
                                               &= \var[(X'X)^{-1}X'e \vert X] \\
                                               &= (X'X)^{-1}X'\var[e \vert X]((X'X)^{-1}X')' \\
                                               &= (X'X)^{-1}X'\sigma^2I X (X'X)^{-1} \\
                                               &= (X'X)^{-1}X'X(X'X)^{-1} \sigma^2 \\
                                               &= (X'X)^{-1}\sigma^2 \\
                \end{align*}
                Thus, though $\hat{\beta}$ may be biased, it has no higher variance with assumption A2$'$ than with the usual assumption A2.
        \end{itemize}
    \item[2.]
        \begin{itemize}
            \item[i.]
                Given that the variable \(wt_i = \beta_0 + e_i\), where each $e_i$ is normally distributed around $0$ and independent of all other $e_j$ such that $j\neq i$, and independent of the variable height, the probability that a model that included height would reduce the residual sum of squares as much as it has or more is $17.31\%$.
            \item[ii.]
                Given that the variable \(wt_i = \beta_0 + \beta_1 ht_i + e_i\), where each $e_i$ is normally distributed around $0$ and independent of all other $e_j$ such that $j\neq i$, the probability that \(\vert \hat\beta_0 \vert \geq 36.8759 \) given a null hypothesis that \(\beta_0=0\) is $58.3\%$.
           \item[iii.]
               Given that the variable \(wt_i = \beta_0 + \beta_1 ht_i + e_i\), where each $e_i$ is normally distributed around $0$ and independent of all other $e_j$ such that $j\neq i$, the probability that \(\vert \hat\beta_1 \vert \geq .5821  \) given a null hypothesis that \(\beta_1=0\) is $17.3\%$.
            \item[iv.]
                You can not compute the probability of the fitted model being the right one, only the probability of seeing the given data and estimates on it for various assumptions of the underlying process that generated it.
        \end{itemize}
    \item[3.]
        All the individuals formulas for the desired quantities are: \\
        \[SSreg = \frac{R^2 RSS}{1 - R^2}= \frac{.2084 \times 351191.4}{1-.2084}=92456.14\]
        \[RSS=(RSE \times df)^2=(2.993 \times 198)^2=351191.4\]
        \[MSreg = \frac{SSreg}{regDF} = \frac{92456.14}{1}= 92456.14\]
        \[\hat\sigma^2 = \frac{RSS}{resDF} = \frac{351191.41}{198} = 1773.694\]
        \[F = \frac{MSreg}{\hat\sigma^2} = \frac{92456.14}{1773.694} = 1773.694 \]
        Which fit in the table as such: \\
        \begin{tabular}{l | l l l l l}
            Regression & df      & ss               & ms       & F     \\
            \hline
            Regression & regDF = 1       & SSreg = 92456.14 & MSreg = 92456.14          & F = 52.126 \\
            Residual   & resDf = 198     & RSS = 351191.4   & $\hat\sigma^2$ = 1773.694 &  
        \end{tabular}
    \item[4.]
        \begin{itemize}
            \item[a)] These are the resulting confidence intervals: \\
                \begin{tabular}{l | r l r l}
                    Confidence Level & $\beta_0$ LB & $\beta_0$ UB & $\beta_1$ LB & $\beta_1$ UP \\
                    \hline
                    66\%             &       (3.802,& 3.828)       &       (-.062,& -.059)     \\
                    90\%             &       (3.792,& 3.837)       &       (-.063,& -.058)     \\
                    95\%             &       (3.788,& 3.841)       &       (-.064,& -.057)
                \end{tabular}
            \item[b)] We get the variance estimate:
            \[1.538\times10^{-5} = \sigma^2 \left[ \begin{array}{c c} 1 & 10 \end{array} \right] \left(X'X\right)^{-1} \left[ \begin{array}{c} 1 \\  10 \end{array} \right] \]
            \item[c)] This is what we end up with, for a 95\% CI: \\
                \begin{tabular}{l | r c l}
                    Type       & LB    & Fit   & UB    \\
                    \hline
                    Mean       & 3.005 & 3.022 & 3.038 \\
                    Prediction & 2.800 & 3.022 & 3.245 
                \end{tabular}
            \item[d)]
                The confidence interval is smallest near the mean of fixed acidity ($8.31$ in the data set). This stands to reason, since the confidence intervals are based on conditioned slices of a multivariate normal distribution; these slices will be steepest towards the center of the distribution (around the means of all variables), and flattest in the tails. In other words, its harder to screw up a prediction where you have a lot of data than where you don't have much data. \\
                \begin{center}
                    \includegraphics[width=12cm]{hw2_plot} 
                \end{center}
        \end{itemize}
        Here is my R code: \\
        \lstinputlisting{hw2.R}
\end{itemize}
\end{document}

### 2014-11-26 stat343/hw7.tex
added homework 2 and 5 tex files

--- 2014-11-25 stat343/hw7.tex
@@
     \item[2.]
+        \begin{itemize}
+            \item[a)]
+                After running the model on the untransformed data, we seem to see a substantial improvement from transformation.
+The log-likelihood for box-cox is maximized right around $\lambda=.25$, with a confidence interval that excluded $1$, so I transformed the data using $.25$.
+The end result is an improvement in $R^2$ from $.681$ to $.713$.
+                \smallskip
+                The untransformed linear model:
+                \FloatBarrier
+                \input{hw7/hw7_2_a}
+                \FloatBarrier                
+                \begin{center}
+                    \includegraphics[width=9cm]{hw7/hw7_2_a} 
+                \end{center}
+                The transformed linear model:
+                \FloatBarrier
+                \input{hw7/hw7_2_a_trans}
+                \FloatBarrier                
+            \item[b)]
+        \end{itemize}
+
     \item[4.]
@@
         \begin{align*}
-            H&= X(X^TX)^{-1}X^T //
-            H&= QR(R^TQ^TQR)^{-1}R^TQ^T //
-            H&= QR(R^TR)^{-1}R^TQ^T //
-            H&= QRR^{-1}(R^T)^{-1}R^TQ^T //
+            H&= X(X^TX)^{-1}X^T \\
+            H&= QR(R^TQ^TQR)^{-1}R^TQ^T \\
+            H&= QR(R^TR)^{-1}R^TQ^T \\
+            H&= QRR^{-1}(R^T)^{-1}R^TQ^T \\
             H&= QQ^T
         \end{align*}
+        As was discussed with added variable plots, for an additional variable $X_i$ added to a regression, if $\hat e_{(-i)}$ are the residuals from the regression excluding $X_i$ and $\hat e$ are the residuals from the regression including it,
+        \begin{align*}
+            \hat e_{(-i)} &= \hat\beta_i (I-H) X_i + \hat e \\ 
+            \hat e_{(-i)} &= \hat\beta_i (I-QQ^T) X_i + \hat e \\ 
+        \end{align*}
+        From simple linear regression, since \(\hat e_{(-i)}\) has mean $0$, it must be the case that \(\hat\beta_i=\frac{\cov((I-QQ^T) X_i,\hat e_{(-i)})}{\var((I-QQ^T) X_i)}\).
+Plugging this in, we get
+        \begin{align*}
+            \hat e_{(-i)} &= \frac{\cov((I-QQ^T) X_i,e_{(-i)})}{\var((I-QQ^T) X_i)} (I-QQ^T) X_i + \hat e \\ 
+            \hat e &= \hat e_{(-i)} - \frac{\cov\left((I-QQ^T) X_i,\hat e_{(-i)}\right)}{\var\left((I-QQ^T) X_i\right)} (I-QQ^T) X_i \\ 
+            \|\hat e\|^2 &= \left\|\hat e_{(-i)} - \frac{\cov\left((I-QQ^T) X_i,\hat e_{(-i)}\right)}{\var\left((I-QQ^T) X_i\right)} (I-QQ^T) X_i \right \|^2 \\
+            RSS &= \left\|\hat e_{(-i)} - \frac{\cov\left((I-QQ^T) X_i,\hat e_{(-i)}\right)}{\var\left((I-QQ^T) X_i\right)} (I-QQ^T) X_i \right \|^2
+        \end{align*}
+        Thus, we can minimize the $RSS$ by minimizing choosing the variable $X_i$ such that the quantity on the right is minimized.
 

### 2014-11-26 stat343/hw7.R

@@ -8,6 +8,8 @@ library('xtable')
 library('nlme')
 library('quantreg')
 library('splines')
+library('rpart')
+library('randomForest')
 
 #digits of precision
 d<- 3
@@ -77,6 +79,74 @@ print(xtable(summary(temp.cs.lm),digits=d),type="latex",file="hw7/hw7_1_e.tex")
 pdf("hw7/hw7_1_e.pdf")
 matplot(aatemp$year, cbind(aatemp$temp,temp.cs.lm$fit),type="pl",pch=20,lty=1,col=1,xlim=c(1850,2000),ylab="Temperature",xlab="Year",main="Cubic Spline Fit")
 dev.off()
+
+#####################
+# Problem 2 
+#####################
+data(ozone)
+
+### a)
+oz.lm<-lm(O3~temp+humidity+ibh, data=ozone)
+print(xtable(summary(oz.lm),digits=d),type="latex",file="hw7/hw7_2_a.tex")
+
+pdf("hw7/hw7_2_a.pdf")
+boxcox(oz.lm,plotit=TRUE, lambda=seq(0,.5,.1))
+title(main="Box-Cox")
+dev.off()
+
+lam<-.25
+oz.t.lm<-lm(O3^lam ~temp+humidity+ibh, data=ozone)
+print(xtable(summary(oz.t.lm),digits=d),type="latex",file="hw7/hw7_2_a_trans.tex")
+
+### b)
+oz.t.rt<-rpart(O3^lam ~temp+humidity+ibh, data=ozone)
+oz.t.rf<-randomForest(O3^lam ~temp+humidity+ibh, data=ozone)
+
+pdf("hw7/hw7_2_b_tree.pdf")
+plot(oz.t.rt,main="Regression Tree")
+text(oz.t.rt)
+dev.off()
+
+#Linear diagnoistics
+pre<-predict(oz.t.lm)
+res<-residuals(oz.t.lm)
+pdf("hw7/hw7_2_b_lm_res.pdf")
+plot(pre,res,xlab="Predicted",ylab="Residual",main="Linear Model Residual Plot")
+lines(predict(loess(res~pre)))
+dev.off()
+pdf("hw7/hw7_2_b_lm_qq.pdf")
+qqnorm(res,ylab="Linear Model Residual Quantiles",main="Linear Model Residual QQ Plot")
+qqline(res)
+dev.off()
+
+#regression tree diagnoistics
[…]

### 2014-11-26 stat343/hw7.tex
have a pretty good poisson model

@@
             \item[b)]
+                Comparing the linear fit to the regression tree and the random forest, it appears that the latter two fit slightly better than the former.
+As we look at the residuals, both the regression tree and the random forest are nearly perfectly distributed around 0 through out the predicted range (for the random tree this is of course by construction), while the residuals for the linear model do not quite seem to be at the lower end of the predicted range.
+As well, while the linear model residuals deviate from a normal distribution towards the upper end of their range, the regression tree and random forest have residuals that are normally distributed pretty well throughout.
+\\
+                \smallskip
+                This is the regression tree I fit:
+                \begin{center}
+                    \includegraphics[width=9cm]{hw7/hw7_2_b_tree} 
+                \end{center}
+                The transformed linear regression: \\
+                    \includegraphics[width=9cm]{hw7/hw7_2_b_lm_res} 
+                    \includegraphics[width=9cm]{hw7/hw7_2_b_lm_qq} \\ 
+                The regression tree: \\
+                    \includegraphics[width=9cm]{hw7/hw7_2_b_rt_res} 
+                    \includegraphics[width=9cm]{hw7/hw7_2_b_rt_qq} \\ 
+                The random forest: \\
+                    \includegraphics[width=9cm]{hw7/hw7_2_b_rf_res} 
+                    \includegraphics[width=9cm]{hw7/hw7_2_b_rf_qq} \\
+
         \end{itemize}
+    \item[3.]
+        \item[a)]
+            I applied backwards elimination, working with a critical value of $.2$. Doing this, I eliminated three variables and kept the rest.
+            \smallskip
+            This was the order I eliminated variables in, and their p-value when I eliminated them:\\
+            \FloatBarrier
+            \input{hw7/hw7_3_a_excl}
+            \FloatBarrier                
+            The end result was this regression: \\
+            \FloatBarrier
+            \input{hw7/hw7_3_a_lm}
+            \FloatBarrier                
+        \item[b)]
+            Going by forward selection, again with a critical value of $.2$, I ended up with the same model gotten previously by backwards selection. 
+            \smallskip
+            This was the order I included variables in, and their p-value when I included them:\\
+            \FloatBarrier
+            \input{hw7/hw7_3_b_incl}
+            \FloatBarrier                
+            The end result was this regression: \\
+            \FloatBarrier
+            \input{hw7/hw7_3_b_lm}
+            \FloatBarrier
 
@@
         \end{align*}
-        As was discussed with added variable plots, for an additional variable $X_i$ added to a regression, if $\hat e_{(-i)}$ are the residuals from the regression excluding $X_i$ and $\hat e$ are the residuals from the regression including it,
+        As was discussed with added variable plots, for an additional variable $X_i$ added to a regression, if $\hat e_{(-i)}$ are the residuals from the regression excluding $X_i$, $\hat e$ are the residuals from the regression including it, and $\beta_i$ is what the coefficient would be on $X_i$ if it was included in the regression, then
         \begin{align*}
@@
         \end{align*}
-        Thus, we can minimize the $RSS$ by minimizing choosing the variable $X_i$ such that the quantity on the right is minimized.
+        Thus, we can minimize the $RSS$ by choosing the variable $X_i$ such that the quantity on the right is minimized.
+    \item[5.]
+        $RSS$ will be increased the least by removing the variable which has the highest p-value. \\
+        \smallskip
+        {\bf Explanation:} If $RSS_{(-i)}$ is the residual sum of squares for a model excluding variable $i$, we want to pick $i$ such that $RSS_{(-i)}-RSS$ is smallest.
+Obviously, the same $i$ will minimize
+        \[ F= \frac{\frac{RSS_{(-i)}-RSS}{p-(p-1)}}{\frac{RSS}{n-p}}\]
+        This will in turn maximize the p-value for the F test comparing the model to the model excluding variable $i$.
+However, as we know, this p-value is equivalent to the p-value from the t test in the regression summary for variable $i$.
+Thus, this variable, when removed, will result in the smallest increase in $RSS$.
 

### 2014-11-26 stat343/hw7.R

@@ -103,7 +103,7 @@ oz.t.rt<-rpart(O3^lam ~temp+humidity+ibh, data=ozone)
 oz.t.rf<-randomForest(O3^lam ~temp+humidity+ibh, data=ozone)
 
 pdf("hw7/hw7_2_b_tree.pdf")
-plot(oz.t.rt,main="Regression Tree")
+plot(oz.t.rt,main="Regression Tree",uniform=TRUE)
 text(oz.t.rt)
 dev.off()
 
@@ -113,9 +113,10 @@ res<-residuals(oz.t.lm)
 pdf("hw7/hw7_2_b_lm_res.pdf")
 plot(pre,res,xlab="Predicted",ylab="Residual",main="Linear Model Residual Plot")
 lines(predict(loess(res~pre)))
+legend('topright',"LOESS Fit",lty=1)
 dev.off()
 pdf("hw7/hw7_2_b_lm_qq.pdf")
-qqnorm(res,ylab="Linear Model Residual Quantiles",main="Linear Model Residual QQ Plot")
+qqnorm(res,ylab="Linear Model Residual Quantiles",xlab="Normal Theoretical Quantiles",main="Linear Model Residual QQ Plot")
 qqline(res)
 dev.off()
 
@@ -125,9 +126,10 @@ res<-residuals(oz.t.rt)
 pdf("hw7/hw7_2_b_rt_res.pdf")
 plot(pre,res,xlab="Predicted",ylab="Residual",main="Regression Tree Residual Plot")
 lines(predict(loess(res~pre)))
+legend('topright',"LOESS Fit",lty=1)
 dev.off()
 pdf("hw7/hw7_2_b_rt_qq.pdf")
-qqnorm(res,ylab="Regression Tree Residual Quantiles",main="Regression Tree Residual QQ Plot")
+qqnorm(res,ylab="Regression Tree Residual Quantiles",xlab="Normal Theoretical Quantiles",main="Regression Tree Residual QQ Plot")
 qqline(res)
 dev.off()
 
@@ -137,11 +139,83 @@ res<-ozone$O3^lam-predict(oz.t.rf)
 pdf("hw7/hw7_2_b_rf_res.pdf")
 plot(pre,res,xlab="Predicted",ylab="Residual",main="Random Forrest Residual Plot")
 lines(predict(loess(res~pre)))
+legend('topright',"LOESS Fit",lty=1)
 dev.off()
 pdf("hw7/hw7_2_b_rf_qq.pdf")
[…]

### 2014-11-26 stat374/final_project.R

@@ -8,9 +8,9 @@ library('reshape2')
 library('chron')
 
 # Turns on different parts of the code
-clean.data <- TRUE 
+clean.data <- FALSE 
 data.sum   <- FALSE
-par.reg    <- FALSE
+par.reg    <- TRUE 
 nonpar.reg <- FALSE
 
 ###### Data cleaning
@@ -37,7 +37,7 @@ if (clean.data) {
     rob.data <- merge(hourly.robberies,hourly.weather,by=c("date","hour"),all.y=TRUE,all.x=FALSE)
     rob.data$robberies[is.na(rob.data$robberies)] <- 0
     # Mark weekend/holidays
-    rob.data$nonworkday <- is.weekend(rob.data$date) | is.holiday(rob.data$date)
+    rob.data$workday <- !(is.weekend(rob.data$date) | is.holiday(rob.data$date))
     # Pick a 20% subset as our test data set
     rob.data$test.set <- FALSE
     rob.data$test.set[sample.int(nrow(rob.data),size=floor(nrow(rob.data)*.2))] <- TRUE 
@@ -47,9 +47,44 @@ if (clean.data) {
 }
 ###### Data Summary
 if (data.sum) {
-    if (!exits("rob.data")) {
+    if (!exists("rob.data")) {
         load(file="final_project/hourly_robberies.RData")
     }
-    #tapply(robberies,list(hour,month),FUN=mean)
+    rob.data$temp.buc<-cut(rob.data$temp,breaks=c(-100,30,50,70,200))
+    rob.data$posprecip<-rob.data$precip>.0
+
+    rd.w <-rob.data[ rob.data$workday,]
+    rd.nw<-rob.data[!rob.data$workday,]
+
+    # Rate of robberies at different temperature ranges
+    tb.w<-tapply(rd.w$robberies,list(rd.w$hour,rd.w$temp.buc),FUN=mean)
+    pdf("final_project/tempbuc_w.pdf")
+    matplot(tb.w,type="l",col=rep(1,4),xlab="Hour (Military Time)",ylab="Mean Robberies Per Hour",main="Rate of Robberies at Different Temperatures, Workdays",xaxp=c(0,24,8),ylim=c(0,3),yaxp=c(0,3,6))
+    legend('bottomright',c("Below 30 Degrees","30-50 Degrees","50-70 Degrees","Above 70 Degrees"),lty=1:4)
+    dev.off()
+
[…]

### 2014-11-27 stat374/final_project.R
Finnished non parametric model

@@ -10,8 +10,12 @@ library('chron')
 # Turns on different parts of the code
 clean.data <- FALSE 
 data.sum   <- FALSE
-par.reg    <- TRUE 
+par.reg    <- FALSE 
 nonpar.reg <- FALSE
+eval.fit   <- TRUE 
+
+# par.reg options
+cross.validate <- FALSE
 
 ###### Data cleaning
 if (clean.data) {
@@ -84,7 +88,64 @@ if (par.reg) {
     if (!exists("rob.data")) {
         load(file="final_project/hourly_robberies.RData")
     }
+
+    ### Poisson model, selected based on Akaike Criteria
     rob.poi<-glm(glm(robberies~as.factor(hour)+as.factor(floor(hour/3))*workday+as.factor(floor(hour/3))*log(temp+10),family=poisson(link=log),data=rob.data,subset=test.set))
+
+    save(rob.poi,file="final_project/poisson_model.RData")
+}
+
+###### Non-Parametric Model
+if (nonpar.reg) {
+    if (!exists("rob.data")) {
+        load(file="final_project/hourly_robberies.RData")
+    }
+    
+    rd<-rob.data[rob.data$test.set,]
+    # Transform the count data
+    rd$logrob<-log(rd$robberies+1)
+    ### Don't want a discontinuity at 12am, so add back in data from prior and next day to get from -24 to 48 hours
+    rd.m <- rd
+    rd.m$date<-rd.m$date+1
+    rd.m$hour<-rd.m$hour-24
+    rd.p <- rd
+    rd.p$date<-rd.p$date-1
+    rd.p$hour<-rd.p$hour+24
+    rd<-rbind(rd.m,rd,rd.p)
+
+    # Want to select bandwith and scale for second variable (effectively two bandwiths). Test a grid of values and pick the best one. Don't care about hours outside 0-23. 
+    if (cross.validate) {
+        sub<-rd$hour>=0 & rd$hour<24
+        cv.score<-function(nn,s2,data,subset) {
+            cv.fit<-fitted(rob.ll<-locfit(logrob~lp(temp,hour,deg=1,nn=nn,scale=c(1,s2)),kern="gauss",data=data),what="coef",cv=TRUE)[subset]
+            return(sum((data$logrob[subset]-cv.fit)^2)/sum(subset))
+        }
+        nnvals<-seq(.02,.06,.005)
+        svals<-seq(.02,.1,.01)
[…]

### 2014-11-27 stat374/final_project.R
Split non-p model into work/weekend

@@ -11,8 +11,8 @@ library('chron')
 clean.data <- FALSE 
 data.sum   <- FALSE
 par.reg    <- FALSE 
-nonpar.reg <- FALSE
-eval.fit   <- TRUE 
+nonpar.reg <- TRUE
+eval.fit   <- FALSE 
 
 # par.reg options
 cross.validate <- FALSE
@@ -101,9 +101,12 @@ if (nonpar.reg) {
         load(file="final_project/hourly_robberies.RData")
     }
     
+    # Data to run regression on
     rd<-rob.data[rob.data$test.set,]
+    
     # Transform the count data
     rd$logrob<-log(rd$robberies+1)
+
     ### Don't want a discontinuity at 12am, so add back in data from prior and next day to get from -24 to 48 hours
     rd.m <- rd
     rd.m$date<-rd.m$date+1
@@ -112,37 +115,47 @@ if (nonpar.reg) {
     rd.p$date<-rd.p$date-1
     rd.p$hour<-rd.p$hour+24
     rd<-rbind(rd.m,rd,rd.p)
+    rd$workday <- !(is.weekend(rd$date) | is.holiday(rd$date))
+    rm(rd.m,rd.p)
 
     # Want to select bandwith and scale for second variable (effectively two bandwiths). Test a grid of values and pick the best one. Don't care about hours outside 0-23. 
     if (cross.validate) {
         sub<-rd$hour>=0 & rd$hour<24
         cv.score<-function(nn,s2,data,subset) {
-            cv.fit<-fitted(rob.ll<-locfit(logrob~lp(temp,hour,deg=1,nn=nn,scale=c(1,s2)),kern="gauss",data=data),what="coef",cv=TRUE)[subset]
+            cv.fit<-fitted(locfit(logrob~lp(temp,hour,deg=1,nn=nn,scale=c(1,s2)),kern="gauss",data=data),what="coef",cv=TRUE)[subset]
             return(sum((data$logrob[subset]-cv.fit)^2)/sum(subset))
         }
-        nnvals<-seq(.02,.06,.005)
-        svals<-seq(.02,.1,.01)
-        cv.mat<-NULL
+        nnvals<-seq(.02,.1,.01)
+        svals<-seq(.06,.2,.01)
+        cv.mat.w<-NULL
+        cv.mat.nw<-NULL
[…]

### 2014-11-29 stat374/final_project.R
Contour plots sorted out

@@ -6,16 +6,17 @@ library('MASS')
 library('xtable')
 library('reshape2')
 library('chron')
+library('graphics')
 
 # Turns on different parts of the code
-clean.data <- FALSE 
-data.sum   <- FALSE
-par.reg    <- FALSE 
-nonpar.reg <- TRUE
-eval.fit   <- FALSE 
+clean.data <- F 
+data.sum   <- F
+par.reg    <- F
+nonpar.reg <- F 
+eval.fit   <- T
 
 # par.reg options
-cross.validate <- FALSE
+cross.validate <- F
 
 ###### Data cleaning
 if (clean.data) {
@@ -47,7 +48,7 @@ if (clean.data) {
     rob.data$test.set[sample.int(nrow(rob.data),size=floor(nrow(rob.data)*.2))] <- TRUE 
     rm(hourly.weather,hourly.robberies)
     save(rob.data,file="final_project/hourly_robberies.RData")
-    
+
 }
 ###### Data Summary
 if (data.sum) {
@@ -90,7 +91,7 @@ if (par.reg) {
     }
 
     ### Poisson model, selected based on Akaike Criteria
-    rob.poi<-glm(glm(robberies~as.factor(hour)+as.factor(floor(hour/3))*workday+as.factor(floor(hour/3))*log(temp+10),family=poisson(link=log),data=rob.data,subset=test.set))
+    rob.poi<-glm(glm(robberies~as.factor(hour)+as.factor(floor(hour/3))*workday+as.factor(floor(hour/3))*log(temp+50),family=poisson(link=log),data=rob.data,subset=test.set))
 
     save(rob.poi,file="final_project/poisson_model.RData")
 }
@@ -119,6 +120,7 @@ if (nonpar.reg) {
     rm(rd.m,rd.p)
 
     # Want to select bandwith and scale for second variable (effectively two bandwiths). Test a grid of values and pick the best one. Don't care about hours outside 0-23. 
+    param.ll<-list() # This will be the list of parameters for fitting the model
     if (cross.validate) {
         sub<-rd$hour>=0 & rd$hour<24
         cv.score<-function(nn,s2,data,subset) {
@@ -133,32 +135,78 @@ if (nonpar.reg) {
[…]

### 2014-11-29 stat374/final_project.R
Getting somewhere with the CI plots

@@ -10,7 +10,7 @@ library('graphics')
 
 # Turns on different parts of the code
 clean.data <- F 
-data.sum   <- F
+data.sum   <- T
 par.reg    <- F
 nonpar.reg <- F 
 eval.fit   <- T
@@ -64,13 +64,13 @@ if (data.sum) {
     # Rate of robberies at different temperature ranges
     tb.w<-tapply(rd.w$robberies,list(rd.w$hour,rd.w$temp.buc),FUN=mean)
     pdf("final_project/tempbuc_w.pdf")
-    matplot(tb.w,type="l",col=rep(1,4),xlab="Hour (Military Time)",ylab="Mean Robberies Per Hour",main="Rate of Robberies at Different Temperatures, Workdays",xaxp=c(0,24,8),ylim=c(0,3),yaxp=c(0,3,6))
+    matplot(0:23,tb.w,type="l",col=rep(1,4),xlab="Hour (Military Time)",ylab="Mean Robberies Per Hour",main="Rate of Robberies at Different Temperatures, Workdays",xaxp=c(0,24,8),ylim=c(0,3),yaxp=c(0,3,6))
     legend('bottomright',c("Below 30 Degrees","30-50 Degrees","50-70 Degrees","Above 70 Degrees"),lty=1:4)
     dev.off()
 
     tb.nw<-tapply(rd.nw$robberies,list(rd.nw$hour,rd.nw$temp.buc),FUN=mean)
     pdf("final_project/tempbuc_nw.pdf")
-    matplot(tb.nw,type="l",col=rep(1,4),xlab="Hour (Military Time)",ylab="Mean Robberies Per Hour",main="Rate of Robberies at Different Temperatures, Weekend/Holiday",xaxp=c(0,24,8),ylim=c(0,3),yaxp=c(0,3,6))
+    matplot(0:23,tb.nw,type="l",col=rep(1,4),xlab="Hour (Military Time)",ylab="Mean Robberies Per Hour",main="Rate of Robberies at Different Temperatures, Weekend/Holiday",xaxp=c(0,24,8),ylim=c(0,3),yaxp=c(0,3,6))
     legend('bottomright',c("Below 30 Degrees","30-50 Degrees","50-70 Degrees","Above 70 Degrees"),lty=1:4)
     dev.off()
 
@@ -78,7 +78,7 @@ if (data.sum) {
[…]

### 2014-11-30 stat374/final_project.tex
Finnished final_project R code, adding tex code

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\setlength{\parindent}{1cm}

\title{Opening a Cold Case \\ {\large Investigating how Temperature Affects The Rate of Robberies}}
\date{December 2, 2014}
\author{Aaron Maurer}

\begin{document}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction and Objective}

With the onset of winter in Chicago, the temperature drops, and residents must concern themselves with a range of new dangers. Roads can be icy and dangerous. Heavy snow can collapse power lines and buildings. And frostbite becomes a concern on exposed skin. However, it may be the case that cold weather mitigates other dangers. When its frigid out, do criminals stay inside as well? \par

To partially investigate this question I have endeavored to quantify how the rate at which robberies occur in the city of Chicago varies with temperature. Robbery, defined as taking property from a person, without their consent, by force or threat of force\footnote{As opposed to theft, which is the taking property without consent but not necessarily by violent means, and burglary, which is breaking into a structure so as to commit a crime (whether or not a crime such as theft is committed)}, includes what is probably the most common outdoor crime, mugging. It would seem natural then that the rate at which robberies occurred would be sensitive to the weather. The this rate is distinct from one's risk of being robbed should they go outside, since it may be the case crime drops when fewer people are outside to be robbed. However, a better understanding of the former should also provide intuition for the latter question.
\par

\section{Data} 
The Chicago city government publishes a data set of which includes every crime reported to the police in the city, going back to 2001\footnote{City of Chicago. (2014). \textit{ Crimes - 2001 to present}. Retrieved from https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2}. I made use of the subset of this data which had a date recorded for the crime during 2011, 2012, or 2013, and which had a primary description of "Robbery", per the Illinois Uniform Crime Reporting Code. These crimes include both successful and attempted armed robbery, unarmed robbery, and vehicular hijacking. For each crime, the data set has additional information on where it occurred in the city, more detailed information about the crime, whether the crime was domestic, and if it resulted in an arrest. I, however, simply aggregated the total number of reported robberies that occurred anywhere in the city during each hour of each day during the time period. \par
I combined this hourly data with top of the hour weather data from O'Hare International Airport\footnote{Midwest Regional Climate Center. (2014). \textit{Unedited Hourly Data - Top of the Hour Observations - O'Hare International Airport}. Retrieved from mrcc.isws.illinois.edu/CLIMATE/ucld/ucld\_hrlyTop\_getdata1.jsp?WBAN=94846}. This data is a series of meteorological data, including temperature, accumulated precipitation, humidity, and wind speed, almost always measured once during a particular hour at 51 minutes after that hour. When there were multiple entries for an hour, I chose the first at 51 minutes which had the temperature recorded, or the first chronologically which had the temperature recorded if that was missing, discarding the rest. This was taken as the temperature for the city for a particular day and hour, and was matched with the hourly robbery counts. The small number ($\approx 10$) of hours for which there was no recorded temperature from O'Hare were excluded. The end result was $26,263$ hourly observations of temperature and reported robbery count, covering almost all of 2011-2013. \par
Examining this data, it is obvious that the most important feature in determining the number of robberies is the time of day, to a lesser extent, time of week. Figure 1 speaks to this effect, displaying an average by hour of the day for different temperature ranges and workday versus holiday/weekend. The mean number of robberies peaks between 6pm and 3am, possibly twice in that span, with its minimum around 6am, with the peak mean 4 to 5 times as high as the trough. This corresponds roughly to typical people's schedules; there are the most robberies during the period where people have gotten off of work for the evening, and the least when the fewest people are awake early in the morning. During the weekend, the trough and peak tend to move later, as people are up later or sleep in. \par
\begin{figure}[h]
    {\bf Figure 1} \\
    \includegraphics[width=9cm]{final_project/tempbuc_w}
    \includegraphics[width=9cm]{final_project/tempbuc_nw}
    \textit{Note: With fewer observations, the Weekend/Holiday series can be expected to be noisier}
\end{figure}
The effect of temperature seems to be less than this, and also varies with the time of day. During the middle of the day (8am to 3pm or so), it appears that the effect of temperature is at most minimal. One of the most interesting features is that in the evening, there actually appear to be fewer robberies between 6pm and 7pm when it is warm, with a much higher rate before and after. This is likely due to the changes in people's schedules that occurs when days are longer and the weather is warm. However, the strongest effect seems to be that there are far fewer robberies in the late evening and early morning when its cold. \par
\begin{figure}[h]
    {\bf Figure 2} \\
    \includegraphics[width=9cm]{final_project/ct_dist_w}
    \includegraphics[width=9cm]{final_project/ct_dist_nw}
\end{figure}
Looking at the distribution for a sample of hours in Figure 2, we once again see the effect of time of day and workday versus weekend/holiday. However, we also see a data set which seems to have, as one would hope for with count data, a Poisson distribution.

\section{Parametric Model} 
This

\end{document}

### 2014-11-30 stat374/final_project.R

@@ -10,13 +10,20 @@ library('graphics')
 
 # Turns on different parts of the code
 clean.data <- F 
-data.sum   <- T
-par.reg    <- F
+data.sum   <- F 
+par.reg    <- F 
 nonpar.reg <- F 
-eval.fit   <- T
+eval.fit   <- T 
 
 # par.reg options
-cross.validate <- F
+cross.validate <- T
+procs <- 2
+
+# global function
+# I'd be shocked if there isn't a time/date function to convert military time to 12 hour time with am/pm, but I'm too lazy to find it
+oclock<-function(h) paste(((as.numeric(h)-1)%%12)+1,ifelse(as.numeric(h) %/%12,"pm","am"),sep="")
+# Returns a sequence of colors along gray scale, since color printing is a scam
+gray.scale<-function(n) gray(seq(1,0,length.out=n))
 
 ###### Data cleaning
 if (clean.data) {
@@ -74,6 +81,17 @@ if (data.sum) {
     legend('bottomright',c("Below 30 Degrees","30-50 Degrees","50-70 Degrees","Above 70 Degrees"),lty=1:4)
     dev.off()
 
+    # Bar plot of counts at different times
+    hour.seq<-seq(2,23,8)
+    rob.ct.w<-table(rd.w$hour[rd.w$hour %in% hour.seq],rd.w$robberies[rd.w$hour %in% hour.seq])
+    rob.ct.nw<-table(rd.nw$hour[rd.nw$hour %in% hour.seq],rd.nw$robberies[rd.nw$hour %in% hour.seq])
+    pdf("final_project/ct_dist_w.pdf")
+    barplot(rob.ct.w,main="Distribution of Robbery Counts, Workdays",xlab="Number of Robberies During Hour",ylab="Occurrences in Data",col=gray.scale(length(hour.seq)),legend=oclock(rownames(rob.ct.w)),beside=TRUE)
+    dev.off()
+    pdf("final_project/ct_dist_nw.pdf")
[…]

### 2014-11-30 stat374/final_project.tex
Working on homework 4

@@
 
-%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
-\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
+%%% Packages
+% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
+% multirow, multicol allow for certain kinds of tables
+% enumerate lets you determine the style of the counter for the enumerate enviorment
+% graphicx lets you include pictures
+% listings lets you stick in blocks of code
+% placeins defines "\FloatBarrier", which stops tables from moving around
+\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
 \newcommand{\Z}{\mathbb{Z}}
@@
 \section{Parametric Model} 
-This
+This observation points to the obvious parametric model to fit to the data, a Poisson regression model\footnote{A generalized linear model with a Poisson distribution and log link fit by maximum likelihood}.
+This is a natural choice, and generally considered the first parametric model to try for count data, if not the default.
+Its assumptions are several fold.
+First, it assumes that data is Poisson distributed around a particular mean, which arises when counting the sum of infrequent and independent occurrences in which a large number of these events could, in theory, occur.
+This assumption seems fairly safe; robberies are infrequent, and though one criminal could be responsible for multiple robberies, this effect overall can be expected to be trivial in such a large city.
+The less safe assumption is that the mean of this distribution can be predicted by the product of exponential functions of the set of predictors.
+It isn't obvious that this is true, but it is mathematically convenient, constricting the mean to positive numbers.
+\par
+Choosing this as my model, I tried a number of specifications on an $80\%$ training subset of the data.
+These included various interactions of hour, as a factor and bucketed, temperature as a continuous variable, and workday versus weekend/holiday as a factor.
+Choosing the specification that had the best trade off of number of parameters versus likelihood, as dictated by the Akaike Information Criteria, I settled on an specification that included as predictors:
+\begin{itemize}
+    \item Hour of the day as a factor 
+    \item Three hour buckets as a factor interacted with workday as a factor
+    \item Three hour buckets as a factor interacted with the log of the sum temperature plus $50$.
+\end{itemize}
+
+Implicit in this set up is that the effect of temperature within a particular hour is assumed to be monotone. \par
+\begin{figure}[h]
+    {\bf Figure 3} \\
+    \includegraphics[width=9cm]{final_project/poi_w_contour_fill}
+    \includegraphics[width=9cm]{final_project/poi_nw_contour_fill}
+\end{figure}
+
+The estimates of the coefficients themselves are not particularly elucidating; the overall effect is best displayed graphically, such as through the heat maps in Figure 3.
+Here, we see the model capturing several of the features we noticed earlier.
+The predicted robbery rate is quite low in the late evening and early morning when it is cold, and rises quickly with the temperature, especially on the weekend.
+A bit later (6am or so), the crime rate is low, and throughout the middle of the day it only rises slightly with temperature.
+The early evening is the only area that seems outright strange; high values are predicted at low temperatures, and drops off somewhat as the temperature increases.
+This questionable, but is likely an artifact of the assumed monotone relationship between temperature and robbery rate paired with the small dip between higher and lower temperatures in this range.
+\par
+This is displayed again for different three hours individually in Figure 4 along with confidence intervals.
+Here, we see again the patterns listed above, but also that the model predicts these big swings in robbery rate due to temperature with a great deal of certainty.
+Particularly, at 2am, the confidence intervals for two predictions with as little as 20 degrees separation don't overlap, asserting that they are different with much higher confidence than $95\%$.
+\par
+
+\begin{figure}[h]
+    {\bf Figure 4} \\
+    \includegraphics[width=9cm]{final_project/poi_w_ci}
+    \includegraphics[width=9cm]{final_project/poi_nw_ci}
+\end{figure}
+
+The main danger with a Poisson regression model, such as this, is that if the predicted mean function may not be able to capture fully how the mean varies.
+This would most likely be due to insufficient predictors being included.
+The result may be larger variance around the predicted mean than allowed for with the Poisson Distribution, invalidating the model.
+In this case, should the rest of the assumptions hold, a negative binomial regression can be used, which can be derived as a certain sum of Poisson distributions with varying means, as could be the case should the mean not be perfectly predicted by the predictors.
+This makes the Poisson model a particular example of a negative binomial, and a formal test can be used to decide if the additional parameter of a negative binomial is needed.
+Such a test might well have indicated that I should have used a negative binomial regression here, but given time constraints and a lack of familiarity, I simply went with the Poisson model.
+\par
+Additionally, as with nonparametric models, a better tool for evaluating the fit of a particular specification of the model may have been cross validation rather than Akaike Information Criteria, since it gives a more direct estimate of over fitting.
+However, this model appears to under fit if anything, so cross validation may not have made such a large difference
+
+\section{Nonparametric Model} 
+To model this data with a nonparametric model, I ran separate local linear regressions on workdays and weekends/holidays, with each predicting the log of the sum number of robberies plus one with hour and temperature simultaneously.
+As with the Poisson model, I fit on an $80\%$ training subset.
+In fitting it, I used a Gaussian kernel, individual bandwidths for each point such that a constant portion of the total data was closer to the point than the bandwidth, and a relative scaling factor for the variables to determine which points were included in the neighborhood (essentially making a second component of the bandwidth for the second variable).
+These two values were chosen from a grid of values based on which resulting model had the lowest cross validation score.
+\par
+The reasoning and justification for these choices are as such:
+\begin{itemize}
+    \item The kernel doesn't make a great deal of difference, so the choice was arbitrary. 
+    \item Since I am particularly interested in how temperatures on the extreme edge of my data effect the robbery rate, I need a model with low boundary bias, making a local linear fit superior to a simpler kernel regression.
+    \item Predicting log robberies nicely constrains the prediction to positive counts, which is what it should be in the real world.
+    \item As the initial plots have shown, there seem to be fairly distinct trends between workdays and not workdays, so two separate models seemed more appropriate than one.
+    \item Also as observed earlier, the relationship between temperature and robberies doesn't seem constant by time of day, so it would likely be insufficient to model it separately, as would be the result from a general additive model that included separate local linear models for hour and temperature
+    \item The observations are not equally spaced out over their range, so one bandwidth for all the points wouldn't smooth enough on the extremes or too much in the middle.
+Instead, by dictating the bandwidth such it defined a neighborhood with a set number of neighbors, the amount of smoothing was comparable across the data points.
+    \item Including a scaling factor for the second variable is essential, since the units of the two variables aren't comparable.
+\end{itemize} \par
+With this model, I am only really relying on two assumptions.
+First, the number or robberies as a particular time and temperature is related to the number of robberies at close time and temperatures, which is obvious.
+Second, I am assuming the true function is sufficiently smooth that some kind of linear fit in a  neighborhood can be used to give a reasonable prediction.
+If, for instance, the function had peaks and troughs more often than I had data, this method would fall apart.
+However, this seems unlikely from observing the data.
+\par
+The parameters
+\begin{figure}[h]
+    {\bf Figure 5} \\
+    \includegraphics[width=9cm]{final_project/ll_w_contour_fill}
+    \includegraphics[width=9cm]{final_project/ll_nw_contour_fill}
+\end{figure}
+\begin{figure}[h]
+    {\bf Figure 6} \\
+    \includegraphics[width=9cm]{final_project/ll_w_ci}
+    \includegraphics[width=9cm]{final_project/ll_nw_ci}
+\end{figure}
 

### 2014-11-30 stat374/hw4.tex
Working on homework 4

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 374, Homework 4 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
\begin{itemize} 
    \item[1.]
        \begin{itemize} 
            \item[(a)] Averaging over $100$ simulations where $500$ points
were drawn from a normal means model with each mean equal to $2$, the average
risk of th mean was $1.016$, and the average risk of the isotonic regression
was $6.159$. As one might expect, since the MLE makes a stronger and correct
assumption, it has lower risk. However, the monotone regression still has
fairly low risk.
            \item[(b)] I estimated the risk for every fourth $n$ betwen 20 and
200. For each, I averged over 500 simulations. It seems that the risks are
proportional; the regression on the linear function seems to have about $25\%$
less risk than the step function for each $n$, with both risks growing at a
slower rate than linear, possibly proportional to \(\frac{log(n)}{n}\).
                \begin{center}
                    \includegraphics[width=9cm]{hw4/1_b} 
                \end{center}

            \item[(c)] Let $c(x_n)$ be the concave sequence of points evaluated at points $x_n$, and let $s(x_n)$ be a corresponding convex sequence. I will attempt to show that if $s(x_n)$ is not linear, it can not be the best fit. We can break this problem down into three cases.
                \begin{itemize}
                    \item[case 1:] If $s(x_n)>c(x_n)$ for all $x$, then there is trivially a better fit $s(x_n)-c$, where $c$ is the minimum distance between the two sequences. This will reduce the problem to the next case.
                    \item[case 2:] $s(x_n)=c(x_n)$ for at least one $n$, but it is never the case $s(x_n)<c(x_n)$. Let $s(x_1)=c(x_1)$. Since $s$ and $c$ are respectively convex and concave, there exists a line passing through $s(x_1)=c(x_1)$ such that $s(x_n)\geq l(x_n) \geq c(x_n) \, \forall n$. If $s(x_n)$ isn't itself a line, then there will be a $n$ for which their is strict inequality on the left hand side, making $l$ a superior fit to $s$.
                    \item[case 3:] $s(x_n)<c(x_n)$ for some $n$. Let us flesh
out $s(x_n)$ and $c(x_n)$, such that $s(x)$ and $c(x)$ are piecewise linear
functions connecting the points of the respective sequences. These functions
are necessarily also convex and concave respectively. We once again have three
cases:
                        \begin{itemize}
                            \item[case 3.1:] If $s(x)$ and $c(x)$ never intersect, a shift as in case 1 will achieve a superior fit and reduce the problem to the next case.
                            \item[case 3.2:]  
                            \item[case 3.3:] If they $s(x)$ and $c(x)$ intersect at multiple points, we can choose two intersection points $a<b$. Let $l$ be the line passing through them. It is a secant line for both $s$ and $c$, so by their respective convexity/concavity, $c(x)\geq l(x) \geq s(x)$ on $[a,b]$ and $c(x)\leq l(x) \leq s(x)$ elsewhere. Thus, if $c(x)\neq l(x)$ for some $x$, then $l$ is a superior fit to the function $s$. Since $s$ and $c$ are linear between the $x_n$, this can only arise if $s(x_n)\neq l(x_n)$ for some $n$, making $l(x_n)$ a superior fit to $c(x_n)$.

                        \end{itemize}
                \end{itemize}
        \end{itemize}
    \item[2.]
        I have computed the mean squared error for each of the methods:
        \FloatBarrier
        \input{hw4_2} 
        \FloatBarrier
        \begin{itemize}
            \item[(a)]
                At a low level of noise, the wavelet methods all preform better than the local linear model. With low noise, their ability to accurately match the fine features of the Doppler function win out. None of them preform terribly though. On the other hand, once the noise increases, the James Stein and Universal Threshold methods get worse much quicker than Sure Shrink and local linear, which are comparatively much better. This is likely since it becomes harder to distinguish feature from noise. In both cases, the James Stein is worse than the Universal Threshold, and Local Linear loses out to Sure Shrink. As the graphs note the distinction between local linear regression and Sure Shrink are hard to pick out by eye  \\
                \includegraphics[width=9cm]{hw4/2_ab_pt1} 
                \includegraphics[width=9cm]{hw4/2_ab_pt01} \\
            \item[(b)]
                On the other hand, once there are occasional spikes in the variance, the local linear model preforms the best. As the graph shows, it preforms more smoothing on the occasional spikes than the Sure Shrink, which likely has them identified as true features.
                \begin{center}
                    \includegraphics[width=9cm]{hw4/2_ab_bimodal} 
                \end{center}
    \item[3.]

        \end{itemize}

\end{itemize}

\end{document}

### 2014-11-30 stat374/hw4.R

@@ -1,0 +1,141 @@
+###### Header
+library('parallel')
+library('locfit')
+library('gam')
+library('MASS')
+library('xtable')
+library('reshape2')
+library('chron')
+library('graphics')
+library('Iso')
+library('wavelets')
+
+### Problems
+problem1<-F
+problem2<-F
+problem3<-T
+
+### Options
+procs<-2
+
+############# Problem 1
+if (problem1) {
+    ### a
+    n<-500
+    th<-2
+    b<-100
+    risk<-function(n,th) {
+        z<-rnorm(n)+th
+        hat<-pava(z)
+        return(c(n*(th-mean(z))^2,sum((th-hat)^2)))
+    }
+    est<-c(0,0)
+    for (i in 1:b){
+        est<-est+risk(n,th)
+    }
+    est<-est/b
+    
+    ### b
+    n<-seq(20,200,4)
+    sig<-.25
+    b<-500 
+    risk.step<-function(n,sig) {
+        step<-c(rep(0,n/2),rep(1,n/2))
+        z<-rnorm(n)*sig+step
+        hat<-pava(z)
+        return(sum((hat-step)^2))
+    }
+    risk.line<-function(n,sig) {
+        line<-1/(n:1)
+        z<-rnorm(n)*sig+line
+        hat<-pava(z)
+        return(sum((hat-line)^2))
+    }
+    risk<-matrix(0,nrow=length(n),ncol=2)
+    for (i in 1:b) {
+        risk[,1]<- risk[,1] + mcmapply(risk.step,n=n,MoreArgs=list(sig=sig),mc.cores=procs)
+        risk[,2]<- risk[,2] + mcmapply(risk.line,n=n,MoreArgs=list(sig=sig),mc.cores=procs)
+    }
+    risk<-risk/b
+    pdf('hw4/1_b.pdf')
+    matplot(x=n,y=risk,col=c(1,1),xlab="N",type="l",ylab=paste("Average Risk Over",b,"Simulations"),main="Monotone Regression Risk")
+    legend('bottomright',c("Step Function","Line"),lty=c(1,2))
+    dev.off()
+    
+}
+############# Problem 2 
+if (problem2) {
+    doppler<-function(x) sqrt(x*(1-x))*sin(2.1*pi/(x+.05))
[…]

### 2014-12-01 stat374/final_project.tex
Project written, homework 4 first three questions done

@@
 \par
-Additionally, as with nonparametric models, a better tool for evaluating the fit of a particular specification of the model may have been cross validation rather than Akaike Information Criteria, since it gives a more direct estimate of over fitting.
-However, this model appears to under fit if anything, so cross validation may not have made such a large difference
+Additionally, as with nonparametric models, a better tool for evaluating the fit of a particular specification of the model may have been cross validation rather than Akaike Information Criteria, since it gives a more direct estimate of over and under fitting.
 
@@
 As with the Poisson model, I fit on an $80\%$ training subset.
-In fitting it, I used a Gaussian kernel, individual bandwidths for each point such that a constant portion of the total data was closer to the point than the bandwidth, and a relative scaling factor for the variables to determine which points were included in the neighborhood (essentially making a second component of the bandwidth for the second variable).
+In fitting it, I used a Gaussian kernel, individual bandwidths for each point such that a constant portion of the total data was closer to the point than the bandwidth, and a relative scaling factor for the second variable (this effectively stretches or shrinks the kernel in one direction).
 These two values were chosen from a grid of values based on which resulting model had the lowest cross validation score.
-\par
+As well, the hour span was expanded to include the previous and next day's data, so there should be no boundary bias towards the beginning and end of days.\par
 The reasoning and justification for these choices are as such:
@@
     \item The kernel doesn't make a great deal of difference, so the choice was arbitrary. 
-    \item Since I am particularly interested in how temperatures on the extreme edge of my data effect the robbery rate, I need a model with low boundary bias, making a local linear fit superior to a simpler kernel regression.
+    \item Since I am particularly interested in how temperatures on the extreme edge of my data effect the robbery rate.
+This necessitates a need a model with low boundary bias along the temperature range, making a local linear fit superior to a simpler kernel regression.
+Higher polynomial fits on the other hand make it increasingly difficult to fit the model well.
     \item Predicting log robberies nicely constrains the prediction to positive counts, which is what it should be in the real world.
     \item As the initial plots have shown, there seem to be fairly distinct trends between workdays and not workdays, so two separate models seemed more appropriate than one. 
-    \item Also as observed earlier, the relationship between temperature and robberies doesn't seem constant by time of day, so it would likely be insufficient to model it separately, as would be the result from a general additive model that included separate local linear models for hour and temperature
+    \item Also as observed earlier, the relationship between temperature and robberies doesn't appear constant by time of day, so it would likely be insufficient to model it separately, as would be the result from a general additive model that included separate local linear models for hour and temperature.
+In this case, there is a sufficient amount of data to still have a reasonable degree of accuracy, so the curse of dimensionality is not prohibitive.
     \item The observations are not equally spaced out over their range, so one bandwidth for all the points wouldn't smooth enough on the extremes or too much in the middle.
-Instead, by dictating the bandwidth such it defined a neighborhood with a set number of neighbors, the amount of smoothing was comparable across the data points.
+Instead, by dictating the bandwidth such that it defined a neighborhood with a fixed number of neighbors, the amount of smoothing was comparable across the data points.
     \item Including a scaling factor for the second variable is essential, since the units of the two variables aren't comparable.
@@
 First, the number or robberies as a particular time and temperature is related to the number of robberies at close time and temperatures, which is obvious.
-Second, I am assuming the true function is sufficiently smooth that some kind of linear fit in a  neighborhood can be used to give a reasonable prediction.
-If, for instance, the function had peaks and troughs more often than I had data, this method would fall apart.
-However, this seems unlikely from observing the data.
+Second, I am assuming sufficient smoothness to achieve convergence of the estimates.
 \par
-The parameters
 \begin{figure}[h]
@@
 \end{figure}
+\FloatBarrier
+The predictions of the models can be seen in the contour plots of Figure 5.
+It is important to note that with less data, bandwidths for the weekend/holiday model were chosen to include a larger portion of the data, and also that the scale was such that this model averaged more strongly over temperature than hour as compared to the workday model.
+This model seems to very clearly portray the patterns we have seen before.
+Peak crime on workdays is highest in the evening, with the main peak later and a second, smaller and earlier, peak developing when the temperature is higher.
+Also, the afternoon and evening crime rate does fall off as the temperature increases beyond a point.
+On non-workdays, the peak is in the earliest hours of the morning, and keeps increasing with the temperature.
+However, the prediction for the highest temperatures should be taken with a grain of salt; their is little if any data where temperatures actually reach that high at that hour of the day.
+\FloatBarrier
+
 \begin{figure}[h]
@@
 \end{figure}
+Evaluating the confidence of the fits, as seen in Figure 6, we see a much more nuanced picture for workdays than we saw from the Poisson model.
+It should be noted that the confidence intervals are based on local estimates of variance, necessitated by the variable bandwidth.
+This model is quite certain that the peak crime rate at $6$pm on workdays occurs when the temperature is between 30 and 50 degrees, and falls off afterwards, as opposed to the monotonic relationship predicted by the Poisson model.
+It also has a large amount of confidence there is a true increase in crime in the early morning, though as I noted, at high temperatures there is much more uncertainty.
+With much more smoothing, particularly over temperature, in the weekend/holiday model, its results look similar to the Poisson models.
+\par
+It seems safe to say this model is the strongest non-parametric regression for this subject.
+Alternate linear smoothers and generalized additive models were discussed earlier.
+Monotonicity with hour or temperature doesn't seem to hold, nor does convexity/concavity, so models making those assumptions aren't appropriate.
+This series doesn't have the quickly varying features that would necessitate a procedure like wavelet regression.
+However, what could be quite a good alternate approach is to not have aggregated robbery counts, and fit a density estimate to the data.
+This could be scaled by how often different temperature and hour combinations occurred to give relative rates of robbery.
+This approach has the advantage of preserving more information about the time a crime occurs, but is computationally more intensive (with more data points), and doesn't easily admit a comparable parametric model.
 
+\section{Discussion} 
+Comparing the two models just based on performance, the nonparametric model wins, though they both do fairly well.
+I calculated the root mean squared log error for each model over the $20\%$ test subset that wasn't used to train the models.
+The local linear model had a RMSLE of $.510$, while the Poisson model had a RMSLE of $.529$.
+By comparison, an unbiased estimate where the mean at each temperature/hour combination in the training data was taken had a RMSLE of $.563$, so both models are a meaningful improvement.
+\par
+Subjectively, they both picked up mostly the same trends.
+The only notable distinction was in how they treated the late afternoon and early evening period, with the Poisson model apparently failing to capture the non-monotone trend.
+However, by including more and varied predictors (from interactions, splines, transformations, etc.), one can get and arbitrary degree of flexibility.
+Choosing a sufficient set of predictors is a slow process to do manually though, making it easier to just start with a non-parametric method which has arbitrary flexibility built in.
+One could meet halfway between the methods with a nonparametric model fit with local likelihood, since the assumption that the error is Poisson or negative binomial distributed seems appropriate.
+\par
+In terms of interpretation, both are pretty convincingly show that the rate of robberies varies with temperature, generally increasing with it.
+Causation is quite unclear though; temperature is highly correlated with a number of other seasonal factors.
+One could reasonably expect the rate of robberies to vary with school holidays, daylight hours, or just the time of the year irrespective of season.
+By including all these factors in a model, it would likely be highly predictive, but the high collinearity can be expected to make it hard to attribute changes to just one factor, modeled parametrically or not.
+Appropriately modeling these effects would be the most important extension of the current model.
+\par
+There are a number of other areas where this model could be expanded though.
+There is a great wealth of geographic data in the Chicago crime database, and there is likely a great deal of heterogeneity in the baseline rate of robberies by geography, as well as the differences in the effect of weather on that rate.
+Different types of crime could also be investigated.
+Different attributes of the weather may also play a roll.
+I initially investigated the effect of precipitation on the robbery rate.
+The presence of precipitation, as showed in Figure 7, does seem to coincide with a small reduction in robberies, at least in during the late evening and early morning.
+This effect appears to vary a lot with the amount of precipitation, and there were relatively little data from when there was enough precipitation to have a meaningful effect.
+With the Poisson model, every version I tried which included precipitation in some form had an inferior AIC, so ultimately I decided not to model it, at least for now.
+Finally, there is slight asymmetry between the early morning and late evening predictions due to how the weekends are defined.
+Friday evening is likely more akin to a Saturday night, with more robberies, pushing up the late evening average of workdays as compared to the early mornings, while Sunday evening likely has the opposite effect on the Weekends/Holidays.
+A modified delineation between the two periods may have done away with this, improving the fit.
+There may be further clustering that could be preformed on the days as well.
+All these areas present opportunity for further refinement.
+
+\begin{figure}[h]
+    {\bf Figure 7} \\
+    \includegraphics[width=9cm]{final_project/precip}
+\end{figure}
 \end{document}

### 2014-12-01 stat374/hw4.tex
Project written, homework 4 first three questions done

@@
             \item[(c)] Let $c(x_n)$ be the concave sequence of points evaluated at points $x_n$, and let $s(x_n)$ be a corresponding convex sequence.
-I will attempt to show that if $s(x_n)$ is not linear, it can not be the best fit.
+I will attempt to show that if $s(x_n)$ is not linear, it is superseded by some linear function.
 We can break this problem down into three cases.
@@
                     \item[case 1:] If $s(x_n)>c(x_n)$ for all $x$, then there is trivially a better fit $s(x_n)-c$, where $c$ is the minimum distance between the two sequences.
-This will reduce the problem to the next case.
+This improved fit is of the next case, so if there is a superseding linear fit for it, there is one for this case as well.
                     \item[case 2:] $s(x_n)=c(x_n)$ for at least one $n$, but it is never the case $s(x_n)<c(x_n)$.
@@
 If $s(x_n)$ isn't itself a line, then there will be a $n$ for which their is strict inequality on the left hand side, making $l$ a superior fit to $s$.
-                    \item[case 3:] $s(x_n)<c(x_n)$ for some $n$. Let us flesh
-out $s(x_n)$ and $c(x_n)$, such that $s(x)$ and $c(x)$ are piecewise linear
-functions connecting the points of the respective sequences. These functions
-are necessarily also convex and concave respectively. We once again have three
-cases:
+                    \item[case 3:] $s(x_n)<c(x_n)$ for some $n$.
+Let us flesh out $s(x_n)$ and $c(x_n)$, such that $s(x)$ and $c(x)$ are piecewise linear functions connecting the points of each sequence.
+These functions are necessarily also convex and concave respectively.
+We once again have three cases:
                         \begin{itemize}
                             \item[case 3.1:] If $s(x)$ and $c(x)$ never intersect, a shift as in case 1 will achieve a superior fit and reduce the problem to the next case.
-                            \item[case 3.2:]  
+                            \item[case 3.2:] If $s(x)$ and $c(x)$ intersect exactly once, we can replace $s(x_n)$ with $s(x_n)+c$ for $n$ such that $s(x_n)<c(x_n)$, where $c$ is the minimum distance between the sequences on this set.
+This action will preserve convexity while creating an improved fit (the distance is strictly reduced), and reduces this case to the next case.
                             \item[case 3.3:] If they $s(x)$ and $c(x)$ intersect at multiple points, we can choose two intersection points $a<b$.
@@
                 \end{center}
+        \end{itemize}
     \item[3.]
+        \begin{itemize}
+            \item[(a)]
+                I have plotted out all of the fits in the graph below.
+The isotonic regression provides a good fit only over through day 25 or so, when the observed values cease to be monotone.
+After that, the shorter fits do a reasonable job over the middle part, which is essentially flat.
+However, the more data the regression is fit on where the trend actually seems to be decreasing, the worse the regression preforms.
+Assuming the error terms are symmetrically distributed, one would expect at worst $50\%$ of the observations to fall below the actual series.
+One could thus generate a test statistic based on how the number of times, as the function is generated step by step, the function is forced to drop in response to a lower observed value within a particular range.
+This statistic would be approximately binomially distributed (though there is additional uncertainty due to the function being noisy), and a critical value could be generated from this to determine whether the range isn't monotone.
+                \begin{center}
+                    \includegraphics[width=9cm]{hw4/3_a} 
+                \end{center}
+            \item[(b)] 
+                The series of local linear fits are generally pretty good.
+However, as they are fit in increasingly noisy parts of the data, they have a greater amount of uncertainty, since the confidence interval is fit with a constant variance assumption.
+                    \includegraphics[width=9cm]{hw4/3_b25day} 
+                    \includegraphics[width=9cm]{hw4/3_b50day}  \\
+                    \includegraphics[width=9cm]{hw4/3_b75day} 
+                    \includegraphics[width=9cm]{hw4/3_b85day}  \\
+                    \includegraphics[width=9cm]{hw4/3_b171day} 
+            \item[(c)]
+                The Haar wavelet fit is below.
+Its fit is comparable on the relatively smooth portion of the series.
+However, in the noisy section to the end, it mostly clings to the large fluctuations in the data, while the others averaged them out in one way or another.
+Since these fluctuations are due to the small number of surviving flies, rather than a reflection of the true rate, this constitutes a worse fit.
+                \begin{center}
+                    \includegraphics[width=9cm]{hw4/3_c} 
+                \end{center}
+                {\textbf My Code:}
+               \lstinputlisting[firstline=181,lastline=210]{hw4.R}
+            \item[(d)]
+                The Daubechies wavelet fit, much like the Haar, clings to closely to the observed data when the noise accounts for large swings, but otherwise has a similar, somewhat smoother, fit over the relatively calm part of the data.
+                \begin{center}
+                    \includegraphics[width=9cm]{hw4/3_d} 
+                \end{center}
+
+        \end{itemize}
+    \item[4.]
+        \begin{itemize}
+            \item[(a)]
+                \begin{align*}
+                    \E(\mu \vert \mathcal{D}_n) &= \int_{\R^d} \mu  p(\mu \vert \mathcal{D}_n) \, d\mu \\
+                    \E(\mu \vert \mathcal{D}_n) &= \frac{\int_{\R^d} \mu  p(\mathcal{D}_n \vert \mu) \pi(\mu) \, d\mu}{\int_{\R^d}p(\mathcal{D}_n \vert \mu) \pi(\mu) \, d\mu} \\
+                    \E(\mu \vert \mathcal{D}_n) &= \frac{\int_{\R^d} \mu \prod_{i=1}^n \frac{1}{2\pi}e^{-\|x_i-\mu\|^2}  \, d\mu}{\int_{\R^d}\prod_{i=1}^n\frac{1}{2\pi} e^{-\|x_i-\mu\|^2} \, d\mu}  \\
+                    \E(\mu \vert \mathcal{D}_n) &= \frac{\int_{\R^d} \mu \frac{1}{2\pi} e^{-\sum_{i=1}^n\|x_i-\mu\|^2}  \, d\mu}{\int_{\R^d}\frac{1}{2\pi}e^{-\sum_{i=1}^n\|x_i-\mu\|^2} \, d\mu}  \\
+                    \E(\mu \vert \mathcal{D}_n) &= \frac{\int_{\R^d} \mu \frac{1}{2\pi} e^{-n\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2 -n\left(\sum_{i=1}^n\frac{\|x_i\|^2}{n}-\left\|\sum_{i=1}^n\frac{x_i}{n}\right\|^2\right)}  \, d\mu}{\int_{\R^d}\frac{1}{2\pi}e^{-n\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2 -n\left(\sum_{i=1}^n\frac{\|x_i\|^2}{n}-\left\|\sum_{i=1}^n\frac{x_i}{n}\right\|^2\right)} \, d\mu}  \\
+                    \E(\mu \vert \mathcal{D}_n) &= \frac{\int_{\R^d} \mu \frac{1}{2\pi} e^{-n\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2}  \, d\mu}{\int_{\R^d}\frac{1}{2\pi}e^{-n\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2} \, d\mu}  \\
+                    \E(\mu \vert \mathcal{D}_n) &= \int_{\R^d} \mu \frac{1}{2\pi} e^{-n\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2}  \, d\mu\\
+                    \E(\mu \vert \mathcal{D}_n) &= \sum_{i=1}^n\frac{x_i}{n}\\
+                \end{align*}
+        \item[(b)]
+                Let \(y_i=\|x_i\|^2\) and \(\mathcal{H}_n={y_1,...,y_n}\).
+Since each component of each $x_i$  is independent of the others (because there is $0$ covariance and the distribution is multivariate normal), we may conclude that the $y_i$ have a $\chi^2$ distribution with $d$ degrees of freedom and a noncentrality parameter of $\|\mu\|^2$.
+Thus,
+            \begin{align*}
+                p(\theta\vert \mathcal{H}_n) &= \frac{p(\mathcal{H}_n \vert \theta)p(\theta)}{p(\mathcal{H}_n)} \\
+                p(\theta\vert \mathcal{H}_n) &= \frac{\prod_{i=1}^n p(y_i \vert \theta)\int_{\R^d} p(\theta \vert \mu) p(\mu) d\mu}{\int_{\R^d}p(\mathcal{H}_n\vert \mu)p(\mu)} \\
+                p(\theta\vert \mathcal{H}_n) &= \frac{\prod_{i=1}^n p(y_i \vert \theta)\int_{\R^d} I_{\{\|x\|^2=\theta\}}(\mu) d\mu}{\int_{\R^d}\prod_{i=1}^n p(y_i \vert \theta) d\mu)} \\
+            \end{align*}
+        \end{itemize}
+    \item[5.]
+        \begin{itemize}
+            \item[(a)]
+                \begin{align*}
+                    \sum_{j=1}^n w_j &= V_1 + \sum_{j=2}^n V_j \prod_{i=1}^{j-1}(1-V_i) \\
+                    \sum_{j=1}^n w_j &= 1 - \prod_{j=1}^{n}(1-V_i) \\
+                    \sum_{j=\infty}^n w_j &= 1 - \prod_{j=1}^{\infty}(1-V_j) \\
+                \end{align*}
 

### 2014-12-01 stat374/hw4.R

@@ -9,14 +9,17 @@ library('chron')
 library('graphics')
 library('Iso')
 library('wavelets')
+library('wavethresh')
 
 ### Problems
 problem1<-F
 problem2<-F
 problem3<-T
+problem4<-F
+problem5<-F
 
 ### Options
-procs<-2
+procs<-4
 
 ############# Problem 1
 if (problem1) {
@@ -136,6 +139,94 @@ if (problem2) {
     
 ############# Problem 3
 if (problem3) {
+    eps<-.001
+    flies<-read.table("hw4/flies.dat",header=T)
+    flies$logm<-log(flies$mort.rate+eps)
+
+    ### a
+    iso.mat<-matrix(,nrow=90,ncol=5)
+    iso.mat[,1]<-flies$logm[1:90]
+    days<-c(25,50,75,85)
+    for (d in 1:4) {
+        func<-pava(flies$logm[1:days[d]],long.out=T,stepfun=T)$h
+        iso.mat[,(d+1)]<-func(0:90)[-1]
+    }
+
+    pdf('hw4/3_a.pdf')
+    matplot(y=iso.mat,x=flies$day[1:90],type="l",col=rep(1,5),lty=1:5,xlab="Day",ylab="Log Mortality Rate",main="Isotonic Fit")
+    legend('bottom',c("Actual",paste(days,"Day Fit")),lty=1:5)
+    dev.off()
+ 
+    ### b
+    cv.score<-function(h) {
+        cv.fit<-fitted(locfit(logm~lp(day,deg=1,h=h),kern="gauss",ev=dat(),data=flies),what="coef",cv=TRUE,maxk=2000)
+        return(sum((y-cv.fit)^2)/n)
+    }
+    hseq<-exp(seq(-2,3,.25))
+    cv.mat.w<-mcmapply(cv.score,h=hseq,mc.cores=procs)
+    h<-hseq[cv.mat.w==min(cv.mat.w)]
+
+    days<-c(25,50,75,85,171)
+    for (d in days) {
+        predmat.ll<-matrix(,nrow=d+1,ncol=4)
+        m.ll<-predict(locfit(logm~lp(day,deg=1,h=h),kern="gauss",ev=dat(),data=flies[1:(d+1),]),se.fit=T)
[…]

### 2014-12-02 stat374/hw4.tex
All but hw4 prbl 5 finnished

@@
                 \begin{align*}
-                    \E(\mu \vert \mathcal{D}_n) &= \int_{\R^d} \mu  p(\mu \vert \mathcal{D}_n) \, d\mu \\
-                    \E(\mu \vert \mathcal{D}_n) &= \frac{\int_{\R^d} \mu  p(\mathcal{D}_n \vert \mu) \pi(\mu) \, d\mu}{\int_{\R^d}p(\mathcal{D}_n \vert \mu) \pi(\mu) \, d\mu} \\
-                    \E(\mu \vert \mathcal{D}_n) &= \frac{\int_{\R^d} \mu \prod_{i=1}^n \frac{1}{2\pi}e^{-\|x_i-\mu\|^2}  \, d\mu}{\int_{\R^d}\prod_{i=1}^n\frac{1}{2\pi} e^{-\|x_i-\mu\|^2} \, d\mu}  \\
-                    \E(\mu \vert \mathcal{D}_n) &= \frac{\int_{\R^d} \mu \frac{1}{2\pi} e^{-\sum_{i=1}^n\|x_i-\mu\|^2}  \, d\mu}{\int_{\R^d}\frac{1}{2\pi}e^{-\sum_{i=1}^n\|x_i-\mu\|^2} \, d\mu}  \\
-                    \E(\mu \vert \mathcal{D}_n) &= \frac{\int_{\R^d} \mu \frac{1}{2\pi} e^{-n\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2 -n\left(\sum_{i=1}^n\frac{\|x_i\|^2}{n}-\left\|\sum_{i=1}^n\frac{x_i}{n}\right\|^2\right)}  \, d\mu}{\int_{\R^d}\frac{1}{2\pi}e^{-n\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2 -n\left(\sum_{i=1}^n\frac{\|x_i\|^2}{n}-\left\|\sum_{i=1}^n\frac{x_i}{n}\right\|^2\right)} \, d\mu}  \\
-                    \E(\mu \vert \mathcal{D}_n) &= \frac{\int_{\R^d} \mu \frac{1}{2\pi} e^{-n\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2}  \, d\mu}{\int_{\R^d}\frac{1}{2\pi}e^{-n\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2} \, d\mu}  \\
-                    \E(\mu \vert \mathcal{D}_n) &= \int_{\R^d} \mu \frac{1}{2\pi} e^{-n\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2}  \, d\mu\\
-                    \E(\mu \vert \mathcal{D}_n) &= \sum_{i=1}^n\frac{x_i}{n}\\
+                    P(\mu \vert \mathcal{D}_n) &= \frac{\mathcal{L}_n(\mu) \pi(\mu)}{\int_{\R^d}\mathcal{L}_n(\mu) \pi(\mu) \, d\mu} \\
+                    P(\mu \vert \mathcal{D}_n) &= \frac{\prod_{i=1}^n \frac{1}{\sqrt{(2\pi)^d}}\exp\left(-\frac{1}{2}\|x_i-\mu\|^2\right)}{\int_{\R^d}\prod_{i=1}^n\frac{1}{\sqrt{(2\pi)^d}} \exp\left(-\frac{1}{2}\|x_i-\mu\|^2\right) \, d\mu}  \\
+                    P(\mu \vert \mathcal{D}_n) &= \frac{ \exp\left(-\frac{1}{2}\sum_{i=1}^n\|x_i-\mu\|^2\right)}{\int_{\R^d}\exp\left(-\frac{1}{2}\sum_{i=1}^n\|x_i-\mu\|^2\right) \, d\mu}  \\
+                    P(\mu \vert \mathcal{D}_n) &= \frac{ \exp\left(-\frac{n}{2}\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2 -\frac{n}{2}\left(\sum_{i=1}^n\frac{\|x_i\|^2}{n}-\left\|\sum_{i=1}^n\frac{x_i}{n}\right\|^2\right)\right)}{\int_{\R^d}\exp\left(-\frac{n}{2}\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2 -\frac{n}{2}\left(\sum_{i=1}^n\frac{\|x_i\|^2}{n}-\left\|\sum_{i=1}^n\frac{x_i}{n}\right\|^2\right)\right) \, d\mu}  \\
+                    P(\mu \vert \mathcal{D}_n) &= \frac{ \exp\left(-\frac{n}{2}\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2\right)  \, d\mu}{\int_{\R^d}\exp\left(-\frac{n}{2}\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2\right)}  \\
+                    P(\mu \vert \mathcal{D}_n) &= \sqrt{\frac{n^d}{(2\pi)^d}} \exp\left(-\frac{n}{2}\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2\right) \\
                 \end{align*}
+                So, we can conclude that the posterior is distributed as \(N\left(\sum_{i=1}^n\frac{x_i}{n},\frac{1}{n}I_d\right)\).
         \item[(b)]
-                Let \(y_i=\|x_i\|^2\) and \(\mathcal{H}_n={y_1,...,y_n}\).
-Since each component of each $x_i$  is independent of the others (because there is $0$ covariance and the distribution is multivariate normal), we may conclude that the $y_i$ have a $\chi^2$ distribution with $d$ degrees of freedom and a noncentrality parameter of $\|\mu\|^2$.
-Thus,
+                By the previous result, the posterior for $\mu$ is normally distributed with a covariance matrix of $\frac{1}{n}I_d$.
+This implies that each $\mu_i$ is independent from the others (since they have covariance $0$ and the function is multivariate normal.
+Accordingly, if $Z\sim N(0,I_d)$, then \(\mu \sim \sum_{i=1}^n\frac{x_i}{n}+\frac{Z}{n}\).
+Thus, $\|n\mu\|^2=n\theta$ has a non-central $\chi^2$ distribution with $d$ degrees of freedom and noncentrality parameter \(\|\sum_{i=1}^n{x_i}\|^2\).
+This in turn means that $\theta\sim \frac{\chi^2}{n}$ distribution with the same offset parameter and degrees of freedom, so its mean is \(\tilde\theta=\frac{\|\sum_{i=1}^n{x_i}\|^2+d}{n}=\|\bar X_n\|^2 + \frac{d}{n}\).
+        \item[(c)]
+            Plugging in the values I've already derived,
             \begin{align*}
-                p(\theta\vert \mathcal{H}_n) &= \frac{p(\mathcal{H}_n \vert \theta)p(\theta)}{p(\mathcal{H}_n)} \\
-                p(\theta\vert \mathcal{H}_n) &= \frac{\prod_{i=1}^n p(y_i \vert \theta)\int_{\R^d} p(\theta \vert \mu) p(\mu) d\mu}{\int_{\R^d}p(\mathcal{H}_n\vert \mu)p(\mu)} \\
-                p(\theta\vert \mathcal{H}_n) &= \frac{\prod_{i=1}^n p(y_i \vert \theta)\int_{\R^d} I_{\{\|x\|^2=\theta\}}(\mu) d\mu}{\int_{\R^d}\prod_{i=1}^n p(y_i \vert \theta) d\mu)} \\
+                \frac{\E_\mu\vert \theta - \tilde \theta\vert^2}{\E_\mu\vert \theta - \hat \theta\vert^2} &= \frac{\E_\mu\vert \theta - \|\bar X_n\|^2 - \frac{d}{n}\vert^2}{\E_\mu\vert \theta - \|\bar X_n\|^2 + \frac{d}{n}\vert^2} \\
+                \frac{\E_\mu\vert \theta - \tilde \theta\vert^2}{\E_\mu\vert \theta - \hat \theta\vert^2} &= \frac{\E_\mu\vert \theta - \|\bar X_n\|^2 + \frac{d}{n} - 2\frac{d}{n}\vert^2}{\E_\mu\vert \theta - \|\bar X_n\|^2 + \frac{d}{n}\vert^2} \\
+                \frac{\E_\mu\vert \theta - \tilde \theta\vert^2}{\E_\mu\vert \theta - \hat \theta\vert^2} &= \frac{\E_\mu\left((\theta - \|\bar X_n\|^2 + \frac{d}{n})^2 - 4\frac{d}{n}(\theta - \|\bar X_n\|^2 + \frac{d}{n})+4\frac{d^2}{n^2}\right)}{2d+4\theta} \\
+                \frac{\E_\mu\vert \theta - \tilde \theta\vert^2}{\E_\mu\vert \theta - \hat \theta\vert^2} &= \frac{2d+4\theta+4\frac{d^2}{n^2}}{2d+4\theta} \\
+                \frac{\E_\mu\vert \theta - \tilde \theta\vert^2}{\E_\mu\vert \theta - \hat \theta\vert^2} &= 1+\frac{4d^2}{n^2(2d+4\theta)} \\
             \end{align*}
+            Which clearly goes to infinity, for any set $n$, as $d$ goes to infinity.
+        \item[(d)]
+                \begin{align*}
+                    P(\mu \vert \mathcal{D}_n) &= \frac{\mathcal{L}_n(\mu) \pi(\mu)}{\int_{\R^d}\mathcal{L}_n(\mu) \pi(\mu) \, d\mu} \\
+                    P(\mu \vert \mathcal{D}_n) &= \frac{\prod_{i=1}^n \frac{1}{\sqrt{(2\pi)^d}}\exp\left(-\frac{1}{2}\|x_i-\mu\|^2\right)\frac{1}{\sqrt{(2\pi\tau^2)^d}}\exp\left(-\frac{1}{2\tau^2}\|\mu\|^2\right)}{\int_{\R^d}\prod_{i=1}^n\frac{1}{\sqrt{(2\pi)^d}} \exp\left(-\frac{1}{2}\|x_i-\mu\|^2\right) \frac{1}{\sqrt{(2\pi\tau^2)^d}}\exp\left(-\frac{1}{2\tau^2}\|\mu\|^2\right)\, d\mu}  \\
+                    P(\mu \vert \mathcal{D}_n) &= \frac{ \exp\left(-\frac{1}{2}\sum_{i=1}^n\|x_i-\mu\|^2-\frac{1}{2\tau^2}\|\mu\|^2\right)}{\int_{\R^d}\exp\left(-\frac{1}{2}\sum_{i=1}^n\|x_i-\mu\|^2-\frac{1}{2\tau^2}\|\mu\|^2\right) \, d\mu}  \\
+                    P(\mu \vert \mathcal{D}_n) &= \frac{ \exp\left(-\frac{n}{2}\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2 -\frac{1}{2\tau^2}\|\mu\|^2 -\frac{n}{2}\left(\sum_{i=1}^n\frac{\|x_i\|^2}{n}-\left\|\sum_{i=1}^n\frac{x_i}{n}\right\|^2\right)\right)}{\int_{\R^d}\exp\left(-\frac{n}{2}\|\mu-\sum_{i=1}^n\frac{x_i}{n}\|^2-\frac{1}{2\tau^2}\|\mu\|^2 -\frac{n}{2}\left(\sum_{i=1}^n\frac{\|x_i\|^2}{n}-\left\|\sum_{i=1}^n\frac{x_i}{n}\right\|^2\right)\right) \, d\mu}  \\
+                    P(\mu \vert \mathcal{D}_n) &= \frac{ \exp\left(-\frac{n+\tau^2}{n2\tau^2}\|\mu-\frac{\tau^2}{n+\tau^2}\sum_{i=1}^n\frac{x_i}{n}\|^2\right)  \, d\mu}{\int_{\R^d}\exp\left(-\frac{n+\tau^2}{n2\tau^2}\|\mu-\frac{\tau^2}{n+\tau^2}\sum_{i=1}^n\frac{x_i}{n}\|^2\right)}  \\
+                    P(\mu \vert \mathcal{D}_n) &= \sqrt{\frac{1}{(2\pi)^d}\left(\frac{n+\tau^2}{n\tau^2}\right)^d} \exp\left(-\frac{n+\tau^2}{n2\tau^2}\left\|\mu-\frac{\tau^2}{n+\tau^2}\sum_{i=1}^n\frac{x_i}{n}\right\|^2\right) \\
+                \end{align*}
+                So, we can conclude that the posterior is distributed as \(N\left(\frac{\tau^2}{n+\tau^2}\sum_{i=1}^n\frac{x_i}{n},\frac{n\tau^2}{n+\tau^2}I_d\right)\).
+This means that \(\frac{n+\tau^2}{n\tau^2}\mu \sim \frac{n+\tau^2}{n\tau^2}\frac{\tau^2}{n+\tau^2}\sum_{i=1}^n\frac{x_i}{n} + Z\), so $\theta$ is distributed as a non-centralized \(\chi^2\frac{n\tau^2}{n+\tau^2}\) distribution with non-centrality parameter \(\left(\frac{\tau^2}{n+\tau^2}\right)^2\left\|\bar X_n\right\|^2\).
+Accordingly, \(\tilde \theta = \left(\frac{\tau^2}{n+\tau^2}\right)^2\left\|\bar X_n\right\|^2+\frac{\tau^2}{n+\tau^2}\frac{d}{n}\).
+Plugging this into the formula from c, we get:
+                \begin{align*}
+                    \frac{\E_\mu\vert \theta - \tilde \theta\vert^2}{\E_\mu\vert \theta - \hat \theta\vert^2} &= \frac{\E_\mu\left\vert \theta - \left(\frac{\tau^2}{n+\tau^2}\right)^2\left\|\bar X_n\right\|^2+\frac{\tau^2}{n+\tau^2}\frac{d}{n}\right\vert^2}{\E_\mu\vert \theta - \|\bar X_n\|^2 + \frac{d}{n}\vert^2}  \\
+                    \frac{\E_\mu\vert \theta - \tilde \theta\vert^2}{\E_\mu\vert \theta - \hat \theta\vert^2} &= \frac{2d+4\theta + \E_\mu\left[\left(\left(1- \left(\frac{\tau^2}{n+\tau^2}\right)^2\right)\left\|\bar X_n\right\|^2+\left(1+\frac{\tau^2}{n+\tau^2}\right)\frac{d}{n}\right)^2\right]}{2d+4\theta}  \\
+                    \frac{\E_\mu\vert \theta - \tilde \theta\vert^2}{\E_\mu\vert \theta - \hat \theta\vert^2} &= \frac{O(d^2)}{2d+4\theta} 
+                \end{align*}
+                This quantity once again goes to infinity for a given $n$ while $d$ goes to infinity.
+        \item[(e)] These are the histograms: \\
+            \includegraphics[width=9cm]{hw4/4_bayes} 
+            \includegraphics[width=9cm]{hw4/4_freq} 
+        \item[(f)] Obviously, the frequentest estimate preforms far better than the Bayesian estimate. It appears to be the case that the mean of a posterior, even with a uniform prior, can be biased.
+
         \end{itemize}
@@
             \item[(a)]
+                Let $\epsilon = .9^{k}$ for some $k\in\N$. Let \(Y_n=\vert\{i:i\leq n,V_i>.1\}\vert\) Since, for any $\alpha$, $P(V_j<.1)<1$,
                 \begin{align*}
@@
                     \sum_{j=1}^n w_j &= 1 - \prod_{j=1}^{n}(1-V_i) \\
-                    \sum_{j=\infty}^n w_j &= 1 - \prod_{j=1}^{\infty}(1-V_j) \\
+                    P\left(\sum_{j=1}^n w_j<1-\epsilon\right) &= P\left(1 - \prod_{j=1}^{n}(1-V_i)<1-\epsilon\right) \\
+                    P\left(\sum_{j=1}^n w_j<1-\epsilon\right) &= P\left(\prod_{j=1}^{n}(1-V_i)>\epsilon\right) \\
+                    P\left(\sum_{j=1}^n w_j<1-\epsilon\right) &\leq P\left(\prod_{j=1}^{n}.9I_{x>.1}(V_j)>\epsilon\right) \\
+                    P\left(\sum_{j=1}^n w_j<1-\epsilon\right) &\leq P\left(.9^{Y_n} >\epsilon\right) \\
+                    P\left(\sum_{j=1}^n w_j<1-\epsilon\right) &\leq P\left(Y_n < k\right) \\
+                    P\left(\sum_{j=1}^n w_j<1-\epsilon\right) &\leq P\left(V_j<.1\right)^{n-k} \\
+                    \lim_{n\to\infty}P\left(\sum_{j=1}^n w_j<1-\epsilon\right) &\leq \lim_{n\to\infty}p\left(V_j<.1\right)^{n-k} \\
+                    \lim_{n\to\infty}P\left(\sum_{j=1}^n w_j<1-\epsilon\right) &\leq 0
                 \end{align*}
-
+                Thus, the weights sum to 1 with probability 1.
         \end{itemize}
+            \item[(b)]
+                If $X\sim F$ and $X_0\sim F_0$, let $F(x)=P(x\leq X)$ and $F_0(x)=P(x\leq X_0)$                
+                    \begin{align*}
+                        \E(F(x)) &= \E(\sum_{j=1}^\infty w_j I{\{z<x\}}(s_j)) \\
+                        \E(F(x)) &= \sum_{j=1}^\infty \E(w_j)\E(I{\{z<x\}}(s_j)) \\
+                        \E(F(x)) &= \sum_{j=1}^\infty \E(w_j)F_0(x) \\
+                        \E(F(x)) &= \E(\sum_{j=1}^\infty w_jF_0(x)) \\
+                        \E(F(x)) &= F_0(x) 
+                    \end{align*}
+                So we may conclude \(\E(F)=F_0\). Additionally, 
+                    \begin{align*}
+                        \var(F(x)) &= \E(F(x)^2) - \E(F(x))^2 \\
+                        \var(F(x)) &= \var\left(\sumj w_jI_{\{z\leq s_j\}}(x)\right) \\
+                        \var(F(x)) &= \E(F(x)^2) - \E(F(x))^2 \\
+                        \var(F(x)) &= \E\left(\left(\sumj w_jI_{\{z\leq s_j\}}(x)\right)^2\right) - F_0(x)^2 \\
+                        \var(F(x)) &= \sumj\sumk \E\left(w_jw_kI_{\{z\leq s_j\}}(x)I_{\{z\leq s_k\}}(x)\right) - F_0(x)^2 
+                    \end{align*}
+                \item[(c)]
+                    \begin{align*}
+                        P\left(\sup_x\left\vert \bar F_n(x)-F(x)\right\vert>\epsilon\right) &=P\left(\sup_x\left\vert \frac{n}{n+\alpha}F_n(x)+\frac{\alpha}{n+\alpha}F_0(x)-F(x)\right\vert>\epsilon\right) \\
+                        P\left(\sup_x\left\vert \bar F_n(x)-F(x)\right\vert>\epsilon\right) &\leq P\left(\sup_x\left\vert \frac{n}{n+\alpha}F_n(x)-\frac{n}{n+\alpha}F(x)\right\vert + \sup_x\left\vert \frac{\alpha}{n+\alpha}F_0(x)-\frac{\alpha}{n+\alpha}F(x)\right\vert>\epsilon\right) \\
+                        P\left(\sup_x\left\vert \bar F_n(x)-F(x)\right\vert>\epsilon\right) &\leq P\left(\frac{n}{n+\alpha}\sup_x\left\vert F_n(x)-F(x)\right\vert + \frac{\alpha}{n+\alpha}\sup_x\left\vert F_0(x)-F(x)\right\vert>\epsilon\right) \\
+                        P\left(\sup_x\left\vert \bar F_n(x)-F(x)\right\vert>\epsilon\right) &\leq P\left(\frac{n}{n+\alpha}\sup_x\left\vert F_n(x)-F(x)\right\vert>\epsilon - \frac{\alpha}{n+\alpha}\sup_x\left\vert F_0(x)-F(x)\right\vert\right)
+                    \end{align*}
+                    Thus, using the DKW inequality,
+                    \[ P\left(\sup_x\left\vert \bar F_n(x)-F(x)\right\vert>\epsilon\right)\leq 
+                        \begin{cases}
+                            2e^{-2n\left(\epsilon - \frac{\alpha}{n+\alpha}\sup_x\left\vert F_0(x)-F(x)\right\vert \right)^2} & \text{if } \epsilon > \frac{\alpha}{n+\alpha}\sup_x\left\vert F_0(x)-F(x)\right\vert \\
+                            1 & \text{otherwise}
+                        \end{cases}
+                    \]
+                    If the prior is really poor and $n$ isn't sufficiently large, the posterior can be arbitrarily bad.
 

### 2014-12-02 stat374/hw4.R

@@ -14,8 +14,8 @@ library('wavethresh')
 ### Problems
 problem1<-F
 problem2<-F
-problem3<-T
-problem4<-F
+problem3<-F
+problem4<-T
 problem5<-F
 
 ### Options
@@ -223,10 +223,27 @@ if (problem3) {
     matplot(y=da.wav.mat,x=fs$day,col=c(1,1),type="l",lty=c(1,2),xlab="Day",ylab="Log Mortality Rate",main="Daubechies Wavelet Fit")
     legend('topright',c("True Rate","Daubechies Wavelet Fit"),lty=c(1,2))
     dev.off()
-
-        
-    
     
 }
+############# Problem 4 
+if (problem4) {
+    n<-10
+    d<-1000
+    b<-1000
+    est<-matrix(,nrow=b,ncol=2)
+    for (i in 1:b) {
+        x<-replicate(n,rnorm(d))
+        xbar<-apply(x,1,mean)
+        est[i,]<-sum(xbar^2)+c(d/n,-d/n)  
+    }
+    pdf('hw4/4_bayes.pdf')
+    lim<-c(min(est),max(est))
+    hist(est[,1],main="Bayes Estimate",xlab="Prediction",xlim=lim)
+    dev.off()
+    pdf('hw4/4_freq.pdf')
+    hist(est[,2],main="Frequentist Estimate",xlab="Prediction",xlim=lim)
+    dev.off()
+}
+
 
 

### 2014-12-02 stat374/hw4.tex
Done hw4

@@
                 Thus, the weights sum to 1 with probability 1.
-        \end{itemize}
             \item[(b)]
@@
                     \end{align*}
-                So we may conclude \(\E(F)=F_0\). Additionally, 
-                    \begin{align*}
-                        \var(F(x)) &= \E(F(x)^2) - \E(F(x))^2 \\
-                        \var(F(x)) &= \var\left(\sumj w_jI_{\{z\leq s_j\}}(x)\right) \\
-                        \var(F(x)) &= \E(F(x)^2) - \E(F(x))^2 \\
-                        \var(F(x)) &= \E\left(\left(\sumj w_jI_{\{z\leq s_j\}}(x)\right)^2\right) - F_0(x)^2 \\
-                        \var(F(x)) &= \sumj\sumk \E\left(w_jw_kI_{\{z\leq s_j\}}(x)I_{\{z\leq s_k\}}(x)\right) - F_0(x)^2 
-                    \end{align*}
+                So we may conclude \(\E(F)=F_0\).
+To see that the prior gets more concentrated around $F_0$ as $\alpha$ increases, we only need demonstrate that the probability the drawn distribution cumulative distribution function function differs by more than $\epsilon$ from $F_0$ is decreasing as $\alpha$ increases.
+To do this, we note that this probability is the sum of the probabilities that the difference is achieved in exactly a certain number of steps of the stick break algorithm.
+This probability, conditioned on a particular sum of weights, decreases as the sum of weights decreases.
+Thus, since the probability of a partial sum of weights having reached a particular length decreases with $\alpha$ (as the expected size of the weights decreases), the probability of the difference occurring in a particular number of steps decreases as well, and in turn the probability of achieving that difference in any number of steps.
                 \item[(c)]
@@
 
+                \item[(d)] These are my plots:
+                    \begin{center}
+                        \includegraphics[width=9cm]{hw4/5_d} 
+                    \end{center}
+                \item[(e)] 
+                    \begin{itemize}
+                        \item[i.] These are the plots, for $\alpha=100$ \\
+                            \includegraphics[width=9cm]{hw4/5_ei10} 
+                            \includegraphics[width=9cm]{hw4/5_ei25}  \\
+                            \includegraphics[width=9cm]{hw4/5_ei100}
+                    \end{itemize}[i.] These are the plots of the posteriors, their CIs, and the draws from them, with $\alpha=100$ \\
+                            \includegraphics[width=9cm]{hw4/5_eii10} 
+                            \includegraphics[width=9cm]{hw4/5_eii25}  \\
+                            \includegraphics[width=9cm]{hw4/5_eii100}
+        \end{itemize}
+
 \end{itemize}

### 2014-12-02 stat374/hw4.R

@@ -15,8 +15,8 @@ library('wavethresh')
 problem1<-F
 problem2<-F
 problem3<-F
-problem4<-T
-problem5<-F
+problem4<-F
+problem5<-T
 
 ### Options
 procs<-4
@@ -245,5 +245,68 @@ if (problem4) {
     dev.off()
 }
 
+############# Problem 5 
+if (problem5) {
+    dirc<-function(rfunc,alpha,n,...) {
+        step.mat<-matrix(,nrow=n,ncol=2)
+        step.mat[,1]<-rfunc(n,...)
+        V<-rbeta(n,1,alpha) 
+        step.mat[,2]<-cumprod(c(1,1-V))[-(n+1)]*V
+        order.mat <- step.mat[order(step.mat[,1]),] 
+        return(stepfun(order.mat[,1],cumsum(c(0,order.mat[,2]))))
+    }
+    
+    ### d
+    alpha<-c(1,1,1,5,5,5,20,20,50,50)
+    n<-200
+    pp<-seq(-4,4,length.out=300)
+    mat<-matrix(,nrow=300,10)
+    for (i in 1:10) {
+        func<-dirc(rnorm,alpha[i],n)
+        mat[,i]<-func(pp)
+    }
+    pdf('hw4/5_d.pdf')
+    matplot(x=pp,y=mat,type="l",col=rep(1,10),lty=c(1,1,1,2,2,2,3,3,4,4),main="Dirchlet Processes",xlab="x",ylab="Cumulative Probability")
+    legend('bottomright',paste("Alpha = ",c(1,5,20,50)),lty=1:4)
+    dev.off()
 
+    ### e
+    pp<-seq(-7,17,length.out=300)
+    type2<-list()
+    for (n in c(10,25,100)) {
+        #i
+        mat<-matrix(,nrow=300,ncol=3)
+        X<-sort(rnorm(n,5,3))
+        mat[,1]<-stepfun(X,(0:n)/n)(pp)
+        e<-sqrt(log(2/.05)/(2*n))
+        mat[,2]<-pmax(mat[,1]-e,0)
+        mat[,3]<-pmin(mat[,1]+e,1)
+        pdf(paste('hw4/5_ei',n,'.pdf',sep=""))
+        matplot(mat,pp,type="l",col=rep(1,3),lty=c(1,3,3),main=paste("Emperical Distribution & DKW, N=",n),xlab="x",ylab="Cumulative Probability")
+        dev.off()
+        
+        #ii 
+        alpha<-100
[…]

### 2014-12-02 stat374/hw2.R
Added homework 2,3 R code

@@ -1,0 +1,190 @@
+library('parallel')
+library('locfit')
+library('gam')
+
+############################################
+# Problem 1
+############################################
+### part c)
+# Input data and transform
+Y<-c(.4,.378,.356,.333,.311,.311,.289,.267,.244,.244,.222,.222,.222,.222,.222,.200,.178,.156)
+p<-c(.346,.298,.276,.222,.273,.270,.263,.201,.269,.230,.264,.256,.303,.264,.226,.285,.316,.200)
+Z<-sqrt(45)*asin(2*Y-1)
+th<-sqrt(45)*asin(2*p-1)
+
+# Compute estimators and calculate risk
+n<-length(Y)
+mean <- mean(Z)
+mle <- Z 
+JS0 <- Z*(1-(n-2)*1/(sum(Z^2)))
+JSm <- mean + (Z-mean)*(1-(n-2)*1/(sum((Z-mean)^2)))
+risk <- c(sum((mle-th)^2),sum((JS0-th)^2),sum((JSm-th)^2))
+
+############################################
+# Problem 2 
+############################################
+
+test<-read.table('test.txt',header=TRUE)
+train<-read.table('train.txt',header=TRUE)
+attach(train)
+
+cas.log<-log(casual+1)
+reg.log<-log(registered+1)
+ct.log<-log(count+1)
+season.f<-factor(season)
+holiday.f<-factor(holiday)
+workingday.f<-factor(workingday)
+month.f<-factor(month)
+weather.f<-factor(weather)
+training<-day<=15
+validate<-day<=19 & day>15
+hour.mod<-abs(hour-17)*abs(hour-8)
+hour.mod.2<-abs(hour-17)
+
+my.data<-data.frame(train,cas.log,reg.log,ct.log,season.f,holiday.f,workingday.f,month.f,weather.f,training,validate,hour.mod,hour.mod.2)
+
+reg.lm <- lm(reg.log~season.f*workingday.f+weather.f+hour.mod.2+daylabel, data=my.data[training,])
+cas.lm <- lm(cas.log~season.f*workingday.f+weather.f+hour.mod.2+daylabel, data=my.data[training,])
+
+reg.er<-sum((my.data$reg.log[my.data$validate] -predict(reg.lm,my.data[validate,]))^2)
[…]

### 2014-12-02 stat374/hw3.R

@@ -1,0 +1,149 @@
+library('parallel')
+library('locfit')
+library('gam')
+library('MASS')
+
+############################################
+# Problem 3 
+############################################
+
+fiji<-read.table('fijiquakes.dat',header=TRUE)
+n<-nrow(fiji)
+
+### a)
+mag.den<-density(fiji$mag,bw="nrd",kernel="gaussian")
+h<-mag.den$bw
+
+pdf("hw3_3_a.pdf")
+plot(mag.den,main=paste("Fiji Earthquake Density Estimates, Bandwidth of",format(h,digits=3)),xlab="Magnitude", ylab="Density",lty=1)
+dev.off()
+
+### b)
+mag.ecdf<-ecdf(fiji$mag)
+conf<-sqrt((1/(2*n))*log(2/.05))
+pdf("hw3_3_b.pdf")
+curve(mag.ecdf,xlim=c(3,7),n=1000,main="Empirical Cumulative Density of Magnitudes",xlab="Magnitude", ylab="Cumulative Density")
+curve(pmin(mag.ecdf(x)+conf,1),xlim=c(3,7),n=1000,add=TRUE,lty=2)
+curve(pmax(mag.ecdf(x)-conf,0),xlim=c(3,7),n=1000,add=TRUE,lty=2)
+dev.off()
+
+### c)
+int.mag.den <- function(x) {
+    mat<-matrix(rep(fiji$mag,length(x)),byrow=TRUE,nrow=length(x))
+    out<-rowMeans(pnorm(x,mat,h))
+    return(out)
+}
+
+int.est <- int.mag.den(4.9) - int.mag.den(4.3)
+ecdf.est <- mag.ecdf(4.9) - mag.ecdf(4.3)
+
+ub<-ecdf.est-pnorm(.025)*ecdf.est*(1-ecdf.est)/sqrt(n)
+lb<-ecdf.est+pnorm(.025)*ecdf.est*(1-ecdf.est)/sqrt(n)
+
+############################################
+# Problem 4 
+############################################
+
+ru<-runif(50)
+th<-max(ru)
+n<-50
+b<-1000
+
+### a)
+p.max.unif <- function(x) n*x^{n-1}
+
+np.boots<-replicate(b,sample(ru,50,replace=TRUE))
[…]

### 2014-12-03 stat343/hw8.tex
Finnished HW8

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

STAT 343 Homework 8 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[(a)]
                As we have proved, if the matix $X$ isn't singular, then the least squares solution from OLS is unique and indeed the optimal fit. Thus, if a $\beta$ acheives the minimum least squares, it must be the OLS coefficient from the regression. Calculating the sum of squares for a coefficient $\hat\beta ^{(-i)}$ on $\tilde Y^{(-i)}$, we get 
                \begin{align*}
                    \|\tilde Y^{(-i)} - X\hat\beta^{(-i)}\|^2 &= (x_i\beta^{(-i)} - x_i\beta^{(-i)})^2 + \|Y^{(-i)} - X^{(-i)}\hat\beta ^{(-i)}\|^2 \\
                    \|\tilde Y^{(-i)} - X\hat\beta^{(-i)}\|^2 &= \|Y^{(-i)} - X^{(-i)}\hat\beta^{(-i)}\|^2
                \end{align*}
                Since the sum on the right is the minimum possible over all $\beta$ (since the coefficient was fit by OLS on $Y^{(-i)}$), $\hat\beta ^{(-i)}$ minimizes the sum of squares for $\tilde Y^{(-i)}$. By our earlier reasoning, it thus must be the coefficient that would result from OLS run predicting $\tilde Y^{(-i)}$ in terms of $X$. 
            \item[(b)]
                The prediction of a linear model with coefficient $\hat\beta ^{(-i)}$ at point $x_i^T$ is $x_i^T\hat\beta ^{(-i)}$, so since that is the coefficient for regressing $\tilde Y^{(-i)}$ on $X$, $x_i^T\hat\beta ^{(-i)}$ will be the fit.

        \end{itemize}
    \item[2.]
        \begin{itemize}
            \item[(a)]
                \begin{align*}
                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- S(x_i)^{(-i)}Y^{(-i)} \\
                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- \frac{1}{\sum_{j\neq i} S_{ij}} \sum_{j\neq i} y_j S_{ij} \\
                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- \frac{\hat f(x_i)-y_iS_{ii}(X)}{1-S_{ii}} \\
                    \hat y_i-\hat f^{(-i)}(x_i) &= \frac{y_i - y_i S_{ii} - \hat f (x) + y_iS_{ii}}{1- S_{ii}} \\
                    \hat y_i-\hat f^{(-i)}(x_i) &= \frac{y_i -\hat f (x) }{1- S_{ii}} \\
                \end{align*}
            \item[(b)]
                Rather than refitting the regression $n$ times to estimate the leave one out cross validation estimate of error, this allows the regression to merely be fit once, saving a lot of costly $QR$ factorization or matrix inversion.
            \item[(c)]
                Since $1\geq1-S_{ii}\geq0$, we have 
                \begin{align*}
                    \vert y_i - \hat f(x_i) \vert &\leq \frac{\vert y_i - \hat f(x_i) \vert }{1-S_{ii}} \\
                    \vert y_i - \hat f(x_i) \vert &\leq \left\vert\frac{y_i - \hat f(x_i) }{1-S_{ii}} \right\vert  \\
                    \vert y_i - \hat f(x_i) \vert &\leq \left\vert y_i - \hat f^{(-i)}(x_i)\right\vert  
                \end{align*}

        \end{itemize}
    \item[3.]
        \begin{itemize}
            \item[(a)]
                Applying backwards elimination with a critical value of $.2$, we only remove the variable status. After that, all predictors have a p-value of less than $.2$, so we stop. \\
                \smallskip

                p-value at the point of exclusion: \\
                \FloatBarrier
                \input{hw8/hw8_3_a_excl} 
                \FloatBarrier

                Resulting model: \\
                \FloatBarrier
                \input{hw8/hw8_3_a_lm} 
                \FloatBarrier

            \item[(b)]
                For each of the following three methods, I used an exhaustive subset search, which determined which set of a particular number of variables resulted in the smallest RSS. The result, which is a boolean indicating whether the predictor should be included at each number of predictors is below. Since this will uniformly be the best model of $n$ predictors, the remaining decision is which $n$ to use. It should be noted that the model in a is indeed the optimal model of $3$ predictors. \\
                \FloatBarrier
                \input{hw8/hw8_3_which} 
                \FloatBarrier
                The model with the lowest AIC also had $3$ predictors, and is thus the same as in a. The different AICs for the best model with $n$ predictors  below. \\
                \FloatBarrier
                \input{hw8/hw8_3_b} 
                \FloatBarrier

            \item[(c)]
                The model with the highest adjusted $R^2$ is this same model with $3$ predictors. Optimal adjusted $R^2$ for different number of predictors are below. \\
                \FloatBarrier
                \input{hw8/hw8_3_c} 
                \FloatBarrier

            \item[(d)]
                Going by Mallow's $C_p$, we once again get the same model. The smallest model that had a $C_p$ smaller than the number of parameters (one more than the number of predictors due to the intercept) was the model with 3 predictors (4 parameters): \\
                \FloatBarrier
                \input{hw8/hw8_3_d} 
                \FloatBarrier
        \end{itemize}
    \item[4.]
        The order of these operations doesn't make much difference. Point 21 is influential both before and after model selection, and the prefered model includes Air.Flow and Water.Temp, while excluding Acid.Conc, whether or not the outlier is included. \\
        \smallskip
        Our initial model is below. Acid.Conc looks ripe to be removed with a high $p-value$
        \FloatBarrier
        \input{hw8/hw8_4_lm.tex} 
        \FloatBarrier
        Indeed, when we test all subsets of the data, it is excluded from every model that has fewer than three predictors:
        \FloatBarrier
        \input{hw8/hw8_4_small_which} 
        \FloatBarrier
        The optimal subset, when going by adjusted $R^2$ is the model with just two predictors ($R^2$ for each optimal subset of a particular size below):
        \FloatBarrier
        \input{hw8/hw8_4_small_adjr2} 
        \FloatBarrier
        Choosing this as our model, we get the following model:
        \FloatBarrier
        \input{hw8/hw8_4_small_lm.tex} 
        \FloatBarrier
        Checking the distribution of studentized residuals for outliers, we see one potential culprit (the largest residual, in the top right, which heavily deviates from T-distribution).
        \begin{center}
            \includegraphics[width=9cm]{hw8/hw8_4_small_rstudent} 
        \end{center}
        It is the 21st data point. With a studentized residual value of $-3.47$, it is just inside conservative Bonferroni critical value, for $\alpha=.5$, of $-3.53$. Thus, we don't quite reject it as an outlier. However, looking at the plot of Cook's distances, it is clear that the same point is highly influential.
        \begin{center}
            \includegraphics[width=9cm]{hw8/hw8_4_small_cook}
        \end{center}
        Now, going back to the full model which included Acid.Conc, we look again for outliers and influential points. Shockingly, the same point is once again the only real canidate. The distribution of studentized residuals on this
        \begin{center}
            \includegraphics[width=9cm]{hw8/hw8_4_rstudent} 
        \end{center}
        Testing it as an outlier as before, it has a studentized residual of $-3.33$ in this model, while the critical value is $-3.54$, so we don't reject it. However, its still influential, so we kick it out for that:
        \begin{center}
            \includegraphics[width=9cm]{hw8/hw8_4_cook}
        \end{center}
        Preforming variable selection on the data set with this point excluded, we still find the optimal model has the two variables besides Acid.Conc. Optimal subsets of each size:
        \FloatBarrier
        \input{hw8/hw8_4_no_which} 
        \FloatBarrier
        Adjusted $R^2$ for these subsets, showing we still prefer the model with two predictors:
        \FloatBarrier
        \input{hw8/hw8_4_no_adjr2} 
        \FloatBarrier
    \item[5.]
        \begin{itemize}
            \item[(a)]
                I have plotted the singular values below, gotten by taking the square roots of the absolute values of the eigenvalues of $XX^T$ (once $X$ had been centered). They quickly fall off to very small values, reflecting that most of the variation in the data set is explained by just a few underlying components.
                \begin{center}
                    \includegraphics[width=9cm]{hw8/5_a}
                \end{center}
            \item[(b)]
                The components of the two largest singular vectors are plotted below. As would be expected, the components of the second vector are generally far smaller than those of the first. It appears that approximately half the time the second component has the same sign as the first, and half the time the opposite sign, as might be expected since they are orthogonal. The vaguely resemble piecies of orthogonal waves.  
                \begin{center}
                    \includegraphics[width=9cm]{hw8/5_b}
                \end{center}

            \item[(c)] 
            We see that the regression on the two largest princple components doesn't actually predict the data as well as the two best predictors from the original data set (for the same number of variables, the latter has higher $R^2$). While this latter model predicts almost as well as the model including all variables, the largest principle components are signficantly less predictive, though still reasonably predictive. The reason for this is that the principle components are a decomposition of variation within the predictors; however, the largest source of variation in these predictors isn't necessarily predictive of the outcome. For instance, if there is a lot of variation among a few variables that aren't predictive of the outcome, they may correspond to large principle components, but these principle components will likely be also unpredictive. On the other hand, since the principle components as a whole define a space equivalent to the orginal space defined by all the predictors, a model including all of them will be equally predictive to including all variables.  \\
            \smallskip
            The regression on the two largest principle components ($R^2=.5469$):
            \FloatBarrier
            \input{hw8/hw8_5_pc_lm} 
            \FloatBarrier
            The regression on all variables ($R^2=.6866$):
            \FloatBarrier
            \input{hw8/hw8_5_all_lm} 
            \FloatBarrier
            The regression on age and height ($R^2=.6562$):
            \FloatBarrier
            \input{hw8/hw8_5_sm_lm} 
            \FloatBarrier
        \end{itemize}

\end{itemize}

\end{document}

### 2014-12-03 stat343/hw8.R

@@ -1,0 +1,153 @@
+library('parallel')
+library('alr3')
+library('faraway')
+library('car')
+library('MASS')
+library('locfit')
+library('xtable')
+library('nlme')
+library('quantreg')
+library('splines')
+library('rpart')
+library('randomForest')
+library('leaps')
+
+prob3<-F
+prob4<-F
+prob5<-T
+
+d<-3
+
+if (prob3) {
+    data(teengamb)
+
+    ### a)
+    crit<-.2
+    exclude<-NULL
+    excl.pval<-NULL
+    go<-TRUE
+    p.lm<-lm(gamble~.,data=teengamb)
+
+    while (go) {
+        p.lm<-lm(gamble~.,data=teengamb[!(names(teengamb) %in% exclude)])
+        pvals<-summary(p.lm)$coefficients[-1,4]
+        max.ind <- which.max(pvals)
+        max.pval <- pvals[max.ind]
+        if (max.pval>crit) {
+            excl.pval<-c(excl.pval,max.pval)
+            exclude<-c(exclude,names(max.ind))
+        }
+        else { 
+            go<-FALSE
+        }
+    }
+    ecl.ma<-cbind(1:length(excl.pval),excl.pval)
+    colnames(ecl.ma)<-c("Order","p-value")
+    print(xtable(summary(p.lm),digits=d),type="latex",file="hw8/hw8_3_a_lm.tex")
+    print(xtable(ecl.ma,digits=d),type="latex",file="hw8/hw8_3_a_excl.tex")
+        
+    ### b,c,d)
+    subs<-regsubsets(gamble~.,data=teengamb)
+    sum<-summary(subs)
+    varname<-colnames(sum$which)
+    print(xtable(sum$which,digits=d),type="latex",file="hw8/hw8_3_which.tex")
+
+    aic<-nrow(teengamb)*log(sum$rss/nrow(teengamb)) + (2:ncol(teengamb))*2
+    aic.min<-min(aic)
+    aic.excl<-!sum$which[which.min(aic),]
+    aic.lm<-lm(gamble~.,data=teengamb[!names(teengamb)%in% varname[aic.excl]])
+    aic.tbl<-rbind(1:(length(varname)-1),aic)
+    rownames(aic.tbl)<-c("Predictors","AIC")
[…]

### 2014-12-05 stat343/final.tex
Started final for 343

--- 2014-12-03 stat343/hw8.tex
@@
 \begin{document}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-STAT 343 Homework 8 \hfill Aaron Maurer
-\vspace{2mm}
-\hrule
-\vspace{2mm}
-
-\begin{itemize}
-    \item[1.]
-        \begin{itemize}
-            \item[(a)]
-                As we have proved, if the matix $X$ isn't singular, then the least squares solution from OLS is unique and indeed the optimal fit.
-Thus, if a $\beta$ acheives the minimum least squares, it must be the OLS coefficient from the regression.
-Calculating the sum of squares for a coefficient $\hat\beta ^{(-i)}$ on $\tilde Y^{(-i)}$, we get
-                \begin{align*}
-                    \|\tilde Y^{(-i)} - X\hat\beta^{(-i)}\|^2 &= (x_i\beta^{(-i)} - x_i\beta^{(-i)})^2 + \|Y^{(-i)} - X^{(-i)}\hat\beta ^{(-i)}\|^2 \\
-                    \|\tilde Y^{(-i)} - X\hat\beta^{(-i)}\|^2 &= \|Y^{(-i)} - X^{(-i)}\hat\beta^{(-i)}\|^2
-                \end{align*}
-                Since the sum on the right is the minimum possible over all $\beta$ (since the coefficient was fit by OLS on $Y^{(-i)}$), $\hat\beta ^{(-i)}$ minimizes the sum of squares for $\tilde Y^{(-i)}$.
-By our earlier reasoning, it thus must be the coefficient that would result from OLS run predicting $\tilde Y^{(-i)}$ in terms of $X$.
-            \item[(b)]
-                The prediction of a linear model with coefficient $\hat\beta ^{(-i)}$ at point $x_i^T$ is $x_i^T\hat\beta ^{(-i)}$, so since that is the coefficient for regressing $\tilde Y^{(-i)}$ on $X$, $x_i^T\hat\beta ^{(-i)}$ will be the fit.
-
-        \end{itemize}
-    \item[2.]
-        \begin{itemize}
-            \item[(a)]
-                \begin{align*}
-                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- S(x_i)^{(-i)}Y^{(-i)} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- \frac{1}{\sum_{j\neq i} S_{ij}} \sum_{j\neq i} y_j S_{ij} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- \frac{\hat f(x_i)-y_iS_{ii}(X)}{1-S_{ii}} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= \frac{y_i - y_i S_{ii} - \hat f (x) + y_iS_{ii}}{1- S_{ii}} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= \frac{y_i -\hat f (x) }{1- S_{ii}} \\
-                \end{align*}
-            \item[(b)]
-                Rather than refitting the regression $n$ times to estimate the leave one out cross validation estimate of error, this allows the regression to merely be fit once, saving a lot of costly $QR$ factorization or matrix inversion.
-            \item[(c)]
-                Since $1\geq1-S_{ii}\geq0$, we have 
-                \begin{align*}
-                    \vert y_i - \hat f(x_i) \vert &\leq \frac{\vert y_i - \hat f(x_i) \vert }{1-S_{ii}} \\
-                    \vert y_i - \hat f(x_i) \vert &\leq \left\vert\frac{y_i - \hat f(x_i) }{1-S_{ii}} \right\vert  \\
-                    \vert y_i - \hat f(x_i) \vert &\leq \left\vert y_i - \hat f^{(-i)}(x_i)\right\vert  
-                \end{align*}
-
-        \end{itemize}
-    \item[3.]
-        \begin{itemize}
-            \item[(a)]
-                Applying backwards elimination with a critical value of $.2$, we only remove the variable status. After that, all predictors have a p-value of less than $.2$, so we stop. \\
-                \smallskip
-
-                p-value at the point of exclusion: \\
-                \FloatBarrier
-                \input{hw8/hw8_3_a_excl} 
-                \FloatBarrier
-
-                Resulting model: \\
-                \FloatBarrier
-                \input{hw8/hw8_3_a_lm} 
-                \FloatBarrier
-
-            \item[(b)]
-                For each of the following three methods, I used an exhaustive subset search, which determined which set of a particular number of variables resulted in the smallest RSS.
-The result, which is a boolean indicating whether the predictor should be included at each number of predictors is below.
-Since this will uniformly be the best model of $n$ predictors, the remaining decision is which $n$ to use.
-It should be noted that the model in a is indeed the optimal model of $3$ predictors.
-\\
-                \FloatBarrier
-                \input{hw8/hw8_3_which} 
-                \FloatBarrier
-                The model with the lowest AIC also had $3$ predictors, and is thus the same as in a. The different AICs for the best model with $n$ predictors  below. \\
-                \FloatBarrier
-                \input{hw8/hw8_3_b} 
-                \FloatBarrier
-
-            \item[(c)]
-                The model with the highest adjusted $R^2$ is this same model with $3$ predictors. Optimal adjusted $R^2$ for different number of predictors are below. \\
-                \FloatBarrier
-                \input{hw8/hw8_3_c} 
-                \FloatBarrier
-
-            \item[(d)]
-                Going by Mallow's $C_p$, we once again get the same model.
-The smallest model that had a $C_p$ smaller than the number of parameters (one more than the number of predictors due to the intercept) was the model with 3 predictors (4 parameters): \\
-                \FloatBarrier
-                \input{hw8/hw8_3_d} 
-                \FloatBarrier
-        \end{itemize}
-    \item[4.]
-        The order of these operations doesn't make much difference.
-Point 21 is influential both before and after model selection, and the prefered model includes Air.Flow and Water.Temp, while excluding Acid.Conc, whether or not the outlier is included.
-\\
-        \smallskip
-        Our initial model is below. Acid.Conc looks ripe to be removed with a high $p-value$
-        \FloatBarrier
-        \input{hw8/hw8_4_lm.tex} 
-        \FloatBarrier
-        Indeed, when we test all subsets of the data, it is excluded from every model that has fewer than three predictors:
-        \FloatBarrier
-        \input{hw8/hw8_4_small_which} 
-        \FloatBarrier
-        The optimal subset, when going by adjusted $R^2$ is the model with just two predictors ($R^2$ for each optimal subset of a particular size below):
-        \FloatBarrier
-        \input{hw8/hw8_4_small_adjr2} 
-        \FloatBarrier
-        Choosing this as our model, we get the following model:
-        \FloatBarrier
-        \input{hw8/hw8_4_small_lm.tex} 
-        \FloatBarrier
-        Checking the distribution of studentized residuals for outliers, we see one potential culprit (the largest residual, in the top right, which heavily deviates from T-distribution).
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_small_rstudent} 
-        \end{center}
-        It is the 21st data point.
-With a studentized residual value of $-3.47$, it is just inside conservative Bonferroni critical value, for $\alpha=.5$, of $-3.53$.
-Thus, we don't quite reject it as an outlier.
-However, looking at the plot of Cook's distances, it is clear that the same point is highly influential.
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_small_cook}
-        \end{center}
-        Now, going back to the full model which included Acid.Conc, we look again for outliers and influential points.
-Shockingly, the same point is once again the only real canidate.
-The distribution of studentized residuals on this
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_rstudent} 
-        \end{center}
-        Testing it as an outlier as before, it has a studentized residual of $-3.33$ in this model, while the critical value is $-3.54$, so we don't reject it.
-However, its still influential, so we kick it out for that:
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_cook}
-        \end{center}
-        Preforming variable selection on the data set with this point excluded, we still find the optimal model has the two variables besides Acid.Conc. Optimal subsets of each size:
-        \FloatBarrier
-        \input{hw8/hw8_4_no_which} 
-        \FloatBarrier
-        Adjusted $R^2$ for these subsets, showing we still prefer the model with two predictors:
-        \FloatBarrier
-        \input{hw8/hw8_4_no_adjr2} 
-        \FloatBarrier
-    \item[5.]
-        \begin{itemize}
-            \item[(a)]
-                I have plotted the singular values below, gotten by taking the square roots of the absolute values of the eigenvalues of $XX^T$ (once $X$ had been centered).
-They quickly fall off to very small values, reflecting that most of the variation in the data set is explained by just a few underlying components.
-                \begin{center}
-                    \includegraphics[width=9cm]{hw8/5_a}
-                \end{center}
-            \item[(b)]
-                The components of the two largest singular vectors are plotted below.
-As would be expected, the components of the second vector are generally far smaller than those of the first.
-It appears that approximately half the time the second component has the same sign as the first, and half the time the opposite sign, as might be expected since they are orthogonal.
-The vaguely resemble piecies of orthogonal waves.
-                \begin{center}
-                    \includegraphics[width=9cm]{hw8/5_b}
-                \end{center}
-
-            \item[(c)] 
-            We see that the regression on the two largest princple components doesn't actually predict the data as well as the two best predictors from the original data set (for the same number of variables, the latter has higher $R^2$).
-While this latter model predicts almost as well as the model including all variables, the largest principle components are signficantly less predictive, though still reasonably predictive.
-The reason for this is that the principle components are a decomposition of variation within the predictors; however, the largest source of variation in these predictors isn't necessarily predictive of the outcome.
-For instance, if there is a lot of variation among a few variables that aren't predictive of the outcome, they may correspond to large principle components, but these principle components will likely be also unpredictive.
-On the other hand, since the principle components as a whole define a space equivalent to the orginal space defined by all the predictors, a model including all of them will be equally predictive to including all variables.
-\\
-            \smallskip
-            The regression on the two largest principle components ($R^2=.5469$):
-            \FloatBarrier
-            \input{hw8/hw8_5_pc_lm} 
-            \FloatBarrier
-            The regression on all variables ($R^2=.6866$):
-            \FloatBarrier
-            \input{hw8/hw8_5_all_lm} 
-            \FloatBarrier
-            The regression on age and height ($R^2=.6562$):
-            \FloatBarrier
-            \input{hw8/hw8_5_sm_lm} 
-            \FloatBarrier
-        \end{itemize}
-
-\end{itemize}
-
 \end{document}

### 2014-12-05 stat343/final.R

@@ -1,0 +1,22 @@
+library('parallel')
+library('alr3')
+library('faraway')
+library('car')
+library('MASS')
+library('locfit')
+library('xtable')
+library('nlme')
+library('quantreg')
+library('splines')
+library('rpart')
+library('randomForest')
+library('leaps')
+
+prob1<-T
+prob2<-F
+prob3<-F
+
+if (prob1) {
+    
+    
+}

### 2014-12-10 stat343/final.tex
Done with final besides some editing

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+STATS 343 Final \hfill Aaron Maurer
+\vspace{2mm}
+\hrule
+\vspace{2mm}
+
+\begin{itemize}
+    \item[1.]
+        \begin{itemize}
+            \item[a)]
+                {\bf Summary:} Fitting a simple linear model, I see that the constant variance assumptions don't hold, which would suggest transformation.
+The normally distributed error also isn't quite true, but the difference is fairly minor.
+The first choice for transformation the Box-Cox series of transformations, but the 95% confidence interval for the maximum likelihood includes 1, suggesting that such a transformation isn't worth while.
+Instead, what ended up working was to normalize all the variables so that their smallest value was $0$ and the largest $1$, and then using the square root of the variables, or the square root of one minus the variable, as predictors.
+This results in improvement, giving residuals with approximately constant variance.
+\par
+
+                {\bf Work:} I ran the simple linear model:
+                \lstinputlisting[firstline=4,lastline=23]{final/1_out.txt}
+                Considering possible problems other than nonlinearity, I considered the QQ plot of the residuals and a plot of the square root of the absolute values of the residuals versus the predicted values: \\
+                \includegraphics[width=9cm]{final/1a_qq} 
+                \includegraphics[width=9cm]{final/1a_sqrt_res_plot} \\ 
+                In the QQ plot, we see that the distribution of the residuals has much somewhat thicker tails than a normal distribution, though stays pretty close otherwise.
+The Shapiro-Wilkes test, with a p-value of $.0003$ has us reject normality, but the predictions are still unbiased if the normality assumption doesn't hold, and the deviation is pretty small.
+The variance of the residuals doesn't look bad, but the kernel regression fit seems to show some meaningful variation over the range of predictions.
+Thus, I ran a new model where, after scaling the variables so they all go from $0$ to $1$, I transformed the predictors, taking the square root of one minus angle, chord length, and thickness, and the square root of velocity.
+This resulted in the following model: \\
+                \lstinputlisting[firstline=40,lastline=60]{final/1_out.txt}
+                Checking the fit again, I see some mild improvement: \\
+                \includegraphics[width=9cm]{final/1a_qq_lm2} 
+                \includegraphics[width=9cm]{final/1a_sqrt_res_plot_lm2} \\ 
+                The Residuals are a tad closer to being normally distributed (the lower tail deviates a little less), but more importantly, as one can see from the linear fit, the variance appears to vary much less over the range of predictions, and looks close to being constant, except maybe at the end.
+            \item[b)]
+                {\bf Summary:} We seem to have some pretty meaningful nonlinearity in the model from part a.
+To improve upon it, I transform the outcome, pressure, by taking it square root, and replace the terms for frequency, angle, and chord length with a fourth degree polynomial of those three variables together.
+This reduced the non-linearity and vastly improved the overall fit, as judged by adjusted R-squared.
+\par
+                {\bf Work:} I checked for non-linearity in the model from part a by plotting the residuals by predicted value and a few of the variables: \\
+                \includegraphics[width=9cm]{final/1b_res_plot_lm2} 
+                \includegraphics[width=9cm]{final/1b_res_plot_tran_ang} \\ 
+                \includegraphics[width=9cm]{final/1b_res_plot_tran_freq}  
+                \includegraphics[width=9cm]{final/1b_res_plot_tran_thi} \\ 
+                As you can see, the local linear fits pretty strongly show that the residuals don't have a mean of $0$ over the range of predicted values nor range of frequency, even if it looks OK for angle and thickness.
+I couldn't find any way to remedy this without interactions; I tried a number of models which had every predictor as a factor and some kind of transformation of the outcome, since this would admit every possible transformation of each predictor, but every model still failed to have residuals with mean $0$ over the range of predicted values.
+Thus, I introduced a set of orthogonal polynomials of degree up to 4 in terms of frequency, angle, and chord length together, as well as took the square root of pressure.
+This is the resulting model summary: \\
+                \lstinputlisting[firstline=76,lastline=127]{final/1_out.txt}
+                There are a lot of variables added, but not that many considering the size of the data set, and we see a big overall improvement in fit, as shown by the adjusted $R^2$ increasing to $.813$ from $.477$.
+Looking at plots of the residuals, we also seem to have mostly mitigated the issues with nonlinearity: \\
+                \includegraphics[width=9cm]{final/1b_res_plot_int} 
+                \includegraphics[width=9cm]{final/1b_res_plot_int_ang} \\ 
+                \includegraphics[width=9cm]{final/1b_res_plot_int_freq}  
+                \includegraphics[width=9cm]{final/1b_res_plot_int_thi} \\ 
+                While there is still some nonlinearity, except in a few spots with sparse data, the mean is consistently right around $0$.
+We do seem to have meaningful non-zero means at lower frequencies, and at the higher end of the predicted values, but these biases are small, and much better than with the previous model.
+Double checking the variance, we seem to still have close to constant variance, with possibly a little less consistency than we had before: \\
+                \begin{center}
+                    \includegraphics[width=9cm]{final/1b_sqrt_res_plot_int} 
+                \end{center}
+            \item[c)]
+                Of these two, each has its own advantages.
+The model from part b has the advantage of simplicity, making it suitable for explanation.
+The confidence intervals obviously won't be perfect, since the predictions are biased and the normality assumption doesn't hold.
+However, the confidence level is high enough that we can still fairly safely say that the predicted positive or negative correlations between the predictors and outcome hold true, and the estimates of how much this is by serves a sufficient ballpark guess.
+On the other hand, the model from b does seem to fulfill the assumptions of a linear regression model mostly (except possibly normality) and has high $R^2$, so it is good for prediction, even if the confidence intervals of the prediction might be slightly off.
+It is just convoluted enough that it is hard to interpret what effect angle, frequency, and chord length have on pressure without making a visual, since there are no single coefficients on these predictors..
+            \item[d)]
+                Fitting a tree model, I pruned the tree at a complexity parameter of $3.31*10^{-6}$.
+Any split which didn't decrease total $R^2$ by at least that much was discarded.
+This was chosen because it was the points at which $10$ fold cross validation estimated there would be the lowest error.
+This gave a tree with $1121$ splits for $1502$ data points, which seems like a lot, but when one calculates the adjusted $R^2$, the result is still an impressive $.999$.
+Between this and the cross validation, we are reasonably assured that this is not an over fit.
+We can look at the distribution of the residuals:\\
+                \includegraphics[width=9cm]{final/1d_res_plot_tree} 
+                \includegraphics[width=9cm]{final/1d_qq_tree} \\
+                As we would expect, by construction, the residuals definitely have mean $0$, with a large number of them, in fact, being $0$.
+Additionally, the residuals are no where close to normally distributed, but this is to be expected with so many of them already at $0$.
+            \item[e)]
+                As we we have seen here, in data that doesn't naturally fit the assumptions necessary for linear regression, its hard to transform it to the point where all the assumptions hold well.
+One can add an arbitrary amount of flexibility to a linear model to make it work, through splines, polynomials, other spanning function bases, and the like, but this has to be done manually.
+By the point enough flexibility has been added, often there is no easy interpretation of coefficients.
+Here in lies the advantage of a tree model; it has arbitrary amounts of flexibility built in by default, making it appropriate and easier for prediction with this kind of messy data.
+If you happen to stumble on a data set that actually has a linear relationship between two variables though, a tree model will never predict quite as well nor be as interpretable as the linear regression.
+
+        \end{itemize}
+
+    \item[2.]
+        \begin{itemize}
+            \item[a)]
+                This is the summary from the linear model: \\ 
+                \lstinputlisting[firstline=5,lastline=21]{final/2_out.txt}
+
+                \begin{itemize}
+                    \item[i.]
+                        This residuals are pretty close to normal for the most part, but is different enough that we can say with a high degree of certainty that weren't actually drawn from normal distribution.
+Looking at the QQ plot of the residuals, the quantiles of residuals line up pretty well with normal quantiles except at the lower tail (perfect correspondence is falling along the line).
+At the lower tail though, there is immense deviation from the normal quantiles.
+With such a huge data set, such a large deviation for so many points is extremely unlikely.
+The end result is that we can expect the estimated confidence intervals (from the normal assumption) will be somewhat off.
+                        \begin{center}
+                            \includegraphics[width=9cm]{final/2ai_qq} 
+                        \end{center}
+
+                    \item[ii.]
+                        There are unquestionably pretty significant outliers.
+We can see this by looking at the Cook distance and the studentized residuals.
+I have plotted a quantile plot of the studentized residuals below.
+The Bonferroni Critical value for the absolute value of studentized residuals an alpha of $.05$, which is the \(\frac{1-.05/2}{n}\) quantile of a t distribution with n-p degrees of freedom, is $4.5581$.
+By comparison, looking at the plot, we can see that there are a number of studentized residuals with an absolute value greater than $4.5581$, going as high as $9.5778$.
+Since this critical value is conservative, we definitely have a number of outliers.
+                        \begin{center}
+                            \includegraphics[width=9cm]{final/2aii_stud_qq} 
+                        \end{center}
+                        Looking at this another way, we consider the Cook distances, plotted against the half-normal quantiles.
+We certainly seem to have some Cook distances which stand out; this further supports the notion we're looking at outliers.
+                        \begin{center}
+                            \includegraphics[width=9cm]{final/2aii_cook} 
+                        \end{center}
+                    \item[iii.]
+                        The linearity assumption seems to hold up for the most part.
+The first check is that the residuals really do have mean $0$ throughout the range of predictions.
+Plotting the residuals against the predicted values and fitting a flexible kernel regression model, we do seem to have very close to mean $0$, except at the lowest end of the predicted range; the local regression is very close to $0$ everywhere else.
+                        \begin{center}
+                            \includegraphics[width=9cm]{final/2aiii_res_plot} 
+                        \end{center}
+                        Additionally, checking the partial residual plot of each variable in turn, all of them the linear fits seem to be extremely good.
+If there isn't a linear relationship between one of the variables and the outcome, the difference isn't much.
+\\
+                        \includegraphics[width=9cm]{final/2aiii_pr_AP} 
+                        \includegraphics[width=9cm]{final/2aiii_pr_AT} \\ 
+                        \includegraphics[width=9cm]{final/2aiii_pr_RH} 
+                        \includegraphics[width=9cm]{final/2aiii_pr_V} \\ 
+                \end{itemize}
+            \item[b)]
+                For both of these statistics, I used a parametric bootstrap methodology to come up with critical values.
+For each boot, I drew a new set of outcomes $Y^*$, each one drawn from a normal distribution whose mean was the prediction from the original linear model for the given observation, and whose variance was the estimated variance from the original linear model.
+Using $Y^*$ and the original $X$, I fit a new linear model, and then calculated the statistic (Cook Distance or absolute studetnized residual) from this model for all the observations, keeping the $q$ largest values.
+After this, the critical value for the $q$th largest value at a p-value of $alpha$ is the $1-\alpha$ percentile of the $q$th largest values from all the boots.
+In other words, it was chosen so it is larger than $q$th largest value for exactly $\alpha$ of the boots.
+                \begin{itemize}
+                    \item[i.]
+                        Using the method described above, run for 2000 total boots, I generated critical values for the $1$st through $10$th largest Cook Distances.
+I have plotted them below for critical values less than $.1$.
+Obviously the critical values should asymptotically approach infinity as they go to $0$.
+This is only a finite simulation though, so little should be read into the estimates extremely close to $0$.
+                        \begin{center}
+                            \includegraphics[width=9cm]{final/2bi_plot} 
+                        \end{center}
+                        Comparing the $1$st-$10$th largest Cook Distances from the original model:\\ 
+                        \lstinputlisting[firstline=24,lastline=25]{final/2_out.txt} 
+                        To the respective critical values: \\
+                        \lstinputlisting[firstline=27,lastline=30]{final/2_out.txt} 
+                        We see that they all fall outside of the critical values for all confidence levels, so we would reject them all as outliers for all confidence levels.
+
+                    \item[ii.]
+                        Using the method described above, once again for 2000 total boots, but this time for studentized residuals, I have once again plotted the critical values.
+The same issue for critical values close to 0 exists.
+                        \begin{center}
+                            \includegraphics[width=9cm]{final/2bii_plot} 
+                        \end{center}
+                        Comparing the $1$st-$10$th studentized residuals from the original model:\\ 
+                        \lstinputlisting[firstline=44,lastline=45]{final/2_out.txt} 
+                        To the respective critical values: \\
+                        \lstinputlisting[firstline=47,lastline=50]{final/2_out.txt} 
+                        We see that they all fall outside of the critical values for all confidence levels, so we would reject them all as outliers for all confidence levels.
+This fits with the simpler Bonferroni Critical value we used before; since it is conservative, and each of these falls above it, it stands to reason a sharper test would reject them to.
+                \end{itemize}
+            \item[c)]
+                From what we've seen, these few extreme points clearly don't fit the assumptions for the simple linear model on the variables we have, which in general seems appropriate for the vast majority of the data, so in that sense they are certainly outliers to that model.
+However, they are not necessarily outliers due to mismeasurement or some other factor such that they should be thrown out as impossible to include in a linear model.
+For instance, it may be the case that they represent periods during which their was a mechanical breakdown of some sort.
+If this were true, we could include them in the model if we had information about mechanical failures at the power plant to include in the model.
+Then they may well fit the assumptions for a standard linear model.
+        \end{itemize}
+    \item[3.]
+        \begin{itemize}
+            \item[a)]
+                Of the shrinkage methods we have discussed, PLS and Lasso are particularly suitable for $n<p$, while PCR is still ok but not as good as the other two.
+PLS and Lasso are best because they attempt to pick a sparse representation of the predictor space which predicts the outcome well.
+PLS does this by selecting components of a orthonormal basis in order of their predictiveness, while LASSO, depending at which point it is cut off, adds the original predictors back in as they increase predictive power, keeping the coefficients for the rest at $0$.
+Both of these processes result in a smaller, predictive subspace, effectively reducing $p$.
+PCR also shrinks the predictor space, but since it selects the orthogonal components which represent the largest variation in the predictor space, rather than are the most predictive of the outcome, it will likely give a subspace of larger dimension.
+For instance, in an extreme case, if the largest $n$ primary components don't predict the outcome at all, PCR will fail.
+Ridge regression is the least suitable option, since it only rarely will out right eliminate a predictor, instead constraining the size of the $\beta$s over all predictors.
+            \item[b)]
+                I fit a PCR model to the data set, using "pcr" in R, and then selected the number of components from the Scree plot and 10 fold cross validation.
+I chose 10 fold cross validation for a few reasons.
+First, with a relatively small number of observations in the test data set, using a smaller $k$ would result in a data set that is small enough to have meaningfully different error properties (since reducing a data set by a set portion increases the prediction error the most when the data set is small).
+Larger $k$ on the other hand would increasingly look like LOO cross validation.
+Besides, that though, the choice is somewhat arbitrary, so I went with common wisdom on the subject.
+The Scree plot and plotted cross validation scores are below: \\
+                \includegraphics[width=9cm]{final/3b_pcr_scree} 
+                \includegraphics[width=9cm]{final/3b_pcr_cv} \\ 
+                The Scree plot vaguely seemed to have an elbow at 10, where the marginal fall off of the standard deviations for each principal component slowed down, so I tried a model with 10 components.
+On the test set, the predictions from this model gave a RMSE of $8.69$.
+Going by $10$ fold cross validation, including $64$ components resulted in the lowest predicted RMSE of $5.58$.
+The actual RMSE of the test data set was $5.65$ for this model.
+            \item[c)]
+                I fit a PLS model, using "plsr" in R. I plotted out the $10$ fold cross validation scores:
+                \begin{center}
+                    \includegraphics[width=9cm]{final/3c_pls_cv} 
+                \end{center}
+                The predicted RMSE was minimized by $23$ components, a much smaller number than with PCR, predicting a RMSE of $6.12$.
+In fact, it turned out the actual RMSE on the test data set was $5.15$.
+            \item[d)]
+                Fitting both of these models again, but this time fitting the number of components by LOO cross validation, we get these predictions for RMSE: \\
+                \includegraphics[width=9cm]{final/3d_pcr_loo} 
+                \includegraphics[width=9cm]{final/3d_pls_loo} \\ 
+                As you can see, they are very similar, though not quite identical, to the results from $10$ fold cross validation.
+For PCR, the minimum is now found at $68$ components, with a predicted RMSE of $5.28$; this results in an actual RMSE on the test data set of $5.64$.
+For PLS, the minimum is at $24$ components, predicting a $RMSE$ of $5.57$ compared to an actual RMSE on the test set of $5.26$.
+\par
+                Both methods result in pretty similar models with similar performance on the test data set; the PLS model chosen from LOO was a bit worse from the one chosen from $10$ fold, but the PCR model was a hair better.
+The differences are small enough to be regarded as mere noise though.
+The only really difference is that LOO predicted the actual RMSE for PLS far better than $10$ fold, though the models end up being almost identical.
+This is about what I expected; the error in fitting a model fit on $n-1$ data points should be more similar to one fit on $n$ data points than one fit on $\frac{9n}{10}$ data points.
+            \item[e)]
+                I fit a LASSO model using the "lars" command in R.
+In the same package, there is "cv.lars", which calculates cross validation scores.
+It subsets the data into $K$ parts and fits a full LASSO model on each of the subsets gotten by removing one of the parts from the data (complete from all coefficients set to 0 to the normal linear regression coefficients).
+Then, for a set of complexity parameters given by the user, calculates the RMSE of each of the models on the part it wasn't fit on at that parameter, and averages over these 10 RMSE to give the predicted out of sample RMSE.
+I made use of this, using $10$ folds, to choose the complexity parameter to use.
+ Below is the graph of the RMSE at different complexity parameters, and a small portion of the LASSO plot: \\   
+                \includegraphics[width=9cm]{final/3e_cv_plot}
+                \includegraphics[width=9cm]{final/3e_lasso_plot} \\ 
+                The best complexity parameter according to $10$ fold cross validation was \(.019=\frac{\|\beta\|_1}{\max\|\beta\|_1}\), with an estimated RMSE of $5.52$.
+This gave 38 variables with positive coefficients.
+On the test data set, this model's predictions had a RMSE of $5.33$.
+\par
+                Of these three methods, I like PLS the best.
+It seems better than PCR, since it gives only the components that are actually predictive, so includes far fewer.
+LASSO results in more predictors, but uses the actual original variables, so it certainly has its value too; I find orthogonal predictors appealing though, even if they are harder to interpret.
+            \item[f)]
+                The biggest advantage of K-fold CV is that, assuming there isn't a simple closed form for LOO CV like with linear smoothers, it is much quicker to run, since a model only has to be run $K$ times, instead of $n$ times, saving a lot of computation.
+Its estimate of the RMSE of the full model on the training set will be biased upwards though; models fit on less data will tend to have somewhat higher error, so it will have more upwards bias than LOO, which is an average fit on models with just one fewer data point rather than $\frac{n}{k}$ fewer.
+It compensates for this by having less variance than LOO.
+Since each model for LOO is fit on a nearly identical data set, the prediction errors will be highly correlated.
+On the other hand, each of the models for $K$ fold are for more distinct, making the prediction errors less correlated, making the resulting mean less noisy.
+
+        \end{itemize}
+\end{itemize}
 

### 2014-12-10 stat343/final.R

@@ -11,12 +11,364 @@ library('splines')
 library('rpart')
 library('randomForest')
 library('leaps')
+library('mgcv')
+options(width=200)
 
 prob1<-T
 prob2<-F
 prob3<-F
 
+res.qq<-function(model,file,title) {
+    if (length(residuals(model))<5000) {
+        p.val<-round(shapiro.test(residuals(model))$p.value,5)
+        sub.text<-paste("Shapiro Test P-Value:",round(p.val,4))
+    }
+    else sub.text<-NULL   
+    pdf(file)
+    qqnorm(residuals(model),ylab="Residual Quantiles",main=title,sub=sub.text)
+    qqline(residuals(model))    
+    dev.off()
+}
+
+stud.res.qq<-function(model,file,title) {
+    rstud<-rstudent(model)
+    max<-abs(rstud[which.max(abs(rstud))])
+    n<-length(rstud)
+    p<-length(coef(model))
+    bf.crit<-abs(qt(.05/(n*2),n-p))
+    pdf(file)
+    qqnorm(rstudent(model),ylab="Residual Quantiles",main=title,sub=paste("Bonferroni Critical Value:",round(bf.crit,4),"Most Extreme Studentized Residual:",round(max,4)))
+    qqline(rstudent(model))    
+    dev.off()
+}
+
+
+res.plot<-function(model,file,title) {
+    pdf(file)
+    n<-length(predict(model))
+    size<-max(.05,min(.5,5/sqrt(n)))
+    plot(residuals(model)~predict(model),ylab="Residuals",xlab="Predicted",main=title,cex=size)
+    plot(locfit(residuals(model)~lp(predict(model),nn=.2,deg=1)),add=T)
+    legend('topleft',"Local Linear Fit",lty=1)
+    dev.off()
+}
+
+var.res.plot<-function(model,data,stub,title,terms) {
+    n<-length(predict(model))
+    size<-max(.05,min(.5,5/sqrt(n)))
+    for (i in terms) {
+        name<-names(data)[i]
+        pdf(paste(stub,name,'.pdf',sep=""))
+        plot(residuals(model)~data[,i],ylab="Residuals",xlab=name,main=paste(title,", ",name,sep=""),cex=size)
[…]

### 2014-12-10 stat343/final.tex
Added in R code for question 2b

@@
                         \lstinputlisting[firstline=27,lastline=30]{final/2_out.txt} 
-                        We see that they all fall outside of the critical values for all confidence levels, so we would reject them all as outliers for all confidence levels.
+                        We see that they all fall outside of the critical values for all confidence levels, so we would reject them all as outliers for all confidence levels. \\
+                        {\bf My R Code:}
+                        \lstinputlisting[firstline=191,lastline=214,breaklines=true]{final.R}
 
@@
 This fits with the simpler Bonferroni Critical value we used before; since it is conservative, and each of these falls above it, it stands to reason a sharper test would reject them to.
+\\
+                        {\bf My R Code:}
+                        \lstinputlisting[firstline=231,lastline=247,breaklines=true]{final.R}
+
                 \end{itemize}

### 2014-12-10 stat343/final.tex
Done with final

--- 2014-12-10 stat343/final.tex
@@
 The normally distributed error also isn't quite true, but the difference is fairly minor.
-The first choice for transformation the Box-Cox series of transformations, but the 95% confidence interval for the maximum likelihood includes 1, suggesting that such a transformation isn't worth while.
+The first choice for transformation the Box-Cox series of transformations, but the $95\%$ confidence interval for the maximum likelihood includes 1, suggesting that such a transformation isn't worth while.
 Instead, what ended up working was to normalize all the variables so that their smallest value was $0$ and the largest $1$, and then using the square root of the variables, or the square root of one minus the variable, as predictors.
@@
                 \lstinputlisting[firstline=4,lastline=23]{final/1_out.txt}
-                Considering possible problems other than nonlinearity, I considered the QQ plot of the residuals and a plot of the square root of the absolute values of the residuals versus the predicted values: \\
+                Considering possible problems other than nonlinearity, I looked at a QQ plot of the residuals and a plot of the square root of the absolute values of the residuals versus the predicted values: \\
                 \includegraphics[width=9cm]{final/1a_qq} 
@@
                 In the QQ plot, we see that the distribution of the residuals has much somewhat thicker tails than a normal distribution, though stays pretty close otherwise.
-The Shapiro-Wilkes test, with a p-value of $.0003$ has us reject normality, but the predictions are still unbiased if the normality assumption doesn't hold, and the deviation is pretty small.
-The variance of the residuals doesn't look bad, but the kernel regression fit seems to show some meaningful variation over the range of predictions.
-Thus, I ran a new model where, after scaling the variables so they all go from $0$ to $1$, I transformed the predictors, taking the square root of one minus angle, chord length, and thickness, and the square root of velocity.
+The Shapiro-Wilkes test, with a p-value of $.0003$ has us reject normality, but the predictions are still unbiased if the normality assumption doesn't hold, and the deviation from normality is pretty small.
+The variation in the variance of the residuals doesn't look terrible, but the local linear regression fit seems to show some meaningful variation over the range of predictions.
+Thus, I ran a new model where, after I scaled the variables so they all go from $0$ to $1$, I transformed the predictors, taking the square root of one minus angle, chord length, and thickness, and the square root of velocity.
 This resulted in the following model: \\
@@
 \par
-                {\bf Work:} I checked for non-linearity in the model from part a by plotting the residuals by predicted value and a few of the variables: \\
+                {\bf Work:} I checked for non-linearity in the model from part a by plotting the residuals by predicted value and a few of the variables (the other two only have a few unique values and looked fine): \\
                 \includegraphics[width=9cm]{final/1b_res_plot_lm2} 
@@
                 \includegraphics[width=9cm]{final/1b_res_plot_int_thi} \\ 
-                While there is still some nonlinearity, except in a few spots with sparse data, the mean is consistently right around $0$.
+                While there is still some nonlinearity, the mean is consistently right around $0$, except in a few spots with sparse data.
 We do seem to have meaningful non-zero means at lower frequencies, and at the higher end of the predicted values, but these biases are small, and much better than with the previous model.
@@
                 Of these two, each has its own advantages.
-The model from part b has the advantage of simplicity, making it suitable for explanation.
+The model from part a has the advantage of simplicity,, while only mildly violating the linear model assumptions, making it suitable for explanation.
 The confidence intervals obviously won't be perfect, since the predictions are biased and the normality assumption doesn't hold.
@@
 This was chosen because it was the points at which $10$ fold cross validation estimated there would be the lowest error.
-This gave a tree with $1121$ splits for $1502$ data points, which seems like a lot, but when one calculates the adjusted $R^2$, the result is still an impressive $.999$.
+This gave a tree with $1121$ splits for $1502$ data points, which seems like a lot, but when one calculates the adjusted $R^2$, the result is still an impressive $.999$, having reduced the $RSS$ to $27.467$ from it original value of $71480.51$.
 Between this and the cross validation, we are reasonably assured that this is not an over fit.
@@
                 This is the summary from the linear model: \\ 
-                \lstinputlisting[firstline=5,lastline=21]{final/2_out.txt}
+                \lstinputlisting[firstline=3,lastline=21]{final/2_out.txt}
 
@@
 I have plotted a quantile plot of the studentized residuals below.
-The Bonferroni Critical value for the absolute value of studentized residuals an alpha of $.05$, which is the \(\frac{1-.05/2}{n}\) quantile of a t distribution with n-p degrees of freedom, is $4.5581$.
+The Bonferroni Critical value for the absolute value of studentized residuals at an alpha of $.05$, which is the \(\frac{1-.05/2}{n}\) quantile of a t distribution with $n-p$ degrees of freedom, is $4.5581$.
 By comparison, looking at the plot, we can see that there are a number of studentized residuals with an absolute value greater than $4.5581$, going as high as $9.5778$.
@@
 After this, the critical value for the $q$th largest value at a p-value of $alpha$ is the $1-\alpha$ percentile of the $q$th largest values from all the boots.
-In other words, it was chosen so it is larger than $q$th largest value for exactly $\alpha$ of the boots.
+In other words, it was chosen so it is larger than $q$th largest value from exactly $\alpha$ of the boots.
                 \begin{itemize}
                     \item[i.]
-                        Using the method described above, run for 2000 total boots, I generated critical values for the $1$st through $10$th largest Cook Distances.
-I have plotted them below for critical values less than $.1$.
+                        {\bf Summary:} Using the method described above, run for 2000 total boots, I generated critical values for the $1$st through $10$th largest Cook Distances.
+We reject the ten largest cook distances as outliers, since they fall above these values.
+\par
+                        {\bf Work:}I have plotted the critical values less than $.1$ below.
 Obviously the critical values should asymptotically approach infinity as they go to $0$.
@@
                         \begin{center}
-                            \includegraphics[width=9cm]{final/2bi_plot} 
+                            \includegraphics[width=12cm]{final/2bi_plot} 
                         \end{center}
@@
                         \lstinputlisting[firstline=24,lastline=25]{final/2_out.txt} 
-                        To the respective critical values: \\
+                        To the respective critical values at p-values of $.01$, $.05$, and $.1$: \\
                         \lstinputlisting[firstline=27,lastline=30]{final/2_out.txt} 
-                        We see that they all fall outside of the critical values for all confidence levels, so we would reject them all as outliers for all confidence levels.
+                        We see that they all fall outside of the critical values for each confidence level, so we would reject them all as outliers for each confidence level. \\
+                        {\bf My R Code:}
+                        \lstinputlisting[firstline=191,lastline=214,breaklines=true]{final.R}
 
                     \item[ii.]
-                        Using the method described above, once again for 2000 total boots, but this time for studentized residuals, I have once again plotted the critical values.
-The same issue for critical values close to 0 exists.
+                        {\bf Summary:} Using the method described above, once again for 2000 total boots, but this time for studentized residuals, I have once generated the critical values.
+We again reject all $10$ largest studentized residuals as outliers, since they fall outside these values.
+\par
+                        {\bf Work:} The critical values are plotted below. The same issue for critical values extremely close to 0 exists.
                         \begin{center}
-                            \includegraphics[width=9cm]{final/2bii_plot} 
+                            \includegraphics[width=12cm]{final/2bii_plot} 
                         \end{center}
-                        Comparing the $1$st-$10$th studentized residuals from the original model:\\ 
+                        Comparing the absolute value of $1$st-$10$th largest studentized residuals from the original model:\\ 
                         \lstinputlisting[firstline=44,lastline=45]{final/2_out.txt} 
-                        To the respective critical values: \\
+                        To the respective critical values at p-values of $.01$, $.05$, and $.1$: \\
                         \lstinputlisting[firstline=47,lastline=50]{final/2_out.txt} 
-                        We see that they all fall outside of the critical values for all confidence levels, so we would reject them all as outliers for all confidence levels.
-This fits with the simpler Bonferroni Critical value we used before; since it is conservative, and each of these falls above it, it stands to reason a sharper test would reject them to.
+                        We see that they all fall outside of the critical values for each confidence level, so we would reject them all as outliers for each confidence level.
+This fits with the simpler Bonferroni Critical value we used before; since it is conservative, and each of the studentized residuals falls above it, it stands to reason a sharper test would reject them to.
+\\
+                        {\bf My R Code:}
+                        \lstinputlisting[firstline=231,lastline=247,breaklines=true]{final.R}
+
                 \end{itemize}
             \item[c)]
-                From what we've seen, these few extreme points clearly don't fit the assumptions for the simple linear model on the variables we have, which in general seems appropriate for the vast majority of the data, so in that sense they are certainly outliers to that model.
-However, they are not necessarily outliers due to mismeasurement or some other factor such that they should be thrown out as impossible to include in a linear model.
+                From what we've seen, these few extreme points clearly don't fit the assumptions for the simple linear model on the variables we have, which in general seems appropriate for the vast majority of the data.
+In that sense they are certainly outliers to that model.
+However, they are not necessarily outliers due to mismeasurement or some other factor such that they should be thrown out as invalid or impossible to include in a linear model.
 For instance, it may be the case that they represent periods during which their was a mechanical breakdown of some sort.
 If this were true, we could include them in the model if we had information about mechanical failures at the power plant to include in the model.
-Then they may well fit the assumptions for a standard linear model.
+Then they might well fit the assumptions for a standard linear model.
         \end{itemize}
@@
 For instance, in an extreme case, if the largest $n$ primary components don't predict the outcome at all, PCR will fail.
-Ridge regression is the least suitable option, since it only rarely will out right eliminate a predictor, instead constraining the size of the $\beta$s over all predictors.
+Ridge regression is the least suitable option, since it only rarely will out right eliminate a predictor, instead constraining the size of $\|\beta\|_2$ over all predictors.
             \item[b)]
-                I fit a PCR model to the data set, using "pcr" in R, and then selected the number of components from the Scree plot and 10 fold cross validation.
+                I fit a PCR model to the permuted training set, using ``pcr" in R, and then selected the number of components from the Scree plot and 10 fold cross validation.
 I chose 10 fold cross validation for a few reasons.
@@
             \item[c)]
-                I fit a PLS model, using "plsr" in R. I plotted out the $10$ fold cross validation scores:
+                I fit a PLS model, using ``plsr" in R. I plotted out the $10$ fold cross validation scores:
                 \begin{center}
@@
 The differences are small enough to be regarded as mere noise though.
-The only really difference is that LOO predicted the actual RMSE for PLS far better than $10$ fold, though the models end up being almost identical.
+The only really distinction between the methods is that LOO predicted the actual RMSE for PLS far better than $10$ fold, though the models end up being almost identical.
 This is about what I expected; the error in fitting a model fit on $n-1$ data points should be more similar to one fit on $n$ data points than one fit on $\frac{9n}{10}$ data points.
             \item[e)]
-                I fit a LASSO model using the "lars" command in R.
-In the same package, there is "cv.lars", which calculates cross validation scores.
-It subsets the data into $K$ parts and fits a full LASSO model on each of the subsets gotten by removing one of the parts from the data (complete from all coefficients set to 0 to the normal linear regression coefficients).
-Then, for a set of complexity parameters given by the user, calculates the RMSE of each of the models on the part it wasn't fit on at that parameter, and averages over these 10 RMSE to give the predicted out of sample RMSE.
-I made use of this, using $10$ folds, to choose the complexity parameter to use.
+                I fit a LASSO model using the ``lars" command in R.
+In the same package, there is ``cv.lars", which calculates cross validation scores.
+It subsets the data into $K$ parts and fits a full LASSO model on each combination of $K-1$ of the $K$ subsets.
+Then, for a set of complexity parameters given by the user, calculates the RMSE of each model on the subset of the data it wasn't fit over.
+The average of the RMSE calculations from the $K$ models is then the predicted out of sample RMSE.
+I made use of this, using $10$ folds, to choose the complexity parameter.
  Below is the graph of the RMSE at different complexity parameters, and a small portion of the LASSO plot: \\   
@@
 It seems better than PCR, since it gives only the components that are actually predictive, so includes far fewer.
-LASSO results in more predictors, but uses the actual original variables, so it certainly has its value too; I find orthogonal predictors appealing though, even if they are harder to interpret.
+LASSO results in more predictors, but uses the actual original variables, so it certainly has its value too; I find the pleasant mathematically properties of orthogonal predictors appealing though, even if they are harder to interpret.
             \item[f)]
-                The biggest advantage of K-fold CV is that, assuming there isn't a simple closed form for LOO CV like with linear smoothers, it is much quicker to run, since a model only has to be run $K$ times, instead of $n$ times, saving a lot of computation.
-Its estimate of the RMSE of the full model on the training set will be biased upwards though; models fit on less data will tend to have somewhat higher error, so it will have more upwards bias than LOO, which is an average fit on models with just one fewer data point rather than $\frac{n}{k}$ fewer.
-It compensates for this by having less variance than LOO.
+                The biggest advantage of K-fold CV is that, assuming there isn't a simple closed form for LOO CV as with linear smoothers, it is much quicker to run; the  model only has to be run $K$ times, instead of $n$ times, saving a lot of computation.
+It also has the advantage of less variance than LOO.
 Since each model for LOO is fit on a nearly identical data set, the prediction errors will be highly correlated.
-On the other hand, each of the models for $K$ fold are for more distinct, making the prediction errors less correlated, making the resulting mean less noisy.
+On the other hand, each of the models for $K$ fold are for more distinct, making the prediction errors less correlated, and the resulting mean less noisy.
+LOO CV will produce a less biased estimate of RMSE though; models fit on less data will tend to have somewhat higher error.
+Thus, identical models fit on $n$ and $n-1$ data points, as with LOO CV, will have a more similar error distribution than identical models fit on $n$ and $n\frac{k-1}{k}$ data points, as with $K$-fold CV (unless $k=n$).
 

### 2014-12-10 stat343/final.R

@@ -188,36 +188,36 @@ if (prob2) {
     }
 
     ### b) bootstrap to determine whether q most extreme points are outliers
-    q<-10
-    b<-2000
-    sigma<-summary(cch.lm)$sigma
-    Yhat<-predict(cch.lm)
-    X<-as.matrix(cch[1:4])
+    q <- 10
+    b <- 2000
+    sigma <- summary(cch.lm)$sigma
+    Yhat <- predict(cch.lm)
+    X <- as.matrix(cch[1:4])
 
     # i. using cook distance
-    cook.boot <-function(sigma,Yhat,X,q) {
-        cooks<-sort(cooks.distance(lm(rnorm(length(Yhat),Yhat,sigma)~X)),decreasing=T)
+    cook.boot  <- function(sigma,Yhat,X,q) {
+        cooks <- sort(cooks.distance(lm(rnorm(length(Yhat),Yhat,sigma)~X)),decreasing=T)
         return(cooks[1:q])
     }
-    cook.mat<-t(replicate(b,cook.boot(sigma,Yhat,X,q)))
-    sort.cook.mat<-apply(cook.mat,2,sort)
-    x<-1-(1:b)/b
-    ymax<-sort.cook.mat[ceiling(.995*b),1]
+    cook.mat <- t(replicate(b,cook.boot(sigma,Yhat,X,q)))
+    sort.cook.mat <- apply(cook.mat,2,sort)
+    x <- 1-(1:b)/b
+    ymax <- sort.cook.mat[ceiling(.995*b),1]
     pdf('final/2bi_plot.pdf')
     matplot(y=sort.cook.mat,x=x,lty=rep(2:6,2),col=rep(1,10),type="l",xlim=c(0,.1),ylim=c(0,ymax),xlab="P-Value",ylab="Critical Value",main="Critical Value, q-th Largest Cook's Distance")
-    xseq<-seq(.01,.1,.01)
-    yseq<-rep(NULL,10)
+    xseq <- seq(.01,.1,.01)
+    yseq <- rep(NULL,10)
     for (i in 1:length(xseq)) {
-        yseq[i]<-sort.cook.mat[ceiling((1-xseq[i])*b),i]
+        yseq[i] <- sort.cook.mat[ceiling((1-xseq[i])*b),i]
     }
     text(x=xseq,y=yseq,pos=rep(4,10),labels=c("1st","2nd","3rd",paste(4:10,"th",sep="")),offset=-.6)
[…]

### 2015-01-08 stat343/midtermcheatsheet.tex
Add folder and description for winter 2015 course work

\documentclass[10pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\corr}{\mathrm{Corr}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
\begin{multicols}{2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Matrix Form Linear Model} 
{\bf Form:} \(Y=X\beta, X\in \R^{n\times p}, Y \in \R^{n}, e \in \R^{n}, \beta \in \R^{p} \)
{\bf Assumptions:} [A1] \(Y=X\beta + e\) [A2] \(\E(e\vert X)=0\) [A3] \(\var(e_i\vert X) = \sigma^2\) [A4] \(\cov(e_i,e_j \vert X) =0\) [A5] \(e\sim N(0,\sigma I_n)\) \\
{\bf Normal Equations:} 
\[RSS(\beta) = \|(Y-X\beta)\|^2, SXX=\|x-\bar x\|^2, SXY = \langle x-\bar x, y-\bar y \rangle \]
\[\hat\beta = (X^TX)^{-1}X^TY, H=X(X^TX)^{-1}X^T, \hat Y = X\hat\beta = H Y\]
\[\hat\beta(e) = \beta + (X^TX)^{-1}X^Te \]
{\bf Properties of H:} (i) \(\hat\epsilon = (I-H)Y\) (ii) $H$, $I-H$ symmetric, (iii) $H$, $I-H$ idempotent ($H^2=H$) (iv) $HX=X$ (v) \(\hat e \perp X\) (vi) $(I-H)X=0$ (vii) $(I-H)H=H(I-H)=0$ (viii) \(\forall a\in \R^n, Ha\perp (I-H)a\) (ix) H only has eigen values 0,1 because $Hx=x$ if $x$ in span $H$. \\
{\bf Variance Estimate:} \(\E(\|\hat e\|) = \E(\hat e^T(I-H)\hat e) = \E(tr(\hat e \hat e^T (I-H))) = n\sigma^2(n-p)\), so
\[\hat\sigma^2 = \frac{\hat e^T\hat e}{n-p} = \frac{RSS}{n-p} \]
{\bf Variance $\hat\beta$:}
\(\var(\hat\beta) = (X^TX)^{-1}X^T\var(Y)X(X^TX)^{-1} = \sigma^2(X^TX)^{-1}X^TX(X^TX)^{-1} = \sigma^2 (X^TX)^{-1} \) \\
{\bf Gauss Markov:} If $a^TY$ is an unbiased estimator of $c^T\beta$, then \(\var(c^T\hat\beta)\leq\var(a^TY)\). Proof: first note \(c^T\beta=\E(a^TY)=a^TX\beta \rightarrow c^T=a^TX\) Thus,
\begin{align*}
    \var(a^TY) - \var(c^T\hat\beta) &= \var(a^T(X\beta+e)) - \var(a^TX\beta) \\
                                    &= \var(a^Te) - \var(a^THY) \\
                                    &= a^T\var(e)a - \var(a^THX\beta + a^THe) \\
                                    &= \sigma^2\|a\|^2 - \var(a^THe) \\
                                    &= \sigma^2\|a\|^2 - Ha\var(e)a^TH \\
                                    &= \sigma^2\|a\|^2 - \sigma^2\|Ha\|^2 \\
\end{align*}
{\bf R-squared:} \(R^2 = 1-\frac{RSS}{SYY} = \corr(\hat y,y) \)

\subsection{Inference} 
{\bf ANOVA Table:}\\
\begin{tabular}{l | l l l l l}
        & df      & ss                       & ms       & F     \\
    \hline
    Reg & p       & $\sum(\hat Y - \bar Y)$  & SS/p                    & $F = \frac{SS/p}{\hat\sigma^2}$ \\
    Res & n-p     & RSS                      & $\hat\sigma^2= \frac{RSS}{n-p}$   &  
\end{tabular}

{\bf Distribution Estimators:} $\hat\beta$ and $\hat\sigma^2$ independent under least squares, \(\hat\beta \sim N_p(\beta,\sigma^2(X^TX)^{-1})\), and \(\frac{\hat\sigma^2}{\sigma^2}(n-p) \sim \chi_{n-p}^2.\). Distribution of $\hat\beta$ follows from it being a linear transformation of $Y$ and variance as said earlier. \\
Proof: Since $(I-H)$ symmetric, for $P$ orthogonal matrix of eigenvalues and $D$ matrix with eigenvalues on diagonal, \(I-H=PDP^T\). All eigenvalues are $0$ or $1$, so get
\[I-H=PDP^T=[P_1 P_2] \left[\begin{array}{cc} I_{n-p} & \mathbf{0}\\\mathbf{0} & \mathbf{0}\end{array} \right][P_1 P_2]^T = P_1 P_1^T\]
So, \(\var(P_1^T\hat e) = \E(P_1^T\hat e \hat e^T P_1) - \E(P_1^T \hat e)^2 = \sigma^2P_1^TP_1 = \sigma^2 I_{n-p}\). This gives us that \(\frac{1}{\sigma^2}\hat e^T \hat e = \frac{1}{\sigma^2}\hat e^T P_1 P_1^T \hat e \sim \chi_{n-p}^2\) \\
{\bf Distribution Standardized Estimators:} \(\hat\beta_i \sim t_{n-p}\). Proof: \(\var(\hat\beta_i) = \sigma^2(X^TX)^{-1}_{ii}\) so \(SE(\hat\beta_i)=\hat\sigma\sqrt{(X^TX)^{-1}}\). Thus 
\[\frac{\hat\beta_i - \beta_i}{SE(\hat\beta_i)} = \frac{\hat\beta_i - \beta_i}{\sqrt{\sigma^2(\hat\beta_i)}}\sqrt{\frac{\sigma^2}{\hat\sigma^2}} \sim N(0,1)\sqrt{\frac{n-p}{\chi_{n-p}^2}} \sim t_{n-p} \]
{\bf t-test:} \(2\mathrm{P}[t_{n-p}> \frac{\hat\beta_i - \beta_i}{SE(\hat\beta_i)}]\) \\
{\bf Prediction Interval:} 
\[\mathrm{P}\left(\hat Y_* \in (x_*^T\hat\beta \pm t_{n-p,\alpha/2}\hat\sigma\sqrt{x_*^T(X^TX)^{-1}x_*})\right)=1-\alpha\]
\[\mathrm{P}\left(Y_* \in (x_*^T\hat\beta \pm t_{n-p,\alpha/2}\hat\sigma\sqrt{1+x_*^T(X^TX)^{-1}x_*})\right)=1-\alpha\]
{\bf F-test:} If you have two models where one is a subset of the other (\(span(H_1)\subset span(H_2)\)), then if \(rank(H_1)=q, rank(H_2)=p\),
\[\frac{\frac{1}{p-q}(\|\hat e_1\|^2-\|\hat e_2\|^2)}{\frac{1}{n-p}\|\hat e_2\|^2} \sim F_{p-q,n-p}\]
This is a one sided test. Good for testing sets of parameters.
{\bf Joint Confidence Interval:} A \(1-\alpha\) confidence region for $\beta$ is
\[\frac{\frac{1}{p}(\hat\beta-\beta)^T(X^TX)(\hat\beta-\beta)}{\hat\sigma^2} \leq p\hat\sigma^2 f_{p,n-p,\alpha} \]
If $R\beta$ has rank $q$, a \(1-\alpha\) confidence region for $R\beta$ is
\[\frac{\frac{1}{p}(R\hat\beta-R\beta)^T(R(X^TX)^{-1}R^T)^{-1}(R\hat\beta-R\beta)}{\hat\sigma^2} \leq p\hat\sigma^2 f_{q,n-p,\alpha} \]

\subsection{Numerical Techniques} 
{\bf Condition Number:} This is something to do with the effect of a small change in $Y$ on $\beta$. With
\[cos(\theta)=\frac{\|\hat Y\|}{\|Y\|} = \frac{\|X\hat\beta\|}{\|Y\|} \]
\[\frac{\|\Delta\hat\beta\|}{\|\hat\beta\|} \leq cond(X)\frac{1}{cos(\theta)}\frac{\|\Delta Y\|}{\|Y\|} \]
{\bf Cholesky Factorization:} If $X$ has rank $n$, $X^TX$ has full rank, and has Cholesky factorization $LL^T$. Thus, $X^TX\hat\beta=X^TY$, which can be solved in stages $Lz=X^TY$ and then $L^T\hat\beta=z$. \\
{\bf QR Factorization:}$\exists Q\in\R^{n\times n}, R\in\R^{p\times p}$ where $Q$ is orthogonal and $R$ is upper triangular such that 
\[X=\left[\begin{array}{c} R\\ \mathbf{0}\end{array} \right] \]
so we get
\[Q^TX\hat\beta = \left[\begin{array}{c} R\\ \mathbf{0}\end{array} \right] \hat\beta \cong \left[\begin{array}{c} f\\ r \end{array} \right] = Q^TY \]
This gives us \(RSS = \|y-X\hat\beta\|^2 = \|Q^Ty-Q^TX\hat\beta\|^2 = \|f-R\hat\beta\|^2 + \|r\|^2 \), which is minimized by \(f=R\hat\beta\).

\subsection{Resampling} 
{\bf Permutation Sampling:} Test significance of set of predictors by shuffling them over outcomes and other predictors a number of times. If $F$ statistic original model higher than all but $\alpha$ of shuffles, significant. \\
{\bf Bootstrap:} Get confidence interval of statistic (possibly $\theta$) by drawing with replacement a number of times and calculating statistic.

\subsection{Designed Experiment}
{\bf Orthogonal Predictor:} If $X_1$, $X_2$ orthogonal, then
\begin{align*}
    \beta=(X^TX)^{-1}X^TY &= \left[\begin{array}{cc} X_1^TX_1 & 0 \\ 0 & X_2^TX_2 \end{array} \right]^{-1}X^TY  \\
                          &= \left[\begin{array}{c} (X_1^TX_1)^{-1}X_1^TY \\ (X_2^TX_2)^{-1}X_2^TY \end{array} \right] 
\end{align*}
Estimates don't change if $X_1$ or $X_w$ removed, both less dependent other non-orthogonal vars. \\
{\bf Randomization:} If $Z$ can't be included in regression, in an experiment, by randomly assigning it to observations, $\cov(X,Z)$ should be $0$, so effect $Z$ part of error. \\
{\bf Lurking Variable} If $Z$ correlated with $X$, then, 
\[\E(Y\vert x, z) = X\beta + \delta z\]
\[\E(Z\vert x) = X\gamma\]
so
\[\E(Y\vert x) = X(\beta+\gamma) \]

\subsection{Diagnostics}
{\bf Non-Constant Variance:} Can Regress $\vert \hat e \vert$ on $\hat Y$ if.
{\bf Transform:} Transform non-linear/non-constant residual data.
\[h(Y)=\log(Y+\delta), h(Y) = \sqrt(Y)\]
{\bf Not Normal:} QQplot, Shapiro-Wilk \\
{\bf Correlated Error:} Durbin-Watson, where $\rho$ autocorrelation:
\[d = \frac{\sum_{i=2}^{n} (\hat e_i - \hat e_{i-1})^2}{\sum_{i=1}^{n} \hat e_i^2} \sim 2(1-\rho) \]
{\bf Leverage:} \(h_i=H_ii=x_i^T(X^TX)^{-1}x_i\). How strongly effects model. \\
{\bf Outlier Test:} $\hat y_{(i)}$ excludes $i$th observation. 
\[t_i = \frac{y_i-\hat y_{(i)}}{\sqrt{\hat\sigma_{(i)}^2(x_i^T(X_{(i)}^TX_{(i)})^{-1}x_i + 1)}}\]
Where \(r_i = \frac{\hat e_i}{\hat \sigma\sqrt{1-h_i}} \) (studentized residuals), this gives us
\[t_i = r_i\sqrt{\frac{n-p-1}{n-p-r_i^2}} \sim t_{n-p-1}\]
Bonferroni Correction: reject only if  \(t_{n-p-1,\alpha/n}>t_i\). \\
{\bf Cook Statistic:} Indicates influential point, whose removal effects fit.
\[D_i = \frac{(\hat\beta - \hat\beta_{(i)})^T (X^TX)(\hat\beta - \hat\beta_{(i)})}{p\hat\sigma^2} = \frac{1}{p}r_i^2\frac{h_i}{1-h_i} \]
{\bf Partial Residual Plots} Fit models, where $X_{(i)}$ excludes column $i$,
\[Y=X_{(i)}\beta_{(i)}+q_i, X_i=X_{(i)}\gamma+s_i \]
Plot $q_i$ in terms $s_i$. Can see leverage of points on $\beta_i$

\subsection{Distributions} 
{\bf Normal Distribution:} \(\phi(x) = \frac{1}{\sigma\sqrt{2\pi}}e^{\frac{-(x-\mu)^2}{2\sigma^2}}\), joint normal vars independent iff \(\cov(Z_1,Z_2)=0\). \\
{\bf Multivariate Normal Distribution:} if \(X \sim N(\mu,\Sigma)\) and $\Sigma$ positive definite,
\[f_X(x) = \frac{1}{(2\pi)^{k/2}\vert\Sigma\Vert^{1/2}}exp\left(-\frac{1}{2}(x-mu)^T\Sigma^{-1}(x-\mu)\right) \]
{\bf Chi Square:} $k$ degree of freedom, then \(\sum_{i=1}^k Z_i^2 \sim \chi^2_k\) if $Z_i$s independent standard normals \\
{\bf Student's T:} \(t_\nu \sim Z\sqrt{\frac{\nu}{\chi_\nu^2}}\) if $Z$ standard normal independent of $\chi_\nu^2$. \\
{\bf F:} \(F_{d_1,d_2} \sim \frac{\chi_{d_1}^2/d_1}{\chi_{d_2}^2/d_2}\) if \(\chi_{d_1}^2\) and \(\chi_{d_2}^2\) independent.

\subsection{Linear Algebra}
{\bf Cauchy Scwarz:} \(\vert\langle x,y \rangle \vert \leq \|x\|\|y\|\). Equality iff linearly independent \\
{\bf Triangle Inequality:} \(\|x+y\|^\leq\|x\|+\|y\|\) \\
{\bf Rank:} Number linearly independent columns/rows. \(rk(A_{m\times n})\leq \min(m,n), rk(AB)\leq \min(rk(A),rk(B)), rk(A+B)\leq rk(A) + rk(B), Rk(AA^T=rk(A))\) \\
{\bf Orthogonal:} $A^TA=I$. Columns $A$ orthonormal basis $R^n$, rotate/reflect vector. \(\langle Ax,Ax\rangle = \langle x, x \rangle\). \\
{\bf Idempotent:} \(AA=A\). Projection matrix. If \(x\in span(A), Ax=x\). \\
{\bf Determinant:} \(\vert AB \vert = \vert A\vert \vert B \vert\), if $A$ orthogonal, \(\vert A\vert = \pm 1, \vert A^TBA\vert = \vert B\vert. \) \\
{\bf Trace:} Sum diagonal entries. \(tr(A)=tr(A^T), tr(A+B)=tr(A)+tr(B), tr(ABC)=tr(CAB)\), if $A$ idempotent, \(rk(A)=tr(A)\), if $A$ nonsingular, \(tr(A^{-1}BA)=tr(B)\). \\
{\bf Eigenvalues:} If $A$ idempotent, \(\lambda =1,0\). If Orthogonal $\lambda$ has modulus 1 (radius in complex plane less than 1). Symmetric matrix has real eigenvalues. \\
{\bf Positive (semi) definite:} \(x^TAx > (\geq) 0 \forall x\). $BB^T$ always positive semi definite. \\
{\bf Eigen Decomposition:} If $A$ symmetric, \(A=P^TDP\), where $P$ matrix eigenvectors and $D$ diagonal matrix of eigenvalues. If $AB$=$BA$ and symmetric, $B=P^TD_BP$ for same $P$. \\
{\bf Diagonally Dominant:} If each diagonal greatest entry in column, $A$ positive semi definite if $A$ symmetric and diagonals positive. Strictly diagonally dominant matrix nonsingular. \\
{\bf Cholesky Decomposition:} If $A$ symmetric and positive definite, \(\exists L\) unique lower diagonal with positive diagonal entries such that $A=LL^T$. If $A$ only positive semi-definite, $L$ may not be unique and may have $0$ diagonal entries.
{\bf Underdetermined Linear System:} Smallest norm solution to $Ax=y$ is \(x=A^T(AA^T)^{-1}y\).
\end{multicols}
\end{document}

### 2015-01-08 stat301/hw0.tex
Homework 0

--- 2014-12-10 stat343/final.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-STATS 343 Final \hfill Aaron Maurer
+STAT 301 Homework 0 \hfill Aaron Maurer
 \vspace{2mm}
@@
 \vspace{2mm}
-
 \begin{itemize}
@@
         \begin{itemize}
-            \item[a)]
-                {\bf Summary:} Fitting a simple linear model, I see that the constant variance assumptions don't hold, which would suggest transformation.
-The normally distributed error also isn't quite true, but the difference is fairly minor.
-The first choice for transformation the Box-Cox series of transformations, but the 95% confidence interval for the maximum likelihood includes 1, suggesting that such a transformation isn't worth while.
-Instead, what ended up working was to normalize all the variables so that their smallest value was $0$ and the largest $1$, and then using the square root of the variables, or the square root of one minus the variable, as predictors.
-This results in improvement, giving residuals with approximately constant variance.
-\par
-
-                {\bf Work:} I ran the simple linear model:
-                \lstinputlisting[firstline=4,lastline=23]{final/1_out.txt}
-                Considering possible problems other than nonlinearity, I considered the QQ plot of the residuals and a plot of the square root of the absolute values of the residuals versus the predicted values: \\
-                \includegraphics[width=9cm]{final/1a_qq} 
-                \includegraphics[width=9cm]{final/1a_sqrt_res_plot} \\ 
-                In the QQ plot, we see that the distribution of the residuals has much somewhat thicker tails than a normal distribution, though stays pretty close otherwise.
-The Shapiro-Wilkes test, with a p-value of $.0003$ has us reject normality, but the predictions are still unbiased if the normality assumption doesn't hold, and the deviation is pretty small.
-The variance of the residuals doesn't look bad, but the kernel regression fit seems to show some meaningful variation over the range of predictions.
-Thus, I ran a new model where, after scaling the variables so they all go from $0$ to $1$, I transformed the predictors, taking the square root of one minus angle, chord length, and thickness, and the square root of velocity.
-This resulted in the following model: \\
-                \lstinputlisting[firstline=40,lastline=60]{final/1_out.txt}
-                Checking the fit again, I see some mild improvement: \\
-                \includegraphics[width=9cm]{final/1a_qq_lm2} 
-                \includegraphics[width=9cm]{final/1a_sqrt_res_plot_lm2} \\ 
-                The Residuals are a tad closer to being normally distributed (the lower tail deviates a little less), but more importantly, as one can see from the linear fit, the variance appears to vary much less over the range of predictions, and looks close to being constant, except maybe at the end.
-            \item[b)]
-                {\bf Summary:} We seem to have some pretty meaningful nonlinearity in the model from part a.
-To improve upon it, I transform the outcome, pressure, by taking it square root, and replace the terms for frequency, angle, and chord length with a fourth degree polynomial of those three variables together.
-This reduced the non-linearity and vastly improved the overall fit, as judged by adjusted R-squared.
-\par
-                {\bf Work:} I checked for non-linearity in the model from part a by plotting the residuals by predicted value and a few of the variables: \\
-                \includegraphics[width=9cm]{final/1b_res_plot_lm2} 
-                \includegraphics[width=9cm]{final/1b_res_plot_tran_ang} \\ 
-                \includegraphics[width=9cm]{final/1b_res_plot_tran_freq}  
-                \includegraphics[width=9cm]{final/1b_res_plot_tran_thi} \\ 
-                As you can see, the local linear fits pretty strongly show that the residuals don't have a mean of $0$ over the range of predicted values nor range of frequency, even if it looks OK for angle and thickness.
-I couldn't find any way to remedy this without interactions; I tried a number of models which had every predictor as a factor and some kind of transformation of the outcome, since this would admit every possible transformation of each predictor, but every model still failed to have residuals with mean $0$ over the range of predicted values.
-Thus, I introduced a set of orthogonal polynomials of degree up to 4 in terms of frequency, angle, and chord length together, as well as took the square root of pressure.
-This is the resulting model summary: \\
-                \lstinputlisting[firstline=76,lastline=127]{final/1_out.txt}
-                There are a lot of variables added, but not that many considering the size of the data set, and we see a big overall improvement in fit, as shown by the adjusted $R^2$ increasing to $.813$ from $.477$.
-Looking at plots of the residuals, we also seem to have mostly mitigated the issues with nonlinearity: \\
-                \includegraphics[width=9cm]{final/1b_res_plot_int} 
-                \includegraphics[width=9cm]{final/1b_res_plot_int_ang} \\ 
-                \includegraphics[width=9cm]{final/1b_res_plot_int_freq}  
-                \includegraphics[width=9cm]{final/1b_res_plot_int_thi} \\ 
-                While there is still some nonlinearity, except in a few spots with sparse data, the mean is consistently right around $0$.
-We do seem to have meaningful non-zero means at lower frequencies, and at the higher end of the predicted values, but these biases are small, and much better than with the previous model.
-Double checking the variance, we seem to still have close to constant variance, with possibly a little less consistency than we had before: \\
-                \begin{center}
-                    \includegraphics[width=9cm]{final/1b_sqrt_res_plot_int} 
-                \end{center}
-            \item[c)]
-                Of these two, each has its own advantages.
-The model from part b has the advantage of simplicity, making it suitable for explanation.
-The confidence intervals obviously won't be perfect, since the predictions are biased and the normality assumption doesn't hold.
-However, the confidence level is high enough that we can still fairly safely say that the predicted positive or negative correlations between the predictors and outcome hold true, and the estimates of how much this is by serves a sufficient ballpark guess.
-On the other hand, the model from b does seem to fulfill the assumptions of a linear regression model mostly (except possibly normality) and has high $R^2$, so it is good for prediction, even if the confidence intervals of the prediction might be slightly off.
-It is just convoluted enough that it is hard to interpret what effect angle, frequency, and chord length have on pressure without making a visual, since there are no single coefficients on these predictors..
-            \item[d)]
-                Fitting a tree model, I pruned the tree at a complexity parameter of $3.31*10^{-6}$.
-Any split which didn't decrease total $R^2$ by at least that much was discarded.
-This was chosen because it was the points at which $10$ fold cross validation estimated there would be the lowest error.
-This gave a tree with $1121$ splits for $1502$ data points, which seems like a lot, but when one calculates the adjusted $R^2$, the result is still an impressive $.999$.
-Between this and the cross validation, we are reasonably assured that this is not an over fit.
-We can look at the distribution of the residuals:\\
-                \includegraphics[width=9cm]{final/1d_res_plot_tree} 
-                \includegraphics[width=9cm]{final/1d_qq_tree} \\
-                As we would expect, by construction, the residuals definitely have mean $0$, with a large number of them, in fact, being $0$.
-Additionally, the residuals are no where close to normally distributed, but this is to be expected with so many of them already at $0$.
-            \item[e)]
-                As we we have seen here, in data that doesn't naturally fit the assumptions necessary for linear regression, its hard to transform it to the point where all the assumptions hold well.
-One can add an arbitrary amount of flexibility to a linear model to make it work, through splines, polynomials, other spanning function bases, and the like, but this has to be done manually.
-By the point enough flexibility has been added, often there is no easy interpretation of coefficients.
-Here in lies the advantage of a tree model; it has arbitrary amounts of flexibility built in by default, making it appropriate and easier for prediction with this kind of messy data.
-If you happen to stumble on a data set that actually has a linear relationship between two variables though, a tree model will never predict quite as well nor be as interpretable as the linear regression.
-
+            \item[(a)] Since $\mu$ is the counting measure, \(\mu(A)=0\Rightarrow A=\emptyset \Rightarrow \nu(A)=0\), since $\nu$ is a measure.
+Thus, $\nu$ is absolutely continious with respect to $\mu$.
+Since $\nu$ is the counting measure, $f$ is the probability mass function \(f(n)=\frac{1}{2^n}\).
+            \item[(b)] \(\mu\left(\R\backslash(0,1)\right)=0\) but \(\nu\left(\R\backslash(0,1)\right)>0\), so \(\mu(A)=0\not\Rightarrow\nu(A)=0\) and $\nu$ is not absolutely continious with respect to $\mu$.
         \end{itemize}
-
     \item[2.]
-        \begin{itemize}
-            \item[a)]
-                This is the summary from the linear model: \\ 
-                \lstinputlisting[firstline=5,lastline=21]{final/2_out.txt}
-
-                \begin{itemize}
-                    \item[i.]
-                        This residuals are pretty close to normal for the most part, but is different enough that we can say with a high degree of certainty that weren't actually drawn from normal distribution.
-Looking at the QQ plot of the residuals, the quantiles of residuals line up pretty well with normal quantiles except at the lower tail (perfect correspondence is falling along the line).
-At the lower tail though, there is immense deviation from the normal quantiles.
-With such a huge data set, such a large deviation for so many points is extremely unlikely.
-The end result is that we can expect the estimated confidence intervals (from the normal assumption) will be somewhat off.
-                        \begin{center}
-                            \includegraphics[width=9cm]{final/2ai_qq} 
-                        \end{center}
-
-                    \item[ii.]
-                        There are unquestionably pretty significant outliers.
-We can see this by looking at the Cook distance and the studentized residuals.
-I have plotted a quantile plot of the studentized residuals below.
-The Bonferroni Critical value for the absolute value of studentized residuals an alpha of $.05$, which is the \(\frac{1-.05/2}{n}\) quantile of a t distribution with n-p degrees of freedom, is $4.5581$.
-By comparison, looking at the plot, we can see that there are a number of studentized residuals with an absolute value greater than $4.5581$, going as high as $9.5778$.
-Since this critical value is conservative, we definitely have a number of outliers.
-                        \begin{center}
-                            \includegraphics[width=9cm]{final/2aii_stud_qq} 
-                        \end{center}
-                        Looking at this another way, we consider the Cook distances, plotted against the half-normal quantiles.
-We certainly seem to have some Cook distances which stand out; this further supports the notion we're looking at outliers.
-                        \begin{center}
-                            \includegraphics[width=9cm]{final/2aii_cook} 
-                        \end{center}
-                    \item[iii.]
-                        The linearity assumption seems to hold up for the most part.
-The first check is that the residuals really do have mean $0$ throughout the range of predictions.
-Plotting the residuals against the predicted values and fitting a flexible kernel regression model, we do seem to have very close to mean $0$, except at the lowest end of the predicted range; the local regression is very close to $0$ everywhere else.
-                        \begin{center}
-                            \includegraphics[width=9cm]{final/2aiii_res_plot} 
-                        \end{center}
-                        Additionally, checking the partial residual plot of each variable in turn, all of them the linear fits seem to be extremely good.
-If there isn't a linear relationship between one of the variables and the outcome, the difference isn't much.
-\\
-                        \includegraphics[width=9cm]{final/2aiii_pr_AP} 
-                        \includegraphics[width=9cm]{final/2aiii_pr_AT} \\ 
-                        \includegraphics[width=9cm]{final/2aiii_pr_RH} 
-                        \includegraphics[width=9cm]{final/2aiii_pr_V} \\ 
-                \end{itemize}
-            \item[b)]
-                For both of these statistics, I used a parametric bootstrap methodology to come up with critical values.
-For each boot, I drew a new set of outcomes $Y^*$, each one drawn from a normal distribution whose mean was the prediction from the original linear model for the given observation, and whose variance was the estimated variance from the original linear model.
-Using $Y^*$ and the original $X$, I fit a new linear model, and then calculated the statistic (Cook Distance or absolute studetnized residual) from this model for all the observations, keeping the $q$ largest values.
-After this, the critical value for the $q$th largest value at a p-value of $alpha$ is the $1-\alpha$ percentile of the $q$th largest values from all the boots.
-In other words, it was chosen so it is larger than $q$th largest value for exactly $\alpha$ of the boots.
-                \begin{itemize}
-                    \item[i.]
-                        Using the method described above, run for 2000 total boots, I generated critical values for the $1$st through $10$th largest Cook Distances.
-I have plotted them below for critical values less than $.1$.
-Obviously the critical values should asymptotically approach infinity as they go to $0$.
-This is only a finite simulation though, so little should be read into the estimates extremely close to $0$.
-                        \begin{center}
-                            \includegraphics[width=9cm]{final/2bi_plot} 
-                        \end{center}
-                        Comparing the $1$st-$10$th largest Cook Distances from the original model:\\ 
-                        \lstinputlisting[firstline=24,lastline=25]{final/2_out.txt} 
-                        To the respective critical values: \\
-                        \lstinputlisting[firstline=27,lastline=30]{final/2_out.txt} 
-                        We see that they all fall outside of the critical values for all confidence levels, so we would reject them all as outliers for all confidence levels.
-
-                    \item[ii.]
-                        Using the method described above, once again for 2000 total boots, but this time for studentized residuals, I have once again plotted the critical values.
-The same issue for critical values close to 0 exists.
-                        \begin{center}
-                            \includegraphics[width=9cm]{final/2bii_plot} 
-                        \end{center}
-                        Comparing the $1$st-$10$th studentized residuals from the original model:\\ 
-                        \lstinputlisting[firstline=44,lastline=45]{final/2_out.txt} 
-                        To the respective critical values: \\
-                        \lstinputlisting[firstline=47,lastline=50]{final/2_out.txt} 
-                        We see that they all fall outside of the critical values for all confidence levels, so we would reject them all as outliers for all confidence levels.
-This fits with the simpler Bonferroni Critical value we used before; since it is conservative, and each of these falls above it, it stands to reason a sharper test would reject them to.
-                \end{itemize}
-            \item[c)]
-                From what we've seen, these few extreme points clearly don't fit the assumptions for the simple linear model on the variables we have, which in general seems appropriate for the vast majority of the data, so in that sense they are certainly outliers to that model.
-However, they are not necessarily outliers due to mismeasurement or some other factor such that they should be thrown out as impossible to include in a linear model.
-For instance, it may be the case that they represent periods during which their was a mechanical breakdown of some sort.
-If this were true, we could include them in the model if we had information about mechanical failures at the power plant to include in the model.
-Then they may well fit the assumptions for a standard linear model.
-        \end{itemize}
-    \item[3.]
-        \begin{itemize}
-            \item[a)]
-                Of the shrinkage methods we have discussed, PLS and Lasso are particularly suitable for $n<p$, while PCR is still ok but not as good as the other two.
-PLS and Lasso are best because they attempt to pick a sparse representation of the predictor space which predicts the outcome well.
-PLS does this by selecting components of a orthonormal basis in order of their predictiveness, while LASSO, depending at which point it is cut off, adds the original predictors back in as they increase predictive power, keeping the coefficients for the rest at $0$.
-Both of these processes result in a smaller, predictive subspace, effectively reducing $p$.
-PCR also shrinks the predictor space, but since it selects the orthogonal components which represent the largest variation in the predictor space, rather than are the most predictive of the outcome, it will likely give a subspace of larger dimension.
-For instance, in an extreme case, if the largest $n$ primary components don't predict the outcome at all, PCR will fail.
-Ridge regression is the least suitable option, since it only rarely will out right eliminate a predictor, instead constraining the size of the $\beta$s over all predictors.
-            \item[b)]
-                I fit a PCR model to the data set, using "pcr" in R, and then selected the number of components from the Scree plot and 10 fold cross validation.
-I chose 10 fold cross validation for a few reasons.
-First, with a relatively small number of observations in the test data set, using a smaller $k$ would result in a data set that is small enough to have meaningfully different error properties (since reducing a data set by a set portion increases the prediction error the most when the data set is small).
-Larger $k$ on the other hand would increasingly look like LOO cross validation.
-Besides, that though, the choice is somewhat arbitrary, so I went with common wisdom on the subject.
-The Scree plot and plotted cross validation scores are below: \\
-                \includegraphics[width=9cm]{final/3b_pcr_scree} 
-                \includegraphics[width=9cm]{final/3b_pcr_cv} \\ 
-                The Scree plot vaguely seemed to have an elbow at 10, where the marginal fall off of the standard deviations for each principal component slowed down, so I tried a model with 10 components.
-On the test set, the predictions from this model gave a RMSE of $8.69$.
-Going by $10$ fold cross validation, including $64$ components resulted in the lowest predicted RMSE of $5.58$.
-The actual RMSE of the test data set was $5.65$ for this model.
-            \item[c)]
-                I fit a PLS model, using "plsr" in R. I plotted out the $10$ fold cross validation scores:
-                \begin{center}
-                    \includegraphics[width=9cm]{final/3c_pls_cv} 
-                \end{center}
-                The predicted RMSE was minimized by $23$ components, a much smaller number than with PCR, predicting a RMSE of $6.12$.
-In fact, it turned out the actual RMSE on the test data set was $5.15$.
-            \item[d)]
-                Fitting both of these models again, but this time fitting the number of components by LOO cross validation, we get these predictions for RMSE: \\
-                \includegraphics[width=9cm]{final/3d_pcr_loo} 
-                \includegraphics[width=9cm]{final/3d_pls_loo} \\ 
-                As you can see, they are very similar, though not quite identical, to the results from $10$ fold cross validation.
-For PCR, the minimum is now found at $68$ components, with a predicted RMSE of $5.28$; this results in an actual RMSE on the test data set of $5.64$.
-For PLS, the minimum is at $24$ components, predicting a $RMSE$ of $5.57$ compared to an actual RMSE on the test set of $5.26$.
-\par
-                Both methods result in pretty similar models with similar performance on the test data set; the PLS model chosen from LOO was a bit worse from the one chosen from $10$ fold, but the PCR model was a hair better.
-The differences are small enough to be regarded as mere noise though.
-The only really difference is that LOO predicted the actual RMSE for PLS far better than $10$ fold, though the models end up being almost identical.
-This is about what I expected; the error in fitting a model fit on $n-1$ data points should be more similar to one fit on $n$ data points than one fit on $\frac{9n}{10}$ data points.
-            \item[e)]
-                I fit a LASSO model using the "lars" command in R.
-In the same package, there is "cv.lars", which calculates cross validation scores.
-It subsets the data into $K$ parts and fits a full LASSO model on each of the subsets gotten by removing one of the parts from the data (complete from all coefficients set to 0 to the normal linear regression coefficients).
-Then, for a set of complexity parameters given by the user, calculates the RMSE of each of the models on the part it wasn't fit on at that parameter, and averages over these 10 RMSE to give the predicted out of sample RMSE.
-I made use of this, using $10$ folds, to choose the complexity parameter to use.
- Below is the graph of the RMSE at different complexity parameters, and a small portion of the LASSO plot: \\   
-                \includegraphics[width=9cm]{final/3e_cv_plot}
-                \includegraphics[width=9cm]{final/3e_lasso_plot} \\ 
-                The best complexity parameter according to $10$ fold cross validation was \(.019=\frac{\|\beta\|_1}{\max\|\beta\|_1}\), with an estimated RMSE of $5.52$.
-This gave 38 variables with positive coefficients.
-On the test data set, this model's predictions had a RMSE of $5.33$.
-\par
-                Of these three methods, I like PLS the best.
-It seems better than PCR, since it gives only the components that are actually predictive, so includes far fewer.
-LASSO results in more predictors, but uses the actual original variables, so it certainly has its value too; I find orthogonal predictors appealing though, even if they are harder to interpret.
-            \item[f)]
-                The biggest advantage of K-fold CV is that, assuming there isn't a simple closed form for LOO CV like with linear smoothers, it is much quicker to run, since a model only has to be run $K$ times, instead of $n$ times, saving a lot of computation.
-Its estimate of the RMSE of the full model on the training set will be biased upwards though; models fit on less data will tend to have somewhat higher error, so it will have more upwards bias than LOO, which is an average fit on models with just one fewer data point rather than $\frac{n}{k}$ fewer.
-It compensates for this by having less variance than LOO.
-Since each model for LOO is fit on a nearly identical data set, the prediction errors will be highly correlated.
-On the other hand, each of the models for $K$ fold are for more distinct, making the prediction errors less correlated, making the resulting mean less noisy.
-
-        \end{itemize}
+        Let \(S_\mu\) and \(S_\nu\) be the supports of $\mu$ and $\nu$.
+If the two measures are mutually singular, \(S_\mu\cap S_\nu=0\) and \(S_\mu,S_\nu\subseteq \Omega\), so \(S_\mu\subseteq \Omega\backslash S_\nu\).
+Thus, \(0\leq\nu(S_\mu)\leq\nu(\Omega\backslash S_\nu)=0\), so $\mu$ is not absolutely continious with respect to $\mu$.
+By an identical argument, $\nu$ is not absolutley continious with respect to $\nu$.
 \end{itemize}

### 2015-01-29 stat345/hw2.tex
Homework 2 finished

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 345 Homework 2 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
{\bf Problem 1}

\begin{itemize}
    \item[a.]
        \begin{align*}
            \E[SS_A] &= \E\left[rb\sum_{j=1}^a (y_{.j.}-y_{...})^2]\right] \\
            \E[SS_A] &= rb\sum_{j=1}^a \E[((t_{j.}+e_{.j.})-(t_{..}+e_{...}))^2] \\
            \E[SS_A] &= rb\sum_{j=1}^a \E[(p_j+e_{.j.}-e_{...})^2] \\
            \E[SS_A] &= rb\sum_{j=1}^a \left( \E[p_j^2] + 2\E[p_je_{.j.}]-2\E[p_je_{...}] + \E[e_{.j.}^2]- 2\E[e_{.j.}e_{...}]+ \E[e_{...}^2] \right) \\
            \E[SS_A] &= rb\sum_{j=1}^a \left(p_j^2 + \E\left[\left(\frac{\sum_{i,k}e_{ijk}}{rb}\right)^2\right]- 2\E\left[\frac{\sum_{i,k}e_{ijk}}{rb}\frac{\sum_{i,j,k}e_{ijk}}{rab}\right]+ \E\left[\left(\frac{\sum_{i,j,k}e_{ijk}}{rab}\right)^2\right] \right) \\
            \E[SS_A] &= rb\sum_{j=1}^a \left(p_j^2 + \frac{\sigma^2}{rb} - \frac{2\sigma^2}{rab} + \frac{\sigma^2}{rab}\right) \\
            \E[SS_A] &= (a-1)\sigma^2 + rb\sum_{j=1}^a p_j^2\\
        \end{align*}
    \item[b.]
        \begin{align*}
            \E[SS_{A\times B}] &= \E\left[r\sum_{j,k} (y_{.jk}-y_{.j.}-y_{..k}+y_{...})^2]\right] \\
            \E[SS_{A\times B}] &= r\sum_{j,k} \E[((pq)_{jk}+e_{.jk}-e_{.j.}-e_{..k}+e_{...})^2] \\
            \E[SS_{A\times B}] &= r\sum_{j,k} \Big[(pq)_{jk}+\E[e_{.jk}^2]+\E[e_{.j.}^2]+\E[e_{..k}^2]+\E[e_{...}^2]-2\E[e_{.jk}e_{.j.}]-2\E[e_{.jk}e_{..k}]+2\E[e_{.jk}e_{...}] \\
                               & \hskip 6em +2\E[e_{.j.}e_{..k}]-2\E[e_{.j.}e_{...}] -2\E[e_{..k}e_{...}] \Big] \\
            \E[SS_{A\times B}] &= r\sum_{j,k} \left[(pq)_{jk} + \frac{\sigma^2}{r} + \frac{\sigma^2}{ra} + \frac{\sigma^2}{rb} + \frac{\sigma^2}{rab} - \frac{2\sigma^2}{rb} - \frac{2\sigma^2}{ra} + \frac{2\sigma^2}{rab} + \frac{2\sigma^2}{rab} - \frac{2\sigma^2}{rab} -  \frac{2\sigma^2}{rab} \right] \\  
            \E[SS_{A\times B}] &= r\sum_{j,k} \left[(pq)_{jk} + \frac{\sigma^2}{r} - \frac{\sigma^2}{ra} - \frac{\sigma^2}{rb} + \frac{\sigma^2}{rab}\right] \\  
            \E[SS_{A\times B}] &= ab\sigma^2 -b\sigma^2 -a\sigma^2 +\sigma^2 r\sum_{j,k} (pq)_{jk} \\  
            \E[SS_{A\times B}] &= (a-1)(b-1)b\sigma^2 + r\sum_{j,k} (pq)_{jk} \\  
        \end{align*}
\end{itemize}
\newpage
\null
\newpage
STAT 345 Homework 2 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

{\bf Problem 2}
\begin{itemize}
    \item[a.]
        To achieve equal variance among the estimated treatment differences, I would have one treatment of each kind in the blocks of four (O,A,B,C), one of each A,B,C in the both blocks of three, and O with each of the other treatments respectively for the final three blocks. In other words, the set of blocks would be:
        \[ \{\{O,A,B,C\},\{O,A,B,C\},\{A,B,C\},\{A,B,C\},\{O,A\},\{O,B\},\{O,C\}\} \]
        This would result in the covariance matrix:
            \FloatBarrier
            \input{hw2/hw2_2a_cov} 
            \FloatBarrier
        And give the following unscaled variances among the estimated differences:
            \FloatBarrier
            \input{hw2/hw2_2a_dif_var} 
            \FloatBarrier
    \item[b.]
        To get the variance estimates for the differences between 0 and the others to be $\frac{2}{3}$ that of the other differences, I used the following block assignment:
        \[ \{\{O,A,B,C\},\{O,O,A,B\},\{O,A,C\},\{O,B,C\},\{O,A\},\{O,B\},\{O,C\}\} \]
        This would result in the covariance matrix:
            \FloatBarrier
            \input{hw2/hw2_2b_cov} 
            \FloatBarrier
        \newpage
        And give the following unscaled variances among the estimated differences:
            \FloatBarrier
            \input{hw2/hw2_2b_dif_var} 
            \FloatBarrier
\end{itemize}
\newpage
STAT 345 Homework 2 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

{\bf Problem 3}
\begin{itemize}
    \item[a.]
        Using the standard model, variety doesn't seem to have much a significant impact on yield. I ran a linear regression with the variety and block as predictors (I used a sum constraint on variety and a treatment constraint on block). I have plotted the coefficients for variety and their confidence intervals. Since I used the sum constraint, $0$ represents the average of all varieties.
            \begin{center}
                \includegraphics[width=16cm]{hw2/hw2_3a_coef_plot}
            \end{center}
        As you can see, most of the confidence intervals overlap $0$, indicating they are not statistically different from the mean across all varieties. One barely doesn't overlap, but the confidence intervals don't account for multiple testing, so a small number of "significant" results are to be expected. Thus, this appears to be a null plot, where variety doesn't actually predict yield significantly. To test this, I generated an ANOVA table:
            \FloatBarrier
            \input{hw2/hw2_3a_anova} 
            \FloatBarrier
        Indeed, the F test on Variety yields a high p-value of $.712$. We can not reject the null hypothesis that different varieties have the same expected yield.
    \item[b.]
        I calculated this estimate as a contrast on the estimated parameters. Where $\beta_i$ was the estimated effect for variety $i$, 
        \[ a_i = \begin{cases} \frac{1}{20} &\mbox{if } i\leq20 \\ 
                               \frac{1}{36} &\mbox{if } i>20 \end{cases} \]
        Thus, my estimate was \(d=\sum_{i=1}^{56} a_i \beta_i=-8.813\). Then, my confidence interval was gotten in the usual way for a linear combination of parameters from a linear regression:
        \[ d\pm t_{165}^{.025} \sqrt{a^t \Sigma a} = [-47.586,29.961] \]
    \item[c.]
        I have plotted the residuals below:
            \begin{center}
                \includegraphics[width=16cm]{hw2/hw2_3c_plot}
            \end{center}
            It appears that the units in the top left corner (high latitude, low longitude), tend not to have mean 0, instead having negative mean. This indicates that the blocking didn't really account for the variation in yield due to geography as the study assumed. Any of the varieties with one plot in that region would be adversely affected, but those that have a plot in both block 3 and 4 which fall in that region would be the worst affected, driving down their predicted yield. This seems to include varieties 30, 26, 52, 4, and possibly a few others.
    \item[d.]
        After including longitude and latitude as predictors in the model, we see some movement in the predicted effects of the varieties:
            \begin{center}
                \includegraphics[width=16cm]{hw2/hw2_3d_coef_plot}
            \end{center}
        From the previous regression, most of the predicted effects stay approximately the same. Some, such as 1, flip signs, though their CI still cover $0$. A few, such as 55 and 53 are nudged over enough so that their CI don't cover 0, but with so many tests that distinction isn't meaningful. The only meaningful change I see is that 30 made a large shift, becoming the most significant effect while before the estimate was very close to 0. Looking at the residuals again, we still see somewhat of the same problem we had before, though less substantial:
            \begin{center}
                \includegraphics[width=16cm]{hw2/hw2_3d_plot}
            \end{center}
        There still remains what looks like a meaningful dip in yield in that top left corner.
    \item[e.]
        With the quadratic function of the coordinates, we finally see substantial differentiation between the varieties:
            \begin{center}
                \includegraphics[width=16cm]{hw2/hw2_3e_coef_plot}
            \end{center}
        Variety 30 appears to be quite distinct from the rest of the varieties; it's p-value is .000368, which is sufficiently small that we can reject the null assumption at an alpha of $5\%$ even with a Bonferoni adjustment.  A few other varieties' CIs now no longer cover 0, such as 39, 40, and 20, but they aren't by very much. Since I am using the sum constraint, this may be a function of variety 30 dragging up the average rather than than these varieties being meaningfully worse than all the varieties excluding 30. Looking at the residuals again, we seem to have finally solved the issue with the top left corner:
            \begin{center}
                \includegraphics[width=16cm]{hw2/hw2_3e_plot}
            \end{center}
            By eyeball, I can not say that that plots in that region do worse than the rest of the plots, and no other area jumps out as doing particularly well or poorly. If there is a lingering geographic effect, it does not seem to be substantial.
    \item[f.]
        I do not think this study design was well chosen. The intent was clearly that the blocks would account for heterogeneity in yield over different areas of the farm, but this did not hold true in practice. Additionally, the blocks allowed replicates of the same variety to be planted relatively closely in the field. For instance, two of variety 30's replicates were right next to each other in the top right corner, and all of 30's replicates were in the four left most columns of plots. A better solution would have been to create smaller blocks which had much closer to homogeneous growing conditions, and then assign varieties to blocks such that they are more evenly spread over the field. For instance, the field might have been broken up into approximately a $4\times4$ grid, where each replicate of a particular variety would be placed in a block such that none of the other replicates of that variety would be in the same row or column of blocks. This would have prevented them being overly clustered together. The cost of this would be that the experiment would no longer have a full factorial design, but with $14$ treatments in each block there would still be substantial grounds to compare varieties.

\end{itemize}

\end{document}

### 2015-02-26 stat345/hw4.tex
Homework 4

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

STAT 345 Homework 4 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
{\bf Problem 1}
\begin{itemize}
    \item[a)]
        Let $Y$ be the vector of $y_{ij}$, and our eventual predictor to be $c'Y$ for $c\in\R^{ab}$. For our predictor to be unbiased, it must be the case that
        \begin{align*}
            0 &= \E(c'Y-\mu-\alpha_1) \\
            0 &= c'\E(Y) - \mu \\
            0 &= \mu c'1 - \mu \\
            1 &= c'1
        \end{align*}
        Furthemore, to make our LUP a BLUP, we need to minimize the variance around our estimate. Thus, making use of our constraint:
        \begin{align*}
            \var(c'Y-\mu-\alpha_1) &= \var\left(\sum_{i,j} c_{ij}(\mu + \alpha_i + \beta_j + e_{ij}) - \mu - \alpha_1\right) \\
                                   &= \var\left(\sum_{i,j} c_{ij}(      \alpha_i + \beta_j + e_{ij}) -     \alpha_1\right) \\
                                   &= \var\left(\sum_{i,j} c_{ij}(\alpha_i + \beta_j + e_{ij})\right) -2\cov\left(\alpha_1,\sum_{i,j} c_{ij}(\alpha_i + \beta_j + e_{ij})\right) + \var(\alpha_1)  \\
                                   &= \var\left(\sum_{i,j} c_{ij}(\alpha_i + \beta_j + e_{ij})\right) -2 \sigma_a^2 \sum_{j} c_{1j} + \sigma_a^2 
        \end{align*}
        If we let \(\Sigma=\cov(Y)=\sigma_e^2I + \sigma_a^2 V_a + \sigma_b^2 V_b\)  and \(I_{\{i=1\}}\) be a vector who has entry $0$ for $ij$ where $i\neq1$ and $1$ when $i=1$, we get

            \begin{align*}
                \var(c'Y-\mu-\alpha_1) &= c'\Sigma c -2\sigma_a^2 c'I_{\{i=1\}} + \sigma_a^2 \\
                \nabla \var(c'Y-\mu-\alpha_1) &= 2\Sigma c -2\sigma_a^2 I_{\{i=1\}}
            \end{align*}
            Which is $0$ when (we know $\sigma$ is invertible due to its block structure):
            \[ c = \sigma_a^2\Sigma^{-1} I_{\{i=1\}} \]
            Making this $c'Y$ the BLUP.
    \item[a)]
        Repeating a similar process, we have a constraint of 
        \begin{align*}
            0 &= \E(c'Y-\mu-\alpha_1 -\beta_1) \\
            0 &= c'\E(Y) - \mu \\
            0 &= \mu c'1 - \mu \\
            1 &= c'1
        \end{align*}
        And find that our estimates variance is
        \begin{align*}
            \var(c'Y-\mu-\alpha_1) &= \var\left(\sum_{i,j} c_{ij}(\mu + \alpha_i + \beta_j + e_{ij}) - \mu - \alpha_1 - \beta_1\right) \\
                                   &= \var\left(\sum_{i,j} c_{ij}(      \alpha_i + \beta_j + e_{ij}) -     \alpha_1 - \beta_1\right) \\
                                   &= \var\left(\sum_{i,j} c_{ij}(\alpha_i + \beta_j + e_{ij})\right) -2\cov\left(\alpha_1+\beta_1,\sum_{i,j} c_{ij}(\alpha_i + \beta_j + e_{ij})\right) + \var(\alpha_1 + \beta_1)  \\
                                   &= \var\left(\sum_{i,j} c_{ij}(\alpha_i + \beta_j + e_{ij})\right) -2 \sigma_a^2 \sum_{j} c_{1j} -2\sigma_b^2 \sum_{i} c_{1j}+ \sigma_a^2 + \sigma_b^2 \\
                                   &= c'\Sigma c -2\sigma_a^2 c' I_{\{i=1\}} -2\sigma_b^2 c' I_{\{j=1\}} + \sigma_a^2 + \sigma_b^2 \\
            \nabla \var(c'Y-\mu-\alpha_1-\beta_1) &= 2\Sigma c -2\sigma_a^2 I_{\{i=1\}} -2\sigma_b^2 I_{\{j=1\}}
        \end{align*}
        Which is $0$ when
        \[ c = \Sigma^{-1}(\sigma_a^2 I_{\{i=1\}} + 2\sigma_b^2  I_{\{j=1\}}) \]
        Making this $c'Y$ the BLUP
\end{itemize}

\newpage
STAT 345 Homework 4 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
{\bf Problem 2}
\begin{itemize}
    \item[a)]
        As we can see from this plot, it appears that the yield of all strains is increased by heavy manure. In addition, yield seems to vary with plot, and it seems that there is an interaction between high yield and strain.
        \begin{center}
            \includegraphics[width=16cm]{hw4/hw4_2a_plot}
        \end{center}
        It would not be appropriate to have fixed effects for each of the three variables, since only the effect of the specific manure level and strain is of interest. On the other hand, no other land manager would care about the particular effect of these specific plots; their inclusion in a model is only so their effect can be factored out of the estimates for manure level and strain and the variance from it can be incorporated.
    \item[b)]
        I fit a model with strain and manure level, along with an interaction between the two, as fixed effects, and plot as a random effect. At eight parameters and four random effects, this is a relatively high number of variables to estimate for a data set of 32 observations, but the manure - strain interaction seems to meaninful to leave out. These are the estimates of the fixed effects: \\
            \FloatBarrier
            \input{hw4/hw4_2b_lme_fixed} 
            \FloatBarrier
            These are the estimated effects for the manure level, strain, and the interaction between the two. They can be thought of as constant for these strains; there was no randomness in their manifestation, and they don't provide information on any other strain or manure level. Then, these are the predictions of random effects for each plot:
            \FloatBarrier
            \input{hw4/hw4_2b_lme_random} 
            \FloatBarrier
            These are the particular manifestation of a random variable which dictates how an arbitrary plot effect yields. These estimates aren't so interesting in themselves, but their distribution is important to understand how crop yield will vary from place to place.

    \item[c)]
        I've calculated the standard errors and p-values two different ways. The first is by using the built in nlme package assumption that the estimated coefficients are t-distributed in the normal fashion with their rules for degrees of freedom. The difference between manure and not for each plant can then be calculated as the sum of the manure and the interaction coefficients (with standard errors based on the covariance matrix):
            \FloatBarrier
            \input{hw4/hw4_2c_lme_par_ttable} 
            \FloatBarrier
            I checked, and the residuals appear to be approximately normally distributed, though with only four estimated random effect predictions, its impossible to say normality holds. Thus, I also estimated the standard errors and p-values using a non-parametric bootstrap. For each boot, I drew randomly drew, with replacement, residuals and random effects from the observed pool of each, and added these on to the model estimates of the mean. Then, for each boot, I estimated the parameters again, using the sample over all boots to get the statistics of interest. As you can see, these standard errors are somewhat smaller (though the p-values are extremely small either way):
            \FloatBarrier
            \input{hw4/hw4_2c_lme_bs_ttable} 
            \FloatBarrier
        This difference is largely due to the small pool of realized random effects being drawn from. With a parametric bootstrap, the estimates are very close (as you would expect). Its hard to say which is better; in both cases we are hampered by the small pool of realized random effects, and we lack enough data to reject the normal assumption. However, with such small p-values either way, we don't have to worry too much about a false positive.
        \item[d)]
            Using the normal assumption again, we can calculate the difference between S23 and NZ when heavily manured as 
            \[\hat d = \mu+\beta_{S23} + \beta_H + \beta_{S23,H} - \mu-\beta_{NZ} - \beta_H - \beta_{NZ,H} = \beta_{S23} +  \beta_{S23,H}-\beta_{NZ} - \beta_{NZ,H} \]
            With the usual variance for a linear combination of a multivariate normal RV. Using the 21 degrees of freedom given by lme, this gives us a confidence interval of:
            \[ \hat d\pm \sqrt{\var(\hat d)} * t_{.025,21} = (-4.783,15.783) \]

\end{itemize}

\end{document}

### 2015-02-26 stat345/hw4.R

@@ -1,0 +1,84 @@
+library('xtable')
+library('stats4')
+library('locfit')
+library('coefplot')
+library('SDMTools')
+library('plotrix')
+library('combinat')
+library('reshape2')
+library('nlme')
+
+### Problem 2
+rye<-read.table("hw4/ryegrass.dat",sep="",header=T)
+rye$Block<-as.factor(rye$Block)
+rye$str.man<-rep(1:8,each=4)
+rye.rs<-dcast(melt(rye,id=c("Strain","Manure","Block","str.man"),variable.name=c("Yield"),measure.vars="Yield"),Strain+Manure+str.man~Block+Yield)
+
+# a)
+pdf('hw4/hw4_2a_plot.pdf')
+matplot(x=rye.rs$str.man[rye.rs$Manure=="A"]-.2,y=rye.rs[rye.rs$Manure=="A",4],axes=F,pch="1",cex=.8,ylim=c(0,450),xlim=c(.9,8.1),col=1,main="Yield by Strain/Manure/Block",xlab="Strain & Manure Level",ylab="Yield")
+matplot(x=rye.rs$str.man[rye.rs$Manure=="H"]-.2,y=rye.rs[rye.rs$Manure=="H",4],cex=.8,col=2,add=T,pch="1")
+for (i in 1:3) {
+    for (l in c("A","H")){
+        c<- ifelse(l=="A",1,2)
+        matplot(x=rye.rs$str.man[rye.rs$Manure==l]-.2+.08*i,y=rye.rs[rye.rs$Manure==l,4+i],cex=.8,col=c,add=T,pch=paste(i+1))
+    }
+}
+mtext("Digit Indicates Block; Red is High Manure, Black is Average Manure")
+axis(1,labels=paste(rep("",8)),at=1:8,pos=0,lwd=0,lwd.ticks=1)
+axis(1,labels=paste(c("S23","","NZ","","X","","Kent",""),rep(c("- High",""),4)),at=1:8,pos=5,lwd=0,lwd.ticks=0)
+axis(1,labels=paste(c("","S23","","NZ","","X","","Kent"),rep(c("","- Avg"),4)),at=1:8,pos=-12.5,lwd=0,lwd.ticks=0)
+axis(1,labels=c("",""),at=c(0,8.5),pos=0,lwd=1,lwd.ticks=0)
+axis(2,at=seq(0,450,by=50))
+dev.off()
+
+# b)
[…]

### 2015-03-15 stat310/ajmaurer_project.m
Added my code for the project, as well as censoring.m, newimage.m, and sparsify.m, which were given by the professor

@@ -1,0 +1,46 @@
+%%% STAT 310 Final Project
+%%% Aaron Maurer
+
+%%% Caller
+setup = 0
+prob = [1,1,1,1,1]
+
+%%% Setup
+if setup
+    cd cvx
+    cvx_setup
+    cd ..
+end
+
+%%% Problem 1
+if prob(1)
+    % set seed and generate matricies/vectors
+    rand('state',0)
+    A = rand(300,100)
+    c = rand(100,1)-1
+    b = .5 * A * ones(100,1)
+
+    % solve the linear programing problem to get optimal y and lower bound l
+    cvx_begin
+        variable y(100)
+        minimize(c'*y)
+        subject to
+            A*y <= b
+            0 <= y <= 1
+    cvx_end
+    l = cvx_optval
+
+    % For values of tau going from 0 to 1 in increments of .01, round values of y above tau to 0 and above tau to 1 to getxtau (I store all 100 vectors in one matrix). Then, calculate objective value of c'x for each tau
+    tau = .01*(1:100)                             % tau comes out as a 1x100 matrix
+    xtau = repmat(y,1,100) >= repmat(tau,100,1) % xtau(r,c) is 1 if y(r)>=tau(c), and otherwise 0
+    obj = c' * xtau
+
+    % Test inequality constraint - generate the largest entry value of A*x -b for each tau
+    con = max(A * xtau - repmat(b,1,100),[],1)
+
+    % Pick the best objective value for which the constraint is satistifed and compare to our initial lower bound
+    u = min(obj .* (con<=0))
+    diff = u - l
+
+end
+

### 2015-03-15 stat310/sparsify.m

@@ -1,0 +1,14 @@
+% STAT 310 Winter 2015
+% Programming Assignment, Problem 5
+% Version 1.2
+
+rand('state',1);
+randn('state',1);
+n = 500;
+k = 20;
+F = 1.5*rand(n,k)-0.5;
+Sigma = 0.5*F*diag(rand(k,1))*F' + diag(0.1*rand(n,1));
+mu0 = 0.2;
+SR = 0.4;
+mu = mu0 + SR*sqrt(diag(Sigma));
+c = 5 + exp(2*randn(n,1));

### 2015-03-15 stat343/hw6.R

@@ -1,169 +1,0 @@
-library('parallel')
-library('alr3')
-library('faraway')
-library('car')
-library('MASS')
-library('locfit')
-library('xtable')
-library('nlme')
-library('quantreg')
-
-#digits of precision
-d<- 3
-
-#############################
-# Problem 1
-#############################
-
-data(ufcgf)
-attach(ufcgf)
-
-### 5.3.1.1
-fir.lm<-lm(Height~Dbh)
-fir.loess<-loess(Height~Dbh)
-
-G<-sum((predict(fir.lm)-predict(fir.loess))^2)/(summary(fir.lm)$sigma^2)
-
-### 5.3.1.2-4
-B<-1000
-Yhat<-predict(fir.lm)
-ehat<-residuals(fir.lm)
-
-res.boot<- function(c,x,y,e) {
-    ystar<-y+sample(e,length(y),replace=TRUE)
-    lm.m    <-lm(ystar~x)
-    loess.m <-loess(ystar~x)
-    Gstar<-sum((predict(lm.m)-predict(loess.m))^2)/(summary(lm.m)$sigma^2)
-    return(Gstar)
-}
-res.boots<-mcmapply(res.boot,c=1:B,MoreArgs=list(x=Dbh,y=Yhat,e=ehat),mc.cores=4)
-fir.p<-sum(res.boots>G)/B
-
-#############################
-# Problem 3 
-#############################
-
-data(divusa)
-attach(divusa)
-n<-nrow(divusa)
-div.lm<-lm(divorce~unemployed+femlab+marriage+birth+military)
-
-### a)
-pdf("hw6_3_a_res.pdf")
-plot(year,residuals(div.lm),ylab="Residual",xlab="Year",main="Residual by Year")
-dev.off()
-
-pdf("hw6_3_a_auto.pdf")
-plot(tail(residuals(div.lm),n-1)~head(residuals(div.lm),n-1),xlab=expression(hat(epsilon)[i]),ylab=expression(hat(epsilon)[i+1]),main="Residual vs Residual With Lag of 1")
-dev.off()
-
-### b)
-div.alm<-gls(divorce~unemployed+femlab+marriage+birth+military,correlation=corAR1(form=~year),method="ML")
-
-print(xtable(summary(div.alm)$tTable,digits=d),type="latex",file="hw6_3_alm_coef.tex")
[…]

### 2015-03-17 stat310/ajmaurer_project.m
Code done, still writting up

@@ -2,8 +2,8 @@
 %%% Aaron Maurer
 
 %%% Caller
-setup = 0
-prob = [1,1,1,1,1]
+setup = 1 
+prob = [0,1,0,0,0]
 
 %%% Setup
 if setup
@@ -30,17 +30,161 @@ if prob(1)
     cvx_end
     l = cvx_optval
 
-    % For values of tau going from 0 to 1 in increments of .01, round values of y above tau to 0 and above tau to 1 to getxtau (I store all 100 vectors in one matrix). Then, calculate objective value of c'x for each tau
-    tau = .01*(1:100)                             % tau comes out as a 1x100 matrix
-    xtau = repmat(y,1,100) >= repmat(tau,100,1) % xtau(r,c) is 1 if y(r)>=tau(c), and otherwise 0
+    % For values of tau going from 0 to 1 in increments of .01, round values 
+    % of y above tau to 0 and above tau to 1 to getxtau (I store all 100 vectors 
+    % in one matrix). Then, calculate objective value of c'x for each tau
+    tau = .01*(1:100)                     % tau comes out as a 1x100 matrix
+    % xtau(r,c) is 1 if y(r)>=tau(c), and otherwise 0
+    xtau = repmat(y,1,100) >= repmat(tau,100,1) 
     obj = c' * xtau
 
-    % Test inequality constraint - generate the largest entry value of A*x -b for each tau
+    % Test inequality constraint: get the largest entry value of A*x -b for each tau
     con = max(A * xtau - repmat(b,1,100),[],1)
 
-    % Pick the best objective value for which the constraint is satistifed and compare to our initial lower bound
-    u = min(obj .* (con<=0))
+    % Pick the best objective value for which the constraint is satistifed and 
+    % compare to our initial lower bound
+    % each entry of obj is negative, so zeroing out some entries doesn't matter
+    u = min(obj .* (con<=0)) 
     diff = u - l
 
+    % Plot objective value vs constraint
+    plot(con,obj,'.k')
+    hold on
+    plot([0,0],[-45,0],'--r')
+    title('Objective Value vs. Constraint')
+    xlabel('Maximum Entry in Constraint Vector = max _i (Ax_\tau - b)_i')
+    ylabel('Objective Value = c^Tx_\tau')
[…]

### 2015-03-17 stat310/project.tex
write up

--- 2014-11-26 stat343/hw2.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Homework 2 \hfill Aaron Maurer
+STAT 310 Programing Assignment \hfill Aaron Maurer
 \vspace{2mm}
@@
 \vspace{2mm}
+
 \begin{itemize}
     \item[1.]
-        \begin{itemize}
-            \item[a)]
+        After randomly generating $A$ and $c$, setting \({\bf b}=\frac{1}{2}A{\bf 1}\), and solving the continuous relaxation of the boolean linear programing problem, I got a \({\bf y}^*\).
+It only had 16 entries with values that weren't 0 or 1.
+\({\bf c}^T{\bf y}=-35.396\), which we know is a lower bound for \({\bf c}^T{\bf x}\), the optimal objective value for the original problem.
+I generated ${\bf x}_\tau$ for $\tau = .01n, n=1$ to $100$, and plotted the objective value and least slack among the rows of $A{\bf x}_\tau\leq{\bf y}$ below.
+        \begin{center}
+            \includegraphics[width=12cm]{plot_1.pdf}
+        \end{center}
+        The feasible points lie to the left of $0$, corresponding to $A{\bf x}_\tau-{\bf y}$ having all negative entries(and thus the constraint being satisfied).
+Since each point in this region is a feasible point for the original boolean program, the corresponding objective value is an upper bound for the minimum objective value, since the minimum must be less than or equal to each of them.
+The smallest of these values is ${\bf c}^T{\bf x}=-34.373$.
+Thus, we have an upper bound on the optimal point of $-34.373$ and a lower bound of $-35.396$, the difference being $1.023$.
+\\
+
+{\bf Note:} Since there are only $16$ values in ${\bf y}$ which aren't either $0$ or $1$, there could be at most unique 17 ${\bf x}_\tau$ (there are actually only 16 since two of the points which aren't 0 or 1 are within .01 of each other).
+This is because ${\bf x}_\tau = {\bf x}_{\tau+.01}$ if none of the entries of ${\bf y}$ lie in $(\tau,\tau+.01)$.
+\\
+
+        {\bf Code:} \\
+        \lstinputlisting[firstline=17,lastline=61]{ajmaurer_project.m}
+
+    \item[2.]
+        If we set \(A={\bf aa}'\), this problem can be equivalently written as, where \(X\in\mathbb{S}^n\),
         \begin{align*}
-                    \E[\hat{\beta} \vert X] &= \E[(X'X)^{-1}X'Y \vert X] \\
-                                            &= \E[(X'X)^{-1}X'(X\beta+e) \vert X] \\
-                                            &= \E[(X'X)^{-1}X'X\beta+(X'X)^{-1}X'e \vert X] \\
-                                            &= \E[\beta \vert X] + (X'X)^{-1}X' \E[e \vert X] \\
-                                            &= \beta + (X'X)^{-1}X' b(X)
+            \text{minimize    }\; & tr(AX) \\
+            \text{subject to  }\; & X \succeq 0 \\
+                                  & X =\left[ \begin{array}{cccc} .2  & x_1 & x_2 & x_3 \\
+                                                                  x_1 & .1  & x_4 & x_5 \\
+                                                                  x_2 & x_4 & .3  & x_6 \\
+                                                                  x_3 & x_5 & x_6 & .1 \end{array} \right]  \\
+                                  & x_1 \geq 0 & x_2 \geq 0 \\
+                                  & x_4 \leq 0 & x_5 \leq 0 \\
+                                  & x_6 \geq 0
         \end{align*}
-                Thus, $\hat{\beta}$ is only unbiased when \((X'X)^{-1}X' b(X) = 0\).
-                In other words, $\hat{\beta}$ is biased if  $b(X)$ is in the span of \((X'X)^{-1}X'\).
-            \item[b)]
-                \begin{align*}
-                    \var[\hat{\beta} \vert X ] &= \var[(X'X)^{-1}X'Y \vert X ] \\
-                                               &= \var[(X'X)^{-1}X'(X\beta+e) \vert X ] \\
-                                               &= \var[(X'X)^{-1}X'X\beta+(X'X)^{-1}X'e \vert X] \\
-                                               &= \var[\beta+(X'X)^{-1}X'e \vert X] \\
-                                               &= \var[(X'X)^{-1}X'e \vert X] \\
-                                               &= (X'X)^{-1}X'\var[e \vert X]((X'X)^{-1}X')' \\
-                                               &= (X'X)^{-1}X'\sigma^2I X (X'X)^{-1} \\
-                                               &= (X'X)^{-1}X'X(X'X)^{-1} \sigma^2 \\
-                                               &= (X'X)^{-1}\sigma^2 \\
-                \end{align*}
-                Thus, though $\hat{\beta}$ may be biased, it has no higher variance with assumption A2$'$ than with the usual assumption A2.
+        Plugging this into our software, we get that
+        \[X^* = \left[ \begin{array}{cccc} .2    & 0   & .2449 &   0 \\
+                                            0    & .1  &    0  & -.1 \\
+                                           .2449 & 0   & .3    &   0 \\
+                                            0    & -.1 &    0  &   .1 \end{array} \right],
+         \; tr(AX^*) = .0013 \]
+        Its interesting to note that our minimal objective value, even with the restrictions placed on $X$, is quite close to the minimum value you would get for arbitrary $X\succeq0$ of 0.
+
+        {\bf Code:} \\
+        \lstinputlisting[firstline=66,lastline=84]{ajmaurer_project.m}
+
 \end{itemize}
-    \item[2.]
-        \begin{itemize}
-            \item[i.]
-                Given that the variable \(wt_i = \beta_0 + e_i\), where each $e_i$ is normally distributed around $0$ and independent of all other $e_j$ such that $j\neq i$, and independent of the variable height, the probability that a model that included height would reduce the residual sum of squares as much as it has or more is $17.31\%$.
-            \item[ii.]
-                Given that the variable \(wt_i = \beta_0 + \beta_1 ht_i + e_i\), where each $e_i$ is normally distributed around $0$ and independent of all other $e_j$ such that $j\neq i$, the probability that \(\vert \hat\beta_0 \vert \geq 36.8759 \) given a null hypothesis that \(\beta_0=0\) is $58.3\%$.
-           \item[iii.]
-               Given that the variable \(wt_i = \beta_0 + \beta_1 ht_i + e_i\), where each $e_i$ is normally distributed around $0$ and independent of all other $e_j$ such that $j\neq i$, the probability that \(\vert \hat\beta_1 \vert \geq .5821  \) given a null hypothesis that \(\beta_1=0\) is $17.3\%$.
-            \item[iv.]
-                You can not compute the probability of the fitted model being the right one, only the probability of seeing the given data and estimates on it for various assumptions of the underlying process that generated it.
-        \end{itemize}
-    \item[3.]
-        All the individuals formulas for the desired quantities are: \\
-        \[SSreg = \frac{R^2 RSS}{1 - R^2}= \frac{.2084 \times 351191.4}{1-.2084}=92456.14\]
-        \[RSS=(RSE \times df)^2=(2.993 \times 198)^2=351191.4\]
-        \[MSreg = \frac{SSreg}{regDF} = \frac{92456.14}{1}= 92456.14\]
-        \[\hat\sigma^2 = \frac{RSS}{resDF} = \frac{351191.41}{198} = 1773.694\]
-        \[F = \frac{MSreg}{\hat\sigma^2} = \frac{92456.14}{1773.694} = 1773.694 \]
-        Which fit in the table as such: \\
-        \begin{tabular}{l | l l l l l}
-            Regression & df      & ss               & ms       & F     \\
-            \hline
-            Regression & regDF = 1       & SSreg = 92456.14 & MSreg = 92456.14          & F = 52.126 \\
-            Residual   & resDf = 198     & RSS = 351191.4   & $\hat\sigma^2$ = 1773.694 &  
-        \end{tabular}
-    \item[4.]
-        \begin{itemize}
-            \item[a)] These are the resulting confidence intervals: \\
-                \begin{tabular}{l | r l r l}
-                    Confidence Level & $\beta_0$ LB & $\beta_0$ UB & $\beta_1$ LB & $\beta_1$ UP \\
-                    \hline
-                    66\%             &       (3.802,& 3.828)       &       (-.062,& -.059)     \\
-                    90\%             &       (3.792,& 3.837)       &       (-.063,& -.058)     \\
-                    95\%             &       (3.788,& 3.841)       &       (-.064,& -.057)
-                \end{tabular}
-            \item[b)] We get the variance estimate:
-            \[1.538\times10^{-5} = \sigma^2 \left[ \begin{array}{c c} 1 & 10 \end{array} \right] \left(X'X\right)^{-1} \left[ \begin{array}{c} 1 \\  10 \end{array} \right] \]
-            \item[c)] This is what we end up with, for a 95\% CI: \\
-                \begin{tabular}{l | r c l}
-                    Type       & LB    & Fit   & UB    \\
-                    \hline
-                    Mean       & 3.005 & 3.022 & 3.038 \\
-                    Prediction & 2.800 & 3.022 & 3.245 
-                \end{tabular}
-            \item[d)]
-                The confidence interval is smallest near the mean of fixed acidity ($8.31$ in the data set).
-This stands to reason, since the confidence intervals are based on conditioned slices of a multivariate normal distribution; these slices will be steepest towards the center of the distribution (around the means of all variables), and flattest in the tails.
-In other words, its harder to screw up a prediction where you have a lot of data than where you don't have much data.
-\\
-                \begin{center}
-                    \includegraphics[width=12cm]{hw2_plot} 
-                \end{center}
-        \end{itemize}
-        Here is my R code: \\
-        \lstinputlisting{hw2.R}
-\end{itemize}
+
 \end{document}

### 2015-03-17 stat310/ajmaurer_project.m
small change

@@ -3,7 +3,7 @@
 
 %%% Caller
 setup = 1 
-prob = [0,1,0,0,0]
+prob = [0,0,1,0,0]
 
 %%% Setup
 if setup
@@ -97,6 +97,12 @@ if prob(3)
     CT = speye([m*(n-1),m*n]) - [sparse(m*(n-1),m),speye(m*(n-1),m*(n-1))] % Top of C matrix
     CB = speye([(m-1)*n,m*n]) - [sparse((m-1)*n,1),speye((m-1)*n,m*n-1)] % bottom of C matrix
     C  = [CT;CB]
+
+    % Show sparse image
+    colormap gray
+    imagesc(abs(X-127*(1-S)))
+    print('image_2_miss.pdf')
+    close
 
     %% part a: L2 variation. 
 

### 2015-03-17 stat310/project.tex
finished with project

@@
     \item[1.]
-        After randomly generating $A$ and $c$, setting \({\bf b}=\frac{1}{2}A{\bf 1}\), and solving the continuous relaxation of the boolean linear programing problem, I got a \({\bf y}^*\).
+        After randomly generating $A$ and ${\bf c}$, setting \({\bf b}=\frac{1}{2}A{\bf 1}\), and solving the continuous relaxation of the boolean linear programing problem, I got a \({\bf y}^*\).
 It only had 16 entries with values that weren't 0 or 1.
-\({\bf c}^T{\bf y}=-35.396\), which we know is a lower bound for \({\bf c}^T{\bf x}\), the optimal objective value for the original problem.
-I generated ${\bf x}_\tau$ for $\tau = .01n, n=1$ to $100$, and plotted the objective value and least slack among the rows of $A{\bf x}_\tau\leq{\bf y}$ below.
+\({\bf c}^T{\bf y}^*=-35.396\), which we know is a lower bound for \({\bf c}^T{\bf x}\), the optimal objective value for the original problem.
+I generated ${\bf x}_\tau$ for $\tau = .01n, n=1$ to $100$, and plotted the objective value against the minimum slack among the rows of $A{\bf x}_\tau\leq{\bf y}$ below.
         \begin{center}
@@
         \end{center}
-        The feasible points lie to the left of $0$, corresponding to $A{\bf x}_\tau-{\bf y}$ having all negative entries(and thus the constraint being satisfied).
+        The feasible points lie to the left of $0$, corresponding to $A{\bf x}_\tau-{\bf y}$ having all negative entries (and thus the constraint being satisfied).
 Since each point in this region is a feasible point for the original boolean program, the corresponding objective value is an upper bound for the minimum objective value, since the minimum must be less than or equal to each of them.
 The smallest of these values is ${\bf c}^T{\bf x}=-34.373$.
-Thus, we have an upper bound on the optimal point of $-34.373$ and a lower bound of $-35.396$, the difference being $1.023$.
+Thus, we have an upper bound on the objective value at optimal point of $-34.373$ and a lower bound of $-35.396$, the difference being $1.023$.
 \\
@@
 
+    \item[3.]
+        We can transform $X$ into a vector ${\bf x}$ such that \(x_{ij}={\bf x}_{ni+j}\) to make this closer to the perfered form for a LP/QP problem.
+Then, we can make a sparse matrix $C$ such that $C{\bf x}=\nabla X$.
+Each row of $C$ will have all $0$s except one $1$ and one $-1$.
+In the first $(m-1)n$ rows of $C$, the $-1$ will be along the main diagonal, and the $1$ will be $n$ spots to the right, making ${\bf c}_i'{\bf x}=x_ij-x_{i-1,j}$.
+In the remaining $m(n-1)$ rows of $C$, there will be $-1$s along the diagonal starting from the top left corner of the section and $1$s one spot to the right, making ${\bf c}_i'{\bf x}=x_ij-x_{i,j-1}$.
+\\
+        \smallskip \\
+        Now, here is the original image on the left, and the original image with the missing pixes greyed out on the right: \\
+        \includegraphics[width=9cm]{image_2_orig.pdf}
+        \includegraphics[width=9cm]{image_2_miss.pdf}
+
+        \begin{itemize}
+            \item[(a)]
+                When we use the $L_2$-variation, our problem can be formatted for CVX as: 
+                \begin{align*}
+                    \text{minimize    }\; & \|C{\bf x}\|_2 \\
+                    \text{subject to  }\; & x_{ij}=a_{ij},\, (i,j)\in S  \\
+                \end{align*}
+                Which has the equivalent solution to the QP problem  
+                \begin{align*}
+                    \text{minimize    }\; & {\bf x}'C'C{\bf x} \\
+                    \text{subject to  }\; & x_{ij}=a_{ij},\, (i,j)\in S  \\
+                \end{align*}
+                Optimizing the first problem in CVX, and transforming ${\bf x}^*$ into $X^*$, we get this image:
+
+                \begin{center}
+                    \includegraphics[width=9cm]{image_2a.pdf}
+                \end{center}
+                It's decently close to the original, but blurry and not sharp at all
+            \item[(b)]
+                When we use the $L_1$-variation, our problem can be formatted for CVX as: 
+                \begin{align*}
+                    \text{minimize    }\; & \|C{\bf x}\|_1 \\
+                    \text{subject to  }\; & x_{ij}=a_{ij},\, (i,j)\in S  \\
+                \end{align*}
+                Which has an equivalent solution to the LP problem  
+                \begin{align*}
+                    \text{minimize    }\; & {\bf 1}'{\bf t} \\
+                    \text{subject to  }\; & {\bf t}\geq C{\bf x} \\
+                                          & {\bf t}\geq -C{\bf x} \\  
+                                          & x_{ij}=a_{ij},\, (i,j)\in S  \\
+                \end{align*}
+                Optimizing the first problem in CVX, and transforming ${\bf x}^*$ into $X^*$, we get this image:
+                \begin{center}
+                    \includegraphics[width=9cm]{image_2b.pdf}
+                \end{center}
+                Which is much better; its much closer to the original and doesn't have nearly as much blurring as with the $L_2$ variation.
+        \end{itemize}
+
+        {\bf Code:} \\
+        \lstinputlisting[firstline=90,lastline=145]{ajmaurer_project.m}
+
+    \item[4.]
+        \begin{itemize}
+            \item[(a)]
+                A simple way to estimate x with a LCQP is to minimize the squared error over the first $k$ terms where $y$ is known under the restriction that the predictions for the remaining terms are higher than $\beta$.
+If \(A_1=[{\bf a}_1,...,{\bf a}_k]', A_2=[{\bf a}_{k+1},...,{\bf a}_n]',\) and \({\bf b}_1=[b_1,...,b_k]\), this gives us the optimization problem
+                \begin{align*}
+                    \text{minimize    }\; & \|A_1{\bf x}-b_1\|_2^2 \\
+                    \text{subject to  }\; & A_2{\bf x}\geq \beta{\bf 1} \\
+                \end{align*}
+                This is a simplification which will probably bias our prediction upwards.
+Just because an observed value is higher than $\beta$ doesn't mean its expectation is, since the error could be greater than \({\bf a}_i'{\bf x}-\beta\).
+If we knew something about the distribution of the error, we could, for instance, instead have the constraint \(A_2x\geq c{\bf 1}\), where $c$ is a constant such that \(P(\max_{k<i\leq n}(\varepsilon_i)\geq \beta-c)\leq \alpha\).
+This would give us $1-\alpha$ confidence that we are allowing for the true values of \(A_2{\bf x}\).
+                \smallskip  \\         
+                We don't have enough information to do this in the data description though, so instead I ran the first version I gave, giving
+                \[{\bf x}_{qp}=
+                    \begin{array}{cccccccccc} 
+                       [-0.22 & -1.69 &  0.40 &  0.18 & -1.06 &  1.28 & 1.13 & -0.01 &  0.30 &  0.38 \\
+                        -0.14 &  0.81 & -0.32 &  2.12 & -0.24 & -0.02 & 0.97 & -0.11 & -0.15 & -0.89]
+                    \end{array} \]
+            \item[(b)]
+                The usual least squares estimator based on just the first $k$ points is given by
+                \begin{align*}
+                    {\bf x}_{ls} &= (A_1'A_1)^{-1}A_1' b_1 \\
+                                 &= \begin{array}{cccccccccc} 
+                                        [-0.35& -1.80&  0.20& 0.17& -0.84&  1.30& 1.83& -0.56&  0.37& -0.05 \\
+                                         -0.11&  1.53& -0.50& 2.42& -0.56& -0.37& 0.99& -0.25& -0.18& -0.43]
+                                    \end{array} \\
+                \end{align*}
+            \item[(c)]
+                Comparing these two estimates to the original, the LCQP has the smaller relative error:
+                \[\frac{\|{\bf x}_{qp}-{\bf x}\|_2}{\|{\bf x}\|_2} = .1538 \]
+                while                         
+                \[\frac{\|{\bf x}_{ls}-{\bf x}\|_2}{\|{\bf x}\|_2} = .3907 \]
+        \end{itemize}
+
+        {\bf Code:} \\
+        \lstinputlisting[firstline=152,lastline=171]{ajmaurer_project.m}
+    \item[5.] 
+        The problem as given is
+        \begin{align*}
+            \text{minimize    }\; & \|{\bf x}\|_1 \\
+            \text{subject to  }\; & \E[(y-{\bf a}'{\bf x})^2]\leq .01 \E(y^2) \\
+        \end{align*}
+        Using a few identities from statistics, we can transform the constraint into a quadratic constraint:
+        \begin{align*}
+            \E[(y-{\bf a}'{\bf x})^2] &\leq .01 \E[y^2] \\
+            (\E[y]-\E[{\bf a}'{\bf x}])^2 + \var(y-{\bf a}'{\bf x}) &\leq .01 (\E[y]^2+\var(y)) \\
+           ({\bf \mu}'{\bf c}-{\bf \mu}'{\bf x})^2 + \var(y)-2\cov(y,{\bf a}'{\bf x})+\var({\bf a}'{\bf x}) &\leq .01 ((\mu'{\bf c})^2+\var(y)) \\
+            ({\bf \mu}'{\bf c}-{\bf \mu}'{\bf x})^2 + {\bf c}'\Sigma {\bf c} - 2{\bf c}'\Sigma {\bf x}+ {\bf x}'\Sigma {\bf x} &\leq .01 (({\bf \mu}'{\bf c})^2+{\bf c}'\Sigma {\bf c}) \\
+            .99({\bf \mu}'{\bf c})^2 - 2{\bf \mu}'{\bf c}{\bf \mu}'{\bf x} + ({\bf \mu}'{\bf x})^2 + .99{\bf c}'\Sigma {\bf c} - 2{\bf c}'\Sigma {\bf x} + {\bf x}'\Sigma {\bf x} &\leq 0 \\
+        \end{align*}
+        Giving us the QCLP
+        \begin{align*}
+            \text{minimize    }\; & \|{\bf x}\|_1 \\
+            \text{subject to  }\; & .99({\bf \mu}'{\bf c})^2 - 2{\bf \mu}'{\bf c}{\bf \mu}'{\bf x} + ({\bf \mu}'{\bf x})^2 + .99{\bf c}'\Sigma {\bf c} - 2{\bf c}'\Sigma {\bf x} + {\bf x}'\Sigma {\bf x} \leq 0 \\
+        \end{align*}
+        Our optimal \({\bf x}^*\) isn't perfectly sparse, but its quite close, with the majority of its entries close to 0.
+Of the $500$ entries, only 50 are larger than .1, 52 are larger than .001, 69 are larger than .0001, and 373 are larger than .00001.
+\\
+        {\bf Code:} \\
+        \lstinputlisting[firstline=179,lastline=199]{ajmaurer_project.m}
+
 \end{itemize}

### 2015-03-17 stat310/ajmaurer_project.m

@@ -3,7 +3,7 @@
 
 %%% Caller
 setup = 1 
-prob = [0,0,1,0,0]
+prob = [0,0,0,0,1]
 
 %%% Setup
 if setup
@@ -31,7 +31,7 @@ if prob(1)
     l = cvx_optval
 
     % For values of tau going from 0 to 1 in increments of .01, round values 
-    % of y above tau to 0 and above tau to 1 to getxtau (I store all 100 vectors 
+    % of y above tau to 0 and above tau to 1 to get xtau (I store all 100 vectors 
     % in one matrix). Then, calculate objective value of c'x for each tau
     tau = .01*(1:100)                     % tau comes out as a 1x100 matrix
     % xtau(r,c) is 1 if y(r)>=tau(c), and otherwise 0
@@ -88,15 +88,21 @@ end
 if prob(3) 
 
     % Get Matricies.
-    newimage
+    newimage        % Given program
 
-    % Convert matricies so this can be formatted as a QP/LP. Matricies are huge but sparse
-    % both times, the objective is a norm of Cx
-    xs = sparse(reshape(S.*X,[m*n,1]))      % elements of X, or 0 if S is 0 (data missing)
+    % Convert matricies so this can be formatted as a QP/LP. Matricies huge but sparse
+    % For both problems, the objective is a norm of Cx
+    xs = sparse(reshape(S.*X,[m*n,1]))      % elements of X, or 0 if data missing
     SD = sparse(diag(reshape(S,[m*n,1])))   % has elements of S along diagonal
-    CT = speye([m*(n-1),m*n]) - [sparse(m*(n-1),m),speye(m*(n-1),m*(n-1))] % Top of C matrix
-    CB = speye([(m-1)*n,m*n]) - [sparse((m-1)*n,1),speye((m-1)*n,m*n-1)] % bottom of C matrix
+    CT = speye([m*(n-1),m*n]) - [sparse(m*(n-1),m),speye(m*(n-1),m*(n-1))] % Top of C
+    CB = speye([(m-1)*n,m*n]) - [sparse((m-1)*n,1),speye((m-1)*n,m*n-1)] % bottom C
     C  = [CT;CB]
+
+    % Show original image
+    colormap gray
+    imagesc(X)
+    print('image_2_orig.pdf')
+    close
 
     % Show sparse image
     colormap gray
@@ -106,7 +112,8 @@ if prob(3)
 
     %% part a: L2 variation. 
 
-    % optimize the problem, in this case a QP
[…]

### 2015-03-17 stat345/hw1.tex

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 345 Homework 1 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[(a)]
                This experiment was designed to test the hypothesis that female platyfish are more attracted to male platyfish who have a sword like tail. 
                To test this, a randomized complete block design was used. \\
                Male platyfish were the experimental units and were paired into six blocks of two based on body size and coloration. 
                Each platyfish within a block was assigned one of two treatments; the control treatment was the surgical attachment of a transparent 25mm plastic "sword" to the back of the fish, while the alternative treatment was the attachment of a yellow and black 25mm plastic "sword" to the back of the fish. In both cases, this was done three days prior to the experiment. It is not stated how assignment of treatment within a block was preformed. 
                Each of these paired platyfish were placed simultaneously into the first and third section of an aquarium divided into equal thirds. A female platyfish was then placed in the middle section for ten minutes. This process was then repeated, except with the sides the males are on switched, resulting in the female being in the aquarium for twenty minutes. The response variable, measured for each male fish independently, was the number of seconds (the observational unit) which the female fish spent in the third of her third of the aquarium adjacent to a particular male fish's section.
                This entire process was replicated for each block with between nine and sixteen different female platyfish.
            \item[(b)] The standard errors in table 1 represent the standard deviation of the sampled mean around the expected time a female fish would spend in the section adjacent to a particular male fish. If we treat the time spent near one fish and another as independent, and assume each is normally distributed, we can get the standard error of the mean difference from the formula
                \[ SE_{\bar X_1 - \bar X_2} = \sqrt{SE_{\bar X_1}^2 + SE_{\bar X_2}^2}\]
                applying this to the six pairs, we get estimates 
                \FloatBarrier
                \input{hw1/p_1_a.tex}
                \FloatBarrier
                I think the standard errors should in fact be larger than this. Since a female fish can only spend time one one side at a time, the response for the two treatments is likely inversely correlated. This would cause larger variance in the difference than one would expect if they were independent.
            \item[(c)]
                If we make a number of assumptions, we can take a similar approach to b to generate a confidence interval using the normal approximation. Assumptions:
                \begin{itemize}
                    \item[1.] As before, the response for the yellow and transparent swords for a particular pair are independent. This is an obviously incorrect assumption, since by construction these variables are negatively correlated, as mentioned before.
                    \item[2.] The means for yellow and transparent respectively are independent between each male and female combination. One could imagine ways in which this could be false, such as females developing preference after being used to test multiple pairs of males, but is probably reasonably safe.
                    \item[3.] The mean difference for yellow and transparent for a particular pair of males is the same as for all other pairs of males. Eyeballing the data, it looks like pairwise their is sufficient data to reject this assumption in at least a few cases (pairs 1 and 4), but this is cherry picking. Whether or not we can show this to be a statistically invalid assumption though, it is unreasonably strong, and one I would not make if I had more complete data to work with.
                    \item[4.] The means for yellow and transparent respectively across all males and females are normally distributed. Given assumption 2, this will be true asymptotically. There are enough trials this is reasonable in itself.
                \end{itemize}
                Given these assumptions, if we first calculate the mean time across all males for yellow and transparent respectively and the associated standard error, we can do a difference of means on that. This will give us a confidence interval of
                \[CI=\sum_{i=1}^6 \bar X_{i,y} - \sum_{i=1}^6 \bar X_{i,t} \pm \frac{1.96}{\sum_{i=1}^6 n_i}\sqrt{\sum_{i=1}^6\left(SE_{i,y}^2+SE_{i,t}^2\right)  n_i^2}=277.7\pm1.96\cdot=(223.5,331.9)\]
            \item[(d)] This additional experiment strengthens the authors argument by using the principle of self-control. When using different fish, no matter how careful the experimenter is in blocking, there is always the possibility that there is some still unaccounted for variance between experimental units. However, if both treatments can be applied to the same unit, as he does here by switching tails, any change in response should be just the result of change in treatment. Since the response changed significantly with the switch in treatment on the same unit, it is a very strong indicator that the treatment is responsible for the change.
            \item[(e)] I generally find the rest of her argument that the preexisting female bias could account for the evolution of the sword in the swordtail convincing, with the caveat that I don't think she precluded that another mechanism could have been more important. Her three points stand to reason; if the females of a species had a preference for a trait before it existed in the males, it would stand to reason this would drive its evolution. If one accepts that platyfish are a good representation of swordtails before the sword evolved, then she has established her three points well through other work and this experiment. However, it could well be the case the male trait which females are biased towards also coincidentally had other evolutionary advantages which are as powerful or more powerful than female attraction. For instance, if the sword made the fish faster, then this could also have been a more powerful force to drive its arising. This always allows the possibility that female preference was an additional motivator, rather than the main one, as the author at times suggests. 
    \end{itemize}

    \item[2.]  
        \begin{itemize}
            \item[(a)] Rejection of the null offers two possibilities: the acceleration of gravity varies over time, or that at least some of the experiments were biased. The former would contradict well established physics theory, which makes it unlikely, so I would tend to concluding the latter.
            \item[(b)] 
                When we calculate the $f$ statistic, 
                \[ f = \frac{\sum_{i=1}^{8} n_i (\bar Y_i - \bar Y)^2/(8-1)}{\sum_{i=1}^{8} \sum_{j=1}^{n_i} (Y_{ij} - \bar Y_i)^2/(81-8)} = 3.56 \]
                which corresponds to a p-value of $.0024$, from a F distribution with 73 and 7 degrees of freedom. On this basis we would reject the null hypothesis that each series has the same mean. This test holds true under the assumptions that each series is normally distributed with identical variance and that each observation is independent of the others.
            \item[(c)] 
                Fitting the two models by max likelihood using the gradient optimization, I got two sets of parameters. For the null, where the mean is equal across series:
                \FloatBarrier
                \input{hw1/p_2_c0.tex}
                \FloatBarrier
                and where it varies by series:
                \FloatBarrier
                \input{hw1/p_2_c1.tex}
                \FloatBarrier
                Plugging the log likelihoods to calculate the log likelihood ratio statistic, we get:
                \[-2\log(LL_0)+2\log(LL_1) = -2\cdot-281.2+2\cdot-271.9=18.7\]
                Assuming this statistic is chi-square distributed with $7$ degrees of freedom, we get a p-value of $.0092$.
            \item[(d)]
                When we plot the series sequentially, several of our assumptions from the prior parts look off:
                \begin{center}
                    \includegraphics[width=12cm]{hw1/p_2_d}
                \end{center}
                from eyeballing this plot, it appears that the different series are, in fact, just sections of one series. This series appears to have decreasing variance over time, and either has a periodic component or has inverse auto-correlation (driving the systematic up and down spikes). Since all the prior tests depended on independence and identical distribution within a series, if this is correct it would invalidate their results. \\
            \item[(e)]
                To deal with the apparent dependence structure and non constant variance, I used a generalized least squares regression to predict the mean. I allowed for a weighting structure where the variance was a constant plus a power function of the number of measurements, and I adjusted for autocorrelation. The end result is an estimate of $78.75$ for the mean deviation, and a standard error of $1.08$
                \lstinputlisting[firstline=2,lastline=30]{hw1/p2e.txt}
        \end{itemize}
\end{itemize}

\end{document}

### 2015-03-17 stat345/hw3.tex

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

STAT 345 Homework 3 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
{\bf Problem 1}
\begin{itemize}
    \item[a)]
        We are using the model \(y_{ij}=\mu+b_i+t_j+e_{ij}\), where $i\in\{1,2,3,4\}$ and $j\in\{A,B,C,D\}$ and, by the treatment constraint, $b_1=t_O=0$. If we define \(Y_{..},Y_{i.},Y_{.j}\) as the sum over the respective blocks and/or treatments, then we get the following equations:
        \begin{itemize}
            \item[1.] \( Y_{..}=13\hat\mu+3(\hat b_2 + \hat b_3 + \hat b_4) + 3(\hat t_A + \hat t_B + \hat t_C)\)
            \item[2.] \( Y_{1.}=4\hat\mu+\hat t_A + \hat t_B + \hat t_C\)
            \item[3.] \( Y_{2.}=3\hat\mu+\hat t_A + \hat t_B + 3\hat b_2 \)
            \item[3.] \( Y_{3.}=3\hat\mu+\hat t_A + \hat t_C + 3\hat b_3 \)
            \item[4.] \( Y_{3.}=3\hat\mu+\hat t_B + \hat t_C + 3\hat b_4 \)
            \item[5.] \( Y_{.O}=4\hat\mu+\hat b_2 + \hat b_3 + \hat b_4 \)
            \item[6.] \( Y_{.A}=3\hat\mu+3\hat t_A \hat\hat b_2 + \hat b_3  \)
            \item[7.] \( Y_{.B}=3\hat\mu+3\hat t_B \hat\hat b_2 + \hat b_4  \)
            \item[8.] \( Y_{.C}=3\hat\mu+3\hat t_C \hat\hat b_3 + \hat b_4  \)
        \end{itemize}
        From this, we get that
            \[4Y_{..} - Y_{1.} + 21Y{.O}=132\hat\mu + 11(\hat t_A + \hat t_B +\hat t_C) + 33(\hat b_2 + \hat b_3 + \hat b_4) \]
        Subtracting the above from 
            \[11Y_{4.} + 33Y_{.A} = 132\hat\mu + 99 \hat t_A + 11\hat t_B + 11\hat t_C + 33(\hat b_2 + \hat b_3 + \hat b_4) \]
        We get 
            \[11Y_{4.} + 33Y_{.A} - 4Y_{..} - Y_{1.} + 21Y{.O} =  88\hat t_A\]
        So, we get estimates of 
        \begin{align*}
            \hat t_A &= \frac{1}{88}(11Y_{4.} + 33Y_{.A} - 4Y_{..} - Y_{1.} + 21Y_{.O}) \\
            \hat t_A &= \frac{1}{88}(40y_{4a}+6y_{4O}+30y_{1A}-24y_{1O}+29(y_{2A}+y_{3A})-25(y_{2O}+y_{3O})+7(y_{4B}+y_{4C})-3(y_{1B}+y_{1C})) \\
            \var(\hat t_A) &=\frac{40^2 + 6^2 + 30^2 +24^2 + 2\times29^2 + 2\times 25^2 + 2\times 7^2 + 2\times 3^2}{88^2}\sigma^2 \\
            \var(\hat t_A) &=\frac{6174}{7744}\sigma^2 \approx .797\sigma^2
        \end{align*}
        Repeating this for $\hat t_B$ and $\hat t_C$, the process is identical, except we subtract \(4Y_{..} - Y_{1.} + 21Y{.O}\) from \(11Y_{3.} + 33Y_{.B}\) and \(11Y_{2.} + 33Y_{.C}\) respectively. This will give us
        \begin{align*}
            \hat t_B &= \frac{1}{88}(11Y_{3.} + 33Y_{.B} - 4Y_{..} - Y_{1.} + 21Y_{.O}) \\
            \var(\hat t_B) &=\frac{6174}{7744}\sigma^2 \approx .797\sigma^2 \\
            \hat t_C &= \frac{1}{88}(11Y_{2.} + 33Y_{.C} - 4Y_{..} - Y_{1.} + 21Y_{.O}) \\
            \var(\hat t_C) &=\frac{6174}{7744}\sigma^2 \approx .797\sigma^2
        \end{align*}
    \item[b)]
        The best design I found was this:  \\
            \FloatBarrier
            \input{hw3/hw3_1a_comb} 
            \FloatBarrier
        As you would want, each treatment appears and equal number of times, there are no treatments appearing multiple times in the same block, and very pair of treatments had either 3 or 4 primary links:
            \FloatBarrier
            \input{hw3/hw3_1a_P} 
            \FloatBarrier
        And those that only had 3 primary links had slightly more secondary links:
            \FloatBarrier
            \input{hw3/hw3_1a_S} 
            \FloatBarrier
\end{itemize}

\newpage
STAT 345 Homework 3 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

{\bf Problem 2}
\begin{itemize}
    \item[a)]
        When I fit a model with an additive fixed effect using all the observations, I get an estimated standard deviation of $\sigma=11.589$. However, when I fit the same model, but only include treatment C and D, the estimated standard deviation is $\sigma=4.275$. When we plot the actual data, we see why:
        \begin{center}
            \includegraphics[width=16cm]{hw3/hw3_2a_plot}
        \end{center}
        Even though the total amount of sleep each participant gets varies, the difference between their sleep under treatment C and D is relatively consistent. Thus, a model just including these two treatments has a relatively low prediction of variance. On the other hand, the difference between the sleep a participant gets on the drugs and without any drugs is not very consistent. Thus, we get a higher prediction of variance including it. This makes sense, because C and D are very similar drugs, so it can be expected that one person won't respond to one of the drugs strongly but not the other, while this may not hold for drug B or sleep without any drugs.

    \item[b)]
        There is evidence of an interaction between person in treatment. As noted above, the difference between each treatment does not seem consistent over all the people. In particular, People who already sleep relatively long hours do not seem to benefit from any of the drugs, while the change in sleep under different drugs does not seem consistent from person to person among those for whom this is not true. The fact that the relatively similar drugs have a consistent difference though suggests that this lack of consistency represents meaningful difference by person, corresponding to a person specific effect for each drug, rather than just random variation.

    \item[c)] The mean effect over the entire population is not of particular importance; as we've seen, those who sleep relatively long periods already do not seem to benefit from the drug, so irrespective of the population mean, its likely not worth giving it to them. Each person's existing sleep habits seems to be more important to predicting effectiveness than the population as a whole. Thus, a population mean by the amount one sleeps unmedicated could be useful. Among those who do not sleep longer periods, there is likely to benefit to medication, but the benefit is random by person, so having some sense of this distribution would be useful.

\end{itemize}

\end{document}

### 2015-03-17 stat345/hw5.tex

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 345 Homework 4 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
{\bf Problem 1}
\begin{itemize}
    \item[a.]
        The distribution of Z is identical to the distribution of the described variance estimator when the error is iid $N(0,1)$. We can see this by the correspondence between its creation and the prescribed procedure:
        \begin{itemize}
            \item[i.] $S_Y^2$ has the same distribution as the residual sum of squares from the model where 2-factor interactions are assumed (giving residuals with five degrees of freedom), and $S_X^2$ has the same distribution as the marginal sum of squares attributable to the two factor interactions. This makes \(\frac{S_X^2}{S_Y^2}\) identically distributed to the F statistic for the comparison of the model without two factor interactions to the model with two factor interactions. 
            \item[ii.] Thus, $\frac{S_X^2}{S_Y^2}<4.95 = F_{.95}(6,5)$ is equivalent to not rejecting the null that there are no two factor interactions. In the procedure, when this occurs the two factor interactions are pooled into the error and the error variance is estimated on 11 degrees of freedom. This is equivalent to \(Z=\frac{5S_Y^2+6S_X^2}{11}\).
            \item[iii.] $\frac{S_X^2}{S_Y^2}\geq4.95$ is equivalent to rejecting the null, so the error is estimated based on the original 5 degrees of freedom. In this case, that is distributed the same as $Z=S_Y^2$.
        \end{itemize}
    \item[b.] Over 1000 simulations, the mean of Z was $.951$, and the mean squared error was $.180$. I simulated it by the same method as described in part a. These values are meaningfully lower than both the true value $\sigma^2=1$ and the mean squared error of $S_Y^2=.4$. We can derive the latter by noting $S_Y^2\sim \frac{\chi_5^2}{5}$, so
        \[ \E((S_Y^2-1)^2)= \E(S_Y^2-1)^2 + \var(S_Y^2) = \frac{\var(\chi_5^2)}{25} = \frac{10}{25} \] 
        The reason we see the underestimate of $\sigma^2$ is that with each $\beta$ set to 0, we are essentially drawing $11$ $\chi^2$ variables, and throwing out the last 6 if they are too much larger than the first 5, then averaging over however many remain. The end result is that the average will be biased downwards. The mean square error is smaller because $95\%$ of the time it is average of $11$ instead of $5$ $\chi^2$ variables, reducing the variance more than enough to compensate for the small additional bias.
    \item[c.] When $\beta_1$ is simulated 1000 times at different values, this is the mean and $MSE$ I got: 
        \FloatBarrier
        \input{hw5/hw5_1c} 
        \FloatBarrier
        As you would expect, initially both the Mean and MSE of Z increase with $\beta_1$. Since the $F$ test has relatively low power for small, positive $\beta$, the null is rarely rejected, and the mean and MSE both go up with $S_x^2$. However once $\beta_1$ is large enough and the power is high, the test rejects more often, so both the mean and MSE drop back down again. We see this clearest at $\beta_1=8$, where the test should reject the vast majority of the time, making it so that $Z\approx S_Y^2$.
    \item[d.]
        The obvious disadvantage of pooling the interactions into the error when a test determines them insignificant is the potential for bias. When there are no interactions, the estimate of the variance is biased downwards, and when there are nonnegative interactions and low power, the variance is significantly biased upwards. The advantage is that for small or non-existent interactions, you still have a better MSE due to the decreased variance of your estimator. My advice is that this is probably safe if your significance test has high power (though this probably makes the pooling unnecessary), but otherwise I would avoid employing this technique; as our tests have shown, it seems that you generally end up with a worse estimator.
\end{itemize}
\newpage
\null
\newpage
STAT 345 Homework 4 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
{\bf Problem 2}
\begin{itemize}
    \item[a.]
        From those two defining relationships, we can figure out what is aliased with the identity via :
        \[E=ABC \, \& \, F=BCD \Longrightarrow I=ABCE \, \& \, I = BCDF\]
        Multiplying ABCE by BCDF gives us the last element aliased with I:
        \[I = ABCE = BCDF = ADEF\]
        Now, multiplying through by remaining terms in the data set gives us the rest of the relationships:
        \begin{alignat*}{4}
            A    &=& BCE   &=& ABCDF  &=& DEF \\
            B    &=& ACE   &=& CDF    &=& ABDEF \\
            AB   &=& CE    &=& ACDF   &=& BDEF \\
            C    &=& ABE   &=& ADF    &=& ADCEF \\
            AC   &=& BE    &=& ABDF   &=& CDEF \\
            BC   &=& AE    &=& ABCDEF &=& DF \\
            ABC  &=& E     &=& BCDEF  &=& ADF \\
            D    &=& ABCDE &=& BCF    &=& AEF \\
            AD   &=& BCDE  &=& ABCF   &=& EF \\
            BD   &=& ACDE  &=& CF     &=& ABEF \\
            ABD  &=& CDE   &=& ACF    &=& BEF \\
            CD   &=& ABDE  &=& BF     &=& ACEF \\
            ACD  &=& BDE   &=& ABF    &=& CEF \\
            BCD  &=& ABCEF &=& F      &=& ADE \\
            ABCD &=& BCEF  &=& AF     &=& DE \\
        \end{alignat*}
    \item[b.]
        As I'll show from the regression model, the best combination seems to be A, B, D at the "-" level, and C, E, and F at the "+" level, though besides A and B, the rest don't seem to have a significant impact. \\
        \indent Running a fully interacted model, we get these estimations (effects are under the sum constraint and for the "-" level):
        \FloatBarrier
        \input{hw5/hw5_2b_rtable} 
        \FloatBarrier
        Based off this, it seems like A- and B- is unquestionably a must, since their individual effects are so large negative, outweighing their positive interaction. Then, with D- having a negative coefficient and A-D- having a relatively large negative coefficient, D- seems a good choice too. A-B-C- has a relatively large positive coefficients, so C+ is the better choice. F and E, as determined by the aliasing relationship in the data, should be thus set to "+". \\
        However, looking at the half normal plot of the absolute value of the coefficients, besides A and B its unclear if any of the estimates are significant enough to worry about:
        \begin{center}
            \includegraphics[width=16cm]{hw5/hw5_2b_halfnorm}
        \end{center}
        Clearly B-, A-, and A-B- have meaningful effects. A-D- and its neighbor, A-C-D- are borderline, only maybe lying above the expected line for normal noise which the remaining coefficients fall on or below. \\
        Finally, I tried to look into whether there was predictable variation in shrinkage. Obviously, with a fully interacted model, we can't estimate the variance at all, but since the interactions generally seemed small, I estimated a model with no interactions (but including E and F, as determined by the aliasing). I then fit a model predicting the log absolute value of the residual:
        \FloatBarrier
        \input{hw5/hw5_2b_lrtable} 
        \FloatBarrier
        The one 'significant' p-value is borderline, and based on the F-statistic comparing this to a mean model, we get a p-value of $.348$. Thus, we don't have evidence to say the shrinkage variation changes with any of our variables.
    \item[c.]
        Having B set to "-" was clearly the most important choice, so I used that as the basis for this analysis. Now, the Aliasing is:
        \begin{alignat*}{4}
            I    &=& ACE   &=& CDF    &=& ADEF \\
            A    &=& CE    &=& ACDF   &=& DEF \\
            C    &=& AE    &=& ADF    &=& ADCEF \\
            AC   &=& E     &=& ADF    &=& CDEF \\
            D    &=& ACDE  &=& CF     &=& AEF \\
            AD   &=& CDE   &=& ACF    &=& EF \\
            CD   &=& ADE   &=& F      &=& ACEF \\
            ACD  &=& DE    &=& AF     &=& CEF \\
        \end{alignat*}
        Repeating our analysis from before, we get the following coefficients from the regression:
        \FloatBarrier
        \input{hw5/hw5_2c_rtable} 
        \FloatBarrier
        Which I have plotted on a half norm plot here:
        \begin{center}
            \includegraphics[width=16cm]{hw5/hw5_2c_halfnorm}
        \end{center}
        With so few data points, its hard to say if any of these are significant; besides the intercept, this could easily be a null plot. However, if we have to make a choice, A-D- still has the biggest negative coefficient, even if the coefficient on just A- has shrunk, while  A-C-D- has an equally large positive coefficient. Thus, The same combination, A-, C+, D-, E+, F+ still seems like the best option.

\end{itemize}

\end{document}

### 2015-04-07 stat302/hw1.tex
Started homework 1

--- 2014-12-03 stat343/hw8.tex
@@
 \begin{document}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-STAT 343 Homework 8 \hfill Aaron Maurer
+STAT 302 Homework 1 \hfill Aaron Maurer
 \vspace{2mm}
@@
             \item[(a)]
-                As we have proved, if the matix $X$ isn't singular, then the least squares solution from OLS is unique and indeed the optimal fit.
-Thus, if a $\beta$ acheives the minimum least squares, it must be the OLS coefficient from the regression.
-Calculating the sum of squares for a coefficient $\hat\beta ^{(-i)}$ on $\tilde Y^{(-i)}$, we get
-                \begin{align*}
-                    \|\tilde Y^{(-i)} - X\hat\beta^{(-i)}\|^2 &= (x_i\beta^{(-i)} - x_i\beta^{(-i)})^2 + \|Y^{(-i)} - X^{(-i)}\hat\beta ^{(-i)}\|^2 \\
-                    \|\tilde Y^{(-i)} - X\hat\beta^{(-i)}\|^2 &= \|Y^{(-i)} - X^{(-i)}\hat\beta^{(-i)}\|^2
-                \end{align*}
-                Since the sum on the right is the minimum possible over all $\beta$ (since the coefficient was fit by OLS on $Y^{(-i)}$), $\hat\beta ^{(-i)}$ minimizes the sum of squares for $\tilde Y^{(-i)}$.
-By our earlier reasoning, it thus must be the coefficient that would result from OLS run predicting $\tilde Y^{(-i)}$ in terms of $X$.
+                We can calculate the type I error as
+                \[P(x\in\{1,2\}|\theta=0) = P(x=1|\theta=0) + P(x=2|\theta=0) = .01\]
+                And the type II error as 
+                \[P(x\not\in\{1,2\}|\theta=1) = P(x=3|\theta=1) =.01\]
             \item[(b)]
-                The prediction of a linear model with coefficient $\hat\beta ^{(-i)}$ at point $x_i^T$ is $x_i^T\hat\beta ^{(-i)}$, so since that is the coefficient for regressing $\tilde Y^{(-i)}$ on $X$, $x_i^T\hat\beta ^{(-i)}$ will be the fit.
+                We can calculate the p-value for $x=1$ as 
+                \[ \sum_{x: P(X=x|H_0)\leq P(X=1|H_0)} \hspace{-4em} P(X=x|H_0) = P(X=1|H_0) + P(X=2|H_0) = .01\]
+                The Bayes Factor for $H_1$ vs. $H_2$ is
+                \[\frac{P(x=1|\theta=1)}{P(x=1|\theta=0)} =\frac{.0049}{.005} = .98\]
+                Qualitatively, this means that $x=1$ is very, very slightly evidence for $H_0$ over $H_1$.
+            \item[(c)]
+                We get a p-value for $x=2$ of 
+                \[ \sum_{x: P(X=x|H_0)\leq P(X=2|H_0)} \hspace{-4em} P(X=x|H_0) = P(X=1|H_0) + P(X=2|H_0) = .01\]
+                But this time the Bayes Factor $H_1$ vs. $H_2$ is
+                \[\frac{P(x=2|\theta=1)}{P(x=2|\theta=0)} =\frac{.9851}{.005} = 197.02\]
+                Which is overwhelming evidence for $H_1$
+            \item[(d)]
+                If we modify our example to \\
+                \begin{table}[h!]
+                \center
+                \begin{tabular}{c|c|c|c}
+                $x$ & 1 & 2 & 3 \\ \hline
+                $p(x|\theta=0)$ & 0.005 & 0.005 & 0.99 \\
+                $p(x|\theta=1)$ & 0.495 & 0.495 & 0.01 
+                \end{tabular}
+                \end{table}
 
+                Then the p-value for $x=1$ is still .01, but now the Bayes Factor is
+                \[\frac{P(x=1|\theta=1)}{P(x=1|\theta=0)} =\frac{.495}{.005} = 99\]
+                Which is a very large and in favor of $H_1$
         \end{itemize}
     \item[2.]
-        \begin{itemize}
-            \item[(a)]
+
+    \item[3.]
+        To get the posterior, we do the usual Bayesian calculation:
         \begin{align*}
-                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- S(x_i)^{(-i)}Y^{(-i)} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- \frac{1}{\sum_{j\neq i} S_{ij}} \sum_{j\neq i} y_j S_{ij} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- \frac{\hat f(x_i)-y_iS_{ii}(X)}{1-S_{ii}} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= \frac{y_i - y_i S_{ii} - \hat f (x) + y_iS_{ii}}{1- S_{ii}} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= \frac{y_i -\hat f (x) }{1- S_{ii}} \\
+            P(\mu|x) &=  P(\mu) P(x|\mu) \\
+                     &\propto \mu^{m-1}e^{-\lambda \mu} \prod_{i=1}^n \frac{\mu^{x_i}}{x_i!}e^{-\mu} \\
+                     &\propto \mu^{m-1+\sum_{i=1}^n x_i}e^{-(\lambda+n) \mu} \prod_{i=1}^n \frac{1}{x_i!} \\
         \end{align*}
-            \item[(b)]
-                Rather than refitting the regression $n$ times to estimate the leave one out cross validation estimate of error, this allows the regression to merely be fit once, saving a lot of costly $QR$ factorization or matrix inversion.
-            \item[(c)]
-                Since $1\geq1-S_{ii}\geq0$, we have 
-                \begin{align*}
-                    \vert y_i - \hat f(x_i) \vert &\leq \frac{\vert y_i - \hat f(x_i) \vert }{1-S_{ii}} \\
-                    \vert y_i - \hat f(x_i) \vert &\leq \left\vert\frac{y_i - \hat f(x_i) }{1-S_{ii}} \right\vert  \\
-                    \vert y_i - \hat f(x_i) \vert &\leq \left\vert y_i - \hat f^{(-i)}(x_i)\right\vert  
-                \end{align*}
-
-        \end{itemize}
-    \item[3.]
-        \begin{itemize}
-            \item[(a)]
-                Applying backwards elimination with a critical value of $.2$, we only remove the variable status. After that, all predictors have a p-value of less than $.2$, so we stop. \\
-                \smallskip
-
-                p-value at the point of exclusion: \\
-                \FloatBarrier
-                \input{hw8/hw8_3_a_excl} 
-                \FloatBarrier
-
-                Resulting model: \\
-                \FloatBarrier
-                \input{hw8/hw8_3_a_lm} 
-                \FloatBarrier
-
-            \item[(b)]
-                For each of the following three methods, I used an exhaustive subset search, which determined which set of a particular number of variables resulted in the smallest RSS.
-The result, which is a boolean indicating whether the predictor should be included at each number of predictors is below.
-Since this will uniformly be the best model of $n$ predictors, the remaining decision is which $n$ to use.
-It should be noted that the model in a is indeed the optimal model of $3$ predictors.
-\\
-                \FloatBarrier
-                \input{hw8/hw8_3_which} 
-                \FloatBarrier
-                The model with the lowest AIC also had $3$ predictors, and is thus the same as in a. The different AICs for the best model with $n$ predictors  below. \\
-                \FloatBarrier
-                \input{hw8/hw8_3_b} 
-                \FloatBarrier
-
-            \item[(c)]
-                The model with the highest adjusted $R^2$ is this same model with $3$ predictors. Optimal adjusted $R^2$ for different number of predictors are below. \\
-                \FloatBarrier
-                \input{hw8/hw8_3_c} 
-                \FloatBarrier
-
-            \item[(d)]
-                Going by Mallow's $C_p$, we once again get the same model.
-The smallest model that had a $C_p$ smaller than the number of parameters (one more than the number of predictors due to the intercept) was the model with 3 predictors (4 parameters): \\
-                \FloatBarrier
-                \input{hw8/hw8_3_d} 
-                \FloatBarrier
-        \end{itemize}
-    \item[4.]
-        The order of these operations doesn't make much difference.
-Point 21 is influential both before and after model selection, and the prefered model includes Air.Flow and Water.Temp, while excluding Acid.Conc, whether or not the outlier is included.
-\\
-        \smallskip
-        Our initial model is below. Acid.Conc looks ripe to be removed with a high $p-value$
-        \FloatBarrier
-        \input{hw8/hw8_4_lm.tex} 
-        \FloatBarrier
-        Indeed, when we test all subsets of the data, it is excluded from every model that has fewer than three predictors:
-        \FloatBarrier
-        \input{hw8/hw8_4_small_which} 
-        \FloatBarrier
-        The optimal subset, when going by adjusted $R^2$ is the model with just two predictors ($R^2$ for each optimal subset of a particular size below):
-        \FloatBarrier
-        \input{hw8/hw8_4_small_adjr2} 
-        \FloatBarrier
-        Choosing this as our model, we get the following model:
-        \FloatBarrier
-        \input{hw8/hw8_4_small_lm.tex} 
-        \FloatBarrier
-        Checking the distribution of studentized residuals for outliers, we see one potential culprit (the largest residual, in the top right, which heavily deviates from T-distribution).
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_small_rstudent} 
-        \end{center}
-        It is the 21st data point.
-With a studentized residual value of $-3.47$, it is just inside conservative Bonferroni critical value, for $\alpha=.5$, of $-3.53$.
-Thus, we don't quite reject it as an outlier.
-However, looking at the plot of Cook's distances, it is clear that the same point is highly influential.
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_small_cook}
-        \end{center}
-        Now, going back to the full model which included Acid.Conc, we look again for outliers and influential points.
-Shockingly, the same point is once again the only real canidate.
-The distribution of studentized residuals on this
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_rstudent} 
-        \end{center}
-        Testing it as an outlier as before, it has a studentized residual of $-3.33$ in this model, while the critical value is $-3.54$, so we don't reject it.
-However, its still influential, so we kick it out for that:
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_cook}
-        \end{center}
-        Preforming variable selection on the data set with this point excluded, we still find the optimal model has the two variables besides Acid.Conc. Optimal subsets of each size:
-        \FloatBarrier
-        \input{hw8/hw8_4_no_which} 
-        \FloatBarrier
-        Adjusted $R^2$ for these subsets, showing we still prefer the model with two predictors:
-        \FloatBarrier
-        \input{hw8/hw8_4_no_adjr2} 
-        \FloatBarrier
-    \item[5.]
-        \begin{itemize}
-            \item[(a)]
-                I have plotted the singular values below, gotten by taking the square roots of the absolute values of the eigenvalues of $XX^T$ (once $X$ had been centered).
-They quickly fall off to very small values, reflecting that most of the variation in the data set is explained by just a few underlying components.
-                \begin{center}
-                    \includegraphics[width=9cm]{hw8/5_a}
-                \end{center}
-            \item[(b)]
-                The components of the two largest singular vectors are plotted below.
-As would be expected, the components of the second vector are generally far smaller than those of the first.
-It appears that approximately half the time the second component has the same sign as the first, and half the time the opposite sign, as might be expected since they are orthogonal.
-The vaguely resemble piecies of orthogonal waves.
-                \begin{center}
-                    \includegraphics[width=9cm]{hw8/5_b}
-                \end{center}
-
-            \item[(c)] 
-            We see that the regression on the two largest princple components doesn't actually predict the data as well as the two best predictors from the original data set (for the same number of variables, the latter has higher $R^2$).
-While this latter model predicts almost as well as the model including all variables, the largest principle components are signficantly less predictive, though still reasonably predictive.
-The reason for this is that the principle components are a decomposition of variation within the predictors; however, the largest source of variation in these predictors isn't necessarily predictive of the outcome.
-For instance, if there is a lot of variation among a few variables that aren't predictive of the outcome, they may correspond to large principle components, but these principle components will likely be also unpredictive.
-On the other hand, since the principle components as a whole define a space equivalent to the orginal space defined by all the predictors, a model including all of them will be equally predictive to including all variables.
-\\
-            \smallskip
-            The regression on the two largest principle components ($R^2=.5469$):
-            \FloatBarrier
-            \input{hw8/hw8_5_pc_lm} 
-            \FloatBarrier
-            The regression on all variables ($R^2=.6866$):
-            \FloatBarrier
-            \input{hw8/hw8_5_all_lm} 
-            \FloatBarrier
-            The regression on age and height ($R^2=.6562$):
-            \FloatBarrier
-            \input{hw8/hw8_5_sm_lm} 
-            \FloatBarrier
-        \end{itemize}
-
 \end{itemize}

### 2015-04-08 stat302/hw1.tex
Done HW1

@@
     \item[2.]
+        I simulated, for one sided p-values for the coefficient from a simple linear regression, how often a value between .04 and .05 actually corresponded to a true positive.
+Each p-value was simulated where the true effect $\beta$ was 0 with probability $1-\pi$ and drawn from an exponential with coefficient $\lambda$ with probability $\pi$.
+This was preformed with all the usual assumptions for a linear regression in place.
+To get each p-value, I simulated as if 10 units received the treatment, 10 received the control, and that underlying variance was 1.
+This was done by drawing \(\hat\beta\sim N(\beta,\frac{1}{\sqrt{10}})\), drawing \(\hat s^2 \sim \frac{\chi^2_{18}}{18}\), and generating the usual p-value from the resulting T score.
+\\
+        \vspace{0mm} \\
+        For a number of combinations of $\lambda$ and $\pi$, I simulated $100,000$ such p-values, and calculated the portion of such values between .04 and .05 for which the true $\beta$ was positive.
+I've include the results below (remember \(\E(\beta)=\frac{1}{\lambda}\)):
+        \FloatBarrier
+        \input{hw1/pr2.tex}
+        \FloatBarrier
+        Though I employ a somewhat different simulation, the main result is the Same as Selke et al; a p-value near .05 is not consistently strong evidence for the alternative hypothesis.
+Indeed, we also see that it provides the best evidence for the alternative with moderate levels of power, since if power is too high one rarely sees p-values near .05 for true positives.
+\\
+
+        {\bf Code:}
+        \lstinputlisting[firstline=22,lastline=50]{hw1.R}
 
@@
         \begin{align*}
-            P(\mu|x) &=  P(\mu) P(x|\mu) \\
+            P(\mu|x) &\propto P(\mu) P(x_1,...,x_n|\mu) \\
                      &\propto \mu^{m-1}e^{-\lambda \mu} \prod_{i=1}^n \frac{\mu^{x_i}}{x_i!}e^{-\mu} \\
                      &\propto \mu^{m-1+\sum_{i=1}^n x_i}e^{-(\lambda+n) \mu} \prod_{i=1}^n \frac{1}{x_i!} \\
+                     &\sim Gamma(m+\sum_{i=1}^n x_i,\lambda+n)
         \end{align*}
+
+    \item[4.]
+        To get the posterior, we do the usual Bayesian calculation, this time substituting in \(\tau=\frac{1}{\sigma^2}\):
+        \begin{align*}
+            P(\tau|x) &\propto P(\tau) P(x|\tau) \\
+                      &\propto e^{-2\tau} \sqrt{\tau}e^{-\frac{1}{2}\tau x^2} \\
+                      &\propto e^{-2\tau} \sqrt{\tau}e^{-\frac{1}{2}\tau x^2} \\
+                      &\propto \tau^\frac{1}{2}e^{-(2+\frac{1}{2}x^2)\tau} \\
+                      &\sim Gamma\left(\frac{3}{2},2+\frac{1}{2}x^2\right) 
+        \end{align*}
+
+    \item[5.]
+        Overall, using this method, I had an error rate of $22.3\%$.
+The main issue I ran into with the method as described was that a number of allele/population combinations didn't appear in the training set.
+This isn't necessarily a problem, but with these 0 likelihoods included, some individuals in the test data set had 0 likelihoods of each population.
+I remedied this by adding one to each of the allele/locus counts in the test data set before calculating frequencies.
+This corresponds to beginning with a weak uniform prior on the allele/locus frequencies.
+Also, when an individual was missing allele data at a particular locus, I omitted that locus from likelihood calculation (in the program I just multiplied each likelihood by 1).
+
+        {\bf Code:}
+        \lstinputlisting[firstline=60,lastline=108]{hw1.R}
+
 \end{itemize}

### 2015-04-09 stat347/hw1.tex

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT347 Homework 1 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[(a)] One would usually want to measure covariates before assigning treatment so that treatments could be assigned on the basis of covariates. By assigning each treatment to a variety of values of covariates (for instance, each unit in a block receiving different treatment), it becomes less likely the effect of the covariate is confounded with the effect of the treatment.
            \item[(b)] Let $v$ be an additional unit such that, where $S$ is the set of all units, \(\forall v' \in S\backslash \{v\}, R(v,v')\neq1\). Since we can not apply the given logic to $v$, it need not be the case \(R(v,v)=1\), so the relation is not necessarily reflexive. 
            \item[(c)] Site, race, sex, and martial status are unordered qualitative factors, since they each take on a finite number of levels which admit not obvious ordering. I would classify age as a quantitative covariate, since it can be measured as a continuous (as opposed to discrete) value; in practice however, it can also be treated as an ordered qualitative factor, since it is usually recorded in discrete year values.
            \item[(d)]
                \begin{itemize}
                    \item[(i)]
                        For the \(y\sim A + {\bf 1}\) model, the model matrix is
                        \[ \left[ \begin{array}{cccc} 1 & 0 & 0 & 1 \\
                                                      1 & 0 & 1 & 0 \\
                                                      1 & 0 & 1 & 0 \\
                                                      1 & 0 & 0 & 1 \\
                                                      1 & 1 & 0 & 0 \\
                                                      1 & 0 & 1 & 0 \\
                                                      1 & 0 & 0 & 1 \\
                                                      1 & 1 & 0 & 0 \\
                                                      1 & 1 & 0 & 0 \\
                                    \end{array} \right] \]
                                    The second column, corresponding to level 1 of A, would be omitted under the default in R (treatment constraint). For the \(y\sim A - {\bf 1}\) model, the model matrix is
                        \[ \left[ \begin{array}{cccc} -1 & 0 & 0 & 1 \\
                                                      -1 & 0 & 1 & 0 \\
                                                      -1 & 0 & 1 & 0 \\
                                                      -1 & 0 & 0 & 1 \\
                                                      -1 & 1 & 0 & 0 \\
                                                      -1 & 0 & 1 & 0 \\
                                                      -1 & 0 & 0 & 1 \\
                                                      -1 & 1 & 0 & 0 \\
                                                      -1 & 1 & 0 & 0 \\
                                    \end{array} \right] \]
                        Once again, the second column would be omitted. These two models are equivalent in the sense they would yield the same predictions. This is because they define the same subspace since \({\bf -1} \in span({\bf 1})\). For \(y \sim A + B\), the model matrix would be
                        \[ \left[ \begin{array}{cccccc}
                                                      0 & 0 & 1 & 1 & 0 & 0 \\
                                                      0 & 1 & 0 & 1 & 0 & 0 \\
                                                      0 & 1 & 0 & 1 & 0 & 0 \\
                                                      0 & 0 & 1 & 0 & 1 & 0 \\
                                                      1 & 0 & 0 & 0 & 1 & 0 \\
                                                      0 & 1 & 0 & 0 & 1 & 0 \\
                                                      0 & 0 & 1 & 0 & 0 & 1 \\
                                                      1 & 0 & 0 & 0 & 0 & 1 \\
                                                      1 & 0 & 0 & 0 & 0 & 1 \\
                                    \end{array} \right] \]
                        Here, the 4th column, corresponding to the $1$ level of B would be omitted, since it is in the span of the rest of the columns.
                    \item[(ii)]
                        If we treat A as a quantitative covariate, then for the \(y\sim A + {\bf 1}\) model, the model matrix is
                        \[ \left[ \begin{array}{cc} 1 & 3 \\
                                                    1 & 2 \\
                                                    1 & 2 \\
                                                    1 & 3 \\
                                                    1 & 1 \\
                                                    1 & 2 \\
                                                    1 & 3 \\
                                                    1 & 1 \\
                                                    1 & 1 \\
                                    \end{array} \right]
                        \]
                        With no columns omitted in R. For the \(y\sim A - {\bf 1}\) model, the model matrix is
                        \[ \left[ \begin{array}{cc}   -1 & 3 \\
                                                      -1 & 2 \\
                                                      -1 & 2 \\
                                                      -1 & 3 \\
                                                      -1 & 1 \\
                                                      -1 & 2 \\
                                                      -1 & 3 \\
                                                      -1 & 1 \\
                                                      -1 & 1 \\
                                    \end{array} \right]
                        \]
                        With no columns omitted in R either. This model still defines an equivalent linear space to the \(y\sim A + {\bf 1}\) model, so would yield identical predictions. Finally, For \(y \sim A + B\), the model matrix would be
                        \[ \left[ \begin{array}{cccc}
                                                      3 & 1 & 0 & 0 \\
                                                      2 & 1 & 0 & 0 \\
                                                      2 & 1 & 0 & 0 \\
                                                      3 & 0 & 1 & 0 \\
                                                      1 & 0 & 1 & 0 \\
                                                      2 & 0 & 1 & 0 \\
                                                      3 & 0 & 0 & 1 \\
                                                      1 & 0 & 0 & 1 \\
                                                      1 & 0 & 0 & 1 \\
                                    \end{array} \right] 
                        \]
                        Which would have no columns omitted.
                    \item[(iii)]
                        For each of these subspaces, the dimension is the same as the dimension of the model matrix as created above. The squared length of the orthogonal projection is, when the model matrix is $X$, \(\|X(X'X)^{-1}X'y\|^2\) and the squared length of the orthogonal complement is \(\|(I-X(X'X)^{-1}X')y\|^2\). This yields:
                        \FloatBarrier
                        \input{hw1/pr1_d_iii}
                        \FloatBarrier
                    \item[(iv)] I would use the sum of squares from A. Specifically, I would compare the marginal difference in sum of squares between the projection onto the A space and onto the ${\bf 1}$ space, and compare this to the sum of squares in the complement space to A. It appears B has little if any effect, so it can probably be safely ignored for this purpose. If there is no treatment effect, these two values should both be $\chi^2$ variables with the respective degrees of freedom, so we can use them to construct a F test:
                        \[ \frac{\frac{330.52-258.89}{3-1}}{\frac{13.99}{6}} = 15.361\]
                        Checking against a $F_{2,6}$ distribution, this corresponds to a p-value of $.0044$.
                \end{itemize}
        \end{itemize}
    \item[2.]
        \begin{itemize}
            \item[(i)]
                This table can be gotten directly from the anova command in R
                \FloatBarrier
                \input{hw1/pr2_i}
                \FloatBarrier
            \item[(ii)]
                This table is a bit more complicated. I generated most of the output from a series of anova commands (Listing the spaces in increasing size so as to get the right quotient spaces), but then recalculated the F and p-values based on the overall residuals from the first model:
                \FloatBarrier
                \input{hw1/pr2_ii}
                \FloatBarrier
            \item[(iii)] This table strongly indicates that temperature an effect on fungal growth, since temperature as a categorical variable has a very low p-value on it. However, lower order polynomials seem to be sufficient to capture the effect; while the linear fit is poor, adding both a 2nd and 3rd degree polynomial seems to cause statistically significant improved fits. After this though, greater complexity does not seem to cause statistically significant improvement, indicating a 3rd degree polynomial is optimal.
            \item[(iv)]
                I've plotted the cubic fit below:
                \FloatBarrier
                \begin{center}
                    \includegraphics[width=9cm]{hw1/pr2_iv}
                \end{center}
                \FloatBarrier
                Based off this fit, one would predict growth rate will be maximized by a temperature of approximately 71.75.

        \end{itemize}
    \item[3.]
        \begin{itemize}
            \item[(i)]
                For each conservative matrix $V$, it must be the case $\forall i,j$ that \(V(i_0,i_1)+V(i_1,i_0)=0\Longrightarrow V(i_0,i_1)=-V(i_1,i_0) \). This means that $V$ is skew-symmetric. On the other hand, for 
                \[ W = \left[ \begin{array}{ccc} 0 & -4 & -1 \\ 4 & 0 & -2 \\ 1 & 2 & 0 \end{array} \right] \]  
                We have that \(W(3,1)+W(1,2)+W(2,3)= 1 -4 -2 \neq 0\), so not all skew-symmetric matrices are conservative. \\
                \vspace{5mm} \\
                If a matrix $V$ is additive, for a path of length $m$ where $i_m=i_0$,
                \[ \sum_{j=0}^{m-1} V(i_j,i_{j+1}) = \alpha_{i_0} + \sum_{j=1}^{m-1}(-\alpha_{i_j}+\alpha_{i_j}) - \alpha_{i_0} = 0 \]
                So $V$ is conservative. We may also conclude the converse by the following logic on any conservative matrix $V$:
                \begin{align*}
                    0 &= V(u_1,u_2) + V(u_2,u_3) + V(u_3,u_4) +V(u_4,u_1) \\
                    \Longrightarrow -V(u_4,u_1) - V(u_2,u_3) &= V(u_1,u_2) + V(u_3,u_4) \\
                    \Longrightarrow \;\;\; V(u_1,u_4) + V(u_3,u_2) &= V(u_1,u_2) + V(u_3,u_4) \\
                \end{align*}
                This is only possible if, for all $u_i, u_j$, $V(u_i,u_j)=f(u_i)+g(u_j)$ for some functions $f, g$. Since 
                \begin{align*}
                    V(u_1,u_3) &= V(u_1,u_2) + V(u_2,u_3) \\
                    f(u_1) + g(u_3) &= f(u_1) + g(u_2) + f(u_2) + g(u_3) \\
                    f(u_2) &= -g(u_2) \\
                \end{align*}
                Proving that $V(u_i,u_j) = f(u_i) - f(u_j)$ and thus that $V$ is additive. \\
                \vspace{5mm} \\
                Finally, the space of $6\times 6$ conservative matrices had dimension $5$. This is because the matrix is minimally defined by \(\{V(1,2),V(1,3),V(1,4),V(1,5),V(1,6)\}\). All the remaining entries can be calculated as a linear combination of these five values: \(V(u,j) = V(u,1)+V(1,j) = -V(1,u)+V(1,j)\). No fewer values will suffice for this purpose. 
            \item[ii]
                We can set this up as a typical least squares problem by transforming the $k\times k$ matrix $Y$ into a $k^2\times 1$ vector $y$. Then, if $R$ is the usual matrix of 1s and 0s for the row as a categorical variable, and $C$ is the same for the column, our model matrix is $X=R-C$. This gives us \(E(y)=X\alpha\). After omitting one of the columns of $X$ due to collinearity, we can then produce the normal least squares estimate $\alpha=(X'X)^{-1}X'y$. \(\alpha_i-\alpha_j = (\alpha_i+c) - (\alpha_j+c)\), so \(\{\alpha_1,\alpha_2,\alpha_3,\alpha_4,\alpha_5\}\) and \(\{\alpha_1+c,\alpha_2+c,\alpha_3+c,\alpha_4+c,\alpha_5+c\}\) are equivalent parameters, demonstrating that one parameter needs to be omitted.\\
                \vspace{5mm} \\
                When we preform this procedure on the data, we get these predictions for the $\alpha$ by electrolyte:
                \FloatBarrier
                \input{hw1/pr3_ii}
                \FloatBarrier
                As noted, the fifth value is a free parameter and was omitted.
            \item[iii]
                These two models are one just including the $\alpha$, and one where the $\alpha$ are interacted with electrolyte. Running these two models and comparing them with ANOVA, we get:
                \FloatBarrier
                \input{hw1/pr3_iii}
                \FloatBarrier
                By the interaction, we don't see quite enough of an improvement to push past the $.05$ p-value threshold, though it is a borderline case.
            \item[iv]
                Our data set only has factor covariates, so no monotone transformation will have an effect on the predictions, and non-monotone transformation wouldn't make sense. If there was evidence of a lack of normality or non-constant variance, a transformation might improve our statistical inference, but checking various residual plots leaves little indication that this is necessary.
        \end{itemize}
    \item[5.]
        This can be formulated as a linear model after transformation: 
        \[ log(Y) ~ \alpha - \beta_T t\] 
        If we are trying to model the expected half-life $h$ for a given sample under this model, we can formulate the problem as
        \begin{align*}
            log(.5) &= \hat\beta_Th \\
            \frac{1}{h} &= \frac{\hat\beta_T}{log(.5)}\\
        \end{align*}
        The value on the right is t-distributed under the assumptions of a normal model, so we can generate a $95\%$ confidence interval by $\hat\beta_T\pm t_{n-p}^{.025}se(\hat\beta)$. Then, we can take the inverse of each boundary to get a confidence interval on the $h$. We must keep in mind though that any value that is $0$ or negative before taking the inverse corresponds to a lack of decay and asorbic acid levels not actually falling. This yields the following confidence intervals for half-life (in weeks):
        \[ \begin{array}{cccc} Temperature & Lower Bound & Estimate & Upper Bound \\
                              \hline  
                              0           & 67.63       & 948.55   & \infty      \\
                              10          & 20.70       & 28.92    &  47.99       \\
                              20          & 4.89        & 5.24     & 5.65        \\
                      \end{array} \]
        It is worth noting that this does not include any prediction error. However, that error's contribution would be relatively small. This model accounts for for the variance among replicates; that variance contributes to the variance in the mean which we are fitting.

\end{itemize}
\end{document}

### 2015-04-14 stat302/hw1.R
Fixed small naming error

@@ -73,7 +73,7 @@ if (problem5) {
     train_freq <- list()
     for (i in 1:12) {
         train_freq[[i]] <- as.data.frame.matrix(compute_freq(train,i))
-        train_freq[[i]]$allele <- as.factor(rownames(test_freq[[i]]))
+        train_freq[[i]]$allele <- as.factor(rownames(train_freq[[i]]))
     }
 
     # Set uniform prior

### 2015-04-14 stat374/final_project.tex

@@
 
-\title{Opening a Cold Case \\ {\large Investigating how Temperature Affects The Rate of Robberies}}
-\date{December 2, 2014}
+\title{Opening a Cold Case \\ {\large Investigating Temperature's Effect on The Rate of Robberies}}
+\date{December 4, 2014}
 \author{Aaron Maurer}
@@
 
-With the onset of winter in Chicago, the temperature drops, and residents must concern themselves with a range of new dangers.
-Roads can be icy and dangerous.
-Heavy snow can collapse power lines and buildings.
-And frostbite becomes a concern on exposed skin.
-However, it may be the case that cold weather mitigates other dangers.
-When its frigid out, do criminals stay inside as well?
+There have been a number of studies on how temperature affects the rate of violent crimes, in particular assault.
+The consensus seems to be that rate of these crimes rises in hot weather, with the inclination being that it also drops off in cold weather.
+Both these changes are thought to occur due to the effect of temperature on human emotion\footnote{Andrea Anderson, ``Links Between Crime \& Chilly Weather is Complex, Controversial", \textit{Scientific American}, January 1st, 2014, Accessed December 4th, 2014.
+http://www.huffingtonpost.com/2014/01/01/weather-crime-temperatures\_n\_4518940.html}.
+However, I was curious what relationship exists between a more practical and presumably less emotional crime such as robbery and the temperature.
 \par
 
-To partially investigate this question I have endeavored to quantify how the rate at which robberies occur in the city of Chicago varies with temperature.
+To investigate this question I have endeavored to quantify how the rate at which robberies occur in the city of Chicago varies with temperature.
 Robbery, defined as taking property from a person, without their consent, by force or threat of force\footnote{As opposed to theft, which is the taking property without consent but not necessarily by violent means, and burglary, which is breaking into a structure so as to commit a crime (whether or not a crime such as theft is committed)}, includes what is probably the most common outdoor crime, mugging.
-It would seem natural then that the rate at which robberies occurred would be sensitive to the weather.
-The this rate is distinct from one's risk of being robbed should they go outside, since it may be the case crime drops when fewer people are outside to be robbed.
-However, a better understanding of the former should also provide intuition for the latter question.
+It would seem natural then that the rate at which these occurred would be sensitive to the weather.
+As well, investigating this relationship will give insight into when one is at the greatest risk for being robbed should they go out, though the number of people out and about also factors in.
 \par
@@
 \section{Data} 
-The Chicago city government publishes a data set of which includes every crime reported to the police in the city, going back to 2001\footnote{City of Chicago.
+The Chicago city government publishes a data set which includes every crime reported to the police in the city, going back to 2001\footnote{City of Chicago.
 (2014).
@@
 Retrieved from https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2}.
-I made use of the subset of this data which had a date recorded for the crime during 2011, 2012, or 2013, and which had a primary description of "Robbery", per the Illinois Uniform Crime Reporting Code.
+I made use of the observations from this data which had a date recorded for the crime during 2011, 2012, or 2013, and which had a primary description of "Robbery", per the Illinois Uniform Crime Reporting Code.
 These crimes include both successful and attempted armed robbery, unarmed robbery, and vehicular hijacking.
@@
 This data is a series of meteorological data, including temperature, accumulated precipitation, humidity, and wind speed, almost always measured once during a particular hour at 51 minutes after that hour.
-When there were multiple entries for an hour, I chose the first at 51 minutes which had the temperature recorded, or the first chronologically which had the temperature recorded if that was missing, discarding the rest.
+When there were multiple entries for an hour, I chose the first at 51 minutes which had the temperature recorded, or the first chronologically which didn't have temperature missing, discarding the rest.
 This was taken as the temperature for the city for a particular day and hour, and was matched with the hourly robbery counts.
 The small number ($\approx 10$) of hours for which there was no recorded temperature from O'Hare were excluded.
-The end result was $26,263$ hourly observations of temperature and reported robbery count, covering almost all of 2011-2013.
+The end result was $26,263$ hourly observations of temperature and reported robbery count, covering almost all of 2011-2013\footnote{My R script is available on github: https://github.com/ajmaurer/Chicago-Course-Work/blob/master/stat374/final\_project.R}.
 \par
-Examining this data, it is obvious that the most important feature in determining the number of robberies is the time of day, to a lesser extent, time of week.
+Examining this data, it is obvious that the most important feature in determining the number of robberies is the time of day, and to a lesser extent, time of week.
 Figure 1 speaks to this effect, displaying an average by hour of the day for different temperature ranges and workday versus holiday/weekend.
-The mean number of robberies peaks between 6pm and 3am, possibly twice in that span, with its minimum around 6am, with the peak mean 4 to 5 times as high as the trough.
-This corresponds roughly to typical people's schedules; there are the most robberies during the period where people have gotten off of work for the evening, and the least when the fewest people are awake early in the morning.
-During the weekend, the trough and peak tend to move later, as people are up later or sleep in.
+The mean number of robberies peaked between 6pm and 3am, occasionally twice in that span.
+The minimum was consistently around 6am, with the mean at the peak 4 to 5 times as high as at the trough.
+This corresponds roughly to typical people's schedules; there were the most robberies during the period where people have gotten off of work for the evening, and the least when the fewest people are awake early in the morning.
+During the weekend, the trough and peak tended to be later, as people were up later or sleep in.
 \par
@@
 \end{figure}
-The effect of temperature seems to be less than this, and also varies with the time of day.
-During the middle of the day (8am to 3pm or so), it appears that the effect of temperature is at most minimal.
-One of the most interesting features is that in the evening, there actually appear to be fewer robberies between 6pm and 7pm when it is warm, with a much higher rate before and after.
-This is likely due to the changes in people's schedules that occurs when days are longer and the weather is warm.
-However, the strongest effect seems to be that there are far fewer robberies in the late evening and early morning when its cold.
+The effect of temperature seems to be smaller than this, and also varies with the time of day.
+During the middle of the day (8am to 3pm or so), the means were similar for all temperatures.
+However, there were fewer robberies in the late evening and early morning when it was cold.
+One of the most interesting features is that in the evening, there were fewer robberies between 6pm and 7pm when it was warm, with a much higher rate before and after.
+This is likely at least partially due to the changes in people's schedules that occurs around summer, when it is both warmer and light later.
 \par
@@
 Looking at the distribution for a sample of hours in Figure 2, we once again see the effect of time of day and workday versus weekend/holiday.
-However, we also see a data set which seems to have, as one would hope for with count data, a Poisson distribution.
+However, we also see a data set which seems to have, as one would hope for with count data, a distribution which is approximately Poisson.
 
@@
 This observation points to the obvious parametric model to fit to the data, a Poisson regression model\footnote{A generalized linear model with a Poisson distribution and log link fit by maximum likelihood}.
-This is a natural choice, and generally considered the first parametric model to try for count data, if not the default.
+This is a natural choice, and  considered the first parametric model to try for count data, if not the default.
 Its assumptions are several fold.
-First, it assumes that data is Poisson distributed around a particular mean, which arises when counting the sum of infrequent and independent occurrences in which a large number of these events could, in theory, occur.
+First, it assumes that data is Poisson distributed around a particular mean, which arises when counting the sum of infrequent and independent occurrences of which a large number of these events could, in theory, occur.
 This assumption seems fairly safe; robberies are infrequent, and though one criminal could be responsible for multiple robberies, this effect overall can be expected to be trivial in such a large city.
@@
     \item Hour of the day as a factor 
-    \item Three hour buckets as a factor interacted with workday as a factor
-    \item Three hour buckets as a factor interacted with the log of the sum temperature plus $50$.
+    \item Buckets of three hour periods as a factor interacted with workday as a factor
+    \item Buckets of three hour periods as a factor interacted with the log of the sum $50$ plus temperature.
 \end{itemize}
 
-Implicit in this set up is that the effect of temperature within a particular hour is assumed to be monotone. \par
+\noindent Implicit in this set up is that the effect of temperature within a particular hour is assumed to be monotone. \par
+
+The estimates of the coefficients themselves are not particularly elucidating; the overall effect is best displayed graphically, such as through the heat maps in Figure 3.
+Here, we see the model capturing several of the features we noticed earlier.
+The predicted robbery rate is quite low in the late evening and early morning when it is cold, and rises quickly with the temperature, especially on the weekend.
+A bit later (6am or so), the crime rate is low, and throughout the middle of the day it only rises slightly with temperature.
+The early evening is the only area that seems outright strange; high values are predicted at low temperatures, and drops off somewhat as the temperature increases.
+This is questionable, but possibly an artifact of the assumed monotone relationship between temperature and robbery rate paired with the small dip between higher and lower temperatures in this range.
+\par
 \begin{figure}[h]
@@
 
-The estimates of the coefficients themselves are not particularly elucidating; the overall effect is best displayed graphically, such as through the heat maps in Figure 3.
-Here, we see the model capturing several of the features we noticed earlier.
-The predicted robbery rate is quite low in the late evening and early morning when it is cold, and rises quickly with the temperature, especially on the weekend.
-A bit later (6am or so), the crime rate is low, and throughout the middle of the day it only rises slightly with temperature.
-The early evening is the only area that seems outright strange; high values are predicted at low temperatures, and drops off somewhat as the temperature increases.
-This questionable, but is likely an artifact of the assumed monotone relationship between temperature and robbery rate paired with the small dip between higher and lower temperatures in this range.
-\par
-This is displayed again for different three hours individually in Figure 4 along with confidence intervals.
+The effect of temperature is displayed again individually for 2am, 10am, and 6pm in Figure 4 along with $95\%$ confidence intervals.
 Here, we see again the patterns listed above, but also that the model predicts these big swings in robbery rate due to temperature with a great deal of certainty.
@@
 
+\FloatBarrier
 \begin{figure}[h]
@@
 In this case, should the rest of the assumptions hold, a negative binomial regression can be used, which can be derived as a certain sum of Poisson distributions with varying means, as could be the case should the mean not be perfectly predicted by the predictors.
-This makes the Poisson model a particular example of a negative binomial, and a formal test can be used to decide if the additional parameter of a negative binomial is needed.
-Such a test might well have indicated that I should have used a negative binomial regression here, but given time constraints and a lack of familiarity, I simply went with the Poisson model.
+This makes the negative binomial model a generilization of the Poisson model, and a formal test can be used to decide if the additional parameter of a negative binomial is needed.
+Such a test might well have indicated that I should have used a negative binomial regression here, but its unlikely the resulting predictions would have been wildly different.
 \par
-Additionally, as with nonparametric models, a better tool for evaluating the fit of a particular specification of the model may have been cross validation rather than Akaike Information Criteria, since it gives a more direct estimate of over and under fitting.
+Additionally, as with nonparametric models, a better tool for evaluating the fit of a particular specification of the model may have been cross validation rather than Akaike Information Criteria, since it gives a more direct estimate of the suitability of the fit.
 
 \section{Nonparametric Model} 
-To model this data with a nonparametric model, I ran separate local linear regressions on workdays and weekends/holidays, with each predicting the log of the sum number of robberies plus one with hour and temperature simultaneously.
+To model this data with a nonparametric model, I ran separate local linear regressions on workdays and non-workdays, with each predicting the log of the sum one plus number of robberies in terms of hour and temperature simultaneously.
 As with the Poisson model, I fit on an $80\%$ training subset.
-In fitting it, I used a Gaussian kernel, individual bandwidths for each point such that a constant portion of the total data was closer to the point than the bandwidth, and a relative scaling factor for the second variable (this effectively stretches or shrinks the kernel in one direction).
-These two values were chosen from a grid of values based on which resulting model had the lowest cross validation score.
-As well, the hour span was expanded to include the previous and next day's data, so there should be no boundary bias towards the beginning and end of days.\par
+In fitting the model, I used a Gaussian kernel, individual bandwidths for each point based on a global nearest neighbor fraction\footnote{The bandwidth is set for each point such that the fraction of the total data less than the bandwidth's distance from the point is the desired nearest neighbor fraction}, and a relative scaling factor for the second variable\footnote{This factor is used to scale the second variable with respect to the first for determining the distance between two points for the weighting function}.
+These two values were chosen from a grid of values based on which resulting model had the lowest leave one out cross validation score.
+As well, the hour span was expanded to include the previous and next day's data, so there should be no boundary bias towards the beginning and end of the day.\par
 The reasoning and justification for these choices are as such:
@@
     \item The kernel doesn't make a great deal of difference, so the choice was arbitrary. 
-    \item Since I am particularly interested in how temperatures on the extreme edge of my data effect the robbery rate.
-This necessitates a need a model with low boundary bias along the temperature range, making a local linear fit superior to a simpler kernel regression.
+    \item To study the effects of temperature at the extreme end of their range, I need a model with low boundary bias, making a local linear fit superior to a simpler kernel regression.
 Higher polynomial fits on the other hand make it increasingly difficult to fit the model well.
-    \item Predicting log robberies nicely constrains the prediction to positive counts, which is what it should be in the real world.
+    \item Predicting log one plus robberies nicely constrains the prediction to positive counts, which is what it should be in the real world.
     \item As the initial plots have shown, there seem to be fairly distinct trends between workdays and not workdays, so two separate models seemed more appropriate than one. 
-    \item Also as observed earlier, the relationship between temperature and robberies doesn't appear constant by time of day, so it would likely be insufficient to model it separately, as would be the result from a general additive model that included separate local linear models for hour and temperature.
-In this case, there is a sufficient amount of data to still have a reasonable degree of accuracy, so the curse of dimensionality is not prohibitive.
+    \item The relationship between temperature and robberies appears to vary by time of day, which would not have been captured by a generalized additive model which modeled the two variables without interaction.
+Since there is sufficient data to achieve satisfactory narrow bandwidth even modeling the two variables together with one local linear model, this choice was preferable.
     \item The observations are not equally spaced out over their range, so one bandwidth for all the points wouldn't smooth enough on the extremes or too much in the middle.
-Instead, by dictating the bandwidth such that it defined a neighborhood with a fixed number of neighbors, the amount of smoothing was comparable across the data points.
+Instead, by using a nearest neighbor fraction, the amount of smoothing was comparable across the data points.
     \item Including a scaling factor for the second variable is essential, since the units of the two variables aren't comparable.
 \end{itemize} \par
-With this model, I am only really relying on two assumptions.
-First, the number or robberies as a particular time and temperature is related to the number of robberies at close time and temperatures, which is obvious.
-Second, I am assuming sufficient smoothness to achieve convergence of the estimates.
+With this model, two assumptions are being made.
+First, the number or robberies as a one time and temperature is correlated with the number of robberies at similar times and temperatures, which is obvious.
+Second, that there sufficient smoothness in the underlying function to achieve convergence of the estimates.
+Having looked at simpler plots, this also seems likely.
 \par
@@
 The predictions of the models can be seen in the contour plots of Figure 5.
-It is important to note that with less data, bandwidths for the weekend/holiday model were chosen to include a larger portion of the data, and also that the scale was such that this model averaged more strongly over temperature than hour as compared to the workday model.
+It is important to note that with less data, bandwidths for the weekend/holiday model were chosen to include a larger portion of the data around each point, and also that the scale was such that this model averaged more strongly over temperature than hour as compared to the workday model.
 This model seems to very clearly portray the patterns we have seen before.
-Peak crime on workdays is highest in the evening, with the main peak later and a second, smaller and earlier, peak developing when the temperature is higher.
+Peak crime on workdays is highest in the evening, with the main peak moving later and a second smaller, earlier peak developing with hot temperatures.
 Also, the afternoon and evening crime rate does fall off as the temperature increases beyond a point.
 On non-workdays, the peak is in the earliest hours of the morning, and keeps increasing with the temperature.
-However, the prediction for the highest temperatures should be taken with a grain of salt; their is little if any data where temperatures actually reach that high at that hour of the day.
+However, the prediction for the highest temperatures in the early morning should be taken with a grain of salt; their is little if any data where temperatures actually reach that high at that hour of the day.
 \FloatBarrier
@@
 It should be noted that the confidence intervals are based on local estimates of variance, necessitated by the variable bandwidth.
-This model is quite certain that the peak crime rate at $6$pm on workdays occurs when the temperature is between 30 and 50 degrees, and falls off afterwards, as opposed to the monotonic relationship predicted by the Poisson model.
+This model is quite certain that the peak crime rate at $6$pm on workdays occurs when the temperature is between 30 and 50 degrees, and falls off on either side, as opposed to the monotonic relationship predicted by the Poisson model.
 It also has a large amount of confidence there is a true increase in crime in the early morning, though as I noted, at high temperatures there is much more uncertainty.
@@
 However, what could be quite a good alternate approach is to not have aggregated robbery counts, and fit a density estimate to the data.
-This could be scaled by how often different temperature and hour combinations occurred to give relative rates of robbery.
+This could be scaled by how often different temperature and hour combinations occurred to give relative rates of robbery at a given temperature and time (per the definition of conditional probability).
 This approach has the advantage of preserving more information about the time a crime occurs, but is computationally more intensive (with more data points), and doesn't easily admit a comparable parametric model.
@@
 \par
-In terms of interpretation, both are pretty convincingly show that the rate of robberies varies with temperature, generally increasing with it.
-Causation is quite unclear though; temperature is highly correlated with a number of other seasonal factors.
-One could reasonably expect the rate of robberies to vary with school holidays, daylight hours, or just the time of the year irrespective of season.
-By including all these factors in a model, it would likely be highly predictive, but the high collinearity can be expected to make it hard to attribute changes to just one factor, modeled parametrically or not.
-Appropriately modeling these effects would be the most important extension of the current model.
-\par
-There are a number of other areas where this model could be expanded though.
-There is a great wealth of geographic data in the Chicago crime database, and there is likely a great deal of heterogeneity in the baseline rate of robberies by geography, as well as the differences in the effect of weather on that rate.
-Different types of crime could also be investigated.
-Different attributes of the weather may also play a roll.
-I initially investigated the effect of precipitation on the robbery rate.
-The presence of precipitation, as showed in Figure 7, does seem to coincide with a small reduction in robberies, at least in during the late evening and early morning.
-This effect appears to vary a lot with the amount of precipitation, and there were relatively little data from when there was enough precipitation to have a meaningful effect.
-With the Poisson model, every version I tried which included precipitation in some form had an inferior AIC, so ultimately I decided not to model it, at least for now.
-Finally, there is slight asymmetry between the early morning and late evening predictions due to how the weekends are defined.
-Friday evening is likely more akin to a Saturday night, with more robberies, pushing up the late evening average of workdays as compared to the early mornings, while Sunday evening likely has the opposite effect on the Weekends/Holidays.
-A modified delineation between the two periods may have done away with this, improving the fit.
-There may be further clustering that could be preformed on the days as well.
-All these areas present opportunity for further refinement.
-
 \begin{figure}[h]
@@
 \end{figure}
+
+In terms of interpretation, both fairly convincingly show that the rate of robberies varies with temperature, generally increasing with it.
+Causation is quite unclear though; temperature is highly correlated with a number of other seasonal factors; one could reasonably expect the rate of robberies to vary with school holidays, daylight hours, or just the time of the year irrespective of season.
+By including all these factors in a model, it would likely be highly predictive, but the high collinearity can be expected to make it hard to attribute changes to just one factor, modeled parametrically or not.
+As well, though three years is a relatively short period for policing and social changes, there is a well known secular decreasing trend of urban crime in America which was not accounted for.
+Should this have synced with the relative severity of winters, it may have skewed results.
+Appropriately modeling these effects would be the most important extension of the current model.
+\par
+There are a number of other areas where this model could be expanded though.
+Crime rates are quite heterogeneous throughout the city, in a fashion that may vary with weather, and there is a great wealth of geographic data in the Chicago crime database to model this.
+Different types of crime could also be investigated.
+Different attributes of the weather may also play a roll; I initially investigated the effect of precipitation on the robbery rate.
+this seemed to show that presence of precipitation, as seen in Figure 7, does seem to coincide with a small reduction in robberies, at least in during the late evening and early morning.
+This effect appears to vary a lot with the amount of precipitation, and there was relatively little data from hours with enough precipitation to have a meaningful effect.
+With the Poisson model, every version I tried which included precipitation in some form had an inferior AIC, so ultimately I decided not to include it, at least for now.
+Finally, there is slight asymmetry between the early morning and late evening predictions due to how the weekends are defined.
+Friday evening is likely more akin to a Saturday night, with more robberies, pushing up the late evening average of workdays as compared to the early mornings.
+Sunday evening likely has the opposite effect on the Weekends/Holidays.
+A modified delineation between the two periods may have done away with this, and there may be further clustering of hours/days within the week that could improve the fit.
+These are all areas which present opportunity for further refinement.
+
 \end{document}

### 2015-04-15 stat302/hw1.R
Finished homework 2, fixed R code for homework 1

@@ -57,23 +57,25 @@ if (problem2) {
 
 ### Problem 5
 if (problem5) {
-    subpops <- c("EelR","FeatherHfa","FeatherHsp","KlamathRfa")
     # Call given code
     source('official/exercises/seeb/train_test.R')
+    subpops <- unique(test$Population)
+    loci <- names(test)[3:26]
+    loci <- locs[!grepl("\\.1",loci)] # get rid of .1 version
 
     # Function to summarize frequency at locus by population (similar to trainc)
     # Add 1 so that no Allele is impossible in each subpopulation
     compute_freq <- function(data,locus){
-        counts <- table(data[,1+2*locus],data$Population) + 
-                    table(data[,2+2*locus],data$Population) + 1
-        return(counts/sum(counts))
+        counts <- table(data[,locus],data$Population) + 
+                    table(data[,paste(locus,".1",sep="")],data$Population) + 1
+        return(counts/colSums(counts))
     }
 
     # Get frequency at each locus
     train_freq <- list()
-    for (i in 1:12) {
-        train_freq[[i]] <- as.data.frame.matrix(compute_freq(train,i))
-        train_freq[[i]]$allele <- as.factor(rownames(train_freq[[i]]))
+    for (loc in loci) {
+        train_freq[[loc]] <- as.data.frame.matrix(compute_freq(train,loc))
+        train_freq[[loc]]$allele <- as.factor(rownames(train_freq[[loc]]))
     }
 
     # Set uniform prior
@@ -83,14 +85,16 @@ if (problem5) {
     # Calculate Log-Likelihood
     log_lks <-paste(subpops,"loglk",sep="_")
     test[,log_lks]<-0
-    for (i in 1:12) {
-        for (j in 1:2) {
-            test <- merge(test,train_freq[[i]],by.x=names(test)[j+i*2],
-                          by.y="allele")
+    for (loc in loci) {
+        for (e in c("",".1")) {
+            locus <- paste(loc,e,sep="")
+            test <- merge(test,train_freq[[loc]],by.x=locus,
+                          by.y="allele",all.x=T,sort=F)
             # Missing data won't effect likelihood
-            test[is.na(test[,j+i*2]),subpops] <- 1  
+            test[is.na(test[,locus]),subpops] <- 1  
             test[,log_lks] <- test[,log_lks] + log(test[,subpops])
             test<-test[,!(names(test) %in% subpops)]
+
         }
     }
 

### 2015-04-15 stat343/hw5.tex
Unscrewed up this repo - files I want are followed, everything else removed from rep

@@
             \item[a)] 
-                The constant variance assumption apiears to hold; we do not seem to have the variance change over the range of predicted values.
+                The constant variance assumption appears to hold; we do not seem to have the variance change over the range of predicted values.
 The plot on the left, which is the residuals versus the fitted values, indicates by the eyeball test that the distribution of residuals is constant across the range of predicted values.
@@
                 \end{center}
+
             \item[d)]
@@
                 \end{align*}
-                So in the expectation, you would get the same $\beta$ irrespective of the variance level.
+                So in the expectation, you would get the same $\beta$ irrespective of the variance level. \\
             \item[d)] 
+                My conclusion is that we really have two distinct sets of data that can't be modeled with one linear regression.
                 \begin{center}
@@
                 \end{itemize}
-                This is essentially two separate models.
+                This is essentially creating a separate model for each half of the data.
 Since the original model corresponds to a subspace of this model (where the coefficient on the first variable is $0$ and the second two hav equal coefficients), we can compare the two models with a $F$ test: \\
@@
         Since $H$ is positive definite, it has an inverse. Since it is symmetric, the inverse is as well. Since \(MM^{-1}=I\), we get that
-        \[HA+eB=I, \, HB+eC=0, \, e^TA+fB^T=0, \, e^TB+fC=I\]
+        \[HA+eB^T=I, \, HB+eC=0, \, e^TA+fB^T=0, \, e^TB+fC=I\]
         From the second equation, using $H^{-1}$, we get \(B=-H^{-1}eC\).
@@
         \[A=H^{-1} + H^{-1}e(f-e^TH^{-1}e)^{-1}e^TH^{-1}\]
-        Plugging everything in and multiplying $MM^{-1}$, I get the desired result, so my initial assumption appears to be correct.
     \item[5)]

### 2015-04-15 stat343/hw1.R

@@ -1,0 +1,125 @@
+library("alr3")
+library("FNN")
+data("htwt")
+sink("hw1.Rout", append=FALSE, split=TRUE)
+
+readline("Problem 2.1.1")
+jpeg('hw1_2_1_1.jpg')
+plot(htwt$Ht,htwt$Wt)
+dev.off()
+readline("I would argue that this data, though seeming to indicate some sort of correlation between height and weight befiting a linear model, is too small considering the amount of noise for a linear regression to make sense. The estimates would be far too noisy to read anything into it. That said, if there was more data it looks like a linear model would make sense.")
+
+readline("Problem 2.1.2")
+n<-length(htwt$Ht)
+xbar<-sum(htwt$Ht)/n
+xbar
+ybar<-sum(htwt$Wt)/n
+ybar
+sxx<-sum((htwt$Ht-xbar)^2)
+sxx
+syy<-sum((htwt$Wt-ybar)^2)
+syy
+sxy<-sum((htwt$Ht-xbar)*(htwt$Wt-ybar))
+sxy
+beta1 <- sxy/sxx
+beta1
+beta0 <- ybar - beta1*xbar
+beta0
+jpeg('hw1_2_1_2.jpg')
+plot(htwt$Ht,htwt$Wt)
+curve(beta0 + x*beta1, add=TRUE)
+dev.off()
+
+readline("Problem 2.1.3")
+rss<-syy-sxy^2/sxx
+sigmasq<-rss/(n-2)
+sigmasq
+se_beta1 <- sqrt(sigmasq/sxx)
+se_beta1
+se_beta0 <- sqrt(sigmasq*(1/n + xbar^2/sxx))
+se_beta0
+cov_betas <- -sigmasq*xbar/sxx
+cov_betas
+t_beta1 <- beta1/se_beta1
+t_beta1
+p_val_beta1 <- 2*pt(-abs(t_beta1),n-2)
+p_val_beta1
+t_beta0 <- beta0/se_beta1
+t_beta0
+p_val_beta0 <- 2*pt(-abs(t_beta0),n-2)
+p_val_beta0
+
+readline("Problem 2.1.4")
+df<-c(1,n-2,n-1)
+ss<-c(syy-rss,rss,syy)
+ms<-c((syy-rss)/1,sigmasq,NA)
+f<-c((syy-rss)/1/sigmasq,NA,NA)
+pval<-pf(-abs(f),1,n-2)
[…]

### 2015-04-15 stat343/hw4.R

@@ -1,0 +1,115 @@
+library('alr3')
+library('faraway')
+library('car')
+library('MASS')
+
+#############################
+# Problem 1
+#############################
+
+### Part d)
+# Load the data and create the matrices 
+data('savings')
+attach(savings)
+n<-nrow(savings)
+x<-as.matrix(data.frame(rep(1,n),savings[c(2,3,4,3,5)]))
+y<-as.matrix(savings[1])
+qr<-qr(x)
+
+R<-qr.R(qr)[1:qr$rank,]
+Q<-qr.Q(qr)[,1:qr$rank]
+W<-R[,qr$pivot]
+beta<-t(W) %*% solve(W%*%t(W)) %*% t(Q) %*% y 
+beta 
+
+### part e)
+
+model<-lm(y~x)
+summary(model)
+
+### part f)
+qr.R(qr)
+
+#############################
+# Problem 3 
+#############################
+
+### part b
+# Linear model and output
+model<-lm(sr ~ pop15 + pop75 + dpi + ddpi)
+X <- model.matrix(model)
+beta<-model$coef
+sigma2 <- summary(model)$sigma^2
+df <- summary(model)$df[2]
+
+# Set up for ellipse
+R<-matrix(c(0,0,1,-1,-1,0,0,0,0,1),nrow=2,ncol=5)
+variance.pred <- R %*% solve(t(X) %*% X) %*% t(R) + diag(2) 
+variance.mean <- R %*% solve(t(X) %*% X) %*% t(R)
+
+mycenter <- R%*%beta
+radius9<- sqrt(2* sigma2 * qf(.1,2,df, lower.tail=FALSE))
+radius95<- sqrt(2* sigma2 * qf(.05,2,df, lower.tail=FALSE))  
+
+# Draw the ellipses
+# Ellipse for 90% CI
+pdf("hw4_3_b_ci.pdf")
+plot.new()
+plot.window(xlim=c(-10,15),ylim=c(-10,15))
+title(main='Prediction Interval of z', xlab='z1', ylab='z2')
+car::ellipse(center=c(mycenter), shape=variance.pred, radius=radius9, lty=2, main='Confidence Interval for Prediction', col=palette()[1])
+car::ellipse(center=c(mycenter), shape=variance.pred, radius=radius95,lty=3, add=TRUE,col=palette()[1])
[…]

### 2015-04-15 stat343/hw5.R

@@ -1,0 +1,198 @@
+library('parallel')
+library('alr3')
+library('faraway')
+library('car')
+library('MASS')
+library('locfit')
+library('xtable')
+
+#############################
+# Problem 1
+#############################
+
+data('sat')
+attach(sat)
+lm.sat<-lm(total~expend + salary + ratio + takers)
+n<-nrow(sat)
+names<-row.names(sat)
+
+### a)
+pdf("hw5_1_a_res.pdf")
+plot(fitted(lm.sat),residuals(lm.sat),xlab="Fitted SAT Score",ylab="Residuals",main="Residuals Plot")
+abline(h=0)
+legend('topleft', legend=c('y=0'), lty=c(1))
+dev.off()
+
+lm.res<-lm(sqrt(abs(residuals(lm.sat)))~predict(lm.sat))
+pdf("hw5_1_a_sqrtres.pdf")
+plot(fitted(lm.sat),sqrt(abs(residuals(lm.sat))),xlab="Fitted SAT Score",ylab="Square Root of Absolute Value of Residuals",main="Square Root Residuals Plot")
+abline(lm.res)
+legend('topleft', legend=c("Linear Fit"), lty=c(1))
+dev.off()
+
+### b)
+pdf("hw5_1_b_qq.pdf")
+qqnorm(residuals(lm.sat),ylab="Residuals",xlab="Theoretical Normal Quantiles", main="Residuals vs. Normal Qunatile Plot")
+qqline(residuals(lm.sat))
+dev.off()
+shap<-shapiro.test(residuals(lm.sat))
+
+### c)
+pdf("hw5_1_c_halfnorm.pdf")
+halfnorm(hatvalues(lm.sat),labs=names,ylab="Leverages",main="Leverages vs. Half-Normal Quantiles")
+dev.off()
+
+### d)
+pdf("hw5_1_d_tqq.pdf")
+qqplot(rt(1000,df=n-5),rstudent(lm.sat),xlab="T Distribution With 45 Degrees of Freedom", ylab="Studentized Residuals", main="QQ Plot Studentized Resisduals vs. T Distribution")
+abline(0,1)
+dev.off()
+stud.max<-rstudent(lm.sat)[which.max(abs(rstudent(lm.sat)))]
+qtpt10<-qt(.1/100,45)
+qtpt05<-qt(.05/100,45)
+
[…]

### 2015-04-15 stat343/hw7.R

@@ -1,0 +1,228 @@
+library('parallel')
+library('alr3')
+library('faraway')
+library('car')
+library('MASS')
+library('locfit')
+library('xtable')
+library('nlme')
+library('quantreg')
+library('splines')
+library('rpart')
+library('randomForest')
+
+#digits of precision
+d<- 3
+
+#####################
+# Problem 1
+#####################
+
+data(aatemp)
+
+### a)
+temp.lm <- lm(temp~year,data=aatemp)
+print(xtable(summary(temp.lm),digits=d),type="latex",file="hw7/hw7_1_a.tex")
+
+pdf("hw7/hw7_1_a.pdf")
+plot(aatemp,xlab="Year",ylab="Temperature",main="Linear Fit",xlim=c(1850,2000),pch=20)
+abline(temp.lm,xlim=c(1850,2000))
+dev.off()
+
+### b)
+temp.alm<-gls(temp~year,correlation=corAR1(form=~year),method="ML",data=aatemp)
+print(xtable(summary(temp.alm)$tTable,digits=d),type="latex",file="hw7/hw7_1_b_ttable.tex")
+print(xtable(intervals(temp.alm)$corStruct,digits=d),type="latex",file="hw7/hw7_1_b_acor.tex")
+
+pdf("hw7/hw7_1_b.pdf")
+matplot(aatemp$year, cbind(aatemp$temp,temp.alm$fit),type="pl",pch=20,lty=1,col=1,xlim=c(1850,2000),ylab="Temperature",xlab="Year",main="Autoregressive Fit")
+dev.off()
+
+### c)
+crit<-.2
+p<-1
+deg<-11
+while (p>crit) {
+    deg <- deg-1
+    temp.poly<-poly(aatemp$year,deg)
+    poly.data<-data.frame(temp=aatemp$temp,temp.poly=temp.poly)
+    temp.poly.lm <- lm(temp ~ .,data=poly.data)
+    p<-coef(summary(temp.poly.lm))[deg+1,4]
+}
+print(xtable(summary(temp.poly.lm),digits=d),type="latex",file="hw7/hw7_1_c.tex")
+
+pdf("hw7/hw7_1_c.pdf")
+plot(aatemp,xlab="Year",ylab="Temperature",main="Polynomial Fit",xlim=c(1850,2000),pch=20)
[…]

### 2015-04-21 stat302/hw3.tex
starting on homework 3

--- 2014-12-03 stat343/hw8.tex
@@
 \begin{document}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-STAT 343 Homework 8 \hfill Aaron Maurer
+STAT 302 Homework 3 \hfill Aaron Maurer
 \vspace{2mm}
@@
             \item[(a)]
-                As we have proved, if the matix $X$ isn't singular, then the least squares solution from OLS is unique and indeed the optimal fit.
-Thus, if a $\beta$ acheives the minimum least squares, it must be the OLS coefficient from the regression.
-Calculating the sum of squares for a coefficient $\hat\beta ^{(-i)}$ on $\tilde Y^{(-i)}$, we get
                 \begin{align*}
-                    \|\tilde Y^{(-i)} - X\hat\beta^{(-i)}\|^2 &= (x_i\beta^{(-i)} - x_i\beta^{(-i)})^2 + \|Y^{(-i)} - X^{(-i)}\hat\beta ^{(-i)}\|^2 \\
-                    \|\tilde Y^{(-i)} - X\hat\beta^{(-i)}\|^2 &= \|Y^{(-i)} - X^{(-i)}\hat\beta^{(-i)}\|^2
-                \end{align*}
-                Since the sum on the right is the minimum possible over all $\beta$ (since the coefficient was fit by OLS on $Y^{(-i)}$), $\hat\beta ^{(-i)}$ minimizes the sum of squares for $\tilde Y^{(-i)}$.
-By our earlier reasoning, it thus must be the coefficient that would result from OLS run predicting $\tilde Y^{(-i)}$ in terms of $X$.
-            \item[(b)]
-                The prediction of a linear model with coefficient $\hat\beta ^{(-i)}$ at point $x_i^T$ is $x_i^T\hat\beta ^{(-i)}$, so since that is the coefficient for regressing $\tilde Y^{(-i)}$ on $X$, $x_i^T\hat\beta ^{(-i)}$ will be the fit.
-
-        \end{itemize}
-    \item[2.]
-        \begin{itemize}
-            \item[(a)]
-                \begin{align*}
-                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- S(x_i)^{(-i)}Y^{(-i)} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- \frac{1}{\sum_{j\neq i} S_{ij}} \sum_{j\neq i} y_j S_{ij} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= y_i- \frac{\hat f(x_i)-y_iS_{ii}(X)}{1-S_{ii}} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= \frac{y_i - y_i S_{ii} - \hat f (x) + y_iS_{ii}}{1- S_{ii}} \\
-                    \hat y_i-\hat f^{(-i)}(x_i) &= \frac{y_i -\hat f (x) }{1- S_{ii}} \\
+                    P(D \st P) &= \frac{P(P\st D)P(D)}{P(P\st D)P(D) +  P(P\st D^c)P(D^c)} \\
+                               &= \frac{p_0p_1}{p_0p_1 +  (1-p_0)p_1} \\
                 \end{align*}
             \item[(b)]
-                Rather than refitting the regression $n$ times to estimate the leave one out cross validation estimate of error, this allows the regression to merely be fit once, saving a lot of costly $QR$ factorization or matrix inversion.
-            \item[(c)]
-                Since $1\geq1-S_{ii}\geq0$, we have 
-                \begin{align*}
-                    \vert y_i - \hat f(x_i) \vert &\leq \frac{\vert y_i - \hat f(x_i) \vert }{1-S_{ii}} \\
-                    \vert y_i - \hat f(x_i) \vert &\leq \left\vert\frac{y_i - \hat f(x_i) }{1-S_{ii}} \right\vert  \\
-                    \vert y_i - \hat f(x_i) \vert &\leq \left\vert y_i - \hat f^{(-i)}(x_i)\right\vert  
-                \end{align*}
-
         \end{itemize}
-    \item[3.]
-        \begin{itemize}
-            \item[(a)]
-                Applying backwards elimination with a critical value of $.2$, we only remove the variable status. After that, all predictors have a p-value of less than $.2$, so we stop. \\
-                \smallskip
-
-                p-value at the point of exclusion: \\
-                \FloatBarrier
-                \input{hw8/hw8_3_a_excl} 
-                \FloatBarrier
-
-                Resulting model: \\
-                \FloatBarrier
-                \input{hw8/hw8_3_a_lm} 
-                \FloatBarrier
-
-            \item[(b)]
-                For each of the following three methods, I used an exhaustive subset search, which determined which set of a particular number of variables resulted in the smallest RSS.
-The result, which is a boolean indicating whether the predictor should be included at each number of predictors is below.
-Since this will uniformly be the best model of $n$ predictors, the remaining decision is which $n$ to use.
-It should be noted that the model in a is indeed the optimal model of $3$ predictors.
-\\
-                \FloatBarrier
-                \input{hw8/hw8_3_which} 
-                \FloatBarrier
-                The model with the lowest AIC also had $3$ predictors, and is thus the same as in a. The different AICs for the best model with $n$ predictors  below. \\
-                \FloatBarrier
-                \input{hw8/hw8_3_b} 
-                \FloatBarrier
-
-            \item[(c)]
-                The model with the highest adjusted $R^2$ is this same model with $3$ predictors. Optimal adjusted $R^2$ for different number of predictors are below. \\
-                \FloatBarrier
-                \input{hw8/hw8_3_c} 
-                \FloatBarrier
-
-            \item[(d)]
-                Going by Mallow's $C_p$, we once again get the same model.
-The smallest model that had a $C_p$ smaller than the number of parameters (one more than the number of predictors due to the intercept) was the model with 3 predictors (4 parameters): \\
-                \FloatBarrier
-                \input{hw8/hw8_3_d} 
-                \FloatBarrier
-        \end{itemize}
-    \item[4.]
-        The order of these operations doesn't make much difference.
-Point 21 is influential both before and after model selection, and the prefered model includes Air.Flow and Water.Temp, while excluding Acid.Conc, whether or not the outlier is included.
-\\
-        \smallskip
-        Our initial model is below. Acid.Conc looks ripe to be removed with a high $p-value$
-        \FloatBarrier
-        \input{hw8/hw8_4_lm.tex} 
-        \FloatBarrier
-        Indeed, when we test all subsets of the data, it is excluded from every model that has fewer than three predictors:
-        \FloatBarrier
-        \input{hw8/hw8_4_small_which} 
-        \FloatBarrier
-        The optimal subset, when going by adjusted $R^2$ is the model with just two predictors ($R^2$ for each optimal subset of a particular size below):
-        \FloatBarrier
-        \input{hw8/hw8_4_small_adjr2} 
-        \FloatBarrier
-        Choosing this as our model, we get the following model:
-        \FloatBarrier
-        \input{hw8/hw8_4_small_lm.tex} 
-        \FloatBarrier
-        Checking the distribution of studentized residuals for outliers, we see one potential culprit (the largest residual, in the top right, which heavily deviates from T-distribution).
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_small_rstudent} 
-        \end{center}
-        It is the 21st data point.
-With a studentized residual value of $-3.47$, it is just inside conservative Bonferroni critical value, for $\alpha=.5$, of $-3.53$.
-Thus, we don't quite reject it as an outlier.
-However, looking at the plot of Cook's distances, it is clear that the same point is highly influential.
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_small_cook}
-        \end{center}
-        Now, going back to the full model which included Acid.Conc, we look again for outliers and influential points.
-Shockingly, the same point is once again the only real canidate.
-The distribution of studentized residuals on this
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_rstudent} 
-        \end{center}
-        Testing it as an outlier as before, it has a studentized residual of $-3.33$ in this model, while the critical value is $-3.54$, so we don't reject it.
-However, its still influential, so we kick it out for that:
-        \begin{center}
-            \includegraphics[width=9cm]{hw8/hw8_4_cook}
-        \end{center}
-        Preforming variable selection on the data set with this point excluded, we still find the optimal model has the two variables besides Acid.Conc. Optimal subsets of each size:
-        \FloatBarrier
-        \input{hw8/hw8_4_no_which} 
-        \FloatBarrier
-        Adjusted $R^2$ for these subsets, showing we still prefer the model with two predictors:
-        \FloatBarrier
-        \input{hw8/hw8_4_no_adjr2} 
-        \FloatBarrier
-    \item[5.]
-        \begin{itemize}
-            \item[(a)]
-                I have plotted the singular values below, gotten by taking the square roots of the absolute values of the eigenvalues of $XX^T$ (once $X$ had been centered).
-They quickly fall off to very small values, reflecting that most of the variation in the data set is explained by just a few underlying components.
-                \begin{center}
-                    \includegraphics[width=9cm]{hw8/5_a}
-                \end{center}
-            \item[(b)]
-                The components of the two largest singular vectors are plotted below.
-As would be expected, the components of the second vector are generally far smaller than those of the first.
-It appears that approximately half the time the second component has the same sign as the first, and half the time the opposite sign, as might be expected since they are orthogonal.
-The vaguely resemble piecies of orthogonal waves.
-                \begin{center}
-                    \includegraphics[width=9cm]{hw8/5_b}
-                \end{center}
-
-            \item[(c)] 
-            We see that the regression on the two largest princple components doesn't actually predict the data as well as the two best predictors from the original data set (for the same number of variables, the latter has higher $R^2$).
-While this latter model predicts almost as well as the model including all variables, the largest principle components are signficantly less predictive, though still reasonably predictive.
-The reason for this is that the principle components are a decomposition of variation within the predictors; however, the largest source of variation in these predictors isn't necessarily predictive of the outcome.
-For instance, if there is a lot of variation among a few variables that aren't predictive of the outcome, they may correspond to large principle components, but these principle components will likely be also unpredictive.
-On the other hand, since the principle components as a whole define a space equivalent to the orginal space defined by all the predictors, a model including all of them will be equally predictive to including all variables.
-\\
-            \smallskip
-            The regression on the two largest principle components ($R^2=.5469$):
-            \FloatBarrier
-            \input{hw8/hw8_5_pc_lm} 
-            \FloatBarrier
-            The regression on all variables ($R^2=.6866$):
-            \FloatBarrier
-            \input{hw8/hw8_5_all_lm} 
-            \FloatBarrier
-            The regression on age and height ($R^2=.6562$):
-            \FloatBarrier
-            \input{hw8/hw8_5_sm_lm} 
-            \FloatBarrier
-        \end{itemize}
-
 \end{itemize}

### 2015-04-21 stat302/hw3.R

@@ -1,0 +1,90 @@
+### Libraries
+library('xtable')
+library('stats4')
+library('locfit')
+library('SDMTools')
+library('plotrix')
+library('reshape2')
+library('nlme')
+library('parallel')
+library('faraway')
+
+### Caller
+problem1 <- TRUE 
+
+### Problem 1
+if (problem1) {
+    ## part (b) - Write a function to calculate credible intervals
+    drawTheta <- function(b,X,n,a) {
+        p <- rbeta(3,a+X,a+n-X)
+        return(p[1]*p[2]/(p[1]*p[2]+(1-p[1])*p[3]))
+    }
+    bootThetaCI <- function(b,X,n,a) {
+        vals<-sort(unlist(mclapply(1:b,drawTheta,X<-X,n<-n,a<-a,mc.cores=4)))
+        return(c(vals[ceiling(b*.025)],vals[ceiling(b*.975)]))
+    }
+
+    ## part (c)/(d) - generate credible intervals for a series of inputs
+
+    formList <- function(li) {
+        return(paste('(',paste(li,collapse=','),')',sep=''))
+    }
+    formCI <- function(ci,d=3) {
+        return(formList(format(ci,digits=d)))
+    }
+
+    scenarios <- list( n=list(20,20,80,80),X=list(c(2,18,2),c(10,18,0),c(20,60,20),c(40,72,8)),CIpt5=list(),CI1=list(),CI2=list())    
+
+    for (i in 1:4) {
+        scenarios$CIpt5[[i]] <- bootThetaCI(10000,scenarios$X[[i]],scenarios$n[[i]],.5)
+        scenarios$CI1[[i]]   <- bootThetaCI(10000,scenarios$X[[i]],scenarios$n[[i]],1)
+        scenarios$CI2[[i]]   <- bootThetaCI(10000,scenarios$X[[i]],scenarios$n[[i]],2)
+    }
+
+    # build the tables
+    table.1.cd = data.frame(
+        n=unlist(scenarios$n),
+        x=unlist(lapply(scenarios$X,formList)),
+        CIpt5=unlist(lapply(scenarios$CIpt5,formCI)),
+        CI1  =unlist(lapply(scenarios$CI1,formCI)),
+        CI2  =unlist(lapply(scenarios$CIpt5,formCI))
+        )
[…]

### 2015-04-22 stat302/hw3.tex
Done with homework 3

@@
             \item[(b)]
+                Here is my R code:
+                \lstinputlisting[firstline=18,lastline=25]{hw3.R}
+            \item[(c)/(d)]
+                I have produced the results for $a=.5,1 \,\&\, 2$ in one table, each time with confidence intervals based off of $10,000$ iterations.
+                \FloatBarrier
+                \input{hw3/pr1_cd.tex}
+                \FloatBarrier
+                We can observe the sensitivity to $a$ by noting that the confidence intervals get pulled towards $.5$ as $a$ increases.
+The distinction isn't huge though; with $n=20$ we note the bounds change by as much as $.1$, but with $n=80$, the change is at most near $.01$.
+            \item[(e)]
+                I have simulated the portion of the time the true $\theta$ (th) was below the CI lower bound (LB) and the portion the time it was above the upper bound (UB) for each $a$ and the given $p$.
+The simulation drew the $X$ $1,000$ times, and for each of these generated a confidence interval based on another $1,000$ random draws based on the given $X$.
+                \FloatBarrier
+                \input{hw3/pr1_e.tex}
+                \FloatBarrier
+                In general, $a=.05$ seems to have proper, or very close to proper, frequentist coverage properties.
+However, in the cases where there are $p$ near $0$ or $1$, this seems to be less true, with the worst case being $p_0=.95, p_1=.95, p_2=0$, where we never see $\theta$ fall below the CI.
+For different values of $a$ though, we seem to consistently see either higher or lower error rates on each side of the CI.
         \end{itemize}
+    \item[2.]
+        \begin{itemize}
+            \item[i)]
+                We can derive $\E(\theta_j)$ as such:
+                \begin{align*}
+                    \E(\theta_j) &= \idotsint_{\sum_{i=1}^k\theta_i=1}\theta_i \frac{\Gamma(\sum_{i=1}^k\alpha_i)}{\prod_{i=1}^k\Gamma(\alpha_i)} \left(\prod_{i=1}^{k}\theta_i^{\alpha_i-1}\right)d\theta_1...d\theta_k \\
+                    \E(\theta_j) &= \frac{\Gamma(1+\alpha_j)\Gamma(\sum_{i=1}^k\alpha_i)}{\Gamma(\alpha_j)\Gamma(1+\sum_{i=1}^k\alpha_i)}   \idotsint_{\sum_{i=1}^k\theta_i=1}\frac{\Gamma(1+\sum_{i=1}^k\alpha_i)}{\Gamma(1+\alpha_j)\prod_{i\neq j}\Gamma(\alpha_i)} \left(\theta_i^{\alpha_i}\prod_{i\neq j}\theta_i^{\alpha_i-1}\right)d\theta_1...d\theta_k \\
+                    \E(\theta_j) &= \frac{\Gamma(1+\alpha_j)\Gamma(\sum_{i=1}^k\alpha_i)}{\Gamma(\alpha_j)\Gamma(1+\sum_{i=1}^k\alpha_i)}\\
+                    \E(\theta_j) &= \frac{\alpha_j}{\sum_{i=1}^k\alpha_i}\\
+                \end{align*}
+                Then, as a first step to get the variance,
+                \begin{align*}
+                    \E(\theta_j^2) &= \idotsint_{\sum_{i=1}^k\theta_i=1} \theta_i^2 \frac{\Gamma(\sum_{i=1}^k\alpha_i)}{\prod_{i=1}^k\Gamma(\alpha_i)} \left(\prod_{i=1}^{k}\theta_i^{\alpha_i-1}\right)d\theta_1...d\theta_k \\
+                    \E(\theta_j^2) &= \frac{\Gamma(2+\alpha_j)\Gamma(\sum_{i=1}^k\alpha_i)}{\Gamma(\alpha_j)\Gamma(2+\sum_{i=1}^k\alpha_i)}   \idotsint_{\sum_{i=1}^k\theta_i=1}\frac{\Gamma(2+\sum_{i=1}^k\alpha_i)}{\Gamma(2+\alpha_j)\prod_{i\neq j}\Gamma(\alpha_i)} \left(\theta_i^{1+\alpha_i}\prod_{i\neq j}\theta_i^{\alpha_i-1}\right)d\theta_1...d\theta_k \\
+                    \E(\theta_j^2) &= \frac{\Gamma(2+\alpha_j)\Gamma(\sum_{i=1}^k\alpha_i)}{\Gamma(\alpha_j)\Gamma(2+\sum_{i=1}^k\alpha_i)}\\
+                    \E(\theta_j^2) &= \frac{\alpha_j(\alpha_j+1)}{\left(\sum_{i=1}^k\alpha_i+1\right)\sum_{i=1}^k\alpha_i}\\
+                \end{align*}
+                Which we can use to get
+                \begin{align*}
+                    \var(\theta_j) &= \E(\theta_j^2) - \E(\theta_j)^2 \\
+                    \var(\theta_j) &= \frac{\alpha_j^2+\alpha_j}{\left(\sum_{i=1}^k\alpha_i+1\right)\sum_{i=1}^k\alpha_i} - \frac{\alpha_j^2}{\left(\sum_{i=1}^k\alpha_i\right)^2} \\
+                    \var(\theta_j) &= \frac{\alpha_j^2\left(\sum_{i=1}^k\alpha_i\right)+\alpha_j\left(\sum_{i=1}^k\alpha_i\right)-\alpha_j^2\left(\sum_{i=1}^k\alpha_i\right) -\alpha_j^2}{\left(\sum_{i=1}^k\alpha_i+1\right)\left(\sum_{i=1}^k\alpha_i\right)^2} \\
+                    \var(\theta_j) &= \frac{\alpha_j\left(\sum_{i=1}^k\alpha_i\right)- \alpha_j^2}{\left(\sum_{i=1}^k\alpha_i+1\right)\left(\sum_{i=1}^k\alpha_i\right)^2}
+                \end{align*}
+                Finally, 
+                \begin{align*}
+                    \cov(\theta_j,\theta_i) &= \E(\theta_j\theta_i) - \E(\theta_j)\E(\theta_i) \\
+                    \cov(\theta_j,\theta_i) &= \frac{\Gamma(1+\alpha_j)\Gamma(1+\alpha_i)\Gamma(\sum_{i=1}^k\alpha_i)}{\Gamma(\alpha_i)\Gamma(\alpha_j)\Gamma(2+\sum_{i=1}^k\alpha_i)} - \frac{\alpha_j\alpha_i}{\left(\sum_{i=1}^k\alpha_i\right)^2} \\
+                    \cov(\theta_j,\theta_i) &= \frac{\alpha_j\alpha_i}{\left(\sum_{i=1}^k\alpha_i+1\right)\sum_{i=1}^k\alpha_i} - \frac{\alpha_j\alpha_i}{\left(\sum_{i=1}^k\alpha_i\right)^2} \\
+                    \cov(\theta_j,\theta_i) &= \frac{-\alpha_j\alpha_i}{\left(\sum_{i=1}^k\alpha_i+1\right)\left(\sum_{i=1}^k\alpha_i\right)^2} \\
+                \end{align*}
+                Qualitatively, as any particular $\alpha_i$ grows with the rest held constant, the probability of high values for $\theta_i$ grows.
+When all the $\alpha_j$ grow together, the distribution becomes more concentrated around the center where all $\theta_j$ are close together, and when all the $\alpha_j$ shrink together, the distribution becomes clustered around high values for one of the $\theta_j$ with the rest close to $0$
+            \item[ii)]
+                Let, for each $1\leq j \leq k$, let \(|\{j;X_j=k\}|=n_j\). Then,
+                \begin{itemize}
+                    \item[a)]
+                        \begin{align*}
+                            P(\theta \st X_1,...,X_n) &\propto P(X_1,...,X_n \st \theta)P(\theta) \\
+                            P(\theta \st X_1,...,X_n) &\propto \left(\prod_{i=1}^k\theta_i^{n_i}\right)\left(\prod_{i=1}^k\theta_i^{\alpha_i-1}\right) \\
+                            P(\theta \st X_1,...,X_n) &\propto \prod_{i=1}^k\theta_i^{n_i+\alpha_i-1}\\
+                        \end{align*}
+                        So \(\theta \st X_1,...,X_n\) has a Dirichlet distribution with parameters \((n_1+\alpha_1,...,n_k+\alpha_k)\)
+                    \item[b)]
+                        \begin{align*}
+                            P(X_{n+1}=j \st X_1,...,X_n) &= \idotsint_{\sum_{i=1}^k\theta_i=1}P(X_{n+1}=j \st X_1,...,X_n,\theta)P(\theta\st X_1,...,X_n)d\theta_1...d\theta_k \\
+                            P(X_{n+1}=j \st X_1,...,X_n) &= \frac{\Gamma(\sum_{i=1}^k n_i +\alpha_i)}{\prod_{i=1}^k\Gamma(n_i+\alpha_i)}\idotsint_{\sum_{i=1}^k\theta_i=1}\theta_j \prod_{i=1}^k\theta_i^{n_i+\alpha_i-1}d\theta_1...d\theta_k \\
+                            P(X_{n+1}=j \st X_1,...,X_n) &= \frac{n_j+\alpha_j}{\sum_{i=1}^kn_i+\alpha_i}
+                        \end{align*}
+                        So we can conclude that \(X_{n+1}=j \st X_1,...,X_n\) has a multinomial distribution with probability vector
+                        \[p = \left( \frac{n_1+\alpha_1}{\sum_{i=1}^kn_i+\alpha_i},...,\frac{n_k+\alpha_k}{\sum_{i=1}^kn_i+\alpha_i}\right) \]
+                \end{itemize}
+        \end{itemize}
+    \item[3.]
+        If $X$ is Poisson with mean $\theta$, then \(P(X=x\st \theta)\propto \theta^x e^{-\theta}\), which makes a prior that is something of the sort \(P(\theta)\propto \theta^a e^{b\theta}\), for some hyperparameters $a$, $b$, a natural choice.
+This is satistified by the gamma distribution, which is the conjugate prior.
+If we parameterize it the usual way with \(P(\theta)\propto \theta^{\alpha-1} e^{-\beta\theta}\), we get the posterior
+        \begin{align*}
+            P(\theta \st X) &\propto P(X \st \theta)P(\theta) \\
+            P(\theta \st X) &\propto \theta^{\sum x_i} e^{-n\theta}\theta^{\alpha-1} e^{-\beta\theta} \\
+            P(\theta \st X) &\propto \theta^{\sum x_i+\alpha-1} e^{-(\beta+n)\theta}\\
+        \end{align*}
+        Giving us a gamma posterior with parameters \((\sum x_i+\alpha,\beta+n)\). \\
+        \vspace{5mm} \\
+        We can get the Jeffreys prior from the Fisher information:
+        \begin{align*}
+            p(\theta) &\propto \sqrt{I(\theta)} \\
+            p(\theta) &\propto \sqrt{-\E\left[\frac{d^2}{d\theta^2} \log(f(X\st \theta))\right]} \\
+            p(\theta) &\propto \sqrt{-\E\left[\frac{d^2}{d\theta^2} x\log(\theta)-\theta+c\right]} \\
+            p(\theta) &\propto \sqrt{-\E\left[\frac{d}{d\theta} \frac{x}{\theta}-1\right]} \\
+            p(\theta) &\propto \sqrt{-\E\left[-\frac{x}{\theta^2}\right]} \\
+            p(\theta) &\propto \sqrt{\frac{1}{\theta}} \\
+            p(\theta) &\propto \theta^{-\frac{1}{2}} \\
+        \end{align*}
+        With the Jeffreys prior, we get the posteriors
+        \begin{align*}
+            P(\theta \st x) &\propto P(x \st \theta)P(\theta) \\
+            P(\theta \st x) &\propto \theta^{x} e^{-\theta}\theta^{-\frac{1}{2}} \\
+            P(\theta \st x) &\propto \theta^{x-\frac{1}{2}} e^{-\theta} \\
+            P(\theta \st cx) &\propto \theta^{cx-\frac{1}{2}} e^{-\theta} \\
+        \end{align*}
+        Versus, for the 'scale invariant' prior
+        \begin{align*}
+            P(\theta \st x) &\propto P(x \st \theta)P(\theta) \\
+            P(\theta \st x) &\propto \theta^{x} e^{-\theta}\theta^{-1} \\
+            P(\theta \st x) &\propto \theta^{x-1} e^{-\theta} \\
+            P(\theta \st cx) &\propto \theta^{cx-1} e^{-\theta} \\
+        \end{align*}
+        The posteriors are similar, but the Jeffreys prior has a mean that is a bit closer to the MLE mean, which probably makes it preferable as far as a noninformative prior.
+As $c$ grows, they both converge to the MLE.
+    \item[4.]
+        We can derive the Jefferies prior from the Fisher Information:
+        \begin{align*}
+            I(p)_{i,j} &= -\E\left[H\left(\log(f(X\st p))\right)_{i,j}\right] \\
+            I(p)_{i,j} &= -\E\left[H\left(c + \sum_{i=1}^k x_i\log(p_i)\right)_{i,j}\right] \\
+            I(p)_{i,j} &= \begin{cases} 
+                                    -\E\left[-\frac{x_i}{p_i^2}\right] & \mbox{if } i=j \\ 
+                                                                     0 & \mbox{if } i\neq j 
+                               \end{cases}  \\
+            I(p)_{i,j} &= \begin{cases} 
+                               \frac{n}{p_i} & \mbox{if } i=j \\ 
+                                       0 & \mbox{if } i\neq j 
+                               \end{cases} 
+        \end{align*}
+        Which makes \(|I(p)|^{\frac{1}{2}} = n^{\frac{k}{2}}\prod_{i=1}^k p_i^{-\frac{1}{2}}\). We can thus conclude
+        \[p(\theta) \propto \prod_{i=1}^k p_i^{-\frac{1}{2}} \]
+        Making it a Dirichlet prior with each $\alpha$ set to $\frac{1}{2}$.
+    \item[5.]
+        \begin{itemize}
+            \item[a)]
+                If we let $X$ be the training data, $G_j$ be the testing data, consisting of gene $g_{l,j}$ at locus $l$ for individual $j$, and $P_j$ denote the population of training sample $j$, then previously our model was:
+                \begin{align*}
+                    P(P_j \st G_j, X) &\propto P(P_j)P(G_j \st P_j, X) \\
+                    P(P_j \st G_j, X) &\propto P(P_j)\prod_{l=1}^{24} P(g_{l,j} \st P_j, X) \\
+                \end{align*}
+                Where $P(g_{l,j} \st P_j=k, X)$ was the portion of the training population $k$ which had gene $g_{l,j}$ at locus $l$. Now however, I implemented a new model, where
+                \begin{align*}
+                    P(P_j \st G_j, X, \alpha) &\propto P(P_j)P(G_j \st P_j, X, \alpha) \\
+                    P(P_j \st G_j, X, \alpha) &\propto P(P_j)\prod_{l=1}^{24} P(g_{l,j} \st P_j, X, \alpha) \\
+                \end{align*}
+                Where $P(g_{l,j} \st P_j, X, \alpha)$ is the posterior prediction distribution based on all the genes at locus $l$ in population $P_j$ with a uniform $\alpha$ Dirichlet distribution (as in question 2.ii.b).
+If $W$ is the set of genes at locus $l$, and $n_{w,P_j}$ is the count from the training sample of gene $w$ at locus $l$ in population $P_j$, then this quantity is:
+                \[ P(g_{l,j}=v \st P_j, X, \alpha) = \frac{\alpha + n_{v,P_j}}{\sum_{w\in W} \alpha +n_{w,P_j}} \]
+            \item[b)]
+                Trying out different values of $\alpha$, I got these empirical error rates
+                \FloatBarrier
+                \input{hw3/pr5_b.tex}
+                \FloatBarrier
+            \item[c)]
+                I calculated the likelihood for a given alpha as the outcome which made our training data most likely. In other words:
+                \begin{align*}
+                    P(G \st X,\alpha) &= \prod_{j=1}^n P(G_j \st X,\alpha) \\ 
+                    P(G \st X,\alpha) &= \prod_{j=1}^n \prod_{l=1}^{24} P(g_{l,j} \st X, \alpha) \\
+                    P(G \st X,\alpha) &= \prod_{j=1}^n \prod_{l=1}^{24}\sum_{k=1}^4 P(P_j=k)P(g_{l,j} \st P_j=k, X, \alpha) \\
+                \end{align*}
+                I maximized $\alpha$ over this quantity, with \(P(g_{l,j} \st P_j=k, X, \alpha)\) defined as above.
+            \item[d)]
+                When I did this, I found the optimal parameter to be $\alpha=.363$, which had a log likelihood of $-13621$ and yielded an error rate of $.231$.
+This rate is only a bit higher than the best rate observed in part b of $.216$.
+            \item[e)]
+                Where A=\(\{0.025, 0.05, 0.1, 0.2, 0.4, 0.8, 1.6, 3.2, 6.4\}\), I implemented the method as such:
+                \begin{align*}
+                    P(P_j \st G_j, X) &\propto P(P_j)P(G_j \st P_j, X) \\
+                    P(P_j \st G_j, X) &\propto \sum_{\alpha\in A} P(\alpha)P(P_j)P(G_j \st P_j, X, \alpha) \\
+                    P(P_j \st G_j, X) &\propto \sum_{\alpha\in A} \frac{1}{|A|}P(P_j)\prod_{l=1}^{24} P(g_{l,j} \st P_j, X, \alpha) \\
+                \end{align*}
+                This time, I got an error rate of $.243$, which is higher than the best $\alpha$ by it self by a few percent.
+This isn't surprising though, we are averaging the probabilities over a few $\alpha$ which we know have higher error rates than the best $\alpha$ we've seen.
+Of course, it may be that this value is really closer to what $\alpha$ "should" be given the true variation in the data.
+
+        \end{itemize}
+
 \end{itemize}

### 2015-04-22 stat302/hw3.R

@@ -10,7 +10,8 @@ library('parallel')
 library('faraway')
 
 ### Caller
-problem1 <- TRUE 
+problem1 <- FALSE
+problem5 <- TRUE
 
 ### Problem 1
 if (problem1) {
@@ -47,9 +48,9 @@ if (problem1) {
         x=unlist(lapply(scenarios$X,formList)),
         CIpt5=unlist(lapply(scenarios$CIpt5,formCI)),
         CI1  =unlist(lapply(scenarios$CI1,formCI)),
-        CI2  =unlist(lapply(scenarios$CIpt5,formCI))
+        CI2  =unlist(lapply(scenarios$CI2,formCI))
         )
-    colnames(table.1.cd) <- c("n0=n1=n2","(x0,x1,x2)","Alpha=.5 95% CI","Alpha=1 95% CI","Alpha=2 95% CI")
+    colnames(table.1.cd) <- c("n0=n1=n2","(x0,x1,x2)","a=.5 95% CI","a=1 95% CI","a=2 95% CI")
     
     print(xtable(table.1.cd),type="latex",file="hw3/pr1_cd.tex")
 
@@ -84,7 +85,132 @@ if (problem1) {
         CI2Low = leftTail[,3],
         CI2Hi = rightTail[,3]
         )
-    print(xtable(coverageData),type="latex",file="hw3/pr1_e.tex")
+    colnames(coverageData) <- c("(p0,p1,p2)",paste("a=",rep(c(.5,1,2),times=1,each=2),c(" th<LB"," UB<th"),sep=''))
+    
+    print(xtable(coverageData,digits=3),type="latex",file="hw3/pr1_e.tex")
 }
+
+### Problem 5
+if (problem5) {
+
+    # Call given code
+    source('official/exercises/seeb/train_test.R')
+    subpops <- unique(test$Population)
+    loci <- names(test)[3:26]
+    loci <- loci[!grepl("\\.1",loci)] # get rid of .1 version
+    alpha=2
+
+    ## part (a)
+    # Function to summarize frequency at locus by population (similar to trainc)
+    # We assume a dirchlet prior with each alpha identical
+    compute_freq <- function(data,locus,alpha){
+        counts <- table(data[,locus],data$Population) + 
+                    table(data[,paste(locus,".1",sep="")],data$Population) + alpha
+        return(counts/colSums(counts))
+    }
+
+    calculate_post <-function(alpha,train,test,subpops,loci){
[…]

### 2015-04-24 masters_paper/knockoffGLM.py
Starting on knockoff for logit in python

@@ -1,0 +1,18 @@
+""" This module is meant to implement and test the knockoff procedure on L1 regularized GLM models. The method was originally designed for linear least squares regression, and there is no theoretical guarantee it will work on other GLMS. Hopefully, with simulation we can discover if it breaks, if so where it breaks, and maybe how to fix it in those situations.
+
+More can be read about the original procedure on least squares linear models here: http://web.stanford.edu/~candes/Knockoffs/index.html
+
+"""
+
+# imports 
+import numpy as np
+import matplotlib.pyplot as plt
+import cvxpy as cvx
+from sklearn import linear_model as lm
+from sklearn.svm import l1_min_c 
+
+
+if __name__ == '___main__':
+    status = main()
+    sys.exit(status)
+

### 2015-04-29 stat302/hw2.tex

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 302 Homework 2 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[i)]
                If $h$ is the number of heads over 1000 flips and $p$ is the probability of any one flip resulting in a head, we can formalize this result as:
                \begin{align*}
                    \frac{P(h=527 \,|\, p = q \neq .5)}{P(h=527 \,|\, p = .5)} &\leq \max_{q\in(0,1)} \frac{P(h=527 \,|\, p = q)}{P(h=527 \,|\, p = .5)} \\
                                                                   &\leq \max_{q\in(0,1)} \frac{{1000 \choose 527}q^{527}(1-q)^{473}}{{1000 \choose 527}.5^{527}.5^{473}} \\
                                                                   &\leq \max_{q\in(0,1)} (2q)^{527}(2-2q)^{473}\\
                                                                   &\leq \left(2\frac{527}{1000}\right)^{527}\left(2-2\frac{527}{1000}\right)^{473}\\
                                                                   &\leq 4.30 \approx 4\\
                \end{align*}
            \item[ii)]
                We can calculate the largest posterior probability that the coin is weighted (to a particular value of $p$) rather than fair using Bayes rule and our previous result
                \begin{align*}
                    \max_{q\in(0,1)} P(p=q \,|\, h=527) &= \frac{P(p=q) P(h=527 \,|\, p=q)}{P(h=527)} \\
                                                        &= \frac{P(p=q) P(h=527 \,|\, p=q)}{P(p=q) P(h=527 \,|\, p=q)+P(p=.5) P(h=527 \,|\, p=.5)} \\
                                                    &= \frac{.5\frac{P(h=527 \,|\, p = q)}{P(h=527 \,|\, p = .5)}P(h=527 \,|\, p = .5)}{.5\frac{P(h=527 \,|\, p = q)}{P(h=527 \,|\, p = .5)}P(h=527 \,|\, p = .5)+.5P(h=527 \,|\, p = .5)} \\
                                                    &= \frac{4.30}{4.30+1} \\
                                                    &= .811
                \end{align*}
            \item[iii)]
                If one started with the prior that coin was far with probability $.5$ and that the true portion was actually the sample mean with probability $.5$, then the Bayes factor in favor of the alternate hypothesis is
                \begin{align*}
                    \frac{P(p =.527 \,|\, h=527)}{P(p =.5 \,|\, h=527)} &= \frac{P(h=527 \,|\, p = .527)P(p=.527)}{P(h=527 \,|\, p = .5)P(p=.5)} \\
                                                                &= \frac{P(h=527 \,|\, p = .527)}{P(h=527 \,|\, p = .5)} \\
                                                                &= 4.30 \\
                \end{align*}
                Which is the answer from part i). If we repeat this analysis with a prior from a Beta(10,10) distribution, we get a Bayes factor of:
                \[ \frac{P(h=527 \,|\, p = .527)P(p=.527)}{P(h=527 \,|\, p = .5)P(p=.5)} = 4.30\frac{f_{10,10}(.527)}{f_{10,10}(.5)} = 4.18 \]
                Then, with a Beta(1,1) distribution:
                \[ \frac{P(h=527 \,|\, p = .527)P(p=.527)}{P(h=527 \,|\, p = .5)P(p=.5)} = 4.30\frac{f_{1,1}(.527)}{f_{1,1}(.5)} = 4.30 \]
                And finally with a Beta(.5,.5) distribution:
                \[ \frac{P(h=527 \,|\, p = .527)P(p=.527)}{P(h=527 \,|\, p = .5)P(p=.5)} = 4.30\frac{f_{.5,.5}(.527)}{f_{.5,.5}(.5)} = 4.31 \]
                I prefer the Beta(10,10) prior, since that is the most heavily concentrated around $p=.5$, meaning that the prior says that the coin is mostly likely fair. This fits my intuition that it is less likely that a coin is heavily biased towards heads or tails than fair or close to fair.
        \end{itemize}
    \item[2.]
        \begin{itemize}
            \item[i)]
                \begin{align*}
                    P(X_i-1.96\leq\theta_i\leq X_i+1.96 \,|\, \theta_i) &= P(\theta_i-1.96\leq X_i\leq\theta_i+1.96) \\
                                                                    &= \Phi(X_i+1.96-\theta_i) - \Phi(X_i-1.96-\theta_i) \\
                                                                    &= .95
                \end{align*}
                So the interval contains $\theta_i$ $95\%$ of the time.
            \item[ii)]
                \begin{align*}
                    P(\theta_i \,|\, X_i, a) &= \frac{P(X_i \,|\, \theta_i, a) P(\theta_i \,|\, a)}{P(X_i \,|\, a)} \\
                                             &= \frac{\phi(X_i-\theta_i) \frac{I_{[-a,a]}(\theta_i)}{2a}}{\int_{-a}^a \phi(X_i-\theta_i) \frac{1}{2a}d\theta} \\
                    &= \frac{\phi(X_i-\theta_i)}{\Phi(X_i+a)-\Phi(X_i-a)}I_{[-a,a]}(\theta_i)
                \end{align*}
                As $a\to \infty$, \(\Phi(X_i+a)-\Phi(X_i-a)\to 1\), so \(P(\theta_i \,|\, X_i, a)\to \phi(X_i-\theta_i)\). Thus
                \[ P(\theta_i \in x \pm 1.96 \,|\, X_i = x,a) \to \Phi(x + 1.96) - \Phi(x - 1.96) = .95 \]
                Making \(x\pm 1.96\) a  $95\%$ credible interval in the limit. 
            \item[iii)]
                Going through the posterior calculations,
                \begin{align*}
                    P(\theta_i \,|\, X_i) &= \frac{P(X_i \,|\, \theta_i) P(\theta_i)}{P(X_i)} \\
                                          &= \frac{\phi(X_i-\theta_i) \phi(\theta_i)}{\int_\R \phi(X_i-\theta_i) \phi(\theta_i)d\theta_i} \\
                \end{align*}
                Which, can be simplified using the fact that
                \begin{align*}
                    \phi(X_i-\theta_i) \phi(\theta_i) &= \frac{1}{2\pi}e^{-(X_i-\theta_i)^2 - \theta_i^2} \\
                                                      &= \frac{1}{2\pi}e^{-2\theta_i^2 - 2\theta_i X_i - X_i^2} \\
                                                      &= \frac{1}{2\pi}e^{-2(\theta_i- \frac{1}{2} X_i)^2 - \frac{1}{2}X_i^2} \\
                                                      &= \frac{1}{\sqrt{4\pi}}e^{- \frac{1}{2}X_i^2}\frac{1}{\sqrt{\pi}}e^{-2(\theta_i- \frac{1}{2} X_i)^2} \\
                                                      &= \phi\left(\frac{X_i}{\sqrt{2}}\right)\phi\left(\sqrt{2}\left(\theta_i-\frac{1}{2}X_i\right)\right) \\
                \end{align*}
                To give us that
                \begin{align*}
                    P(\theta_i \,|\, X_i) &= \frac{\phi\left(\frac{X_i}{\sqrt{2}}\right)\phi\left(\sqrt{2}\left(\theta_i-\frac{1}{2}X_i\right)\right)}{\int_\R \phi\left(\frac{X_i}{\sqrt{2}}\right)\phi\left(\sqrt{2}\left(\theta_i-\frac{1}{2}X_i\right)\right)d\theta_i} \\
                    P(\theta_i \,|\, X_i) &= \phi\left(\sqrt{2}\left(\theta_i-\frac{1}{2}X_i\right)\right) \\
                \end{align*}
                Thus, \( \left[-1.16+\frac{X_i}{2},\infty\right)\) Is a one sided $95\%$ CI for $\theta_i$. This is because 
                \begin{align*}
                    P\left(\theta_i\geq-1.16+\frac{X_i}{2}\right) &= 1-\Phi\left(\sqrt{2}\left(-1.16+\frac{1}{2}X_i-\frac{1}{2}X_i\right)\right) \\
                                                       &=  1-\Phi\left(-1.64\right) \\
                                                       &= .95
                \end{align*}
                Thus, we will say we are ``confident" that $\theta_i$ is positive when \( X_i\geq 2.32\). Thus, we can calculate the portion of the time we are confident that $\theta_i>0$ but are actually incorrect as
                \begin{align*}
                    P(\theta_i<0 \,|\, X_i \geq 2.32) &= \frac{P(x_i \geq 2.32, \theta_i<0)}{P(x_i \geq 2.32)} \\
                                                      &= \frac{\int_{-\infty}^0 \Phi(\theta_i-2.32)\phi(\theta_i)d\theta_i}{\int_{-\infty}^\infty \Phi(\theta_i-2.32)\phi(\theta_i)d\theta_i} \\
                                                      &= .025
                \end{align*}
                The last value is achieved by numerical integration. When I test this by simulation, drawing $\theta_i$ and then $X_i$ for $n=100,000$, I found that
                \[ \frac{|\{X_i>2.32 \,\&\, \theta_i<0\}|}{|\{X_i>2.32\}|} = \frac{129}{5024} = .026 \]
                Confirming my result
            \item[iv)]
                Part ii) confirms, as discussed in class, that a confidence interval can be viewed as the limit of a credible interval as the variance of a prior goes to infinity (assuming the support of the prior also covers the confidence interval). 
        \end{itemize}
    \item[3.]
        \begin{itemize}
            \item[i)]
                \begin{align*}
                    P(\pi \,|\, X_i,...,X_n) &\propto P( X_i,...,X_n \,|\, \pi )P(\pi) \\
                                             &\propto \pi^{\sum_{i=1}^n X_i}(1-\pi)^{n-\sum_{i=1}^n X_i}\pi^{\alpha-1}(1-\pi)^{\beta-1} \\
                                             &\propto \pi^{\alpha + \sum_{i=1}^n X_i - 1}(1-\pi)^{\beta+n-\sum_{i=1}^n X_i - 1} \\
                                             &\sim \mathrm{Beta}\left(\alpha + \sum_{i=1}^n X_i, \beta+n-\sum_{i=1}^n X_i\right)
                \end{align*}
                Thus, we can derive the distribution for the prediction of $X_{n+1}$
                \begin{align*}
                    P(X_{n+1}=1 \,|\,  X_i,...,X_n) &= \int_{0}^1 P(X_{n+1}=1 \,|\,\pi, X_i,...,X_n)P( \pi\,|\, X_i,...,X_n)d\pi \\
                                                    &= \int_{0}^1 P(X_{n+1}=1 \,|\,\pi)P( \pi\,|\, X_i,...,X_n)d\pi \\
                                                    &=\int_0^1 \pi\pi^{\alpha + \sum_{i=1}^n X_i - 1}(1-\pi)^{\beta+n-\sum_{i=1}^n X_i - 1}d\pi \\
                                                    &= \frac{\alpha + \sum_{i=1}^n X_i}{\alpha + \sum_{i=1}^n X_i+\beta+n-\sum_{i=1}^n X_i} \\
                                                    &= \frac{\alpha + \sum_{i=1}^n X_i}{\alpha +\beta+n} \\
                \end{align*}
                So \(X_{n+1} \,|\,  X_i,...,X_n \sim \mathrm{Bernoulli}\left(\frac{\alpha + \sum_{i=1}^n X_i}{\alpha +\beta+n}\right)\).

            \item[ii)]
                \begin{align*}
                    P(\pi \,|\, X_i,...,X_n,w,\alpha,\beta) &\propto P( X_i,...,X_n \,|\, \pi )P(\pi \,|\, w,\alpha,\beta) \\
                                                            &\propto  \pi^{\sum_{i=1}^n X_i}(1-\pi)^{n-\sum_{i=1}^n X_i} \sum_{j=1}^k w_j\pi^{\alpha_j-1}(1-\pi)^{\beta_j-1} \\
                                                            &\propto  \sum_{j=1}^k w_j\pi^{\alpha_j+\sum_{i=1}^n X_i-1}(1-\pi)^{\beta_j+n-\sum_{i=1}^n X_i-1} \\
                                                            &\sim \sum_{j=1}^k w_j \mathrm{Beta}\left(\alpha_j+\sum_{i=1}^n X_i,\beta_j+n-\sum_{i=1}^n X_i-1\right)\\
                \end{align*}
                Giving us a mixture of Betas for the posterior. The posterior mixing weights have remained the same, but the posterior hyperparameters have been updated as they would normally be for a Beta prior.
        \end{itemize}
    \item[4.]
        \begin{itemize}
            \item[(a)] I swapped code with Jonathan Eskreiswinkler.
            \item[(b/c)] To start with, we seemed to have a large discrepancy - I had an accuracy around $77\%$ and he had an accuracy around $84\%$. The difference turned out to be a few bugs on both our parts. \\
                \begin{itemize}
                    \item I accidentally was dropping observations which were missing the allele at a particular locus \\
                    \item For my likelihood, I was using P(Population $|$ Allele) rather than P(Allele $|$ Population) \\
                    \item Jon was multiplying the population likelihoods by the wrong value for a few locus due to a bug in matching the appropriate values \\
                    \item Jon accidentally multiplied by the likelihood for the first version of the locus twice, rather than the first and the second version (Ogo2 and Ogo2.1 for example).
                \end{itemize}
                In the end, both my version of the code and Jon's with my fixes had an accuracy of $78.0\% $.
            \item[(d)] Besides the bugs, both Jon and I attempted to do the theoretical aspects identically. In terms of programing though, we had somewhat different approaches. I essentially had two arrays, one with all the probabilities from the training data set, and one with all the test data and associated priors, likelihoods, posteriors, and predictions. Jon generally stored these in separate arrays. Also, I tried to preform my calculations over the whole array at the same time to the extent possible, while Jon looped over each individual entry to do calculations. This made his code somewhat simpler, but mine is much more compact and runs quicker. I find my version a little easier to read, but then again, I wrote it. I didn't learn any R tricks from Jon. \\

        \end{itemize}
    {\bf Improved Code:}
\end{itemize}
    \lstinputlisting[firstline=60,lastline=114]{hw1.R}

\end{document}

### 2015-04-29 stat343/hw3.tex

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Homework 3 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[a)]
                The columns of $X$ are linearly independent, so \(rank(X)=p\). This in turn dictates that \(rank(QR)=p\). Since the rank of the product of two matrices can be no greater than the smallest rank of the factors, we must conclude that \(rank(Q)=rank(R)=p\), since they both can't have greater rank than $p$. Since $R$ has full rank, it must be invertible.
            \item[b)]
                Multiplying both sides of the linear system through by $Q$, we get that
                \begin{align*}
                    R \tilde\beta &= Q^TY \\
                    QR \tilde\beta &= QQ^TY \\
                    X\tilde\beta &= Y \\
                \end{align*}
                So we may deduce that $Y$ is in the span of $X$. Thus, for $H=X(X^TX)^{-1}X^T$, $HY=Y$, since $HZ=Z$ for all vectors $Z$ in the span of $X$. Starting with this, we get that
                \begin{align*}
                    HY &=Y \\
                    X(X^TX)^{-1}X^TY &= Y \\
                    X \hat\beta &= Y
                \end{align*}
                Since, $\tilde\beta$ is unique, we may conclude that $\tilde\beta=\hat\beta$.
            \item[c)]
                The "normal equation" to calculate $\hat\beta$ ran into a singularity at $\epsilon=1e-5$, while the QR method of calculation still worked. The R code used for this calculation is attached in the back.
            \item[d)]
                The QR method and the linear model returned the following results for $\epsilon=1e-5$: \\
                \begin{tabular}{l | l l}
                    Coefficient & QR Method & Linear Model  \\
                    \hline
                    Intercept           &  2.862e+1    &   2.862e+1\\
                    pop15               & -4.620e-1    &  -4.620e-1\\
                    pop75               & -1.151e+4    &  -1.151e+4\\
                    dpi                 & -3.356e-4    &  -3.356e-4\\
                    ddpi                &  4.088e-1    &   4.088e-1\\
                    pop75 + $u\epsilon$ &  1.151e+4    &   1.151e+4
                \end{tabular} \\
                It is not surprising that they returned the same estimates, since we were told that the linear model in R is using the QR method of calculation. The one surprising result is that the two pop75 variables had such big coefficients, but this is due to them being highly collinear, which is why their coefficients added together are almost $0$.  
        \end{itemize}
    \item[2.]
        %Since $A$ and $B$ are both positive semidefinite and symmetric, we can decompose both using Cholesky decomposition. Let $K$ and $L$ be lower triangular matrices such $A=KK^T$ and $B=LL^T$. This gives us $Y^TKK^TY=\|Y^TK\|^2$ and $Y^TLL^TY=\|L^TY\|^2$. These two are independent if $L^TY$ and $Y^TK$ are independent. Since each of these variables is a linear combination of the random vector $Y$, which has a multivariate normal distribution, these two sets of variables are pairwise jointly normally distributed. Thus, all it takes to prove normality is that their covariance is $0$. 
        %\[ \cov(Y^TK \]

        $AY$ and $BY$ are jointly normally distributed, since they are both linear combinations of $Y$, which is drawn from a multivariate normal distribution. We can also show that their covariance is $0$:
        \begin{align*}
            \cov[AY,BY]&=\E[AY(BY)^T] \\
                       &=\E[AYY^TB^T] \\
                       &=\E[\|Y\|^2AB] \\
                       &=\E[0] \\
                       &=0
        \end{align*}
        Since $AY$ and $BY$ are jointly normally distributed, this is sufficient to prove their independence. Now, since $YAY$ and $YBY$ are merely linear combinations of the variables in $AY$ and $BY$, which are pairwise independent, $YAY$ and $YBY$ must also be linearly independent.

    \item[3.]
        Since $A$ is symmetric, we can eigendecompose it, so that $A=PDP^T$, where $D$ is a diagonal matrix of $A$'s eigenvalues, and $P$ is an orthogonal matrix of $A$'s eigenvectors. It is the case that, where $L$ is the set of all eigenvalues of $A$:
        \[ p=tr(A)=tr(D)=\sum_{\lambda\in L} \lambda \]
        Since all of $A$'s eigenvalues are either $0$ or $1$, this means that $p$ is the number of non-zero eigenvalues of $A$. Also,
        \begin{align*}
            e^TAe &= tr(e^TAe) \\
                  &= tr(e^TPDP^Te) \\
                  &= tr(e^TDP^TPe) \\
                  &= tr(e^TDe) \\
                  &= e^TDe  \\
                  &= \sum_{i=1}^n e_i^2 I_{\{j:d_{jj}=1\}}(i) \\
        \end{align*}
        The last line is based on the fact that the eigen values of $A$, which are also the diagonal of $D$, are all either $0$ or $1$. Clearly, we will have $p$ terms in our summand, since that is how many eigenvalues of $1$ there are. Since each $e_i$ is independent of all other $e_j$, this means $e^TAe$ is the sum of $p$ independent squared normals, or in other words
        \[ e^TAe \sim \chi_p^2\]

    \item[4.]
        \begin{align*}
            \cov(\hat e, \hat Y) &= \E\left[\hat e \hat Y^T\right] - \E\left[\hat e\right] \E\left[\hat Y\right]^T & \\
                                 &= \E\left[0_{n\times n}\right] - 0_{n \times 1} Y^T          & \mbox{Since } \hat e \mbox{ and } \hat Y \mbox{are orthogonal} \\
                                 &= 0_{n \times n}
        \end{align*}
    \item[5.]
        \begin{itemize}
            \item[3.1]
                \begin{itemize}
                    \item[a)]
                        After running the model and getting the coefficient for age and its standard error, the confidence intervals are:
                        \[ 90\% \mbox{ CI} = \beta_{age} \pm SE \cdot  t_{rdf}^{.05} = -.0196 \pm .0112 \cdot 1.6624 = (-.0382,-.0012) \]
                        \[ 95\% \mbox{ CI} = \beta_{age} \pm SE \cdot  t_{rdf}^{.025} = -.0196 \pm .0112 \cdot 1.9873 = (-.0418,0026) \]
                        Since the $90\%$ CI doesn't include $0$, but the $95\%$ CI does, we can conclude the pvalue for $\beta_{age}$ is between $.05$ and $.1$.
                    \item[c)]
                        The simulation, the code for which is in my code supplement, run on $1000$ permutations of lpsa, had $8.8\%$ of the results with a higher t value for $\beta_{age}$.
                    \item[d)]
                        Running the F test with the null hypothesis that a model just with lcavol, lweight, and svi is the real model, one would expect to see at least as large an improvement from adding four variables as we saw by adding lbph, age, lcp, and gleason $21.67\%$ of the time. Thus , we can not reject the null hypothesis, and prefer the simpler model.
                \end{itemize}
            \item[3.2]
                \begin{itemize}
                    \item[a)]
                        Running a linear model predicting taste in terms of the other three variables, only H2S and Lactic are statistically significant. 
                    \item[b)]
                        Transforming Acetic and H2S back from the log scale, H2S ceases to be statistically significant, Acetic remains insignificant, and Lactic remains significant. 
                \end{itemize}

        \end{itemize}
\end{itemize}

\newpage
\hrule
\vspace{2mm}
R Code Supplement
\vspace{2mm}
\hrule
\lstinputlisting{hw3.R}

\end{document}

### 2015-04-29 stat347/hw2.tex

--- 2015-04-21 stat302/hw3.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-STAT 302 Homework 3 \hfill Aaron Maurer
+
+STAT 347 Homework 2, Part 1 \hfill Aaron Maurer
 \vspace{2mm}
@@
 \vspace{2mm}
+\begin{itemize}
+    \item[A1:]
+        Filling out the table, I got:
+        \FloatBarrier
+        \input{hw2/pr1}
+        \FloatBarrier
+        Overall, the majority of the variation not accounted for by the mean can be explained by the Fac(T) space, even though the other variables can account for part of this total variation too.
+We can see this by noting that (if $S$ is the entire space)
+        \[ SS_{T/1} > SS_{t/1} + SS_{R/1}+ SS_{(t:r)/(T+R+r)} \geq SS_{S/T} \]
 
+        Additionally we notice the spaces defined by each of three variables are not orthogonal to each other; otherwise, the sum of squares in the space defined by each would be identical to the quotient space with the other two variables removed.
+Additionally, a significant amount of the covariation between each subspace and the Prothrombin level is due to the intersection of the three subspaces.
+This said, each quotient space still accounts for its own covariation with Prothombin level.
+Finally, there is variation unaccounted for by the joint span of the three subspaces due to each variable, as indicated by the variation of the quotient space of the interaction space and the span of the three variable spaces.
+    \item[A2:]
+        Let us say the number of rows is $R$ \\
         \begin{itemize}
-    \item[1.]
-        \begin{itemize}
-            \item[(a)]
-                \begin{align*}
-                    P(D \st P) &= \frac{P(P\st D)P(D)}{P(P\st D)P(D) +  P(P\st D^c)P(D^c)} \\
-                               &= \frac{p_0p_1}{p_0p_1 +  (1-p_0)p_1} \\
-                \end{align*}
-            \item[(b)]
+            \item[(i)] 
+                The subspace will have dimension $R$.
+                Algebraically this can be expressed as \(\E(Y_{ij})=\alpha_i+\alpha_b\). 
+                This might be used to model a variable assigned to pairs, where order doesn't matter and each member contributes an amount independently.
+For instance, it might be used to model the energy released by burning one ounce of one fuel and one ounce of another fuel.
+            \item[(ii)] 
+                The subspace will have dimension $R$ and be orthogonal to the subspace in the prior problem.
+Algebraically it can be expressed as \(\E(Y_{ij})=\alpha_i-\alpha_b\).
+This might be used to model the difference between the first item in a pair and the second.
+For instance, this was the appropriate model for the voltage difference between two metals in an electrolytic solution.
+            \item[(iii)] This subspace will have dimension ${R \choose 2}$.
+Algebraically, it can be expressed as $\E(Y_{ij})=\alpha_{ij}$, where $\alpha_{ij}=\alpha_{ji}$.
+This might be used to model an effect which is unique to particular unordered pairs.
+For instance, one might use this model the energy released from the reaction of one ounce of one chemical mixed with one ounce of another chemical.
+            \item[(iv)] This subspace will have dimension ${R \choose 2}$.
+Algebraically, it can be expressed as $\E(Y_{ij})=\alpha_{ij}$, where $\alpha_{ij}=-\alpha_{ji}$ when $i\neq j$.
+This might be used to model an effect which is unique for particular pairs, with the sign flipped depending on the order.
+For instance, this might be used to model the difference in score between the first player to move and the second player to move in a game of go (we'll say someone can play themselves).
         \end{itemize}

### 2015-04-30 stat347/hw2b.tex
finished homework 2b

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\sd}{\mathrm{SD}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 347 Homework 2 Part 2 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}
\begin{itemize}
    \item[A0.]
        In this paper, the authors attempt to measure whether raising two sets of flies on different feed for several generations results in the two groups being less likely to mate with each other. To test this, the researchers performed two experiments. For both experiments, the distinct populations were raised on the respective feed for a number of generations. Then, a new generation from each population were both raised on one feed. Finally, a number of different wells were filled with one male and one female fly from the final generation of each population. The mating in each well was observed, and the number of pairings was tallied by male and female population. In the first experiment, these counts were summed across all the wells for one generation. In the second experiment, this process was repeated after different numbers of generations had been raised on the two feed, and the results were summed across wells and generations. \\
        \vspace{-2mm} \\
        To analyze this statistics, the researchers assumed that a mating represented an independent Bernoulli random draw of whether the mating occurred between flies of the same population. Thus, for $n_h$, the number of matings with both members from the same population, and $p$, the probability of homogamic mating given mating, $n_h\sim Bin(n,p)$. They transformed this into a statistic 
        \[ SII = \frac{n_h - (n-n_h)}{n} = \frac{2n_h}{n} - 1\]
        Then, they calculated its standard error based on the binomial assumption:
        \begin{align*}
            \sd(SII) &= \sd(\frac{2n_h}{n}) \\
                     &= \sqrt{\frac{2}{n}\frac{n_h}{n}\left(1-\frac{n_h}{n}\right)} \\
                     &= \sqrt{\frac{1}{n}\left(1-\left(\frac{n_h - (n-n_h)}{n}\right)^2\right)} \\
                     &= \sqrt{\frac{1}{n}\left(1-SII^2\right)} \\
        \end{align*}
        For the first experiment, with $n_h=29, n=38$ they calculated an SII of $.53$ with a standard error of $.14$. Also, they attempted to calculate a p-value as $P(n_h\geq 29 \st p =.5)$, which should have been $.00083$, but which they reported as $.0012$. In the second experiment, $n=900, n_h=571$, yielding a SII of $.27$, a SE of $.02$, and a p-value less than $.0001$. They accordingly concluded that the different diets reduced inter-mating. \\
        \vspace{-2mm} \\
        The biggest issue with this analysis was the binomial assumption. The fact that there was only two males and two females in each well prevented independence; while one pair of flies is mating, the other two can only possibly be mating with each other. This means that two homogamic matings or two heterogamic matings would be more common than one of each in a well (indeed, depending on how they counted mating, which was not clear, it may be impossible). This means that their confidence intervals and p-values are invalid. \\
        \vspace{-2mm} \\
        Other set of potential issues stem from the pooling of data across generation in the second experiment. Its questionable whether SII is constant across the generations. When I fit a quadratic regression by least squares, our p-value from an F test is $.1534$, but as you see below, it does appear to be a better fit than just a constant. 
        \begin{center}
            \includegraphics[width=10cm]{hw2/A0_SII} 
        \end{center}
        Thankfully, since flies which were placed in wells to mate were permanently removed from the population, there is little concern of dependence among mating between generation.\\
        \vspace{-2mm} \\
        The one remaining issue then is whether the last three sets of tests in the second experiment, which were from a parallel pair of populations, is comparable to the rest of the data set. If we look at the number of matings per well, the answer seems to be no. As one would expect, the total number of matings grows linearly with wells. However, the last three test groups do not seem to fit the pattern from the first 15:
        \begin{center}
            \includegraphics[width=10cm]{hw2/A0_wells} 
        \end{center}
        When we test for the probability of seeing the mean of the three points fall so far away from the prediction under the linear model, we get a p-value of $.013$, which is sufficient to reject that they follow the same distribution. (I used the usual formula for prediction distribution of a linear model taking into account the smaller standard error for the mean of three observations).
    \item[A3.]
        Fitting a logistic model to the data, we get this model:
        \FloatBarrier
        \input{hw2/A3_logit}
        \FloatBarrier
        Superimposed over the data, it seems to be a very good fit:
        \begin{center}
            \includegraphics[width=10cm]{hw2/A3_plot1} 
        \end{center}
        We can obtain an estimate of $\log_2LD_{50}$ by solving this equation:
        \begin{align*}
            .5 &= (1+e^{-1.01x+2.54})^{-1} \\
            0  &= -1.01x+2.54 \\
            x  &= 2.51 \\
        \end{align*}
        Where $\hat\alpha$ is in the intercept and $\hat\beta$ is the coefficient on $x$, we want a confidence interval on $-\frac{\hat\alpha}{\hat\beta}$, which will set the linear predictor to $0$ and yield $\log_2LD_{50}$. Thus, we can apply Fieller's method.
        \begin{align*} 
            1.96^2 &> \frac{(\hat\alpha - x\hat\beta)^2}{\hat\sigma_\alpha^2 - 2x\hat\sigma_{\alpha,\beta} + x^2\sigma_\beta^2} \\
            U,L &= \frac{(\hat\alpha\hat\beta - 1.96^2\hat\sigma_{\alpha,\beta}) \pm \sqrt{(\hat\alpha\hat\beta - 1.96^2\hat\sigma_{\alpha,\beta})^2 - (\hat\beta^2 - 1.96^2\sigma_{\hat\beta})(\hat\alpha^2 - 1.96^2\sigma_{\hat\alpha})}}{\hat\beta^2 - 1.96^2 \sigma_{\hat\alpha}} \\
            U&=-1.66, L=-3.31
        \end{align*}
        Since this is the confidence interval on $\frac{\hat\alpha}{\hat\beta}$ while we want $-\frac{\hat\alpha}{\hat\beta}$, our confidence interval for $\log_2LD_{50}$ is $(1.66,3.31)$. Its worth noting that this ignores the uncertainty in the covariance matrix. \\
        \vspace{-2mm} \\
        Fitting a model under the constraint that $\log_2LD_{50}=4$ means that the linear predictor must be $0$ at $x$. Thus, we fit the model $E(y) = logit(\beta(x-4))$. This yields the model
        \FloatBarrier
        \input{hw2/A3_logit4}
        \FloatBarrier
        Calculating the likelihood ratio statistic for this model as compared to the original model, we get a value of $11.59$, which corresponds to a p-value of $.00066$ assuming a $\chi^2_1$ distribution. Extending this concept, we might calculate a confidence interval set based on the $.95$ percentile of the $\chi^2_1$ distribution, including only values of $\log_2LD_{50}$ whose likelihood ratio statistic would fall below it. Indeed, this is plotted below, with the dotted line representing the cutoff:
        \begin{center}
            \includegraphics[width=10cm]{hw2/A3_lkl} 
        \end{center}
        The region which has log likelihoods above the cut off corresponds to $\log_2LD_{50}\in(1.75,3.21)$. This is pretty close to the confidence interval we calculated earlier.

\end{itemize}

\end{document}

### 2015-04-30 stat347/hw2b.R

@@ -1,0 +1,120 @@
+### Libraries
+library('xtable')
+library('stats4')
+library('locfit')
+library('SDMTools')
+library('plotrix')
+library('reshape2')
+library('nlme')
+library('parallel')
+library('faraway')
+
+### callers
+problem0 <- FALSE 
+problem3 <- TRUE
+
+### A0
+if (problem0) {
+    #Mating counts used in the PNAS 2010 paper by
+    #Sharon, Segal, Ringo, Hefetz, Zilber-Rosenberg and Rosenberg
+    #"Commensal bacteria play a role in mating preference of Drosophila melanogaster"
+    #Data provided by the authors on request
+    # First column is generation no 
+    # Last four columns are mating counts by type CxC CxS SxC SxS
+    # Last three rows taken from a separate expt of a similar design)
+    matings <- matrix(c(
+    2, 12, 8, 9, 16,
+    6, 10, 5, 9, 10,
+    7, 17, 9, 9, 15,
+    9, 8, 7, 7, 9,
+    10, 18, 13, 5, 12,
+    11, 12, 5, 7, 14,
+    13, 14, 9, 8, 12,
+    15, 18, 9, 7, 15,
+    16, 14, 5, 5, 10,
+    17, 31, 22, 12, 27,
+    20, 23, 13, 10, 20,
+    21, 13, 7, 5, 14,
+    26, 30, 19, 12, 21,
+    31, 9, 7, 3, 10,
+    37, 20, 14, 11, 17,
+    11, 18, 11, 7, 16,
+    12, 16, 11, 8, 15,
+    13, 22, 13, 8, 13), ncol=5, byrow=T)
+    colnames(matings) <- c("gen", "CxC", "CxS", "SxC", "SxS")
+    
+    wells <- c(NA,NA,NA,24, 39, 20, 24, 36, 23, 70, 46, 24, 45, 23, 48, 48, 48, 48)
+    totmat <- apply(matings[,2:5],1,sum)
+    SII <- (matings[,2]+matings[,5]-matings[,3]-matings[,4])/totmat
+
+    matings <- data.frame(cbind(wells,matings,totmat,SII))
+
+    genpoly = poly(matings$gen,2)
+    genlm <- lm(matings$SII~genpoly)
+    xseq <- seq(0,40,length.out=500)
[…]

### 2015-05-01 masters_paper/knockoffGLM.py
added in counter for which variables entered, some things to analyze output

@@ -104,23 +104,17 @@ def find_w(y,X,X_ko):
 
     X_lrg = np.concatenate((X,X_ko), axis=1)
 
-    clen = p*10
-    cs = l1_min_c(X_lrg, y, loss='log') * 10 * np.logspace(0, 3,clen)
-    clf = lm.LogisticRegression(C=1.0, penalty='l1', tol=1e-6)
-    coefs_ = []
-    for c in cs:
-        clf.set_params(C=c)
-        clf.fit(X_lrg, y)
-        coefs_.append(clf.coef_.ravel().copy())
-
+    clen = int(np.log(p)*50)
+    coefs_,cs = lm.logistic_regression_path(X_lrg,y, penalty='l1', tol=1e-6,fit_intercept=False,solver='liblinear',Cs=clen)
     coefsMat = np.vstack(coefs_)
     coefsMat[abs(coefsMat)<1e-6] = 0
     Z   = np.zeros(2*p) 
     for i in range(clen):
         Z[np.all(np.vstack((Z==0,coefsMat[i,:]!=0)),axis=0)] = cs[i]
     w = np.min((Z[:p],Z[p:2*p]),axis=0) * np.sign(Z[p:2*p]-Z[:p])
+    ent = np.min((Z[:p],Z[p:2*p]),axis=0)>0
 
-    return w 
+    return w, ent
 
 def generate_logit_w(y,X,knockoff='equicor',randomize=False):
     """ Generates the w statistic for a logistic model predicting y~X"""
@@ -134,15 +128,32 @@ def generate_logit_w(y,X,knockoff='equicor',randomize=False):
 
     return find_w(y,X,X_ko)
 
-def iid_normal_null_test(n,p,q):
-    """ Test run with n observations. p covariates, each iid Normal. Base rate q"""
-    y = npran.binomial(1,.2,n)
-    X = npran.randn(n,p)
-    w = generate_logit_w(y,X)
-    print '%.2f of null knockoffs came in first; %.2f of variables never entered' % (np.sum(w>0)/float(np.sum(w!=0)),np.sum(w==0)/float(p))
+def analyze_knockoff(X_1,X_null,y):
+    if X_1 == None:
+        n,p = X_null.shape
+        p0,p1 = p,0
+        X = X_null
+    else:
+        n,p0 = X_null.shape
+        n,p1 = X_1.shape
+        p = p0+p1
+        X = np.concatenate((X_1,X_null),axis=1)
+    w,ent = generate_logit_w(y,X)
[…]

### 2015-05-13 stat302/hw4.tex
Finished problem 1, hw4

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\newcommand{\hwhead}[1]{#1 \hfill Aaron Maurer \vspace{2mm} \hrule \vspace{2mm}}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hwhead{STAT 302 Homework 4}
\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[i)]
                Together, (1) and (2) define a hierarchical model since they relate observed values $D_{ij}$ drawn independently based on parameters $\beta_i$ to a set of hyperparameters $\pi_0$ and $\sigma_b$ in such a way that the $D_{ij}$ can provide inference on each other through the hyperparameters. In other words, even though the observations are conditionally independent, they are still informative on each other through the hyperparameters.Since the $\beta_j$ are iid given the hyperparameters, they are necessarily exchangeable. Not conditioned on $\pi_0, \sigma_b$, the $\beta_j$ should still be exchangeable assuming the same value of $\pi_0$ and $\sigma_b$ are used to draw each $\beta_j$; there is just additional variability in the initial draw of the hyperparameters. $D_{ij}\st \beta, \sigma$ will be exchangeable for all $i$ since they are iid, but $D_{ij}\st \beta_j$ won't be exchangeable across $j$ since \(P(D_{ij}\st \beta_j=a,D_{ij'}\st \beta_{j'}=b)=P(D_{ij'}\st \beta_{j'}=a,D_{ij}\st \beta_j=b)$ only if $\beta_{j'}=\beta_j$. Finally, the $D_{ij}$ should be exchangable accross $i$ and $j$ if unconditioned, or conditioned on $\pi,\sigma_b$, since then they will be iid.
            \item[ii)]
                This command will calculate $\beta$, $D$, and also the p-values.
                \lstinputlisting[firstline=18,lastline=28]{hw4.R}
            \item[iii)] I calculated the usual two sided t-test p-value using the R command. They were distributed as such: \\
                \includegraphics[width=9cm]{hw4/pr1_iii_hist_a}
                \includegraphics[width=9cm]{hw4/pr1_iii_hist_b} \\
                \includegraphics[width=9cm]{hw4/pr1_iii_hist_c} \\
                As one would expect, for the first one where the $\beta$ are all $0$, the p-values are uniformly distributed. In the second, with $\pi_0=.5$, we see a big lump of low p-values near $0$ corresponding to most of the $\beta_j$ which are non-zero, and a approximately uniform distribution over the rest of the p-values corresponding to the $\beta_j$ which are $0$. Finally, in the last histogram, where none of the $\beta_j$ are $0$, we see overwhelmingly low p-values corresponding to the strong evidence against $\beta_j=0$.
            \item[iv)] This is my function to apply Benjamini-Hochberg:
                \lstinputlisting[firstline=44,lastline=50]{hw4.R}
            \item[v)] This is my function to computer the empirical FDR (and pFDR):
                \lstinputlisting[firstline=53,lastline=58]{hw4.R}
            \item[vi)]
                To simulate $\E(V/R)$, I generated the dataset D $1000$ times for each scenario, and averaged the emperical FDR for each $alpha$ over the data sets. The results are ploted below: \\
                \includegraphics[width=9cm]{hw4/pr1_vi_hist_a}
                \includegraphics[width=9cm]{hw4/pr1_vi_hist_b} \\
                \includegraphics[width=9cm]{hw4/pr1_vi_hist_c} \\
                In the first scenario, despite the fact that all the discoveries are false, our FDR statistic is less than 1, which is due to the definiton. In any event, The FDR rate is properly controlled, coming in under $\alpha$. For the second scenario, the FDR rate seems to come in at almost exactly half of $\alpha$ for all levels, providing conservative control. For the final scenario, since all $\beta\neq0$, the FDR is necessarily $0$, so FDR is inherently controlled.  
            \item[vii)] I repeated the same procedure as above, but this time caclulated the pFDR. The results are below:
                \includegraphics[width=9cm]{hw4/pr1_vii_hist_a}
                \includegraphics[width=9cm]{hw4/pr1_vii_hist_b} \\
                \includegraphics[width=9cm]{hw4/pr1_vii_hist_c} \\
                In the first case, with all discoveries false, the pFDR is naturally at 1 for all $\alpha$ and not controlled. The second case is nearly identical as with FDR, controlled at about half of $\alpha$. Again, for the final case, since no discoveries are false, the pFDR is controlled perfectly at 0.

        \end{itemize}
\end{itemize}
\end{document}

### 2015-05-13 stat302/hw4.R

@@ -1,0 +1,108 @@
+### Libraries
+library('xtable')
+library('stats4')
+library('locfit')
+library('SDMTools')
+library('plotrix')
+library('reshape2')
+library('nlme')
+library('parallel')
+library('faraway')
+
+### Caller
+problem1 <- TRUE
+
+### Problem 1
+if (problem1){
+    ## ii - simulate simple hierarchical model
+    p.values <- function(x){
+        return(t.test(x)$p.value)
+    }
+
+    simulateD <- function(pi0,sigb,n=10,m=1000,sigj=1){
+        output <- list()
+        output$beta     <- ifelse(rbinom(m,1,pi0),0,rnorm(m,0,sigb))
+        output$D        <- replicate(n,rnorm(m,output$beta,sigj))
+        output$p.values <- apply(output$D,1,p.values)
+        return(output)
+    }
+   
+    ## iii - generate simple p-values
+    scenarios <- list(c(1,1),c(.5,3),c(0,3))    # each is a pi0, sigb pair
+    simulations <- list()
+    for (i in 1:3){
+        pi0  <- scenarios[[i]][1]
+        sigb <- scenarios[[i]][2]
+        simulations[[i]] <- simulateD(pi0,sigb)
+
+        pdf(paste('hw4/pr1_iii_hist_',letters[i],'.pdf',sep=''))
+        hist(simulations[[i]]$p.values,breaks=20,freq=F,main=bquote("Histogram, "~pi[0] == .(pi0)~','~ sigma[b]== .(sigb)),xlab='p-value',ylab='Density')
+        dev.off()
+    }
+
+    ## iv - implement Benjamini-Hochberg
+    BH <- function(pvals,alpha){
+        m    <- length(pvals)
+        ord  <- order(pvals)
+        test <- pvals[ord]< (1:m)*alpha/m
+        test[ord] <- (1:m)<Position(identity,test,right=T,nomatch=0)
+        return(test)
+    }
+
+    ## v - implement FDR
+    FDR <- function(beta,gamma,pFDR=FALSE){
+        R <- sum(gamma)
+        V <- sum(gamma & beta==0)
+        FDR <- ifelse(R>0,V/R,ifelse(pFDR,NA,0))
+        return(FDR)
+    }
+
+    ## vi - simulate actual FDR
+    k <- 1000
[…]

### 2015-05-13 stat302/hw4.tex
Finished problem 2, hw4

@@
                 To simulate $\E(V/R)$, I generated the dataset D $1000$ times for each scenario, and averaged the emperical FDR for each $alpha$ over the data sets. The results are ploted below: \\
-                \includegraphics[width=9cm]{hw4/pr1_vi_hist_a}
-                \includegraphics[width=9cm]{hw4/pr1_vi_hist_b} \\
-                \includegraphics[width=9cm]{hw4/pr1_vi_hist_c} \\
+                \includegraphics[width=9cm]{hw4/pr1_vi_a}
+                \includegraphics[width=9cm]{hw4/pr1_vi_b} \\
+                \includegraphics[width=9cm]{hw4/pr1_vi_c} \\
                 In the first scenario, despite the fact that all the discoveries are false, our FDR statistic is less than 1, which is due to the definiton.
@@
             \item[vii)] I repeated the same procedure as above, but this time caclulated the pFDR. The results are below:
-                \includegraphics[width=9cm]{hw4/pr1_vii_hist_a}
-                \includegraphics[width=9cm]{hw4/pr1_vii_hist_b} \\
-                \includegraphics[width=9cm]{hw4/pr1_vii_hist_c} \\
+                \includegraphics[width=9cm]{hw4/pr1_vii_a}
+                \includegraphics[width=9cm]{hw4/pr1_vii_b} \\
+                \includegraphics[width=9cm]{hw4/pr1_vii_c} \\
                 In the first case, with all discoveries false, the pFDR is naturally at 1 for all $\alpha$ and not controlled.
@@
 Again, for the final case, since no discoveries are false, the pFDR is controlled perfectly at 0.
+        \end{itemize}
+    \item[2.]
+        \begin{itemize}
+            \item[i)]        
+                Repeating the same procedure as above, except with reject decisions based on q values, I get these results: \\
+                \includegraphics[width=9cm]{hw4/pr2_i_FDR_a}
+                \includegraphics[width=9cm]{hw4/pr2_i_pFDR_a} \\
+                In this first scenario, the FDR seems to be controlled at pretty much exactly the desired rate, while the pFDR is still necessarily 1.\\
+                \includegraphics[width=9cm]{hw4/pr2_i_FDR_b}
+                \includegraphics[width=9cm]{hw4/pr2_i_pFDR_b} \\
+                In this second scenario, we have control of both FDR and pFDR at the desired rate.\\
+                \includegraphics[width=9cm]{hw4/pr2_i_FDR_c}
+                \includegraphics[width=9cm]{hw4/pr2_i_pFDR_c} \\
+                With no false discoveries, FDR and pFDR are still 0.
+            \item[2)]
+                I have plotted the estimated $\pi_0$ vs actual $\pi_0$ over a range of values for $\sigma_b=.5,1,2,4$. Each estimate is the average over 100 simulations. \\
+                \includegraphics[width=9cm]{hw4/pr2_ii_pt5}
+                \includegraphics[width=9cm]{hw4/pr2_ii_1}\\
+                As we can see above and below, the estimates get increasingly good as the variance goes up and the $\beta$ tend to fall farther from $0$.
+The predictions are uniformly higher than the actual $\pi_0$, except for values near 1, and the worst predictions are for low values of $\pi_0$.
+In general, this method seem to provide a pretty solid estimate.
+\\
+                \includegraphics[width=9cm]{hw4/pr2_ii_2}
+                \includegraphics[width=9cm]{hw4/pr2_ii_4}\\
 

### 2015-05-13 stat302/hw4.R

@@ -8,25 +8,27 @@ library('reshape2')
 library('nlme')
 library('parallel')
 library('faraway')
+library("qvalue")
 
 ### Caller
-problem1 <- TRUE
+problem1 <- FALSE 
+problem2 <- TRUE
 
 ### Problem 1
-if (problem1){
-    ## ii - simulate simple hierarchical model
-    p.values <- function(x){
-        return(t.test(x)$p.value)
-    }
+## ii - simulate simple hierarchical model
+p.values <- function(x){
+    return(t.test(x)$p.value)
+}
 
-    simulateD <- function(pi0,sigb,n=10,m=1000,sigj=1){
-        output <- list()
-        output$beta     <- ifelse(rbinom(m,1,pi0),0,rnorm(m,0,sigb))
-        output$D        <- replicate(n,rnorm(m,output$beta,sigj))
-        output$p.values <- apply(output$D,1,p.values)
-        return(output)
-    }
+simulateD <- function(pi0,sigb,n=10,m=1000,sigj=1){
+    output <- list()
+    output$beta     <- ifelse(rbinom(m,1,pi0),0,rnorm(m,0,sigb))
+    output$D        <- replicate(n,rnorm(m,output$beta,sigj))
+    output$p.values <- apply(output$D,1,p.values)
+    return(output)
+}
    
+if (problem1){
     ## iii - generate simple p-values
     scenarios <- list(c(1,1),c(.5,3),c(0,3))    # each is a pi0, sigb pair
     simulations <- list()
@@ -39,24 +41,26 @@ if (problem1){
         hist(simulations[[i]]$p.values,breaks=20,freq=F,main=bquote("Histogram, "~pi[0] == .(pi0)~','~ sigma[b]== .(sigb)),xlab='p-value',ylab='Density')
         dev.off()
     }
+}
 
-    ## iv - implement Benjamini-Hochberg
-    BH <- function(pvals,alpha){
-        m    <- length(pvals)
-        ord  <- order(pvals)
-        test <- pvals[ord]< (1:m)*alpha/m
-        test[ord] <- (1:m)<Position(identity,test,right=T,nomatch=0)
-        return(test)
-    }
+## iv - implement Benjamini-Hochberg
+BH <- function(pvals,alpha){
+    m    <- length(pvals)
+    ord  <- order(pvals)
[…]

### 2015-05-13 stat302/hw4.tex
Handing in hw4, but there is some bug with how I'm calculating the likelihood that screws up problem 3

@@
 % placeins defines "\FloatBarrier", which stops tables from moving around
-\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
+\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins, bigints} 
 \newcommand{\Z}{\mathbb{Z}}
@@
 
+\newcommand{\p}{\mathrm{P}}
 \newcommand{\E}{\mathrm{E}}
@@
                 \includegraphics[width=9cm]{hw4/pr2_ii_4}\\
-
+        \end{itemize}
+    \item[3.]
+        \begin{itemize}
+            \item[i)]
+                We can prove this by using the conditional independence of the $D_ij$ and decomposing the likelihood with $\sigma$ fixed:
+                \begin{align*}
+                    \p(D\st\beta) &= \prod_{j=1}^m \prod_{i=1}^n \p(D_{ij}\st\beta_j) \\
+                    \p(D\st\beta) &\propto \prod_{j=1}^m \prod_{i=1}^n \exp\left\{-\frac{1}{2\sigma^2}(\beta_j^2-2\beta_jD_{ij})\right\} \\
+                    \p(D\st\beta) &\propto \prod_{j=1}^m \exp\left\{-\frac{n}{2\sigma^2}\left(\beta_j^2-2\beta_j\sum_{j=1}^n\frac{D_{ij}}{n}\right)\right\} \\
+                    \p(D\st\beta) &\propto \prod_{j=1}^m \exp\left\{-\frac{n}{2\sigma^2}\left(\beta_j^2-2\beta_j\bar D_{j}\right)\right\} \\
+                    \p(D\st\beta) &\propto \prod_{j=1}^m \exp\left\{-\frac{1}{2\frac{\sigma^2}{n}}\left(\beta_j^2-2\beta_j\bar D_{j}\right)\right\} \\
+                    \p(D\st\beta) &\propto \prod_{j=1}^m \p(\bar D_j \st \beta_j) \\
+                    \p(D\st\beta) &\propto \p(\bar D \st \beta) \\
+                \end{align*}
+                Proving that $\bar D$ is a sufficient statistic for the full matrix $D$.
+            \item[ii)]
+                We can derive the log-likelihood by first integrating out $\beta$ to get the pdf of $\bar D$, and then taking the log. Let $\sigma_n^2 = \frac{\sigma^2}{n}$.
+                \begin{align*}
+                    \p(\bar D\st \pi_0,\sigma_b) &= \prod_{j=1}^m \p(\bar D_j\st \pi_0,\sigma_b)\\
+                    \p(\bar D\st \pi_0,\sigma_b) &= \prod_{j=1}^m\int_\R \p(\bar D_j\st \beta_j) d\p_{\beta_j \st \pi_0,\sigma_b}\\
+                    \p(\bar D\st \pi_0,\sigma_b) &= \prod_{j=1}^m\left(\pi_0\p(\bar D_j\st \beta_j=0) + \frac{1-\pi_0}{\sigma_b\sqrt{2\pi}}\bigintsss_\R \p(\bar D_j\st \beta_j)\exp\left\{-\frac{\beta_j^2}{2\sigma_b^2}\right\}d\beta_j\right) \\
+                    \p(\bar D\st \pi_0,\sigma_b) &\propto \prod_{j=1}^m\left(\pi_0 + \frac{1-\pi_0}{\sigma_b\sqrt{2\pi}}\bigintsss_\R \exp\left\{-\frac{1}{2\sigma_n^2}\left(\beta_j^2-2\beta_j\bar D_{j}\right)\right\}\exp\left\{-\frac{\beta_j^2}{2\sigma_b^2}\right\}d\beta_j\right) \\
+                    \p(\bar D\st \pi_0,\sigma_b) &\propto \prod_{j=1}^m\left(\pi_0 + \frac{1-\pi_0}{\sigma_b\sqrt{2\pi}}\bigintsss_\R \exp\left\{-\frac{\sigma_n^2+\sigma_b^2}{2\sigma_n^2\sigma_b^2}\left(\beta_j^2-2\frac{\sigma_b^2}{\sigma_n^2+\sigma_b^2}\beta_j\bar D_{j}\right)\right\}d\beta_j\right)\\
+                    \p(\bar D\st \pi_0,\sigma_b) &\propto \prod_{j=1}^m\left(\pi_0 + \frac{1-\pi_0}{\sigma_b\sqrt{2\pi}}\exp\left\{\frac{\sigma_b^2}{2\sigma_n^2(\sigma_n^2+\sigma_b^2)}\bar D_j^2\right\}          \bigintsss_\R \exp\left\{-\frac{\sigma_n^2+\sigma_b^2}{2\sigma_n^2\sigma_b^2}\left(\beta_j-\frac{\sigma_b^2}{\sigma_n^2+\sigma_b^2}\bar D_{j}\right)^2\right\}d\beta_j\right)\\
+                    \p(\bar D\st \pi_0,\sigma_b) &\propto \prod_{j=1}^m\left(\pi_0 + \frac{1-\pi_0}{\sigma_b\sqrt{2\pi}}\exp\left\{\frac{\sigma_b^2}{2\sigma_n^2(\sigma_n^2+\sigma_b^2)}\bar D_j^2\right\}  \sqrt{\frac{2\pi\sigma_n^2\sigma_b^2}{\sigma_n^2+\sigma_b^2}}\right) \\
+                    \p(\bar D\st \pi_0,\sigma_b) &\propto \prod_{j=1}^m\left(\pi_0 + \frac{(1-\pi_0)\sigma_n}{\sqrt{\sigma_n^2+\sigma_b^2}}\exp\left\{\frac{\sigma_b^2}{2\sigma_n^2(\sigma_n^2+\sigma_b^2)}\bar D_j^2\right\}  \right) \\
+                    l(\pi_0,\sigma_b) &= c +  \sum_{j=1}^m\log\left(\pi_0 + \frac{(1-\pi_0)\sigma_n}{\sqrt{\sigma_n^2+\sigma_b^2}}\exp\left\{\frac{\sigma_b^2}{2\sigma_n^2(\sigma_n^2+\sigma_b^2)}\bar D_j^2\right\}  \right) \\
+                \end{align*}
+            \item[iii)]
+                I've implemented a method that maximizes the likelihood given above.
+In testing, it converged fine, but there must be some mistake in my derivation or my code I can't figure out, since it yields consistantly the wrong answer.
+None the less, this is my code:
+                \lstinputlisting[firstline=186,lastline=218]{hw4.R}
+            \item[iv)]
+                Using the expression for likelihood we derived at the second to last step of ii, we can calculate \\ $\p(\beta_j=0\st \bar D_j, \pi_0, \sigma_b)$ as 
+                \begin{align*}
+                    \p(\beta_j=0\st \bar D_j, \pi_0, \sigma_b) &= \frac{\p(\beta_j=0\st \pi_0, \sigma_b)\p(\bar D_j \st \beta_j=0)}{\p( \bar D_j \st \pi_0, \sigma_b)} \\
+                    \p(\beta_j=0\st \bar D_j, \pi_0, \sigma_b) &= \frac{\pi_0}{ \pi_0 + \frac{(1-\pi_0)\sigma_n}{\sqrt{\sigma_n^2+\sigma_b^2}}\exp\left\{\frac{\sigma_b^2}{2\sigma_n^2(\sigma_n^2+\sigma_b^2)}\bar D_j^2\right\} } \\
+                    \p(\beta_j=0\st \bar D_j, \pi_0, \sigma_b) &:= \pi_1 \\
+                \end{align*}
+                We can then derive $\p(\beta_j\st \bar D_j, \pi_0, \sigma_b,\beta_j\neq0)$ as 
+                \begin{align*}
+                    \p(\beta_j\st \bar D_j, \pi_0, \sigma_b, \beta_j\neq 0) &\propto  \p(\bar D_j\st \beta_j) \p(\beta_j \st \pi_0, \sigma_b, \beta_j\neq 0) \\
+                    \p(\beta_j\st \bar D_j, \pi_0, \sigma_b, \beta_j\neq 0) &\propto  \exp\left\{-\frac{1}{2\sigma_n^2}\left(\beta_j^2-2\beta_j\bar D_j\right) -\frac{1}{2\sigma_b^2}\beta_j^2\right\} \\
+                    \p(\beta_j\st \bar D_j, \pi_0, \sigma_b, \beta_j\neq 0) &\propto  \exp\left\{-\frac{\sigma_n^2+\sigma_b^2}{2\sigma_n^2\sigma_b^2}\left(\beta_j^2-2\frac{\sigma_b^2}{\sigma_n^2+\sigma_b^2}\bar \beta_jD_{j}\right)\right\} \\
+                    \p(\beta_j\st \bar D, \pi_0, \sigma_b, \beta_j\neq 0) &= \frac{1}{\sqrt{\frac{\sigma_n^2\sigma_b^2}{\sigma_n^2+\sigma_b^2}2\pi}} \exp\left\{-\frac{\sigma_n^2+\sigma_b^2}{2\sigma_n^2\sigma_b^2}\left(\beta_j-\frac{\sigma_b^2}{\sigma_n^2+\sigma_b^2}\bar D_{j}\right)^2\right\} \\
+                    \p(\beta_j\st \bar D, \pi_0, \sigma_b, \beta_j\neq 0) &= \sqrt{\frac{1}{\sigma_b^2}+\frac{1}{\sigma_n^2}} \frac{1}{2\pi} \exp\left\{-\left(\frac{1}{\sigma_b^2}+\frac{1}{\sigma_n^2}\right)\left(\beta_j-\frac{\sigma_b^2}{\sigma_n^2+\sigma_b^2}\bar D_{j}\right)^2\right\}
+                \end{align*}
+                Giving us a total estimate of 
+                \[ \p(\beta_j\st\bar D, \pi_0, \sigma_b) = \begin{cases} 
+                        (1-\pi_1)\sqrt{\frac{1}{\sigma_b^2}+\frac{1}{\sigma_n^2}} \frac{1}{2\pi} \exp\left\{-\left(\frac{1}{\sigma_b^2}+\frac{1}{\sigma_n^2}\right)\left(\beta_j-\frac{\sigma_b^2}{\sigma_n^2+\sigma_b^2}\bar D_{j}\right)^2\right\} &\mbox{if } \beta_j\neq 0 \\
+                        \pi_1 &\mbox{if } \beta_j=0 \end{cases} \]
+            \item[v)] Since my likelihood is screwed up, I didn't actually use the optimized parameters and cheated.
+However, I still use the likelihood in part of my calculation of p-values as stated above, so the end result seems to still be that something is off.
+Non withstanding, here are my results: \\
+                \includegraphics[width=9cm]{hw4/pr3_v_FDR_a}
+                \includegraphics[width=9cm]{hw4/pr3_v_pFDR_a} \\
+                \includegraphics[width=9cm]{hw4/pr3_v_FDR_b}
+                \includegraphics[width=9cm]{hw4/pr3_v_pFDR_b} \\
+                \includegraphics[width=9cm]{hw4/pr3_v_FDR_c}
+                \includegraphics[width=9cm]{hw4/pr3_v_pFDR_c} \\
+                I think my mistake made my test way overly conservative in a way I don't quite understand.
         \end{itemize}

### 2015-05-13 stat302/hw4.R

@@ -12,7 +12,8 @@ library("qvalue")
 
 ### Caller
 problem1 <- FALSE 
-problem2 <- TRUE
+problem2 <- FALSE
+problem3 <- TRUE
 
 ### Problem 1
 ## ii - simulate simple hierarchical model
@@ -177,4 +178,100 @@ if (problem2){
 
 }
 
+
+### Problem 3
+if (problem3){
+    ##part iii - r function to compute log-likelihood for hyperparameters
+
+    # compute log-likelihood for single observation in the hierarchical model
+    hloglik.sin <- function(Dbar,pi,sig.b,sig.n=1/sqrt(10)) {
+
+        # fraction part second term in log
+        frac  <- (1-pi)*sig.n/sqrt(sig.b^2+sig.n^2)
+        # exponential part second term in log
+        expon <- exp(sig.b^2*Dbar^2/(2*sig.n^4 + 2*sig.b^2)) 
+
+        # actual log-likelihood
+        loglik <- log(pi + frac*expon)  
+        return(loglik)
+    }
+
+    # Vectorize the log-likelihood for a single term
+    v.hloglik.sin <- Vectorize(hloglik.sin,c('pi','sig.b'))
+
+    # function to return the entire log-likelihood
+    hloglik <- function(Dbar,pi,sig.b) {
+       return(sum(v.hloglik.sin(Dbar,pi=pi,sig.b=sig.b)))
+    }
+
+    # wrapper of above for optim
+    hloglik.optim <- function(param,Dbar){
+        # convert to form in equation
+        pi    <- inv.logit(param[1])
+        sig.b <- exp(param[1])
+        return(hloglik(Dbar,pi,sig.b))
+    }
+
+    # converts output optim back into normal form
+    conv.param <- function(param) {
+        return(c(inv.logit(param[1]),exp(param[2])))
+    }
+
+    Dbar <- apply(simulateD(.2,2)$D,1,mean)
+    opt.param <- conv.param(optim(c(.5,2),hloglik.optim,Dbar=Dbar)$par)
+
+    ## part v - estimate pvalues with posterior
+    post.pval <- function(Dbar,pi,sig.b,sig.n=1/sqrt(10)){
+        denom <- exp(hloglik.sin(Dbar,pi,sig.b,sig.n=1/sqrt(10)))
+        return(pi/denom)
+    }
[…]

### 2015-05-14 stat347/hw3a.tex
working on hw3 a

--- 2015-04-09 stat347/hw1.tex
@@
 
+\newcommand{\hwhead}[1]{#1 \hfill Aaron Maurer \vspace{2mm} \hrule \vspace{2mm}}
+
 \begin{document}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-STAT347 Homework 1 \hfill Aaron Maurer
-\vspace{2mm}
-\hrule
-\vspace{2mm}
-
+\hwhead{STAT 347 Homework 3, Part 1}
 \begin{itemize}
-    \item[1.]
-        \begin{itemize}
-            \item[(a)] One would usually want to measure covariates before assigning treatment so that treatments could be assigned on the basis of covariates.
-By assigning each treatment to a variety of values of covariates (for instance, each unit in a block receiving different treatment), it becomes less likely the effect of the covariate is confounded with the effect of the treatment.
-            \item[(b)] Let $v$ be an additional unit such that, where $S$ is the set of all units, \(\forall v' \in S\backslash \{v\}, R(v,v')\neq1\).
-Since we can not apply the given logic to $v$, it need not be the case \(R(v,v)=1\), so the relation is not necessarily reflexive.
-            \item[(c)] Site, race, sex, and martial status are unordered qualitative factors, since they each take on a finite number of levels which admit not obvious ordering.
-I would classify age as a quantitative covariate, since it can be measured as a continuous (as opposed to discrete) value; in practice however, it can also be treated as an ordered qualitative factor, since it is usually recorded in discrete year values.
-            \item[(d)]
+    \item[0.]
         \begin{itemize}
             \item[(i)]
-                        For the \(y\sim A + {\bf 1}\) model, the model matrix is
-                        \[ \left[ \begin{array}{cccc} 1 & 0 & 0 & 1 \\
-                                                      1 & 0 & 1 & 0 \\
-                                                      1 & 0 & 1 & 0 \\
-                                                      1 & 0 & 0 & 1 \\
-                                                      1 & 1 & 0 & 0 \\
-                                                      1 & 0 & 1 & 0 \\
-                                                      1 & 0 & 0 & 1 \\
-                                                      1 & 1 & 0 & 0 \\
-                                                      1 & 1 & 0 & 0 \\
-                                    \end{array} \right] \]
-                                    The second column, corresponding to level 1 of A, would be omitted under the default in R (treatment constraint).
-For the \(y\sim A - {\bf 1}\) model, the model matrix is
-                        \[ \left[ \begin{array}{cccc} -1 & 0 & 0 & 1 \\
-                                                      -1 & 0 & 1 & 0 \\
-                                                      -1 & 0 & 1 & 0 \\
-                                                      -1 & 0 & 0 & 1 \\
-                                                      -1 & 1 & 0 & 0 \\
-                                                      -1 & 0 & 1 & 0 \\
-                                                      -1 & 0 & 0 & 1 \\
-                                                      -1 & 1 & 0 & 0 \\
-                                                      -1 & 1 & 0 & 0 \\
-                                    \end{array} \right] \]
-                        Once again, the second column would be omitted.
-These two models are equivalent in the sense they would yield the same predictions.
-This is because they define the same subspace since \({\bf -1} \in span({\bf 1})\).
-For \(y \sim A + B\), the model matrix would be
-                        \[ \left[ \begin{array}{cccccc}
-                                                      0 & 0 & 1 & 1 & 0 & 0 \\
-                                                      0 & 1 & 0 & 1 & 0 & 0 \\
-                                                      0 & 1 & 0 & 1 & 0 & 0 \\
-                                                      0 & 0 & 1 & 0 & 1 & 0 \\
-                                                      1 & 0 & 0 & 0 & 1 & 0 \\
-                                                      0 & 1 & 0 & 0 & 1 & 0 \\
-                                                      0 & 0 & 1 & 0 & 0 & 1 \\
-                                                      1 & 0 & 0 & 0 & 0 & 1 \\
-                                                      1 & 0 & 0 & 0 & 0 & 1 \\
-                                    \end{array} \right] \]
-                        Here, the 4th column, corresponding to the $1$ level of B would be omitted, since it is in the span of the rest of the columns.
+                We can calculate the joint distribution easily from the description:
+                \[\p(Y,N) = \begin{cases} 
+                              \;            \frac{1}{3} &\mbox{if } Y=0,N=0 \\
+                              \;   \frac{1}{3}(1-\pi_1) &\mbox{if } Y=0,N=1 \\
+                              \;       \frac{1}{3}\pi_1 &\mbox{if } Y=1,N=1 \\
+                              \;   \frac{1}{3}(1-\pi_2) &\mbox{if } Y=0,N=2 \\
+                              \;       \frac{1}{3}\pi_2 &\mbox{if } Y=2,N=2 \\
+                              \;            0           &\mbox{otherwise } \\
+                            \end{cases} \]
             \item[(ii)]
-                        If we treat A as a quantitative covariate, then for the \(y\sim A + {\bf 1}\) model, the model matrix is
-                        \[ \left[ \begin{array}{cc} 1 & 3 \\
-                                                    1 & 2 \\
-                                                    1 & 2 \\
-                                                    1 & 3 \\
-                                                    1 & 1 \\
-                                                    1 & 2 \\
-                                                    1 & 3 \\
-                                                    1 & 1 \\
-                                                    1 & 1 \\
-                                    \end{array} \right]
-                        \]
-                        With no columns omitted in R. For the \(y\sim A - {\bf 1}\) model, the model matrix is
-                        \[ \left[ \begin{array}{cc}   -1 & 3 \\
-                                                      -1 & 2 \\
-                                                      -1 & 2 \\
-                                                      -1 & 3 \\
-                                                      -1 & 1 \\
-                                                      -1 & 2 \\
-                                                      -1 & 3 \\
-                                                      -1 & 1 \\
-                                                      -1 & 1 \\
-                                    \end{array} \right]
-                        \]
-                        With no columns omitted in R either.
-This model still defines an equivalent linear space to the \(y\sim A + {\bf 1}\) model, so would yield identical predictions.
-Finally, For \(y \sim A + B\), the model matrix would be
-                        \[ \left[ \begin{array}{cccc}
-                                                      3 & 1 & 0 & 0 \\
-                                                      2 & 1 & 0 & 0 \\
-                                                      2 & 1 & 0 & 0 \\
-                                                      3 & 0 & 1 & 0 \\
-                                                      1 & 0 & 1 & 0 \\
-                                                      2 & 0 & 1 & 0 \\
-                                                      3 & 0 & 0 & 1 \\
-                                                      1 & 0 & 0 & 1 \\
-                                                      1 & 0 & 0 & 1 \\
-                                    \end{array} \right] 
-                        \]
-                        Which would have no columns omitted.
+                $Y/N \st N=1$ and $Y/N \st N=2$ are both Bernoulli random variables with probability with probabilities $\pi_1$, $\pi_2$, so 
+                \[\E(Y/N \st N=1) = \pi_1,\; \var(Y/N \st N=1) = \pi_1(1-\pi_1),\;\E(Y/N \st N=2) = \pi_2,\; \var(Y/N \st N=1) = \pi_2(1-\pi_2) \]
+                If $\pi_1=\pi_2$, $Y/N \st N=1 \sim Y/N \st N=2$, so $Y/N$ and $N$ are uncorrelated.
             \item[(iii)]
-                        For each of these subspaces, the dimension is the same as the dimension of the model matrix as created above.
-The squared length of the orthogonal projection is, when the model matrix is $X$, \(\|X(X'X)^{-1}X'y\|^2\) and the squared length of the orthogonal complement is \(\|(I-X(X'X)^{-1}X')y\|^2\).
-This yields:
-                        \FloatBarrier
-                        \input{hw1/pr1_d_iii}
-                        \FloatBarrier
-                    \item[(iv)] I would use the sum of squares from A.
-Specifically, I would compare the marginal difference in sum of squares between the projection onto the A space and onto the ${\bf 1}$ space, and compare this to the sum of squares in the complement space to A.
-It appears B has little if any effect, so it can probably be safely ignored for this purpose.
-If there is no treatment effect, these two values should both be $\chi^2$ variables with the respective degrees of freedom, so we can use them to construct a F test:
-                        \[ \frac{\frac{330.52-258.89}{3-1}}{\frac{13.99}{6}} = 15.361\]
-                        Checking against a $F_{2,6}$ distribution, this corresponds to a p-value of $.0044$.
+                Since the wells are independent, $Y\sim B(26,\pi_1)+2\times B(15,\pi_2)$, so
+                \[ \E[Y] = \E[B(26,\pi_1)] + 2\E[B(15,\pi_2)] = 26\pi_1 + 30\pi_2\]
+                and
+                \[ \var(Y) = \var[B(26,\pi_1)] + 4 \var[B(15,\pi_2)] = 26\pi_1(1-\pi_1) + 60\pi_2(1-\pi_2)\]
+            \item[(iv)]
+                Fitting this distribution $F$ to the data, with $m=26\times1 + 15\times 2 = 56$, we would estimate 
+                \[ \mu = m\pi = \E[Y] = 26\pi_1 + 30\pi_2 \Longrightarrow \pi = \frac{26\pi_1 + 30\pi_2}{56}\] 
+                so we would calculate $\sigma^2$ as
+                \begin{align*}
+                    \sigma^2m\pi(1-\pi) &= \var[y] \\
+                    \sigma^256\frac{26\pi_1 + 30\pi_2}{56}\left(1-\frac{26\pi_1 + 30\pi_2}{56}\right) &= 26\pi_1(1-\pi_1) + 60\pi_2(1-\pi_2)\\
+                    \sigma^2&= \frac{56(26\pi_1(1-\pi_1) + 60\pi_2(1-\pi_2))}{(26\pi_1 + 30\pi_2)(56-26\pi_1 - 30\pi_2)}\\
+                \end{align*}
+                Which, in the case where $\pi_1=\pi_2=\pi$, reduces to 
+                \[ \sigma^2 = \frac{76}{56} = \frac{19}{14}\]
+            \item[(v)]
+                Let $Y_1$ be the number of homogamic matings in single-mating wells out of $M_1$ total,  and $Y_2$ and $M_2$ the same for two mating wells. If we assume that $\pi_1=\pi_2=\pi$, then 
+                \[ Y_2/2 \sim B(M_2/2,\pi) \quad \&\quad Y_1 \sim B(M_1,\pi)\]
+                so 
+                \[\var(Y_2) = \var(2B(M_2/2,\pi)) = 2M_2\pi(1-\pi)\]
+                and 
+                \[\var(Y_1) = \var(B(M_1,\pi)) = M_1\pi(1-\pi)\]
+                so, if $2M_1=2M_2=M$,
+                \begin{align*}
+                    \var(Y) &= 2M_2\pi(1-\pi)+M_1\pi(1-\pi) \\
+                    \sigma^2M\pi(1-\pi) &= \frac{3}{2}\pi(1-\pi) \\
+                    \sigma^2 &= \frac{3}{2} \\
+                \end{align*}
+            \item[(vi)]
+                This mechanism can't explain the chi-square test, which indicated under dispersion.
+If it was the case that one couldn't see one homogamic and one heterogamic mating in one well as we simulated, then the result is higher variance than the binomial model anticipates.
+This is the opposite of what is observed in the actual data set, suggesting this mechanism is not the explanation.
+            \item[(vii)]
+                They would have concluded that the data does not follow a multinomial distribution by looking at the variance of frequencies among replicated experiments, for instance through the same chi square test as above.
+Since these frequencies had much lower variance than one would anticipate with a binomial model, they concluded the frequencies weren't binomial.
+This would also stand to reason, since with only two flies of each gender in a well, one pair mating would necessarily effect the probability of the other pair mating, invalidating independence among the mating events.
+The data not being binomial would invalidate the standard errors and p-values, which were made under an assumed binomial mode.
+            \item[(viii)] By only analyzing the first mating in each well, they were assuring there was no dependence among mating events they counted.
+Since there is independence between wells, this first mating must necessarily be independent of the other counted matings.
+With this fix, the binomial model should hold again, allowing for simpler inference.
         \end{itemize}
 \end{itemize}
-    \item[2.]
-        \begin{itemize}
-            \item[(i)]
-                This table can be gotten directly from the anova command in R
-                \FloatBarrier
-                \input{hw1/pr2_i}
-                \FloatBarrier
-            \item[(ii)]
-                This table is a bit more complicated.
-I generated most of the output from a series of anova commands (Listing the spaces in increasing size so as to get the right quotient spaces), but then recalculated the F and p-values based on the overall residuals from the first model:
-                \FloatBarrier
-                \input{hw1/pr2_ii}
-                \FloatBarrier
-            \item[(iii)] This table strongly indicates that temperature an effect on fungal growth, since temperature as a categorical variable has a very low p-value on it.
-However, lower order polynomials seem to be sufficient to capture the effect; while the linear fit is poor, adding both a 2nd and 3rd degree polynomial seems to cause statistically significant improved fits.
-After this though, greater complexity does not seem to cause statistically significant improvement, indicating a 3rd degree polynomial is optimal.
-            \item[(iv)]
-                I've plotted the cubic fit below:
-                \FloatBarrier
-                \begin{center}
-                    \includegraphics[width=9cm]{hw1/pr2_iv}
-                \end{center}
-                \FloatBarrier
-                Based off this fit, one would predict growth rate will be maximized by a temperature of approximately 71.75.
 
-        \end{itemize}
-    \item[3.]
-        \begin{itemize}
-            \item[(i)]
-                For each conservative matrix $V$, it must be the case $\forall i,j$ that \(V(i_0,i_1)+V(i_1,i_0)=0\Longrightarrow V(i_0,i_1)=-V(i_1,i_0) \).
-This means that $V$ is skew-symmetric.
-On the other hand, for
-                \[ W = \left[ \begin{array}{ccc} 0 & -4 & -1 \\ 4 & 0 & -2 \\ 1 & 2 & 0 \end{array} \right] \]  
-                We have that \(W(3,1)+W(1,2)+W(2,3)= 1 -4 -2 \neq 0\), so not all skew-symmetric matrices are conservative. \\
-                \vspace{5mm} \\
-                If a matrix $V$ is additive, for a path of length $m$ where $i_m=i_0$,
-                \[ \sum_{j=0}^{m-1} V(i_j,i_{j+1}) = \alpha_{i_0} + \sum_{j=1}^{m-1}(-\alpha_{i_j}+\alpha_{i_j}) - \alpha_{i_0} = 0 \]
-                So $V$ is conservative. We may also conclude the converse by the following logic on any conservative matrix $V$:
-                \begin{align*}
-                    0 &= V(u_1,u_2) + V(u_2,u_3) + V(u_3,u_4) +V(u_4,u_1) \\
-                    \Longrightarrow -V(u_4,u_1) - V(u_2,u_3) &= V(u_1,u_2) + V(u_3,u_4) \\
-                    \Longrightarrow \;\;\; V(u_1,u_4) + V(u_3,u_2) &= V(u_1,u_2) + V(u_3,u_4) \\
-                \end{align*}
-                This is only possible if, for all $u_i, u_j$, $V(u_i,u_j)=f(u_i)+g(u_j)$ for some functions $f, g$. Since 
-                \begin{align*}
-                    V(u_1,u_3) &= V(u_1,u_2) + V(u_2,u_3) \\
-                    f(u_1) + g(u_3) &= f(u_1) + g(u_2) + f(u_2) + g(u_3) \\
-                    f(u_2) &= -g(u_2) \\
-                \end{align*}
-                Proving that $V(u_i,u_j) = f(u_i) - f(u_j)$ and thus that $V$ is additive. \\
-                \vspace{5mm} \\
-                Finally, the space of $6\times 6$ conservative matrices had dimension $5$.
-This is because the matrix is minimally defined by \(\{V(1,2),V(1,3),V(1,4),V(1,5),V(1,6)\}\).
-All the remaining entries can be calculated as a linear combination of these five values: \(V(u,j) = V(u,1)+V(1,j) = -V(1,u)+V(1,j)\).
-No fewer values will suffice for this purpose.
-            \item[ii]
-                We can set this up as a typical least squares problem by transforming the $k\times k$ matrix $Y$ into a $k^2\times 1$ vector $y$.
-Then, if $R$ is the usual matrix of 1s and 0s for the row as a categorical variable, and $C$ is the same for the column, our model matrix is $X=R-C$.
-This gives us \(E(y)=X\alpha\).
-After omitting one of the columns of $X$ due to collinearity, we can then produce the normal least squares estimate $\alpha=(X'X)^{-1}X'y$.
-\(\alpha_i-\alpha_j = (\alpha_i+c) - (\alpha_j+c)\), so \(\{\alpha_1,\alpha_2,\alpha_3,\alpha_4,\alpha_5\}\) and \(\{\alpha_1+c,\alpha_2+c,\alpha_3+c,\alpha_4+c,\alpha_5+c\}\) are equivalent parameters, demonstrating that one parameter needs to be omitted.\\
-                \vspace{5mm} \\
-                When we preform this procedure on the data, we get these predictions for the $\alpha$ by electrolyte:
-                \FloatBarrier
-                \input{hw1/pr3_ii}
-                \FloatBarrier
-                As noted, the fifth value is a free parameter and was omitted.
-            \item[iii]
-                These two models are one just including the $\alpha$, and one where the $\alpha$ are interacted with electrolyte. Running these two models and comparing them with ANOVA, we get:
-                \FloatBarrier
-                \input{hw1/pr3_iii}
-                \FloatBarrier
-                By the interaction, we don't see quite enough of an improvement to push past the $.05$ p-value threshold, though it is a borderline case.
-            \item[iv]
-                Our data set only has factor covariates, so no monotone transformation will have an effect on the predictions, and non-monotone transformation wouldn't make sense.
-If there was evidence of a lack of normality or non-constant variance, a transformation might improve our statistical inference, but checking various residual plots leaves little indication that this is necessary.
-        \end{itemize}
-    \item[5.]
-        This can be formulated as a linear model after transformation: 
-        \[ log(Y) ~ \alpha - \beta_T t\] 
-        If we are trying to model the expected half-life $h$ for a given sample under this model, we can formulate the problem as
-        \begin{align*}
-            log(.5) &= \hat\beta_Th \\
-            \frac{1}{h} &= \frac{\hat\beta_T}{log(.5)}\\
-        \end{align*}
-        The value on the right is t-distributed under the assumptions of a normal model, so we can generate a $95\%$ confidence interval by $\hat\beta_T\pm t_{n-p}^{.025}se(\hat\beta)$.
-Then, we can take the inverse of each boundary to get a confidence interval on the $h$.
-We must keep in mind though that any value that is $0$ or negative before taking the inverse corresponds to a lack of decay and asorbic acid levels not actually falling.
-This yields the following confidence intervals for half-life (in weeks):
-        \[ \begin{array}{cccc} Temperature & Lower Bound & Estimate & Upper Bound \\
-                              \hline  
-                              0           & 67.63       & 948.55   & \infty      \\
-                              10          & 20.70       & 28.92    &  47.99       \\
-                              20          & 4.89        & 5.24     & 5.65        \\
-                      \end{array} \]
-        It is worth noting that this does not include any prediction error.
-However, that error's contribution would be relatively small.
-This model accounts for for the variance among replicates; that variance contributes to the variance in the mean which we are fitting.
-
-\end{itemize}
 \end{document}

### 2015-05-14 stat347/hw3a.tex
approaching done

@@
                 This mechanism can't explain the chi-square test, which indicated under dispersion.
-If it was the case that one couldn't see one homogamic and one heterogamic mating in one well as we simulated, then the result is higher variance than the binomial model anticipates.
+If it was the case that one couldn't see one homogamic and one heterogamic mating in one well as we simulated, then the result is higher variance than the multinomial model anticipates.
 This is the opposite of what is observed in the actual data set, suggesting this mechanism is not the explanation.
@@
                 They would have concluded that the data does not follow a multinomial distribution by looking at the variance of frequencies among replicated experiments, for instance through the same chi square test as above.
-Since these frequencies had much lower variance than one would anticipate with a binomial model, they concluded the frequencies weren't binomial.
+Since these frequencies had much lower variance than one would anticipate with a multinomial model, they concluded the frequencies weren't multinomial.
 This would also stand to reason, since with only two flies of each gender in a well, one pair mating would necessarily effect the probability of the other pair mating, invalidating independence among the mating events.
-The data not being binomial would invalidate the standard errors and p-values, which were made under an assumed binomial mode.
+The data not being multinomial would invalidate the standard errors and p-values, which were made under an assumed multinomial mode.
             \item[(viii)] By only analyzing the first mating in each well, they were assuring there was no dependence among mating events they counted.
 Since there is independence between wells, this first mating must necessarily be independent of the other counted matings.
-With this fix, the binomial model should hold again, allowing for simpler inference.
+With this fix, the multinomial model should hold again, allowing for simpler inference.
+            \item[(ix)]
+                It appears counting just the first mating solves the dispersion problem.
+When performing a Chi-Square test for the presence of mating pair-generation effects, we now get a p-value of $.107$.
+If we then combine the counts for all homogamic and heterogamic matings in each generation and repeat the test, we get a p-value of .785.
+This suggests that we have the proper dispersion for a multinomial and that generations don't have an effect on mating preference.
+\vspace{2mm}
+
+                If subsequent matings had been analyzed, aside from the mentioned statistical issues due to dependence, the conclusion would likely have been less well supported.
+First, we can imagine in general, the first mating to occur is in general the one representing the strongest preference.
+Even if this is not the case though, due to female flies refractory period, after one mating both male flies will likely attempt to mate with the remaining female fly.
+One can imagine that the desire to mate is stronger than a particular preference if there is only one option, making a preference harder to discern.
+            \item[(x)]
+                Testing the significance of homogamic versus heterogamic pairings, I ran a logistic regression on the counts:
+                \FloatBarrier
+                \input{hw3/A0_ix_logit}
+                \FloatBarrier
+                There is extremely high significance on the intercept, strongly supporting that the probability of the first mating being homogamic is high.
+This supports the authors' conclusion that homogamic prefernce is preferred, though one could imagine a case where preference changes after some exposure time and an initial mating.
+Also, I included a second degree polynomial on generation as a second test for whether it effected mating preference; as we can see, it does not.
+
+        \end{itemize}
+    \item[3.]
+        \begin{itemize}
+            \item[(a)]
+                These are the parameter estimates from the main-effects linear logistic model:
+                \FloatBarrier
+                \input{hw3/A3_a_logit}
+                \FloatBarrier
+                For the deviance, the residual degrees of freedom are calculated as the total number of Bernoulli trials represented in the data set minus the minus dimension of the linear space of predictors.
+Another way of saying this is the sum of all positive and negative outcomes minus the number of variables included in the model (after removing linearly dependent variables).
+            \item[(b)]
+                With an estimate of $.1239$ on sex2, the model is estimating a small increase in the rate of affected for women, though it is not statistically significant.
+We can estimate a confidence interval for the odds ratio by calculating a confidence interval for the parameter based on asymptotic normality and transforming it.
+To start, a $90\%$ CI for the parameter is:
+                \[(.1239+z_{.05}.2288,1.239+z_{.95}.2288)\quad=\quad(-.2524,.5002)\]
+                Then, we transform this by exponentiating it to yields an odds ratio CI:
+                \[(e^{-.2524},e^{.5002})\quad=\quad (.7769,1.6491)\]
+                Thus, we can say with $90\%$ confidence that the odds ratio for women as compared to men is between $.7769$ and $1.6491$.
+            \item[(c)]
+                I dropped sex first, and then race from the original model to get the model below.
+I didn't drop dust2 because thought it wasn't significant, it was part of a factor which was extremely significant.
+                \FloatBarrier
+                \input{hw3/A3_c_logit}
+                \FloatBarrier
+                Each of the remaining factors is estimated to cause a statistically significant increase in the risk of byssinosis.
+The third level of dust is estimated to cause a far bigger increase in risk than any of the other variables, causing a large increase over the first and second level, which are not statistically significantly different.
+The smoking level 2 and employment levels 2 and 3 both are estimated to increase the risk of the disease over the first level of each.
+
         \end{itemize}

### 2015-05-14 stat347/hw3a.R

@@ -1,0 +1,40 @@
+### Libraries
+library('xtable')
+library('stats4')
+library('locfit')
+library('SDMTools')
+library('plotrix')
+library('reshape2')
+library('nlme')
+library('parallel')
+library('faraway')
+
+### Problem 0
+comm2   <- read.table('hw3/commensal2.txt',header=T)[1:15,]
+comm2.col <- data.frame(Generation=as.numeric(comm2[,1]),homo=comm2[,2]+comm2[,5],hetero=comm2[,3]+comm2[,4])
+
+comm.glm <- glm(cbind(homo,hetero)~poly(Generation,2),family=binomial,data=comm2.col)
+comm.glm.sum <-summary(comm.glm)
+print(xtable(comm.glm.sum),type="latex",file="hw3/A0_ix_logit.tex")
+
+comm.cs <- chisq.test(comm2[1:15,2:5])
+comm.col.cs <- chisq.test(comm2.col[1:15,2:3])
+
+### Problem 3 
+byss <- read.table('hw3/byss.txt',header=T)
+byss.col <- colnames(byss)
+for (c in 3:7) {
+    byss[,byss.col[c]] <- as.factor(byss[,byss.col[c]])
+}
+## (a)
+byss.glm <- glm(cbind(affected,not_affected)~race+sex+smok+empl+dust,family=binomial,data=byss)
+print(xtable(summary(byss.glm)),type="latex",file="hw3/A3_a_logit.tex")
+
+## (b) estimate odds ratio CI for sex2
+sex2.ci <- .1239+c(-1,1)*qnorm(.05)*.2288
+sex2.od.ci <- exp(sex2.ci)
+
+## (c) drop sex then race, in that order
+byss.glm2 <- glm(cbind(affected,not_affected)~smok+empl+dust,family=binomial,data=byss)
+print(xtable(summary(byss.glm2)),type="latex",file="hw3/A3_c_logit.tex")
+

### 2015-05-14 stat347/hw3a.tex
finished

@@
                 This mechanism can't explain the chi-square test, which indicated under dispersion.
-If it was the case that one couldn't see one homogamic and one heterogamic mating in one well as we simulated, then the result is higher variance than the multinomial model anticipates.
+If it was the case, as we simulated, that  homogamic and one heterogamic matings in the same well are impossible, then the result would be higher variance than the multinomial model anticipates.
 This is the opposite of what is observed in the actual data set, suggesting this mechanism is not the explanation.
             \item[(vii)]
-                They would have concluded that the data does not follow a multinomial distribution by looking at the variance of frequencies among replicated experiments, for instance through the same chi square test as above.
+                They concluded that the data does not follow a multinomial distribution by looking at the variance of frequencies among replicated experiments, for instance through the same chi square test as above.
 Since these frequencies had much lower variance than one would anticipate with a multinomial model, they concluded the frequencies weren't multinomial.
@@
                 \FloatBarrier
-                For the deviance, the residual degrees of freedom are calculated as the total number of Bernoulli trials represented in the data set minus the minus dimension of the linear space of predictors.
-Another way of saying this is the sum of all positive and negative outcomes minus the number of variables included in the model (after removing linearly dependent variables).
+                For the deviance, the residual degrees of freedom are calculated as the observed unique combinations of predictors minus dimension of the linear space of predictors.
+Another way of saying this is the number of observations in the model after aggregating by predictors minus number of variable in the model (after removing linearly dependent variables).
             \item[(b)]
@@
 The smoking level 2 and employment levels 2 and 3 both are estimated to increase the risk of the disease over the first level of each.
+            \item[(d)]
+                Of all the ten models, the only model with an interaction which was statistically significantly better by the likelihood ratio test (p-value of .02294) was the one with sex interacted with dust.
+After running that model, race remained insignificant, so was removed.
+This left the following model, in which there were no insignificant main effects which weren't part of the interaction:
+                \FloatBarrier
+                \input{hw3/A3_d_logit}
+                \FloatBarrier
+                The model now predicts that while dust level 3 still causes a big increase in risk over dust level 1, this increase in risk is much smaller for women.
+As well, the model now has the point estimate that dust 2 causes a small decrease in risk for men and a small increase in risk for women, though neither the main effect or interaction are statistically significant.
+Besides that, the predictions of the model by factor are largely similar to before, with every other variable causing a small increase in risk.
+            \item[(e)]
+                Based on the final version of the model, where sex and dust level are interacted, I have plotted the odds ratios for men and women at different dust levels below:
+                \begin{center}
+                    \includegraphics[width=12cm]{hw3/A3_e_plot} 
+                \end{center}
+                As you can see, the cotton dust level can greatly increase a workers risk of getting Byssinosis.
+When exposed to high levels, males are estimated to be 17.8 times more likely to get the disease then men exposed to low levels; women are 6.54 times as likely to get the disease than males at low levels.
+Though there are large error bounds on the risk for women, both men and women have statistically significantly increased risk.
+The big increase in risk seems to be predicated on high level exposure though.
+At medium levels, neither men or women are estimated to have significant increased in risk over men at low levels of dust, and there is not a statistically significant difference between men and women at the low level.
+            \item[(f)]
+                None of our inference for particular effects shows a statistically significant difference in risk of byssinosis by race and sex.
+However, we do see statistically significant improvement in the overall model when we include the interaction of sex and dust level.
+The significance wasn't so high as to preclude this improvement  being random, but this does suggest that gender has some effect on ones risk.
+The reason could well not be biological though; it could, for instance, be due to men and women being in different roles in the factory in a way not accounted for in our model.
+We don't however have any evidence that race is a factor, since neither the inclusion of it as a variable nor any interaction with it was meaningful.
 

### 2015-05-14 stat347/hw3a.R

@@ -8,6 +8,8 @@ library('reshape2')
 library('nlme')
 library('parallel')
 library('faraway')
+library('lmtest')
+library('Hmisc')
 
 ### Problem 0
 comm2   <- read.table('hw3/commensal2.txt',header=T)[1:15,]
@@ -37,4 +39,36 @@ sex2.od.ci <- exp(sex2.ci)
 ## (c) drop sex then race, in that order
 byss.glm2 <- glm(cbind(affected,not_affected)~smok+empl+dust,family=binomial,data=byss)
 print(xtable(summary(byss.glm2)),type="latex",file="hw3/A3_c_logit.tex")
+
+## (d) largest interaction - turns out to be sex*dust
+byss.glm.int <- glm(cbind(affected,not_affected)~race+sex+smok+empl+dust+sex:dust,family=binomial,data=byss)
+byss.lrtest <- lrtest(byss.glm.int,byss.glm)
+byss.glm.int2 <- glm(cbind(affected,not_affected)~sex+smok+empl+dust+sex:dust,family=binomial,data=byss)
+print(xtable(summary(byss.glm.int2)),type="latex",file="hw3/A3_d_logit.tex")
+
+## (e) summary
+beta <- coef(byss.glm.int2)
+sig <- vcov(byss.glm.int2)
+lincombo <- matrix(,nrow=5,ncol=9)
+lincombo[1,] <- c(0,0,0,0,0,1,0,0,0) # man, dust 2
+lincombo[2,] <- c(0,0,0,0,0,0,1,0,0) # man, dust 3 
+lincombo[3,] <- c(0,1,0,0,0,0,0,0,0) # woman, dust 1
+lincombo[4,] <- c(0,1,0,0,0,1,0,1,0) # woman, dust 2
+lincombo[5,] <- c(0,1,0,0,0,0,1,0,1) # woman, dust 3 
+
+# estimates and standard errors for given subgroups
+est <- lincombo %*% beta
+std <- sqrt(diag(lincombo %*% sig %*% t(lincombo) ))
+odds_est <-exp(cbind(est,est-1.96*std,est+1.96*std))
+odds_est <-cbind(rbind(c(1,1,1),odds_est[1:2,]),odds_est[3:5,])
+
+pdf('hw3/A3_e_plot.pdf')
[…]

### 2015-05-15 masters_paper/knockoffGLM.py
Now using glmnet, averaging over many runs for simulations

@@ -15,8 +15,16 @@ import scipy.linalg as splin
 import scipy.stats as spstat
 import matplotlib.pyplot as plt
 import cvxpy as cvx
-from sklearn import linear_model as lm
-from sklearn.svm import l1_min_c 
+from glmnet import LogisticNet
+
+def get_index_z(coef_vector):
+    """This returns the first index for a variable from the glmnet coef matrix which isn't zero"""
+    n     = coef_vector.size
+    index = 0 
+    while index<n: 
+        if coef_vector[index]!=0: return index
+        index+=1
+    return np.nan 
 
 def solve_sdp(G):
     """ Solves the SDP problem:
@@ -45,88 +53,94 @@ def solve_sdp(G):
     # Return as array, not as matrix.
     return np.asarray(s.value).flatten()
 
-def create_SDP(X,randomize):
-    """ Creates the SDP knockoff of X"""
+class knockoff_logit(object):
+    def __init__(self,y,X,knockoff='equicor',randomize=False):
+        self.y = y
+        self.X = X
+        self.n,self.p = self.X.shape
+        self.knockoff_type = knockoff
+        self.randomize=False
+        self.tol = 1E-5
 
-    # Check for rank deficiency (will add later).
-    tol = 1e-5
+    def _create_SDP(self):
+        """ Creates the SDP knockoff of X"""
+ 
+        # Check for rank deficiency (will add later).
+ 
+        # SVD and come up with perpendicular matrix
+        U, d, V = nplin.svd(self.X,full_matrices=True) 
+        d[d<0] = 0
+        U_perp = U[:,self.p:(2*self.p)]
+        if self.randomize:
+            U_perp = np.dot(U_perp,splin.orth(npran.randn(self.p,self.p)))
+ 
+        # Compute the Gram matrix and its (pseudo)inverse.
+        G     = np.dot(V.T * d**2 ,V)
+        G_inv = np.dot(V.T * d**-2,V)
+ 
+        # Optimize the parameter s of Equation 1.3 using SDP.
+        s = solve_sdp(G)
+        s[s <= self.tol] = 0
+ 
+        # Construct the knockoff according to Equation 1.4:
+        C_U,C_d,C_V = nplin.svd(2*np.diag(s) - (s * G_inv.T).T * s)
[…]

### 2015-05-15 masters_paper/simulate.py

@@ -1,0 +1,165 @@
+# Imports
+from __future__ import division
+import numpy as np
+import numpy.linalg as nplin
+import numpy.random as npran
+import scipy as sp
+import scipy.linalg as splin
+import scipy.stats as spstat
+import matplotlib.pyplot as plt
+import requests
+from multiprocessing import Pool
+from bs4 import BeautifulSoup
+
+import knockoffGLM as ko
+
+# logit/invlogit
+def logit(p):
+    return np.log(p) - np.log(1 - p)
+
+def inv_logit(p):
+    return np.exp(p) / (1 + np.exp(p))
+
+def getRandomIntegers(num=1000,min=1,max=1000000000,base=10):
+    """
+    This will query random.org to get a bunch of truly random integers. I want this so I can get seeds for parallel pseudo-random number generation, since you don't want to use a pseudo-random numbers or a sequence as seeds.
+    
+    Since I'm not using random.org's API (do I look like I'm made out of money?!?), a maximum of 10,000 random numbers can be generated, and they min and max must be within +/- 1,000,000,000
+    """
+    address = 'https://www.random.org/integers/?num=%d&min=%d&max=%d&col=1&base=%d&format=html&rnd=new' % (num,min,max,base)
+    r       = requests.get(address)
+    uniList = BeautifulSoup(r.text).find('pre',class_='data').text.split()
+    numList = map(int,uniList)
+    return numList
+
+def cutoff(array,min=-50,max=50):
+    shape = array.shape
+    cut_array = np.min((max*np.ones(shape),array),axis=0)
+    cut_array = np.max((min*np.ones(shape),cut_array),axis=0)
+    return cut_array
+
+def genXy_normal_X_beta(seed,n,p1,pnull,base_prob=.25,beta_sd=1):
+    """ The X are normal. p1 predictive vars, pnull null vars. beta on the p1 vars is ~normal(0,beta_sd) and the intercept is logit(base_prob)"""
+    if not seed == None:
+        npran.seed(seed)
+    X_1    = npran.randn(n,p1)
+    X_null = npran.randn(n,pnull)
+    X      = np.concatenate((X_1,X_null),axis=1)
+    beta   = npran.randn(p1)*beta_sd
+    if p1>0:
+        eta    = cutoff(np.dot(X_1,beta)+logit(base_prob))
+        y      = npran.binomial(1,inv_logit(eta),n)
+    else:
+        y      = npran.binomial(1,base_prob,n)
+    return X,y
+
[…]

### 2015-05-21 stat347/hw3b.tex
Just need to edit in the morning

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\newcommand{\hwhead}[1]{#1 \hfill Aaron Maurer \vspace{2mm} \hrule \vspace{2mm}}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hwhead{STAT 347 Homework 3b}
\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[(a)] \& (b) \\
                I've plotted the given rates and the fits below. The fits are from a model with the formulae \(chem +dose\), where the log of dose was not taken.
                \begin{center}
                    \includegraphics[width=12cm]{hw3/A1_ab} 
                \end{center}
            \item[(c)]
                We want to compare to logistic models, one with the formulae \(chem + ldose\), where the relationship for each chemical is identical except for different intercepts, and \(chem*ldose\), where the relationship for each chemical is linear in log dose with different intercept and different slope. Since the first model is fit on a subset of the second, we can use a likelihood ratio test to evaluate whether the larger model is significantly better. In this case, with a p-value of .18, we can not conclude so, and default to the smaller model being superior.
            \item[(d)]
                These two models define the same subspace, but in slightly different ways, with the second omitting a constant. The coefficients on $ldose$ will be identical, while the coefficients on $chem$ in the second model will be those in the first plus the intercept from before. If $\Sigma_1$ and $\Sigma_2$ are the covariance matrices, then
                \[\left[\begin{array}{cccc} 1 & 0 & 0 & 0 \\
                                            1 & 1 & 0 & 0 \\
                                            1 & 0 & 1 & 0 \\
                                            0 & 0 & 0 & 1 \\
                           \end{array}\right] \Sigma_1
                   \left[\begin{array}{cccc} 1 & 0 & 0 & 0 \\
                                             1 & 1 & 0 & 0 \\
                                             1 & 0 & 1 & 0 \\
                                             0 & 0 & 0 & 1 \\
                           \end{array}\right] = \Sigma_2 \]

            \item[(e)]
                I for the potency of the combo relative to DDT, I got an estimate of $2.45$ with a $90\%$ CI of \((2.12,2.89)\). Relative to BHC, I got an estimate of $3.08$ with a CI of \((2.64,3.67)\).
            \item[(f)]
                Fitting the models with alternate links, we only seem to get a mildly better fit with the log-log link, raising the log-likelihood from -43.0 to -42.0. All the others have lower (worse) log-likelihoods. When we estimate the potency of the combo relative to the individual chemicals, for DDT I got an estimate of 2.37 with a CI of \((2.07,2.74)\) and for BHC I got an estimate of 3.05 with a CI of \((2.63,3.63)\).
            \item[(g)]
                We will have a 99\% kill rate for the combo with dose $x$ if 
                \begin{align*}
                    .99 &= \frac{1}{1-e^{-\alpha_{combo} - \beta x}} \\
                    \log\left(\frac{.99}{.01}\right) &= \alpha_{combo} + \beta x \\
                    4.59 &= \alpha_{combo} + \beta x \\
                    x &= \frac{4.59-\alpha_{combo}}{\beta} \\
                \end{align*}
                We can estimate a CI on this by fitting a model with an offset of $-4.59$, and then using Fieller's method on the coefficients for the combo and the log dose. Doing so gives us an estimate of $10.59$ with a CI of \((7.17,17.35)\).
            \item[(h)] From our analysis, a few things are clear. For one, the combination dose is more effective than either of the individual insecticides, having much higher potency than each of them. Additionally, from the high significance on dosage, we can see more insecticide kills more bugs, at least up to a point, but the effect of increasing the dosage doesn't change much or at all between insecticides. Also, DDT seems to be somewhat stronger than BHC when each are used on their own.
        \end{itemize}
    \item[2.]
        I fit a log-linear Poisson model predicting the total number of children ($mean \times sample size$) based on the three given variables with an offset of $\log(sample size)$. Log-linear is appropriate since we're fitting counts, and the weighting should account for the varying sample size. A Poisson model seems appropriate here since the variation will increase with the expected number of children (mostly due to more mothers in the sample). A quasi-Poission model might also be appropriate, but when one fits such a model, the dispersion parameter isn't that much smaller at $.8$. Since the data is, if anything, under dispersed, worst case is the model is conservative. \vspace{2mm} \\
        The variables included were Education, a factor indicating no education, lower elementary, upper elementary, or secondary and up, years since marriage in five year buckets as a factor, and whether the mother lived in a rural area or not. I used the model formula $Years + Education + Rural$. Both dropping education, the least significant of the factors, and interacting years and rural, the most significant variables, were inferior choices according to the likelihood ratio test. Fertility seemed to be heavily effected by years since first marriage and rural setting. The highest education levels was also highly significant, but the differences between the other three levels was marginal. \vspace{2mm}\\
        The model gives a prediction that an urban woman with upper elementary education will have 3.60 children on average after 10 years of marriage, with a $95\%$ CI of \((3.34,3.89)\) \vspace{2mm} \\
        To get the lifetime average prediction for a rural women with secondary education, I predicted for the oldest group, 25+ years after marriage. This ignores that some women wouldn't live this long, and thus the lifetime average would be somewhat lower on the whole. None the less, the prediction is 6.02 children with a $90\%$ CI of \((5.42,6.69)\).

\end{itemize}

\end{document}

### 2015-05-21 stat347/hw3b.R

@@ -1,0 +1,116 @@
+### Libraries
+library('xtable')
+library('stats4')
+library('locfit')
+library('SDMTools')
+library('plotrix')
+library('reshape2')
+library('nlme')
+library('parallel')
+library('faraway')
+library('lmtest')
+library('Hmisc')
+
+
+### Problem 1
+# Data
+dose        <- c(   2, 2.64,3.48,4.59,6.06,   8)
+ddt_num     <- c(   3,    5,  19,  19,  24,  35)
+ddt_denom   <- c(  50,   49,  47,  38,  49,  50)
+bhc_num     <- c(   2,   14,  20,  27,  41,  40)
+bhc_denom   <- c(  50,   49,  50,  50,  50,  50)
+both_num    <- c(  28,   37,  46,  48,  48,  50)
+both_denom  <- c(  50,   50,  50,  50,  50,  50)
+
+insect      <- data.frame(
+                            dose        = rep(dose,3),
+                            insecticide = rep(c('DDT','BHC','DDT+BHC'),each=6),
+                            died        = c(ddt_num,bhc_num,both_num),
+                            pop         = c(ddt_denom,bhc_denom,both_denom)
+                         )
+insect$lived <- insect$pop - insect$died
+
+## (a) & (b)
+ins.logit.b <- glm(cbind(died,lived)~insecticide+dose,family=binomial,data=insect)
+n           <- 500
+xseq        <- seq(0,9,length.out=n)
+
+pdf('hw3/A1_ab.pdf')
+matplot(dose,cbind(ddt_num/ddt_denom,bhc_num/bhc_denom,both_num/both_denom),lty=c(1,1,1),col=2:4,pch=c(1,1,1),main="Insect Death Rate With Logistic Fit",ylab="Death Rate",xlab=expression(paste('Dose Insecticide (mg/10c',m^2,')')))
+points(xseq,predict(ins.logit.b,data.frame(dose=xseq,insecticide=rep('DDT',n)    ),type='response'),type='l',col=2)
[…]

### 2015-05-22 stat347/hw3b.tex
More simulations and a demo for Reina

@@
             \item[(c)]
-                We want to compare to logistic models, one with the formulae \(chem + ldose\), where the relationship for each chemical is identical except for different intercepts, and \(chem*ldose\), where the relationship for each chemical is linear in log dose with different intercept and different slope.
+                We want to compare to logistic models, one with the formulae \(chem + ldose\), where the linear predictor for each chemical is identical except for different intercepts, and \(chem*ldose\), where the linear predictor for each chemical is linear in log dose with different intercept and different slope.
 Since the first model is fit on a subset of the second, we can use a likelihood ratio test to evaluate whether the larger model is significantly better.
@@
                            \end{array}\right] = \Sigma_2 \]
-
+                The matrix is, of course, the linear transformation to turn the first space into the second.       
             \item[(e)]
@@
 A quasi-Poission model might also be appropriate, but when one fits such a model, the dispersion parameter isn't that much smaller at $.8$.
-Since the data is, if anything, under dispersed, worst case is the model is conservative.
+Since the data is, if anything, under dispersed, worst case is the model has conservative confidence intervals.
 \vspace{2mm} \\
@@
 I used the model formula $Years + Education + Rural$.
-Both dropping education, the least significant of the factors, and interacting years and rural, the most significant variables, were inferior choices according to the likelihood ratio test.
+Both dropping education, the least significant of the factors, and interacting years and rural, the most significant variables, resulted in inferior models by the likelihood ratio test.
 Fertility seemed to be heavily effected by years since first marriage and rural setting.
-The highest education levels was also highly significant, but the differences between the other three levels was marginal.
+The highest education levels was also highly significant, but the differences between the other three levels was marginal or non-existent.
 \vspace{2mm}\\
         The model gives a prediction that an urban woman with upper elementary education will have 3.60 children on average after 10 years of marriage, with a $95\%$ CI of \((3.34,3.89)\) \vspace{2mm} \\
-        To get the lifetime average prediction for a rural women with secondary education, I predicted for the oldest group, 25+ years after marriage.
-This ignores that some women wouldn't live this long, and thus the lifetime average would be somewhat lower on the whole.
-None the less, the prediction is 6.02 children with a $90\%$ CI of \((5.42,6.69)\).
+        To get the lifetime average prediction for a rural women with secondary education, I predicted for the oldest group, for whom 25+ years had passed since there first marriage.
+This ignores that some women wouldn't live this long, driving up the average, and that some women in this group might still have children, driving down the average.
+None the less, with no other obvious way to answer this question, the prediction is 6.02 children with a $90\%$ CI of \((5.42,6.69)\).
 

### 2015-05-22 masters_paper/knockoffGLM.py

@@ -16,15 +16,16 @@ import scipy.stats as spstat
 import matplotlib.pyplot as plt
 import cvxpy as cvx
 from glmnet import LogisticNet
+from sklearn.preprocessing import normalize
 
 def get_index_z(coef_vector):
     """This returns the first index for a variable from the glmnet coef matrix which isn't zero"""
     n     = coef_vector.size
     index = 0 
     while index<n: 
-        if coef_vector[index]!=0: return index
+        if abs(coef_vector[index])>1e-6: return index
         index+=1
-    return np.nan 
+    return n+1
 
 def solve_sdp(G):
     """ Solves the SDP problem:
@@ -56,7 +57,7 @@ def solve_sdp(G):
 class knockoff_logit(object):
     def __init__(self,y,X,knockoff='equicor',randomize=False):
         self.y = y
-        self.X = X
+        self.X = normalize(X.astype(float),norm='l2',axis=0)
         self.n,self.p = self.X.shape
         self.knockoff_type = knockoff
         self.randomize=False
@@ -114,7 +115,7 @@ class knockoff_logit(object):
 
         # initialize the glmnet object
         self.lognet = LogisticNet(alpha=1) 
-        self.lognet.fit(X_lrg,self.y)
+        self.lognet.fit(X_lrg,self.y,normalize=False,include_intercept=False)
 
         self.lambdas = self.lognet.out_lambdas
         self.var_index_ent = np.sort(self.lognet._indices)
@@ -127,11 +128,11 @@ class knockoff_logit(object):
     def _get_z(self): 
         """ Given the coefficient matrix from a glmnet object, returns the Z, the first non-zero entries"""
         self.z_rank  = np.array(map(get_index_z,self.coef_matrix))
-        self.z_value = np.array(map(lambda x: self.lambdas[x] if not np.isnan(x) else 0,self.z_rank))
+        self.z_value = np.array(map(lambda x: self.lambdas[x] if x<self.lambdas.size else 0,self.z_rank))
 
     def _get_w(self): 
         """Produces the w values using the z values"""
-        self.w_filter = np.sign(self.z_value[0:self.p]-self.z_value[self.p:(2*self.p)])
+        self.w_filter = np.sign(self.z_rank[self.p:(2*self.p)]-self.z_rank[0:self.p])
[…]

### 2015-05-22 masters_paper/simulate.py

@@ -148,9 +148,10 @@ class ko_test(object):
             plt.plot(np.arange(self.p)+1,self.true_orig_rank_rate,'b-', label = 'True Predictors')
         if self.pnull>0:
             plt.plot(np.arange(self.p)+1,self.null_orig_rank_rate,'g-', label = 'Null Predictors')
+            end_val = self.null_orig_rank_rate[-1]
         plt.xlabel('Rank Variable Entered Model')
         plt.ylabel('Cumulative % Original Variables')
-        plt.title('Percentage Original Variables')
+        plt.title('Percentage Original Variables\n %.3f of Null Originals Came in First By End' % end_val)
         plt.axis([1,self.p,0,1])
         plt.legend(loc='lower right')
         plt.show()

### 2015-06-03 masters_paper/knockoffGLM.py
Done with homework 7

@@ -15,6 +15,9 @@ import scipy.linalg as splin
 import scipy.stats as spstat
 import matplotlib.pyplot as plt
 import cvxpy as cvx
+import rpy2 as rp 
+import rpy2.robjects.numpy2ri
+from rpy2.robjects.packages import importr
 from glmnet import LogisticNet
 from sklearn.preprocessing import normalize
 
@@ -56,8 +59,9 @@ def solve_sdp(G):
 
 class knockoff_logit(object):
     def __init__(self,y,X,knockoff='equicor',randomize=False):
-        self.y = y
-        self.X = normalize(X.astype(float),norm='l2',axis=0)
+        self.y      = y
+        self.X      = normalize(X.astype(float),norm='l2',axis=0)
+        self.X_orig = X
         self.n,self.p = self.X.shape
         self.knockoff_type = knockoff
         self.randomize=False
@@ -109,6 +113,89 @@ class knockoff_logit(object):
         s_diff = 2*s - (s/d)**2
         s_diff[s_diff<0]=0 # can be negative due to numerical error
         self.X_ko = np.dot(U*(d-s/d) + U_perp*(np.sqrt(s_diff)) , V)
+
+    def _binary_knockoff(self):
+        ''' This creates knockoffs which are random multivariate bernoulli which should have, in expectation,
+        the same first two moments as X. Only will work if X is all binaray '''
+        # First, calculate s
+        # Largely replicates begining of self._create_equicor()/self._create_SDP()
+
+        # SVD and come up with perpendicular matrix
+        U, d, V = nplin.svd(self.X,full_matrices=True) 
+        d[d<0] = 0
+        if   self.knockoff_type == 'SDP': 
+            # Compute the Gram matrix and its (pseudo)inverse.
+            G     = np.dot(V.T * d**2 ,V)
+            G_inv = np.dot(V.T * d**-2,V)
+         
+            # Optimize the parameter s of Equation 1.3 using SDP.
+            s = solve_sdp(G)
+            s[s <= self.tol] = 0
+
+        # Same as first part self._create_equicor 
+        elif self.knockoff_type == 'equicor':
+            # Set s = min(2 * smallest eigenvalue of X'X, 1), so that all the correlations
+            # have the same value 1-s.
+            lambda_min = min(d)**2
+            s = np.ones(self.p)*min(2*lambda_min,1)
+
+        # Generate the large covariance matrix we want
[…]

### 2015-06-03 stat302/hw7.R

@@ -1,0 +1,19 @@
+
+### Problem 1
+util <- function(r) .62*log(.004*r+1)
+exp.util.1a <- util(500)/3
+exp.util.1b <- 2*util(-100)/3 + util(400)/3
+
+### Problem 3
+brier <-  function(p,act) (p-act)^2 + (-p+act)^2
+lloss <-  function(p,act) -act*log(p) - (1-act)*log(1-p)
+pred1 <-c(.59,.79,.54,.36,.77,.62,.37,.18,.39,.14)
+pred2 <-c(.32,.57,.60,.53,.85,.21,.46,.07,.45,.29)
+actual <- c(1,1,0,0,1,1,1,0,1,0)
+
+brier1.3 <- sum(brier(pred1,actual))/10
+brier2.3 <- sum(brier(pred2,actual))/10
+lloss1.3 <- sum(lloss(pred1,actual))/10 
+lloss2.3 <- sum(lloss(pred2,actual))/10 
+
+

### 2015-06-03 stat347/hw4.tex
Started working on homework 4

--- 2015-05-21 stat347/hw3b.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hwhead{STAT 347 Homework 3b}
+\hwhead{STAT 347 Homework 4}
 \begin{itemize}
@@
         \begin{itemize}
-            \item[(a)] \& (b) \\
-                I've plotted the given rates and the fits below. The fits are from a model with the formulae \(chem +dose\), where the log of dose was not taken.
-                \begin{center}
-                    \includegraphics[width=12cm]{hw3/A1_ab} 
-                \end{center}
-            \item[(c)]
-                We want to compare to logistic models, one with the formulae \(chem + ldose\), where the relationship for each chemical is identical except for different intercepts, and \(chem*ldose\), where the relationship for each chemical is linear in log dose with different intercept and different slope.
-Since the first model is fit on a subset of the second, we can use a likelihood ratio test to evaluate whether the larger model is significantly better.
-In this case, with a p-value of .18, we can not conclude so, and default to the smaller model being superior.
-            \item[(d)]
-                These two models define the same subspace, but in slightly different ways, with the second omitting a constant.
-The coefficients on $ldose$ will be identical, while the coefficients on $chem$ in the second model will be those in the first plus the intercept from before.
-If $\Sigma_1$ and $\Sigma_2$ are the covariance matrices, then
-                \[\left[\begin{array}{cccc} 1 & 0 & 0 & 0 \\
-                                            1 & 1 & 0 & 0 \\
-                                            1 & 0 & 1 & 0 \\
-                                            0 & 0 & 0 & 1 \\
-                           \end{array}\right] \Sigma_1
-                   \left[\begin{array}{cccc} 1 & 0 & 0 & 0 \\
-                                             1 & 1 & 0 & 0 \\
-                                             1 & 0 & 1 & 0 \\
-                                             0 & 0 & 0 & 1 \\
-                           \end{array}\right] = \Sigma_2 \]
-
-            \item[(e)]
-                I for the potency of the combo relative to DDT, I got an estimate of $2.45$ with a $90\%$ CI of \((2.12,2.89)\).
-Relative to BHC, I got an estimate of $3.08$ with a CI of \((2.64,3.67)\).
-            \item[(f)]
-                Fitting the models with alternate links, we only seem to get a mildly better fit with the log-log link, raising the log-likelihood from -43.0 to -42.0.
-All the others have lower (worse) log-likelihoods.
-When we estimate the potency of the combo relative to the individual chemicals, for DDT I got an estimate of 2.37 with a CI of \((2.07,2.74)\) and for BHC I got an estimate of 3.05 with a CI of \((2.63,3.63)\).
-            \item[(g)]
-                We will have a 99\% kill rate for the combo with dose $x$ if 
-                \begin{align*}
-                    .99 &= \frac{1}{1-e^{-\alpha_{combo} - \beta x}} \\
-                    \log\left(\frac{.99}{.01}\right) &= \alpha_{combo} + \beta x \\
-                    4.59 &= \alpha_{combo} + \beta x \\
-                    x &= \frac{4.59-\alpha_{combo}}{\beta} \\
-                \end{align*}
-                We can estimate a CI on this by fitting a model with an offset of $-4.59$, and then using Fieller's method on the coefficients for the combo and the log dose.
-Doing so gives us an estimate of $10.59$ with a CI of \((7.17,17.35)\).
-            \item[(h)] From our analysis, a few things are clear.
-For one, the combination dose is more effective than either of the individual insecticides, having much higher potency than each of them.
-Additionally, from the high significance on dosage, we can see more insecticide kills more bugs, at least up to a point, but the effect of increasing the dosage doesn't change much or at all between insecticides.
-Also, DDT seems to be somewhat stronger than BHC when each are used on their own.
+            \item[(i)]
+                After calculating the mean vector and the two covariance matricies, I found that
+                \[ 
+                    \hat\mu = \left[\begin{array}{c} 5.843 \\ 3.057 \\ 3.758 \\ 1.199 \\ \end{array}\right], \quad
+                    S_b = \left[\begin{array}{cccc}  
+                              31.606 & -9.976 & 82.624 & 35.640 \\ 
+                              -9.976 & 5.672 & -28.620 & -11.466 \\ 
+                              82.624 & -28.620 & 218.551 & 93.387 \\ 
+                              35.640 & -11.466 & 93.387 & 40.207 \\ 
+                        \end{array}\right]
+                    \]
+                and
+                \[
+                    S_w = \left[\begin{array}{cccc}
+                            0.265 & 0.093 & 0.168 & 0.038 \\ 
+                            0.093 & 0.115 & 0.055 & 0.033 \\ 
+                            0.168 & 0.055 & 0.185 & 0.043 \\ 
+                            0.038 & 0.033 & 0.043 & 0.042 \\
+                        \end{array}\right]
+                \]
+            \item[(ii)]
+                The eigenvalues of this matrix are necessarily real because it is the product of two PSD matricies (the PSD cone is closed under taking the inverse).
+H is not an orthogonal matrix though because the product is not symmetric.
         \end{itemize}
-    \item[2.]
-        I fit a log-linear Poisson model predicting the total number of children ($mean \times sample size$) based on the three given variables with an offset of $\log(sample size)$.
-Log-linear is appropriate since we're fitting counts, and the weighting should account for the varying sample size.
-A Poisson model seems appropriate here since the variation will increase with the expected number of children (mostly due to more mothers in the sample).
-A quasi-Poission model might also be appropriate, but when one fits such a model, the dispersion parameter isn't that much smaller at $.8$.
-Since the data is, if anything, under dispersed, worst case is the model is conservative.
-\vspace{2mm} \\
-        The variables included were Education, a factor indicating no education, lower elementary, upper elementary, or secondary and up, years since marriage in five year buckets as a factor, and whether the mother lived in a rural area or not.
-I used the model formula $Years + Education + Rural$.
-Both dropping education, the least significant of the factors, and interacting years and rural, the most significant variables, were inferior choices according to the likelihood ratio test.
-Fertility seemed to be heavily effected by years since first marriage and rural setting.
-The highest education levels was also highly significant, but the differences between the other three levels was marginal.
-\vspace{2mm}\\
-        The model gives a prediction that an urban woman with upper elementary education will have 3.60 children on average after 10 years of marriage, with a $95\%$ CI of \((3.34,3.89)\) \vspace{2mm} \\
-        To get the lifetime average prediction for a rural women with secondary education, I predicted for the oldest group, 25+ years after marriage.
-This ignores that some women wouldn't live this long, and thus the lifetime average would be somewhat lower on the whole.
-None the less, the prediction is 6.02 children with a $90\%$ CI of \((5.42,6.69)\).
-
 \end{itemize}

### 2015-06-03 stat347/hw4.R

@@ -1,0 +1,27 @@
+### Libraries
+library('xtable')
+library('stats4')
+library('locfit')
+library('SDMTools')
+library('plotrix')
+library('reshape2')
+library('nlme')
+library('parallel')
+library('faraway')
+
+### Problem 1
+source('hw4/iris.R')
+data.1 <- data.frame(data,variety)
+
+## (i)
+mu   <- apply(data,2,mean)
+mu.b <- apply(data,2,function(x) tapply(x,variety,mean))
+data.b <- matrix(rep(mu.b - matrix(rep(mu,each=3),ncol=4,byrow=F),each=50),ncol=4)
+S.b  <- t(data.b) %*% data.b / 2
+data.w <- data - matrix(rep(mu.b,each=50),ncol=4,byrow=F)
+S.w  <- t(data.w) %*% data.w / 147 
+
+## (ii)
+eig <- eigen(solve(S.w) %*% S.b)
+
+H <- eig$vectors

### 2015-06-04 stat347/hw4.tex
Done homework 4

@@
 \newcommand{\E}{\mathrm{E}}
+\newcommand{\p}{\mathrm{P}}
 \newcommand{\var}{\mathrm{Var}}
@@
 \newcommand{\st}{ \; \big | \:}
+\newcommand{\car}[1]{\left\vert #1 \right\vert}
 
@@
             \item[(ii)]
-                The eigenvalues of this matrix are necessarily real because it is the product of two PSD matricies (the PSD cone is closed under taking the inverse).
-H is not an orthogonal matrix though because the product is not symmetric.
+                The eigenvalues of $S_w^{-1}S_b$ are 
+                \[\lambda = 2366.107,\; 20.976, \; 0.000, \; 0.000\]
+                This tells us that variation among blocks largely lies on a 1 or 2 dimmensional manifold.
+In particular, it can overwheminly be captured as occuring on a line.
+That $S_w^{-1}S_b$ follows from $S_w$ being positive definite.
+Due to this, we can say $S_w=BB$ For some positive definite matrix $B$.
+Then,
+                \begin{align*}
+                    S_w^{-1}S_b     &= BBS_b \\
+                    B^{-1}S_w^{-1}S_bB &= BS_bB
+                \end{align*}
+                $BS_bB$ must be symmetric and thus have real eigenvalues.
+Since I have established $S_w^{-1}S_b$ is similar to a $BS_bB$, it too must have real eigenvalues.
+Since the eigenvalues are real, so are the eigenvectors, H.
+However, the $S_w^{-1}S_b$ need not be symmetric, so nothing is forcing $H$ to be orthogonal.
+Let $h_1$ be the first eigenvector.
+The portion of the variance due to the clustering in $Yh_1$ is smaller than for any other linear combination.
+            \item[(iii)]
+                To show that $\E(S_w)=\Sigma$, we observe
+                \begin{align*}
+                    \Sigma &= \E\left[\left(Y_i-\E[Y_i\st b_i]\right)'\left(Y_i-\E[Y_i\st b_i]\right)\right] \\
+                           &= \E\left[\frac{1}{n-\car{B}}\sum_{i=1}^n\left(Y_i-\frac{1}{n_b}\sum_{j\in b} Y_j\right)'\left(Y_i-\frac{1}{n_b}\sum_{j\in b} Y_j\right)     \right] \\
+                           &= \E\left[Y'Y + Y'\bar B Y\right] \\
+                           &= \E\left[Y'(I+\bar B)Y\right] \\
+                           &= \E\left[S_w\right] 
+                \end{align*}
+                Now, to show the result for $\E(S_b)$
+                \begin{align*}
+                    (\car{B}-1)S_b &= \sum(Y_iY_i') - n\bar Y\bar Y' - Y'(I-J)Y  \\
+                    (\car{B}-1)S_b &= \sum(Y_iY_i'-\bar Y\bar Y') - Y'(I-J)Y  \\
+                    (\car{B}-1)S_b &= \sum(Y_iY_i'-\frac{1}{n}\sum_b n_b\bar Y_b\bar Y_b') - Y'(I-J)Y  \\
+                    (\car{B}-1)S_b &= \sum_b\left(\frac{1}{n}\sum_{i\in b}Y_iY_i'- \bar Y_b\bar Y_b'\right) - Y'(I-J)Y  \\
+                    \E(S_b) &= \Sigma +\theta\Sigma\frac{n^2-\sum n^2_b}{n(\car{B}-1)}
+                \end{align*}
+                Using the result above, we can estimate the variance ratio based on 
+                \[ \E\left[S_b\right] = \E\left[S_w\right]\left(1+\theta\frac{n^2-\sum n_b^2}{n(\car{B}-1}\right) \]
+                Allowing us to derive an estimator
+                \begin{align*}
+                    tr(S_b) &= tr(S_w)\left(1+\hat\theta\frac{n^2-\sum n_b^2}{n(\car{B}-1)}\right) \\
+                    \frac{tr(S_b)}{tr(S_w)} &=1+\hat\theta\frac{n^2-\sum n_b^2}{n(\car{B}-1)} \\
+                    \frac{tr(S_b)}{tr(S_w)}-1 &=\hat\theta\frac{n^2-\sum n_b^2}{n(\car{B}-1)} \\
+                    \hat\theta &= \frac{n(\car{B}-1)}{n^2-\sum n_b^2}\left(\frac{tr(S_b)}{tr(S_w)}-1\right) 
+                \end{align*}
+                Using this estimator, I get an estimate of $\hat\theta = 9.727$ on the original scale, and $\hat\theta = 18.210$ when the log is taken of the data.
+            \item[(iv)]
+                Calculating the predictive probabilities with $\lambda=1$ and $\theta=10$, I got this result:
+                \FloatBarrier
+                \input{hw4/A1_iv}
+                \FloatBarrier
+            \item[(v)]
+                Now, with $\lambda=1$ and $\theta=5$, I got
+                \FloatBarrier
+                \input{hw4/A1_v_5}
+                \FloatBarrier
+                And with  $\lambda=1$ and $\theta=50$, I got
+                \FloatBarrier
+                \input{hw4/A1_v_50}
+                \FloatBarrier
+            The predictions do seem to be somewhat sensitive to $\theta$.
+In general, $\theta$ effects the relative probabilities of the existing clusters only slightly, but has a large effect on the probability of a new cluster.
+            \item[(vi)]
+                As $\theta\to0$
+                \[ n_b\, \phi_4\!\left(y(u') - \frac{\mu+n_b\theta \bar y_b}{1+n_b\theta};\Sigma\left(1+\frac{\theta}{1+n_b\theta}\right)\right) \to n_b\, \phi_4\!\left(y(u') - \mu;\Sigma\right) \]
+                and 
+                \[ \lambda\, \phi_4\!\left(y(u') - \mu;\Sigma\left(1+\theta\right)\right) \to \lambda\, \phi_4\!\left(y(u') - \mu;\Sigma\right) \]
+                This means that, since $\phi_4\!\left(y(u') - \mu;\Sigma\right)$ is a common factor,
+                \[ \p(u' \mapsto b\st ...) \to \begin{cases} \frac{n_b}{\lambda+n} & b\in B \\
+                                                         \frac{\lambda}{\lambda+n} & b = \emptyset 
+                                                 \end{cases} \]
+                Which is the Ewens process and independent of the feature vector. On the other hand, if $\theta\to\infty$, we see that
+                \[ n_b\, \phi_4\!\left(y(u') - \frac{\mu+n_b\theta \bar y_b}{1+n_b\theta};\Sigma\left(1+\frac{\theta}{1+n_b\theta}\right)\right) \to n_b\, \phi_4\!\left(y(u') - \bar y_b;\Sigma\left(1+\frac{1}{n_b}\right)\right) \]
+                and 
+                \[ \lambda\, \phi_4\!\left(y(u') - \mu;\Sigma\left(1+\theta\right)\right) \to 0 \]
+                so the predictive probability for the new class tends to zero and 
+                \[ \p(u' \mapsto b \st b\in B ...) \propto n_b\phi_4\!\left(y(u') - \bar y_b;\Sigma\right) \]
+                which is the Bayes optimal solution based on the Fisher discriminant and class frequencies based on sample frequencies. 
+            \item[(vii)]
+                Making the scatter plot of the projection:
+                \begin{center}
+                    \includegraphics[width=15cm]{hw4/A1_vii} 
+                \end{center}
+                Looking at this plot, its unsurprising that none of the values we predicted had high probabilities for both Setosa and another existing cluster; it is far more distinct from the other two than they are from each other.
+If its probability is of similar magnitude to one of the other clusters, then that means that the point is in the big gap in the middle, and the most likely prediction is probably a new cluster.
+Now, predicting $\bar y$, I got
+                \FloatBarrier
+                \input{hw4/A1_vii}
+                \FloatBarrier
+                The overall mean is much closer to the Versicolor mean than the other means, so these probabilities are very skewed towards Versicolor.
+            \item[(viii)]
+                Here is the alternate plot of the data
+                \begin{center}
+                    \includegraphics[width=15cm]{hw4/A1_viii} 
+                \end{center}
+                This is a plot of the data's projection into the first two principal components of the data matrix (though without subtracting out the mean or normalizing).
+We are still capturing the majority of the variation in the data; we're just using a different basis.
+            \item[(ix)]
+                One might want to use a log scale since all the measurements by their nature must be positive, which defies the normality assumption.
+When one makes predictions based on the log scale, these are the new probabilities.
+                \FloatBarrier
+                \input{hw4/A1_ix}
+                \FloatBarrier
+                Besides the second point, the predictions are quite similar as before.
+        \end{itemize}
+    \item[5.]
+        \begin{itemize}
+            \item[(i)] {\bf Show this is a probability distribution:}
+                This follows from this manipulation
+                \begin{align*}
+                    \p(T=t;m,n) &= \frac{\prod_i n_i! \prod_j m_j!}{n_*! \prod_{ij} t_{ij}!} \\
+                                &= \frac{n_*!\prod_i n_i! \prod_j m_j!}{n_*!n_*! \prod_{ij} t_{ij}!} \\
+                                &= \frac{\prod_i n_i!}{n_*!}\frac{\prod_j m_j}{n_*!}\frac{n_*!}{ \prod_{ij} t_{ij}!} \\
+                                &= \frac{{{n_*}\choose {t_{11},...,t_{cr}}}}{{{n_*}\choose {n_1,...,n_r}}{{n_*}\choose {m_1,...,m_c}}}
+                \end{align*}
+                The numerator is the number of ways to fill fill each cell with the desired $t_{ij}$ from $n_*$ observations, and the denominator is the number of all possible tables which satisfy the marginal distributions.
+Thus, it is clear that summing the probability over all possible $T$ which satisfy the marginals will yield 1, making this a proper probability distribution.
+            \item[(ii)] {\bf Show $T=\tilde A'P\tilde B$ has a hypergeometric distribution:} Since $PB$ is just a permutation of the rows of $B$, the column sums will remain the same.
+This means that the marginals for the new factor mapping to $PB$ is the same as for the factor for $B$.
+In turn, the resulting table $A'PB$ will have the same marginals in both the rows and columns, so clearly there is the proper conditioning.
+Further, each permutation is chosen with probability $\frac{1}{n_*!}$.
+However, most of these permutations result in the same table.
+In particular, if one permutes the values of $B$ matched with each particular value of $A$, then the table is the same.
+There are $\prod_i n_i!$ such permutations.
+Similarly, permutations of a particular value of $B$, of which there are $\prod_j m_j$ will also result in the same table.
+Then, there are $\prod_{ij}t_{ij}!$ permutations that do both of these things at once.
+It follows that there are
+                \[ \frac{\prod_i n_i! \prod_j m_j!}{\prod_{ij} t_{ij}!} \]
+                permutations total that result in the same table. Thus, we conclude that the probability under this model of drawing a particular table is
+                \[\frac{\prod_i n_i! \prod_j m_j!}{n_*! \prod_{ij} t_{ij}!}\]
+                Making it a hypergeometric distribution.
+            \item[(iii)] {\bf Compute Pearson's chi-squared test and deviance statistic:} 
+                I found that the nominal Pearson Chi-square statistic is 109.883, and the deviance statistic is
+                \[ 2\sum_{i,j}t_{ij}\left(\log(t_{ij})-\log\left(n_*\frac{m_j}{n_*}\frac{n_i}{n_*}\right)\right) = 130.599 \]
+            \item[(iv)] {\bf Estimate the p-value by simulation:} 
+                Over $10000$ simulations, I found that the Perason's chi-squared test statistic was exceeded 7723 times, and the deviance statistic was exceeded 6689 times.
+These correspond to p-values of .772 and .669 respectively.
+Below I've plotted both sets of bootstraps versus the density of the expected chi-square with 121 degrees of freedom: \\
+                \includegraphics[width=9cm]{hw4/A5_chi} 
+                \includegraphics[width=9cm]{hw4/A5_dev} 
+                The Pearson statistic is extremely close the Chi-sq, as one would hope for. However, the deviance statistic is noticeably shifted to the right. This 
+            \item[(v)] {\bf Aggregate the data by birth death month difference:} 
+                After aggregating the expected and actual birth and death combinations by the difference in month, this is the resulting table:
+                \FloatBarrier
+                \input{hw4/A5_sum_diff}
+                \FloatBarrier
+                We seem to see fewer than expected deaths in the months leading up to someone's birthday, and more deaths than expected afterwards.
+Using the Pearson Chi-squared test to evaluate this, I get a p-value of $.027$, which is evidence, though not overwhelming, supporting that this is statistically significant.
+
+            \item[(vi)] {\bf Explain how to extend the hypergeometric simulation to three dimensions where each one dimensional table fixed:} 
+                For each cell, the expected count in each cell under the null should still be the product of the marginal probabilities times the total $n$.
+Thus, for for the indicator matrices $\tilde A$, $\tilde B$, $\tilde C$, we can simulate a random null matrix fitting the marginals using random permutation matrices $P_1$ and $P_2$.
+A cell of the table is
+                \[ T_{i,j,k} = \sum_{m=1}^n \tilde A_{mi} (P_1\tilde B)_{mj} (P_2\tilde C)_{mk} \]
+                with expectation, under the null, of
+                \[ \E[T_{i,j,k}\st H_o] = n\frac{\|\tilde A_i\|_1}{n}\frac{\|\tilde B_j\|_1}{n}\frac{\|\tilde C_k\|_1}{n}\]
+                From here, the simulation is the same, bootstrapping a test statistic over many simulated tables and comparing it to the observed value.
+            \item[(vii)] {\bf Explain how to extend the hypergeometric simulation to three dimensions with two-dimension tables fixed:} 
+                Here, the null is that the cells are conditionally independent. In other words,
+                \begin{align*}
+                    \p(A_m=i,B_m=j,C_m=k) &= \p(A_m=i,B_m=j\st C_m=k)\p( C_m=k) \\
+                                          &= \p(A_m=i\st C_m=k)\p(B_m=j\st C_m=k)\p( C_m=k) \\
+                \end{align*}
+                This can be tested by the same method as with two way tables, just induced on all the subtables $T_{*,*,k}$ at the same time.
+For each subset of $A_k$ and $B_k$, which are the values of $A$ and $B$ for which $C=k$, draw a random permutation $P$, and then $\tilde A_k' P \tilde B_k$ should be hypergeometric.
+A test statistic is then generated on all the two way tables for all $k$ by comparing the seen cells $T_{i,j,k}$ to
+                \[ \E(T_{i,j,k}) = \frac{1}{u_k}\sum_{i'} T_{i',j,k}\sum_{j'}T_{i,j',k} \]
+                Where $u_k$ is the number of times level $k$ of $C$ occurs.
         \end{itemize}

### 2015-06-04 stat347/hw4.R

@@ -8,20 +8,150 @@ library('reshape2')
 library('nlme')
 library('parallel')
 library('faraway')
+library('mvtnorm')
 
 ### Problem 1
 source('hw4/iris.R')
 data.1 <- data.frame(data,variety)
 
 ## (i)
-mu   <- apply(data,2,mean)
-mu.b <- apply(data,2,function(x) tapply(x,variety,mean))
-data.b <- matrix(rep(mu.b - matrix(rep(mu,each=3),ncol=4,byrow=F),each=50),ncol=4)
-S.b  <- t(data.b) %*% data.b / 2
-data.w <- data - matrix(rep(mu.b,each=50),ncol=4,byrow=F)
-S.w  <- t(data.w) %*% data.w / 147 
+calc <- function(data,variety) {
+    mu   <- apply(data,2,mean)
+    mu.b <- apply(data,2,function(x) tapply(x,variety,mean))
+    data.b <- matrix(rep(mu.b - matrix(rep(mu,each=3),ncol=4,byrow=F),each=50),ncol=4)
+    S.b  <- t(data.b) %*% data.b / 2
+    data.w <- data - matrix(rep(mu.b,each=50),ncol=4,byrow=F)
+    S.w  <- t(data.w) %*% data.w / 147 
+    return(list(mu,mu.b,S.b,S.w))
+}
+calc.reg <- calc(data     ,variety)
+calc.log <- calc(log(data),variety)
 
 ## (ii)
-eig <- eigen(solve(S.w) %*% S.b)
+eig <- eigen(solve(calc.reg[[4]]) %*% calc.reg[[3]])
+H   <- eig$vectors
+lam <- eig$values
 
-H <- eig$vectors
+## (iii)
+trace <- function(X) return(sum(diag(X)))
+
+est.theta <- function(calc.list) {
+    return((trace(calc.list[[3]])/trace(calc.list[[4]])-1)*150*2/(150^2-3*50^2))
+}
+theta.reg <- est.theta(calc.reg)
+theta.log <- est.theta(calc.log) 
+
+## (iv)
+pred.data <- rbind(c(6.3,2.9,4.9,1.7),
+                   c(5.5,3.1,2.9,0.8),
+                   c(5.8,3.2,3.5,1.1),
+                   c(5.6,3.2,3.2,1.0)
+                   )
[…]

### 2015-06-08 stat302/paper_report.tex
Started on paper report; finished overview

--- 2015-06-03 stat347/hw4.tex
@@
 \documentclass[11pt]{article}
-\usepackage[paper=letterpaper, margin=.5in]{geometry}
+\usepackage[paper=letterpaper, margin=.75in]{geometry}
+\linespread{1.5}
 \pdfpagewidth 8.5in
 \pdfpageheight 11in
-\setlength\parindent{0in}
 
@@
 \begin{document}
+\title{Literature Report on DP-means}
+\author{Aaron Maurer}
+\date{STAT 302, Spring Quarter 2015}
+\maketitle
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hwhead{STAT 347 Homework 4}
-\begin{itemize}
-    \item[1.]
-        \begin{itemize}
-            \item[(i)]
-                After calculating the mean vector and the two covariance matricies, I found that
-                \[ 
-                    \hat\mu = \left[\begin{array}{c} 5.843 \\ 3.057 \\ 3.758 \\ 1.199 \\ \end{array}\right], \quad
-                    S_b = \left[\begin{array}{cccc}  
-                              31.606 & -9.976 & 82.624 & 35.640 \\ 
-                              -9.976 & 5.672 & -28.620 & -11.466 \\ 
-                              82.624 & -28.620 & 218.551 & 93.387 \\ 
-                              35.640 & -11.466 & 93.387 & 40.207 \\ 
-                        \end{array}\right]
-                    \]
-                and
-                \[
-                    S_w = \left[\begin{array}{cccc}
-                            0.265 & 0.093 & 0.168 & 0.038 \\ 
-                            0.093 & 0.115 & 0.055 & 0.033 \\ 
-                            0.168 & 0.055 & 0.185 & 0.043 \\ 
-                            0.038 & 0.033 & 0.043 & 0.042 \\
-                        \end{array}\right]
-                \]
-            \item[(ii)]
-                The eigenvalues of this matrix are necessarily real because it is the product of two PSD matricies (the PSD cone is closed under taking the inverse).
-H is not an orthogonal matrix though because the product is not symmetric.
-        \end{itemize}
-\end{itemize}
+\section{Overview}
+For this report, I chose to read "Revisiting k-means: New Algorithms via Bayesian Nonparametrics"\footnote{Brian Kulis and Michael I.
+Jordan, ``Revisiting k-means: New Algorithms via Bayesian Nonparametrics'', \textit{CoRR} (2011): http://arxiv.org/abs/1111.0352}.
+This article introduces a alternative to the k-means clustering algorithm called DP-means, where instead of specifying a number of clusters at the start, then working to optimize their fit, new clusters are introduced when a point in the data set is too far from one of the existing clusters.
+This in itself is a `hard' clustering algorithm, which outputs an assignment to clusters, rather than a posterior distribution one might normally expect from a Bayesian method.
+However, this paper demonstrates that this algorithm is the limit of a Bayesian model where the total distribution is a mixture of Gaussian distributions, with the number of components coming from a Dirichlet process.
+This is the source of the name of the algorithm: Dirichlet process means or DP-means.
+\par
+In addition to a description and derivation of this algorithm, the paper offers a few extensions and simulations.
+The most interesting of theses is a hierarchical clustering method where, over multiple data sets of the same variables, local clusters are simultaneously fit on each data set so as to match a set of global clusters across all of the data sets.
+This is again a hard clustering algorithm, but it can be derived as well in a Bayesian fashion by taking the limit when the parameters of the Dirichlet process which generates the components in each data set arises from a prior global Dirichlet process.
+The next extension shows how DP-means extends to spectral clustering; where one would spectrally cluster using k-means by performing k-means on the first k eigenvectors of a similarity matrix for the data, instead with DP-means one takes all eigenvectors where the eigenvalues are above a given threshold, then cluster via DP-means using that same threshold to determine when to add additional components.
+Similarly, the author also shows that DP-means can be extended to graph cut problems.
+The paper concludes with a few simulations demonstrating the effectiveness of DP-means and its multiple data set extension.
+\par
+This article offers obvious extensions of how the classical Bayesian statistics we learned can be extended to a machine learning algorithm.
+The prior Dirichlet process is used to guide the posterior number of components.
+The method for fitting the model is the limit of a Gibbs sampling algorithm.
+Finally, the multiple data set version of DP-means is a hierarchical model, based on the exchangeability of the data sets.
+All together, this forms an interesting competitor to k-means, being similarly easy to compute, but built on classic Bayesian principles.
+
+\section{Paper Contents}
+\subsection{DP-Means Algorithm}
 

### 2015-06-08 stat302/paper_report.tex
In second section

@@
 
+\DeclareMathOperator*{\argmin}{arg\,min}
+
 \newcommand{\hwhead}[1]{#1 \hfill Aaron Maurer \vspace{2mm} \hrule \vspace{2mm}}
@@
 \subsection{DP-Means Algorithm}
+The idea of clustering arises, in statistical terms, from the idea that a set of random variables are drawn from a mixture distribution.
+In other words, a random variable $X_i$ is drawn in a two stage process.
+First, a multinomial variable $z_i$ with $k$ possible outcomes is drawn.
+Then, $X_i\st z_i\sim F_{z_i}$, where \(\left\{F_j\st j\in\{0,...,k\}\right\}\) is some set of probability distributions.
+The goal of a clustering method is then to impute the $z_i$ based on the $X_i$, revealing important underlying structure in the data.
+Bayesian models provide a natural way to fit probability distributions to $z_i$, in particular without choosing $k$ a priori, but there issue is that they, classically, are complicated to compute on large data sets and don't necessarily scale well.
+\par
+Thus, the k-means algorithm, where $k$ must be specified beforehand, remains the most commonly used algorithm.
+This method can be thought of as designed for a model where $z_i$ is multinomial $k$, for fixed $k$, and then $X\st z_i \sim N_p(\mu_j,\sigma I_p)$ for some set of mean vectors \(\left\{\mu_j\st j\in\{0,...,k\}\right\}\).
+To find $\{\mu_i\}$ using k-means, one picks initial guesses for the $\hat\mu_j$, then alternates assigning \(\hat z_i = \argmin_j \|X_i - \hat\mu_j\|\) and
+\[\hat\mu_j = \frac{1}{\big | \{i\st z_i=j\} \big |}\sum_{i\st z_i=j }X_i \] 
+until the $\mu_j$ and $z_i$ have converged.
+This method allows little flexibility, but is relatively easy to compute, even for large data sets.
+The authors of the paper notes that the fitting process is the limit of what the EM algorithm would do if $\sigma$ goes to 0.
+\par
+
+In designing the DP-means algorithm, the authors ``attempt to achieve the best of both worlds by designing scalable hard clustering algorithms from a Bayesian nonparametric viewpoint." To this end, they start with a purely Bayesian model for how the data would arise which puts a prior on $k$, and then similarly derives a simple algorithm for hard clustering that represents the limit of fitting this Bayesian algorithm as a variance parameter is sent to 0.
 

### 2015-06-08 masters_paper/notes.tex
Working on notes

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\linespread{1.5}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\p}{\mathrm{P}}
\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\newcommand{\hwhead}[1]{#1 \hfill Aaron Maurer \vspace{2mm} \hrule \vspace{2mm}}

\begin{document}
\title{Binary Knockoffs Notes}
\author{Aaron Maurer}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminaries} 
Some early investigation into how the deterministic knockoffs as described in Rina's paper work with regularized logistic regression revealed that the answer is "not very well". Even when $X$ is a null predictor of $y$, the $X$ still tend to enter the model prior to $\tilde X$. The issue is that even when $X_i\sim N_p({\bf 0}, \Sigma)$ for some $\Sigma\succeq 0$, $\tilde X$ is not normally distributed. This can be seen from producing qq plots of $X_i$ vs $\tilde X_j$Of course, when $X_i$ is a binary vector, $\tilde X_i$ completely doesn't match its distribution, causing the original $X$ to beat the knockoffs into the model. This indicates that a new method of generating $\tilde X$ must be created to use with FDR via knockoffs for regularized logistic regression.

\section{Probabilistic Random Bernoulli Knockoffs}
My idea is to generate $\tilde X$ randomly such that, approximately, $\tilde X_i\sim X_i$. In particular, they should have similar marginal densities, expectations, and first moments. However, $\tilde X\st X$ should also have desired knockoff property that $\E(\tilde X'X\st X) = X'X-s$, where $\diag(X'X)-s$ is small. In the general case, this is likely infeasible; however, if $X$ is a binary vector, as is often the case, we know we are dealing with a much more limited class of random variables, and it should be possible to randomly generate $\tilde X\st X$ so as to have the desired properties. At worst, this method will provide a suitable replacement for deterministic $\tilde X$ as described in Rina's paper for LASSO, and if we are lucky, it will work reasonably for other regularized GLMs.

\section{Random Bernoulli Generation}
Thankfully, there has been a reasonable amount of work on how one can generate random Bernoulli vectors with some kind of correlation among among the values. A random Bernoulli vector $X$ can be easiest represented with a mean vector $\E(X)=m\in(0,1)^p$ and $\E(XX')=M\in(0,1)^{p\times p}$, called the cross moment matrix. Obviously, $m_i = \p(X_i=1)$, $M_{ij} = \p(X_i=X_j=1)$, and $m=\diag(M)$. For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-\diag(M)\diag(M)'$ must be PSD, and
\[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
for all $i\neq j$\footnote{``On parametric families for sampling binary data with specified mean and correlation" - http://arxiv.org/abs/1111.0576}. Given a qualifying $M$, or observed $X$, there are a few ways of generating more random $X$.

\subsection{Gaussian Copula Family}
Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $Z~N_p({\bf 0},\Sigma)$ such that,for $X_j = I(Z_j<0)$, $X$ has the desired properties. There are a number of ways to do this\footnote{``On the Generation of Correlated Artificial Binary Data" - http://epub.wu.ac.at/286/1/document.pdf}\footnote{``On parametric families for sampling binary data with specified mean and correlation"}, but it turns out that there is only certain to exist a working $\Sigma$ in the bivariate case.

\subsection{$\mu$-Conditionals family}
It turns out that there exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as 
\[ X_{j}\st X_{j},...,X_{j-1} \sim \mathrm{B}\left(1,\mu\left(a_{jj}+\sum_{k=1}^{j-1}a_{kj}X_j\right)\right) \]
for some monotone function $\mu:\R\to[0,1]$. This is essentially a binomial family GLM for a link function $\mu$. If one takes all of the $a_{kj}$, they can form a lower triangular matrix $A$, and then the joint density can be expressed as 
\[ \p(X_j=\gamma) \propto \mu(\gamma'A\gamma)\]
If $\mu$ is chose such that it is a bijection and differentiable, there is a unique $M$ such that $\E(X_iX_i')=M$\footnote{``On parametric families for sampling binary data with specified mean and correlation"}. As one might guess, the natural $\mu$ is the logistic link function, which is the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
\[A_{ij} = \log\left(\frac{\p(X_j=X_k=1)\p(X_j=X_k=0)}{\p(X_j=0,X_k=1)\p(X_j=1,X_k=0)}\right) \]
when $i\neq j$. I think this dictates that if $A_{jk}=0$, then $X_k$ and $X_j$ are independent.

\section{Generating Knockoffs}

\end{document}

### 2015-06-10 stat302/paper_report.tex
Done first draft of paper report

@@
 \newcommand{\cov}{\mathrm{Cov}}
+\newcommand{\trace}{\mathrm{trace}}
+\newcommand{\diag}{\mathrm{diag}}
 \newcommand{\ST}{mbox{ s.t. }}
@@
 \newcommand{\st}{ \; \big | \:}
+\newcommand{\car}[1]{ \left\vert #1 \right\vert}
 
@@
 \section{Overview}
-For this report, I chose to read "Revisiting k-means: New Algorithms via Bayesian Nonparametrics"\footnote{Brian Kulis and Michael I.
+For this report, I chose to read ``Revisiting k-means: New Algorithms via Bayesian Nonparametrics"\footnote{Brian Kulis and Michael I.
 Jordan, ``Revisiting k-means: New Algorithms via Bayesian Nonparametrics'', \textit{CoRR} (2011): http://arxiv.org/abs/1111.0352}.
-This article introduces a alternative to the k-means clustering algorithm called DP-means, where instead of specifying a number of clusters at the start, then working to optimize their fit, new clusters are introduced when a point in the data set is too far from one of the existing clusters.
-This in itself is a `hard' clustering algorithm, which outputs an assignment to clusters, rather than a posterior distribution one might normally expect from a Bayesian method.
-However, this paper demonstrates that this algorithm is the limit of a Bayesian model where the total distribution is a mixture of Gaussian distributions, with the number of components coming from a Dirichlet process.
+This article introduces an alternative to the k-means clustering algorithm called DP-means, where instead of specifying a number of clusters at the start, then working to optimize their fit, new clusters are introduced when a point in the data set is too far from one of the existing clusters.
+This in itself is a `hard' clustering algorithm, which outputs an assignment to clusters rather than a posterior distribution across clusters typical of a Bayesian model.
+However, this paper demonstrates that this algorithm is the limit of a Bayesian model.
+In the model, the total distribution is a mixture of Gaussian distributions, with the number of components coming from a Dirichlet process.
 This is the source of the name of the algorithm: Dirichlet process means or DP-means.
@@
 In addition to a description and derivation of this algorithm, the paper offers a few extensions and simulations.
-The most interesting of theses is a hierarchical clustering method where, over multiple data sets of the same variables, local clusters are simultaneously fit on each data set so as to match a set of global clusters across all of the data sets.
-This is again a hard clustering algorithm, but it can be derived as well in a Bayesian fashion by taking the limit when the parameters of the Dirichlet process which generates the components in each data set arises from a prior global Dirichlet process.
-The next extension shows how DP-means extends to spectral clustering; where one would spectrally cluster using k-means by performing k-means on the first k eigenvectors of a similarity matrix for the data, instead with DP-means one takes all eigenvectors where the eigenvalues are above a given threshold, then cluster via DP-means using that same threshold to determine when to add additional components.
+The most interesting of theses is an extension of DP-means called the ``hard Gaussian hierarchical Dirichlet process", over multiple data sets, local clusters are simultaneously fit on each data set to match a set of global clusters.
+This is again a hard clustering algorithm, but it can be derived as well in a Bayesian fashion by taking the limit of a Bayesian model.
+The next extension shows how DP-means extends to spectral clustering in a similar fashion to how k-means does.
 Similarly, the author also shows that DP-means can be extended to graph cut problems.
@@
 This article offers obvious extensions of how the classical Bayesian statistics we learned can be extended to a machine learning algorithm.
-The prior Dirichlet process is used to guide the posterior number of components.
+The DP-means algorithm is a hierarchical model which uses a Dirichlet process as the prior on the number of clusters.
 The method for fitting the model is the limit of a Gibbs sampling algorithm.
-Finally, the multiple data set version of DP-means is a hierarchical model, based on the exchangeability of the data sets.
+The hard Gaussian HDP extension is a typical Bayesian extension of the original model, where an additional level prior distributions is added to an existing model.
 All together, this forms an interesting competitor to k-means, being similarly easy to compute, but built on classic Bayesian principles.
@@
 Thus, the k-means algorithm, where $k$ must be specified beforehand, remains the most commonly used algorithm.
-This method can be thought of as designed for a model where $z_i$ is multinomial $k$, for fixed $k$, and then $X\st z_i \sim N_p(\mu_j,\sigma I_p)$ for some set of mean vectors \(\left\{\mu_j\st j\in\{0,...,k\}\right\}\).
-To find $\{\mu_i\}$ using k-means, one picks initial guesses for the $\hat\mu_j$, then alternates assigning \(\hat z_i = \argmin_j \|X_i - \hat\mu_j\|\) and
-\[\hat\mu_j = \frac{1}{\big | \{i\st z_i=j\} \big |}\sum_{i\st z_i=j }X_i \] 
-until the $\mu_j$ and $z_i$ have converged.
+This method can be thought of as designed for a model where $z_i$ is multinomial $k$, for fixed $k$, and then $X_i\st z_i \sim N_p(\mu_c,\sigma I_p)$ for some set of mean vectors \(\left\{\mu_c\st c\in\{0,...,k\}\right\}\).
+To find $\{\mu_i\}$ using k-means, one picks initial guesses for the $\hat\mu_c$, then alternates assigning \(\hat z_i = \argmin_c \|X_i - \hat\mu_c\|\) and
+\[\hat\mu_c = \bar X_c = \frac{1}{\car{\{i\st z_i=c\}}}\sum_{i\st z_i=c }X_i \] 
+until the $\mu_c$ and $z_i$ have converged.
 This method allows little flexibility, but is relatively easy to compute, even for large data sets.
-The authors of the paper notes that the fitting process is the limit of what the EM algorithm would do if $\sigma$ goes to 0.
+The authors of the paper notes that the fitting process is the limit of what the EM algorithm would do for a mixture of $k$ Gaussians if the variance of each, $\sigma$, goes to 0.
 \par
@@
 In designing the DP-means algorithm, the authors ``attempt to achieve the best of both worlds by designing scalable hard clustering algorithms from a Bayesian nonparametric viewpoint." To this end, they start with a purely Bayesian model for how the data would arise which puts a prior on $k$, and then similarly derives a simple algorithm for hard clustering that represents the limit of fitting this Bayesian algorithm as a variance parameter is sent to 0.
+The Bayesian model for how the data arises sequentially is something like this:
+\begin{itemize}
+    \item The $z_i$ arise from a Dirichlet process. This means that given a parameter $\alpha$, for the $i$th draw, if we have $k$ clusters already
+        \begin{itemize}
+            \item $z_i=c$ for $c\leq k$ with probability \(\frac{\car{\{ j\vert j<i, z_j=c\}}}{\alpha + i -1}\)
+            \item $z_i=k+1$ with probability \(\frac{\alpha}{\alpha + i -1}\)
+        \end{itemize}
+    \item The cluster centers $\mu_1,...,\mu_k\sim N(0,\rho I)$ for some variance parameter $\rho$.
+    \item Finally, \(X_i\st z_i \sim N(\mu_{z_i},\sigma I)\) for another variance parameter $\sigma$.
+\end{itemize}
+The authors fix $\rho$ and $\sigma$, and then set $\alpha=\left(1+\frac{\rho}{\sigma}\right)^{d/2}\exp\left(-\frac{\lambda}{2\sigma}\right)$, where $d$ is the dimension of the $X_i$.
+To set up a Gibbs sampling algorithm, we can calculate the conditional distributions of the unknown parameters $z_1,...,z_n$ and $\mu_1,...,\mu_k$.
+Let us say that there are $k$ clusters, excluding $z_i,X_i$, at any given point.
+Also, let the number of points assigned to cluster $c$ excluding $z_i,X_i$ is $n_{-i,c}$.
+This yields, for the conditional distribution of the $z_i$,
+\begin{align*} 
+    \p(z_i=c \st X_{1:n},z_{-i},\mu_{1:k},\rho,\sigma) &= \p(z_i=c \st X_i,z_{-i},\mu_{1:k},\rho,\sigma) \\
+                                           &\propto \begin{cases} 
+                                                \p(X_i\st z_i=c,\sigma,\rho)\p(z_i=c\st z_{-i})& \mbox{if } c=k+1 \\ 
+                                                \p(X_i\st z_i=c,\mu_c,\sigma)\p(z_i=c\st z_{-i})& \mbox{if } c\leq k \\
+                                            \end{cases} \\
+                                           &\propto \begin{cases} 
+                                                (2\pi\sigma)^{-d/2} \exp\left(-\frac{\lambda}{2\sigma}-\frac{\|X_i\|^2}{2(\rho+\sigma)}\right) & \mbox{if } c=k+1 \\
+                                                n_{-i,c}(2\pi\sigma)^{-d/2} \exp\left(-\frac{\|X_i-\mu_c\|^2}{2\sigma}\right) & \mbox{if } c\leq k \\
+                                            \end{cases}\\
+                                           &\propto \begin{cases} 
+                                                \exp\left(-\frac{\lambda}{2\sigma}-\frac{\|X_i\|^2}{2(\rho+\sigma)}\right) & \mbox{if } c=k+1 \\
+                                                n_{-i,c}\exp\left(-\frac{\|X_i-\mu_c\|^2}{2\sigma}\right) & \mbox{if } c\leq k \\
+                                            \end{cases}
+\end{align*}
+For the conditional distribution of the $\mu_c$ given everything else, it yields, where $\bar X_c$ is the mean over all $X_i\st z_i=c$,
+\begin{align*}
+    \p(\mu_c=\tau\st X_{1:n},z_{1:n},\mu_{-i},\rho,\sigma) &\propto \p( \bar X_c \st n_c,\mu_c, \sigma)\p(\mu_c=\tau\st \rho) \\
+    &\propto \exp\left(\frac{1}{(\rho^{-2}+n\sigma^{-2})}\left\|\bar X_c - \frac{n_c\bar X_c}{\sigma^2(\rho^{-2}+n\sigma^{-2})}\right\|^2\right)
+\end{align*}
+Now, as one sends $\sigma$ to $0$, both of these probabilities become point masses.
+For the first probability, $\p(z_i=c \st X_{1:n},z_{-i},\mu_{1:k},\rho,\sigma)$, it is a point mass on $z_i=c\leq k$ if $\|\mu_c-X_i\|$ is minimal and less than $\lambda$, otherwise it is a point mass on $z_i=k+1$.
+The second distribution, of $ \mu_c\st X_{1:n},z_{1:n},\mu_{-i},\rho,\sigma$, is just a point mass on $\bar X_c$.
+Thus, the Gibbs sampling algorithm initialized with one cluster at the global mean turns into the DP-means algorithm, which is iterated till convergence:
+\begin{itemize}
+    \item Initialize $k=1$, $\mu_1 = \bar X$, and $z_i=1 \;\forall i$
+    \item For each $X_i$
+        \begin{itemize}
+            \item {\bf If $X_i$ is too far from a cluster, introduce a new one:} if \(\lambda \leq \|X_i-\mu_c\|^2 \; \forall c\), set $k=k+1$, $z_i=k$ and $\mu_k=X_i$
+            \item {\bf Otherwise, have it join the nearest cluster:} Otherwise $z_i=\argmin_c\|X_i-\mu_c\|^2$
+        \end{itemize}
+    \item {\bf Set the cluster centers to the mean of the cluster:} For each $c$, set $\mu_c = \bar X_c$
+\end{itemize}
+\subsection{Hierarchical DP-means}
+As is common with Bayesian models, a hierarchical model such as DP-means can be expanded to a more complicated model by adding an additional level of hierarchy by putting a probability distribution on some of the hyperparameters.
+This is exactly what the authors of this paper do to DP-means.
+They replace the original fixed Dirichlet process with a mixture of Dirichlet processes as determined by an additional Dirichlet process.
+This is called a hierarchical Dirichlet process.
+While this doesn't yield anything interesting for a single data set, this allows for a clustering algorithm that clustering over multiple data sets.
+The idea is that one might have multiple data sets of different observations, but with identical measurements taken for each, where each data set has a set of local clusters within it that don't necessarily align with the local clusters of another data set.
+This method will allow the simultaneous identification of a set of global clusters while the data sets borrow strength from each other to identify the local clusters within each data set.
+\par
+The ``Revisiting k-means:..." paper doesn't go into the full derivation of this model from a Bayesian setup, instead pointing to the original paper on the hierarchical Dirichlet process\footnote{Yee Teh, Micheal Jordan, Matthew Beal, David Blei, ``Hierarchical Dirichlet Processes", JASA, 1001(476):1566-1581, 2006.} and indicating that ``the limiting process described earlier for the standard DP can be straightforwardly extended to the HDP".
+To start, the Bayesian model for how the data arises goes something like this:
+\begin{itemize}
+    \item The $z_{ij}$ which are the local indicators of the group for the $i$th observation in the $j$th data set, are drawn first.
+Lets say, up to the current $z_{ij}$ being drawn, there are $k_{j.}$ groups in the data set, each group has $n_{jc}$ observations in it, and $z_{ij}$ is the $n_{j.}$th observation drawn.
+$z_{ij}$ is drawn locally by a Dirichlet process with parameter $\alpha$.
+        \begin{itemize}
+            \item As before, $z_{ij}=c\leq k_{j.}$ with probability $\frac{n_{jc}}{\alpha+n_{j.}-1}$
+            \item $z_{ij}=k_j+1$ with probability $\frac{\alpha}{\alpha+n_{j.}-1}$
+        \end{itemize}
+
+    \item $v_{jc}$, which is the association of the $c$th group in the $j$th data set with a global cluster, is drawn next.
+Let us say that there are $g$ global clusters, and $k_{.p}$ is the number of groups accross all  data set $j$ for which $v_{jc}=p$.
+$v_{jc}$ is drawn globally by a second Dirichlet process with parameter $\gamma$:
+        \begin{itemize}
+            \item $v_{jc}=p\leq g$ with probability $\frac{k_{.p}}{\gamma+k_{..}-1}$
+            \item $v_{jc}=g+1$ with probability $\frac{\gamma}{\gamma+k_{..}-1}$
+        \end{itemize}
+    \item All the global cluster centers $\mu_p$ are drawn iid from $N(0,\rho I)$ for some variance parameter $\rho$
+    \item Finally, each observed data point $X_{ij}\st z_{ij},v_{jz_{ij}} \sim N(\mu_{v_{jz_{ij}}}, \sigma I)$ for another variance parameter $\sigma$
+\end{itemize}
+After fixing $\sigma$ and $\rho$ in a similar manner to before, the conditionals can be constructed for a Gibbs sampler.
+Then $\alpha$ and $\gamma$ are in terms of local and global parameter $\lambda_l$ and $\lambda_g$.
+In the same way, as $\sigma$ is sent to $0$, this makes each conditional distribution becomes a point mass.
+This turns the Gibbs sampler into the "Hard Gaussian HDP", which is the hard clustering algorithm to fit the local and global clusters across data sets.
+The algorithm is this, iterated till convergence.
+\begin{itemize}
+    \item Initialize $g=1$, $k_j=1$, $\mu_1$ = $\bar X$, $z_{ij}=1$, $v_{j1}=1$ for all $i$ and $j$.
+    \item For each $X_{ij}$:
+        \begin{itemize}
+            \item {\bf If the point is too far from a global cluster, introduce a new one:}  \\
+                If
+                    \[\lambda_l<\|X_{ij}-\mu_{v_{jc}}\|^2 \;\forall c<k_j \mbox{ and } \lambda_l+\lambda_p<\|X_{ij}-\mu_{p}\|^2 \;\forall p\leq g\]
+                    set $k_j=k_j+1$, $g=g+1$, $z_{ij}=k_j$, $v_{jk_j}=g$, $\mu_g=X_{ij}$
+            \item {\bf If the point is too far away from a local cluster, but near a global cluster, add a new local cluster:} \\
+                If 
+                \[\min_c\|X_{ij}-\mu_{v_{jc}}\|^2>\min_p\|X_{ij}-\mu_{p}\|^2+\lambda_l\]
+                set \(k_j=k_j+1$, $z_{ij}=k_j$, and $v_{jz_{ij}}=\argmin_p\|X_{ij}-\mu_{p}\|^2$
+            \item {\bf Otherwise, match to one of the nearest clusters:}  \\
+                Set $z_{ij}=\argmin_c\|X_{ij}-\mu_{v_{jc}}\|^2$
+        \end{itemize}
+    \item For each local cluster $c_j$:
+        \begin{itemize}
+            \item{\bf If the local cluster is too far from a global cluster, introduce a new one:}
+                If 
+                \[\min_p\sum_{\{X_{ij}\st z_{ij}=c_j\}}\hspace{-1em} \|x-\mu_p\|^2\;>\;\lambda_l+ \hspace{-1em} \sum_{\{X_{ij}\st z_{ij}=c_j\}}\hspace{-1em}\|x-\bar X_{jc_j}\| \]
+                Set $g=g+1$, $v_{jc_j}=g$, and $\mu_g=\bar X_{jc_j}$
+            \item{\bf Otherwise, match each the local cluster to the nearest global cluster:}
+                Set 
+                \[v_{jc_j} = \argmin_p\sum_{\{X_{ij}\st z_{ij}=c_j\}}\hspace{-1em} \|x-\mu_p\|^2\]
+        \end{itemize}
+    \item For each global cluster $p$:
+        \begin{itemize}
+            \item {\bf Recenter each global cluster:}\\
+                Set
+                \[ \mu_p = \sum_{\{X_{ij}\st v_{jz_{ij}}=p\}}\hspace{-1em} x \]
+        \end{itemize}
+\end{itemize}
+Its worth noting that the final $z_{ij}$ aren't meaningful in the end, since observations are allocated to local clusters only based on the associated global cluster center.
+This means that the partition of points into local clusters which map to the same global cluster is nonidentifiable.
+Thus, the important value is the global cluster association of each point, $v_{jz_{ij}}$.
+\subsection{Spectral Clustering with DP-means}
+A convieient way of thinking of clustering is that one is trying to find the clustering such that when one replaces the observations with the clusters, the result is the minimal distortion of the data.
+Another way of saying this is that we want to project our data onto a set of vectors in the fashion that reduces the sum of the norms, or sum of squares, the least.
+This is akin to linear regression, where we are looking to project our data onto a lower dimmensional plane in a similar fashion.
+Just as with linear regression, where this projection was performed with the hat matrix $H$, for each clustering we can use a weight matrix $W$ to do this.
+$W_{ij}=\frac{1}{n_c}$ if $X_i$ and $X_j$ are to be placed in the same cluster, and $0$ otherwise.
+This means $WX$ replaces each row of $X$, so we can calculate are total distortion as $\trace(X'X)-\trace(W'X'XW)$.
+Since all the rows in $WX$ will be identical within a given cluster, we can replace this by a weighted sum of the unique cluster centers.
+Thus, $\trace(W'X'XW)=\trace(Y'X'XY)$, where $Y$ is the normalized $n\times k$ clustering matrix, which will have $Y_{ic}=\frac{1}{\sqrt n_c}$ if observation $i$ is in cluster $c$, and 0 otherwise.
+\par
+Both clustering algorithms, k-means and DP-means, attempt to minimize this distortion, or maximize $\trace(Y'X'XY)$, under a certain criteria.
+For k-means, that criteria is over all normalized clustering matricies $Y$ with $k$ is fixed, while for DP-means, it turns out the criteria is to minimize $\trace(Y'X'XY)-\lambda k$ over all normalized clustering matricies $Y$ and corresponding $k$.
+It turns out that a natural relaxation, or simplification of both of these targets is to let $Y$ be any orthonormal matrix rather than just a normalized clustering matrix.
+This is where the spectral part comes in; let $X'X=Q\Lambda Q'$ be the eigen decomposition of $X$, where $Q$ is the orthonormal matrix of eigenvalues and $\Lambda=\diag(\lambda_1,...,\lambda_n)$ is the diagonal matrix of eigenvalues in decreasing order.
+Its obvious that the optimal $Y$ in both cases will be a subset of the columns of $Q$.
+For k-means, it will be the the first $k$ columns, corresponding to the largest $k$ eignvalues, while for DP-means, it will be all columns such that $\lambda_p>\lambda$.
+\par
+The reason this relaxation is useful is two fold.
+First, when one has the relaxed matrix $Y$ above, they can perform a potentially simpler clustering on it, treating it as a new data matrix.
+This clustering will also work very well on the original data set, since $Y$ is a set of principle components from it.
+More importantly though, if one thinks of $X'X$ as a matrix meauring similarity between all the observations in a particular fashion, it can be replaced with another similarity matrix $K$ which uses a different critiera, for instance $K_{ij} = \exp\left(-\|X_i-X_j\|^2\right)$.
+Then, one similarly can make a matrix $Y$ which is the first $k$ eigenvectors for k means or all eigenvectors with eigenvalues greater than $\lambda$ for DP-means.
+The advantage of this is when $X$ is not a nice mix of gaussian clusters, as needed for k-means or DP-means to work well, one can use a different measure of similarity to generate $K$ and $Y$ such that $Y$ is approximately a nice mixture of Gaussians.
+
+\subsection{Remaining Content}
+The remaining part of the paper had two sections.
+The first of these has to do with graph clustering.
+The paper only gives a very brief desciription of the result, without enough detail to understand any of it without knowing the area already.
+The gist is that there is a natural extension of DP-means to the graph cut problem, just as there is with k-means.
+It turns out that, as has been demonstrated similarlily in the three earlier sections, DP-means ends up searching for an equivalent optimal result as k-means given the right parameterization.
+\par
+The final section deals with a number of simulations that the authors conducted comparing fitting models by k-means, DP-means, and a full Dirichlet process mixture with Gibbs sampling.
+On individual data sets, they all generally produce similar quality clusters.
+On large data sets though, Gibbs sampling becomes computationally infeasible while both k-means and DP-means run in similar time.
+The impressive result though come from the hard Gaussian HDP.
+To test it, they created 15 cluster centers, then created 50 data sets which each had 5 data points generated from 5 randomly chosen clusters.
+The hard Gaussian HDP clustering, fit accross all data sets, performed considerably better than k-means or DP-means fit on all the data sets together or each data set individually.
+
+\section{Similarities to Class Material}
+The central idea which this paper uses that we learned in class is that of the hiearchical model.
+In the based DP-algorithm, the underlying Bayesian model treats the observations $X_i$ as exchangable, allowing the $X_i$, although independent conditioned on their clusters, to lend strength to each other in infering both their cluster associations and the cluster means.
+This concept, where the observations share information, is a central one we saw in many contexts in class.
+The hard Gaussian HDP model displayed another precept of heiarchical models we saw inclass; Bayesian approaches are accomodating to additional levels of hiearchy being added to models.
+This is exactly how the hard Gaussian HDP was developed.
+They added an additional level of hiearchy to the DP-means model so that observations wouldn't just lend strength to each other within a data set but also accross data sets.\par
+The second important idea we saw before in class was fitting the model by a MCMC process, in particular Gibbs sampling.
+There model was complicated to the point that deriving an analytical solution would be at best infeasible.
+Since the marginal distributions of parameters could be expressed in closed form, this allowed a Gibbs sampler to be set up.
+An interesting twist I thought was stretching the Gibbs Sampler till it was deterministic so as to derive a hard clustering algorithm.
+While it wasn't like the other markov chains we saw which were irreducible, instead having an absorbing state, it still proved useful in fitting the model.
+\par
+The final idea from class that stuck out was the idea of an uniformative prior.
+While this isn't how it was described in the paper, several of the priors struck me as being set up to be largely uninformative.
+For instance, by sending the within cluster variance to 0 while the cluster mean variance was fixed, the result was essentially a uniform prior on $\R^p$ for the cluster means, such that within a cluster the center was just set to the mean.
+Certainly, taking the limit of a distribution as a parmeter goes to 0 or infinity was a common trick we saw to set up uninformative priors.
+\section{My Thoughts}
+On the whole, I found this paper very interesting to read, and learned a lot from it and the related papers I looked up.
+However, as a practical matter, they established DP-means as a largely equivalent method to k-means rather than a superior method.
+Throught the paper, they showed that they the two methods acheived largely similar results in similar times in several different areas where they could both be used.
+The main difference was in how they are parameterized.
+In practice, with either one, I would want to try several different values of the parameter and compare the fit.
+I'm not sure why DP-means would make deciding on the proper parameter easier.
+I would just be looking for a number of parameters to map to the same number of clusters with DP-means, while with k-means I would be looking to a falloff in the marginal improvement in distortion.
+I like that its parameter, $\lambda$ is akin to the complexity parameter we see in a lot of machine learning algorithms, but that is more from an aesthetic point of view than anything else.
+\par
+The value of the Bayesian approach, for me, comes with the hard Gaussian HDP algorithm.
+As mentioned before, the adding hiearchy to Bayesian models is natural to the extent it often isn't with other models, which is why there is not, or I am unaware of, a k-means based algorithm which can similarly be fit simaltaneously accross many data sets.
+I would certainly choose this algorithm if I ran into a situation where I was trying to fit clusters accross many data sets and had reason to expect that not all clusters were within all data sets.
+That said, I think that situation is a somewhat rare occurance.
+Either way, it is a good tool to have, and learning about hiearchical Dirichlet processes through this algorithm was interesting.
+\par
+An extension which I would have liked to see in the paper would be an algorithm for hiearchical clustering in the more traditional sense, where the base clusters within one data set are clustered into meta clusters within that one data set.
+For instance, one might imagine cluster centers $\mu_{ij}\sim N(\nu_i,\sigma I)$ with $\nu_i\sim N(0,\rho I)$.
+This is fertile grounds for a similar Bayesian algorithm based on $DP-means$, and it seems like something which wouldn't be hard to develop.
+\section{Big Picture}
+This is a fairly recent paper, adding to the relatively recent area of Bayesian nonparametrics which I know at least one of the authors, Micheal Jordan, has been fairly central to.
+The uniting thread of this field seems to be attempts to model objects which can have infinite dimmension with a countably infinite, nonparametric discrete distrubtion.
+The dirichlet process, which is the conjugate prior to this family of distributions, thus tends to pop up a lot, as it does here.
+This process, though a newer development, is based on the Dirichlet distribution, which has long been a tool of traditional Bayesian statistics.
+Bayesian nonparametrics is also closely tied to the more general statistics/computer science field of machine learning, with many of the methods of the former being adopted by the latter.
+This paper was, in large part, an attempt of the authors to push this process further by offering a bayesian nonparametric algorithm which has appealing properties to machine learning, mainly scalability.
 

### 2015-06-10 stat302/paper_report.tex
Finished paper

@@
 
-In designing the DP-means algorithm, the authors ``attempt to achieve the best of both worlds by designing scalable hard clustering algorithms from a Bayesian nonparametric viewpoint." To this end, they start with a purely Bayesian model for how the data would arise which puts a prior on $k$, and then similarly derives a simple algorithm for hard clustering that represents the limit of fitting this Bayesian algorithm as a variance parameter is sent to 0.
+In designing the DP-means algorithm, the authors ``attempt to achieve the best of both worlds by designing scalable hard clustering algorithms from a Bayesian nonparametric viewpoint." To this end, they start with a purely Bayesian model for how the data would arise.
+Here, a prior is put on $k$, the weights of each cluster, the distribution of the centers, and the conditional distributions of the $X_i$ given the centers.
+Then, as with k-means, a hard clustering algorithm is the limit of a Gibbs sampler as a variance parameter goes to 0.
 The Bayesian model for how the data arises sequentially is something like this:
@@
 The authors fix $\rho$ and $\sigma$, and then set $\alpha=\left(1+\frac{\rho}{\sigma}\right)^{d/2}\exp\left(-\frac{\lambda}{2\sigma}\right)$, where $d$ is the dimension of the $X_i$.
-To set up a Gibbs sampling algorithm, we can calculate the conditional distributions of the unknown parameters $z_1,...,z_n$ and $\mu_1,...,\mu_k$.
-Let us say that there are $k$ clusters, excluding $z_i,X_i$, at any given point.
-Also, let the number of points assigned to cluster $c$ excluding $z_i,X_i$ is $n_{-i,c}$.
+To set up a Gibbs sampling algorithm, the authors calculate the conditional distributions of the unknown parameters $z_1,...,z_n$ and $\mu_1,...,\mu_k$.
+Let us say that there are $k$ clusters, excluding observation $i$, at any given point.
+Also, let the number of points assigned to cluster $c$ excluding point $i$ be $n_{-i,c}$.
 This yields, for the conditional distribution of the $z_i$,
@@
         \begin{itemize}
-            \item {\bf If $X_i$ is too far from a cluster, introduce a new one:} if \(\lambda \leq \|X_i-\mu_c\|^2 \; \forall c\), set $k=k+1$, $z_i=k$ and $\mu_k=X_i$
-            \item {\bf Otherwise, have it join the nearest cluster:} Otherwise $z_i=\argmin_c\|X_i-\mu_c\|^2$
+            \item {\bf If $X_i$ is too far from a cluster, introduce a new one:} \\
+                if 
+                \[\lambda \leq \|X_i-\mu_c\|^2 \; \forall c\]
+                set $k=k+1$, $z_i=k$ and $\mu_k=X_i$
+            \item {\bf Otherwise, have it join the nearest cluster:} \\
+                Otherwise 
+                \[z_i=\argmin_c\|X_i-\mu_c\|^2\]
         \end{itemize}
@@
 This is called a hierarchical Dirichlet process.
-While this doesn't yield anything interesting for a single data set, this allows for a clustering algorithm that clustering over multiple data sets.
-The idea is that one might have multiple data sets of different observations, but with identical measurements taken for each, where each data set has a set of local clusters within it that don't necessarily align with the local clusters of another data set.
-This method will allow the simultaneous identification of a set of global clusters while the data sets borrow strength from each other to identify the local clusters within each data set.
+While this doesn't yield anything interesting for a single data set, this allows for a clustering algorithm that clusters over multiple data sets.
+The idea is that one might have multiple data sets of different observations, but with identical measurements taken for each.
+In this setting each data set might have a set of local clusters partially, but not fully, align with local clusters of each other data set.
+This method will allow the simultaneous identification of a set of global clusters while the data sets borrow strength from each other to identify the shared local clusters within each data set.
 \par
 The ``Revisiting k-means:..." paper doesn't go into the full derivation of this model from a Bayesian setup, instead pointing to the original paper on the hierarchical Dirichlet process\footnote{Yee Teh, Micheal Jordan, Matthew Beal, David Blei, ``Hierarchical Dirichlet Processes", JASA, 1001(476):1566-1581, 2006.} and indicating that ``the limiting process described earlier for the standard DP can be straightforwardly extended to the HDP".
-To start, the Bayesian model for how the data arises goes something like this:
+To partially describe what this would look like, the Bayesian model for how the data arises goes something like this:
 \begin{itemize}
-    \item The $z_{ij}$ which are the local indicators of the group for the $i$th observation in the $j$th data set, are drawn first.
-Lets say, up to the current $z_{ij}$ being drawn, there are $k_{j.}$ groups in the data set, each group has $n_{jc}$ observations in it, and $z_{ij}$ is the $n_{j.}$th observation drawn.
-$z_{ij}$ is drawn locally by a Dirichlet process with parameter $\alpha$.
+    \item The $z_{ij}$, which are the local indicators of the group for the $i$th observation in the $j$th data set, are drawn first.
+Lets say, up to the current $z_{ij}$ being drawn, there are $k_{j.}$ local clusters in the data set, each group has $n_{jc}$ observations in it, and $z_{ij}$ is the $n_{j.}$th observation drawn.
+In this setting, $z_{ij}$ is drawn locally by a Dirichlet process with parameter $\alpha$.
         \begin{itemize}
@@
     \item $v_{jc}$, which is the association of the $c$th group in the $j$th data set with a global cluster, is drawn next.
-Let us say that there are $g$ global clusters, and $k_{.p}$ is the number of groups accross all  data set $j$ for which $v_{jc}=p$.
-$v_{jc}$ is drawn globally by a second Dirichlet process with parameter $\gamma$:
+Let us say that there are $g$ global clusters,$k_{..}$ is the total number of local clusters, and $k_{.p}$ is the number of local clusters, in all data sets, for which $v_{jc}=p$.
+In this setting, $v_{jc}$ is drawn globally by a second Dirichlet process with parameter $\gamma$:
         \begin{itemize}
@@
 \end{itemize}
+Its important to note that the particular local cluster matters only so far as which global cluster it maps to. All the $X_i$ that are map to the same global cluster are iid. \par
 After fixing $\sigma$ and $\rho$ in a similar manner to before, the conditionals can be constructed for a Gibbs sampler.
@@
                 If
-                    \[\lambda_l<\|X_{ij}-\mu_{v_{jc}}\|^2 \;\forall c<k_j \mbox{ and } \lambda_l+\lambda_p<\|X_{ij}-\mu_{p}\|^2 \;\forall p\leq g\]
+                    \[\lambda_l<\|X_{ij}-\mu_{v_{jc}}\|^2 \;\;\forall c<k_j \mbox{ and } \lambda_l+\lambda_p<\|X_{ij}-\mu_{p}\|^2 \;\;\forall p\leq g\]
                     set $k_j=k_j+1$, $g=g+1$, $z_{ij}=k_j$, $v_{jk_j}=g$, $\mu_g=X_{ij}$
@@
             \item {\bf Otherwise, match to one of the nearest clusters:}  \\
-                Set $z_{ij}=\argmin_c\|X_{ij}-\mu_{v_{jc}}\|^2$
+                Set 
+                \[z_{ij}=\argmin_c\|X_{ij}-\mu_{v_{jc}}\|^2\]
+                There may be multiple $c$ which fit this criteria, in which case one should be chosen randomly.
         \end{itemize}
@@
 \end{itemize}
-Its worth noting that the final $z_{ij}$ aren't meaningful in the end, since observations are allocated to local clusters only based on the associated global cluster center.
-This means that the partition of points into local clusters which map to the same global cluster is nonidentifiable.
+As mentioned above, the local clustering is irrelevant in the end.
+This means that the partition of points which map to the same global cluster into local clusters is nonidentifiable.
 Thus, the important value is the global cluster association of each point, $v_{jz_{ij}}$.
+
 \subsection{Spectral Clustering with DP-means}
-A convieient way of thinking of clustering is that one is trying to find the clustering such that when one replaces the observations with the clusters, the result is the minimal distortion of the data.
+A convieient way of thinking of fitting clusters is that one is trying find the set of cluster centers which subsititue for the original observations with minimal distortion.
 Another way of saying this is that we want to project our data onto a set of vectors in the fashion that reduces the sum of the norms, or sum of squares, the least.
-This is akin to linear regression, where we are looking to project our data onto a lower dimmensional plane in a similar fashion.
+This is akin to linear regression, where we are looking to project our data onto a plane.
 Just as with linear regression, where this projection was performed with the hat matrix $H$, for each clustering we can use a weight matrix $W$ to do this.
-$W_{ij}=\frac{1}{n_c}$ if $X_i$ and $X_j$ are to be placed in the same cluster, and $0$ otherwise.
-This means $WX$ replaces each row of $X$, so we can calculate are total distortion as $\trace(X'X)-\trace(W'X'XW)$.
+Its definition is that $W_{ij}=\frac{1}{n_c}$ if $X_i$ and $X_j$ are to be placed in the same cluster, and $0$ otherwise.
+This means $WX$ replaces each row of $X$ with its cluster mean.
+We can accordingly calculate a clustering's total distortion as $\trace(X'X)-\trace(W'X'XW)$.
 Since all the rows in $WX$ will be identical within a given cluster, we can replace this by a weighted sum of the unique cluster centers.
@@
 For k-means, that criteria is over all normalized clustering matricies $Y$ with $k$ is fixed, while for DP-means, it turns out the criteria is to minimize $\trace(Y'X'XY)-\lambda k$ over all normalized clustering matricies $Y$ and corresponding $k$.
-It turns out that a natural relaxation, or simplification of both of these targets is to let $Y$ be any orthonormal matrix rather than just a normalized clustering matrix.
-This is where the spectral part comes in; let $X'X=Q\Lambda Q'$ be the eigen decomposition of $X$, where $Q$ is the orthonormal matrix of eigenvalues and $\Lambda=\diag(\lambda_1,...,\lambda_n)$ is the diagonal matrix of eigenvalues in decreasing order.
+A natural relaxation, or simplification of both of these targets is to let $Y$ be any orthonormal matrix rather than just a normalized clustering matrix.
+This is where the spectral part comes in; let $X'X=Q\Lambda Q'$ be the eigen decomposition of $X'X$, where $Q$ is the orthonormal matrix of eigenvalues and $\Lambda=\diag(\lambda_1,...,\lambda_n)$ is the diagonal matrix of eigenvalues in decreasing order.
 Its obvious that the optimal $Y$ in both cases will be a subset of the columns of $Q$.
@@
 This clustering will also work very well on the original data set, since $Y$ is a set of principle components from it.
-More importantly though, if one thinks of $X'X$ as a matrix meauring similarity between all the observations in a particular fashion, it can be replaced with another similarity matrix $K$ which uses a different critiera, for instance $K_{ij} = \exp\left(-\|X_i-X_j\|^2\right)$.
+More importantly though, if one thinks of $X'X$ as a matrix meauring similarity between all the observations in a particular fashion, it can be replaced with another similarity matrix $K$ which uses a different critiera.
+For instance $K_{ij} = \exp\left(-\|X_i-X_j\|^2\right)$ is common.
 Then, one similarly can make a matrix $Y$ which is the first $k$ eigenvectors for k means or all eigenvectors with eigenvalues greater than $\lambda$ for DP-means.
-The advantage of this is when $X$ is not a nice mix of gaussian clusters, as needed for k-means or DP-means to work well, one can use a different measure of similarity to generate $K$ and $Y$ such that $Y$ is approximately a nice mixture of Gaussians.
+The advantage of this is when $X$ is not a nice gaussian mixture, as needed for k-means or DP-means to work well, one can use a different measure of similarity to generate $K$ and $Y$ such that $Y$ is a proper mixture.
 
@@
 To test it, they created 15 cluster centers, then created 50 data sets which each had 5 data points generated from 5 randomly chosen clusters.
-The hard Gaussian HDP clustering, fit accross all data sets, performed considerably better than k-means or DP-means fit on all the data sets together or each data set individually.
+The hard Gaussian HDP clustering, fit accross all data sets, performed considerably better than k-means or DP-means fit on all the data sets pooled together or each data set individually.
 
 \section{Similarities to Class Material}
-The central idea which this paper uses that we learned in class is that of the hiearchical model.
-In the based DP-algorithm, the underlying Bayesian model treats the observations $X_i$ as exchangable, allowing the $X_i$, although independent conditioned on their clusters, to lend strength to each other in infering both their cluster associations and the cluster means.
+The central idea from our class which this paper uses is that of the hiearchical model.
+In the basic DP-algorithm, the underlying Bayesian model treats the observations $X_i$ as exchangable.
+This allows the $X_i$, although independent conditioned on their clusters, to lend strength to each other in infering both their cluster associations and the cluster means.
 This concept, where the observations share information, is a central one we saw in many contexts in class.
-The hard Gaussian HDP model displayed another precept of heiarchical models we saw inclass; Bayesian approaches are accomodating to additional levels of hiearchy being added to models.
+The hard Gaussian HDP model displayed another precept of heiarchical models we saw; Bayesian approaches are accomodating to additional levels of hiearchy being added to models.
 This is exactly how the hard Gaussian HDP was developed.
-They added an additional level of hiearchy to the DP-means model so that observations wouldn't just lend strength to each other within a data set but also accross data sets.\par
+They added an additional level of hiearchy to the DP-means model by putting a prior on the Dirichlet process for each data set.
+This allows for different mixtures in each data set, while still lending strength accross all data sets on several levels.\par
 The second important idea we saw before in class was fitting the model by a MCMC process, in particular Gibbs sampling.
@@
 While this isn't how it was described in the paper, several of the priors struck me as being set up to be largely uninformative.
-For instance, by sending the within cluster variance to 0 while the cluster mean variance was fixed, the result was essentially a uniform prior on $\R^p$ for the cluster means, such that within a cluster the center was just set to the mean.
+For instance, by sending the within cluster variance to 0 while the cluster mean variance was fixed, the result was essentially a uniform prior on $\R^p$ for the cluster centers.
+The result was that the cluster centers were estimated as the mean, rather than a mix of the data mean and prior mean.
 Certainly, taking the limit of a distribution as a parmeter goes to 0 or infinity was a common trick we saw to set up uninformative priors.
@@
 However, as a practical matter, they established DP-means as a largely equivalent method to k-means rather than a superior method.
-Throught the paper, they showed that they the two methods acheived largely similar results in similar times in several different areas where they could both be used.
+Through out the paper, they showed that they the two methods acheived similar results in similar times, and were both applicable in the same contexts.
 The main difference was in how they are parameterized.
@@
 I'm not sure why DP-means would make deciding on the proper parameter easier.
-I would just be looking for a number of parameters to map to the same number of clusters with DP-means, while with k-means I would be looking to a falloff in the marginal improvement in distortion.
+I would just be looking for a number of parameters to map to the same number of clusters with DP-means, while with k-means I would be looking for a falloff in the marginal improvement in distortion.
 I like that its parameter, $\lambda$ is akin to the complexity parameter we see in a lot of machine learning algorithms, but that is more from an aesthetic point of view than anything else.
@@
 The value of the Bayesian approach, for me, comes with the hard Gaussian HDP algorithm.
-As mentioned before, the adding hiearchy to Bayesian models is natural to the extent it often isn't with other models, which is why there is not, or I am unaware of, a k-means based algorithm which can similarly be fit simaltaneously accross many data sets.
+As mentioned before, adding the hiearchy to Bayesian models is natural in a fashion it often isn't with other types of models.
+I assume this is why there is not, or I am unaware of, a k-means based algorithm which can similarly be fit simaltaneously accross many data sets.
 I would certainly choose this algorithm if I ran into a situation where I was trying to fit clusters accross many data sets and had reason to expect that not all clusters were within all data sets.
@@
 For instance, one might imagine cluster centers $\mu_{ij}\sim N(\nu_i,\sigma I)$ with $\nu_i\sim N(0,\rho I)$.
-This is fertile grounds for a similar Bayesian algorithm based on $DP-means$, and it seems like something which wouldn't be hard to develop.
+This is fertile grounds for a similar Bayesian algorithm based on DP-means, and it seems like something which wouldn't be hard to develop.
 \section{Big Picture}
-This is a fairly recent paper, adding to the relatively recent area of Bayesian nonparametrics which I know at least one of the authors, Micheal Jordan, has been fairly central to.
-The uniting thread of this field seems to be attempts to model objects which can have infinite dimmension with a countably infinite, nonparametric discrete distrubtion.
+This is a recent paper, only published in 2012, which adds to the relatively recent area of Bayesian nonparametrics which I know at least one of the authors, Micheal Jordan, has been central to.
+The uniting thread of this field seems to be attempts to model objects which can have infinite dimmension with a countably infinite, nonparametric discrete distrubtion prior.
 The dirichlet process, which is the conjugate prior to this family of distributions, thus tends to pop up a lot, as it does here.
+Another example is that one can put a prior on the space of cumulative probability functions by drawing an countably infinite number of point masses and assigning weight to them based on a dirichlet process.
 This process, though a newer development, is based on the Dirichlet distribution, which has long been a tool of traditional Bayesian statistics.
 Bayesian nonparametrics is also closely tied to the more general statistics/computer science field of machine learning, with many of the methods of the former being adopted by the latter.
-This paper was, in large part, an attempt of the authors to push this process further by offering a bayesian nonparametric algorithm which has appealing properties to machine learning, mainly scalability.
+This paper was, in large part, an attempt of the authors to push this process further by offering a bayesian nonparametric algorithm which has appealing properties to machine learning, in particular scalability.
 

### 2015-06-17 masters_paper/notes.tex
Finishing the generating random binary section

@@
 \pdfpageheight 11in
-\setlength\parindent{0in}
 
@@
 
+\newcommand{\x}{\bf x}
+
 \newcommand{\Example}{\noindent {\bf Example. \quad} }
@@
 \section{Preliminaries} 
-Some early investigation into how the deterministic knockoffs as described in Rina's paper work with regularized logistic regression revealed that the answer is "not very well".
-Even when $X$ is a null predictor of $y$, the $X$ still tend to enter the model prior to $\tilde X$.
-The issue is that even when $X_i\sim N_p({\bf 0}, \Sigma)$ for some $\Sigma\succeq 0$, $\tilde X$ is not normally distributed.
-This can be seen from producing qq plots of $X_i$ vs $\tilde X_j$Of course, when $X_i$ is a binary vector, $\tilde X_i$ completely doesn't match its distribution, causing the original $X$ to beat the knockoffs into the model.
-This indicates that a new method of generating $\tilde X$ must be created to use with FDR via knockoffs for regularized logistic regression.
+{\em Note:} I will try to hold to the convention that $X$ is the $n\times p$ data matrix, while $\bf x$ is the random vector variable from which each row of $X$ was drawn.
+Accordingly, $\tilde X$ will be the knockoff matrix while $\bf \tilde x$ is a random vector variable.
+$x_i$ will be the random variable corresponding to the $i$th entry of $\bf x$, while $X_i$ is the vector of observations drawn from $x_i$ in the data matrix.
+\\
+
+Some early investigation into deterministic knockoffs (as described in the original knockoff paper) reveal that they don't perform well in L1 regularized logistic regression.
+Even when $X_i$ is a null predictor of $y$, the $X_i$ still tend to enter the model prior to $\tilde X_i$.
+The issue is that even when $x\sim N_p({\bf 0}, \Sigma)$ for some $\Sigma\succeq 0$, $\tilde x$ is not normally distributed.
+This can be seen from producing qq plots of $X_i$ vs $\tilde X_i$ for each $i$.
+Of course, when $X$ is a binary vector, $\tilde X$ completely doesn't match its distribution, causing the original $X$ to beat the knockoffs into the model.
+This indicates that a new method of generating $\tilde X$ must be created to control FDR via knockoffs with regularized logistic regression.
 
 \section{Probabilistic Random Bernoulli Knockoffs}
-My idea is to generate $\tilde X$ randomly such that, approximately, $\tilde X_i\sim X_i$.
-In particular, they should have similar marginal densities, expectations, and first moments.
-However, $\tilde X\st X$ should also have desired knockoff property that $\E(\tilde X'X\st X) = X'X-s$, where $\diag(X'X)-s$ is small.
-In the general case, this is likely infeasible; however, if $X$ is a binary vector, as is often the case, we know we are dealing with a much more limited class of random variables, and it should be possible to randomly generate $\tilde X\st X$ so as to have the desired properties.
-At worst, this method will provide a suitable replacement for deterministic $\tilde X$ as described in Rina's paper for LASSO, and if we are lucky, it will work reasonably for other regularized GLMs.
+My idea is to generate $\bf \tilde x$ randomly such that, approximately, $\bf \tilde x\sim \bf x$.
+In particular, both variables should have similar marginal densities, expectations, and second moments.
+However, $\bf \tilde x\st \bf x$ should also have desired knockoff property that $\E(\bf \tilde x'\bf x\st \bf x) = \bf x'\bf x-s$, where $\diag(\bf x'\bf x)-s$ is small.
+In the general case, this is likely infeasible; however, if $\bf x$ is a binary vector, as is often the case, we know we are dealing with a much more limited class of random variables, and it should be possible to randomly generate $\bf \tilde x\st \bf x$ so as to have the desired properties.
+At worst, this method will provide a suitable replacement for deterministic $\bf \tilde x$ for use with LASSO, and if we are lucky, it will work reasonably for other regularized GLMs.
 
@@
 Thankfully, there has been a reasonable amount of work on how one can generate random Bernoulli vectors with some kind of correlation among among the values.
-A random Bernoulli vector $X$ can be easiest represented with a mean vector $\E(X)=m\in(0,1)^p$ and $\E(XX')=M\in(0,1)^{p\times p}$, called the cross moment matrix.
-Obviously, $m_i = \p(X_i=1)$, $M_{ij} = \p(X_i=X_j=1)$, and $m=\diag(M)$.
-For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-\diag(M)\diag(M)'$ must be PSD, and
+A random Bernoulli vector $\bf x$ can be summarized by its first two moments: a mean vector $\E(\bf x)=m\in(0,1)^p$ and cross-moment matrix $\E(\bf x\bf x')=M\in(0,1)^{p\times p}$.
+Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $m=\diag(M)$.
+For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-mm'\succeq0$, and
 \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
 for all $i\neq j$\footnote{``On parametric families for sampling binary data with specified mean and correlation" - http://arxiv.org/abs/1111.0576}.
-Given a qualifying $M$, or observed $X$, there are a few ways of generating more random $X$.
+Given a qualifying $M$, or observed $X$, there are a few ways of generating more random $\bf x$.
 
 \subsection{Gaussian Copula Family}
-Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $Z~N_p({\bf 0},\Sigma)$ such that,for $X_j = I(Z_j<0)$, $X$ has the desired properties.
+Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $z\sim N_p({\bf 0},\Sigma)$ such that,for $x_i = I(z_i<0)$, $x$ has the desired properties.
 There are a number of ways to do this\footnote{``On the Generation of Correlated Artificial Binary Data" - http://epub.wu.ac.at/286/1/document.pdf}\footnote{``On parametric families for sampling binary data with specified mean and correlation"}, but it turns out that there is only certain to exist a working $\Sigma$ in the bivariate case.
@@
 It turns out that there exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as
-\[ X_{j}\st X_{j},...,X_{j-1} \sim \mathrm{B}\left(1,\mu\left(a_{jj}+\sum_{k=1}^{j-1}a_{kj}X_j\right)\right) \]
+
+\[ \p(x_{i}=1\st x_{1},...,x_{i-1}) = \mu\left(a_{ii}+\sum_{k=1}^{i-1}a_{ik}x_i\right) \]
+
 for some monotone function $\mu:\R\to[0,1]$.
@@
 If one takes all of the $a_{kj}$, they can form a lower triangular matrix $A$, and then the joint density can be expressed as
-\[ \p(X_j=\gamma) \propto \mu(\gamma'A\gamma)\]
-If $\mu$ is chose such that it is a bijection and differentiable, there is a unique $M$ such that $\E(X_iX_i')=M$\footnote{``On parametric families for sampling binary data with specified mean and correlation"}.
-As one might guess, the natural $\mu$ is the logistic link function, which is the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
-\[A_{ij} = \log\left(\frac{\p(X_j=X_k=1)\p(X_j=X_k=0)}{\p(X_j=0,X_k=1)\p(X_j=1,X_k=0)}\right) \]
-when $i\neq j$. I think this dictates that if $A_{jk}=0$, then $X_k$ and $X_j$ are independent.
+
+\[ \p(\mathbf x=\mathbf \gamma) \propto \mu(\mathbf{\gamma}' A\mathbf \gamma)\]
+
+If $\mu$ is chose such that it is a bijection and differentiable, there is a unique $M$ such that $\E(x_ix_i')=M$\footnote{``On parametric families for sampling binary data with specified mean and correlation"}.
+It turns out that the natural choice for $\mu$ is the logistic link function, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
+\[a_{ij} = \log\left(\frac{\p(x_j=x_k=1)\p(x_j=x_k=0)}{\p(x_j=0,x_k=1)\p(x_j=1,x_k=0)}\right) \]
+when $i\neq j$. I think this dictates that if $\bf x$ is generated from this model with $a_{ij}=0$, then $x_i$ and $x_j$ are independent.  \par
+
+There is no closed form to calculate the entries in $A$ if $p>1$, but they can be derived numerically two ways.  
+\begin{enumerate}
+    \item If one is attempting to replicate the empirical cross-moments from a data matrix $X$, $a_{1i}$ to $a_{ii}$ can be derived from fitting successive logistic regressions of $X_i$ on $X_{1} \ldots X_{i-1}$ using maximum likelihood.
+$a_{ji}$ for $i\neq j$ will then be the coefficient on $X_j$ while $a_{ii}$ is the intercept of the regression.
+    \item If one is just working with a desired cross-moment matrix $M$, the successive rows of $A$ can be fit via Newton-Raphson. \par
+        Let us say that the first $i-1$ rows have already been fit, resulting in the upper left $(i-1)\times(i-1)$ sub matrix $A_{-i}$ of $A$.
+Let us say that $\mathbf a_i$ is the first $i$ entries of the $i$th row of $A$ (the rest will be 0 anyway).
+As well, let $\mathbf m_i$ be similarly the first $i$ entries of the $i$th row of $M$.
+In other words, $\mathbf{m}_i = [\E(x_i x_j)]_{j=1}^{i}$.
+Finally, let us say that $\mathbf x_{-i}$ is the first $i-1$ entries of $\mathbf x$.
+We want to solve for $\mathbf a_i$ such that
+        \begin{align*}
+            \mathbf m_i &= \E\left(x_i \left[\begin{array}{c} \mathbf x_{-i} \\ x_i \end{array}\right]\right) \\
+            \mathbf m_i &= \E\left(\E\left(x_i \left[\begin{array}{c} \mathbf x_{-i} \\ x_i \end{array}\right]\;\bigg \vert\; \mathbf x_{-i}\right)\right)\\
+            \mathbf m_i &= \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \p(\mathbf x_{-i})\p(x_i=1\st \mathbf x_{-i}) \left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \\
+    \mathbf m_i &= \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}\mu\left(\mathbf x_{-i}'A_{-i}\mathbf x_{-i}\right) \mu\left(\mathbf{a}_i'\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \\
+        \end{align*}
+        Where $c$ is the appropriate normalizing constant.
+Let us define the quantity on the right in the last line as $f(\mathbf a_i)$.
+We can solve for $f(\mathbf a_i)=\mathbf{m}_i$ by successive Newton-Raphson iterations defined by
+        \[\mathbf a_i^{(k+1)} = \left[H f\left(\mathbf a_i^{(k)}\right)\right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
+        The Hessian matrix is calculated as 
+    \[ H f\left(\mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}\mu \left(\mathbf x_{-i}'A_{-i}\mathbf x_{-i}\right) \mu'\left(\mathbf{a}_i'\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}' & 1 \end{array}\right] \]
+        With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive.
+Instead, with a series of values $\mathbf x_{-i}^{(k)}\sim \mathbf x_{-i}$, we can approximate
+        \[f\left(\mathbf a_i\right) \approx \frac{1}{N}\sum_{k=1}^N \mu\left(\mathbf{a}_i'\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right] \]
+        and
+        \[H f\left(\mathbf a_i\right) \approx \frac{1}{N}\sum_{k=1}^N \mu'\left(\mathbf{a}_i'\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]' & 1 \end{array}\right] \]
+
+\end{enumerate}
 

### 2015-06-19 masters_paper/notes.tex
First draft of notes done

@@
 $x_i$ will be the random variable corresponding to the $i$th entry of $\bf x$, while $X_i$ is the vector of observations drawn from $x_i$ in the data matrix.
-\\
+$\hat \Sigma = \frac{1}{n}X^TX$ is the empirical covariance matrix associated with $X$, while $\Sigma=\E(\mathbf x \mathbf x^T)$ is the theoretical covariance matrix associate with $\mathbf x$.
+Also, for a square matrix $A$, $\diag(A)$ if the vector of values along the diagonal, while for a vector $\mathbf a$, $\diag(\mathbf a)$ is a square diagonal matrix with $\mathbf a$ along the diagonal.\\
 
@@
 
-\section{Probabilistic Random Bernoulli Knockoffs}
+\section{Random Bernoulli Knockoffs}
     My idea is to generate $\bf \tilde x$ randomly such that, approximately, $\bf \tilde x\sim \bf x$.
 In particular, both variables should have similar marginal densities, expectations, and second moments.
-However, $\bf \tilde x\st \bf x$ should also have desired knockoff property that $\E(\bf \tilde x'\bf x\st \bf x) = \bf x'\bf x-s$, where $\diag(\bf x'\bf x)-s$ is small.
+However, $\bf \tilde x\st \bf x$ should also have desired knockoff property that $\E(\mathbf{\tilde x}^T\mathbf x\st \mathbf x) = \Sigma-\mathbf s$, where $\|\diag(\Sigma)-s\|$ is small.
 In the general case, this is likely infeasible; however, if $\bf x$ is a binary vector, as is often the case, we know we are dealing with a much more limited class of random variables, and it should be possible to randomly generate $\bf \tilde x\st \bf x$ so as to have the desired properties.
@@
     Thankfully, there has been a reasonable amount of work on how one can generate random Bernoulli vectors with some kind of correlation among among the values.
-A random Bernoulli vector $\bf x$ can be summarized by its first two moments: a mean vector $\E(\bf x)=m\in(0,1)^p$ and cross-moment matrix $\E(\bf x\bf x')=M\in(0,1)^{p\times p}$.
+A random Bernoulli vector $\bf x$ can be summarized by its first two moments: a mean vector $\E(\bf x)=m\in(0,1)^p$ and cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$.
 Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $m=\diag(M)$.
-For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-mm'\succeq0$, and
+For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-mm^T\succeq0$, and
     \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
@@
 
-\[ \p(\mathbf x=\mathbf \gamma) \propto \mu(\mathbf{\gamma}' A\mathbf \gamma)\]
+    \[ \p(\mathbf x=\mathbf \gamma) \propto \mu(\mathbf{\gamma}^T A\mathbf \gamma)\]
 
-If $\mu$ is chose such that it is a bijection and differentiable, there is a unique $M$ such that $\E(x_ix_i')=M$\footnote{``On parametric families for sampling binary data with specified mean and correlation"}.
+    If $\mu$ is chose such that it is a bijection and differentiable, there is a unique $M$ such that $\E(x_ix_i^T)=M$\footnote{``On parametric families for sampling binary data with specified mean and correlation"}.
 It turns out that the natural choice for $\mu$ is the logistic link function, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
@@
                 \mathbf m_i &= \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \p(\mathbf x_{-i})\p(x_i=1\st \mathbf x_{-i}) \left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \\
-    \mathbf m_i &= \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}\mu\left(\mathbf x_{-i}'A_{-i}\mathbf x_{-i}\right) \mu\left(\mathbf{a}_i'\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \\
+        \mathbf m_i &= \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}\mu\left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) \mu\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \\
             \end{align*}
@@
             The Hessian matrix is calculated as 
-    \[ H f\left(\mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}\mu \left(\mathbf x_{-i}'A_{-i}\mathbf x_{-i}\right) \mu'\left(\mathbf{a}_i'\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}' & 1 \end{array}\right] \]
+        \[ H f\left(\mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}\mu \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
             With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive.
 Instead, with a series of values $\mathbf x_{-i}^{(k)}\sim \mathbf x_{-i}$, we can approximate
-        \[f\left(\mathbf a_i\right) \approx \frac{1}{N}\sum_{k=1}^N \mu\left(\mathbf{a}_i'\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right] \]
+            \[f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K \mu\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right] \]
             and
-        \[H f\left(\mathbf a_i\right) \approx \frac{1}{N}\sum_{k=1}^N \mu'\left(\mathbf{a}_i'\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]' & 1 \end{array}\right] \]
-
+            \[H f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
+            Though in theory $A$ should always exist, in practice numerical issues may compound to the point that the Newton-Raphson method won't converge.
+In this case, one can solve instead for $\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
+    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & \E(x_i) \end{array} \right]^T \]
+            When $\tau=0$, this yields the original problem, while when $\tau=1$, it is treating $x_i$ as independent of $\mathbf x_{-i}$. The latter will always have the solution
+            \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{\E(x_i)}{1-\E(x_i)}\right) \end{array} \right]^T \]
+            The hope is that for some $\tau$ close to $0$, convergence can be achieved, only causing a slight distortion from the desired cross moments.
     \end{enumerate}
@@
 \section{Generating Knockoffs}
+    My method for generating binary knockoffs broadly involves two steps:
+    \begin{enumerate}
+        \item I use either the equal correlation method or the SDP method described in the original knockoff method to find $\mathbf s$ such that $\diag(\hat\Sigma)-\mathbf s$ is small and 
+        \[ \Sigma_L = \left[\begin{array}{cc}  \hat\Sigma & \hat\Sigma - \diag(\mathbf s )\\ \hat\Sigma - \diag(\mathbf s) & \hat\Sigma \end{array}\right] \succeq 0 \]
+        I use a subscript $L$ for large to denote items associated with the joint distribution of $[\mathbf x \; \mathbf{\tilde x}]$.
+If $\mathbf m_L = \E\left([\mathbf x \; \mathbf{\tilde x}]\right)^T = [\mathbf m \; \mathbf m]^T$, then the desired cross moment matrix of the joint distribution is
+        \[ M_L = \Sigma_L + \mathbf m_L\mathbf m_L^T\]
+        To ensure that this is a valid cross moment matrix for a binary random vector is that 
+        \[\max\{0,\mathbf m_{L,i} + \mathbf m_{L,j} -1\} \leq M_{L,ij} \leq \min\{\mathbf m_{L,i}, \mathbf m_{L,j}\} \]
+        I've built a check into the code for this, but in practice it shouldn't be a worry.
+This condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$.
+Since the we are either keeping the value $M_{L,ij}$ from a valid cross moment matrix or minimizing $\vert M_{L,ij} - \mathbf m_{L,j}\mathbf m_{L,i}\vert$, it would be very surprising if this condition was violated.
+
+        \item I can fit the matrix $A$ that will generate random binary variables similarly to the method described in section $3.2$ which have cross moments $X_L$.
+This can be used to generate the $\tilde x_i$ sequentially as $\tilde x_i \st \mathbf x,\tilde x_1,\ldots,\tilde x_{i-1}$ so as to create $\tilde X \st X$.
+    \end{enumerate}
+    \subsection{More Detail on Fitting $A$}
+        I could fit $A$ based on $M_L$ exactly as described in the second method of $3.2$, however, this isn't exactly what I do.
+First off, with $p$ being potentially large, the simulation method for estimating $f(\mathbf a)$ and $H f(\mathbf a)$ was the obvious choice.
+This involves fitting $\mathbf a_i$ based on the conditional distribution of $\mathbf x_{L,i}$ given randomly drawn partial vectors $\mathbf x_{L,-i}$.
+There is no need to simulate the marginal distribution of $\mathbf x$ though, since the simulation is only approximate and we already have a number of realizations in $X$.
+Thus, I only fit the lower half of $A$, using this process:
+        \begin{enumerate}
+            \item To get a simulation of size at least $K$, I create a matrix $X_F$ ($F$ for fixed) which is initially $X$ stacked up until it has $K'\geq K$ rows. 
+            \item For each $p<i\leq 2p$,
+                \begin{itemize}
+                    \item Where $\mathbf x_F^{(k)}$ is the $k$th row of $X_F$, the rows $\mathbf a_i$ are fit sequentially by Newton-Raphson iterations with 
+                    \[f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right] \]
+                    and
+                    \[H f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
+                    \item If the iterations won't converge, I attempt to $f(\mathbf a_i)=m_i^*(\tau)$ for increasing values of $\tau$ until it converges.
+                    \item After $\mathbf a_i$ is fit, a new column $X_i$ is drawn as independent Bernoulli with probability vector $\mu(X_F \mathbf a_i)$.   
+                    \item $X_F$ is updated to 
+                    \[ X_F = [X_F X_i] \]
+                \end{itemize}
+            \item At the end, the first $n$ rows of $X_F$ are taken as $[X \tilde X]$, though the rows corresponding to any copy of $X$ would work equally well.
+        \end{enumerate}
+        Some thoughts/concerns/explanations:
+        \begin{itemize}
+            \item Since $x_{-i}^{(k)}$ is replaced with $x_F^{(k)}$, which is a hybrid of real and simulated data, I am not sure there is the same theoretical guarantee that a unique $A$ matrix exists.
+In practice though it still worked pretty well, and has the advantage that we are deriving $\tilde X$ such that $\E(X'\tilde X) = \hat \Sigma- \diag(\mathbf s)$.
+            \item A hybrid of simulating $x_{-i}^{(k)}$ and keeping fixed $x_F^{(k)}$ would be to draw from the rows of $X$ $K$ times, with replacement, then simulating the rest of the $x_{-i}^{(k)}$ vector.
+I'm not convinced there is a good reason to do this.
+            \item By not simulating, there is the advantage of not needing to fit the upper half of $A$.
+            \item By not redrawing $x_{-i}^{(k)}$ each time, there is less computation for the computer. As well, the multiplication
+                \[ \left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
+            needn't be redone for each iteration, and only partially recalculated for each $i$ (though I don't have this implemented yet).
+The downside might be that error is getting compounded over each $i$.
+        \end{itemize}
+
+\section{Further Work and Simulations}
+As I see it, further simulation work for my paper breaks down into three logical groups: comparison of random Bernoulli knockoffs to the original Knockoffs in LASSO, evaluation of Bernoulli knockoffs in other L1 regularized GLMs, and expanding random knockoffs to more general sorts of variables.
+
+\subsection{Comparison to Original Knockoffs in LASSO}
+The basic idea is to compare how the two sorts of knockoffs compare when $X$ is binary and we are fitting a LASSO regression.
+Do they select the same variables?
+Do the binary knockoffs control FDR more or less conservatively?
+Situations to model:
+\begin{itemize}
+    \item $X$ arises out of the assumed Isling model, with no high order interactions. 
+    \item $X$ does not arise out of this model, and does have higher order interactions.
+Could possibly model this by drawing the vector $\mathbf x$ or subsets of it as multinomial with probabilities from a Dirichlet distribution.
+    \item $X$ is from a real world data set, likely something in genetics.
+    \item $y$ is normally distributed around $X\beta$.
+    \item $y$ is drawn from some other distribution, possibly skewed, heavy tailed, or light tailed.
+    \item $y$ is from a real world data set.
+\end{itemize}
+
+\subsection{Evaluation of Binary Knockoffs in other GLMS}
+The one of the most interest would be logits. It would be good to see how successful the binary knockoffs are in controlling FDR without the theoretical guarantees that LASSO provides.
+\begin{itemize}
+    \item $X$ arises out of the Isling model, in which case it seems like the knockoffs should control FDR.
+    \item $X$ has higher order interactions, which might make the knockoffs perform poorly.
+    \item $X$ is from a real world data set, likely something in genetics.
+    \item $y$ is simulated based on the assumptions of the model.
+    \item $y$ is simulated to violate assumptions of the model.
+    \item $y$ is from a real world data set.
+\end{itemize}
+
+\subsection{Generalizations/Harebrained Ideas}
+I have two ideas for extensions if I have enough time.
+\begin{enumerate}
+    \item Still with binary data, one might be able to simulate binary variables with higher order interactions in the generation of $X$ by including higher order interactions in the regression of $X_i$ on $X_{-i}$.
+This would lead to $A$ being a matrix of higher dimension.
+Even if this worked, extending the method to knockoffs may not be obvious.
+    \item In the general case where $X$ is not binary, I can almost imagine a method set up along similar lines to the binary case.
+    \begin{itemize}
+        \item The desired covariance matrix would arise as before.
+        \item Each $x_i$'s marginal would be approximated by a kernel density estimate on $X_i$.
+        \item A $x_i \st x_1, \ldots, x_{i-1}$ would be drawn from some reweighing of this marginal to achieve the proper covariance.
+For instance, if $F_i^{-1}$ is the inverse CDF of the marginal kernel density for $x_i$ and $u_i\st x_1, \ldots, x_{i-1}$ is a RV on $(0,1)$, then $x_i=F_i^{-1}(u_i)$.
+    \end{itemize}
+\end{enumerate}
 

### 2015-06-20 masters_paper/notes.tex
Notes in relatively good shape

@@
 A random Bernoulli vector $\bf x$ can be summarized by its first two moments: a mean vector $\E(\bf x)=m\in(0,1)^p$ and cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$.
-Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $m=\diag(M)$.
+Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag(M)$.
 For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-mm^T\succeq0$, and
@@
 \subsection{Gaussian Copula Family}
-    Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $z\sim N_p({\bf 0},\Sigma)$ such that,for $x_i = I(z_i<0)$, $x$ has the desired properties.
+    Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $\mathbf z\sim N_p({\bf 0},\Sigma)$ such that,for $x_i = I(z_i<0)$, $x$ has the desired properties.
 There are a number of ways to do this\footnote{``On the Generation of Correlated Artificial Binary Data" - http://epub.wu.ac.at/286/1/document.pdf}\footnote{``On parametric families for sampling binary data with specified mean and correlation"}, but it turns out that there is only certain to exist a working $\Sigma$ in the bivariate case.
@@
 \subsection{$\mu$-Conditionals family}
-    It turns out that there exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as
+    There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as
 
@@
 
-    for some monotone function $\mu:\R\to[0,1]$.
+    for some monotone function $\mu:\R\to(0,1)$.
 This is essentially a binomial family GLM for a link function $\mu$.
@@
 
-    If $\mu$ is chose such that it is a bijection and differentiable, there is a unique $M$ such that $\E(x_ix_i^T)=M$\footnote{``On parametric families for sampling binary data with specified mean and correlation"}.
-It turns out that the natural choice for $\mu$ is the logistic link function, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
+    If $\mu$ is chosen such that it is a bijection and differentiable, there is a unique $A$ such that $\E(\mathbf x \mathbf x^T)=M$ when generated from this model\footnote{``On parametric families for sampling binary data with specified mean and correlation"}.
+The natural choice for $\mu$ is the logistic link function, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
     \[a_{ij} = \log\left(\frac{\p(x_j=x_k=1)\p(x_j=x_k=0)}{\p(x_j=0,x_k=1)\p(x_j=1,x_k=0)}\right) \]
@@
 In this case, one can solve instead for $\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
-    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & \E(x_i) \end{array} \right]^T \]
+    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T \]
             When $\tau=0$, this yields the original problem, while when $\tau=1$, it is treating $x_i$ as independent of $\mathbf x_{-i}$. The latter will always have the solution
-            \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{\E(x_i)}{1-\E(x_i)}\right) \end{array} \right]^T \]
+            \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{M_{ii}}{1-M_{ii}}\right) \end{array} \right]^T \]
             The hope is that for some $\tau$ close to $0$, convergence can be achieved, only causing a slight distortion from the desired cross moments.
@@
     \begin{enumerate}
-        \item I use either the equal correlation method or the SDP method described in the original knockoff method to find $\mathbf s$ such that $\diag(\hat\Sigma)-\mathbf s$ is small and 
+        \item I use either the equal correlation method or the SDP method described in the original knockoff paper to find $\mathbf s$ such that $\|\diag(\hat\Sigma)-\mathbf s\|$ is small and 
         \[ \Sigma_L = \left[\begin{array}{cc}  \hat\Sigma & \hat\Sigma - \diag(\mathbf s )\\ \hat\Sigma - \diag(\mathbf s) & \hat\Sigma \end{array}\right] \succeq 0 \]
@@
                     \[H f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
-                    \item If the iterations won't converge, I attempt to $f(\mathbf a_i)=m_i^*(\tau)$ for increasing values of $\tau$ until it converges.
+                    \item If the iterations won't converge, I attempt to fit $f(\mathbf a_i)=m_i^*(\tau)$ instead for increasing values of $\tau$ until it converges.
                     \item After $\mathbf a_i$ is fit, a new column $X_i$ is drawn as independent Bernoulli with probability vector $\mu(X_F \mathbf a_i)$.   
                     \item $X_F$ is updated to 
-                    \[ X_F = [X_F X_i] \]
+                    \[ X_F = [X_F \; X_i] \]
                 \end{itemize}
-            \item At the end, the first $n$ rows of $X_F$ are taken as $[X \tilde X]$, though the rows corresponding to any copy of $X$ would work equally well.
+            \item At the end, the first $n$ rows of $X_F$ are taken as $[X \; \tilde X]$, though the rows corresponding to any copy of $X$ would work equally well, since these all should have the desired distribution.
         \end{enumerate}
@@
 Do the binary knockoffs control FDR more or less conservatively?
-Situations to model:
+Situations to test:
 \begin{itemize}
@@
     \item $X$ is from a real world data set, likely something in genetics.
-    \item $y$ is simulated based on the assumptions of the model.
-    \item $y$ is simulated to violate assumptions of the model.
+    \item $y$ is simulated based on the assumptions of the regression model.
+    \item $y$ is simulated to violate assumptions of the regression model.
     \item $y$ is from a real world data set.
@@
 \subsection{Generalizations/Harebrained Ideas}
-I have two ideas for extensions if I have enough time.
+I have two seeds of ideas for extensions if I have enough time.
 \begin{enumerate}
@@
     \begin{itemize}
-        \item The desired covariance matrix would arise as before.
-        \item Each $x_i$'s marginal would be approximated by a kernel density estimate on $X_i$.
+        \item The desired covariance matrix could be chosen as in the original knockoff paper.
+        \item Each $x_i$'s marginal distribution would be approximated by a kernel density estimate on $X_i$.
         \item A $x_i \st x_1, \ldots, x_{i-1}$ would be drawn from some reweighing of this marginal to achieve the proper covariance.
 For instance, if $F_i^{-1}$ is the inverse CDF of the marginal kernel density for $x_i$ and $u_i\st x_1, \ldots, x_{i-1}$ is a RV on $(0,1)$, then $x_i=F_i^{-1}(u_i)$.
+        \item Maybe fit generalized additive model for each successive $x_i$ with kernel regression for $x_1,\ldots,x_{i-1}$ to predict mean. Then, skew marginal of $x_i$ until it has that mean. 
     \end{itemize}

### 2015-07-10 masters_paper/notes.tex
Added FDR calculations

@@
 \documentclass[11pt]{article}
-\usepackage[paper=letterpaper, margin=.5in]{geometry}
+\usepackage[paper=letterpaper, margin=1in]{geometry}
 \linespread{1.5}
@@
 \begin{document}
-\title{Binary Knockoffs Notes}
+\title{Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate}
 \author{Aaron Maurer}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Preliminaries} 
+\section{Introduction}
+    Basic overview of what paper is about
+
+\subsection{Knockoff Filter}
+    Describe the original knockoff filter,
+
+\subsection{Binary Knockoffs}
+    Overview of my addition
+
+\section{Issues With Deterministic Knockoffs} 
     {\em Note:} I will try to hold to the convention that $X$ is the $n\times p$ data matrix, while $\bf x$ is the random vector variable from which each row of $X$ was drawn.
-Accordingly, $\tilde X$ will be the knockoff matrix while $\bf \tilde x$ is a random vector variable.
-$x_i$ will be the random variable corresponding to the $i$th entry of $\bf x$, while $X_i$ is the vector of observations drawn from $x_i$ in the data matrix.
+Accordingly, $\tilde X$ will be the knockoff matrix while $\bf \tilde x$ is a random vector variable, at least when $\tilde X$ is generated randomly.
+$x_i$ will be the random variable corresponding to the $i$th entry of $\bf x$, while $X_i$ is $i$th column of $X$, with observations drawn from $x_i$.
 $\hat \Sigma = \frac{1}{n}X^TX$ is the empirical covariance matrix associated with $X$, while $\Sigma=\E(\mathbf x \mathbf x^T)$ is the theoretical covariance matrix associate with $\mathbf x$.
-Also, for a square matrix $A$, $\diag(A)$ if the vector of values along the diagonal, while for a vector $\mathbf a$, $\diag(\mathbf a)$ is a square diagonal matrix with $\mathbf a$ along the diagonal.\\
+Also, for a square matrix $A$, $\diag(A)$ is the vector of values along the diagonal, while for a vector $\mathbf a$, $\diag(\mathbf a)$ is a square diagonal matrix with $\mathbf a$ along the diagonal.\\
 
-    Some early investigation into deterministic knockoffs (as described in the original knockoff paper) reveal that they don't perform well in L1 regularized logistic regression.
+    Simulations with deterministic knockoffs reveal that they don't perform well in L1 regularized logistic regression.
 Even when $X_i$ is a null predictor of $y$, the $X_i$ still tend to enter the model prior to $\tilde X_i$.
-The issue is that even when $x\sim N_p({\bf 0}, \Sigma)$ for some $\Sigma\succeq 0$, $\tilde x$ is not normally distributed.
+The issue is that even when $x\sim N_p({\bf 0}, \Sigma)$ for some $\Sigma\succeq 0$, the $\tilde X_i$ do not fit a normal distribution.
+This can be seen below in figure 1, where normal Q-Q plots of simulated variables $X_i$ and corresponding knockoffs $\tilde X_i$ are provided.
+    \begin{figure}[h!]
+        \includegraphics[width=20cm]{images/normalQQ}
+        \caption{Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
+    \end{figure}
+
     This can be seen from producing qq plots of $X_i$ vs $\tilde X_i$ for each $i$.

### 2015-07-10 masters_paper/knockoffGLM.py

@@ -8,6 +8,7 @@ The method for binary models is my development, based on the paper "On parametri
 """
 
 # imports 
+from __future__ import division
 import numpy as np
 import numpy.linalg as nplin
 import numpy.random as npran
@@ -67,10 +68,11 @@ def solve_sdp(G):
     return np.asarray(s.value).flatten()
 
 class knockoff_logit(object):
-    def __init__(self,y,X,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,tol=1E-5,maxiter=100,full_A=False):
+    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,tol=1E-5,maxiter=100,full_A=False):
         self.y      = y
         self.X      = normalize(X.astype(float),norm='l2',axis=0)
         self.X_orig = X
+        self.q      = q                # Level we want to control FDR at
         self.n,self.p = self.X.shape
         self.knockoff_type = knockoff  # what type of knockoff are we generating - binary or the original deterministic ones?
         self.cov_type = cov_method     # how are we generating the desired covariance matrix - equicor or SDP?
@@ -301,12 +303,23 @@ class knockoff_logit(object):
         self.w_rank   = self.w_filter * np.nanmin((self.z_rank[self.p:(2*self.p)],self.z_rank[:self.p]),axis=0)
         self.w_value  = self.w_filter * np.max((self.z_value[self.p:(2*self.p)],self.z_value[:self.p]),axis=0)
 
+    def _get_T(self):
+        """ Calculates the data-dependent threshold for the statistics """
+        Ws = set(np.abs(self.w_value))
+        self.T = np.amin(np.append([t for t in Ws if np.sum(self.w_value<=-t)/max(1,np.sum(self.w_value>=t)) <= self.q],np.inf))
+
+    def _get_S(self):
+        """ Calculate the vector selecting features S """
+        self.S = self.w_value>=self.T
+
     def fit(self):
         if   self.knockoff_type == 'deterministic': self._deterministic_knockoff()
         elif self.knockoff_type == 'binary':        self._binary_knockoff()
         self._fit_lognet()
         self._get_z()
         self._get_w()
+        self._get_T()
+        self._get_S()
 
 def main(n,p,q):
     print 'nobody lives here right now'

### 2015-07-10 masters_paper/knockoffGLM.py
Added knockoff lasso via breaking class into parent and child classes

@@ -21,7 +21,7 @@ import rpy2 as rp
 import rpy2.robjects.numpy2ri
 import statsmodels.api as sm 
 from rpy2.robjects.packages import importr
-from glmnet import LogisticNet
+from glmnet import LogisticNet, ElasticNet
 from sklearn.preprocessing import normalize
 from scipy.special import expit as invlogit
 from scipy.special import logit
@@ -67,7 +67,8 @@ def solve_sdp(G):
     # Return as array, not as matrix.
     return np.asarray(s.value).flatten()
 
-class knockoff_logit(object):
+class knockoff_net(object):
+    """ Parent class for knockoff lasso and knockoff logistic regression. Defines everything besides fitting the particular GlmNet object """
     def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,tol=1E-5,maxiter=100,full_A=False):
         self.y      = y
         self.X      = normalize(X.astype(float),norm='l2',axis=0)
@@ -130,14 +131,15 @@ class knockoff_logit(object):
         s_diff[s_diff<0]=0 # can be negative due to numerical error
         self.X_ko = np.dot(U*(d-s/d) + U_perp*(np.sqrt(s_diff)) , V)
 
-    def _deterministic_knockoff(self):
+    def _original_knockoff(self):
+        """ Creates the original style knockoffs"""
         if self.cov_type == 'equicor':
             self._create_equicor()
         else: 
             self._create_SDP()
 
     def _binary_knockoff(self):
-        ''' This creates knockoffs which are random multivariate bernoulli which should have, in expectation,
+        ''' This creates the new binary knockoffs, which are random multivariate bernoulli which should have, in expectation,
         the same first two moments as X. Only will work if X is all binaray '''
 
         ###############################################
@@ -277,21 +279,6 @@ class knockoff_logit(object):
         # hang onto A
         self.A = A
 
-    def _fit_lognet(self):
-        X_lrg = np.concatenate((self.X,self.X_ko), axis=1)
-
-        # initialize the glmnet object
-        self.lognet = LogisticNet(alpha=1) 
-        self.lognet.fit(X_lrg,self.y,normalize=False,include_intercept=False)
-
-        self.lambdas = self.lognet.out_lambdas
-        self.var_index_ent = np.sort(self.lognet._indices)
-        self.coef_matrix = np.zeros((2*self.p,self.lognet.n_lambdas))
[…]

### 2015-07-10 masters_paper/paper.tex
Adding the paper

--- 2015-07-10 masters_paper/notes.tex
@@
 \documentclass[11pt]{article}
-\usepackage[paper=letterpaper, margin=1in]{geometry}
+\usepackage[paper=letterpaper, margin=.5in]{geometry}
 \linespread{1.5}
@@
 \subsection{Knockoff Filter}
-    Describe the original knockoff filter,
+    Describe the original knockoff filter, and introduce some notation. Original=deterministic
 
@@
 This can be seen below in figure 1, where normal Q-Q plots of simulated variables $X_i$ and corresponding knockoffs $\tilde X_i$ are provided.
+Of course, when $X$ is a binary vector, $\tilde X$ completely doesn't match its distribution.
     \begin{figure}[h!]
@@
     \end{figure}
-
-    This can be seen from producing qq plots of $X_i$ vs $\tilde X_i$ for each $i$.
-Of course, when $X$ is a binary vector, $\tilde X$ completely doesn't match its distribution, causing the original $X$ to beat the knockoffs into the model.
-This indicates that a new method of generating $\tilde X$ must be created to control FDR via knockoffs with regularized logistic regression.
+    These
 
 \section{Random Bernoulli Knockoffs}
-    My idea is to generate $\bf \tilde x$ randomly such that, approximately, $\bf \tilde x\sim \bf x$.
-In particular, both variables should have similar marginal densities, expectations, and second moments.
+    My solution to the issues with the
+
+     generate $\bf \tilde x$ randomly such that it has the desired knockoff properties in expectation.
+In particular, both  should have similar marginal densities, expectations, and second moments.
 However, $\bf \tilde x\st \bf x$ should also have desired knockoff property that $\E(\mathbf{\tilde x}^T\mathbf x\st \mathbf x) = \Sigma-\mathbf s$, where $\|\diag(\Sigma)-s\|$ is small.
@@
 
-\section{Random Bernoulli Generation}
+\subsection{Random Bernoulli Generation}
     Thankfully, there has been a reasonable amount of work on how one can generate random Bernoulli vectors with some kind of correlation among among the values.
@@
 
-\subsection{Gaussian Copula Family}
+\subsubsection{Gaussian Copula Family}
     Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $\mathbf z\sim N_p({\bf 0},\Sigma)$ such that,for $x_i = I(z_i<0)$, $x$ has the desired properties.
@@
 
-\subsection{$\mu$-Conditionals family}
+\subsubsection{$\mu$-Conditionals family}
     There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as
@@
 
-\section{Generating Knockoffs}
+\subsection{Generating Binary Knockoffs}
     My method for generating binary knockoffs broadly involves two steps:
@@
 
+\section{Bernoulli Knockoff Performance}
+For each of thses, exaimine these situatations:
+\begin{itemize}
+    \item $X$ arises out of the assumed Isling model, with no high order interactions. 
+    \item $X$ does not arise out of this model, and does have higher order interactions.
+Could possibly model this by drawing the vector $\mathbf x$ or subsets of it as multinomial with probabilities from a Dirichlet distribution.
+    \item $X$ is from a real world data set, likely something in genetics.
+    \item $y$ is normally distributed around $X\beta$.
+    \item $y$ is drawn from some other distribution, possibly skewed, heavy tailed, or light tailed.
+    \item $y$ is from a real world data set.
+\end{itemize}
+\subsection{Gram Matrix Comparison}
+\subsection{Binary vs. Determinist Knockoffs in LASSO}
+\subsection{Binary Knockoffs in other Regularlized GLMS}
+
+\section{Discussion}
+\subsection{Areas for Further Work}
+
 \section{Further Work and Simulations}

### 2015-07-13 masters_paper/paper.tex
Rough version section 1

@@
 \newcommand{\cov}{\mathrm{Cov}}
-\newcommand{\diag}{\mathrm{diag}}
+\newcommand{\diag}[1]{\mathrm{diag}\{#1\}}
 \newcommand{\ST}{mbox{ s.t. }}
@@
 \section{Introduction}
-    Basic overview of what paper is about
+    Variable selection is an essential problem to fitting a regression model.
+For linear regression, the knockoff filter (Barber and Cand\`{e}s, 2014) offers a method of exact FDR control.
+However, the method as originally formulated does not extend to other generalized linear models such as logistic regression.
+This paper offers an extension of this method which will work with binary features in the context of linear regression, but also will extend to other GLMs.
 
 \subsection{Knockoff Filter}
-    Describe the original knockoff filter, and introduce some notation. Original=deterministic
+    Let us consider the consider the usual setting for linear regression, where $n$ observations of a variable of interest $y$ arise from the model
+    \[\mathbf{y} = X\beta + \mathbf{z}\]
+    where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise.
+For the purpose of this paper, only the case where $2p\leq n$ will be considered, though the methods can be extended to $p\leq n$.
+I will to refer the $j$th column of $X$ as $X_j$, which is $n$ observations of the random variable $x_j$.
+The random variable $x_j$ is in turn $j$th entry of the random vector valued variable $\mathbf x$.
+In the situation where $\beta$ is sparse and there is reason to expect that $x_j$ and $y$ are uncorrelated for many $j$, the knockoff filter is a method to select a subset of the $x_j$ which are likely correlated with $y$.
+\par
+    The method specifically seeks to control the false discovery rate (FDR), which is the expected portion of some selection of $x_j$ which are, in fact, uncorrelated with $y$.
+For a procedure which chooses a subset of variables $\hat S\subseteq \{1,\ldots,p\}$, this is defined as
+    \[ \textrm{FDR} = \E\left[\frac{\vert{\{j: \beta_j=0 \; \& \; j\in\hat S\}}\vert}{\max\{\vert{\hat S}\vert,1\}} \right] \]
+    FDR is controlled at level $q$ if FDR is less than $q$ irrespective of the coefficients $\beta$. \par
+
+    The knockoff filter achieves FDR control in several steps, the first of which is creating a set of `knockoff' features $\tilde X_j$ which imitate the original $X_j$ while being less correlated with $y$.
+In particular, the matrix of knockoffs $\tilde X$ has the same internal correlation structure as $X$, the knockoff feature $\tilde X_j$ has the same correlation with other $X_i$ as $X_j$ does, but the correlation between $\tilde X_j$ and $X_j$ is minimized.
+In other words, where $\hat \Sigma := X^TX$,
+    \[ \tilde X^T\tilde X=\hat\Sigma \textrm{ and } \tilde X^T X = \Sigma - \diag{\mathbf s}\]
+    for some vector $s\in \R^p$ such that, where $\diag{\Sigma}$ is the vector made from the main diagonal of $\Sigma$, $\diag{\Sigma} - \mathbf s$ is small.
+Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$.
+In the case where $X_j$ is uncorrelated with $\mathbf y$, $\tilde X_j$ will also be uncorrelated with $\mathbf y$.
+How these knockoffs are actually generated will be described shortly.\par
+    The second step is to fit a series of Lasso model of $\mathbf y$ on the combined design matrix $[X \; \tilde X]$ so as to determine the largest value $\lambda$ at which the coefficient for each $X_j$ and $\tilde X_j$ is nonzero.\footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} Recall, for a given $\lambda$, the estimated coefficient from the Lasso regression will be
+    \[ \beta(\lambda) = \textrm{argmin}_\mathbf b \left\{\frac{1}{2}\|\mathbf y - [X\; \tilde X]\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]
+    In the general case, the coefficient for one feature in a Lasso regression being first nonzero for a larger value of $\lambda$ than another feature is an indication that the first feature is a stronger predictor of the outcome than the second.
+Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to have nonzero coefficients sooner than the knockoff features when the original is a valid predictor.
+On the other hand, since $\beta(\lambda)$ only depends on $[X\; \tilde X]$ through the sufficient statistics $[X\; \tilde X]^T[X\; \tilde X]$ and $[X\; \tilde X]^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
+This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $[X\; \tilde X]^T[X\; \tilde X]$ will be unaltered, while $\E\left[[X\; \tilde X]^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$.
+\par
+   This observation leads to the final step.
+Let $Z_j$ be the largest $\lambda$ such that feature $j$ has a nonzero coefficient and $\tilde Z_j$ the largest $\lambda$ such that knockoff $j$ has a nonzero coefficient.
+Then, define $W_j$ as
+   \[ W_j = \begin{cases}   Z_j         &\mbox{if } Z_j>\tilde Z_j \\ 
+                            -\tilde Z_j &\mbox{if } Z_j<\tilde Z_j \\
+                            0           &\mbox{if } Z_j=\tilde Z_j \\
+            \end{cases} \]
+            When $X_j$ is a null predictor, $W_j$ will be symmetrically distributed around zero, and when $X_j$ is true predictor, it will have a distribution skewed positive.
+Since $W_j$ with large positive values are more likely to be true predictors, the knockoff filter selects the variables $\{j:W_j\geq T\}$, where $T$ is a threshold defined as
+            \[ T = \min\left\{ t>0 \;: \; \frac{\vert\{j:W_j\leq -t\}\vert}{\max\{\vert\{j:W_j\geq t\}\vert,1\}}\leq q \right\} \]
+            Basic logic of this method is fairly simple.
+Let $N_h(t)$ be the random variable for the number of null predictors with $W_j\leq t$, $N_l(t)$ be the number of null predictors with $W_j\geq t$, and $V_l(t)$ \& $V_h(t)$ be the similar number of true predictors.
+Since the null $W_j$ are symmetrically distributed around $0$, $\E[N_l(t)]=\E[N_h(t)]$.
+Thus,
+            \begin{align*}
+                q &\geq \E\left[ \frac{\vert\{j:W_j\leq -t\}\vert}{\max\{\vert\{j:W_j\geq t\}\vert,1\}} \right] \\
+                q &\geq \E\left[ \frac{N_l(t) + V_l(t)}{\max\{N_h(t) + V_h(t),1\}} \right] \\
+                q &\geq \E\left[ \frac{N_l(t)}{\max\{N_h(t) + V_h(t),1\}} \right] \\
+                q &\geq \E\left[ \frac{N_h(t)}{\max\{N_h(t) + V_h(t),1\}} \right] \\
+                q &\geq \textrm{FDR}
+            \end{align*}
+            So FDR is controlled, and since $V_l(t)$ will tend to be small, it should be controlled fairly tightly.
+
+\subsection{Original Knockoff Features}
+    The original formulation of the method offers two similar methods of constructing knockoffs. Both of these will, by construction, have exactly the property that
+    \[ \tilde X^T\tilde X=\hat\Sigma \textrm{ and } \tilde X^T X = \Sigma - \diag{\mathbf s}.\]
+    For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. Then, let the Gram matrix of $[X\; \tilde X]$ be
+    \[ G:= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} \Sigma & \Sigma - \diag{\mathbf s} \\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \]
+    $A$, the Schur complement of $\Sigma$ in $G$ can be calculated as 
+    \[ A = 2\,\diag{\mathbf s} - \diag{\mathbf s}\Sigma^{-1}\diag{\mathbf s} \]
+    For $G$ to exist, $G$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if\footnote{A fleshed out version of this argument can be found in the original knockoff paper}
+    \[ \diag{\mathbf s} \succeq 0  \; \textrm{ and } \; 2\Sigma - \diag{\mathbf s} \succeq 0 \]
+    Given this is true, $A$ can be factored as $A=C^TC$.
+Combining this with a satisfactory $\mathbf s$ and an orthonormal matrix $\tilde U\in\R^{n\times p}$ such that $\tilde U^T X = 0$, a $\tilde X$ fulfilling the desired properties can be calculated as
+    \[ \tilde X = X(I-\Sigma^{-1}\,\diag{\mathbf s}) + \tilde UC\]
+    As mentioned above, $\mathbf s$ should be chosen so as to make $\diag{\Sigma}-\mathbf s$ small.
+Since each $X_j$ has been normalized, this means that $\diag{\Sigma}=\mathbf 1$, so $\mathbf 1 - \mathbf s$ should be minimized in accordance with the restrictions on $\mathbf s$.
+The two methods differ in how $\mathbf s$ is chosen:
+    \begin{itemize}
+        \item \textit{Equi-correlated knockoffs}: Each original features is set to have the same correlation with its knockoff by setting $\mathbf s = 2\min\{\lambda_{min}(\Sigma),1\}\mathbf 1$.
+For all simulations in this paper, this was the method used.
+        \item \textit{SDP knockoffs}: The $\mathbf s$ which minimizes the average correlation between knockoff and original features can be found via a semi-definite programing problem:
+            \begin{center}
+                \begin{tabular}{r l}
+                    minimize & $\|\mathbf1-\mathbf s\|_1$ \\
+                    subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2\Sigma $
+                \end{tabular} 
+            \end{center}
+            This method is significantly more computationally intensive. 
+    \end{itemize}
 
 \subsection{Binary Knockoffs}
-    Overview of my addition
+    Though the ``original" knockoff method, described above, achieves the exact desired knockoff properties, the individual values in the vector $X_j$ will have little relation to the individual values in $X_j$.
+In particular, these values will often have very different empirical distributions.
+This effect is particularly noticeable when the random variable $x_j$ is discrete but $\tilde X_j$ does not even consist of the same set of values.
+The end result is that for variable selection for other generalized linear models, which do not have the same sufficient statistics as linear regression which knockoffs are designed to hit, the original knockoffs will fail to control FDR.
+\par
+    Thus, this paper offers a new method of generating knockoffs which should offer superior performance with other GLMs.
+In this new method, the matrix of knockoff features $\tilde X$ will be generated by row from a random vector valued variable $\mathbf{\tilde x}$, each entry $\tilde x_j$ of which will have the property $\tilde x_j \sim x_j$.
+This random variable will also abide by a relaxation of the original knockoff condition:
+    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=\hat\Sigma \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = \Sigma - \diag{\mathbf s}\]
+    This will ensure that, at the very least, these knockoffs are a suitable replacement for the original knockoffs for linear regression. \par
+    Generating such knockoffs for a completely general case of any $\mathbf x$ is likely infeasible.
+However, in the specific case where $\mathbf x$ is a random binary vector variable, this paper will demonstrate a method to generate such knockoff variables.
+Since binary data is extremely common in many areas, this will hopefully serve as a useful tool.
+
+\subsection{Paper Organization}
+The rest of this paper will be organized in the following fashion:
+\begin{itemize}
+    \item Section 2 will go into more detail about how the original knockoffs breakdown with generalized linear models. In particular, several simulations will demonstrate how they breakdown.
+    \item Section 3 will develop the method and theory for generating binary knockoffs.
+    \item Section 4 will discuss tests of Binary knockoffs in simulation, both in comparison to the original knockoffs with Lasso and for logistic regression by itself.
+    \item Section 5 is the final section, and will contain discussion of the results as well as areas for further work.
+\end{itemize}
 
@@
 $\hat \Sigma = \frac{1}{n}X^TX$ is the empirical covariance matrix associated with $X$, while $\Sigma=\E(\mathbf x \mathbf x^T)$ is the theoretical covariance matrix associate with $\mathbf x$.
-Also, for a square matrix $A$, $\diag(A)$ is the vector of values along the diagonal, while for a vector $\mathbf a$, $\diag(\mathbf a)$ is a square diagonal matrix with $\mathbf a$ along the diagonal.\\
+Also, for a square matrix $A$, $\diag A$ is the vector of values along the diagonal, while for a vector $\mathbf a$, $\diag \mathbf a$ is a square diagonal matrix with $\mathbf a$ along the diagonal.\\
 
@@
 In particular, both  should have similar marginal densities, expectations, and second moments.
-However, $\bf \tilde x\st \bf x$ should also have desired knockoff property that $\E(\mathbf{\tilde x}^T\mathbf x\st \mathbf x) = \Sigma-\mathbf s$, where $\|\diag(\Sigma)-s\|$ is small.
+However, $\bf \tilde x\st \bf x$ should also have desired knockoff property that $\E(\mathbf{\tilde x}^T\mathbf x\st \mathbf x) = \Sigma-\mathbf s$, where $\|\diag{\Sigma}-s\|$ is small.
 In the general case, this is likely infeasible; however, if $\bf x$ is a binary vector, as is often the case, we know we are dealing with a much more limited class of random variables, and it should be possible to randomly generate $\bf \tilde x\st \bf x$ so as to have the desired properties.
@@
 A random Bernoulli vector $\bf x$ can be summarized by its first two moments: a mean vector $\E(\bf x)=m\in(0,1)^p$ and cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$.
-Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag(M)$.
+Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag{M}$.
 For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-mm^T\succeq0$, and
@@
     \begin{enumerate}
-        \item I use either the equal correlation method or the SDP method described in the original knockoff paper to find $\mathbf s$ such that $\|\diag(\hat\Sigma)-\mathbf s\|$ is small and 
-        \[ \Sigma_L = \left[\begin{array}{cc}  \hat\Sigma & \hat\Sigma - \diag(\mathbf s )\\ \hat\Sigma - \diag(\mathbf s) & \hat\Sigma \end{array}\right] \succeq 0 \]
+        \item I use either the equal correlation method or the SDP method described in the original knockoff paper to find $\mathbf s$ such that $\|\diag{\hat\Sigma}-\mathbf s\|$ is small and 
+        \[ \Sigma_L = \left[\begin{array}{cc}  \hat\Sigma & \hat\Sigma - \diag{\mathbf s }\\ \hat\Sigma - \diag{\mathbf s} & \hat\Sigma \end{array}\right] \succeq 0 \]
         I use a subscript $L$ for large to denote items associated with the joint distribution of $[\mathbf x \; \mathbf{\tilde x}]$.
@@
             \item Since $x_{-i}^{(k)}$ is replaced with $x_F^{(k)}$, which is a hybrid of real and simulated data, I am not sure there is the same theoretical guarantee that a unique $A$ matrix exists.
-In practice though it still worked pretty well, and has the advantage that we are deriving $\tilde X$ such that $\E(X'\tilde X) = \hat \Sigma- \diag(\mathbf s)$.
+In practice though it still worked pretty well, and has the advantage that we are deriving $\tilde X$ such that $\E(X'\tilde X) = \hat \Sigma- \diag{\mathbf s}$.
             \item A hybrid of simulating $x_{-i}^{(k)}$ and keeping fixed $x_F^{(k)}$ would be to draw from the rows of $X$ $K$ times, with replacement, then simulating the rest of the $x_{-i}^{(k)}$ vector.

### 2015-07-13 masters_paper/simulate.py
Added images

@@ -164,7 +164,7 @@ class ko_test(object):
             # The cumulitive rate of originals coming in first up to a given rank
             self.true_orig_rank_rate = self.true_or_rank_ct/(self.true_or_rank_ct+self.true_ko_rank_ct)
 
-    def plot_ko_rank_rate(self):
+    def plot_ko_rank_rate(self,title='Percentage Original Variables'):
         self.knockoff_rank_rate()
         if self.p1>0:
             plt.plot(np.arange(self.p)+1,self.true_orig_rank_rate,'b-', label = 'True Predictors')
@@ -173,7 +173,7 @@ class ko_test(object):
             end_val = self.null_orig_rank_rate[-1]
         plt.xlabel('Rank Variable Entered Model')
         plt.ylabel('Cumulative % Original Variables')
-        plt.title('Percentage Original Variables\n %.3f of Null Originals Came in First By End' % end_val)
+        plt.title(title+'\n %.3f of Null Originals Came in First By End' % end_val)
         plt.axis([1,self.p,0,1])
         plt.legend(loc='lower right')
         plt.show()

### 2015-07-14 masters_paper/paper.tex
Draft of section 2

@@
 % placeins defines "\FloatBarrier", which stops tables from moving around
-\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
+\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins, caption} 
+\captionsetup{font={small,singlespacing},margin=1cm}
 \newcommand{\Z}{\mathbb{Z}}
@@
 This random variable will also abide by a relaxation of the original knockoff condition:
-    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=\hat\Sigma \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = \Sigma - \diag{\mathbf s}\]
+    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=\Sigma \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = \Sigma - \diag{\mathbf s}\]
     This will ensure that, at the very least, these knockoffs are a suitable replacement for the original knockoffs for linear regression. \par
@@
 \section{Issues With Deterministic Knockoffs} 
-    {\em Note:} I will try to hold to the convention that $X$ is the $n\times p$ data matrix, while $\bf x$ is the random vector variable from which each row of $X$ was drawn.
-Accordingly, $\tilde X$ will be the knockoff matrix while $\bf \tilde x$ is a random vector variable, at least when $\tilde X$ is generated randomly.
-$x_i$ will be the random variable corresponding to the $i$th entry of $\bf x$, while $X_i$ is $i$th column of $X$, with observations drawn from $x_i$.
-$\hat \Sigma = \frac{1}{n}X^TX$ is the empirical covariance matrix associated with $X$, while $\Sigma=\E(\mathbf x \mathbf x^T)$ is the theoretical covariance matrix associate with $\mathbf x$.
-Also, for a square matrix $A$, $\diag A$ is the vector of values along the diagonal, while for a vector $\mathbf a$, $\diag \mathbf a$ is a square diagonal matrix with $\mathbf a$ along the diagonal.\\
+    The knockoff filter and $W$ statistics have a very natural extension to generalized linear models.
+Here, where $l(\beta\st X, \tilde X)$ is the log likelihood of coefficients $\beta$ given the model, the $L1$ regularized regression model will have estimated coefficient vector
+    \[ \beta(\lambda) = \textrm{argmin}_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
+    The $Z_j$ and $\tilde Z_j$ are then once again the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before.
+The problem with this is that if the given GLM doesn't have the same sufficient statistics as linear regression, there is no theoretical guarantee that the knockoff filter will control FDR.
+This is since, for null $X_j$, there is no theory to guarantee the $Z_j$ and $\tilde Z_j$ are exchangeable, and thus $W_j$ is not guaranteed to be symmetrically distributed around $0$.
+\par
+    \begin{figure}[h]
+        \begin{center}
+        \includegraphics[width=15cm]{images/entryrate_original_logit}
+    \end{center}
+        \caption{The mean cumulative portion of $W$ statistic which are positive by absolute rank of the value.
+The mean is over 1000 simulations.
+On the left side, the features are drawn from a random standard normal distribution.
+On the right side, they are drawn as Bernoulli with probability $.25$ of being 1.
+The top has 50 features uncorrelated with the outcome, while the bottom has 40 features uncorrelated with the outcome and 10 correlated with it.}
+    \end{figure}
 
-    Simulations with deterministic knockoffs reveal that they don't perform well in L1 regularized logistic regression.
-Even when $X_i$ is a null predictor of $y$, the $X_i$ still tend to enter the model prior to $\tilde X_i$.
-The issue is that even when $x\sim N_p({\bf 0}, \Sigma)$ for some $\Sigma\succeq 0$, the $\tilde X_i$ do not fit a normal distribution.
-This can be seen below in figure 1, where normal Q-Q plots of simulated variables $X_i$ and corresponding knockoffs $\tilde X_i$ are provided.
-Of course, when $X$ is a binary vector, $\tilde X$ completely doesn't match its distribution.
-    \begin{figure}[h!]
-        \includegraphics[width=20cm]{images/normalQQ}
+    \FloatBarrier
+    To test how bad this problem is, I simulated the method with the original knockoffs with logistic regression.
+The results can be seen in figure 1, which shows cumulatively, starting with the largest absolute value, what portion of the $W_j$ are positive.
+For the method to work, for the null predictors, this should be a consistent $50\%$, indicative of the desired symmetric distribution.
+This is particularly important for high values, which play the largest roll in selecting a threshold.
+We see that in the top left corner, when all features are null and normally distributed, that we almost achieve this, with only a small, but consistent, deviation from $50\%$.
+However, in the top right, when the predictors are binary, the null originals enter first far more often, which will the FDR too high.
+\par
+    This gets much worse when some true predictors are included, in the bottom row of the figure.
+For both normal and binary predictors, the originals come in first far more often in the beginning.
+This will completely ruin FDR control, since the portion of null predictors will be assumed far lower than it actually is.
+It is worth noting that the particular shape of the curve in the diagram, with the plateau for the null predictors through 10, is partially an artifact of how the totals were tabulated.
+The mean through the first ten is the mean for null predictors that came in first out of both null and true predictors; since the largest ten $W_j$ will tend to be true predictors, the average over the null predictors represent the few extreme $W_j$ which managed to beat out some number of true predictors.
+\par
+    \begin{figure}[h]
+        \includegraphics[width=15cm]{images/normalQQ}
         \caption{Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
     \end{figure}
-    These
+
+    Upon examination, the issue seems to be that the original style knockoff features have a very different empirical distribution than the features they are imitating.
+This is obviously true with the binary features, but even with the normal features, the distributions differ a lot.
+In figure 2, the distribution of four feature and the original style knockoff are plotted against a normal distribution in Q-Q plots.
+As you can see, they don't match at all.
+This is suggestive of a solution; if the knockoff variables match the marginal distribution of the original features better, the result may be more suitable for FDR control in GLMs.
+\par
+    \FloatBarrier
 

### 2015-07-14 masters_paper/knockoffGLM.py

@@ -168,23 +168,46 @@ class knockoff_net(object):
             lambda_min = min(d)**2
             s = np.ones(self.p)*min(2*lambda_min,1)
 
-        # Generate the large covariance matrix we want
+        # Return to the original X
         self.X = self.X_orig              # Now that we've adjusted the variance, revert back to the original X 
-        Sigma     = np.cov(self.X,rowvar=0)
-        s         = s*np.diag(Sigma)        # scale s back up to original variance
-        Sigma_lrg = np.hstack(((np.vstack((Sigma,Sigma-np.diag(s)))),np.vstack((Sigma-np.diag(s),Sigma))))
 
-        # Generate the crossmoment matrix 
+        # Calculate first moments 
         mu       = np.mean(self.X,axis=0)
         mu_lrg   = np.concatenate((mu,mu))
-        self.M   = Sigma_lrg + np.outer(mu_lrg,mu_lrg)
 
-        # Check condition 2.1 from Schafer - make sure its a valid crossmoment matrix
-        cond1 = self.M <= np.minimum(np.outer(mu_lrg,np.ones(2*self.p)),np.outer(np.ones(2*self.p),mu_lrg))
-        cond2 = np.outer(mu_lrg,np.ones(2*self.p))+np.outer(np.ones(2*self.p),mu_lrg) -1 <= self.M
-        test = np.logical_or(np.logical_and(cond1,cond2),np.diag(np.ones(2*self.p)))  # ignore diagonal
-        if np.min(test) == False:
-            print "The cross momement matrix violates necessary conditions for being random Bernoulli"
+        # Scale s up to original variance
+
+        # Generate the covariance matrix Sigma
+        M    = np.dot(self.X.T,self.X)/self.n
+        Sigma = M - np.outer(mu,mu)
+        s     = s*np.diag(Sigma)        # scale s back up to original variance
+
+
+        # Generate the large Sigma matrix, and keep shrinking s until its a valid crossmoment matrix
+        test = False 
+        shrink = 1 
+        while not test:
+            # s will be shrunk a little more each time the test is failed
+            s = s*shrink
+
+            # Generate large Sigma and crossmoment matrix
+            Sigma_lrg = np.hstack(((np.vstack((Sigma,Sigma-np.diag(s)))),np.vstack((Sigma-np.diag(s),Sigma))))
+            self.M   = Sigma_lrg + np.outer(mu_lrg,mu_lrg)
+
+            # Check condition 2.1 from Schafer - make sure its a valid crossmoment matrix
[…]

### 2015-07-14 masters_paper/simulate.py

@@ -43,6 +43,79 @@ def cutoff(array,min=-50,max=50):
     cut_array = np.max((min*np.ones(shape),cut_array),axis=0)
     return cut_array
 
+def genXy_given_X_norm_y(seed,data,n,p1,pnull,sd=1,beta_sd=1):
+    ''' X is binary from the isling model, with the coefficients drawn from a normal. Y is binary, with beta's coefficients also from a normal '''
+    if not seed == None:
+        npran.seed(seed)
+    p = p1 + pnull
+    h,w = data.shape
+    rows = npran.choice(h,n)
+    X      = data[rows,:][:,npran.choice(w,p)]
+    X_1    = X[:,:p1]
+    X_null = X[:,p1:]
+    beta   = npran.randn(p1)*beta_sd
+    if p1>0:
+        eta    = np.dot(X_1,beta)+logit(base_prob)
+        y      = eta + npran.normal(0,sd,n)
+    else:
+        y      = npran.normal(0,sd,n)
+    return X,y
+
+def genXy_binary_X_norm_y(seed,n,p1,pnull,base_prob=.25,beta_sd=1,A_base_diag=-1,A_sd=.2):
+    ''' X is binary from the isling model, with the coefficients drawn from a normal. Y is binary, with beta's coefficients also from a normal '''
+    if not seed == None:
+        npran.seed(seed)
+    p = p1 + pnull
+    A = npran.normal(0,.2,(p,p))-np.diag(A_base_diag*np.ones(p))
+    X = draw_random_binary(n,A)
+    X_1    = X[:,:p1]
+    X_null = X[:,p1:]
+    beta   = npran.randn(p1)*beta_sd
+    if p1>0:
+        eta    = np.dot(X_1,beta)
+        y      = eta + npran.normal(0,sd,n)
+    else:
+        y      = npran.normal(0,sd,n)
+    return X,y
+
+
+
+
+def genXy_given_X_norm_beta(seed,data,n,p1,pnull,base_prob=.25,beta_sd=1):
+    ''' X is binary from the isling model, with the coefficients drawn from a normal. Y is binary, with beta's coefficients also from a normal '''
+    if not seed == None:
+        npran.seed(seed)
+    p = p1 + pnull
+    h,w = data.shape
+    rows = npran.choice(h,n)
+    X      = data[rows,:][:,npran.choice(w,p)]
+    X_1    = X[:,:p1]
+    X_null = X[:,p1:]
[…]

### 2015-07-15 masters_paper/knockoffGLM.py
Reorganized, added data, made compare functions

@@ -69,7 +69,7 @@ def solve_sdp(G):
 
 class knockoff_net(object):
     """ Parent class for knockoff lasso and knockoff logistic regression. Defines everything besides fitting the particular GlmNet object """
-    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,tol=1E-5,maxiter=100,full_A=False):
+    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,tol=1E-5,maxiter=40,full_A=False):
         self.y      = y
         self.X      = normalize(X.astype(float),norm='l2',axis=0)
         self.X_orig = X
@@ -199,15 +199,19 @@ class knockoff_net(object):
             cond2 = np.outer(mu_lrg,np.ones(2*self.p))+np.outer(np.ones(2*self.p),mu_lrg) -1 <= self.M
             testmat = np.logical_or(np.logical_and(cond1,cond2),np.diag(np.ones(2*self.p)))  # ignore diagonal
             test = np.min(testmat)
+
+            shrink_s = False
+            if shrink_s:
+                if shrink<.25:
+                    test = True
                 
-            #if test and shrink<1:
-            #    print "s had to be shrunk by a factor of %.2f" % shrink
+                if test and shrink<1:
+                    print "s had to be shrunk by a factor of %.2f" % shrink
 
-            if shrink<.05:
-                print "Still won't work"
-                test = True
-
-            shrink = shrink*.90
+                shrink = shrink*.90
+            elif not test:
+                test = True 
+                print "Cross moment test failed, didn't do anything about it"
 
         ####################################################
         # Fit the A matrix for the original variables first. 

### 2015-07-15 masters_paper/lasso_compare.py

@@ -1,0 +1,124 @@
+#imports
+from __future__ import division
+import numpy as np
+import numpy.linalg as nplin
+import numpy.random as npran
+import scipy as sp
+import scipy.linalg as splin
+import scipy.stats as spstat1gt
+import matplotlib.pyplot as plt
+import json
+import inspect
+import glmnet
+import knockoffGLM as ko
+import simulate as sim
+
+def lasso_compare(n,b,p,q,p1s):
+    # since I'm parrallel computing, need to make sure I'm using different seeeds to generate random numbers
+    # these come from random.org, and should be truely random
+    # the function is a tad spotty due parsing a web page isntead of using the API, so may need to be run a few times
+    # also, it errors out when the request is rejected due to server load
+    seeds = sim.getRandomIntegers(num=10000,min=1,max=1000000000)
+
+    # Load the genetic data
+    genes = np.genfromtxt('data/SNPdata.txt', delimiter=',')
+    # replace all the 2s with 1s
+    np.place(genes,genes!=0,1)
+
+    isl_data_lasso = np.empty((5,len(p1s)))
+    gen_data_lasso = np.empty((5,len(p1s)))
+
+    for i in range(len(p1s)):
+        p1 = p1s[i]
+        isl_data_lasso[:,i] = sim.ko_test(procs=4,
+                                          seeds=seeds[(i*b):(i*b+b)],
+                                          q=q,
+                                          knockoff='both',
+                                          model='lasso',
+                                          func=sim.genXy_binary_X_norm_y,
+                                          n=n,
+                                          p1=p1,
+                                          pnull=p-p1,
+                                          beta_sd=1,
+                                          A_base_diag=-1,
+                                          A_sd=.2
+                                ).calc_fdr_power()
+        print "Done Ising %d" % (i+1)
+        gen_data_lasso[:,i] = sim.ko_test(procs=4,
+                                          seeds=seeds[(i*b):(i*b+b)],
+                                          q=q,
+                                          knockoff='both',
+                                          model='lasso',
[…]

### 2015-07-15 masters_paper/logit_compare.py

@@ -1,0 +1,104 @@
+#imports
+from __future__ import division
+import numpy as np
+import numpy.linalg as nplin
+import numpy.random as npran
+import scipy as sp
+import scipy.linalg as splin
+import scipy.stats as spstat1gt
+import matplotlib.pyplot as plt
+import json
+import inspect
+import glmnet
+import knockoffGLM as ko
+import simulate as sim
+
+def logit_compare(n,b,p,q,p1s):
+    # since I'm parrallel computing, need to make sure I'm using different seeeds to generate random numbers
+    # these come from random.org, and should be truely random
+    # the function is a tad spotty due parsing a web page isntead of using the API, so may need to be run a few times
+    # also, it errors out when the request is rejected due to server load
+    seeds = sim.getRandomIntegers(num=10000,min=1,max=1000000000)
+
+    # Load the genetic data
+    genes = np.genfromtxt('data/SNPdata.txt', delimiter=',')
+    # replace all the 2s with 1s
+    np.place(genes,genes!=0,1)
+
+    isl_data_logit = np.empty((5,len(p1s)))
+    gen_data_logit = np.empty((5,len(p1s)))
+
+    for i in range(len(p1s)):
+        p1 = p1s[i]
+        isl_data_logit[:,i] = sim.ko_test(procs=4,
+                                          seeds=seeds[(i*b):(i*b+b)],
+                                          q=q,
+                                          knockoff='binary',
+                                          model='logit',
+                                          func=sim.genXy_binary_X_norm_beta,
+                                          n=n,
+                                          p1=p1,
+                                          pnull=p-p1,
+                                          beta_sd=1,
+                                          A_base_diag=-1,
+                                          A_sd=.2
+                                ).calc_fdr_power()
+        print "Done Ising %d" % (i+1)
+        gen_data_logit[:,i] = sim.ko_test(procs=4,
+                                          seeds=seeds[(i*b):(i*b+b)],
+                                          q=q,
+                                          knockoff='binary',
+                                          model='logit',
+                                          func=sim.genXy_given_X_norm_beta,
+                                          n=n,
[…]

### 2015-07-15 masters_paper/simulate.py

@@ -7,7 +7,7 @@ import scipy as sp
 import scipy.linalg as splin
 import scipy.stats as spstat
 import matplotlib.pyplot as plt
-import requests
+import urllib3, certifi
 from scipy.special import expit as invlogit
 from scipy.special import logit
 
@@ -31,8 +31,13 @@ def getRandomIntegers(num=1000,min=1,max=1000000000,base=10):
     
     Since I'm not using random.org's API (do I look like I'm made out of money?!?), a maximum of 10,000 random numbers can be generated, and they min and max must be within +/- 1,000,000,000
     """
+
+    http = urllib3.PoolManager(
+        cert_reqs='CERT_REQUIRED', # Force certificate check.
+        ca_certs=certifi.where(),  # Path to the Certifi bundle.
+    )
     address = 'https://www.random.org/integers/?num=%d&min=%d&max=%d&col=1&base=%d&format=html&rnd=new' % (num,min,max,base)
-    r       = requests.get(address)
+    r       = http.request('get',address)
     uniList = BeautifulSoup(r.text).find('pre',class_='data').text.split()
     numList = map(int,uniList)
     return numList
@@ -55,13 +60,13 @@ def genXy_given_X_norm_y(seed,data,n,p1,pnull,sd=1,beta_sd=1):
     X_null = X[:,p1:]
     beta   = npran.randn(p1)*beta_sd
     if p1>0:
-        eta    = np.dot(X_1,beta)+logit(base_prob)
+        eta    = np.dot(X_1,beta)
         y      = eta + npran.normal(0,sd,n)
     else:
         y      = npran.normal(0,sd,n)
     return X,y
 
-def genXy_binary_X_norm_y(seed,n,p1,pnull,base_prob=.25,beta_sd=1,A_base_diag=-1,A_sd=.2):
+def genXy_binary_X_norm_y(seed,n,p1,pnull,sd=1,beta_sd=1,A_base_diag=-1,A_sd=.2):
     ''' X is binary from the isling model, with the coefficients drawn from a normal. Y is binary, with beta's coefficients also from a normal '''
     if not seed == None:
         npran.seed(seed)
@@ -196,13 +201,14 @@ def gen_w_s(kwargs):
         elif kwargs['model']=='lasso':
             knockoff_model = ko.knockoff_lasso(y,X,kwargs['q'],knockoff=kwargs['knockoff'])
         knockoff_model.fit()
-        return (knockoff_model.w_value, knockoff_model.S)
[…]

### 2015-07-15 masters_paper/lasso_compare.py
Small typo

@@ -115,7 +115,7 @@ def lasso_compare(n,b,p,q,p1s):
 
 
 def main():
-    lasso_compare(n=1000,b=100,p=100,q=.2,p1s[5,10,15,20,25])
+    lasso_compare(n=1000,b=100,p=100,q=.2,p1s=[5,10,15,20,25])
 
 if __name__ == '___main__':
     status = main()

### 2015-07-15 masters_paper/logit_compare.py

@@ -95,7 +95,7 @@ def logit_compare(n,b,p,q,p1s):
 
 
 def main():
-    logit_compare(n=1000,b=100,p=100,q=.2,p1s[5,10,15,20,25])
+    logit_compare(n=1000,b=100,p=100,q=.2,p1s=[5,10,15,20,25])
 
 if __name__ == '___main__':
     status = main()

### 2015-07-15 masters_paper/lasso_compare.py
Fixed simulation code

@@ -7,18 +7,21 @@ import scipy as sp
 import scipy.linalg as splin
 import scipy.stats as spstat1gt
 import matplotlib.pyplot as plt
-import json
-import inspect
 import glmnet
 import knockoffGLM as ko
 import simulate as sim
 
-def lasso_compare(n,b,p,q,p1s):
+def lasso_compare(id,n,b,p,q,p1s,new_seeds=False,procs=4,plot=False):
     # since I'm parrallel computing, need to make sure I'm using different seeeds to generate random numbers
-    # these come from random.org, and should be truely random
-    # the function is a tad spotty due parsing a web page isntead of using the API, so may need to be run a few times
-    # also, it errors out when the request is rejected due to server load
-    seeds = sim.getRandomIntegers(num=10000,min=1,max=1000000000)
+    if new_seeds:
+        # these come from random.org, and should be truely random
+        # the function is a tad spotty due parsing a web page isntead of using the API, so may need to be run a few times
+        # also, it errors out when the request is rejected due to server load
+        seeds = sim.getRandomIntegers(num=10000,min=1,max=1000000000)
+    else:
+        # failing that, backup seeds gotten by good ol copy and paste
+        f = open('data/backup_seeds.txt','r')
+        seeds = [int(seed) for seed in f.read().split()]
 
     # Load the genetic data
     genes = np.genfromtxt('data/SNPdata.txt', delimiter=',')
@@ -30,7 +33,7 @@ def lasso_compare(n,b,p,q,p1s):
 
     for i in range(len(p1s)):
         p1 = p1s[i]
-        isl_data_lasso[:,i] = sim.ko_test(procs=4,
+        isl_data_lasso[:,i] = sim.ko_test(procs=procs,
                                           seeds=seeds[(i*b):(i*b+b)],
                                           q=q,
                                           knockoff='both',
@@ -44,7 +47,12 @@ def lasso_compare(n,b,p,q,p1s):
                                           A_sd=.2
                                 ).calc_fdr_power()
         print "Done Ising %d" % (i+1)
[…]

### 2015-07-15 masters_paper/logit_compare.py

@@ -7,30 +7,33 @@ import scipy as sp
 import scipy.linalg as splin
 import scipy.stats as spstat1gt
 import matplotlib.pyplot as plt
-import json
-import inspect
 import glmnet
 import knockoffGLM as ko
 import simulate as sim
 
-def logit_compare(n,b,p,q,p1s):
+def logit_compare(id,n,b,p,q,p1s,new_seeds=False,procs=4,plot=False):
     # since I'm parrallel computing, need to make sure I'm using different seeeds to generate random numbers
-    # these come from random.org, and should be truely random
-    # the function is a tad spotty due parsing a web page isntead of using the API, so may need to be run a few times
-    # also, it errors out when the request is rejected due to server load
-    seeds = sim.getRandomIntegers(num=10000,min=1,max=1000000000)
+    if new_seeds:
+        # these come from random.org, and should be truely random
+        # the function is a tad spotty due parsing a web page isntead of using the API, so may need to be run a few times
+        # also, it errors out when the request is rejected due to server load
+        seeds = sim.getRandomIntegers(num=10000,min=1,max=1000000000)
+    else:
+        # failing that, backup seeds gotten by good ol copy and paste
+        f = open('data/backup_seeds.txt','r')
+        seeds = [int(seed) for seed in f.read().split()]
 
     # Load the genetic data
     genes = np.genfromtxt('data/SNPdata.txt', delimiter=',')
     # replace all the 2s with 1s
     np.place(genes,genes!=0,1)
 
-    isl_data_logit = np.empty((5,len(p1s)))
-    gen_data_logit = np.empty((5,len(p1s)))
+    isl_data_logit = np.empty((2,len(p1s)))
+    gen_data_logit = np.empty((2,len(p1s)))
 
     for i in range(len(p1s)):
         p1 = p1s[i]
-        isl_data_logit[:,i] = sim.ko_test(procs=4,
+        isl_data_logit[:,i] = sim.ko_test(procs=procs,
                                           seeds=seeds[(i*b):(i*b+b)],
                                           q=q,
                                           knockoff='binary',
@@ -43,8 +46,13 @@ def logit_compare(n,b,p,q,p1s):
                                           A_base_diag=-1,
                                           A_sd=.2
                                 ).calc_fdr_power()
+        # save data
[…]

### 2015-07-15 masters_paper/simulate.py

@@ -38,7 +38,7 @@ def getRandomIntegers(num=1000,min=1,max=1000000000,base=10):
     )
     address = 'https://www.random.org/integers/?num=%d&min=%d&max=%d&col=1&base=%d&format=html&rnd=new' % (num,min,max,base)
     r       = http.request('get',address)
-    uniList = BeautifulSoup(r.text).find('pre',class_='data').text.split()
+    uniList = BeautifulSoup(r.data).find('pre',class_='data').text.split()
     numList = map(int,uniList)
     return numList
 

### 2015-07-16 masters_paper/paper.tex
Hopefully fixed choosing Simga

@@
 In other words, where $\hat \Sigma := X^TX$,
-    \[ \tilde X^T\tilde X=\hat\Sigma \textrm{ and } \tilde X^T X = \Sigma - \diag{\mathbf s}\]
+    \[ \tilde X^T\tilde X=\Sigma \textrm{ and } \tilde X^T X = \Sigma - \diag{\mathbf s}\]
     for some vector $s\in \R^p$ such that, where $\diag{\Sigma}$ is the vector made from the main diagonal of $\Sigma$, $\diag{\Sigma} - \mathbf s$ is small.
@@
 How these knockoffs are actually generated will be described shortly.\par
-    The second step is to fit a series of Lasso model of $\mathbf y$ on the combined design matrix $[X \; \tilde X]$ so as to determine the largest value $\lambda$ at which the coefficient for each $X_j$ and $\tilde X_j$ is nonzero.\footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} Recall, for a given $\lambda$, the estimated coefficient from the Lasso regression will be
-    \[ \beta(\lambda) = \textrm{argmin}_\mathbf b \left\{\frac{1}{2}\|\mathbf y - [X\; \tilde X]\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]
+    The second step is to fit a series of Lasso model of $\mathbf y$ on the combined design matrix $X_L=[X \; \tilde X]$ so as to determine the largest value $\lambda$ at which the coefficient for each $X_j$ and $\tilde X_j$ is nonzero.
+The subscript $L$ refers to the large design matrix with both features and knockoffs.
+\footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} Recall, for a given $\lambda$, the estimated coefficient from the Lasso regression will be
+    \[ \beta(\lambda) = \textrm{argmin}_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]
     In the general case, the coefficient for one feature in a Lasso regression being first nonzero for a larger value of $\lambda$ than another feature is an indication that the first feature is a stronger predictor of the outcome than the second.
 Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to have nonzero coefficients sooner than the knockoff features when the original is a valid predictor.
-On the other hand, since $\beta(\lambda)$ only depends on $[X\; \tilde X]$ through the sufficient statistics $[X\; \tilde X]^T[X\; \tilde X]$ and $[X\; \tilde X]^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
-This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $[X\; \tilde X]^T[X\; \tilde X]$ will be unaltered, while $\E\left[[X\; \tilde X]^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$.
+On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $\Sigma_L=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
+This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $\Sigma_L$ will be unaltered, while $\E\left[X_L^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$.
 \par
@@
     The original formulation of the method offers two similar methods of constructing knockoffs. Both of these will, by construction, have exactly the property that
-    \[ \tilde X^T\tilde X=\hat\Sigma \textrm{ and } \tilde X^T X = \Sigma - \diag{\mathbf s}.\]
-    For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. Then, let the Gram matrix of $[X\; \tilde X]$ be
-    \[ G:= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} \Sigma & \Sigma - \diag{\mathbf s} \\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \]
-    $A$, the Schur complement of $\Sigma$ in $G$ can be calculated as 
+    \[ \tilde X^T\tilde X=\Sigma \textrm{ and } \tilde X^T X = \Sigma - \diag{\mathbf s}.\]
+    For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. Then, let the Gram matrix of $X_L$ is
+    \[ \Sigma_L= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} \Sigma & \Sigma - \diag{\mathbf s} \\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \]
+    $A$, the Schur complement of $\Sigma$ in $\Sigma_L$ can be calculated as 
     \[ A = 2\,\diag{\mathbf s} - \diag{\mathbf s}\Sigma^{-1}\diag{\mathbf s} \]
-    For $G$ to exist, $G$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if\footnote{A fleshed out version of this argument can be found in the original knockoff paper}
+    For $\Sigma_L$ to exist, $\Sigma_L$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if\footnote{A fleshed out version of this argument can be found in the original knockoff paper}
     \[ \diag{\mathbf s} \succeq 0  \; \textrm{ and } \; 2\Sigma - \diag{\mathbf s} \succeq 0 \]
@@
     \[ \beta(\lambda) = \textrm{argmin}_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
-    The $Z_j$ and $\tilde Z_j$ are then once again the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before.
+    The $Z_j$ and $\tilde Z_j$ are then, once again, the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before.
 The problem with this is that if the given GLM doesn't have the same sufficient statistics as linear regression, there is no theoretical guarantee that the knockoff filter will control FDR.
@@
         \begin{center}
-        \includegraphics[width=15cm]{images/entryrate_original_logit}
+        \includegraphics[width=14cm]{images/entryrate_original_logit}
     \end{center}
@@
 
-    \FloatBarrier
-    To test how bad this problem is, I simulated the method with the original knockoffs with logistic regression.
+    To test how bad this problem is, I tested the method using original style knockoffs in logistic regression.
 The results can be seen in figure 1, which shows cumulatively, starting with the largest absolute value, what portion of the $W_j$ are positive.
@@
 We see that in the top left corner, when all features are null and normally distributed, that we almost achieve this, with only a small, but consistent, deviation from $50\%$.
-However, in the top right, when the predictors are binary, the null originals enter first far more often, which will the FDR too high.
+However, in the top right, when the predictors are binary, the null originals enter first far more often, which will result in the FDR being too high.
 \par
-    This gets much worse when some true predictors are included, in the bottom row of the figure.
+    \FloatBarrier
+    This gets much worse when some true predictors are included, as shown in the bottom row of the figure.
 For both normal and binary predictors, the originals come in first far more often in the beginning.
@@
     \begin{figure}[h]
-        \includegraphics[width=15cm]{images/normalQQ}
+        \includegraphics[width=14cm]{images/normalQQ}
         \caption{Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
@@
 
-\section{Random Bernoulli Knockoffs}
-    My solution to the issues with the
+\section{Random Binary Knockoffs}
+Random binary knockoffs $\mathbf{\tilde x}$, which maintain the correlation condition in expectation, fit the bill in this regard.
+The method to produce them is based heavily on more general methodology to random binary vector variables with internal correlation.
 
-    generate $\bf \tilde x$ randomly such that it has the desired knockoff properties in expectation.
-In particular, both  should have similar marginal densities, expectations, and second moments.
-However, $\bf \tilde x\st \bf x$ should also have desired knockoff property that $\E(\mathbf{\tilde x}^T\mathbf x\st \mathbf x) = \Sigma-\mathbf s$, where $\|\diag{\Sigma}-s\|$ is small.
-In the general case, this is likely infeasible; however, if $\bf x$ is a binary vector, as is often the case, we know we are dealing with a much more limited class of random variables, and it should be possible to randomly generate $\bf \tilde x\st \bf x$ so as to have the desired properties.
-At worst, this method will provide a suitable replacement for deterministic $\bf \tilde x$ for use with LASSO, and if we are lucky, it will work reasonably for other regularized GLMs.
-
-\subsection{Random Bernoulli Generation}
-Thankfully, there has been a reasonable amount of work on how one can generate random Bernoulli vectors with some kind of correlation among among the values.
-A random Bernoulli vector $\bf x$ can be summarized by its first two moments: a mean vector $\E(\bf x)=m\in(0,1)^p$ and cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$.
+\subsection{Random Binary Vector Generation}
+The full class of random binary vector variables on $\{0,1\}^p$ can be specified as multinomial on the $2^p$ elements of the set.
+However, in practice, it is generally impractical to specify or estimate all $2^p$ necessary parameters, so it is useful to, as is desirable for knockoffs, pick a method to match the first two moments.
+These are the mean vector $\E(\bf x)=m\in(0,1)^p$ and the cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$.
 Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag{M}$.
-For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-mm^T\succeq0$, and
+For an arbitrary symmetric $M$ to be valid cross-moment matrix for some random binary vector variable, $M-mm^T\succeq0$, and
     \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
-    for all $i\neq j$\footnote{``On parametric families for sampling binary data with specified mean and correlation" - http://arxiv.org/abs/1111.0576}.
-Given a qualifying $M$, or observed $X$, there are a few ways of generating more random $\bf x$.
+    for all $i\neq j.$ Given a qualifying $M$, or observed $X$, there are then a few methods for generating additional random values.
 
 \subsubsection{Gaussian Copula Family}
-    Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $\mathbf z\sim N_p({\bf 0},\Sigma)$ such that,for $x_i = I(z_i<0)$, $x$ has the desired properties.
-There are a number of ways to do this\footnote{``On the Generation of Correlated Artificial Binary Data" - http://epub.wu.ac.at/286/1/document.pdf}\footnote{``On parametric families for sampling binary data with specified mean and correlation"}, but it turns out that there is only certain to exist a working $\Sigma$ in the bivariate case.
+Since multivariate normal distributions are easy to randomly draw from and defined by their first two moments, an obvious choice is to use them to generate random binary vectors.
+This might be done by finding $\mu$ and $\Sigma$ such that for $\mathbf z\sim N_p(\mu,\Sigma)$, $x_i$ can be generated as $x_i = \mathrm{sign}(z_i)$, and have the desired first two moments.
+However, this is only guaranteed to be feasible in the bivariate case.
+For higher dimensions, this method can generally only provide random binary vectors with approximately the right moments, and the quality of the approximation quickly degenerates as $p$ increases.
 
@@
     There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as 
-
     \[ \p(x_{i}=1\st x_{1},...,x_{i-1}) = \mu\left(a_{ii}+\sum_{k=1}^{i-1}a_{ik}x_i\right) \]
-
     for some monotone function $\mu:\R\to(0,1)$.
@@
 
-    If $\mu$ is chosen such that it is a bijection and differentiable, there is a unique $A$ such that $\E(\mathbf x \mathbf x^T)=M$ when generated from this model\footnote{``On parametric families for sampling binary data with specified mean and correlation"}.
-The natural choice for $\mu$ is the logistic link function, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
+    If $\mu$ is chosen such that it is a bijection and differentiable, there is a unique $A$ such that $\E(\mathbf x \mathbf x^T)=M$ when generated from this model.
+The natural choice for $\mu$ is the logistic link function $L(x)=\frac{1}{1-e^{-x}}$, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
     \[a_{ij} = \log\left(\frac{\p(x_j=x_k=1)\p(x_j=x_k=0)}{\p(x_j=0,x_k=1)\p(x_j=1,x_k=0)}\right) \]
-    when $i\neq j$. I think this dictates that if $\bf x$ is generated from this model with $a_{ij}=0$, then $x_i$ and $x_j$ are independent.  \par
+    when $i\neq j$. When $\bf x$ is generated from this model with $a_{ij}=0$, then $x_i$ and $x_j$ are conditionally independent.  \par
 
     There is no closed form to calculate the entries in $A$ if $p>1$, but they can be derived numerically two ways.
-    \begin{enumerate}
-        \item If one is attempting to replicate the empirical cross-moments from a data matrix $X$, $a_{1i}$ to $a_{ii}$ can be derived from fitting successive logistic regressions of $X_i$ on $X_{1} \ldots X_{i-1}$ using maximum likelihood.
+The first of these, when one is attempting to replicate the empirical cross-moments from a data matrix $X$, $a_{1i}$ to $a_{ii}$ is by fitting successive binomial family GLM regressions.
+The link function is the given function $\mu$, and $X_i$ is regressed on $X_{1} \ldots X_{i-1}$ using maximum likelihood.
 $a_{ji}$ for $i\neq j$ will then be the coefficient on $X_j$ while $a_{ii}$ is the intercept of the regression.
-        \item If one is just working with a desired cross-moment matrix $M$, the successive rows of $A$ can be fit via Newton-Raphson. \par
-            Let us say that the first $i-1$ rows have already been fit, resulting in the upper left $(i-1)\times(i-1)$ sub matrix $A_{-i}$ of $A$.
-Let us say that $\mathbf a_i$ is the first $i$ entries of the $i$th row of $A$ (the rest will be 0 anyway).
-As well, let $\mathbf m_i$ be similarly the first $i$ entries of the $i$th row of $M$.
+\par
+    Otherwise, if one is just working with a desired cross-moment matrix $M$, the successive rows of $A$ can be fit via Newton-Raphson.
+This is performed successively on each row of $A$.
+If the first $i-1$ rows have been fit, then the upper left $(i-1)\times(i-1)$ sub matrix $A_{-i}$ of $A$ has already been filled.
+Next, $\mathbf a_i$, the first $i$ entries of the $i$th row of $A$ must be fit, while the rest of the row will be 0.
+This corresponds to $\mathbf m_i$, the first $i$ entries of the $i$th row of $M$.
 In other words, $\mathbf{m}_i = [\E(x_i x_j)]_{j=1}^{i}$.
 Finally, let us say that $\mathbf x_{-i}$ is the first $i-1$ entries of $\mathbf x$.
-We want to solve for $\mathbf a_i$ such that
+$\mathbf a_i$ must satisfy
             \begin{align*}
@@
                 \mathbf m_i &= \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \p(\mathbf x_{-i})\p(x_i=1\st \mathbf x_{-i}) \left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \\
-        \mathbf m_i &= \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}\mu\left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) \mu\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \\
             \end{align*}
+            When $\mu$ is the logistic link function $L$, this is
+            \[ \mathbf m_i = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L\left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \]
             Where $c$ is the appropriate normalizing constant.
 Let us define the quantity on the right in the last line as $f(\mathbf a_i)$.
-We can solve for $f(\mathbf a_i)=\mathbf{m}_i$ by successive Newton-Raphson iterations defined by
+This equation $f(\mathbf a_i)=\mathbf{m}_i$ can be solved by successive Newton-Raphson iterations defined by
             \[\mathbf a_i^{(k+1)} = \left[H f\left(\mathbf a_i^{(k)}\right)\right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
-            The Hessian matrix is calculated as 
-        \[ H f\left(\mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}\mu \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
+            To do so, the Hessian matrix is calculated as 
+        \[ H f\left(\mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
             With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive.
-Instead, with a series of values $\mathbf x_{-i}^{(k)}\sim \mathbf x_{-i}$, we can approximate
-            \[f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K \mu\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right] \]
+Instead, with a series of values $\mathbf x_{-i}^{(k)}\sim \mathbf x_{-i}$, the simulated values
+            \[f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right] \]
             and
-            \[H f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
-            Though in theory $A$ should always exist, in practice numerical issues may compound to the point that the Newton-Raphson method won't converge.
+            \[H f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
+            can be used. \par
+            Though, in theory, $A$ should always exist, in practice numerical issues may compound to the point that the Newton-Raphson method won't converge.
 In this case, one can solve instead for $\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
     \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T \]
-            When $\tau=0$, this yields the original problem, while when $\tau=1$, it is treating $x_i$ as independent of $\mathbf x_{-i}$. The latter will always have the solution
+            When $\tau=0$, this yields the original problem, while when $\tau=1$, it is treats $x_i$ as independent of $\mathbf x_{-i}$. The latter will always have the solution
             \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{M_{ii}}{1-M_{ii}}\right) \end{array} \right]^T \]
             The hope is that for some $\tau$ close to $0$, convergence can be achieved, only causing a slight distortion from the desired cross moments.
-    \end{enumerate}
 
 \subsection{Generating Binary Knockoffs}
-    My method for generating binary knockoffs broadly involves two steps:
+This theory provides the tools to generate random binary knockoffs.
+This is done is done sequentially as two steps, the first of which is to come up with the target covariance matrix for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$.
+Much as before with the original knockoffs, both the equi-correlated and SDP method can be used to select  $\mathbf s$  such that $\diag{\Sigma}-\mathbf s$ is small and
+        \[ \Sigma_L = \left[\begin{array}{cc}  \Sigma & \Sigma - \diag{\mathbf s }\\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \succeq 0 \]
+        Additionally though, the moment condition must hold for this to correspond to a proper random binary vector.
+If $\mathbf m_L = \E\left(\mathbf x_L\right) = [\mathbf m \; \mathbf m]$, then the desired cross moment matrix of the joint distribution is
+        \[ M_L = \Sigma_L + \mathbf m_L\mathbf m_L^T\]
+        So it must be the case that  
+        \[\max\{0,\mathbf m_{L,i} + \mathbf m_{L,j} -1\} \leq \Sigma_{L,ij} + \mathbf m_i \mathbf m_j \leq \min\{\mathbf m_{L,i}, \mathbf m_{L,j}\} \]
+        However, since $\Sigma$ and $\mathbf m$ are from an actual random binary vector, the entries of $\Sigma$ satisfy this property by default.
+Only off diagonal entries which have had $\mathbf s$ subtracted from them could possibly violate this condition.
+Furthermore, for these entries, the goal is to make $\Sigma_{L,ij}$ as close to $0$ as possible.
+Since the condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$, this implies that this method should normally be satisfied.
+For the SDP method though, the restriction can be built directly into the optimization problem, yielding the updated problem
+            \begin{center}
+                \begin{tabular}{r l}
+                    minimize & $\|\mathbf 1-\diag{\Sigma}^{-1}\mathbf s\|_1$ \\
+                    subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2\Sigma $ \\
+                               & $ 2\mathbf m_i -1 \leq \Sigma_{ii}-\mathbf s_i + \mathbf m_i^2 \leq \mathbf m_i  \quad \forall i$
+                \end{tabular} 
+            \end{center} 
+        replacing $\mathbf s$ with $\diag{\Sigma}^{-1}\mathbf s$ in the objective is to account for not normalizing $X$, as is usually done with the original knockoffs. \par
+        Once a proper cross moment matrix $M_L$ has been selected, one could simply apply the Newton-Raphson method to fit a corresponding $A$ matrix, and then generate $\mathbf{ \tilde x}$ sequentially conditioned on $x$, as $\tilde x_i \st \mathbf x,\tilde x_1,\ldots,\tilde x_{i-1}$.
+However, there are a few computational wrinkles which can be chosen to speed up the process:
         \begin{enumerate}
-        \item I use either the equal correlation method or the SDP method described in the original knockoff paper to find $\mathbf s$ such that $\|\diag{\hat\Sigma}-\mathbf s\|$ is small and 
-        \[ \Sigma_L = \left[\begin{array}{cc}  \hat\Sigma & \hat\Sigma - \diag{\mathbf s }\\ \hat\Sigma - \diag{\mathbf s} & \hat\Sigma \end{array}\right] \succeq 0 \]
-        I use a subscript $L$ for large to denote items associated with the joint distribution of $[\mathbf x \; \mathbf{\tilde x}]$.
-If $\mathbf m_L = \E\left([\mathbf x \; \mathbf{\tilde x}]\right)^T = [\mathbf m \; \mathbf m]^T$, then the desired cross moment matrix of the joint distribution is
-        \[ M_L = \Sigma_L + \mathbf m_L\mathbf m_L^T\]
-        To ensure that this is a valid cross moment matrix for a binary random vector is that 
-        \[\max\{0,\mathbf m_{L,i} + \mathbf m_{L,j} -1\} \leq M_{L,ij} \leq \min\{\mathbf m_{L,i}, \mathbf m_{L,j}\} \]
-        I've built a check into the code for this, but in practice it shouldn't be a worry.
-This condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$.
-Since the we are either keeping the value $M_{L,ij}$ from a valid cross moment matrix or minimizing $\vert M_{L,ij} - \mathbf m_{L,j}\mathbf m_{L,i}\vert$, it would be very surprising if this condition was violated.
+            \item If the whole $A$ matrix is being fit, then the upper half of the matrix, corresponding to $\mathbf x$, can be fit via the regression method, since they arise from real data $X$.
+This will tend to be quicker, since the calculation lends itself to parallel computing, and the data set will often be smaller than the space of all binary vectors or a simulation of that space.
+The second half of the matrix still must be fit via Newton-Raphson.
+            \item One can avoid fitting the first half of the $A$ matrix entirely though if performing Newton-Raphoson with by using the actual observed values in $X$ to simulate $\mathbf x$.
+This can be done by either bootstrapping rows from $X$ or using some number of copies of $X$ exactly.
+The advantage of this is the saved computation and that $\tilde x$ should be built to match the correlation condition more exactly.
+The downside is that, since these values won't be from the Ising model, the theoretical guarantee of a suitable $A$ may no longer hold.
+            \item As $A$ is being fit, even the parts of $\mathbf x_{-i}$ corresponding to knockoffs need not be  redrawn for each Newton-Raphson iteration or each successive $i$.
+This saves computation, and the multiplication
+            \[ \left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [\mathbf x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
+                doesn't need to be recomputed for each Newton-Raphson iteration, and the upper left portion, $\mathbf x_{-i}^{(k)}[\mathbf{x}_{-i}^{(k)}]^T$ doesn't have to be recomputed for successive $i$.
+This may result in slight deviations in any particular draw compounding into a more severe error over time.
+        \end{enumerate}
+        Combining the second and third trick leads to the final formulation.
+Rather then $K$ simulations of $\mathbf x$, the simulation is started with the matrix $X_F$, which is $X$ stacked up $m$ times to get $n\times m=K'\geq K$ rows.
+The following process is then repeated for $p<i\leq 2p$:
 
-        \item I can fit the matrix $A$ that will generate random binary variables similarly to the method described in section $3.2$ which have cross moments $X_L$.
-This can be used to generate the $\tilde x_i$ sequentially as $\tilde x_i \st \mathbf x,\tilde x_1,\ldots,\tilde x_{i-1}$ so as to create $\tilde X \st X$.
-    \end{enumerate}
-    \subsection{More Detail on Fitting $A$}
-        I could fit $A$ based on $M_L$ exactly as described in the second method of $3.2$, however, this isn't exactly what I do.
-First off, with $p$ being potentially large, the simulation method for estimating $f(\mathbf a)$ and $H f(\mathbf a)$ was the obvious choice.
-This involves fitting $\mathbf a_i$ based on the conditional distribution of $\mathbf x_{L,i}$ given randomly drawn partial vectors $\mathbf x_{L,-i}$.
-There is no need to simulate the marginal distribution of $\mathbf x$ though, since the simulation is only approximate and we already have a number of realizations in $X$.
-Thus, I only fit the lower half of $A$, using this process:
-        \begin{enumerate}
-            \item To get a simulation of size at least $K$, I create a matrix $X_F$ ($F$ for fixed) which is initially $X$ stacked up until it has $K'\geq K$ rows. 
-            \item For each $p<i\leq 2p$,
          \begin{itemize}
-                    \item Where $\mathbf x_F^{(k)}$ is the $k$th row of $X_F$, the rows $\mathbf a_i$ are fit sequentially by Newton-Raphson iterations with 
+             \item Where $\mathbf x_F^{(k)}$ is the $k$th row of $X_F$, the row $\mathbf a_i$ is fit by Newton-Raphson iterations with 
              \[f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right] \]
@@
              \[H f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
-                    \item If the iterations won't converge, I attempt to fit $f(\mathbf a_i)=m_i^*(\tau)$ instead for increasing values of $\tau$ until it converges.
+             \item If the iterations won't converge, the relaxed equation $f(\mathbf a_i)=m_i^*(\tau)$ is attempted instead for increasing values of $\tau$ until it converges.
              \item After $\mathbf a_i$ is fit, a new column $X_i$ is drawn as independent Bernoulli with probability vector $\mu(X_F \mathbf a_i)$.   
@@
          \end{itemize}
-            \item At the end, the first $n$ rows of $X_F$ are taken as $[X \; \tilde X]$, though the rows corresponding to any copy of $X$ would work equally well, since these all should have the desired distribution.
-        \end{enumerate}
-        Some thoughts/concerns/explanations:
-        \begin{itemize}
-            \item Since $x_{-i}^{(k)}$ is replaced with $x_F^{(k)}$, which is a hybrid of real and simulated data, I am not sure there is the same theoretical guarantee that a unique $A$ matrix exists.
-In practice though it still worked pretty well, and has the advantage that we are deriving $\tilde X$ such that $\E(X'\tilde X) = \hat \Sigma- \diag{\mathbf s}$.
-            \item A hybrid of simulating $x_{-i}^{(k)}$ and keeping fixed $x_F^{(k)}$ would be to draw from the rows of $X$ $K$ times, with replacement, then simulating the rest of the $x_{-i}^{(k)}$ vector.
-I'm not convinced there is a good reason to do this.
-            \item By not simulating, there is the advantage of not needing to fit the upper half of $A$.
-            \item By not redrawing $x_{-i}^{(k)}$ each time, there is less computation for the computer. As well, the multiplication
-                \[ \left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
-            needn't be redone for each iteration, and only partially recalculated for each $i$ (though I don't have this implemented yet).
-The downside might be that error is getting compounded over each $i$.
-        \end{itemize}
+         At the end, the any $n$ rows of $X_F$ corresponding to the rows of $X$ can be taken as $X_L$. Any such set should have the desired distribution for $\tilde X$.
 

### 2015-07-16 masters_paper/knockoffGLM.py

@@ -137,6 +137,7 @@ class knockoff_net(object):
             self._create_equicor()
         else: 
             self._create_SDP()
+        self.X_lrg = np.concatenate((self.X_orig,self.X_ko), axis=1)
 
     def _binary_knockoff(self):
         ''' This creates the new binary knockoffs, which are random multivariate bernoulli which should have, in expectation,
@@ -148,17 +149,21 @@ class knockoff_net(object):
 
         # First, calculate s
         # Largely replicates begining of self._create_equicor()/self._create_SDP()
+        # However, we now want to perform calculations on the covariance matrix
+        # The goal is minimize the correlation between X_i and ~X_i
+        # Generate the second moment matrix Sigma
+        M    = np.dot(self.X_orig.T,self.X_orig)/self.n
+        Cov  = np.cov(self.X_orig,rowvar=0,bias=1)      # I'm doing redundant calculations here
+        Corr = np.corrcoef(self.X_orig,rowvar=0,bias=1)
 
         # SVD and come up with perpendicular matrix
-        U, d, V = nplin.svd(self.X,full_matrices=True) 
+        d, V = nplin.eig(Corr)
+        #U, d, V = nplin.svd(self.X,full_matrices=True) 
         d[d<0] = 0
         if   self.cov_type == 'SDP': 
-            # Compute the Gram matrix and its (pseudo)inverse.
-            G     = np.dot(V.T * d**2 ,V)
-            G_inv = np.dot(V.T * d**-2,V)
          
             # Optimize the parameter s of Equation 1.3 using SDP.
-            s = solve_sdp(G)
+            s = solve_sdp(Corr)
             s[s <= self.zerotol] = 0
 
         # Same as first part self._create_equicor 
@@ -168,20 +173,12 @@ class knockoff_net(object):
             lambda_min = min(d)**2
             s = np.ones(self.p)*min(2*lambda_min,1)
 
-        # Return to the original X
-        self.X = self.X_orig              # Now that we've adjusted the variance, revert back to the original X 
+        # Scale s up to original variance
+        s     = s*np.diag(Cov)        # scale s back up to original variance
 
         # Calculate first moments 
-        mu       = np.mean(self.X,axis=0)
+        mu       = np.mean(self.X_orig,axis=0)
         mu_lrg   = np.concatenate((mu,mu))
-
[…]

### 2015-07-16 masters_paper/lasso_compare.py

@@ -48,7 +48,7 @@ def lasso_compare(id,n,b,p,q,p1s,new_seeds=False,procs=4,plot=False):
                                 ).calc_fdr_power()
         print "Done Ising %d" % (i+1)
         # save data
-        fi = open('data/lasso_genetic_' +id + '.npy','w')
+        fi = open('data/lasso_isling_' +id + '.npy','w')
         np.save(fi,isl_data_lasso)
         fi.close()
 
@@ -66,7 +66,7 @@ def lasso_compare(id,n,b,p,q,p1s,new_seeds=False,procs=4,plot=False):
                                ).calc_fdr_power()
         print "Done Genetic %d" % (i+1)
         # save data 
-        fg = open('data/lasso_ising_' +id + '.npy','w')
+        fg = open('data/lasso_genetic_' +id + '.npy','w')
         np.save(fg,gen_data_lasso)
         fg.close()
 
@@ -116,7 +116,7 @@ def lasso_compare(id,n,b,p,q,p1s,new_seeds=False,procs=4,plot=False):
  
         subplts[1].legend(bbox_to_anchor=(1.05, 0), loc='lower left', borderaxespad=0.)
  
-        plt.savefig('images/lasso_corr'+id+'.pdf',bbox_inches='tight')  
+        plt.savefig('images/lasso_corr_'+id+'.pdf',bbox_inches='tight')  
 
 
 def main():

### 2015-07-16 masters_paper/logit_compare.py

@@ -66,7 +66,7 @@ def logit_compare(id,n,b,p,q,p1s,new_seeds=False,procs=4,plot=False):
                                ).calc_fdr_power()
         print "Done Genetic %d" % (i+1)
         # save data 
-        fg = open('data/logit_ising_' +id + '.npy','w')
+        fg = open('data/logit_genetic_' +id + '.npy','w')
         np.save(fg,gen_data_logit)
         fg.close()
 

### 2015-07-16 masters_paper/knockoffGLM.py
Should fix correlation

@@ -155,6 +155,8 @@ class knockoff_net(object):
         M    = np.dot(self.X_orig.T,self.X_orig)/self.n
         Cov  = np.cov(self.X_orig,rowvar=0,bias=1)      # I'm doing redundant calculations here
         Corr = np.corrcoef(self.X_orig,rowvar=0,bias=1)
+        Corr[np.isnan(Corr)]=0
+        Corr = np.max((Corr,np.identity(2*self.p)),axis=0) 
 
         # SVD and come up with perpendicular matrix
         d, V = nplin.eig(Corr)

### 2015-07-16 masters_paper/knockoffGLM.py
Should fix correlation

@@ -156,7 +156,7 @@ class knockoff_net(object):
         Cov  = np.cov(self.X_orig,rowvar=0,bias=1)      # I'm doing redundant calculations here
         Corr = np.corrcoef(self.X_orig,rowvar=0,bias=1)
         Corr[np.isnan(Corr)]=0
-        Corr = np.max((Corr,np.identity(2*self.p)),axis=0) 
+        Corr = np.max((Corr,np.identity(self.p)),axis=0) 
 
         # SVD and come up with perpendicular matrix
         d, V = nplin.eig(Corr)

### 2015-07-18 masters_paper/paper.tex
Whole ton of updates. Model should finally work

@@
             \[\mathbf a_i^{(k+1)} = \left[H f\left(\mathbf a_i^{(k)}\right)\right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
-            To do so, the Hessian matrix is calculated as 
-        \[ H f\left(\mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
+            To do so, the Jacobian matrix is calculated as 
+        \[ J f\left(\mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
             With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive.
@@
             and
-            \[H f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
+            \[J f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
             can be used. \par
@@
 This theory provides the tools to generate random binary knockoffs.
-This is done is done sequentially as two steps, the first of which is to come up with the target covariance matrix for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$.
+This is done is done sequentially as two steps.
+The first of these is to come up with the target correlation matrix $C$ for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$.
+
+the first of which is to come up with the target covariance matrix for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$.
 Much as before with the original knockoffs, both the equi-correlated and SDP method can be used to select  $\mathbf s$  such that $\diag{\Sigma}-\mathbf s$ is small and
@@
              and
-             \[H f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
+             \[J f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
              \item If the iterations won't converge, the relaxed equation $f(\mathbf a_i)=m_i^*(\tau)$ is attempted instead for increasing values of $\tau$ until it converges.
@@
 
-\section{Bernoulli Knockoff Performance}
-For each of thses, exaimine these situatations:
-\begin{itemize}
-    \item $X$ arises out of the assumed Isling model, with no high order interactions. 
-    \item $X$ does not arise out of this model, and does have higher order interactions.
-Could possibly model this by drawing the vector $\mathbf x$ or subsets of it as multinomial with probabilities from a Dirichlet distribution.
-    \item $X$ is from a real world data set, likely something in genetics.
-    \item $y$ is normally distributed around $X\beta$.
-    \item $y$ is drawn from some other distribution, possibly skewed, heavy tailed, or light tailed.
-    \item $y$ is from a real world data set.
-\end{itemize}
-\subsection{Gram Matrix Comparison}
-\subsection{Binary vs. Determinist Knockoffs in LASSO}
-\subsection{Binary Knockoffs in other Regularlized GLMS}
+\section{Binary Knockoff Performance}
+    To evaluate how useful these binary knockoffs will be in practice, it is useful to evaluate their empirical performance under simulation.
 
-\section{Discussion}
-\subsection{Areas for Further Work}
-
-\section{Further Work and Simulations}
-As I see it, further simulation work for my paper breaks down into three logical groups: comparison of random Bernoulli knockoffs to the original Knockoffs in LASSO, evaluation of Bernoulli knockoffs in other L1 regularized GLMs, and expanding random knockoffs to more general sorts of variables.
-
-\subsection{Comparison to Original Knockoffs in LASSO}
-The basic idea is to compare how the two sorts of knockoffs compare when $X$ is binary and we are fitting a LASSO regression.
-Do they select the same variables?
-Do the binary knockoffs control FDR more or less conservatively?
-Situations to test:
-\begin{itemize}
-    \item $X$ arises out of the assumed Isling model, with no high order interactions. 
-    \item $X$ does not arise out of this model, and does have higher order interactions.
-Could possibly model this by drawing the vector $\mathbf x$ or subsets of it as multinomial with probabilities from a Dirichlet distribution.
-    \item $X$ is from a real world data set, likely something in genetics.
-    \item $y$ is normally distributed around $X\beta$.
-    \item $y$ is drawn from some other distribution, possibly skewed, heavy tailed, or light tailed.
-    \item $y$ is from a real world data set.
-\end{itemize}
-
-\subsection{Evaluation of Binary Knockoffs in other GLMS}
-The one of the most interest would be logits. It would be good to see how successful the binary knockoffs are in controlling FDR without the theoretical guarantees that LASSO provides.
-\begin{itemize}
-    \item $X$ arises out of the Isling model, in which case it seems like the knockoffs should control FDR.
-    \item $X$ has higher order interactions, which might make the knockoffs perform poorly.
-    \item $X$ is from a real world data set, likely something in genetics.
-    \item $y$ is simulated based on the assumptions of the regression model.
-    \item $y$ is simulated to violate assumptions of the regression model.
-    \item $y$ is from a real world data set.
-\end{itemize}
-
-\subsection{Generalizations/Harebrained Ideas}
-I have two seeds of ideas for extensions if I have enough time.
-\begin{enumerate}
-    \item Still with binary data, one might be able to simulate binary variables with higher order interactions in the generation of $X$ by including higher order interactions in the regression of $X_i$ on $X_{-i}$.
-This would lead to $A$ being a matrix of higher dimension.
-Even if this worked, extending the method to knockoffs may not be obvious.
-    \item In the general case where $X$ is not binary, I can almost imagine a method set up along similar lines to the binary case.
-    \begin{itemize}
-        \item The desired covariance matrix could be chosen as in the original knockoff paper.
-        \item Each $x_i$'s marginal distribution would be approximated by a kernel density estimate on $X_i$.
-        \item A $x_i \st x_1, \ldots, x_{i-1}$ would be drawn from some reweighing of this marginal to achieve the proper covariance.
-For instance, if $F_i^{-1}$ is the inverse CDF of the marginal kernel density for $x_i$ and $u_i\st x_1, \ldots, x_{i-1}$ is a RV on $(0,1)$, then $x_i=F_i^{-1}(u_i)$.
-        \item Maybe fit generalized additive model for each successive $x_i$ with kernel regression for $x_1,\ldots,x_{i-1}$ to predict mean. Then, skew marginal of $x_i$ until it has that mean. 
-    \end{itemize}
-\end{enumerate}
+\subsection{Convergence to Theoretical Covariance}
 

### 2015-07-18 masters_paper/compare.py

@@ -1,0 +1,106 @@
+#imports
+from __future__ import division
+import numpy as np
+import numpy.linalg as nplin
+import numpy.random as npran
+import scipy as sp
+import scipy.linalg as splin
+import scipy.stats as spstat1gt
+import matplotlib.pyplot as plt
+import knockoffGLM as ko
+import simulate as sim
+from multiprocessing import Pool
+
+def gen_logit(input):
+    """ Testing function, all parameters are in a single tuple """
+    seed,gen,p0,p1 = input
+    if gen.lower() == "ising":
+        X,y = sim.genXy_binary_X_norm_beta(seed,1000,p1,p0,base_prob=.25,beta_sd=1,A_base_diag=-1,A_sd=.2)
+    elif gen.lower() == "genetic":
+        genes = np.genfromtxt('data/SNPdata.txt', delimiter=',')
+        np.place(genes,genes!=0,1)
+        X,y = sim.genXy_given_X_norm_beta(seed,genes,1000,p1,p0,base_prob=.25,beta_sd=1)
+
+    model = ko.knockoff_logit(y,X,.2,
+                              knockoff='binary',
+                              MCsize=25000,
+                              tol=1E-5,
+                              fresh_sim=True,
+                              opt_method='anderson',
+                              intercept=True
+                              )
+    model.fit()
+    trueS = (np.arange(p0+p1)<p1).astype(int)
+    FDR   = np.dot(model.S,1-trueS)/max(np.sum(model.S),1)
+    power = np.dot(model.S,trueS)  /max(p1,1)
+
+    with open('data/logit_test_'+str(p0+p1)+'.txt','a') as f:
+        f.write("%d\t%s\t%d\t%.5f\t%.5f\t%.5f\t%.5f\n" % (seed,gen,p1,model.M_distortion,np.mean(model.emp_ko_corr),FDR,power))
+
+def gen_lasso(input):
+    """ Testing function, all parameters are in a single tuple """
+    seed,gen,p0,p1 = input
+    if gen.lower() == "ising":
+        X,y = sim.genXy_binary_X_norm_y(seed,1000,p1,p0,beta_sd=1,A_base_diag=-1,A_sd=.2)
+    elif gen.lower() == "genetic":
+        genes = np.genfromtxt('data/SNPdata.txt', delimiter=',')
+        np.place(genes,genes!=0,1)
+        X,y = sim.genXy_given_X_norm_y(seed,genes,1000,p1,p0,beta_sd=1)
+
[…]

### 2015-07-18 masters_paper/knockoffGLM.py

@@ -16,16 +16,14 @@ import scipy as sp
 import scipy.linalg as splin
 import scipy.stats as spstat
 import matplotlib.pyplot as plt
-import cvxpy as cvx
-import rpy2 as rp 
-import rpy2.robjects.numpy2ri
-import statsmodels.api as sm 
-from rpy2.robjects.packages import importr
+#import cvxpy as cvx 
+import statsmodels.api as sm
 from glmnet import LogisticNet, ElasticNet
 from sklearn.preprocessing import normalize
 from scipy.special import expit as invlogit
 from scipy.special import logit
 from sys import float_info
+from scipy.optimize import root 
 
 def dinvlogit(x):
     '''Derivative of logit function at each point of vextor x'''
@@ -39,6 +37,15 @@ def get_index_z(coef_vector):
         if abs(coef_vector[index])>1e-6: return index
         index+=1
     return n+1
+
+def draw_random_binary(n,A):
+    ''' If p is the size of the square, lower triangular A, creates a n by p matrix of random binary vectors using Schafer's method '''
+    m,p = A.shape
+    ones   = np.ones((n,1))
+    output = np.empty((n,p))
+    for i in np.arange(0,p):
+        output[:,i] = npran.binomial(1,invlogit(np.dot(np.hstack((output[:,0:i],ones)),A[i,0:(i+1)])))
+    return output
 
 def solve_sdp(G):
     """ Solves the SDP problem:
@@ -69,7 +76,7 @@ def solve_sdp(G):
 
 class knockoff_net(object):
     """ Parent class for knockoff lasso and knockoff logistic regression. Defines everything besides fitting the particular GlmNet object """
-    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,tol=1E-5,maxiter=40,full_A=False):
+    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,tol=1E-5,fresh_sim=True,pseudocount=0,opt_method='anderson',intercept=False):
         self.y      = y
         self.X      = normalize(X.astype(float),norm='l2',axis=0)
         self.X_orig = X
@@ -79,10 +86,12 @@ class knockoff_net(object):
         self.cov_type = cov_method     # how are we generating the desired covariance matrix - equicor or SDP?
         self.randomize=False
         self.tol = tol
-        self.maxiter = maxiter
         self.MCsize = MCsize   # how many values for Monte Carlo simulation
[…]

### 2015-07-18 masters_paper/lasso_compare.py

@@ -1,129 +1,0 @@
-#imports
-from __future__ import division
-import numpy as np
-import numpy.linalg as nplin
-import numpy.random as npran
-import scipy as sp
-import scipy.linalg as splin
-import scipy.stats as spstat1gt
-import matplotlib.pyplot as plt
-import glmnet
-import knockoffGLM as ko
-import simulate as sim
-
-def lasso_compare(id,n,b,p,q,p1s,new_seeds=False,procs=4,plot=False):
-    # since I'm parrallel computing, need to make sure I'm using different seeeds to generate random numbers
-    if new_seeds:
-        # these come from random.org, and should be truely random
-        # the function is a tad spotty due parsing a web page isntead of using the API, so may need to be run a few times
-        # also, it errors out when the request is rejected due to server load
-        seeds = sim.getRandomIntegers(num=10000,min=1,max=1000000000)
-    else:
-        # failing that, backup seeds gotten by good ol copy and paste
-        f = open('data/backup_seeds.txt','r')
-        seeds = [int(seed) for seed in f.read().split()]
-
-    # Load the genetic data
-    genes = np.genfromtxt('data/SNPdata.txt', delimiter=',')
-    # replace all the 2s with 1s
-    np.place(genes,genes!=0,1)
-
-    isl_data_lasso = np.empty((5,len(p1s)))
-    gen_data_lasso = np.empty((5,len(p1s)))
-
-    for i in range(len(p1s)):
-        p1 = p1s[i]
-        isl_data_lasso[:,i] = sim.ko_test(procs=procs,
-                                          seeds=seeds[(i*b):(i*b+b)],
-                                          q=q,
-                                          knockoff='both',
-                                          model='lasso',
-                                          func=sim.genXy_binary_X_norm_y,
-                                          n=n,
-                                          p1=p1,
-                                          pnull=p-p1,
-                                          beta_sd=1,
-                                          A_base_diag=-1,
-                                          A_sd=.2
-                                ).calc_fdr_power()
-        print "Done Ising %d" % (i+1)
-        # save data
-        fi = open('data/lasso_isling_' +id + '.npy','w')
-        np.save(fi,isl_data_lasso)
-        fi.close()
-
[…]

### 2015-07-18 masters_paper/logit_compare.py

@@ -1,107 +1,0 @@
-#imports
-from __future__ import division
-import numpy as np
-import numpy.linalg as nplin
-import numpy.random as npran
-import scipy as sp
-import scipy.linalg as splin
-import scipy.stats as spstat1gt
-import matplotlib.pyplot as plt
-import glmnet
-import knockoffGLM as ko
-import simulate as sim
-
-def logit_compare(id,n,b,p,q,p1s,new_seeds=False,procs=4,plot=False):
-    # since I'm parrallel computing, need to make sure I'm using different seeeds to generate random numbers
-    if new_seeds:
-        # these come from random.org, and should be truely random
-        # the function is a tad spotty due parsing a web page isntead of using the API, so may need to be run a few times
-        # also, it errors out when the request is rejected due to server load
-        seeds = sim.getRandomIntegers(num=10000,min=1,max=1000000000)
-    else:
-        # failing that, backup seeds gotten by good ol copy and paste
-        f = open('data/backup_seeds.txt','r')
-        seeds = [int(seed) for seed in f.read().split()]
-
-    # Load the genetic data
-    genes = np.genfromtxt('data/SNPdata.txt', delimiter=',')
-    # replace all the 2s with 1s
-    np.place(genes,genes!=0,1)
-
-    isl_data_logit = np.empty((2,len(p1s)))
-    gen_data_logit = np.empty((2,len(p1s)))
-
-    for i in range(len(p1s)):
-        p1 = p1s[i]
-        isl_data_logit[:,i] = sim.ko_test(procs=procs,
-                                          seeds=seeds[(i*b):(i*b+b)],
-                                          q=q,
-                                          knockoff='binary',
-                                          model='logit',
-                                          func=sim.genXy_binary_X_norm_beta,
-                                          n=n,
-                                          p1=p1,
-                                          pnull=p-p1,
-                                          beta_sd=1,
-                                          A_base_diag=-1,
-                                          A_sd=.2
-                                ).calc_fdr_power()
-        # save data
-        fi = open('data/logit_ising_' +id + '.npy','w')
-        np.save(fi,isl_data_logit)
-        fi.close()
-        print "Done Ising %d" % (i+1)
-
-        gen_data_logit[:,i] = sim.ko_test(procs=procs,
[…]

### 2015-07-18 masters_paper/compare.py
Added code to ensure linear independence

@@ -10,16 +10,81 @@ import matplotlib.pyplot as plt
 import knockoffGLM as ko
 import simulate as sim
 from multiprocessing import Pool
+from scipy.special import expit as invlogit
+from scipy.special import logit
+
+def cutoff(array,min=-50,max=50):
+    shape = array.shape
+    cut_array = np.min((max*np.ones(shape),array),axis=0)
+    cut_array = np.max((min*np.ones(shape),cut_array),axis=0)
+    return cut_array
+
+def make_X_ind(X,q=.02):
+    """ Add some random noise to make X linearly independent """
+    n,p = X.shape
+    v = np.min(nplin.svd(X)[1])
+    if v<1E-5:
+        for i in range(p):
+            v = np.min(nplin.svd(X[:,:(i+1)])[1])
+            while v<1E-5:
+                # Keep flipping a few bits until we have linear independence
+                X[:,i] = np.where(npran.binomial(1,q*np.ones(n)),1-X[:,i],X[:,i])
+                v = np.min(nplin.svd(X[:,:(i+1)])[1])
+                lmin = np.min(v)
+    return X
+
+def ising_X(p,n,A_base_diag=-1,A_sd=.2):
+    """ generate X from ising model """
+    A = npran.normal(0,.2,(p,p))-np.diag(A_base_diag*np.ones(p))
+    m,p = A.shape
+    ones   = np.ones((n,1))
+    X = np.empty((n,p))
+    for i in np.arange(0,p):
+        X[:,i] = npran.binomial(1,invlogit(np.dot(np.hstack((X[:,0:i],ones)),A[i,0:(i+1)])))
+    return make_X_ind(X)
+
+def given_X(p,n,data):
+    h,w = data.shape
+    X      = data[npran.choice(h,n),:][:,npran.choice(w,p)]
+    return make_X_ind(X)
+
+def norm_y(X,p1,sd=1,beta_sd=1):
+    """ Generate a normal Y from the X """
+    n,p = X.shape
+    X_1    = X[:,:p1]
+    beta   = npran.randn(p1)*beta_sd
+    if p1>0:
+        eta    = np.dot(X_1,beta)
+        y      = eta + npran.normal(0,sd,n)
+    else:
+        y      = npran.normal(0,sd,n)
+    return y
+
+def bern_y(X,p1,base_prob=.25,beta_sd=1):
+    n,p = X.shape
+    X_1    = X[:,:p1]
+    v = 0 
[…]

### 2015-07-18 masters_paper/compare.py
Perfect seperation issue should be fixed, missing correlation and high distortion still in the air

@@ -9,6 +9,8 @@ import scipy.stats as spstat1gt
 import matplotlib.pyplot as plt
 import knockoffGLM as ko
 import simulate as sim
+import statsmodels.api as sm
+from statsmodels.tools.sm_exceptions import PerfectSeparationError
 from multiprocessing import Pool
 from scipy.special import expit as invlogit
 from scipy.special import logit
@@ -19,18 +21,18 @@ def cutoff(array,min=-50,max=50):
     cut_array = np.max((min*np.ones(shape),cut_array),axis=0)
     return cut_array
 
-def make_X_ind(X,q=.02):
+def make_X_ind(X,q=.05):
     """ Add some random noise to make X linearly independent """
     n,p = X.shape
-    v = np.min(nplin.svd(X)[1])
-    if v<1E-5:
-        for i in range(p):
-            v = np.min(nplin.svd(X[:,:(i+1)])[1])
-            while v<1E-5:
+    for i in range(p):
+        while True:
+            try:
+                model = sm.GLM(self.X_orig[:,i],np.hstack((np.ones((n,1)),X[:,:i])),family=sm.families.Binomial()).fit()
+            except:
                 # Keep flipping a few bits until we have linear independence
                 X[:,i] = np.where(npran.binomial(1,q*np.ones(n)),1-X[:,i],X[:,i])
-                v = np.min(nplin.svd(X[:,:(i+1)])[1])
-                lmin = np.min(v)
+                v = np.min(nplin.svd(np.hstack((np.ones((n,1)),X[:,:(i+1)])))[1])
+                break
     return X
 
 def ising_X(p,n,A_base_diag=-1,A_sd=.2):
@@ -77,6 +79,7 @@ def bern_y(X,p1,base_prob=.25,beta_sd=1):
 def gen_logit(input):
     """ Testing function, all parameters are in a single tuple """
     seed,gen,p0,p1 = input
+    npran.seed(seed)
     if gen.lower() == "ising":
         X = ising_X(p1+p0,1000)
         y = bern_y(X,p1)
@@ -105,6 +108,7 @@ def gen_logit(input):
 def gen_lasso(input):
     """ Testing function, all parameters are in a single tuple """
     seed,gen,p0,p1 = input
+    npran.seed(seed)
     if gen.lower() == "ising":
         X = ising_X(p1+p0,1000)
[…]

### 2015-07-18 masters_paper/knockoffGLM.py

@@ -341,6 +341,8 @@ class knockoff_net(object):
         self.M_distortion = nplin.norm(self.M-np.dot(self.X_lrg.T,self.X_lrg)/self.n)/nplin.norm(self.M)
 
         self.emp_ko_corr= np.corrcoef(self.X_lrg,rowvar=0,bias=1)[:self.p,self.p:2*self.p][np.identity(self.p)==1]
+        if np.sum(np.isnan(self.emp_ko_corr))>0:
+            print "There were %d out of %d variables who had missing correlation" % (np.sum(np.isnan(self.emp_ko_corr)),self.p)
 
     def _vector_objective(self,a,X_fix,m):
         return np.mean(invlogit(np.dot(X_fix,a))[:,np.newaxis]*X_fix,axis=0) - m

### 2015-07-20 masters_paper/compare.py
Reducded A_sd again

@@ -35,7 +35,7 @@ def make_X_ind(X,q=.05):
                 break
     return X
 
-def ising_X(p,n,A_base_diag=-1,A_sd=.5):
+def ising_X(p,n,A_base_diag=-1,A_sd=.2):
     """ generate X from ising model """
     A = npran.normal(0,A_sd,(p,p))+np.diag(A_base_diag*np.ones(p))
     m,p = A.shape

### 2015-07-20 masters_paper/knockoffGLM.py

@@ -294,7 +294,7 @@ class knockoff_net(object):
                         a,
                         args=(X_fix,m),
                         method='anderson',
-                        options = {'maxiter':(i*2+150),'fatol':1E-5,'jac_options':{'M':25}}
+                        options = {'maxiter':(i*2+150),'fatol':1E-5,'jac_options':{'M':20}}
                         )
 
                 # update a to most recent estimate, even without convergence

### 2015-07-21 masters_paper/compare.py
Implemented approximate knockoff generation

@@ -43,12 +43,12 @@ def ising_X(p,n,A_base_diag=-1,A_sd=.2):
     X = np.empty((n,p))
     for i in np.arange(0,p):
         X[:,i] = npran.binomial(1,invlogit(np.dot(np.hstack((X[:,0:i],ones)),A[i,0:(i+1)])))
-    return make_X_ind(X)
+    return X
 
 def given_X(p,n,data):
     h,w = data.shape
     X      = data[npran.choice(h,n),:][:,npran.choice(w,p)]
-    return make_X_ind(X)
+    return X
 
 def norm_y(X,p1,sd=1,beta_sd=1):
     """ Generate a normal Y from the X """
@@ -78,7 +78,7 @@ def bern_y(X,p1,base_prob=.25,beta_sd=1):
 
 def generate(input):
     """ Testing function, all parameters are in a single tuple """
-    seed,gen,p0,p1,MCsize = input
+    seed,gen,p0,p1= input
     npran.seed(seed)
     if gen.lower() == "ising":
         X    = ising_X(p1+p0,1000)
@@ -93,8 +93,7 @@ def generate(input):
     # Logit
     model = ko.knockoff_logit(ybin,X,.2,
                               knockoff='binary',
-                              MCsize=MCsize,
-                              fresh_sim=True,
+                              method='approx',
                               intercept=True
                               )
     model.fit()
@@ -109,8 +108,7 @@ def generate(input):
     # LASSO
     bin_model = ko.knockoff_lasso(ynor,X,.2,
                               knockoff='binary',
-                              MCsize=MCsize,
-                              fresh_sim=True,
+                              method='approx',
                               intercept=True
                               )
     bin_model.fit(model.X_lrg)
@@ -130,7 +128,7 @@ def generate(input):
     with open('data/lasso_test_'+str(p0+p1)+'.txt','a') as f:
         f.write("%d\t%s\t%d\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\n" % (seed, gen, p1, model.M_distortion, np.mean(ko_corr), bin_FDR, bin_power, np.mean(ori_model.emp_ko_corr), ori_FDR, ori_power, corr))
 
-def batch_compare(b,p,p1s,gens,start_seed,MCsize=200000,procs=4):
+def batch_compare(b,p,p1s,gens,start_seed,procs=4):
     with open('data/backup_seeds.txt','r') as f:
[…]

### 2015-07-21 masters_paper/knockoffGLM.py

@@ -16,7 +16,7 @@ import scipy as sp
 import scipy.linalg as splin
 import scipy.stats as spstat
 import matplotlib.pyplot as plt
-#import cvxpy as cvx 
+import cvxpy as cvx 
 import statsmodels.api as sm
 from glmnet import LogisticNet, ElasticNet
 from sklearn.preprocessing import normalize
@@ -74,9 +74,17 @@ def solve_sdp(G):
     # Return as array, not as matrix.
     return np.asarray(s.value).flatten()
 
+def cutoff(array,min=0,max=1):
+    shape = array.shape
+    cut_array = np.min((max*np.ones(shape),array),axis=0)
+    cut_array = np.max((min*np.ones(shape),cut_array),axis=0)
+    return cut_array
+
+
+
 class knockoff_net(object):
     """ Parent class for knockoff lasso and knockoff logistic regression. Defines everything besides fitting the particular GlmNet object """
-    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,fresh_sim=True,pseudocount=0,intercept=False):
+    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,method='fresh_sim',pseudocount=0,intercept=False):
         self.y      = y
         self.X      = normalize(X.astype(float),norm='l2',axis=0)
         self.X_orig = X
@@ -87,7 +95,7 @@ class knockoff_net(object):
         self.randomize=False
         self.MCsize = MCsize   # how many values for Monte Carlo simulation
         self.zerotol = 1E-5
-        self.fresh_sim = fresh_sim 
+        self.method = method
         self.pseudocount=pseudocount
         self.intercept = intercept 
 
@@ -238,7 +246,7 @@ class knockoff_net(object):
         A = np.zeros((2*self.p,2*self.p))
 
         # Simulate fresh x based on the original data
-        if self.fresh_sim: 
+        if self.method == 'fresh_sim': 
             # Fit the upperhalf of A on the actual data, making this easier
             A[1,1] = logit(self.mu_lrg[0])
             for i in np.arange(0,self.p):
@@ -250,79 +258,134 @@ class knockoff_net(object):
             nMC = self.MCsize
         
         # just repeat X a bunch of times
-        else:
-            # Rather than simulate entirely new X at each stage, I will used a fixed set of X_1 ... X_(i-1)
[…]

### 2015-07-21 masters_paper/paper.tex
Fixed bug with X_lrg for original knockoffs

@@
     where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise.
+It is important to note that a intercept vector $\mathbf 1$ is not included in $X$.
 For the purpose of this paper, only the case where $2p\leq n$ will be considered, though the methods can be extended to $p\leq n$.
@@
 In particular, the matrix of knockoffs $\tilde X$ has the same internal correlation structure as $X$, the knockoff feature $\tilde X_j$ has the same correlation with other $X_i$ as $X_j$ does, but the correlation between $\tilde X_j$ and $X_j$ is minimized.
-In other words, where $\hat \Sigma := X^TX$,
-    \[ \tilde X^T\tilde X=\Sigma \textrm{ and } \tilde X^T X = \Sigma - \diag{\mathbf s}\]
-    for some vector $s\in \R^p$ such that, where $\diag{\Sigma}$ is the vector made from the main diagonal of $\Sigma$, $\diag{\Sigma} - \mathbf s$ is small.
+In other words, where $ G := X^TX$,
+    \[ \tilde X^T\tilde X=G \textrm{ and } \tilde X^T X = G - \diag{\mathbf s}\]
+    for some vector $s\in \R^p$ such that, where $\diag{G}$ is the vector made from the main diagonal of $G$, $\diag{G} - \mathbf s$ is small.
 Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$.
@@
 Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to have nonzero coefficients sooner than the knockoff features when the original is a valid predictor.
-On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $\Sigma_L=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
-This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $\Sigma_L$ will be unaltered, while $\E\left[X_L^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$.
+On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $G_L=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
+This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $G_L$ will be unaltered, while $\E\left[X_L^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$.
 \par
@@
     The original formulation of the method offers two similar methods of constructing knockoffs. Both of these will, by construction, have exactly the property that
-    \[ \tilde X^T\tilde X=\Sigma \textrm{ and } \tilde X^T X = \Sigma - \diag{\mathbf s}.\]
+    \[ \tilde X^T\tilde X=G \textrm{ and } \tilde X^T X = G - \diag{\mathbf s}.\]
     For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. Then, let the Gram matrix of $X_L$ is
-    \[ \Sigma_L= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} \Sigma & \Sigma - \diag{\mathbf s} \\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \]
-    $A$, the Schur complement of $\Sigma$ in $\Sigma_L$ can be calculated as 
-    \[ A = 2\,\diag{\mathbf s} - \diag{\mathbf s}\Sigma^{-1}\diag{\mathbf s} \]
-    For $\Sigma_L$ to exist, $\Sigma_L$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if\footnote{A fleshed out version of this argument can be found in the original knockoff paper}
-    \[ \diag{\mathbf s} \succeq 0  \; \textrm{ and } \; 2\Sigma - \diag{\mathbf s} \succeq 0 \]
+    \[ G_L= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right] \]
+    $A$, the Schur complement of $G$ in $G_L$ can be calculated as 
+    \[ A = 2\,\diag{\mathbf s} - \diag{\mathbf s}G^{-1}\diag{\mathbf s} \]
+    For $G_L$ to exist, $G_L$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if\footnote{A fleshed out version of this argument can be found in the original knockoff paper}
+    \[ \diag{\mathbf s} \succeq 0  \; \textrm{ and } \; 2G - \diag{\mathbf s} \succeq 0 \]
     Given this is true, $A$ can be factored as $A=C^TC$.
 Combining this with a satisfactory $\mathbf s$ and an orthonormal matrix $\tilde U\in\R^{n\times p}$ such that $\tilde U^T X = 0$, a $\tilde X$ fulfilling the desired properties can be calculated as
-    \[ \tilde X = X(I-\Sigma^{-1}\,\diag{\mathbf s}) + \tilde UC\]
-    As mentioned above, $\mathbf s$ should be chosen so as to make $\diag{\Sigma}-\mathbf s$ small.
-Since each $X_j$ has been normalized, this means that $\diag{\Sigma}=\mathbf 1$, so $\mathbf 1 - \mathbf s$ should be minimized in accordance with the restrictions on $\mathbf s$.
+    \[ \tilde X = X(I-G^{-1}\,\diag{\mathbf s}) + \tilde UC\]
+    As mentioned above, $\mathbf s$ should be chosen so as to make $\diag{G}-\mathbf s$ small.
+Since each $X_j$ has been normalized, this means that $\diag{G}=\mathbf 1$, so $\mathbf 1 - \mathbf s$ should be minimized in accordance with the restrictions on $\mathbf s$.
 The two methods differ in how $\mathbf s$ is chosen:
     \begin{itemize}
-        \item \textit{Equi-correlated knockoffs}: Each original features is set to have the same correlation with its knockoff by setting $\mathbf s = 2\min\{\lambda_{min}(\Sigma),1\}\mathbf 1$.
+        \item \textit{Equi-correlated knockoffs}: Each original features is set to have the same correlation with its knockoff by setting $\mathbf s = 2\min\{\lambda_{min}(G),1\}\mathbf 1$.
 For all simulations in this paper, this was the method used.
@@
                     minimize & $\|\mathbf1-\mathbf s\|_1$ \\
-                    subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2\Sigma $
+                    subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2G $
                 \end{tabular} 
@@
 This random variable will also abide by a relaxation of the original knockoff condition:
-    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=\Sigma \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = \Sigma - \diag{\mathbf s}\]
+    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=G \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = G - \diag{\mathbf s}\]
     This will ensure that, at the very least, these knockoffs are a suitable replacement for the original knockoffs for linear regression. \par
@@
 This theory provides the tools to generate random binary knockoffs.
-This is done is done sequentially as two steps.
-The first of these is to come up with the target correlation matrix $C$ for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$.
-
-the first of which is to come up with the target covariance matrix for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$.
+This is done is done sequentially as two steps, the first of which is to come up with the target covariance matrix $\Sigma_L$ for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$.
+This is extremely similar, though not identical, to the method with the original knockoffs, where $X_L^TX_L$ after normalization was targeted.
+However, much the same method as before can be used.
+Let the vector of empirical means of the columns of $X$ is $\mathbf m = \frac{1}{n}X^T\mathbf 1$ and the empirical covariance matrix of $X$ is $\Sigma = \frac{1}{n}X^TX - \mathbf{m}\mathbf{m}^T$.
 Much as before with the original knockoffs, both the equi-correlated and SDP method can be used to select  $\mathbf s$  such that $\diag{\Sigma}-\mathbf s$ is small and
         \[ \Sigma_L = \left[\begin{array}{cc}  \Sigma & \Sigma - \diag{\mathbf s }\\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \succeq 0 \]
-        Additionally though, the moment condition must hold for this to correspond to a proper random binary vector.
+    While with the original knockoffs, the result was small values for $X_i^T\tilde X_i$, now the result is low covariance between $x_i$ and $\tilde x_i$.\par
+        Additionally though, the moment condition must hold for $\Sigma$ to correspond to a proper random binary vector.
 If $\mathbf m_L = \E\left(\mathbf x_L\right) = [\mathbf m \; \mathbf m]$, then the desired cross moment matrix of the joint distribution is
@@
 Since the condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$, this implies that this method should normally be satisfied.
+This is the principle advantage of targeting low covariance between $X_i$ and $\tilde X_i$ rathervh.
 For the SDP method though, the restriction can be built directly into the optimization problem, yielding the updated problem

### 2015-07-21 masters_paper/knockoffGLM.py

@@ -123,7 +123,7 @@ class knockoff_net(object):
         C_U,C_d,C_V = nplin.svd(2*np.diag(s) - (self.s * G_inv.T).T * self.s)
         C_d[C_d < 0] = 0
         X_ko = self.X - np.dot(self.X,G_inv*s) + np.dot(U_perp*np.sqrt(C_d),C_V)
-        self.X_lrg = np.concatenate((self.X_orig,X_ko), axis=1)
+        self.X_lrg = np.concatenate((self.X,X_ko), axis=1)
 
     def _create_equicor(self):
         """ Creates the equal correlation knockoff of X"""
@@ -146,7 +146,7 @@ class knockoff_net(object):
         s_diff = 2*self.s - (self.s/d)**2
         s_diff[s_diff<0]=0 # can be negative due to numerical error
         X_ko = np.dot(U*(d-self.s/d) + U_perp*(np.sqrt(s_diff)) , V)
-        self.X_lrg = np.concatenate((self.X_orig,X_ko), axis=1)
+        self.X_lrg = np.concatenate((self.X,X_ko), axis=1)
 
     def _original_knockoff(self):
         """ Creates the original style knockoffs"""
@@ -274,15 +274,18 @@ class knockoff_net(object):
         ###################################################
 
         if self.method== 'approx':
+            covinv   = np.diag(self.M - np.outer(self.mu_lrg,self.mu_lrg))**-1
+            upwt = 1
             X_tmp = np.hstack((self.X_orig,np.ones((self.n,1)))).T/self.n
             for i in np.arange(self.p,2*self.p):
                 m = np.append(self.M[i,0:self.p],self.M[i,i])
+                wt = np.diag(np.append(np.append(np.ones(i-self.p),upwt),np.ones(2*self.p-i)))*np.diag(np.append(covinv[:self.p],covinv[i]))
          
                 ps = cvx.Variable(self.n)
-                objective = cvx.Minimize(cvx.norm(X_tmp * ps - m ,2))
+                objective = cvx.Minimize(cvx.norm( wt*(X_tmp * ps - m) ,2))
                 constraints = [0 <= ps, ps<=1]
                 prob = cvx.Problem(objective,constraints)
-                prob.solve()
+                prob.solve(solver=cvx.SCS,max_iters=100)
          
                 X_fix = np.hstack((X_fix,ps.value))
          
@@ -292,20 +295,19 @@ class knockoff_net(object):
                     X_tmp = X_fix
[…]

### 2015-07-22 masters_paper/knockoffGLM.py
Added plots

@@ -319,33 +319,18 @@ class knockoff_net(object):
          
          
             for i in np.arange(self.p,2*self.p):
-                # m are the cross moments we are trying to fit
-                m = self.M[i,0:(i+1)]
-                # X_tmp is the list of (x_k,1) vectors
-                #X_fix     = np.hstack((X_fix,np.ones((nMC,1))))
-         
-                #X_tmp_out = X_fix[:,:,np.newaxis]*X_fix[:,np.newaxis,:]
-         
                 # Now, the Newton-Raphson steps
                 # If the hessian becomes singular, we will relax the cross moment requirements, as described in Schafer 5.1 point 2
                 # the idea is that the problem is relaxed until X_i is independent of all prior vars
                 # as por increases, covariance drops
                 # a is the row we are adding to A. Initialize with values as if independent all other vars
+
                 a = np.append(np.zeros(i),logit(self.mu_lrg[i]))
-         
-         
-                X_fix     = np.hstack((X_fix,np.ones((self.n,1))))
+                X_fix     = np.hstack((X_fix,np.ones((nMC,1))))
          
                 for por in por_seq:
                     # m are the cross moments we are trying to fit
                     m = (1-por)*self.M[i,0:(i+1)] + por*self.M[i,i]*np.append(np.diag(self.M)[0:i],1)
-         
-                    # Linear program
-                    ps = cvx.Variable(self.n)
-                    objective = cvx.Minimize(cvx.norm(ps-self.M[i,i],1))
-                    constraints = [0 <= ps, ps<=1, (X_fix.T/self.n) * ps == m]
-                    prob = cvx.Problem(objective,constraints)
-                    prob.solve()
          
                     # Minimize the actual difference vector
                     opt = root(self._vector_objective,

### 2015-07-22 masters_paper/compare.py
Fixed fresh_sim, updated compare

@@ -78,7 +78,7 @@ def bern_y(X,p1,base_prob=.25,beta_sd=1):
 
 def generate(input):
     """ Testing function, all parameters are in a single tuple """
-    seed,gen,p0,p1= input
+    seed,gen,p0,p1,method,MCsize= input
     npran.seed(seed)
     if gen.lower() == "ising":
         X    = ising_X(p1+p0,1000)
@@ -91,44 +91,53 @@ def generate(input):
     ynor = norm_y(X,p1)
 
     # Logit
-    model = ko.knockoff_logit(ybin,X,.2,
+    bin_logit = ko.knockoff_logit(ybin,X,.2,
                               knockoff='binary',
-                              method='approx',
+                              method=method,
+                              MCsize=MCsize,
                               intercept=True
                               )
-    model.fit()
-    trueS = (np.arange(p0+p1)<p1).astype(int)
-    FDR   = np.dot(model.S,1-trueS)/max(np.sum(model.S),1)
-    power = np.dot(model.S,trueS)  /max(p1,1)
-    ko_corr = [cor for cor in model.emp_ko_corr if not np.isnan(cor)]
-
-    with open('data/logit_test_'+str(p0+p1)+'.txt','a') as f:
-        f.write("%d\t%s\t%d\t%.5f\t%.5f\t%.5f\t%.5f\n" % (seed,gen,p1,model.M_distortion,np.mean(ko_corr),FDR,power))
-
-    # LASSO
-    bin_model = ko.knockoff_lasso(ynor,X,.2,
-                              knockoff='binary',
-                              method='approx',
-                              intercept=True
-                              )
-    bin_model.fit(model.X_lrg)
-    ori_model = ko.knockoff_lasso(ynor,X,.2,
+    bin_logit.fit()
+    ori_logit = ko.knockoff_logit(ybin,X,.2,
                               knockoff='original',
                               intercept=False
                               )
-    ori_model.fit()
+    ori_logit.fit()
     trueS = (np.arange(p0+p1)<p1).astype(int)
-    bin_FDR   = np.dot(bin_model.S,1-trueS)/max(np.sum(bin_model.S),1)
[…]

### 2015-07-22 masters_paper/knockoffGLM.py

@@ -344,23 +344,20 @@ class knockoff_net(object):
                     a = opt.x
          
                     # Stop once optimal has been reached
-                    if prob.status=="optimal":
-                        p = ps.value
+                    if opt.success:
                         self.por = np.append(self.por,por)
                         if por>0:
                             print "Variable %d relaxed by tau=%.2f" % (i-self.p+1,por)
-                        print "Variable %d success" % (i-self.p+1)
                         break
          
-                if not prob.status=="optimal":
+                if not opt.success:
                     a = np.append(np.zeros(i),logit(self.mu_lrg[i]))
-                    p = np.ones(self.n)*self.M[i,i]
                     self.por = np.append(self.por,1)
                     print "Variable %d fully relaxed" % (i-self.p+1)
          
                 # put a into A matrix, draw X_i for 'fixed' matrix, update X_out_fix
                 A[i,0:(i+1)] = a
-                X_fix[:,-1] = p.squeeze()
+                X_fix[:,-1] = npran.binomial(1,invlogit(np.dot(X_fix,a)))
 
         # hang onto A
         self.A = A

### 2015-07-23 masters_paper/paper.tex
Small tweaks
Almost done updating section 3

@@
             To do so, the Jacobian matrix is calculated as 
-        \[ J f\left(\mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
+        \[ J\left( f\left(\mathbf a_i\right)\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
             With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive.
@@
             and
-            \[J f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
+            \[J\left( f\left(\mathbf a_i\right)\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
             can be used. \par
-            Though, in theory, $A$ should always exist, in practice numerical issues may compound to the point that the Newton-Raphson method won't converge.
-In this case, one can solve instead for $\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
+            Though, in theory, $A$ should always exist, in practice limited accuracy may dictate that Newton-Raphson method won't converge.
+In the simulation method in particular, if $K$ is not large enough, the equations can be very difficult or impossible to solve.
+When the intial problem can't be solved, besides increasing the simulation size or numerical accuracy, one can instead solve the relaxed problem  $f(\mathbf a)=\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
     \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T \]
@@
     While with the original knockoffs, the result was small values for $X_i^T\tilde X_i$, now the result is low covariance between $x_i$ and $\tilde x_i$.\par
-        Additionally though, the moment condition must hold for $\Sigma$ to correspond to a proper random binary vector.
+    Additionally though, the moment condition must hold for $\Sigma_L$ to correspond to a proper random binary vector.
 If $\mathbf m_L = \E\left(\mathbf x_L\right) = [\mathbf m \; \mathbf m]$, then the desired cross moment matrix of the joint distribution is
@@
 Since the condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$, this implies that this method should normally be satisfied.
-This is the principle advantage of targeting low covariance between $X_i$ and $\tilde X_i$ rathervh.
-For the SDP method though, the restriction can be built directly into the optimization problem, yielding the updated problem
+This is the principle advantage minimizing the $\textrm{cor}(x_i,\tilde x_i)$ rather than $\E[x_i\tilde x_i]$.
+For the SDP method though, the restriction can be built directly into the optimization problem.
+If $\sigma = \diag{\Sigma}^{\frac{1}{2}}$ is the vector of standard deviations, the updated problem is
             \begin{center}
                 \begin{tabular}{r l}
-                    minimize & $\|\mathbf 1-\diag{\Sigma}^{-1}\mathbf s\|_1$ \\
+                    minimize & $\|\mathbf 1-\diag{\sigma}^{-1}\mathbf s\|_1$ \\
                     subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2\Sigma $ \\
@@
             \end{center} 
-        replacing $\mathbf s$ with $\diag{\Sigma}^{-1}\mathbf s$ in the objective is to account for not normalizing $X$, as is usually done with the original knockoffs. \par
-        Once a proper cross moment matrix $M_L$ has been selected, one could simply apply the Newton-Raphson method to fit a corresponding $A$ matrix, and then generate $\mathbf{ \tilde x}$ sequentially conditioned on $x$, as $\tilde x_i \st \mathbf x,\tilde x_1,\ldots,\tilde x_{i-1}$.
-However, there are a few computational wrinkles which can be chosen to speed up the process:
+            the factor of $\diag{\sigma}^{-1}$ in the objective is to account for the $x_i$ not being normalized, as was the case with the original knockoffs. \par
+\subsubsection{Fitting via $\mu$-conditionals}
+        Once a proper cross moment matrix $M_L$ has been selected, the most obvious approach to generate knockoffs is to apply the Newton-Raphson method to fit the matching $A$ matrix.
+This can be used to generate $\mathbf{ \tilde x}$ sequentially conditioned on $x$, as $\tilde x_i \st \mathbf x,\tilde x_1,\ldots,\tilde x_{i-1}$.
+However, there are a few variations of this process which are preferable depending on the situation.
         \begin{enumerate}
@@
 The second half of the matrix still must be fit via Newton-Raphson.
-            \item One can avoid fitting the first half of the $A$ matrix entirely though if performing Newton-Raphoson with by using the actual observed values in $X$ to simulate $\mathbf x$.
+            \item One can avoid fitting the first half of the $A$ matrix entirely though if performing Newton-Raphoson with by using the actual observed values in $X$ to simulate the first half of $\mathbf x_L$.
 This can be done by either bootstrapping rows from $X$ or using some number of copies of $X$ exactly.
@@
 The downside is that, since these values won't be from the Ising model, the theoretical guarantee of a suitable $A$ may no longer hold.
+As well, if $n$ isn't big enough, the equations will quickly become difficult to solve, since the observed samples will provide a noisy and incomplete sample of the probability space.
             \item As $A$ is being fit, even the parts of $\mathbf x_{-i}$ corresponding to knockoffs need not be redrawn for each Newton-Raphson iteration or each successive $i$.
+To fit $\mathbf a$ for feature $i$, the $\mathbf x_{-i}^{(k)}$ will remain constant; then, once $\mathbf a$ has been fit, the $x_i^{(k)}$ need only be drawn once.
 This saves computation, and the multiplication
             \[ \left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [\mathbf x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
-                doesn't need to be recomputed for each Newton-Raphson iteration, and the upper left portion, $\mathbf x_{-i}^{(k)}[\mathbf{x}_{-i}^{(k)}]^T$ doesn't have to be recomputed for successive $i$.
+                needs only be partially recalculated for each Newton-Raphson iteration since the upper left portion, $\mathbf x_{-i}^{(k)}[\mathbf{x}_{-i}^{(k)}]^T$, won't change.
 This may result in slight deviations in any particular draw compounding into a more severe error over time.
+            \item As $p$ grows large, and $K$ necessarily along with it, the cost of calculating the Jacobian matrix grows far quicker than the cost of calculating $f(\mathbf a)$.
+This is since it has $O(Kp^2)$ complexity compared to $O(Kp)$ complexity for $f(\mathbf a)$.
+This can quickly become prohibitive.
+One might instead estimate the full Jacobian based on first differences, but in practice Anderson Mixing, a quasi-Newtonian method, works well.
+In general, quasi-Newtonian methods perform Newton-Raphson iterations, but replace the full inverted Jacobian with an approximation.
+{\bf MORE DETAILS}
         \end{enumerate}
-        Combining the second and third trick leads to the final formulation.
-Rather then $K$ simulations of $\mathbf x$, the simulation is started with the matrix $X_F$, which is $X$ stacked up $m$ times to get $n\times m=K'\geq K$ rows.
-The following process is then repeated for $p<i\leq 2p$:
+        In practice, it seemed that combining the first, third, and fourth variation was the most successful technique.
+However, should $n$ be sufficiently large, the second variation also saves a lot of time.
 
-         \begin{itemize}
-             \item Where $\mathbf x_F^{(k)}$ is the $k$th row of $X_F$, the row $\mathbf a_i$ is fit by Newton-Raphson iterations with 
-             \[f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right] \]
-             and
-             \[J f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
-             \item If the iterations won't converge, the relaxed equation $f(\mathbf a_i)=m_i^*(\tau)$ is attempted instead for increasing values of $\tau$ until it converges.
-             \item After $\mathbf a_i$ is fit, a new column $X_i$ is drawn as independent Bernoulli with probability vector $\mu(X_F \mathbf a_i)$.   
-             \item $X_F$ is updated to 
-             \[ X_F = [X_F \; X_i] \]
-         \end{itemize}
-         At the end, the any $n$ rows of $X_F$ corresponding to the rows of $X$ can be taken as $X_L$. Any such set should have the desired distribution for $\tilde X$.
+\subsubsection{Fitting via Quadratic Programing}
+
+        Random binary knockoffs which produce the desired cross moment matrix $M_L$ need not arise from a generative model for $\mathbf{\tilde x}$ however.
+It is sufficient to have a random matrix $\tilde X \in \{0,1\}^{n\times p}$ such that
+        \[ \E_{\tilde X}(X^T \tilde X) = \Sigma - \diag{s} \quad \& \quad \E_{\tilde X}(\tilde X^T \tilde X) = \Sigma\]
+        The simplest way to do this is based on a probability matrix $P\in (0,1)^{n\times p}$ where each $X_{ij}$ is drawn independently as $X_{ij}\sim\textrm{Bernoulli}(P_{ij})$.
+This requirement can be stated as the quadratic feasibility problem
+            \begin{center}
+                \begin{tabular}{r l}
+                    exists     & $P$ \\
+                    subject to & $X^TP=\Sigma-\diag{s} $ \\
+                               & $P^TP=\Sigma $ \\
+                               & $0 \leq P \leq 1$
+                \end{tabular} 
+            \end{center}
+        If such a matrix exists and can be found, then $\tilde X$ can easily be generated.
+However, the constraints likely result in infeasibility much of the time.
+Instead, one can try to minimize how far the expected moments given $P$ would be from those desired for $M_L$.
+Ideally, this would take the form
+            \begin{center}
+                \begin{tabular}{r l}
+                    minimize     & $\|X^TP-(\Sigma-\diag{s})\|_{fro}^2 + \|P^TP-\Sigma\|_{fro}^2 $\\
+                    subject to   & $ \mathbf 1^T P = \mathbf m_L $ \\
+                                 & $0 \leq P \leq 1$
+                \end{tabular} 
+            \end{center}
+            Here, the marginal distribution of $\tilde X_i$ will be the same as $X_i$, and the crossmoments shouldn't deviate to far from $M_L$.
+This problem can be stated as the quadratic program via the introduction of slack variables:
+            \begin{center}
+                \begin{tabular}{r l}
+                    minimize     & $\|W\|_{fro}^2 + \|V\|_{fro}^2 $ \\
+                    subject to   & $ X^TP-(\Sigma-\diag{s})\leq W $ \\
+                                 & $ P^TP-\Sigma \leq V $ \\
+                                 & $ \mathbf 1^T P = \mathbf m_L $ \\
+                                 & $0 \leq P \leq 1$
+                \end{tabular} 
+            \end{center}
+            Though there are nice tools and theory for solving quadratic programs, $n\times p$ would have to be fairly small for this optimization problem to be computationally reasonable in practice.
+Ideally, there would be a method to breakdown this problem into a series of smaller optimization problems which would yield the same solution.
+A optimistic numerical analyst would hope that iterating through the columns, optimizing each in individually, would achieve this.
 

### 2015-07-25 masters_paper/compare.py
Allows X to be stacked

@@ -78,7 +78,7 @@ def bern_y(X,p1,base_prob=.25,beta_sd=1):
 
 def generate(input):
     """ Testing function, all parameters are in a single tuple """
-    seed,gen,p0,p1,method,MCsize= input
+    seed,gen,p0,p1,method,MCsize,stacks= input
     npran.seed(seed)
     if gen.lower() == "ising":
         X    = ising_X(p1+p0,1000)
@@ -95,7 +95,8 @@ def generate(input):
                               knockoff='binary',
                               method=method,
                               MCsize=MCsize,
-                              intercept=True
+                              intercept=True,
+                              stacks = stacks
                               )
     bin_logit.fit()
     ori_logit = ko.knockoff_logit(ybin,X,.2,
@@ -119,7 +120,8 @@ def generate(input):
                               knockoff='binary',
                               method=method,
                               MCsize=MCsize,
-                              intercept=True
+                              intercept=True,
+                              stacks = stacks
                               )
     bin_lasso.fit(bin_logit.X_lrg)
     ori_lasso = ko.knockoff_lasso(ynor,X,.2,
@@ -137,7 +139,7 @@ def generate(input):
     with open('data/lasso_test_'+str(p0+p1)+'.txt','a') as f:
         f.write("%d\t%s\t%d\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\n" % (seed, gen, p1, bin_logit.M_distortion, np.mean(ko_corr), bin_FDR, bin_power, np.mean(ori_lasso.emp_ko_corr), ori_FDR, ori_power, corr))
 
-def batch_compare(b,p,p1s,gens,start_seed,method='fresh_sim',MCsize=100000,procs=4):
+def batch_compare(b,p,p1s,gens,start_seed,method='fresh_sim',MCsize=100000,stacks=1,procs=4):
     with open('data/backup_seeds.txt','r') as f:
         seeds = [int(seed) for seed in f.read().split()]
     inputs = []
@@ -145,7 +147,7 @@ def batch_compare(b,p,p1s,gens,start_seed,method='fresh_sim',MCsize=100000,procs
     for b in range(b):
         for p1 in p1s:
             for gen in gens:
-                inputs.append((seeds[i+start_seed],gen,p-p1,p1,method,MCsize))
+                inputs.append((seeds[i+start_seed],gen,p-p1,p1,method,MCsize,stacks))
                 i += 1
 
     pool = Pool(processes=procs)

### 2015-07-25 masters_paper/knockoffGLM.py

@@ -80,11 +80,9 @@ def cutoff(array,min=0,max=1):
     cut_array = np.max((min*np.ones(shape),cut_array),axis=0)
     return cut_array
 
-
-
 class knockoff_net(object):
     """ Parent class for knockoff lasso and knockoff logistic regression. Defines everything besides fitting the particular GlmNet object """
-    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,method='fresh_sim',pseudocount=0,intercept=False):
+    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,method='fresh_sim',pseudocount=0,intercept=False,stacks=1):
         self.y      = y
         self.X      = normalize(X.astype(float),norm='l2',axis=0)
         self.X_orig = X
@@ -97,6 +95,7 @@ class knockoff_net(object):
         self.zerotol = 1E-5
         self.method = method
         self.pseudocount=pseudocount
+        self.stacks = stacks
         self.intercept = intercept 
 
     def _create_SDP(self):
@@ -370,7 +369,7 @@ class knockoff_net(object):
 
         # If we freshly simulated x, we need to draw ~x based on x
         if self.method=='fresh_sim': 
-            self.X_lrg = np.hstack((self.X_orig,np.empty((self.n,self.p))))
+            self.X_lrg = np.hstack((np.repeat(self.X_orig,self.stacks,0),np.empty((self.n*self.stacks,self.p))))
             for i in np.arange(self.p,2*self.p):
                 # need to make sure the knockoff isn't uniformly 0 or 1
                 count = 0

### 2015-07-26 masters_paper/paper.tex
more paper

@@
 \newtheorem{example}[theorem]{Example}
+\DeclareMathOperator*{\argmin}{arg\min}
 
@@
 \footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} Recall, for a given $\lambda$, the estimated coefficient from the Lasso regression will be
-    \[ \beta(\lambda) = \textrm{argmin}_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]
+    \[ \beta(\lambda) = \argmin_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]
     In the general case, the coefficient for one feature in a Lasso regression being first nonzero for a larger value of $\lambda$ than another feature is an indication that the first feature is a stronger predictor of the outcome than the second.
@@
 Here, where $l(\beta\st X, \tilde X)$ is the log likelihood of coefficients $\beta$ given the model, the $L1$ regularized regression model will have estimated coefficient vector
-    \[ \beta(\lambda) = \textrm{argmin}_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
+    \[ \beta(\lambda) = \argmin_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
     The $Z_j$ and $\tilde Z_j$ are then, once again, the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before.
@@
 This equation $f(\mathbf a_i)=\mathbf{m}_i$ can be solved by successive Newton-Raphson iterations defined by
-            \[\mathbf a_i^{(k+1)} = \left[H f\left(\mathbf a_i^{(k)}\right)\right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
-            To do so, the Jacobian matrix is calculated as 
-        \[ J\left( f\left(\mathbf a_i\right)\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
+            \[\mathbf a_i^{(k+1)} = \mathbf a_i^{(k)} - \left[J\left(\mathbf a_i^{(k)}\right) \right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
+            where $J$ is the Jacobian of $f$. To do so, the Jacobian matrix is calculated as 
+        \[ J\left( \mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
             With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive.
@@
             and
-            \[J\left( f\left(\mathbf a_i\right)\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
+            \[J\left( \mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
             can be used. \par
@@
 One might instead estimate the full Jacobian based on first differences, but in practice Anderson Mixing, a quasi-Newtonian method, works well.
-In general, quasi-Newtonian methods perform Newton-Raphson iterations, but replace the full inverted Jacobian with an approximation.
-{\bf MORE DETAILS}
+In general, quasi-Newtonian methods perform Newton-Raphson iterations but replace the full inverted Jacobian at each itteration $k$ with an approximation $G_k$.
+Thus, to solve for a variable $\mathbf z$ such that $f(\mathbf z)=\mathbf 0$, the best approximation for the zero is updated as $\mathbf z_{k+1}=\mathbf z_k - G_k f(\mathbf{z}_k)$.
+\par
+    Anderson mixing does this implicitly, with no extra recalculation of $f$, by approximating $f$ with the affine space spanning the previous $m$ evaluations of $f$, at $\mathbf z_{k}$ through $\mathbf z_{k-m+1}$.
+The best approximation for $f(\mathbf z)=0$ in this space is defined by
+    \[ \mathbf \alpha = \argmin_{\|\mathbf w\|_1=1} \|\sum_{i=1}^m f(\mathbf z_{k-i+1}) w_i \|_2 \]
+    This leads to the updated approximation
+    \[ \mathbf z_{k+1} = \sum_{i=1}^{m} w_i \mathbf z_{k-i+1} \]
+    There is no explicit calculation of $G_k$ needed here, and to do so in practice would be slower. However, it's presence can be seen through an equivalent formulation. Let 
+    \[ \mathcal F = [(f(\mathbf z_k) - f(\mathbf z_{k-1})) \ldots (f(\mathbf z_k) - f(\mathbf z_{k-m}))] \]
+    and
+    \[ \mathcal Z = [(\mathbf z_k - \mathbf z_{k-1})\ldots (\mathbf z_k - \mathbf z_{k-m})] \]
+    This means $\mathbf \alpha$ is now
+    \[ \mathbf \alpha \; = \;\argmin_{\|\mathbf w\|_1=1} \|f(\mathbf z_{k}) - \mathcal  F \mathbf w\|_2\; = \;(\mathcal F^T \mathcal F)^{-1}\mathcal F^T\mathbf z_k \]
+    so 
+    \[ \mathbf z_{k+1} \;= \;\mathbf z_{k+1} - \mathcal Z \alpha\; = \;z_{k+1} + \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T\mathbf z_k \]
+    Which yields the approximated inverse Jacobian $G_k= \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T$
         \end{enumerate}
@@
                     minimize     & $\|W\|_{fro}^2 + \|V\|_{fro}^2 $ \\
-                    subject to   & $ X^TP-(\Sigma-\diag{s})\leq W $ \\
-                                 & $ P^TP-\Sigma \leq V $ \\
+                    subject to   & $ -W \leq X^TP-(\Sigma-\diag{s})\leq W $ \\
+                                 & $ -V \leq P^TP-\Sigma \leq V $ \\
                                  & $ \mathbf 1^T P = \mathbf m_L $ \\
@@
 Ideally, there would be a method to breakdown this problem into a series of smaller optimization problems which would yield the same solution.
-A optimistic numerical analyst would hope that iterating through the columns, optimizing each in individually, would achieve this.
+An optimist might hope that iterating through the columns of $P$, optimizing each to hit its desired crossmoments individually conditioned on the estimate of the rest of $P$, would achieve this.
+In practice, this does seem to yield reasonable results, and is far quicker than other methods for large $p$.In particular, this method seemed to work well:
+    \begin{enumerate}
+        \item Optimize initial $P_i$ as such that $X_j^T P_i = M_ij$ if $i\neq j$ and $X_i^T P_i = \sigma_i(1-s)$.   
+            \begin{center}
+                \begin{tabular}{r l}
+                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_ij)^2$ \\
+                    subject to   & $ \frac{1}{n}\mathbf 1^T P_i = M_{ii} $ \\
+                                 & $0 \leq P_i \leq 1$
+                \end{tabular} 
+            \end{center}
+        \item Draw initial $\tilde X_i$ as independent Bernoulli with probability $P_i$.
+        \item Iterate through $1\leq i \leq p$, fitting new $P_i$ such that 
+            \begin{center}
+                \begin{tabular}{r l}
+                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_ij)^2 + \sum_{i\neq j}(\frac{1}{n}\tilde X_j^T P_i - M_ij)^2$  \\
+                    subject to   & $ \frac{1}{n}\mathbf 1^T P_i = M_{ii} $ \\
+                                 & $0 \leq P_i \leq 1$
+                \end{tabular} 
+            \end{center}
+    \end{enumerate}
 

### 2015-07-27 masters_paper/paper.tex
Onto last two sections

@@
             \item As $p$ grows large, and $K$ necessarily along with it, the cost of calculating the Jacobian matrix grows far quicker than the cost of calculating $f(\mathbf a)$.
-This is since it has $O(Kp^2)$ complexity compared to $O(Kp)$ complexity for $f(\mathbf a)$.
-This can quickly become prohibitive.
-One might instead estimate the full Jacobian based on first differences, but in practice Anderson Mixing, a quasi-Newtonian method, works well.
-In general, quasi-Newtonian methods perform Newton-Raphson iterations but replace the full inverted Jacobian at each itteration $k$ with an approximation $G_k$.
-Thus, to solve for a variable $\mathbf z$ such that $f(\mathbf z)=\mathbf 0$, the best approximation for the zero is updated as $\mathbf z_{k+1}=\mathbf z_k - G_k f(\mathbf{z}_k)$.
+This is since it has $O(Kp^2)$ complexity compared to $O(Kp)$ complexity for $f(\mathbf a)$, quickly becoming prohibitive.
+One might instead estimate the full Jacobian based on first differences, but in practice an even simpler method called Anderson mixing, which avoids additional computation of $f(\mathbf a)$, works well.
+Anderson mixing is one of several related quasi-Newtonian methods, which performs Newton-Raphson iteration $k$ with an approximation of the inverted inverted Jacobian labeled $G_k$.
+Thus, to solve for a variable $\mathbf z$ such that $f(\mathbf z)=\mathbf 0$, the best approximation for the zero is updated as $\mathbf z_{k+1}=\mathbf z_k - G_k f(\mathbf{z}_k)$ at each iteration.
 \par
-    Anderson mixing does this implicitly, with no extra recalculation of $f$, by approximating $f$ with the affine space spanning the previous $m$ evaluations of $f$, at $\mathbf z_{k}$ through $\mathbf z_{k-m+1}$.
-The best approximation for $f(\mathbf z)=0$ in this space is defined by
-    \[ \mathbf \alpha = \argmin_{\|\mathbf w\|_1=1} \|\sum_{i=1}^m f(\mathbf z_{k-i+1}) w_i \|_2 \]
+    Anderson mixing does this implicitly, with no extra recalculation of $f$, by approximating $f$ with the affine space spanning the previous $m$ evaluations of $f$.
+These values are $f(\mathbf z_{k})$ through $f(\mathbf z_{k-m+1})$.
+The best approximation for $f(\mathbf z)=0$ in the affine space which spans them is defined by
+    \[ \mathbf \alpha = \argmin_{\mathbf w:\, \sum_{i=1}^m w_i=1} \left\|\sum_{i=1}^m  w_i f(\mathbf z_{k-i+1}) \right\|_2 \]
     This leads to the updated approximation
-    \[ \mathbf z_{k+1} = \sum_{i=1}^{m} w_i \mathbf z_{k-i+1} \]
-    There is no explicit calculation of $G_k$ needed here, and to do so in practice would be slower. However, it's presence can be seen through an equivalent formulation. Let 
-    \[ \mathcal F = [(f(\mathbf z_k) - f(\mathbf z_{k-1})) \ldots (f(\mathbf z_k) - f(\mathbf z_{k-m}))] \]
+    \[ \mathbf z_{k+1} = \sum_{i=1}^{m} \alpha_{i} \mathbf z_{k-i+1} \]
+    $G_k$ doesn't need to be calculated here, and the quickest implementation finds $\alpha$ without calculating it.
+However, it's presence can be seen through an equivalent formulation.
+Let $\Delta \mathbf f_i = f(\mathbf z_k) - f(\mathbf z_{k-i})$, $\Delta \mathbf z_i = \mathbf z_k - \mathbf z_{k-i}$,
+    \[ \mathcal F = [\Delta \mathbf f_1 \ldots \Delta \mathbf f_{m-1}] \]
     and
-    \[ \mathcal Z = [(\mathbf z_k - \mathbf z_{k-1})\ldots (\mathbf z_k - \mathbf z_{k-m})] \]
-    This means $\mathbf \alpha$ is now
-    \[ \mathbf \alpha \; = \;\argmin_{\|\mathbf w\|_1=1} \|f(\mathbf z_{k}) - \mathcal  F \mathbf w\|_2\; = \;(\mathcal F^T \mathcal F)^{-1}\mathcal F^T\mathbf z_k \]
+    \[ \mathcal Z = [\Delta \mathbf z_1 \ldots \Delta \mathbf z_{m-1}] \]
+    This means that $\mathbf\alpha$ has been replaced by $\mathbf\theta$ such that
+    \[ \mathbf \theta \; = \;\argmin_{\mathbf w} \left\|f(\mathbf z_{k}) - \mathcal  F \mathbf w\right\|_2\; = \;(\mathcal F^T \mathcal F)^{-1}\mathcal F^Tf(\mathbf z_k) \]
     so 
-    \[ \mathbf z_{k+1} \;= \;\mathbf z_{k+1} - \mathcal Z \alpha\; = \;z_{k+1} + \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T\mathbf z_k \]
+    \[ \mathbf z_{k+1} \;= \;\mathbf z_{k} - \mathcal Z \mathbf \theta\; = \; \mathbf z_{k} - \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T f(\mathbf z_k) \]
     Which yields the approximated inverse Jacobian $G_k= \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T$
         \end{enumerate}
-        In practice, it seemed that combining the first, third, and fourth variation was the most successful technique.
-However, should $n$ be sufficiently large, the second variation also saves a lot of time.
+        In practice, it seemed that combining the first, third, and fourth of these variations leads to the most successful technique.
+However, should $n$ be large enough to provide a good approximation of the probability space, the second variation also saves a lot of time.
 
@@
 Instead, one can try to minimize how far the expected moments given $P$ would be from those desired for $M_L$.
-Ideally, this would take the form
+This would take the form
             \begin{center}
@@
 An optimist might hope that iterating through the columns of $P$, optimizing each to hit its desired crossmoments individually conditioned on the estimate of the rest of $P$, would achieve this.
-In practice, this does seem to yield reasonable results, and is far quicker than other methods for large $p$.In particular, this method seemed to work well:
+In practice, this does seem to yield reasonable results, and is far quicker than other methods for large $p$.
+In particular, the following method seemed to work well:
     \begin{enumerate}
-        \item Optimize initial $P_i$ as such that $X_j^T P_i = M_ij$ if $i\neq j$ and $X_i^T P_i = \sigma_i(1-s)$.   
+        \item Optimize initial $P_i$ as such that $X_j^T P_i = M_{ij}$ if $i\neq j$ and $X_i^T P_i = \sigma_i(1-s)$.   
             \begin{center}
                 \begin{tabular}{r l}
-                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_ij)^2$ \\
+                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_{ij})^2$ \\
                     subject to   & $ \frac{1}{n}\mathbf 1^T P_i = M_{ii} $ \\
@@
                 \begin{tabular}{r l}
-                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_ij)^2 + \sum_{i\neq j}(\frac{1}{n}\tilde X_j^T P_i - M_ij)^2$  \\
+                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_{ij})^2 + \sum_{i\neq j}(\frac{1}{n}\tilde X_j^T P_i - M_{ij})^2$  \\
                     subject to   & $ \frac{1}{n}\mathbf 1^T P_i = M_{ii} $ \\
@@
             \end{center}
+            After each $P_i$ is fit, redraw $\tilde X_i$ as independent Bernoulli with probability $P_i$.
     \end{enumerate}
+    Drawing the $\tilde X_i$ after each step, rather than just leaving the $P_i$, seems counter intuitive, but under experimentation, it seemed to provide better results.
+As well, though the third step may be repeated several times, in practice this didn't lead to improvement.
 
 \section{Binary Knockoff Performance}
-    To evaluate how useful these binary knockoffs will be in practice, it is useful to evaluate their empirical performance under simulation.
+    While the theory for original knockoffs and random binary variable generation suggests at how the binary knockoffs should perform, seeing their performance in practice still provides invaluable insight to their uses and limitations.
+There are three major areas in which this is useful; how well do the binary knockoffs' empirical cross moment matrices converge to their expectation, how well is FDR controlled for linear regression, and how well, if at all, is FDR controlled for other GLMs?
 
-\subsection{Convergence to Theoretical Covariance}
+\subsection{Convergence to Theoretical Crossmoments} 
+    \FloatBarrier
+        In a certain sense, the question of the convergence of the binary knockoffs' empirical crossmoments is the least interesting of these questions.
+Assuming the Ising model coefficient $A$ or the probability matrix $P$ was successfully fit, the crossmoment matrices' will converge to their expectation in probability for increasingly large values of $n$.
+However, especially with the limited numerical accuracy available to fit $A$ or $P$, the rate at which this convergence occurs and the size of the deviation is an issue of paramount importance.
+The original knockoff theory only provides guarantees when the crossmoments are replicated exactly.
+Thus, further examination of this convergence for various $n$ and $p$ is essential.
+
+    \FloatBarrier
+\subsection{Performance in Linear Regression} 
+    \FloatBarrier
+
+    \begin{figure}[h]
+        \begin{center}
+        \includegraphics[width=14cm]{images/lasso_FDR_power_50}
+    \end{center}
+    \caption{The mean FDR and power of binary and original knockoffs.
+Taken over $200$ simulations of linear regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
+$X$ generated from Ising model while $y$ normally distributed about $X\beta$.}
+    \end{figure}
+
+    \begin{figure}[h]
+        \begin{center}
+        \includegraphics[width=7cm]{images/lasso_corr_50}
+    \end{center}
+    \caption{The mean correlation between variables selected using binary and original knockoffs.
+Taken over $200$ simulations of linear regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
+$X$ generated from Ising model while $y$ normally distributed about $X\beta$.}
+    \end{figure}
+
+    \FloatBarrier
+\subsection{Performance in Logistic Regression} 
+    \FloatBarrier
+    \begin{figure}[h]
+        \begin{center}
+        \includegraphics[width=14cm]{images/logit_FDR_power_50}
+    \end{center}
+    \caption{The mean FDR and power of binary and original knockoffs.
+Taken over $200$ simulations of logistic regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
+$X$ generated from Ising model while $y$ Bernoulli with $p=L(X\beta)$.}
+    \end{figure}
+
+    \begin{figure}[h]
+        \begin{center}
+        \includegraphics[width=7cm]{images/logit_corr_50}
+    \end{center}
+    \caption{The mean correlation between variables selected using binary and original knockoffs.
+Taken over $200$ simulations of logistic regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
+$X$ generated from Ising model while $y$ Bernoulli with $p=L(X\beta)$.}
+    \end{figure}
+
+    \FloatBarrier
+
+\section{Discussion}
+    Binary knockoffs, as a practical tool for statistical inference, are still a work in progress.
+Currently, they work best for high values of $n$, for which the empirical crossmoment matrix deviates little, and relatively small values of $p$, for which the computational time is not prohibitive.
+Considering that FDR control in feature selection is most essential when $p$ is large, this is not a edifying state of affairs.
+However, the approach suggests paths forward to a much more complete tool.
+
+\subsection{Dealing With Small $n$}
+    With random generation of knockoffs, random variation from desired crossmoments is necessarily going to be an issue.
+However, there are likely ways to mitigate this problem by using multiple draws of $\tilde X$.
+For instance, if the $A$ or $P$ matrices were fit once, several draws of $\tilde X$ could be made from it, and a variable selection could be made for each of the corresponding $X_L$.
+Then these could be combined into a single variable selection, most obviously by selecting the variables which were selected for the majority of the $X_L$.
+
+\subsection{Computational Issues}
+    The biggest drawback of this method is its computational requirements.
+As $p$ increases, a larger and larger matrix $A$ must be fit for the Ising model as well as a larger $P$ for the quadratic program.
+For the Ising model, $K$ must similarly grow to provide accurate estimates of the probability space.
+This causes an explosion in computation time.
+Some of this can likely be mitigated by improvements in implementation, and the most time intensive issue, calculating $f(\mathbf a)$, can be easily parallelized.
+A simpler method would still be desirable though.
+A method based on optimizing each column of $P$ individually is thus appealing.
+Certainly, in cases such as the EM algorithm, methods have been developed which have strong guarantees for global optimization using such piecewise methods.
+If a similar development could be made here, or merely a method that hits the desired crossmoments more closely in practice, the result would be far more useful tool.
+
+\subsection{Random Binary Variables With Higher Order Interactions}
 

### 2015-07-27 masters_paper/compare.py
Undoing stacking, because it didn't work

@@ -78,7 +78,7 @@ def bern_y(X,p1,base_prob=.25,beta_sd=1):
 
 def generate(input):
     """ Testing function, all parameters are in a single tuple """
-    seed,gen,p0,p1,method,MCsize,stacks= input
+    seed,gen,p0,p1,method,MCsize= input
     npran.seed(seed)
     if gen.lower() == "ising":
         X    = ising_X(p1+p0,1000)
@@ -95,8 +95,7 @@ def generate(input):
                               knockoff='binary',
                               method=method,
                               MCsize=MCsize,
-                              intercept=True,
-                              stacks = stacks
+                              intercept=True
                               )
     bin_logit.fit()
     ori_logit = ko.knockoff_logit(ybin,X,.2,
@@ -120,8 +119,7 @@ def generate(input):
                               knockoff='binary',
                               method=method,
                               MCsize=MCsize,
-                              intercept=True,
-                              stacks = stacks
+                              intercept=True
                               )
     bin_lasso.fit(bin_logit.X_lrg)
     ori_lasso = ko.knockoff_lasso(ynor,X,.2,
@@ -139,7 +137,7 @@ def generate(input):
     with open('data/lasso_test_'+str(p0+p1)+'.txt','a') as f:
         f.write("%d\t%s\t%d\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\n" % (seed, gen, p1, bin_logit.M_distortion, np.mean(ko_corr), bin_FDR, bin_power, np.mean(ori_lasso.emp_ko_corr), ori_FDR, ori_power, corr))
 
-def batch_compare(b,p,p1s,gens,start_seed,method='fresh_sim',MCsize=100000,stacks=1,procs=4):
+def batch_compare(b,p,p1s,gens,start_seed,method='fresh_sim',MCsize=100000,procs=4):
     with open('data/backup_seeds.txt','r') as f:
         seeds = [int(seed) for seed in f.read().split()]
     inputs = []
@@ -147,7 +145,7 @@ def batch_compare(b,p,p1s,gens,start_seed,method='fresh_sim',MCsize=100000,stack
     for b in range(b):
         for p1 in p1s:
             for gen in gens:
-                inputs.append((seeds[i+start_seed],gen,p-p1,p1,method,MCsize,stacks))
+                inputs.append((seeds[i+start_seed],gen,p-p1,p1,method,MCsize))
                 i += 1
 
     pool = Pool(processes=procs)

### 2015-07-27 masters_paper/knockoffGLM.py

@@ -80,9 +80,11 @@ def cutoff(array,min=0,max=1):
     cut_array = np.max((min*np.ones(shape),cut_array),axis=0)
     return cut_array
 
+
+
 class knockoff_net(object):
     """ Parent class for knockoff lasso and knockoff logistic regression. Defines everything besides fitting the particular GlmNet object """
-    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,method='fresh_sim',pseudocount=0,intercept=False,stacks=1):
+    def __init__(self,y,X,q,knockoff='binary',cov_method='equicor',randomize=False,MCsize=10000,method='fresh_sim',pseudocount=0,intercept=False):
         self.y      = y
         self.X      = normalize(X.astype(float),norm='l2',axis=0)
         self.X_orig = X
@@ -95,7 +97,6 @@ class knockoff_net(object):
         self.zerotol = 1E-5
         self.method = method
         self.pseudocount=pseudocount
-        self.stacks = stacks
         self.intercept = intercept 
 
     def _create_SDP(self):
@@ -369,7 +370,7 @@ class knockoff_net(object):
 
         # If we freshly simulated x, we need to draw ~x based on x
         if self.method=='fresh_sim': 
-            self.X_lrg = np.hstack((np.repeat(self.X_orig,self.stacks,0),np.empty((self.n*self.stacks,self.p))))
+            self.X_lrg = np.hstack((self.X_orig,np.empty((self.n,self.p))))
             for i in np.arange(self.p,2*self.p):
                 # need to make sure the knockoff isn't uniformly 0 or 1
                 count = 0

### 2015-07-27 masters_paper/compare.py
Added code to analyze distortion

@@ -152,3 +152,50 @@ def batch_compare(b,p,p1s,gens,start_seed,method='fresh_sim',MCsize=100000,procs
     pool.map(generate,inputs)
     pool.close()
 
+def distortion(input):
+    """ Testing function, all parameters are in a single tuple """
+    seed,p,n= input
+    npran.seed(seed)
+    X    = ising_X(p,n)
+    ybin = bern_y(X,1)
+
+    if   p<=10:
+        MCsize =10000
+    elif p<=20:
+        MCsize =30000
+    elif p<=30:     
+        MCsize =70000
+    elif p<=40:    
+        MCsize =100000
+    else:    
+        MCsize =150000
+
+    # Logit
+    bin_logit = ko.knockoff_logit(ybin,X,.2,
+                              knockoff='binary',
+                              method='fresh_sim',
+                              MCsize=MCsize,
+                              intercept=True
+                              )
+    bin_logit._binary_knockoff()
+
+    with open('data/distortion.txt','a') as f:
+        f.write("%d\t%d\t%d\t%.5f\n" % (seed, p, n, bin_logit.M_distortion))
+
+def batch_distortion(b,ps,ns,start_seed,procs=4):
+    with open('data/backup_seeds.txt','r') as f:
+        seeds = [int(seed) for seed in f.read().split()]
+    inputs = []
+    i = 0
+    for b in range(b):
+        for p in ps:
+            for n in ns:
+                inputs.append((seeds[i+start_seed],p,n))
+                i += 1
+
+    pool = Pool(processes=procs)
+    pool.map(distortion,inputs)
+    pool.close()
+
+
+

### 2015-07-27 masters_paper/paper.tex
Done with section 4

@@
             \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{M_{ii}}{1-M_{ii}}\right) \end{array} \right]^T \]
-            The hope is that for some $\tau$ close to $0$, convergence can be achieved, only causing a slight distortion from the desired cross moments.
+            The hope is that for some $\tau$ close to $0$, convergence can be achieved, only causing a slight distortion from the desired cross-moments.
 
@@
     Additionally though, the moment condition must hold for $\Sigma_L$ to correspond to a proper random binary vector.
-If $\mathbf m_L = \E\left(\mathbf x_L\right) = [\mathbf m \; \mathbf m]$, then the desired cross moment matrix of the joint distribution is
+If $\mathbf m_L = \E\left(\mathbf x_L\right) = [\mathbf m \; \mathbf m]$, then the desired cross-moment matrix of the joint distribution is
         \[ M_L = \Sigma_L + \mathbf m_L\mathbf m_L^T\]
@@
 \subsubsection{Fitting via $\mu$-conditionals}
-        Once a proper cross moment matrix $M_L$ has been selected, the most obvious approach to generate knockoffs is to apply the Newton-Raphson method to fit the matching $A$ matrix.
+        Once a proper cross-moment matrix $M_L$ has been selected, the most obvious approach to generate knockoffs is to apply the Newton-Raphson method to fit the matching $A$ matrix for the Ising model.
 This can be used to generate $\mathbf{ \tilde x}$ sequentially conditioned on $x$, as $\tilde x_i \st \mathbf x,\tilde x_1,\ldots,\tilde x_{i-1}$.
@@
 
-        Random binary knockoffs which produce the desired cross moment matrix $M_L$ need not arise from a generative model for $\mathbf{\tilde x}$ however.
+        Random binary knockoffs which produce the desired cross-moment matrix $M_L$ need not arise from a generative model for $\mathbf{\tilde x}$ however.
 It is sufficient to have a random matrix $\tilde X \in \{0,1\}^{n\times p}$ such that
@@
             \end{center}
-            Here, the marginal distribution of $\tilde X_i$ will be the same as $X_i$, and the crossmoments shouldn't deviate to far from $M_L$.
+            Here, the marginal distribution of $\tilde X_i$ will be the same as $X_i$, and the cross-moments shouldn't deviate to far from $M_L$.
 This problem can be stated as the quadratic program via the introduction of slack variables:
@@
 Ideally, there would be a method to breakdown this problem into a series of smaller optimization problems which would yield the same solution.
-An optimist might hope that iterating through the columns of $P$, optimizing each to hit its desired crossmoments individually conditioned on the estimate of the rest of $P$, would achieve this.
+An optimist might hope that iterating through the columns of $P$, optimizing each to hit its desired cross-moments individually conditioned on the estimate of the rest of $P$, would achieve this.
 In practice, this does seem to yield reasonable results, and is far quicker than other methods for large $p$.
@@
     While the theory for original knockoffs and random binary variable generation suggests at how the binary knockoffs should perform, seeing their performance in practice still provides invaluable insight to their uses and limitations.
-There are three major areas in which this is useful; how well do the binary knockoffs' empirical cross moment matrices converge to their expectation, how well is FDR controlled for linear regression, and how well, if at all, is FDR controlled for other GLMs?
+There are three major areas in which this is useful; how well do the binary knockoffs' empirical cross-moment matrices converge to their expectation, how well is FDR controlled for linear regression, and how well, if at all, is FDR controlled for other GLMs?
 
-\subsection{Convergence to Theoretical Crossmoments} 
+\subsection{Convergence to Theoretical cross-moments} 
     \FloatBarrier
-        In a certain sense, the question of the convergence of the binary knockoffs' empirical crossmoments is the least interesting of these questions.
-Assuming the Ising model coefficient $A$ or the probability matrix $P$ was successfully fit, the crossmoment matrices' will converge to their expectation in probability for increasingly large values of $n$.
+        In a certain sense, the question of the convergence of the binary knockoffs' empirical cross-moments is the least interesting of these questions.
+Assuming the Ising model coefficient $A$ or the probability matrix $P$ was successfully fit, the cross-moment matrices' will converge to their expectation in probability for increasingly large values of $n$.
 However, especially with the limited numerical accuracy available to fit $A$ or $P$, the rate at which this convergence occurs and the size of the deviation is an issue of paramount importance.
-The original knockoff theory only provides guarantees when the crossmoments are replicated exactly.
-Thus, further examination of this convergence for various $n$ and $p$ is essential.
+As is seen in Figure 3, this convergence occurs at similar rates irrespective of $p$, with fairly meaningful deviation until $n$ has reached a few thousand.
+    \begin{figure}[h]
+        \begin{center}
+        \includegraphics[width=9cm]{images/sigma_fit}
+    \end{center}
+    \caption{Measures how much far the empirical cross-moment matrix $\frac{1}{n}X_L^TX_L$ is from the desired cross moment matrix $M_L$ for random binary knockoffs for various $n$ and $p$, averaged over $50$ simulations of $X$ from the Ising model.
+Where $W = \Sigma - \diag{\mathbf s} + \mathbf m \mathbf m^T$ is the top quadrant of $M_L$, deviation is calculated as \\ $(\|X^T\tilde X - W\|_{fro} + \|\tilde X^T \tilde X - M\|_{fro})/(\|W\|_{fro}+\|M\|_{fro})$}
+    \end{figure}
 
@@
     \FloatBarrier
+    When the knockoff filter is used in linear regression, we see that the original knockoffs are still superior to the binary knockoffs.
+As figure 4 shows, while they both fail to control the FDR when there are only a few true predictors\footnote{This issue is addressed in the original paper; a slightly more conservative filter called knockoff+ avoids this problem}, the original knockoffs properly control FDR for more valid predictors.
+That it does so rather conservatively is a reflection of the relatively low number of total features compared to the sparsity.
+On the other hand, though mirroring the trend with the original knockoffs, the binary knockoffs fail to properly control FDR.
+They select too many variables, which gives higher power at the expense of proper FDR control.
+As figure 5 shows, both methods make related, but no where near identical variable selections.
+\par
 
@@
 Taken over $200$ simulations of linear regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ generated from Ising model while $y$ normally distributed about $X\beta$.}
+$X$ is generated from Ising model while $y$ is normally distributed about $X\beta$.}
     \end{figure}
@@
 Taken over $200$ simulations of linear regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ generated from Ising model while $y$ normally distributed about $X\beta$.}
+$X$ is generated from Ising model while $y$ is normally distributed about $X\beta$.}
     \end{figure}
+    This is disappointing, but shouldn't be surprising.
+The binary knockoffs only replicate the correlation structure of the original features in expectation, rather than exactly.
+Thus, for any given value $n$, the empirical correlation of the combined design matrix $X_L$ won't exactly be $\Sigma_L$, and the method won't work perfectly.
+However, even here, the actual FDR is only somewhat hire than the desired $.2$, and for larger values of $n$, the empirical correlation will get closer to $\Sigma_L$, and the FDR will be controlled better.
 
@@
     \FloatBarrier
+    In logistic regression, the binary knockoffs do seem to provide an improvement over the original knockoffs, as can be seen in figure 6.
+Though neither control FDR properly, using binary knockoffs result both in higher power and lower FDR.
+As well, while the original knockoffs do considerably worse in logistic regression than in linear regression, the binary knockoffs perform only a little worse, having similar FDR and somewhat lower power.
+This suggests that much of the excess FDR is due the deviation from the desired correlation matrix.
+It should be noted though that this simulation is very favorable towards the binary knockoffs; the design matrix $X$ is generated from the Ising model, which is also the model used to generate the binary knockoffs.
+It is quite likely that for $X$ generated with significant high order interactions, the binary knockoffs wouldn't perform as well, though still likely better than the original knockoffs.
     \begin{figure}[h]
@@
 Taken over $200$ simulations of logistic regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ generated from Ising model while $y$ Bernoulli with $p=L(X\beta)$.}
+$X$ is generated from Ising model while $y$ is Bernoulli with $p=L(X\beta)$.}
     \end{figure}
@@
     Binary knockoffs, as a practical tool for statistical inference, are still a work in progress.
-Currently, they work best for high values of $n$, for which the empirical crossmoment matrix deviates little, and relatively small values of $p$, for which the computational time is not prohibitive.
+Currently, they work best for high values of $n$, for which the empirical cross-moment matrix deviates little, and relatively small values of $p$, for which the computational time is not prohibitive.
 Considering that FDR control in feature selection is most essential when $p$ is large, this is not a edifying state of affairs.
@@
 \subsection{Dealing With Small $n$}
-    With random generation of knockoffs, random variation from desired crossmoments is necessarily going to be an issue.
+    With random generation of knockoffs, random variation from desired cross-moments is necessarily going to be an issue.
 However, there are likely ways to mitigate this problem by using multiple draws of $\tilde X$.
@@
 Certainly, in cases such as the EM algorithm, methods have been developed which have strong guarantees for global optimization using such piecewise methods.
-If a similar development could be made here, or merely a method that hits the desired crossmoments more closely in practice, the result would be far more useful tool.
+If a similar development could be made here, or merely a method that hits the desired cross-moments more closely in practice, the result would be far more useful tool.
 

### 2015-07-27 masters_paper/compare.py

@@ -78,15 +78,15 @@ def bern_y(X,p1,base_prob=.25,beta_sd=1):
 
 def generate(input):
     """ Testing function, all parameters are in a single tuple """
-    seed,gen,p0,p1,method,MCsize= input
+    seed,gen,p0,p1,n,method,MCsize= input
     npran.seed(seed)
     if gen.lower() == "ising":
-        X    = ising_X(p1+p0,1000)
+        X    = ising_X(p1+p0,n)
         ynor = norm_y(X,p1)
     elif gen.lower() == "genetic":
         genes = np.genfromtxt('data/SNPdata.txt', delimiter=',')
         np.place(genes,genes!=0,1)
-        X = given_X(p1+p0,1000,genes)
+        X = given_X(p1+p0,n,genes)
     ybin = bern_y(X,p1)
     ynor = norm_y(X,p1)
 
@@ -111,8 +111,8 @@ def generate(input):
     corr      = np.corrcoef(ori_logit.S,bin_logit.S)[0,1]
     ko_corr = [cor for cor in bin_logit.emp_ko_corr if not np.isnan(cor)]
 
-    with open('data/logit_test_'+str(p0+p1)+'.txt','a') as f:
-        f.write("%d\t%s\t%d\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\n" % (seed, gen, p1, bin_logit.M_distortion, np.mean(ko_corr), bin_FDR, bin_power, np.mean(ori_logit.emp_ko_corr), ori_FDR, ori_power, corr))
+    with open('data/logit_test_'+str(p0+p1)+'_w_n.txt','a') as f:
+        f.write("%d\t%s\t%d\t%d\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\t%.5f\n" % (seed, gen, p1, n, bin_logit.M_distortion, np.mean(ko_corr), bin_FDR, bin_power, np.mean(ori_logit.emp_ko_corr), ori_FDR, ori_power, corr))
 
     # LASSO
     bin_lasso = ko.knockoff_lasso(ynor,X,.2,
@@ -134,8 +134,8 @@ def generate(input):
     ori_power = np.dot(ori_lasso.S,trueS)  /max(p1,1)
[…]

### 2015-07-27 masters_paper/paper.tex
Almost done with discussion

@@
 \usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins, caption} 
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage[style=numeric,backend=biber]{biblatex}
+\addbibresource{bibli.bib}
+
 \captionsetup{font={small,singlespacing},margin=1cm}
@@
 \begin{document}
-\title{Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate}
-\author{Aaron Maurer}
-\maketitle
+
+\noindent
+\thispagestyle{empty}
+\underline{\bf Master's Paper of the Department of Statistics, the University of Chicago} 
+    \\~~(Internal document only, not for circulation)
+
+\vspace{1.8in}
+\begin{center}
+    {\bf\LARGE Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate}
+
+    \vspace{1.4in}
+    {\Large Aaron Maurer}
+
+    \vspace{1.3in}
+    {\Large Advisor: Rina Foygel Barber }
+
+\end{center}
+
+\vspace{.6in}
+{\Large Approved} ~\underline{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}
+
+\vspace{.2in}
+{\Large Date} ~\underline{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}
+
+\vfill
+\begin{center}{\large August-31, 2015}\end{center}
+
+\newpage
+\pagestyle{plain}
+\setcounter{page}{1}
+
+\begin{abstract}
+
+    \vspace{7mm}\noindent  Variable selection for regression is a key problem in applied statistics.
+The knockoff filter method provides one method of variable selection for linear regression.
+It relies on generating 'knockoff' features, which replicate the correlation structure of the original variable; when the full path of LASSO regression is fit, the point at which a a null variable and its knockoff have nonzero coefficients will be exchangeable.
+However, for other GLMs, the method breaks down.
+I will provide an alternative method of randomly generating knockoffs for binary variables which will will satisfy the original correlation condition in expectation and offer improved performance for other GLMs.
+\end{abstract}
+
+\newpage
+\vspace{1.5in}
+\tableofcontents
+
+\newpage
+
+%\title{Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate}
+%\author{Aaron Maurer}
+%\maketitle
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     Binary knockoffs, as a practical tool for statistical inference, are still a work in progress.
-Currently, they work best for high values of $n$, for which the empirical cross-moment matrix deviates little, and relatively small values of $p$, for which the computational time is not prohibitive.
-Considering that FDR control in feature selection is most essential when $p$ is large, this is not a edifying state of affairs.
-However, the approach suggests paths forward to a much more complete tool.
-
-\subsection{Dealing With Small $n$}
-    With random generation of knockoffs, random variation from desired cross-moments is necessarily going to be an issue.
-However, there are likely ways to mitigate this problem by using multiple draws of $\tilde X$.
-For instance, if the $A$ or $P$ matrices were fit once, several draws of $\tilde X$ could be made from it, and a variable selection could be made for each of the corresponding $X_L$.
-Then these could be combined into a single variable selection, most obviously by selecting the variables which were selected for the majority of the $X_L$.
-
-\subsection{Computational Issues}
-    The biggest drawback of this method is its computational requirements.
+Currently, they provide approximate FDR control for LASSO and seem like they also provide approximate FDR control for logistic regression.
+However, there are no theoretical guarantees for logistic regression, and the quality of FDR control for LASSO depends on how far the realized cross-moment matrix deviates from the desired $M_L$.
+In cases where $n$ is large, this should only be a small issue, but for smaller $n$ FDR control fails in a meaningful way.
+As well, computational issues make binary knockoffs hard to generate for large $p$; as $p$ increases, the Ising model to generate the knockoffs gets rapidly more computationally intensive to fit, as do the methods based on quadratic programs.
+This leaves the method, as currently developed, only suitable for relatively high values of $n$ and low values of $p$.
+Considering FDR control for variable selection is most important for data sets with large $p$, this is not terribly satisfying.
+\par
+    However, this method, with the real possibility of extending the knockoff filter to GLMs, offers an interesting development if these issues can be resolved.
+For the issue with the deviation from the desired cross-moment matrix, a solution would likely take the form of an ensemble method.
+Rather than fit the $A$ and $P$ matrix to generate one $\tilde X$, several $\tilde X$ might be generated and used to make variable selections $S_m$.
+A finally variable selection $S$ could be made based on theses, perhaps only selecting the variables selected by the majority of the $S_m$.
+This, or some other sort of average across multiple realizations of $\tilde X$, could likely be used to dampen the effect of variance in the generation of $\tilde X$.
+This would at least make binary knockoffs reliable for Lasso.
+\par
+    The computational issues will likely be more difficult to deal with.
 As $p$ increases, a larger and larger matrix $A$ must be fit for the Ising model as well as a larger $P$ for the quadratic program.
-For the Ising model, $K$ must similarly grow to provide accurate estimates of the probability space.
+For the Ising model, the simulation size $K$ must similarly grow to provide accurate estimates of the probability space.
 This causes an explosion in computation time.
 Some of this can likely be mitigated by improvements in implementation, and the most time intensive issue, calculating $f(\mathbf a)$, can be easily parallelized.
-A simpler method would still be desirable though.
+A less computationally intensive method to derive either $A$ or $P$ is necessary to make $p$ in the hundreds or thousands feasible.
 A method based on optimizing each column of $P$ individually is thus appealing.
-Certainly, in cases such as the EM algorithm, methods have been developed which have strong guarantees for global optimization using such piecewise methods.
+So far, only approximate solutions have resulted from this method.
+However, in similar situations methods, such as the EM algorithm, have been developed which have strong guarantees for global optimization using such piecewise optimization.
 If a similar development could be made here, or merely a method that hits the desired cross-moments more closely in practice, the result would be far more useful tool.
+\par
+    The final area of concern is how binary knockoffs will occur when the $X$ matrix is not fit well by the Ising model due to significant high order interactions.
+Hopefully, having matched the first and second moments of $X$, binary knockoffs will continue to provide reasonable FDR control.
+An avenue to deal with this though would be to explicitly model these higher order interactions.
+One might introduce higher order moments into the fitting of $\mu$-conditionals.
+For instance, if $A$ is a three dimensional matrix whose entry $a_{ijk}$ would be added to the linear predictor when $x_i=x_j=x_k=1$.
+Just as the $\mu$-conditionals are based on a regression of $X_i$ on $X_{1:i-1}$ with no interactions, this would be akin to adding two way interactions.
+This can be used to replicate arbitrary binary vectors to an arbitrary desired level of precision.
+However, in practice this has big limitations.
+For one, adding additional levels of interactions will make the computation much more difficult very quickly.
+As well, as far as generating knockoffs, it will become more and more difficult to reduce the correlation between $x_i$ and $\tilde x_i$ as higher order interactions are insisted on.
+Nonetheless, this may be an avenue for further investigation.
 
-\subsection{Random Binary Variables With Higher Order Interactions}
-
+\nocite{*}
+\printbibliography[heading=bibintoc]
 \end{document}

### 2015-07-27 masters_paper/simulate.py
Done with discussion

@@ -293,6 +293,7 @@ class ko_test(object):
             plot.set_xlim(1,self.p)
             plot.set_ylim(0,1)
             plot.legend(loc='lower right')
+            plot.plot((0,self.p),(.5,.5),'k--')
 
     def calc_fdr_power(self):
         trueS = (np.arange(self.p)<self.p1).astype(int)

### 2015-07-29 masters_paper/presentation.tex
working on presentation

--- 2015-07-27 masters_paper/paper.tex
@@
-\documentclass[11pt]{article}
-\usepackage[paper=letterpaper, margin=.5in]{geometry}
-\linespread{1.5}
-\pdfpagewidth 8.5in
-\pdfpageheight 11in
+\documentclass{beamer}
+\usetheme{umbc1}
 
@@
 % placeins defines "\FloatBarrier", which stops tables from moving around
-\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins, caption} 
-\captionsetup{font={small,singlespacing},margin=1cm}
+\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
 \newcommand{\Z}{\mathbb{Z}}
@@
 
-\newcommand{\p}{\mathrm{P}}
 \newcommand{\E}{\mathrm{E}}
@@
 \newcommand{\cov}{\mathrm{Cov}}
-\newcommand{\diag}[1]{\mathrm{diag}\{#1\}}
 \newcommand{\ST}{mbox{ s.t. }}
-
-\newcommand{\x}{\bf x}
-
-\newcommand{\Example}{\noindent {\bf Example. \quad} }
-\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
-\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
-\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
-\newcommand{\Case}{\noindent {\underline{Case} \quad} }
 
@@
 \newcommand{\ddisc}{d_{\mathrm disc}}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{corollary}[theorem]{Corollary}
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
+
+\newcommand{\hwhead}[1]{#1 \hfill Aaron Maurer \vspace{2mm} \hrule \vspace{2mm}}
+\newcommand{\diag}[1]{\mathrm{diag}\{#1\}}
 \DeclareMathOperator*{\argmin}{arg\min}
 
-\newcommand{\hwhead}[1]{#1 \hfill Aaron Maurer \vspace{2mm} \hrule \vspace{2mm}}
+\title{Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate}
+\author{Aaron Maurer}
+\date{July 29th, 2015}
 
 \begin{document}
-\title{Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate}
-\author{Aaron Maurer}
-\maketitle
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-    Variable selection is an essential problem to fitting a regression model.
-For linear regression, the knockoff filter (Barber and Cand\`{e}s, 2014) offers a method of exact FDR control.
-However, the method as originally formulated does not extend to other generalized linear models such as logistic regression.
-This paper offers an extension of this method which will work with binary features in the context of linear regression, but also will extend to other GLMs.
+\begin{frame}[plain]
+    \titlepage
+\end{frame}
 
-\subsection{Knockoff Filter}
-    Let us consider the consider the usual setting for linear regression, where $n$ observations of a variable of interest $y$ arise from the model
+\begin{frame}{Overview}
+    \begin{enumerate} 
+        \item Original Knockoffs: What They Do and Where They Fail
+        \item Making Knockoffs Work With GLMs
+        \item Random Binary Knockoffs: The Theory
+        \item Random Binary Knockoffs: Performance
+        \item Where to next?
+    \end{enumerate}
+\end{frame}
+
+\begin{frame}{Variable Selection in Linear Regression}
+    Assume
      \[\mathbf{y} = X\beta + \mathbf{z}\]
-    where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise.
-It is important to note that a intercept vector $\mathbf 1$ is not included in $X$.
-For the purpose of this paper, only the case where $2p\leq n$ will be considered, though the methods can be extended to $p\leq n$.
-I will to refer the $j$th column of $X$ as $X_j$, which is $n$ observations of the random variable $x_j$.
-The random variable $x_j$ is in turn $j$th entry of the random vector valued variable $\mathbf x$.
-In the situation where $\beta$ is sparse and there is reason to expect that $x_j$ and $y$ are uncorrelated for many $j$, the knockoff filter is a method to select a subset of the $x_j$ which are likely correlated with $y$.
-\par
-    The method specifically seeks to control the false discovery rate (FDR), which is the expected portion of some selection of $x_j$ which are, in fact, uncorrelated with $y$.
-For a procedure which chooses a subset of variables $\hat S\subseteq \{1,\ldots,p\}$, this is defined as
+    where $\mathbf{y}\in\R^n$, $X \in \R^{n\times p}$, $\beta\in\R^p$, and $\mathbf z$ is Gaussian noise. Also, assume sparsity:
+    \[\beta_i = 0 \quad \forall i\not\in S\]
+    How do we pick estimate $\hat S$?
+\end{frame}
+
+\begin{frame}{False Discover Rate}
+    A common goal for a method that generates $\hat S$ is to control the false discovery rate
     \[ \textrm{FDR} = \E\left[\frac{\vert{\{j: \beta_j=0 \; \& \; j\in\hat S\}}\vert}{\max\{\vert{\hat S}\vert,1\}} \right] \]
-    FDR is controlled at level $q$ if FDR is less than $q$ irrespective of the coefficients $\beta$. \par
+    In other words, control portion of elements in $\hat S$ which aren't in $S$. \par
+    \vspace{1cm}
+    FDR is controlled at level $q$ if $q<$FDR irrespective of true $\beta$.
 
-    The knockoff filter achieves FDR control in several steps, the first of which is creating a set of `knockoff' features $\tilde X_j$ which imitate the original $X_j$ while being less correlated with $y$.
-In particular, the matrix of knockoffs $\tilde X$ has the same internal correlation structure as $X$, the knockoff feature $\tilde X_j$ has the same correlation with other $X_i$ as $X_j$ does, but the correlation between $\tilde X_j$ and $X_j$ is minimized.
-In other words, where $ G := X^TX$,
-    \[ \tilde X^T\tilde X=G \textrm{ and } \tilde X^T X = G - \diag{\mathbf s}\]
-    for some vector $s\in \R^p$ such that, where $\diag{G}$ is the vector made from the main diagonal of $G$, $\diag{G} - \mathbf s$ is small.
-Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$.
-In the case where $X_j$ is uncorrelated with $\mathbf y$, $\tilde X_j$ will also be uncorrelated with $\mathbf y$.
-How these knockoffs are actually generated will be described shortly.\par
-    The second step is to fit a series of Lasso model of $\mathbf y$ on the combined design matrix $X_L=[X \; \tilde X]$ so as to determine the largest value $\lambda$ at which the coefficient for each $X_j$ and $\tilde X_j$ is nonzero.
-The subscript $L$ refers to the large design matrix with both features and knockoffs.
-\footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} Recall, for a given $\lambda$, the estimated coefficient from the Lasso regression will be
+\end{frame}
+
+\begin{frame}{Knockoffs}
+    Knockoff variables can be used to control FDR in linear regression. 
+    \begin{itemize}
+        \item The idea is to create a forgery of each variable; if the forgeries seem about as good predictors as the originals, the originals are lousy predictors.
+        \item For each variable $X_i$, create a knockoff feature $\tilde X_i$. Such that. 
+            \[ \tilde X^T \tilde X = X^T X \quad \& \quad X^T \tilde X = X^T X - \diag{\mathbf s} \]
+            Where $\diag{X^T X} - s$ is small but $\diag{X^T X} - s\succeq0$
+        \item $\tilde X_i$ and $X_i$ will have same correlation with other variables, but only low correlation with each other.
+        \item Given $\mathbf s$, $\tilde X$ can be generated via a rotation of $X$.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Knockoff Filter}
+    These knockoffs can be used in the knockoff filter method. 
+    \begin{itemize}
+        \item Fit full path of LASSO regression on $[\,X\,\tilde X\,]$.
     \[ \beta(\lambda) = \argmin_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]
-    In the general case, the coefficient for one feature in a Lasso regression being first nonzero for a larger value of $\lambda$ than another feature is an indication that the first feature is a stronger predictor of the outcome than the second.
-Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to have nonzero coefficients sooner than the knockoff features when the original is a valid predictor.
-On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $G_L=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
-This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $G_L$ will be unaltered, while $\E\left[X_L^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$.
-\par
-   This observation leads to the final step.
-Let $Z_j$ be the largest $\lambda$ such that feature $j$ has a nonzero coefficient and $\tilde Z_j$ the largest $\lambda$ such that knockoff $j$ has a nonzero coefficient.
-Then, define $W_j$ as
-   \[ W_j = \begin{cases}   Z_j         &\mbox{if } Z_j>\tilde Z_j \\ 
-                            -\tilde Z_j &\mbox{if } Z_j<\tilde Z_j \\
-                            0           &\mbox{if } Z_j=\tilde Z_j \\
-            \end{cases} \]
-            When $X_j$ is a null predictor, $W_j$ will be symmetrically distributed around zero, and when $X_j$ is true predictor, it will have a distribution skewed positive.
-Since $W_j$ with large positive values are more likely to be true predictors, the knockoff filter selects the variables $\{j:W_j\geq T\}$, where $T$ is a threshold defined as
+        \item $Z_i$, $\tilde Z_i$ largest $\lambda$ such that $X_i$, $\tilde X_i$ have nonzero coefficient.
+        \item $W_i= Z_i$ if $Z_i>\tilde Z_i$, otherwise $W_i = -\tilde Z_i$.
+        \item Since $[\, X \, \tilde X\,]^T[\, X \, \tilde X\,]$ \& $[\, X \, \tilde X\,]^T\mathbf y$ sufficient statistics for $\beta(\lambda)$, $W_i$ symmetrically distributed around $0$ when $X_i$ null predictor.
+        \item Thus, FDR controlled when $\hat S = \{i:W_i>T\}$ for 
             \[ T = \min\left\{ t>0 \;: \; \frac{\vert\{j:W_j\leq -t\}\vert}{\max\{\vert\{j:W_j\geq t\}\vert,1\}}\leq q \right\} \]
-            Basic logic of this method is fairly simple.
-Let $N_h(t)$ be the random variable for the number of null predictors with $W_j\leq t$, $N_l(t)$ be the number of null predictors with $W_j\geq t$, and $V_l(t)$ \& $V_h(t)$ be the similar number of true predictors.
-Since the null $W_j$ are symmetrically distributed around $0$, $\E[N_l(t)]=\E[N_h(t)]$.
-Thus,
-            \begin{align*}
-                q &\geq \E\left[ \frac{\vert\{j:W_j\leq -t\}\vert}{\max\{\vert\{j:W_j\geq t\}\vert,1\}} \right] \\
-                q &\geq \E\left[ \frac{N_l(t) + V_l(t)}{\max\{N_h(t) + V_h(t),1\}} \right] \\
-                q &\geq \E\left[ \frac{N_l(t)}{\max\{N_h(t) + V_h(t),1\}} \right] \\
-                q &\geq \E\left[ \frac{N_h(t)}{\max\{N_h(t) + V_h(t),1\}} \right] \\
-                q &\geq \textrm{FDR}
-            \end{align*}
-            So FDR is controlled, and since $V_l(t)$ will tend to be small, it should be controlled fairly tightly.
+    \end{itemize}
+\end{frame}
 
-\subsection{Original Knockoff Features}
-    The original formulation of the method offers two similar methods of constructing knockoffs. Both of these will, by construction, have exactly the property that
-    \[ \tilde X^T\tilde X=G \textrm{ and } \tilde X^T X = G - \diag{\mathbf s}.\]
-    For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. Then, let the Gram matrix of $X_L$ is
-    \[ G_L= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right] \]
-    $A$, the Schur complement of $G$ in $G_L$ can be calculated as 
-    \[ A = 2\,\diag{\mathbf s} - \diag{\mathbf s}G^{-1}\diag{\mathbf s} \]
-    For $G_L$ to exist, $G_L$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if\footnote{A fleshed out version of this argument can be found in the original knockoff paper}
-    \[ \diag{\mathbf s} \succeq 0  \; \textrm{ and } \; 2G - \diag{\mathbf s} \succeq 0 \]
-    Given this is true, $A$ can be factored as $A=C^TC$.
-Combining this with a satisfactory $\mathbf s$ and an orthonormal matrix $\tilde U\in\R^{n\times p}$ such that $\tilde U^T X = 0$, a $\tilde X$ fulfilling the desired properties can be calculated as
-    \[ \tilde X = X(I-G^{-1}\,\diag{\mathbf s}) + \tilde UC\]
-    As mentioned above, $\mathbf s$ should be chosen so as to make $\diag{G}-\mathbf s$ small.
-Since each $X_j$ has been normalized, this means that $\diag{G}=\mathbf 1$, so $\mathbf 1 - \mathbf s$ should be minimized in accordance with the restrictions on $\mathbf s$.
-The two methods differ in how $\mathbf s$ is chosen:
+\begin{frame}{Where Knockoff Filter Fails}
+    Knockoff filter don't work for other GLMs. \\
+    \begin{center}
+        \includegraphics[width=8cm]{images/entryrate_original_logit}
+    \end{center}
+\end{frame}
+
+\begin{frame}{Can Knockoffs Be Fixed for GLMs?}
     \begin{itemize}
-        \item \textit{Equi-correlated knockoffs}: Each original features is set to have the same correlation with its knockoff by setting $\mathbf s = 2\min\{\lambda_{min}(G),1\}\mathbf 1$.
-For all simulations in this paper, this was the method used.
-        \item \textit{SDP knockoffs}: The $\mathbf s$ which minimizes the average correlation between knockoff and original features can be found via a semi-definite programing problem:
-            \begin{center}
-                \begin{tabular}{r l}
-                    minimize & $\|\mathbf1-\mathbf s\|_1$ \\
-                    subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2G $
-                \end{tabular} 
-            \end{center}
-            This method is significantly more computationally intensive. 
+        \item Other GLMs don't have the same sufficient statistics as linear regression.
+        \item Original Knockoffs don't remotely have same distribution as $X$, so ``look'' different than real variables.
+        \item Knockoffs will likely work better if they have the same marginal distribution as originals. 
+        \item For $X_i$ with arbitrary distribution, unclear how this might be accomplished.
     \end{itemize}
+\end{frame}
 
-\subsection{Binary Knockoffs}
-    Though the ``original" knockoff method, described above, achieves the exact desired knockoff properties, the individual values in the vector $X_j$ will have little relation to the individual values in $X_j$.
-In particular, these values will often have very different empirical distributions.
-This effect is particularly noticeable when the random variable $x_j$ is discrete but $\tilde X_j$ does not even consist of the same set of values.
-The end result is that for variable selection for other generalized linear models, which do not have the same sufficient statistics as linear regression which knockoffs are designed to hit, the original knockoffs will fail to control FDR.
-\par
-    Thus, this paper offers a new method of generating knockoffs which should offer superior performance with other GLMs.
-In this new method, the matrix of knockoff features $\tilde X$ will be generated by row from a random vector valued variable $\mathbf{\tilde x}$, each entry $\tilde x_j$ of which will have the property $\tilde x_j \sim x_j$.
-This random variable will also abide by a relaxation of the original knockoff condition:
-    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=G \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = G - \diag{\mathbf s}\]
-    This will ensure that, at the very least, these knockoffs are a suitable replacement for the original knockoffs for linear regression. \par
-    Generating such knockoffs for a completely general case of any $\mathbf x$ is likely infeasible.
-However, in the specific case where $\mathbf x$ is a random binary vector variable, this paper will demonstrate a method to generate such knockoff variables.
-Since binary data is extremely common in many areas, this will hopefully serve as a useful tool.
-
-\subsection{Paper Organization}
-The rest of this paper will be organized in the following fashion:
+\begin{frame}{Random Binary Knockoffs}
     \begin{itemize}
-    \item Section 2 will go into more detail about how the original knockoffs breakdown with generalized linear models. In particular, several simulations will demonstrate how they breakdown.
-    \item Section 3 will develop the method and theory for generating binary knockoffs.
-    \item Section 4 will discuss tests of Binary knockoffs in simulation, both in comparison to the original knockoffs with Lasso and for logistic regression by itself.
-    \item Section 5 is the final section, and will contain discussion of the results as well as areas for further work.
+        \item A useful 
     \end{itemize}
-
-\section{Issues With Deterministic Knockoffs} 
-    The knockoff filter and $W$ statistics have a very natural extension to generalized linear models.
-Here, where $l(\beta\st X, \tilde X)$ is the log likelihood of coefficients $\beta$ given the model, the $L1$ regularized regression model will have estimated coefficient vector
-    \[ \beta(\lambda) = \argmin_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
-    The $Z_j$ and $\tilde Z_j$ are then, once again, the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before.
-The problem with this is that if the given GLM doesn't have the same sufficient statistics as linear regression, there is no theoretical guarantee that the knockoff filter will control FDR.
-This is since, for null $X_j$, there is no theory to guarantee the $Z_j$ and $\tilde Z_j$ are exchangeable, and thus $W_j$ is not guaranteed to be symmetrically distributed around $0$.
-\par
-    \begin{figure}[h]
-        \begin{center}
-        \includegraphics[width=14cm]{images/entryrate_original_logit}
-    \end{center}
-        \caption{The mean cumulative portion of $W$ statistic which are positive by absolute rank of the value.
-The mean is over 1000 simulations.
-On the left side, the features are drawn from a random standard normal distribution.
-On the right side, they are drawn as Bernoulli with probability $.25$ of being 1.
-The top has 50 features uncorrelated with the outcome, while the bottom has 40 features uncorrelated with the outcome and 10 correlated with it.}
-    \end{figure}
-
-    To test how bad this problem is, I tested the method using original style knockoffs in logistic regression.
-The results can be seen in figure 1, which shows cumulatively, starting with the largest absolute value, what portion of the $W_j$ are positive.
-For the method to work, for the null predictors, this should be a consistent $50\%$, indicative of the desired symmetric distribution.
-This is particularly important for high values, which play the largest roll in selecting a threshold.
-We see that in the top left corner, when all features are null and normally distributed, that we almost achieve this, with only a small, but consistent, deviation from $50\%$.
-However, in the top right, when the predictors are binary, the null originals enter first far more often, which will result in the FDR being too high.
-\par
-    \FloatBarrier
-    This gets much worse when some true predictors are included, as shown in the bottom row of the figure.
-For both normal and binary predictors, the originals come in first far more often in the beginning.
-This will completely ruin FDR control, since the portion of null predictors will be assumed far lower than it actually is.
-It is worth noting that the particular shape of the curve in the diagram, with the plateau for the null predictors through 10, is partially an artifact of how the totals were tabulated.
-The mean through the first ten is the mean for null predictors that came in first out of both null and true predictors; since the largest ten $W_j$ will tend to be true predictors, the average over the null predictors represent the few extreme $W_j$ which managed to beat out some number of true predictors.
-\par
-    \begin{figure}[h]
-        \includegraphics[width=14cm]{images/normalQQ}
-        \caption{Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
-    \end{figure}
-
-    Upon examination, the issue seems to be that the original style knockoff features have a very different empirical distribution than the features they are imitating.
-This is obviously true with the binary features, but even with the normal features, the distributions differ a lot.
-In figure 2, the distribution of four feature and the original style knockoff are plotted against a normal distribution in Q-Q plots.
-As you can see, they don't match at all.
-This is suggestive of a solution; if the knockoff variables match the marginal distribution of the original features better, the result may be more suitable for FDR control in GLMs.
-\par
-    \FloatBarrier
-
-\section{Random Binary Knockoffs}
-Random binary knockoffs $\mathbf{\tilde x}$, which maintain the correlation condition in expectation, fit the bill in this regard.
-The method to produce them is based heavily on more general methodology to random binary vector variables with internal correlation.
-
-\subsection{Random Binary Vector Generation}
-The full class of random binary vector variables on $\{0,1\}^p$ can be specified as multinomial on the $2^p$ elements of the set.
-However, in practice, it is generally impractical to specify or estimate all $2^p$ necessary parameters, so it is useful to, as is desirable for knockoffs, pick a method to match the first two moments.
-These are the mean vector $\E(\bf x)=m\in(0,1)^p$ and the cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$.
-Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag{M}$.
-For an arbitrary symmetric $M$ to be valid cross-moment matrix for some random binary vector variable, $M-mm^T\succeq0$, and
-    \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
-    for all $i\neq j.$ Given a qualifying $M$, or observed $X$, there are then a few methods for generating additional random values.
-
-\subsubsection{Gaussian Copula Family}
-Since multivariate normal distributions are easy to randomly draw from and defined by their first two moments, an obvious choice is to use them to generate random binary vectors.
-This might be done by finding $\mu$ and $\Sigma$ such that for $\mathbf z\sim N_p(\mu,\Sigma)$, $x_i$ can be generated as $x_i = \mathrm{sign}(z_i)$, and have the desired first two moments.
-However, this is only guaranteed to be feasible in the bivariate case.
-For higher dimensions, this method can generally only provide random binary vectors with approximately the right moments, and the quality of the approximation quickly degenerates as $p$ increases.
-
-\subsubsection{$\mu$-Conditionals family}
-    There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as 
-    \[ \p(x_{i}=1\st x_{1},...,x_{i-1}) = \mu\left(a_{ii}+\sum_{k=1}^{i-1}a_{ik}x_i\right) \]
-    for some monotone function $\mu:\R\to(0,1)$.
-This is essentially a binomial family GLM for a link function $\mu$.
-If one takes all of the $a_{kj}$, they can form a lower triangular matrix $A$, and then the joint density can be expressed as
-
-    \[ \p(\mathbf x=\mathbf \gamma) \propto \mu(\mathbf{\gamma}^T A\mathbf \gamma)\]
-
-    If $\mu$ is chosen such that it is a bijection and differentiable, there is a unique $A$ such that $\E(\mathbf x \mathbf x^T)=M$ when generated from this model.
-The natural choice for $\mu$ is the logistic link function $L(x)=\frac{1}{1-e^{-x}}$, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
-    \[a_{ij} = \log\left(\frac{\p(x_j=x_k=1)\p(x_j=x_k=0)}{\p(x_j=0,x_k=1)\p(x_j=1,x_k=0)}\right) \]
-    when $i\neq j$. When $\bf x$ is generated from this model with $a_{ij}=0$, then $x_i$ and $x_j$ are conditionally independent.  \par
-
-    There is no closed form to calculate the entries in $A$ if $p>1$, but they can be derived numerically two ways.
-The first of these, when one is attempting to replicate the empirical cross-moments from a data matrix $X$, $a_{1i}$ to $a_{ii}$ is by fitting successive binomial family GLM regressions.
-The link function is the given function $\mu$, and $X_i$ is regressed on $X_{1} \ldots X_{i-1}$ using maximum likelihood.
-$a_{ji}$ for $i\neq j$ will then be the coefficient on $X_j$ while $a_{ii}$ is the intercept of the regression.
-\par
-    Otherwise, if one is just working with a desired cross-moment matrix $M$, the successive rows of $A$ can be fit via Newton-Raphson.
-This is performed successively on each row of $A$.
-If the first $i-1$ rows have been fit, then the upper left $(i-1)\times(i-1)$ sub matrix $A_{-i}$ of $A$ has already been filled.
-Next, $\mathbf a_i$, the first $i$ entries of the $i$th row of $A$ must be fit, while the rest of the row will be 0.
-This corresponds to $\mathbf m_i$, the first $i$ entries of the $i$th row of $M$.
-In other words, $\mathbf{m}_i = [\E(x_i x_j)]_{j=1}^{i}$.
-Finally, let us say that $\mathbf x_{-i}$ is the first $i-1$ entries of $\mathbf x$.
-$\mathbf a_i$ must satisfy
-            \begin{align*}
-                \mathbf m_i &= \E\left(x_i \left[\begin{array}{c} \mathbf x_{-i} \\ x_i \end{array}\right]\right) \\
-                \mathbf m_i &= \E\left(\E\left(x_i \left[\begin{array}{c} \mathbf x_{-i} \\ x_i \end{array}\right]\;\bigg \vert\; \mathbf x_{-i}\right)\right)\\
-                \mathbf m_i &= \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \p(\mathbf x_{-i})\p(x_i=1\st \mathbf x_{-i}) \left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \\
-            \end{align*}
-            When $\mu$ is the logistic link function $L$, this is
-            \[ \mathbf m_i = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L\left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \]
-            Where $c$ is the appropriate normalizing constant.
-Let us define the quantity on the right in the last line as $f(\mathbf a_i)$.
-This equation $f(\mathbf a_i)=\mathbf{m}_i$ can be solved by successive Newton-Raphson iterations defined by
-            \[\mathbf a_i^{(k+1)} = \mathbf a_i^{(k)} - \left[J\left(\mathbf a_i^{(k)}\right) \right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
-            where $J$ is the Jacobian of $f$. To do so, the Jacobian matrix is calculated as 
-        \[ J\left( \mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
-            With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive.
-Instead, with a series of values $\mathbf x_{-i}^{(k)}\sim \mathbf x_{-i}$, the simulated values
-            \[f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right] \]
-            and
-            \[J\left( \mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
-            can be used. \par
-            Though, in theory, $A$ should always exist, in practice limited accuracy may dictate that Newton-Raphson method won't converge.
-In the simulation method in particular, if $K$ is not large enough, the equations can be very difficult or impossible to solve.
-When the intial problem can't be solved, besides increasing the simulation size or numerical accuracy, one can instead solve the relaxed problem  $f(\mathbf a)=\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
-    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T \]
-            When $\tau=0$, this yields the original problem, while when $\tau=1$, it is treats $x_i$ as independent of $\mathbf x_{-i}$. The latter will always have the solution
-            \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{M_{ii}}{1-M_{ii}}\right) \end{array} \right]^T \]
-            The hope is that for some $\tau$ close to $0$, convergence can be achieved, only causing a slight distortion from the desired cross moments.
-
-\subsection{Generating Binary Knockoffs}
-This theory provides the tools to generate random binary knockoffs.
-This is done is done sequentially as two steps, the first of which is to come up with the target covariance matrix $\Sigma_L$ for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$.
-This is extremely similar, though not identical, to the method with the original knockoffs, where $X_L^TX_L$ after normalization was targeted.
-However, much the same method as before can be used.
-Let the vector of empirical means of the columns of $X$ is $\mathbf m = \frac{1}{n}X^T\mathbf 1$ and the empirical covariance matrix of $X$ is $\Sigma = \frac{1}{n}X^TX - \mathbf{m}\mathbf{m}^T$.
-Much as before with the original knockoffs, both the equi-correlated and SDP method can be used to select  $\mathbf s$  such that $\diag{\Sigma}-\mathbf s$ is small and
-        \[ \Sigma_L = \left[\begin{array}{cc}  \Sigma & \Sigma - \diag{\mathbf s }\\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \succeq 0 \]
-    While with the original knockoffs, the result was small values for $X_i^T\tilde X_i$, now the result is low covariance between $x_i$ and $\tilde x_i$.\par
-    Additionally though, the moment condition must hold for $\Sigma_L$ to correspond to a proper random binary vector.
-If $\mathbf m_L = \E\left(\mathbf x_L\right) = [\mathbf m \; \mathbf m]$, then the desired cross moment matrix of the joint distribution is
-        \[ M_L = \Sigma_L + \mathbf m_L\mathbf m_L^T\]
-        So it must be the case that  
-        \[\max\{0,\mathbf m_{L,i} + \mathbf m_{L,j} -1\} \leq \Sigma_{L,ij} + \mathbf m_i \mathbf m_j \leq \min\{\mathbf m_{L,i}, \mathbf m_{L,j}\} \]
-        However, since $\Sigma$ and $\mathbf m$ are from an actual random binary vector, the entries of $\Sigma$ satisfy this property by default.
-Only off diagonal entries which have had $\mathbf s$ subtracted from them could possibly violate this condition.
-Furthermore, for these entries, the goal is to make $\Sigma_{L,ij}$ as close to $0$ as possible.
-Since the condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$, this implies that this method should normally be satisfied.
-This is the principle advantage minimizing the $\textrm{cor}(x_i,\tilde x_i)$ rather than $\E[x_i\tilde x_i]$.
-For the SDP method though, the restriction can be built directly into the optimization problem.
-If $\sigma = \diag{\Sigma}^{\frac{1}{2}}$ is the vector of standard deviations, the updated problem is
-            \begin{center}
-                \begin{tabular}{r l}
-                    minimize & $\|\mathbf 1-\diag{\sigma}^{-1}\mathbf s\|_1$ \\
-                    subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2\Sigma $ \\
-                               & $ 2\mathbf m_i -1 \leq \Sigma_{ii}-\mathbf s_i + \mathbf m_i^2 \leq \mathbf m_i  \quad \forall i$
-                \end{tabular} 
-            \end{center} 
-            the factor of $\diag{\sigma}^{-1}$ in the objective is to account for the $x_i$ not being normalized, as was the case with the original knockoffs. \par
-\subsubsection{Fitting via $\mu$-conditionals}
-        Once a proper cross moment matrix $M_L$ has been selected, the most obvious approach to generate knockoffs is to apply the Newton-Raphson method to fit the matching $A$ matrix.
-This can be used to generate $\mathbf{ \tilde x}$ sequentially conditioned on $x$, as $\tilde x_i \st \mathbf x,\tilde x_1,\ldots,\tilde x_{i-1}$.
-However, there are a few variations of this process which are preferable depending on the situation.
-        \begin{enumerate}
-            \item If the whole $A$ matrix is being fit, then the upper half of the matrix, corresponding to $\mathbf x$, can be fit via the regression method, since they arise from real data $X$.
-This will tend to be quicker, since the calculation lends itself to parallel computing, and the data set will often be smaller than the space of all binary vectors or a simulation of that space.
-The second half of the matrix still must be fit via Newton-Raphson.
-            \item One can avoid fitting the first half of the $A$ matrix entirely though if performing Newton-Raphoson with by using the actual observed values in $X$ to simulate the first half of $\mathbf x_L$.
-This can be done by either bootstrapping rows from $X$ or using some number of copies of $X$ exactly.
-The advantage of this is the saved computation and that $\tilde x$ should be built to match the correlation condition more exactly.
-The downside is that, since these values won't be from the Ising model, the theoretical guarantee of a suitable $A$ may no longer hold.
-As well, if $n$ isn't big enough, the equations will quickly become difficult to solve, since the observed samples will provide a noisy and incomplete sample of the probability space.
-            \item As $A$ is being fit, even the parts of $\mathbf x_{-i}$ corresponding to knockoffs need not be redrawn for each Newton-Raphson iteration or each successive $i$.
-To fit $\mathbf a$ for feature $i$, the $\mathbf x_{-i}^{(k)}$ will remain constant; then, once $\mathbf a$ has been fit, the $x_i^{(k)}$ need only be drawn once.
-This saves computation, and the multiplication
-            \[ \left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [\mathbf x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
-                needs only be partially recalculated for each Newton-Raphson iteration since the upper left portion, $\mathbf x_{-i}^{(k)}[\mathbf{x}_{-i}^{(k)}]^T$, won't change.
-This may result in slight deviations in any particular draw compounding into a more severe error over time.
-            \item As $p$ grows large, and $K$ necessarily along with it, the cost of calculating the Jacobian matrix grows far quicker than the cost of calculating $f(\mathbf a)$.
-This is since it has $O(Kp^2)$ complexity compared to $O(Kp)$ complexity for $f(\mathbf a)$, quickly becoming prohibitive.
-One might instead estimate the full Jacobian based on first differences, but in practice an even simpler method called Anderson mixing, which avoids additional computation of $f(\mathbf a)$, works well.
-Anderson mixing is one of several related quasi-Newtonian methods, which performs Newton-Raphson iteration $k$ with an approximation of the inverted inverted Jacobian labeled $G_k$.
-Thus, to solve for a variable $\mathbf z$ such that $f(\mathbf z)=\mathbf 0$, the best approximation for the zero is updated as $\mathbf z_{k+1}=\mathbf z_k - G_k f(\mathbf{z}_k)$ at each iteration.
-\par
-    Anderson mixing does this implicitly, with no extra recalculation of $f$, by approximating $f$ with the affine space spanning the previous $m$ evaluations of $f$.
-These values are $f(\mathbf z_{k})$ through $f(\mathbf z_{k-m+1})$.
-The best approximation for $f(\mathbf z)=0$ in the affine space which spans them is defined by
-    \[ \mathbf \alpha = \argmin_{\mathbf w:\, \sum_{i=1}^m w_i=1} \left\|\sum_{i=1}^m  w_i f(\mathbf z_{k-i+1}) \right\|_2 \]
-    This leads to the updated approximation
-    \[ \mathbf z_{k+1} = \sum_{i=1}^{m} \alpha_{i} \mathbf z_{k-i+1} \]
-    $G_k$ doesn't need to be calculated here, and the quickest implementation finds $\alpha$ without calculating it.
-However, it's presence can be seen through an equivalent formulation.
-Let $\Delta \mathbf f_i = f(\mathbf z_k) - f(\mathbf z_{k-i})$, $\Delta \mathbf z_i = \mathbf z_k - \mathbf z_{k-i}$,
-    \[ \mathcal F = [\Delta \mathbf f_1 \ldots \Delta \mathbf f_{m-1}] \]
-    and
-    \[ \mathcal Z = [\Delta \mathbf z_1 \ldots \Delta \mathbf z_{m-1}] \]
-    This means that $\mathbf\alpha$ has been replaced by $\mathbf\theta$ such that
-    \[ \mathbf \theta \; = \;\argmin_{\mathbf w} \left\|f(\mathbf z_{k}) - \mathcal  F \mathbf w\right\|_2\; = \;(\mathcal F^T \mathcal F)^{-1}\mathcal F^Tf(\mathbf z_k) \]
-    so 
-    \[ \mathbf z_{k+1} \;= \;\mathbf z_{k} - \mathcal Z \mathbf \theta\; = \; \mathbf z_{k} - \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T f(\mathbf z_k) \]
-    Which yields the approximated inverse Jacobian $G_k= \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T$
-        \end{enumerate}
-        In practice, it seemed that combining the first, third, and fourth of these variations leads to the most successful technique.
-However, should $n$ be large enough to provide a good approximation of the probability space, the second variation also saves a lot of time.
-
-\subsubsection{Fitting via Quadratic Programing}
-
-        Random binary knockoffs which produce the desired cross moment matrix $M_L$ need not arise from a generative model for $\mathbf{\tilde x}$ however.
-It is sufficient to have a random matrix $\tilde X \in \{0,1\}^{n\times p}$ such that
-        \[ \E_{\tilde X}(X^T \tilde X) = \Sigma - \diag{s} \quad \& \quad \E_{\tilde X}(\tilde X^T \tilde X) = \Sigma\]
-        The simplest way to do this is based on a probability matrix $P\in (0,1)^{n\times p}$ where each $X_{ij}$ is drawn independently as $X_{ij}\sim\textrm{Bernoulli}(P_{ij})$.
-This requirement can be stated as the quadratic feasibility problem
-            \begin{center}
-                \begin{tabular}{r l}
-                    exists     & $P$ \\
-                    subject to & $X^TP=\Sigma-\diag{s} $ \\
-                               & $P^TP=\Sigma $ \\
-                               & $0 \leq P \leq 1$
-                \end{tabular} 
-            \end{center}
-        If such a matrix exists and can be found, then $\tilde X$ can easily be generated.
-However, the constraints likely result in infeasibility much of the time.
-Instead, one can try to minimize how far the expected moments given $P$ would be from those desired for $M_L$.
-This would take the form
-            \begin{center}
-                \begin{tabular}{r l}
-                    minimize     & $\|X^TP-(\Sigma-\diag{s})\|_{fro}^2 + \|P^TP-\Sigma\|_{fro}^2 $\\
-                    subject to   & $ \mathbf 1^T P = \mathbf m_L $ \\
-                                 & $0 \leq P \leq 1$
-                \end{tabular} 
-            \end{center}
-            Here, the marginal distribution of $\tilde X_i$ will be the same as $X_i$, and the crossmoments shouldn't deviate to far from $M_L$.
-This problem can be stated as the quadratic program via the introduction of slack variables:
-            \begin{center}
-                \begin{tabular}{r l}
-                    minimize     & $\|W\|_{fro}^2 + \|V\|_{fro}^2 $ \\
-                    subject to   & $ -W \leq X^TP-(\Sigma-\diag{s})\leq W $ \\
-                                 & $ -V \leq P^TP-\Sigma \leq V $ \\
-                                 & $ \mathbf 1^T P = \mathbf m_L $ \\
-                                 & $0 \leq P \leq 1$
-                \end{tabular} 
-            \end{center}
-            Though there are nice tools and theory for solving quadratic programs, $n\times p$ would have to be fairly small for this optimization problem to be computationally reasonable in practice.
-Ideally, there would be a method to breakdown this problem into a series of smaller optimization problems which would yield the same solution.
-An optimist might hope that iterating through the columns of $P$, optimizing each to hit its desired crossmoments individually conditioned on the estimate of the rest of $P$, would achieve this.
-In practice, this does seem to yield reasonable results, and is far quicker than other methods for large $p$.
-In particular, the following method seemed to work well:
-    \begin{enumerate}
-        \item Optimize initial $P_i$ as such that $X_j^T P_i = M_{ij}$ if $i\neq j$ and $X_i^T P_i = \sigma_i(1-s)$.   
-            \begin{center}
-                \begin{tabular}{r l}
-                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_{ij})^2$ \\
-                    subject to   & $ \frac{1}{n}\mathbf 1^T P_i = M_{ii} $ \\
-                                 & $0 \leq P_i \leq 1$
-                \end{tabular} 
-            \end{center}
-        \item Draw initial $\tilde X_i$ as independent Bernoulli with probability $P_i$.
-        \item Iterate through $1\leq i \leq p$, fitting new $P_i$ such that 
-            \begin{center}
-                \begin{tabular}{r l}
-                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_{ij})^2 + \sum_{i\neq j}(\frac{1}{n}\tilde X_j^T P_i - M_{ij})^2$  \\
-                    subject to   & $ \frac{1}{n}\mathbf 1^T P_i = M_{ii} $ \\
-                                 & $0 \leq P_i \leq 1$
-                \end{tabular} 
-            \end{center}
-            After each $P_i$ is fit, redraw $\tilde X_i$ as independent Bernoulli with probability $P_i$.
-    \end{enumerate}
-    Drawing the $\tilde X_i$ after each step, rather than just leaving the $P_i$, seems counter intuitive, but under experimentation, it seemed to provide better results.
-As well, though the third step may be repeated several times, in practice this didn't lead to improvement.
-
-\section{Binary Knockoff Performance}
-    While the theory for original knockoffs and random binary variable generation suggests at how the binary knockoffs should perform, seeing their performance in practice still provides invaluable insight to their uses and limitations.
-There are three major areas in which this is useful; how well do the binary knockoffs' empirical cross moment matrices converge to their expectation, how well is FDR controlled for linear regression, and how well, if at all, is FDR controlled for other GLMs?
-
-\subsection{Convergence to Theoretical Crossmoments} 
-    \FloatBarrier
-        In a certain sense, the question of the convergence of the binary knockoffs' empirical crossmoments is the least interesting of these questions.
-Assuming the Ising model coefficient $A$ or the probability matrix $P$ was successfully fit, the crossmoment matrices' will converge to their expectation in probability for increasingly large values of $n$.
-However, especially with the limited numerical accuracy available to fit $A$ or $P$, the rate at which this convergence occurs and the size of the deviation is an issue of paramount importance.
-The original knockoff theory only provides guarantees when the crossmoments are replicated exactly.
-Thus, further examination of this convergence for various $n$ and $p$ is essential.
-
-    \FloatBarrier
-\subsection{Performance in Linear Regression} 
-    \FloatBarrier
-
-    \begin{figure}[h]
-        \begin{center}
-        \includegraphics[width=14cm]{images/lasso_FDR_power_50}
-    \end{center}
-    \caption{The mean FDR and power of binary and original knockoffs.
-Taken over $200$ simulations of linear regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ generated from Ising model while $y$ normally distributed about $X\beta$.}
-    \end{figure}
-
-    \begin{figure}[h]
-        \begin{center}
-        \includegraphics[width=7cm]{images/lasso_corr_50}
-    \end{center}
-    \caption{The mean correlation between variables selected using binary and original knockoffs.
-Taken over $200$ simulations of linear regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ generated from Ising model while $y$ normally distributed about $X\beta$.}
-    \end{figure}
-
-    \FloatBarrier
-\subsection{Performance in Logistic Regression} 
-    \FloatBarrier
-    \begin{figure}[h]
-        \begin{center}
-        \includegraphics[width=14cm]{images/logit_FDR_power_50}
-    \end{center}
-    \caption{The mean FDR and power of binary and original knockoffs.
-Taken over $200$ simulations of logistic regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ generated from Ising model while $y$ Bernoulli with $p=L(X\beta)$.}
-    \end{figure}
-
-    \begin{figure}[h]
-        \begin{center}
-        \includegraphics[width=7cm]{images/logit_corr_50}
-    \end{center}
-    \caption{The mean correlation between variables selected using binary and original knockoffs.
-Taken over $200$ simulations of logistic regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ generated from Ising model while $y$ Bernoulli with $p=L(X\beta)$.}
-    \end{figure}
-
-    \FloatBarrier
-
-\section{Discussion}
-    Binary knockoffs, as a practical tool for statistical inference, are still a work in progress.
-Currently, they work best for high values of $n$, for which the empirical crossmoment matrix deviates little, and relatively small values of $p$, for which the computational time is not prohibitive.
-Considering that FDR control in feature selection is most essential when $p$ is large, this is not a edifying state of affairs.
-However, the approach suggests paths forward to a much more complete tool.
-
-\subsection{Dealing With Small $n$}
-    With random generation of knockoffs, random variation from desired crossmoments is necessarily going to be an issue.
-However, there are likely ways to mitigate this problem by using multiple draws of $\tilde X$.
-For instance, if the $A$ or $P$ matrices were fit once, several draws of $\tilde X$ could be made from it, and a variable selection could be made for each of the corresponding $X_L$.
-Then these could be combined into a single variable selection, most obviously by selecting the variables which were selected for the majority of the $X_L$.
-
-\subsection{Computational Issues}
-    The biggest drawback of this method is its computational requirements.
-As $p$ increases, a larger and larger matrix $A$ must be fit for the Ising model as well as a larger $P$ for the quadratic program.
-For the Ising model, $K$ must similarly grow to provide accurate estimates of the probability space.
-This causes an explosion in computation time.
-Some of this can likely be mitigated by improvements in implementation, and the most time intensive issue, calculating $f(\mathbf a)$, can be easily parallelized.
-A simpler method would still be desirable though.
-A method based on optimizing each column of $P$ individually is thus appealing.
-Certainly, in cases such as the EM algorithm, methods have been developed which have strong guarantees for global optimization using such piecewise methods.
-If a similar development could be made here, or merely a method that hits the desired crossmoments more closely in practice, the result would be far more useful tool.
-
-\subsection{Random Binary Variables With Higher Order Interactions}
+\end{frame}
 

### 2015-07-29 masters_paper/presentation.tex
Getting to mu conditionals

@@
 
+\begin{frame}{Random Binary Notation}
+    \begin{itemize}
+        \item Binary data is common in data analysis and a much more manageable family of distributions.
+        \item We can think of observations in $X$ as observations of random binary vector $\mathbf x\in\{0,1\}^p$.
+        \item The full family for $\mathbf x$ is multinomial on $2^p$ outcomes.
+        \item Still useful to consider first two moments: 
+            \[ \E(\mathbf x) = \mathbf{m} \in [0,1]^p \quad \& \quad \E(\mathbf{x}\mathbf{x}^T) = M \in [0,1]^{p\times p} \]
+        \item For arbitrary $M$ to correspond to a random binary vector, must be case that $ M-\mathbf{m}\mathbf{m}^T = \Sigma \succeq0$
+            \[\max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\} \]
+
+    \end{itemize}
+\end{frame}
+
 \begin{frame}{Random Binary Knockoffs}
     \begin{itemize}
-        \item A useful 
+        \item Integer programing is np-hard, making finding finding $\tilde X\in\{0,1\}^{n\times p}$ to fit correlations exactly difficult. 
+        \item Instead, introduce a relaxed problem where $\tilde X\st X$ is drawn randomly such that, where $\Sigma=\cov(\mathbf x)$
+            \[ \cov(\mathbf{\tilde x}, \mathbf x) = \Sigma - \diag{\mathbf s} \quad \& \quad \cov(\mathbf{\tilde x}) = \Sigma \]
+        \item Almost same correlation condition as before, just only holds in expectation.
+        \item Switch from Gramian matrix to correlation matrix makes moment condition less likely to be violated. 
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Quadratic Programing}
+    \begin{itemize} 
+        \item Simplest approach to Random Binary Knockoffs is to draw the entries of $\tilde X$ independently based on $P\in[0,1]^{n\times p}$.
+        \item The best possible $P$ for the task would satisfy
+            \begin{center}
+                \begin{tabular}{r l}
+                    minimize     & $\|X^TP-(M-\diag{s})\|_{fro}^2 + \sum_{i\neq j}(P_i^T P_j - M_{ij})^2 $\\
+                    subject to   & $ \mathbf 1^T P = \mathbf m $ \\
+                                 & $0 \leq P \leq 1$
+                \end{tabular} 
+            \end{center}
+        \item Can be formulated as a quadratic program with slack variables
+            \begin{center}
+                \begin{tabular}{r l}
+                    minimize     & $\|W\|_{fro}^2 + \|V\|_{fro}^2 $ \\
+                    subject to   & $ -W \leq X^TP-(M-\diag{s})\leq W $ \\
+                                 & $ -V_{ij} \leq P_i^T P_j - M_{ij} \leq V_{ij} \quad \forall i\neq j$ \\
+                                 & $ \mathbf 1^T P = \mathbf m $ \\
+                                 & $0 \leq P \leq 1$
+                \end{tabular} 
+            \end{center}
+        \item Huge optimization problem, likely computationally impractical.
+
     \end{itemize}

### 2015-07-29 masters_paper/presentation.tex
Done presentation

@@
 \title{Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate}
-\author{Aaron Maurer}
+\author{Aaron Maurer \\ Advisor: Rina Foygel Barber}
 \date{July 29th, 2015}
@@
 
-\begin{frame}{Knockoffs}
+\begin{frame}{Knockoff Features}
     Knockoff variables can be used to control FDR in linear regression. 
@@
         \item The idea is to create a forgery of each variable; if the forgeries seem about as good predictors as the originals, the originals are lousy predictors.
-        \item For each variable $X_i$, create a knockoff feature $\tilde X_i$. Such that. 
-            \[ \tilde X^T \tilde X = X^T X \quad \& \quad X^T \tilde X = X^T X - \diag{\mathbf s} \]
-            Where $\diag{X^T X} - s$ is small but $\diag{X^T X} - s\succeq0$
+        \item For each variable $X_i$, create a knockoff feature $\tilde X_i$ such that, where $X^TX=G$, $\diag{X^T X} - s$ and 
+            \[ \tilde X^T \tilde X = G \quad \& \quad X^T \tilde X = G - \diag{\mathbf s} \]
+        \item For $\tilde X$ to exist, it must be the case that
+            \[ G_L =[X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right] \succeq 0 \]
         \item $\tilde X_i$ and $X_i$ will have same correlation with other variables, but only low correlation with each other.
@@
         \item $W_i= Z_i$ if $Z_i>\tilde Z_i$, otherwise $W_i = -\tilde Z_i$.
-        \item Since $[\, X \, \tilde X\,]^T[\, X \, \tilde X\,]$ \& $[\, X \, \tilde X\,]^T\mathbf y$ sufficient statistics for $\beta(\lambda)$, $W_i$ symmetrically distributed around $0$ when $X_i$ null predictor.
+        \item Since $G_L$ \& $[\, X \, \tilde X\,]^T\mathbf y$ sufficient statistics for $\beta(\lambda)$, $W_i$ symmetrically distributed around $0$ when $X_i$ null predictor.
         \item Thus, FDR controlled when $\hat S = \{i:W_i>T\}$ for 
@@
     \end{itemize}
+\end{frame}
+
+\begin{frame}{Variable Selection in GLMs}
+    Knockoffs work great for linear regression, but what about GLMs? \\
+    \vspace{1cm}
+    Now, assume, for some link function $g$ and $y_i,\ldots,y_n$ from a exponential family distribution,
+     \[\E(\mathbf{y}) = g(X\beta)\]
+    where $X \in \R^{n\times p}$ and $\beta\in\R^p$. Also, assume sparsity:
+    \[\beta_i = 0 \quad \forall i\not\in S\]
+    How do we pick estimate $\hat S$?
 \end{frame}
@@
     \begin{itemize}
-        \item Binary data is common in data analysis and a much more manageable family of distributions.
+        \item Binary data is common in data analysis and a much more manageable family of distributions for $X$.
         \item We can think of observations in $X$ as observations of random binary vector $\mathbf x\in\{0,1\}^p$.
@@
             \[ \cov(\mathbf{\tilde x}, \mathbf x) = \Sigma - \diag{\mathbf s} \quad \& \quad \cov(\mathbf{\tilde x}) = \Sigma \]
+        \item For this to correspond to a random binary vector, must be the case
+            \[ \Sigma_L =\cov([\,\mathbf x\,\mathbf{\tilde x}\,]) = \left[ \begin{array}{cc} \Sigma & \Sigma - \diag{\mathbf s} \\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \succeq 0 \]
         \item Almost same correlation condition as before, just only holds in expectation.
@@
             \end{center}
-        \item Huge optimization problem, likely computationally impractical.
+        \item Huge optimization problem, likely computationally difficult.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Ising Model}
+    \begin{itemize}
+        \item Instead, what if we found a random binary vector variable $\mathbf{x}_L$ that had cross-moments $M_L$ corresponding to $\Sigma_L$?
+        \item The Ising model can match any $M_L$. If $A$ lower triangular matrix and $L$ logistic link
+            \[ \p(\mathbf x=\mathbf \gamma) \propto L(\mathbf{\gamma}^T A\mathbf \gamma)\]
+        \item The Ising model binary analog of normal distribution; maximum entropy for given covariance matrix.
+        \item Very easy to draw successive entries
+            \[ \p(x_{i}=1\st x_{1},...,x_{i-1}) = L\left(A_{ii}+\sum_{k=1}^{i-1}A_{ik}x_i\right) \]
+        \item Once fit, can draw $\mathbf{\tilde x}\st \mathbf x$ easily.
 
@@
 
+\begin{frame}{Fitting Ising Model}
+    \begin{itemize}
+        \item If we were just trying to fit $A$ $X$, we could do so via successive logistic regression to fit row $\mathbf a_i$.
+        \item Instead, simulate $\mathbf m_i = f(\mathbf a_i)$ and fit via Newton-Raphson. Let $\mathbf x_{-i}^{(k)}\sim \mathbf x_{-i}$
+            \[f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right] \]
+            \[J\left( \mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
+        \item Make successive updates
+            \[\mathbf a_i^{(k+1)} = \mathbf a_i^{(k)} - \left[J\left(\mathbf a_i^{(k)}\right) \right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Computational Issues}
+    \begin{itemize}
+        \item $K$ must be very large for big $p$ and high correlation.
+        \item This makes $J\left( \mathbf a_i\right)$ and even $f\left(\mathbf a_i\right)$ very expensive to calculate.
+        \item This makes quasi-Newtonian methods, where $J\left( \mathbf a_i\right)$ is approximated appealing.
+        \item In particular, Anderson Mixing, where $f$ approximated with secant hyperplane through $\mathbf a_i^{k},\ldots,\mathbf a_i^{(k-h+1)}$ works well
+        \item When $K$ too small, can instead solve relaxed problem
+    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T \]
+        \item $n$ doesn't matter, but this method is also fairly impractical for large $p$.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Random Binary Knockoffs in Linear Regression}
+    Unfortunately, random binary Knockoffs only provide approximate FDR control for linear regression. 
+    \begin{center}
+        \includegraphics[width=11cm]{images/lasso_FDR_power_50}
+    \end{center}
+\end{frame}
+
+\begin{frame}{Distortion from $M_L$}
+    Since $\tilde X$ is randomly generated $\frac{1}{n}X_L^TX_L$ deviates from $M_L$.
+    \begin{center}
+        \includegraphics[width=9cm]{images/sigma_fit}
+    \end{center}
+\end{frame}
+
+\begin{frame}{Random Binary Knockoffs in Logistic Regression}
+    Still, Random Binary Knockoffs seem to do way better than original knockoffs in logistic regression.
+    \begin{center}
+        \includegraphics[width=11cm]{images/logit_FDR_power_50}
+    \end{center}
+\end{frame}
+
+\begin{frame}{Discussion}
+    \begin{itemize}
+        \item Random Binary Knockoffs seem to offer promise as a useful technique, but have outstanding issues.
+        \item Seem to offer a method to extend Knockoffs for one type of variable to GLMs.
+        \item Computational complexity prohibitive; simpler method, perhaps by good approximation of $P$, would be helpful.
+        \item Random distortion from desired cross-moments might be compensated for by ensemble method based on multiple $\tilde X$.
+        \item Might build higher order interactions into Ising model to allow for nasty data.
+    \end{itemize}
+\end{frame}
+
 \end{document}

### 2015-07-29 masters_paper/presentation.tex
Edited presentation

@@
     \[\beta_i = 0 \quad \forall i\not\in S\]
-    How do we pick estimate $\hat S$?
+    How do we choose a model $\hat S$?
 \end{frame}
 
-\begin{frame}{False Discover Rate}
+\begin{frame}{False Discovery Rate}
     A common goal for a method that generates $\hat S$ is to control the false discovery rate
     \[ \textrm{FDR} = \E\left[\frac{\vert{\{j: \beta_j=0 \; \& \; j\in\hat S\}}\vert}{\max\{\vert{\hat S}\vert,1\}} \right] \]
-    In other words, control portion of elements in $\hat S$ which aren't in $S$. \par
+    In other words, control proportion of elements in $\hat S$ which aren't in $S$. \par
     \vspace{1cm}
-    FDR is controlled at level $q$ if $q<$FDR irrespective of true $\beta$.
+    FDR is controlled at level $q$ if $q\leq$FDR irrespective of true $\beta$.
 
@@
     \begin{itemize}
-        \item Fit full path of LASSO regression on $[\,X\,\tilde X\,]$.
+        \item Fit full path of LASSO regression on $[\,X\,\tilde X\,]:=X_L$.
     \[ \beta(\lambda) = \argmin_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]
-        \item $Z_i$, $\tilde Z_i$ largest $\lambda$ such that $X_i$, $\tilde X_i$ have nonzero coefficient.
+        \item $Z_i$, $\tilde Z_i$ are the largest $\lambda$ such that $X_i$, $\tilde X_i$ have nonzero coefficient.
         \item $W_i= Z_i$ if $Z_i>\tilde Z_i$, otherwise $W_i = -\tilde Z_i$.
-        \item Since $G_L$ \& $[\, X \, \tilde X\,]^T\mathbf y$ sufficient statistics for $\beta(\lambda)$, $W_i$ symmetrically distributed around $0$ when $X_i$ null predictor.
-        \item Thus, FDR controlled when $\hat S = \{i:W_i>T\}$ for 
+        \item Since $G_L$ \& $[\, X \, \tilde X\,]^T\mathbf y$ are sufficient statistics for $\beta(\lambda)$, $W_i$ symmetrically distributed around $0$ when $X_i$ null predictor.
+        \item Thus, FDR controlled when $\hat S = \{i:W_i\geq T\}$ for 
             \[ T = \min\left\{ t>0 \;: \; \frac{\vert\{j:W_j\leq -t\}\vert}{\max\{\vert\{j:W_j\geq t\}\vert,1\}}\leq q \right\} \]
@@
     \vspace{1cm}
-    Now, assume, for some link function $g$ and $y_i,\ldots,y_n$ from a exponential family distribution,
+    Now, assume, for some link function $g$ and $y_1,\ldots,y_n$ from a exponential family distribution,
      \[\E(\mathbf{y}) = g(X\beta)\]
@@
     \[\beta_i = 0 \quad \forall i\not\in S\]
-    How do we pick estimate $\hat S$?
+    How do we choose a model $\hat S$?
 \end{frame}
@@
 \begin{frame}{Where Knockoff Filter Fails}
-    Knockoff filter don't work for other GLMs. \\
+    Knockoff filter doesn't work for other GLMs. \\
     \begin{center}
@@
         \item The full family for $\mathbf x$ is multinomial on $2^p$ outcomes.
-        \item Still useful to consider first two moments: 
-            \[ \E(\mathbf x) = \mathbf{m} \in [0,1]^p \quad \& \quad \E(\mathbf{x}\mathbf{x}^T) = M \in [0,1]^{p\times p} \]
+        \item Can summarize with the first two moments: 
+            \[ \E(\mathbf{x}\mathbf{x}^T) = M \in [0,1]^{p\times p} \quad \& \quad \diag{M} = \E(\mathbf x) = \mathbf{m} \in [0,1]^p  \]
         \item For arbitrary $M$ to correspond to a random binary vector, must be case that $ M-\mathbf{m}\mathbf{m}^T = \Sigma \succeq0$
@@
     \begin{itemize}
-        \item Integer programing is np-hard, making finding finding $\tilde X\in\{0,1\}^{n\times p}$ to fit correlations exactly difficult. 
+        \item Integer programing is NP-hard, making finding $\tilde X\in\{0,1\}^{n\times p}$ to fit correlations exactly difficult. 
         \item Instead, introduce a relaxed problem where $\tilde X\st X$ is drawn randomly such that, where $\Sigma=\cov(\mathbf x)$
@@
             \[ \Sigma_L =\cov([\,\mathbf x\,\mathbf{\tilde x}\,]) = \left[ \begin{array}{cc} \Sigma & \Sigma - \diag{\mathbf s} \\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \succeq 0 \]
-        \item Almost same correlation condition as before, just only holds in expectation.
+        \item Almost same correlation condition as before, but only holds in expectation.
         \item Switch from Gramian matrix to correlation matrix makes moment condition less likely to be violated. 
@@
     \begin{itemize} 
-        \item Simplest approach to Random Binary Knockoffs is to draw the entries of $\tilde X$ independently based on $P\in[0,1]^{n\times p}$.
-        \item The best possible $P$ for the task would satisfy
+        \item Simplest approach to Random Binary Knockoffs is to draw the entries of $\tilde X$ independently based on probabilities $P\in[0,1]^{n\times p}$.
+        \item $P$ should satisfy 
             \begin{center}
@@
     \begin{itemize}
-        \item Instead, what if we found a random binary vector variable $\mathbf{x}_L$ that had cross-moments $M_L$ corresponding to $\Sigma_L$?
-        \item The Ising model can match any $M_L$. If $A$ lower triangular matrix and $L$ logistic link
+        \item Alternatively, find a random binary vector variable $\mathbf{x}_L$ that has cross-moments $M_L$ corresponding to $\Sigma_L$.
+        \item The Ising model can match any proper $M_L$. If $A$ is a lower triangular matrix and $L$ the logistic link function
             \[ \p(\mathbf x=\mathbf \gamma) \propto L(\mathbf{\gamma}^T A\mathbf \gamma)\]
@@
     \begin{itemize}
-        \item If we were just trying to fit $A$ $X$, we could do so via successive logistic regression to fit row $\mathbf a_i$.
-        \item Instead, simulate $\mathbf m_i = f(\mathbf a_i)$ and fit via Newton-Raphson. Let $\mathbf x_{-i}^{(k)}\sim \mathbf x_{-i}$
+        \item If we were just trying to fit $A$ to $X$, we could do so via successive logistic regression to fit row $\mathbf a_i$.
+        \item Instead, simulate $\mathbf m_i = f(\mathbf a_i)$ and fit via Newton-Raphson. 
+        \item Draw $K$ samples $\mathbf x_{-i}^{(k)}\sim \mathbf x_{-i}$. Then
             \[f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right] \]
@@
         \item This makes $J\left( \mathbf a_i\right)$ and even $f\left(\mathbf a_i\right)$ very expensive to calculate.
-        \item This makes quasi-Newtonian methods, where $J\left( \mathbf a_i\right)$ is approximated appealing.
+        \item This makes quasi-Newtonian methods, where $J\left( \mathbf a_i\right)$ is approximated, appealing.
         \item In particular, Anderson Mixing, where $f$ approximated with secant hyperplane through $\mathbf a_i^{k},\ldots,\mathbf a_i^{(k-h+1)}$ works well
@@
 \begin{frame}{Random Binary Knockoffs in Linear Regression}
-    Unfortunately, random binary Knockoffs only provide approximate FDR control for linear regression. 
+    Random binary Knockoffs only provide approximate FDR control for linear regression. 
     \begin{center}
@@
     \end{center}
+    They select too many variables, causing higher power at the expense of FDR control.
 \end{frame}
@@
 \begin{frame}{Distortion from $M_L$}
-    Since $\tilde X$ is randomly generated $\frac{1}{n}X_L^TX_L$ deviates from $M_L$.
+    Since $\tilde X$ is randomly generated, $\frac{1}{n}X_L^TX_L$ deviates from $M_L$.
     \begin{center}
-        \includegraphics[width=9cm]{images/sigma_fit}
+        \includegraphics[width=8cm]{images/sigma_fit}
     \end{center}
+    As $n$ increases, this effect diminishes. 
 \end{frame}
@@
 \begin{frame}{Random Binary Knockoffs in Logistic Regression}
-    Still, Random Binary Knockoffs seem to do way better than original knockoffs in logistic regression.
+    Random Binary Knockoffs perform much better than original knockoffs in logistic regression.
     \begin{center}
@@
     \end{center}
+    They have lower FDR and higher power.
 \end{frame}

### 2015-07-29 masters_paper/presentation.tex
Edited presentation 2

@@
         \item The idea is to create a forgery of each variable; if the forgeries seem about as good predictors as the originals, the originals are lousy predictors.
-        \item For each variable $X_i$, create a knockoff feature $\tilde X_i$ such that, where $X^TX=G$, $\diag{X^T X} - s$ and 
+        \item For each variable $X_i$, create a knockoff feature $\tilde X_i$ such that, where $X^TX=G$, $\diag{X^T X} - s$ is small and 
             \[ \tilde X^T \tilde X = G \quad \& \quad X^T \tilde X = G - \diag{\mathbf s} \]
+        \item $\tilde X_i$ and $X_i$ will have same correlation with other variables, but only low correlation with each other.
         \item For $\tilde X$ to exist, it must be the case that
             \[ G_L =[X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right] \succeq 0 \]
-        \item $\tilde X_i$ and $X_i$ will have same correlation with other variables, but only low correlation with each other.
         \item Given $\mathbf s$, $\tilde X$ can be generated via a rotation of $X$.
@@
     \begin{itemize}
-        \item Integer programing is NP-hard, making finding $\tilde X\in\{0,1\}^{n\times p}$ to fit correlations exactly difficult. 
+        \item Integer programing is NP-hard, making finding $\tilde X\in\{0,1\}^{n\times p}$ to fit correlations optimally difficult. 
         \item Instead, introduce a relaxed problem where $\tilde X\st X$ is drawn randomly such that, where $\Sigma=\cov(\mathbf x)$
@@
     \begin{itemize} 
-        \item Simplest approach to Random Binary Knockoffs is to draw the entries of $\tilde X$ independently based on probabilities $P\in[0,1]^{n\times p}$.
+        \item Simplest approach to is to draw the entries of $\tilde X$ independently based on probabilities $P\in[0,1]^{n\times p}$.
         \item $P$ should satisfy 

### 2015-07-30 masters_paper/paper.tex
Done with discussion
Updating section 1

@@
 Variable selection is an essential problem to fitting a regression model.
-For linear regression, the knockoff filter (Barber and Cand\`{e}s, 2014) offers a method of exact FDR control.
-However, the method as originally formulated does not extend to other generalized linear models such as logistic regression.
-This paper offers an extension of this method which will work with binary features in the context of linear regression, but also will extend to other GLMs.
+For linear regression, the knockoff filter \cite{2014arXiv1404.5609F} offers a method of exact FDR control.
+However, the method as originally formulated does not extend to other generalized linear models (GLMs) such as logistic regression.
+This paper offers an extension of this method for binary features in the context of both GLMs as well as linear regression.
 
-\subsection{Knockoff Filter}
-    Let us consider the consider the usual setting for linear regression, where $n$ observations of a variable of interest $y$ arise from the model
+\subsection{Background: the Knockoff Filter}
+    Let us consider the usual setting for linear regression, where $n$ observations of a variable of interest $y$ arise from the model
     \[\mathbf{y} = X\beta + \mathbf{z}\]
-    where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise.
-It is important to note that a intercept vector $\mathbf 1$ is not included in $X$.
+    where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise.It is important to note that a intercept vector $\mathbf 1$ is not included in $X$.
 For the purpose of this paper, only the case where $2p\leq n$ will be considered, though the methods can be extended to $p\leq n$.
@@
 The random variable $x_j$ is in turn $j$th entry of the random vector valued variable $\mathbf x$.
-In the situation where $\beta$ is sparse and there is reason to expect that $x_j$ and $y$ are uncorrelated for many $j$, the knockoff filter is a method to select a subset of the $x_j$ which are likely correlated with $y$.
+Let $\beta$ be sparse, implying only a subset of the features $x_j$ have an effect on  $y$.
+In other words, there is a true model $S\subseteq \{1,\ldots,p\}$ such that $\beta_i\neq0$ if and only if $i\in S$.
 \par
-    The method specifically seeks to control the false discovery rate (FDR), which is the expected portion of some selection of $x_j$ which are, in fact, uncorrelated with $y$.
-For a procedure which chooses a subset of variables $\hat S\subseteq \{1,\ldots,p\}$, this is defined as
-    \[ \textrm{FDR} = \E\left[\frac{\vert{\{j: \beta_j=0 \; \& \; j\in\hat S\}}\vert}{\max\{\vert{\hat S}\vert,1\}} \right] \]
-    FDR is controlled at level $q$ if FDR is less than $q$ irrespective of the coefficients $\beta$. \par
+     In this context, the knockoff filter is designed to choose a model $\hat S$ so as to control the false discovery rate (FDR), or the portion of variables chosen by $\hat S$ which aren't in the true model $S$:
+    \[ \textrm{FDR} = \E\left[\frac{\vert{\{j: j\not\in S \; \& \; j\in\hat S\}}\vert}{\max\{\vert{\hat S}\vert,1\}} \right] \]
+    In other words, FDR is the portion of features which are thought to have an effect on $y$ but actually have no effect.
+FDR is controlled at level $q$ if FDR is less than $q$ irrespective of the coefficients $\beta$.
+\par
 
-    The knockoff filter achieves FDR control in several steps, the first of which is creating a set of `knockoff' features $\tilde X_j$ which imitate the original $X_j$ while being less correlated with $y$.
-In particular, the matrix of knockoffs $\tilde X$ has the same internal correlation structure as $X$, the knockoff feature $\tilde X_j$ has the same correlation with other $X_i$ as $X_j$ does, but the correlation between $\tilde X_j$ and $X_j$ is minimized.
-In other words, where $ G := X^TX$,
+    The knockoff filter achieves FDR control in several steps, the first of which is creating a set of `knockoff' features $\tilde X_j$ which imitate the original $X_j$ while being no more correlated with $y$.
+In particular, the matrix of knockoffs $\tilde X$ has the same internal correlation structure as $X$ and the knockoff feature $\tilde X_j$ has the same correlation with other $X_i$ as $X_j$ does, but the correlation between $\tilde X_j$ and $X_j$ is minimized.
+In other words, for $G := X^TX$,
     \[ \tilde X^T\tilde X=G \textrm{ and } \tilde X^T X = G - \diag{\mathbf s}\]
-    for some vector $s\in \R^p$ such that, where $\diag{G}$ is the vector made from the main diagonal of $G$, $\diag{G} - \mathbf s$ is small.
+    for some vector $\mathbf{s}\in \R^p$ such that, writing $\diag{G}$ as the vector made from the main diagonal of $G$, $\diag{G} - \mathbf s$ is small.
 Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$.
-In the case where $X_j$ is uncorrelated with $\mathbf y$, $\tilde X_j$ will also be uncorrelated with $\mathbf y$.
+In the case where $X_j$ is a null predictor of $\mathbf y$, $\tilde X_j$ will thus also be a null predictor of $\mathbf y$.
 How these knockoffs are actually generated will be described shortly.\par
-    The second step is to fit a series of Lasso model of $\mathbf y$ on the combined design matrix $X_L=[X \; \tilde X]$ so as to determine the largest value $\lambda$ at which the coefficient for each $X_j$ and $\tilde X_j$ is nonzero.
-The subscript $L$ refers to the large design matrix with both features and knockoffs.
-\footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} Recall, for a given $\lambda$, the estimated coefficient from the Lasso regression will be
-    \[ \beta(\lambda) = \argmin_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]
-    In the general case, the coefficient for one feature in a Lasso regression being first nonzero for a larger value of $\lambda$ than another feature is an indication that the first feature is a stronger predictor of the outcome than the second.
-Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to have nonzero coefficients sooner than the knockoff features when the original is a valid predictor.
-On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $G_L=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
+    The second step is to fit a series of LASSO models of $\mathbf y$ on the combined design matrix $X_L:=[X \; \tilde X]$ so as to determine the largest value $\lambda$ at which the coefficient for each of $X_j$ and $\tilde X_j$ is nonzero.\footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} The subscript $L$ refers to the large design matrix with both features and knockoffs.
+Recall, for a given $\lambda$, the estimated coefficient from the LASSO regression will be
+    \[ \beta(\lambda) = \argmin_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|\mathbf{b}\|_1 \right\}\]
+    In the general case, the ``LASSO path" refers to the path of solutions as $\lambda$ is decreased from $\infty$ to $0$, and we say that a features ``enters the path" at $\lambda$ when its coefficient is first estimated to be nonzero.
+One feature entering the LASSO path before another is an indication that the first feature is a stronger predictor of the outcome than the second.
+Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to enter the LASSO path sooner than the knockoff features when the original is a valid predictor.
+On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $G_L:=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
 This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $G_L$ will be unaltered, while $\E\left[X_L^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$.
@@
     \item Section 3 will develop the method and theory for generating binary knockoffs.
-    \item Section 4 will discuss tests of Binary knockoffs in simulation, both in comparison to the original knockoffs with Lasso and for logistic regression by itself.
+    \item Section 4 will discuss tests of Binary knockoffs in simulation, both in comparison to the original knockoffs with LASSO and for logistic regression by itself.
     \item Section 5 is the final section, and will contain discussion of the results as well as areas for further work.
@@
 On the right side, they are drawn as Bernoulli with probability $.25$ of being 1.
-The top has 50 features uncorrelated with the outcome, while the bottom has 40 features uncorrelated with the outcome and 10 correlated with it.}
+The top has 50 null features, while the bottom has 40 null features with the outcome and 10 correlated with it.}
     \end{figure}
@@
 This, or some other sort of average across multiple realizations of $\tilde X$, could likely be used to dampen the effect of variance in the generation of $\tilde X$.
-This would at least make binary knockoffs reliable for Lasso.
+This would at least make binary knockoffs reliable for LASSO.
 \par
@@
 Nonetheless, this may be an avenue for further investigation.
+\par
 

### 2015-07-30 masters_paper/paper.tex
Penultimate for sections 1-3

@@
 How these knockoffs are actually generated will be described shortly.\par
-    The second step is to fit a series of LASSO models of $\mathbf y$ on the combined design matrix $X_L:=[X \; \tilde X]$ so as to determine the largest value $\lambda$ at which the coefficient for each of $X_j$ and $\tilde X_j$ is nonzero.\footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} The subscript $L$ refers to the large design matrix with both features and knockoffs.
+    The second step is to fit a series of LASSO models of $\mathbf y$ on the combined design matrix $X_{aug}:=[X \; \tilde X]$ so as to determine the largest value $\lambda$ at which the coefficient for each of $X_j$ and $\tilde X_j$ is nonzero.\footnote{Other statistics besides the $\lambda$s can also be used in the knockoff method, but are not considered in this paper} The subscript $aug$ refers to the augmented design matrix with both features and knockoffs.
 Recall, for a given $\lambda$, the estimated coefficient from the LASSO regression will be
-    \[ \beta(\lambda) = \argmin_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|\mathbf{b}\|_1 \right\}\]
+    \[ \beta(\lambda) = \argmin_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_{aug}\mathbf b\|^2_2 + \lambda\|\mathbf{b}\|_1 \right\}\]
     In the general case, the ``LASSO path" refers to the path of solutions as $\lambda$ is decreased from $\infty$ to $0$, and we say that a features ``enters the path" at $\lambda$ when its coefficient is first estimated to be nonzero.
@@
 Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to enter the LASSO path sooner than the knockoff features when the original is a valid predictor.
-On the other hand, since $\beta(\lambda)$ only depends on $X_L$ through the sufficient statistics $G_L:=X_L^TX_L$ and $X_L^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
-This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $G_L$ will be unaltered, while $\E\left[X_L^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$.
+On the other hand, since $\beta(\lambda)$ only depends on $X_{aug}$ through the sufficient statistics $G_{aug}:=X_{aug}^TX_{aug}$ and $X_{aug}^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
+This can be seen by switching a null $X_j$ with $\tilde X_j$; by construction, $G_{aug}$ will be unaltered, while $\E\left[X_{aug}^T\mathbf y\right]$ will also be unchanged, since $\E[X_j^T\mathbf y]=\E[\tilde X_j^T\mathbf y]=0$.
 \par
-   This observation leads to the final step.
-Let $Z_j$ be the largest $\lambda$ such that feature $j$ has a nonzero coefficient and $\tilde Z_j$ the largest $\lambda$ such that knockoff $j$ has a nonzero coefficient.
-Then, define $W_j$ as
+   This observation leads to the final step. Let $Z_j$ be the point at which $X_j$ entered the path and $\tilde Z_j$ the point at which $\tilde X_j$ entered the path. Then, define $W_j$ as 
    \[ W_j = \begin{cases}   Z_j         &\mbox{if } Z_j>\tilde Z_j \\ 
@@
             \end{cases} \]
-            When $X_j$ is a null predictor, $W_j$ will be symmetrically distributed around zero, and when $X_j$ is true predictor, it will have a distribution skewed positive.
-Since $W_j$ with large positive values are more likely to be true predictors, the knockoff filter selects the variables $\{j:W_j\geq T\}$, where $T$ is a threshold defined as
+            $W_j$ is positive when $Z_j$ entered the path first and negative when $\tilde X_j$ entered the path first.
+Large, positive values for $W_j$ will indicate that $X_j$ is a strong predictor, since $X_j$ entered the path early and before $\tilde X_j$.
+Thus, we will make a selection $\hat S=\{j:W_j\geq T\}$ for some threshold $T$.
+When $X_j\not\in S$, $X_j$ and $\tilde X_j$ are both equally likely to enter the path first, so $W_j$ is equally as likely to be positive as negative for $j\in S$.
+Thus,
+            \[ \vert\{j:W_j\geq T \; \& \; j\not\in S\}\vert \approx \vert\{j:W_j\leq -T \; \& \; j\not\in S\}\vert] \]
+            The quantity on the left is, of course, the number of false discoveries made by choosing a particular threshold $T$. Also, it must be the case
+            \[\vert \{j:W_j\leq -T \; \& \; j\not\in S\}\vert \leq \vert\{j:W_j\leq -T\}\vert\]
+            This inequality should be pretty tight, since a true predictor should be a better predictor than its knockoff and unlikely to enter the path after it.
+Accordingly, the expected portion of false discoveries out of all discoveries is approximately
+            \[FDR \approx \E\left[\frac{\vert\{j:W_j\leq -T\}\vert}{\max\{\vert\{j:W_j\geq T\}\vert,1\}}\right]\] 
+            So, when $T$ is selected as
             \[ T = \min\left\{ t>0 \;: \; \frac{\vert\{j:W_j\leq -t\}\vert}{\max\{\vert\{j:W_j\geq t\}\vert,1\}}\leq q \right\} \] 
-            Basic logic of this method is fairly simple.
-Let $N_h(t)$ be the random variable for the number of null predictors with $W_j\leq t$, $N_l(t)$ be the number of null predictors with $W_j\geq t$, and $V_l(t)$ \& $V_h(t)$ be the similar number of true predictors.
-Since the null $W_j$ are symmetrically distributed around $0$, $\E[N_l(t)]=\E[N_h(t)]$.
-Thus,
-            \begin{align*}
-                q &\geq \E\left[ \frac{\vert\{j:W_j\leq -t\}\vert}{\max\{\vert\{j:W_j\geq t\}\vert,1\}} \right] \\
-                q &\geq \E\left[ \frac{N_l(t) + V_l(t)}{\max\{N_h(t) + V_h(t),1\}} \right] \\
-                q &\geq \E\left[ \frac{N_l(t)}{\max\{N_h(t) + V_h(t),1\}} \right] \\
-                q &\geq \E\left[ \frac{N_h(t)}{\max\{N_h(t) + V_h(t),1\}} \right] \\
-                q &\geq \textrm{FDR}
-            \end{align*}
-            So FDR is controlled, and since $V_l(t)$ will tend to be small, it should be controlled fairly tightly.
+            FDR should be controlled at level $q$.
 
@@
     \[ \tilde X^T\tilde X=G \textrm{ and } \tilde X^T X = G - \diag{\mathbf s}.\]
-    For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. Then, let the Gram matrix of $X_L$ is
-    \[ G_L= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right] \]
-    $A$, the Schur complement of $G$ in $G_L$ can be calculated as 
+    For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. The Gram matrix of $X_{aug}$ is
+    \[ G_{aug}= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right] \]
+    $A$, the Schur complement of $G$ in $G_{aug}$ can be calculated as 
     \[ A = 2\,\diag{\mathbf s} - \diag{\mathbf s}G^{-1}\diag{\mathbf s} \]
-    For $G_L$ to exist, $G_L$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if\footnote{A fleshed out version of this argument can be found in the original knockoff paper}
+    For $G_{aug}$ to exist, $G_{aug}$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if 
     \[ \diag{\mathbf s} \succeq 0  \; \textrm{ and } \; 2G - \diag{\mathbf s} \succeq 0 \]
@@
 \subsection{Binary Knockoffs}
-    Though the ``original" knockoff method, described above, achieves the exact desired knockoff properties, the individual values in the vector $X_j$ will have little relation to the individual values in $X_j$.
+    Though the ``original" knockoff method, described above, achieves the exact desired correlation properties, the individual values in the vector $X_j$ will have little relation to the individual values in $X_j$.
 In particular, these values will often have very different empirical distributions.
@@
 \par
-    Thus, this paper offers a new method of generating knockoffs which should offer superior performance with other GLMs.
-In this new method, the matrix of knockoff features $\tilde X$ will be generated by row from a random vector valued variable $\mathbf{\tilde x}$, each entry $\tilde x_j$ of which will have the property $\tilde x_j \sim x_j$.
-This random variable will also abide by a relaxation of the original knockoff condition:
-    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=G \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = G - \diag{\mathbf s}\]
-    This will ensure that, at the very least, these knockoffs are a suitable replacement for the original knockoffs for linear regression. \par
-    Generating such knockoffs for a completely general case of any $\mathbf x$ is likely infeasible.
-However, in the specific case where $\mathbf x$ is a random binary vector variable, this paper will demonstrate a method to generate such knockoff variables.
-Since binary data is extremely common in many areas, this will hopefully serve as a useful tool.
+    Thus, this paper offers a new method of generating knockoffs for binary data which should offer superior performance with other GLMs.
+In this new method, the matrix of knockoff features $\tilde X$ will also be binary, generated randomly so as to hit the new correlation condition
+    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=\Sigma \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = \Sigma - \diag{\mathbf s}\]
+    Where $\Sigma$ is the covariance matrix of $X$.
+This is essentially relaxation of the original correlation condition, requiring the desired covariance to only hold in expectation.
+For large values of $n$, the observed covariance will converge to its expectation, making this distinction minor.
 
@@
 \begin{itemize}
-    \item Section 2 will go into more detail about how the original knockoffs breakdown with generalized linear models. In particular, several simulations will demonstrate how they breakdown.
-    \item Section 3 will develop the method and theory for generating binary knockoffs.
-    \item Section 4 will discuss tests of Binary knockoffs in simulation, both in comparison to the original knockoffs with LASSO and for logistic regression by itself.
+    \item Section 2 will go into more detail about how the original knockoffs break down with generalized linear models.
+In particular, several simulations will demonstrate how they fail to control FDR.
+    \item Section 3 will develop the method for generating binary knockoffs.
+    \item Section 4 will discuss tests of binary knockoffs in simulation, comparing their performance to the original knockoffs for both linear and logistic regression. 
     \item Section 5 is the final section, and will contain discussion of the results as well as areas for further work.
@@
 Here, where $l(\beta\st X, \tilde X)$ is the log likelihood of coefficients $\beta$ given the model, the $L1$ regularized regression model will have estimated coefficient vector
-    \[ \beta(\lambda) = \argmin_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
+    \[ \beta(\lambda) = \argmin_\mathbf b \left\{l(\mathbf{b}\st X, \tilde X) + \lambda\|\mathbf{b}\|_1 \right\}\]
     The $Z_j$ and $\tilde Z_j$ are then, once again, the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before.
@@
 On the right side, they are drawn as Bernoulli with probability $.25$ of being 1.
-The top has 50 null features, while the bottom has 40 null features with the outcome and 10 correlated with it.}
+The top has 50 null features, while the bottom has 10 features in the true model, and 40 null features.}
     \end{figure}
 
-    To test how bad this problem is, I tested the method using original style knockoffs in logistic regression.
+    To evaluate whether this made a difference in practice, I tested the method using original style knockoffs in logistic regression.
 The results can be seen in figure 1, which shows cumulatively, starting with the largest absolute value, what portion of the $W_j$ are positive.
@@
 This is particularly important for high values, which play the largest roll in selecting a threshold.
-We see that in the top left corner, when all features are null and normally distributed, that we almost achieve this, with only a small, but consistent, deviation from $50\%$.
-However, in the top right, when the predictors are binary, the null originals enter first far more often, which will result in the FDR being too high.
+In the top left corner, under the ``global null", where all features are null and normally distributed, that we almost achieve this.
+There is just a small, but consistent, deviation from $50\%$.
+However, in the top right, when the predictors are binary but still all null, the originals enter first far more often, which will result in the FDR being too high.
 \par
@@
 This will completely ruin FDR control, since the portion of null predictors will be assumed far lower than it actually is.
-It is worth noting that the particular shape of the curve in the diagram, with the plateau for the null predictors through 10, is partially an artifact of how the totals were tabulated.
+It is worth noting that the particular shape of the curve in the diagram, with the plateau for the null predictors through 10, is partially an artifact of there being 10 true signals.
 The mean through the first ten is the mean for null predictors that came in first out of both null and true predictors; since the largest ten $W_j$ will tend to be true predictors, the average over the null predictors represent the few extreme $W_j$ which managed to beat out some number of true predictors.
@@
     \begin{figure}[h]
+        \begin{center}
             \includegraphics[width=14cm]{images/normalQQ}
             \caption{Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
+        \end{center}
     \end{figure}
-
     Upon examination, the issue seems to be that the original style knockoff features have a very different empirical distribution than the features they are imitating.
-This is obviously true with the binary features, but even with the normal features, the distributions differ a lot.
-In figure 2, the distribution of four feature and the original style knockoff are plotted against a normal distribution in Q-Q plots.
-As you can see, they don't match at all.
+This is obviously true with the binary features, but even with the normal features, the distributions differ substantially.
+In figure 2, the distribution of four features and their original style knockoff are plotted against a normal distribution in Q-Q plots.
+As you can see, the distributions differ substantially.
 This is suggestive of a solution; if the knockoff variables match the marginal distribution of the original features better, the result may be more suitable for FDR control in GLMs.
@@
 \section{Random Binary Knockoffs}
-Random binary knockoffs $\mathbf{\tilde x}$, which maintain the correlation condition in expectation, fit the bill in this regard.
-The method to produce them is based heavily on more general methodology to random binary vector variables with internal correlation.
+Coming up with knockoffs which would match the original features' marginal distributions better for arbitrary $X$ is daunting problem.
+Limiting ourselves to binary data, which is common in real world applications, presents a far more manageable case.
+However, solving for a binary $\tilde X$ which matches the correlation condition exactly or closely is an integer programing problem, which are generally NP-hard.
+Thus, instead, this section will develop methods of generating $\tilde X$ randomly such that the correlation condition holds in expectation.
+Mainly, $\tilde X\st X$ should be a random matrix such that
+        \[ \E(X^T \tilde X \st X) = X^TX - \diag{s} \quad \& \quad \E(\tilde X^T \tilde X \st X) = X^TX\]
 
-\subsection{Random Binary Vector Generation}
+\subsection{Random Binary Vector Variables}
+We can imagine each row of $X$ being an observation of a random binary vector variable $\mathbf x$.
 The full class of random binary vector variables on $\{0,1\}^p$ can be specified as multinomial on the $2^p$ elements of the set.
-However, in practice, it is generally impractical to specify or estimate all $2^p$ necessary parameters, so it is useful to, as is desirable for knockoffs, pick a method to match the first two moments.
-These are the mean vector $\E(\bf x)=m\in(0,1)^p$ and the cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$.
+However, in practice, it is generally impractical to specify or estimate all $2^p-1$ necessary parameters, so it is useful to pick a method to match the first two moments, as is desirable for knockoffs.
+These are the mean vector $\E(\bf x)=m\in[0,1]^p$ and the cross-moment matrix $\E(\bf x\bf x^T)=M\in[0,1]^{p\times p}$.
 Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag{M}$.
-For an arbitrary symmetric $M$ to be valid cross-moment matrix for some random binary vector variable, $M-mm^T\succeq0$, and
+For an arbitrary symmetric $M$ to be a valid cross-moment matrix for some random binary vector variable, $M-mm^T\succeq0$, and
     \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
-    for all $i\neq j.$ Given a qualifying $M$, or observed $X$, there are then a few methods for generating additional random values.
+    for all $i\neq j$\cite{2011arXiv1111.0576S}.
 
-\subsubsection{Gaussian Copula Family}
-Since multivariate normal distributions are easy to randomly draw from and defined by their first two moments, an obvious choice is to use them to generate random binary vectors.
-This might be done by finding $\mu$ and $\Sigma$ such that for $\mathbf z\sim N_p(\mu,\Sigma)$, $x_i$ can be generated as $x_i = \mathrm{sign}(z_i)$, and have the desired first two moments.
-However, this is only guaranteed to be feasible in the bivariate case.
-For higher dimensions, this method can generally only provide random binary vectors with approximately the right moments, and the quality of the approximation quickly degenerates as $p$ increases.
-
-\subsubsection{$\mu$-Conditionals family}
-    There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as 
-    \[ \p(x_{i}=1\st x_{1},...,x_{i-1}) = \mu\left(a_{ii}+\sum_{k=1}^{i-1}a_{ik}x_i\right) \]
-    for some monotone function $\mu:\R\to(0,1)$.
-This is essentially a binomial family GLM for a link function $\mu$.
-If one takes all of the $a_{kj}$, they can form a lower triangular matrix $A$, and then the joint density can be expressed as
-
-    \[ \p(\mathbf x=\mathbf \gamma) \propto \mu(\mathbf{\gamma}^T A\mathbf \gamma)\]
-
-    If $\mu$ is chosen such that it is a bijection and differentiable, there is a unique $A$ such that $\E(\mathbf x \mathbf x^T)=M$ when generated from this model.
-The natural choice for $\mu$ is the logistic link function $L(x)=\frac{1}{1-e^{-x}}$, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
-    \[a_{ij} = \log\left(\frac{\p(x_j=x_k=1)\p(x_j=x_k=0)}{\p(x_j=0,x_k=1)\p(x_j=1,x_k=0)}\right) \]
-    when $i\neq j$. When $\bf x$ is generated from this model with $a_{ij}=0$, then $x_i$ and $x_j$ are conditionally independent.  \par
-
-    There is no closed form to calculate the entries in $A$ if $p>1$, but they can be derived numerically two ways.
-The first of these, when one is attempting to replicate the empirical cross-moments from a data matrix $X$, $a_{1i}$ to $a_{ii}$ is by fitting successive binomial family GLM regressions.
-The link function is the given function $\mu$, and $X_i$ is regressed on $X_{1} \ldots X_{i-1}$ using maximum likelihood.
-$a_{ji}$ for $i\neq j$ will then be the coefficient on $X_j$ while $a_{ii}$ is the intercept of the regression.
-\par
-    Otherwise, if one is just working with a desired cross-moment matrix $M$, the successive rows of $A$ can be fit via Newton-Raphson.
-This is performed successively on each row of $A$.
-If the first $i-1$ rows have been fit, then the upper left $(i-1)\times(i-1)$ sub matrix $A_{-i}$ of $A$ has already been filled.
-Next, $\mathbf a_i$, the first $i$ entries of the $i$th row of $A$ must be fit, while the rest of the row will be 0.
-This corresponds to $\mathbf m_i$, the first $i$ entries of the $i$th row of $M$.
-In other words, $\mathbf{m}_i = [\E(x_i x_j)]_{j=1}^{i}$.
-Finally, let us say that $\mathbf x_{-i}$ is the first $i-1$ entries of $\mathbf x$.
-$\mathbf a_i$ must satisfy
-            \begin{align*}
-                \mathbf m_i &= \E\left(x_i \left[\begin{array}{c} \mathbf x_{-i} \\ x_i \end{array}\right]\right) \\
-                \mathbf m_i &= \E\left(\E\left(x_i \left[\begin{array}{c} \mathbf x_{-i} \\ x_i \end{array}\right]\;\bigg \vert\; \mathbf x_{-i}\right)\right)\\
-                \mathbf m_i &= \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \p(\mathbf x_{-i})\p(x_i=1\st \mathbf x_{-i}) \left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \\
-            \end{align*}
-            When $\mu$ is the logistic link function $L$, this is
-            \[ \mathbf m_i = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L\left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right] \]
-            Where $c$ is the appropriate normalizing constant.
-Let us define the quantity on the right in the last line as $f(\mathbf a_i)$.
-This equation $f(\mathbf a_i)=\mathbf{m}_i$ can be solved by successive Newton-Raphson iterations defined by
-            \[\mathbf a_i^{(k+1)} = \mathbf a_i^{(k)} - \left[J\left(\mathbf a_i^{(k)}\right) \right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
-            where $J$ is the Jacobian of $f$. To do so, the Jacobian matrix is calculated as 
-        \[ J\left( \mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]
-            With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive.
-Instead, with a series of values $\mathbf x_{-i}^{(k)}\sim \mathbf x_{-i}$, the simulated values
-            \[f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right] \]
-            and
-            \[J\left( \mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
-            can be used. \par
-            Though, in theory, $A$ should always exist, in practice limited accuracy may dictate that Newton-Raphson method won't converge.
-In the simulation method in particular, if $K$ is not large enough, the equations can be very difficult or impossible to solve.
-When the intial problem can't be solved, besides increasing the simulation size or numerical accuracy, one can instead solve the relaxed problem  $f(\mathbf a)=\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
-    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T \]
-            When $\tau=0$, this yields the original problem, while when $\tau=1$, it is treats $x_i$ as independent of $\mathbf x_{-i}$. The latter will always have the solution
-            \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{M_{ii}}{1-M_{ii}}\right) \end{array} \right]^T \]
-            The hope is that for some $\tau$ close to $0$, convergence can be achieved, only causing a slight distortion from the desired cross-moments.
-
-\subsection{Generating Binary Knockoffs}
-This theory provides the tools to generate random binary knockoffs.
-This is done is done sequentially as two steps, the first of which is to come up with the target covariance matrix $\Sigma_L$ for $\mathbf x_L = [\mathbf x\; \mathbf{\tilde x}]$.
-This is extremely similar, though not identical, to the method with the original knockoffs, where $X_L^TX_L$ after normalization was targeted.
+\subsection{Relaxed Correlation Condition}
+As before with the original knockoffs, the first step of generating knockoffs is to decide on the target correlation structure for $X_{aug}$.
+For binary knockoffs, this is done by selecting the target covariance matrix $\Sigma_{aug}$.
+This is extremely similar, though not identical, to the method with the original knockoffs, where $X_{aug}^TX_{aug}$ after normalization was targeted.
 However, much the same method as before can be used.
-Let the vector of empirical means of the columns of $X$ is $\mathbf m = \frac{1}{n}X^T\mathbf 1$ and the empirical covariance matrix of $X$ is $\Sigma = \frac{1}{n}X^TX - \mathbf{m}\mathbf{m}^T$.
+Let the vector of empirical means of the columns of $X$ be $\mathbf m = \frac{1}{n}X^T\mathbf 1$ and the empirical covariance matrix of $X$ be $\Sigma = \frac{1}{n}X^TX - \mathbf{m}\mathbf{m}^T$.
 Much as before with the original knockoffs, both the equi-correlated and SDP method can be used to select  $\mathbf s$  such that $\diag{\Sigma}-\mathbf s$ is small and
-        \[ \Sigma_L = \left[\begin{array}{cc}  \Sigma & \Sigma - \diag{\mathbf s }\\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \succeq 0 \]
-    While with the original knockoffs, the result was small values for $X_i^T\tilde X_i$, now the result is low covariance between $x_i$ and $\tilde x_i$.\par
-    Additionally though, the moment condition must hold for $\Sigma_L$ to correspond to a proper random binary vector.
-If $\mathbf m_L = \E\left(\mathbf x_L\right) = [\mathbf m \; \mathbf m]$, then the desired cross-moment matrix of the joint distribution is
-        \[ M_L = \Sigma_L + \mathbf m_L\mathbf m_L^T\]
+        \[ \Sigma_{aug} = \left[\begin{array}{cc}  \Sigma & \Sigma - \diag{\mathbf s }\\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \succeq 0 \]
+    While with the original knockoffs, the result was small values for $X_i^T\tilde X_i$, now the result is low covariance between $X_i$ and $\tilde X_i$.\par
+    Additionally though, the moment condition must hold for $\Sigma_{aug}$ to correspond to a proper random binary vector. If 
+    \[\mathbf m_{aug} = \E\left(\mathbf x_{aug}\right) = \left[\begin{array}{c}\mathbf m \\ \mathbf m \end{array}\right] \]
+    then the desired cross-moment matrix of the joint distribution is 
+        \[ M_{aug} = \Sigma_{aug} + \mathbf m_{aug}\mathbf m_{aug}^T\]
         So it must be the case that  
-        \[\max\{0,\mathbf m_{L,i} + \mathbf m_{L,j} -1\} \leq \Sigma_{L,ij} + \mathbf m_i \mathbf m_j \leq \min\{\mathbf m_{L,i}, \mathbf m_{L,j}\} \]
+        \[\max\{0,\mathbf m_{aug,i} + \mathbf m_{aug,j} -1\} \leq \Sigma_{aug,ij} + \mathbf m_i \mathbf m_j \leq \min\{\mathbf m_{aug,i}, \mathbf m_{aug,j}\} \]
         However, since $\Sigma$ and $\mathbf m$ are from an actual random binary vector, the entries of $\Sigma$ satisfy this property by default.
 Only off diagonal entries which have had $\mathbf s$ subtracted from them could possibly violate this condition.
-Furthermore, for these entries, the goal is to make $\Sigma_{L,ij}$ as close to $0$ as possible.
-Since the condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$, this implies that this method should normally be satisfied.
+Furthermore, for these entries, the goal is to make $\Sigma_{aug,ij}$ as close to $0$ as possible.
+Since the condition is always satisfied in a neighborhood of $M_{aug,ij} = m_{aug,j}m_{aug,i}$, this implies that this method should normally be satisfied.
 This is the principle advantage minimizing the $\textrm{cor}(x_i,\tilde x_i)$ rather than $\E[x_i\tilde x_i]$.
@@
                     subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2\Sigma $ \\
-                               & $ 2\mathbf m_i -1 \leq \Sigma_{ii}-\mathbf s_i + \mathbf m_i^2 \leq \mathbf m_i  \quad \forall i$
+                               & $ \max(2m_i -1,0) \leq \Sigma_{ii}-\mathbf s_i + m_i^2 \leq m_i  \quad \forall i$
                 \end{tabular} 
@@
             the factor of $\diag{\sigma}^{-1}$ in the objective is to account for the $x_i$ not being normalized, as was the case with the original knockoffs. \par
-\subsubsection{Fitting via $\mu$-conditionals}
-        Once a proper cross-moment matrix $M_L$ has been selected, the most obvious approach to generate knockoffs is to apply the Newton-Raphson method to fit the matching $A$ matrix for the Ising model.
-This can be used to generate $\mathbf{ \tilde x}$ sequentially conditioned on $x$, as $\tilde x_i \st \mathbf x,\tilde x_1,\ldots,\tilde x_{i-1}$.
-However, there are a few variations of this process which are preferable depending on the situation.
-        \begin{enumerate}
-            \item If the whole $A$ matrix is being fit, then the upper half of the matrix, corresponding to $\mathbf x$, can be fit via the regression method, since they arise from real data $X$.
-This will tend to be quicker, since the calculation lends itself to parallel computing, and the data set will often be smaller than the space of all binary vectors or a simulation of that space.
-The second half of the matrix still must be fit via Newton-Raphson.
-            \item One can avoid fitting the first half of the $A$ matrix entirely though if performing Newton-Raphoson with by using the actual observed values in $X$ to simulate the first half of $\mathbf x_L$.
-This can be done by either bootstrapping rows from $X$ or using some number of copies of $X$ exactly.
-The advantage of this is the saved computation and that $\tilde x$ should be built to match the correlation condition more exactly.
-The downside is that, since these values won't be from the Ising model, the theoretical guarantee of a suitable $A$ may no longer hold.
-As well, if $n$ isn't big enough, the equations will quickly become difficult to solve, since the observed samples will provide a noisy and incomplete sample of the probability space.
-            \item As $A$ is being fit, even the parts of $\mathbf x_{-i}$ corresponding to knockoffs need not be redrawn for each Newton-Raphson iteration or each successive $i$.
-To fit $\mathbf a$ for feature $i$, the $\mathbf x_{-i}^{(k)}$ will remain constant; then, once $\mathbf a$ has been fit, the $x_i^{(k)}$ need only be drawn once.
-This saves computation, and the multiplication
-            \[ \left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [\mathbf x_{-i}^{(k)}]^T & 1 \end{array}\right] \]
-                needs only be partially recalculated for each Newton-Raphson iteration since the upper left portion, $\mathbf x_{-i}^{(k)}[\mathbf{x}_{-i}^{(k)}]^T$, won't change.
-This may result in slight deviations in any particular draw compounding into a more severe error over time.
-            \item As $p$ grows large, and $K$ necessarily along with it, the cost of calculating the Jacobian matrix grows far quicker than the cost of calculating $f(\mathbf a)$.
-This is since it has $O(Kp^2)$ complexity compared to $O(Kp)$ complexity for $f(\mathbf a)$, quickly becoming prohibitive.
-One might instead estimate the full Jacobian based on first differences, but in practice an even simpler method called Anderson mixing, which avoids additional computation of $f(\mathbf a)$, works well.
-Anderson mixing is one of several related quasi-Newtonian methods, which performs Newton-Raphson iteration $k$ with an approximation of the inverted inverted Jacobian labeled $G_k$.
-Thus, to solve for a variable $\mathbf z$ such that $f(\mathbf z)=\mathbf 0$, the best approximation for the zero is updated as $\mathbf z_{k+1}=\mathbf z_k - G_k f(\mathbf{z}_k)$ at each iteration.
-\par
-    Anderson mixing does this implicitly, with no extra recalculation of $f$, by approximating $f$ with the affine space spanning the previous $m$ evaluations of $f$.
-These values are $f(\mathbf z_{k})$ through $f(\mathbf z_{k-m+1})$.
-The best approximation for $f(\mathbf z)=0$ in the affine space which spans them is defined by
-    \[ \mathbf \alpha = \argmin_{\mathbf w:\, \sum_{i=1}^m w_i=1} \left\|\sum_{i=1}^m  w_i f(\mathbf z_{k-i+1}) \right\|_2 \]
-    This leads to the updated approximation
-    \[ \mathbf z_{k+1} = \sum_{i=1}^{m} \alpha_{i} \mathbf z_{k-i+1} \]
-    $G_k$ doesn't need to be calculated here, and the quickest implementation finds $\alpha$ without calculating it.
-However, it's presence can be seen through an equivalent formulation.
-Let $\Delta \mathbf f_i = f(\mathbf z_k) - f(\mathbf z_{k-i})$, $\Delta \mathbf z_i = \mathbf z_k - \mathbf z_{k-i}$,
-    \[ \mathcal F = [\Delta \mathbf f_1 \ldots \Delta \mathbf f_{m-1}] \]
-    and
-    \[ \mathcal Z = [\Delta \mathbf z_1 \ldots \Delta \mathbf z_{m-1}] \]
-    This means that $\mathbf\alpha$ has been replaced by $\mathbf\theta$ such that
-    \[ \mathbf \theta \; = \;\argmin_{\mathbf w} \left\|f(\mathbf z_{k}) - \mathcal  F \mathbf w\right\|_2\; = \;(\mathcal F^T \mathcal F)^{-1}\mathcal F^Tf(\mathbf z_k) \]
-    so 
-    \[ \mathbf z_{k+1} \;= \;\mathbf z_{k} - \mathcal Z \mathbf \theta\; = \; \mathbf z_{k} - \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T f(\mathbf z_k) \]
-    Which yields the approximated inverse Jacobian $G_k= \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T$
-        \end{enumerate}
-        In practice, it seemed that combining the first, third, and fourth of these variations leads to the most successful technique.
-However, should $n$ be large enough to provide a good approximation of the probability space, the second variation also saves a lot of time.
+            Given the $\sigma_{aug}$ and corresponding $M_{aug}$, all that is needed is some process to generate $\tilde X$ that replicates the desired empirical cross moments.
 
-\subsubsection{Fitting via Quadratic Programing}
-
-        Random binary knockoffs which produce the desired cross-moment matrix $M_L$ need not arise from a generative model for $\mathbf{\tilde x}$ however.
-It is sufficient to have a random matrix $\tilde X \in \{0,1\}^{n\times p}$ such that
-        \[ \E_{\tilde X}(X^T \tilde X) = \Sigma - \diag{s} \quad \& \quad \E_{\tilde X}(\tilde X^T \tilde X) = \Sigma\]
-        The simplest way to do this is based on a probability matrix $P\in (0,1)^{n\times p}$ where each $X_{ij}$ is drawn independently as $X_{ij}\sim\textrm{Bernoulli}(P_{ij})$.
-This requirement can be stated as the quadratic feasibility problem
+\subsection{Fitting via Quadratic Programing}
+        The simplest way to do this is based on a probability matrix $P\in (0,1)^{n\times p}$, where each $X_{ij}$ is drawn independently as $X_{ij}\sim\textrm{Bernoulli}(P_{ij})$.
+Such a $P$, to give as close as possible to the desired cross-moments, should satisfy
             \begin{center}
                 \begin{tabular}{r l}
-                    exists     & $P$ \\
-                    subject to & $X^TP=\Sigma-\diag{s} $ \\
-                               & $P^TP=\Sigma $ \\
+                    minimize     & $\|X^TP-(M-\diag{s})\|_{fro}^2 + \sum_{i\neq j}(P_i^T P_j - M_{ij})^2 $\\
+                    subject to   & $ \mathbf 1^T P = \mathbf m $ \\
                                  & $0 \leq P \leq 1$
@@
             \end{center}
-        If such a matrix exists and can be found, then $\tilde X$ can easily be generated.
-However, the constraints likely result in infeasibility much of the time.
-Instead, one can try to minimize how far the expected moments given $P$ would be from those desired for $M_L$.
-This would take the form
+            The first part of the objective corresponds to the deviation of $\E[X^T\tilde X]$ from $M-\diag{s}$ and the second part to the deviation of $\E[\tilde X^T \tilde X]$ from $M$, excluding the diagonal values which are matched by $ \mathbf 1^T P = \mathbf m $.
+This problem can be stated as a quadratic program via the introduction of slack variables:
             \begin{center}
                 \begin{tabular}{r l}
-                    minimize     & $\|X^TP-(\Sigma-\diag{s})\|_{fro}^2 + \|P^TP-\Sigma\|_{fro}^2 $\\
-                    subject to   & $ \mathbf 1^T P = \mathbf m_L $ \\
+                    minimize     & $\|W\|_{fro}^2 + \|V\|_{fro}^2 $ \\
+                    subject to   & $ -W \leq X^TP-(M-\diag{s})\leq W $ \\
+                                 & $ -V_{ij} \leq P_i^T P_j - M_{ij} \leq V_{ij} \quad \forall i\neq j$ \\
+                                 & $ \mathbf 1^T P = \mathbf m $ \\
                                  & $0 \leq P \leq 1$
@@
             \end{center}
-            Here, the marginal distribution of $\tilde X_i$ will be the same as $X_i$, and the cross-moments shouldn't deviate to far from $M_L$.
-This problem can be stated as the quadratic program via the introduction of slack variables:
-            \begin{center}
-                \begin{tabular}{r l}
-                    minimize     & $\|W\|_{fro}^2 + \|V\|_{fro}^2 $ \\
-                    subject to   & $ -W \leq X^TP-(\Sigma-\diag{s})\leq W $ \\
-                                 & $ -V \leq P^TP-\Sigma \leq V $ \\
-                                 & $ \mathbf 1^T P = \mathbf m_L $ \\
-                                 & $0 \leq P \leq 1$
-                \end{tabular} 
-            \end{center}
-            Though there are nice tools and theory for solving quadratic programs, $n\times p$ would have to be fairly small for this optimization problem to be computationally reasonable in practice.
-Ideally, there would be a method to breakdown this problem into a series of smaller optimization problems which would yield the same solution.
+
+            Though there are powerful tools and theory for solving quadratic programs, $n\times p$ would have to be fairly small for this optimization problem to be computationally reasonable in practice.
+Ideally, there would be a method to split up this problem into a series of smaller optimization problems which would yield the same solution.
 An optimist might hope that iterating through the columns of $P$, optimizing each to hit its desired cross-moments individually conditioned on the estimate of the rest of $P$, would achieve this.
@@
 
+\subsection{Fitting via Conditional Random Binary Vectors }
+An alternate approach to generating $\tilde X$ that will satisfy the relaxed correlation condition is to find a random binary vector $\mathbf{x}_{aug}\in\{0,1\}^{2p}$ that has the desired cross-moments $M_{aug}$.
+Let $\mathbf{x}_{aug,1}$ be the first $p$ entries of $\mathbf{x}_{aug}$ and $\mathbf{x}_{aug,2}$ be the remainder.
+Then, for a row $\xi$ of $X$, the corresponding row $\tilde \xi$ of $\tilde X$ would be drawn from $\mathbf{x}_{aug,2}\st \mathbf{x}_{aug,1}=\xi$.
+This necessitates a family of binary vectors which can match any cross-moment matrix and is easy to conditionally sample from.
+
+\subsubsection{Gaussian Copula Family}
+Since multivariate normal distributions are quite easy to sample from and are defined by their first two moments, an obvious choice is to use them to generate random binary vectors.
+This might be done by finding $\mu$ and $\Sigma$ such that for $\mathbf z\sim N_p(\mu,\Sigma)$, where $x_i$ is defined as $x_i = \mathrm{sign}(z_i)$, $\mathbf x$ has the desired first two moments.
+However, this is only guaranteed to be feasible in the bivariate case.
+For higher dimensions, this method can generally only provide random binary vectors with approximately the right moments, and the quality of the approximation quickly degenerates as $p$ increases\cite{2011arXiv1111.0576S}.
+Therefore, we do not explore this direction further.
+
+\subsubsection{$\mu$-Conditionals family}
+There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals\cite{2011arXiv1111.0576S}.
+The basic idea is that the entries of a binary vector variable $\mathbf x$ can be sampled sequentially as
+    \[ \p(x_{i}=1\st x_{1},...,x_{i-1}) = \mu\left(a_{ii}+\sum_{k=1}^{i-1}a_{ik}x_i\right) \]
+    for some monotone function $\mu:\R\to(0,1)$.
+This is essentially a binomial family GLM for a link function $\mu$.
+If one takes all of the $a_{kj}$, they can form a lower triangular matrix $A$, and then the joint density can be expressed as
+
+    \[ \p(\mathbf x=\mathbf \xi) \propto \mu(\mathbf{\xi}^T A\mathbf \xi)\]
+
+    If $\mu$ is chosen such that it is a bijection and differentiable, there is a unique $A$ such that $\E(\mathbf x \mathbf x^T)=M$ when generated from this model.
+The natural choice for $\mu$ is the logistic link function $L(x)=\frac{1}{1-e^{-x}}$, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
+    \[a_{ij} = \log\left(\frac{\p(x_j=x_k=1)\p(x_j=x_k=0)}{\p(x_j=0,x_k=1)\p(x_j=1,x_k=0)}\right) \]
+    when $i\neq j$. When $\bf x$ is generated from this model with $a_{ij}=0$, then $x_i$ and $x_j$ are conditionally independent.  \par
+
+    There is no closed form to calculate the entries in $A$ if $p>1$, but they can be derived numerically two ways.
+The first of these, for when one is attempting to replicate the empirical cross-moments from a data matrix $X$, is to fit $a_{1i}$ to $a_{ii}$ by successive binomial family GLM regressions.
+The link function is the given function $\mu$, and $X_i$ is regressed on $X_{1} \ldots X_{i-1}$ using maximum likelihood.
+$a_{ji}$ for $i\neq j$ will then be the coefficient on $X_j$ while $a_{ii}$ is the intercept of the regression.
+\par
+    Otherwise, if one is just working with a desired cross-moment matrix $M$, the successive rows of $A$ can be fit via Newton-Raphson\cite{2011arXiv1111.0576S}.
+This is performed successively on each row of $A$.
+If the first $i-1$ rows have been fit, then the upper left $(i-1)\times(i-1)$ sub matrix $A_{-i}$ of $A$ has already been filled.
+Next, $\mathbf a_i$, the first $i$ entries of the $i$th row of $A$ must be fit, while the rest of the row will be 0.
+This corresponds to $\mathbf m_i$, the first $i$ entries of the $i$th row of $M$.
+In other words, $\mathbf{m}_i = [\E(x_i x_j)]_{j=1}^{i}$.
+Finally, let us say that $\mathbf x_{-i}$ is the first $i-1$ entries of $\mathbf x$.
+$\mathbf a_i$ must satisfy
+            \begin{align*}
+                \mathbf m_i &= \E\left(x_i \left[\begin{array}{c} \mathbf x_{-i} \\ x_i \end{array}\right]\right) \\
+                 &= \E\left(\E\left(x_i \left[\begin{array}{c} \mathbf x_{-i} \\ x_i \end{array}\right]\;\bigg \vert\; \mathbf x_{-i}\right)\right)\\
+                 &= \sum_{\xi_{-i} \in \{0,1\}^{i-1}} \p(\xi_{-i})\p(x_i=1\st \mathbf x_{-i}=\xi_{-i}) \left[\begin{array}{c} \xi_{-i} \\ 1 \end{array}\right] \\
+            \end{align*}
+            When $\mu$ is the logistic link function $L$, this is
+            \[ \mathbf m_i = \sum_{\xi_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L\left(\mathbf \xi_{-i}^TA_{-i}\xi_{-i}\right) L\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf \xi_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \xi_{-i} \\ 1 \end{array}\right] \]
+            where $c$ is the appropriate normalizing constant.
+Let us define the quantity on the right in the last line as $f(\mathbf a_i)$.
+This equation $f(\mathbf a_i)=\mathbf{m}_i$ can be solved by successive Newton-Raphson iterations defined by
+            \[\mathbf a_i^{(k+1)} = \mathbf a_i^{(k)} - \left[J\left(\mathbf a_i^{(k)}\right) \right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
+            where $J$ is the Jacobian of $f$. To do so, the Jacobian matrix is calculated as 
+        \[ J\left( \mathbf a_i\right) = \sum_{\xi_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L\left(\mathbf \xi_{-i}^TA_{-i}\xi_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf \xi_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \xi_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \xi_{-i}^T & 1 \end{array}\right] \]
+            With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive. Instead, with K iid draws $\xi_{-i}^{(k)}\sim \mathbf x_{-i}$, the simulated values 
+            \[f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L\left(\mathbf{a}_i^T\left[\begin{array}{c} \xi_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \xi_{-i}^{(k)} \\ 1 \end{array}\right] \]
+            and
+            \[J\left( \mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \xi_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \xi_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} [\xi_{-i}^{(k)}]^T & 1 \end{array}\right] \]
+            can be used. \par
+            Though, in theory, $A$ should always exist, in practice limited accuracy may dictate that Newton-Raphson method won't converge.
+In the simulation method in particular, if $K$ is not large enough, the equations can be very difficult or impossible to solve.
+When the initial problem can't be solved, besides increasing the simulation size or numerical accuracy, one can instead solve the relaxed problem  $f(\mathbf a)=\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
+    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T \]
+            When $\tau=0$, this yields the original problem, while when $\tau=1$, it is treats $x_i$ as independent of $\mathbf x_{-i}$. The latter will always have the solution
+            \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{M_{ii}}{1-M_{ii}}\right) \end{array} \right]^T \]
+            The hope is that for some $\tau$ close to $0$, convergence can be achieved, only causing a slight distortion from the desired cross-moments.
+
+\subsubsection{Generating Binary Knockoffs from $\mu$-conditionals}
+This suggests that an obvious choice for $\mathbf{x}_{aug}$ is to fit the $A$ for the Ising model which corresponds to $M_{aug}$ via the Newton-Raphson method.
+$A$ can then be used to generate a row $\tilde \xi$ of $\tilde X$ sequentially conditioned on the corresponding row $\xi$ of $X$ as $\tilde \xi_i \st \mathbf \xi,\tilde \xi_1,\ldots,\tilde \xi_{i-1}$.
+However, there are a few variations of this process which are preferable depending on the situation.
+        \begin{enumerate}
+            \item If the whole $A$ matrix is being fit, then the upper half of the matrix, corresponding to $\mathbf x$, can be fit via the regression method, since they arise from real data $X$.
+This will tend to be quicker, since the calculation lends itself to parallel computing, and the data set will often be smaller than the space of all binary vectors or a simulation of that space.
+The second half of the matrix still must be fit via Newton-Raphson.
+            \item One can avoid fitting the first half of the $A$ matrix entirely though by using the actual observed values in $X$ to simulate the first half of $\mathbf x_{aug}$, if performing Newton-Raphoson.
+This can be done by either bootstrapping rows from $X$ or using some number of copies of $X$ exactly.
+The advantage of this is the saved computation and that $\mathbf{\tilde x}$ should be built to match the correlation condition more exactly.
+The downside is that, since these values won't be from the Ising model, the theoretical guarantee of a suitable $A$ may no longer hold.
+As well, if $n$ isn't big enough, the equations will quickly become difficult to solve, since the observed samples will provide a noisy and incomplete sample of the probability space.
+            \item As $A$ is being fit, even the parts of $\mathbf x_{-i}$ corresponding to knockoffs need not be redrawn for each Newton-Raphson iteration or each successive $i$.
+To fit $\mathbf a_i$, the $\xi_{-i}^{(k)}$ will remain constant; then, once $\mathbf a_i$ has been fit, the $\xi_i^{(k)}$ need only be drawn once.
+This saves computation, and the multiplication
+            \[ \left[\begin{array}{c} \xi_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} [\xi_{-i}^{(k)}]^T & 1 \end{array}\right] \]
+                needs only be partially recalculated for each Newton-Raphson iteration since the upper left portion, $\xi_{-i}^{(k)}[\xi_{-i}^{(k)}]^T$, won't change.
+This may result in slight deviations in any particular draw compounding into a more severe error over time.
+            \item As $p$ grows large, and $K$ necessarily along with it, the cost of calculating the Jacobian matrix grows far quicker than the cost of calculating $f(\mathbf a)$.
+This is since it has $O(Kp^2)$ complexity compared to $O(Kp)$ complexity for $f(\mathbf a)$, which quickly becomes prohibitive.
+One might instead estimate the full Jacobian based on first differences, but in practice an even simpler method called Anderson mixing\cite{Fang07twoclasses}, which avoids additional computation of $f(\mathbf a)$, works well.
+Anderson mixing is one of several related quasi-Newtonian methods, which performs Newton-Raphson iteration $k$ with an approximation of the inverted inverted Jacobian labeled $G_k$.
+Thus, to solve for a variable $\mathbf z$ such that $f(\mathbf z)=\mathbf 0$, the best approximation for the root is updated as $\mathbf z_{k+1}=\mathbf z_k - G_k f(\mathbf{z}_k)$ at each iteration.
+\par
+    Anderson mixing does this implicitly, with no extra recalculation of $f$, by approximating $f$ with the affine space spanning the previous $m$ evaluations of $f$.
+These values are $f(\mathbf z_{k})$ through $f(\mathbf z_{k-m+1})$.
+The best approximation for $f(\mathbf z)=0$ in the affine space which spans them is defined by
+    \[ \mathbf \alpha = \argmin_{\mathbf w:\, \sum_{i=1}^m w_i=1} \left\|\sum_{i=1}^m  w_i f(\mathbf z_{k-i+1}) \right\|_2 \]
+    This leads to the updated approximation
+    \[ \mathbf z_{k+1} = \sum_{i=1}^{m} \alpha_{i} \mathbf z_{k-i+1} \]
+    $G_k$ doesn't need to be calculated here, and the quickest implementation finds $\alpha$ without calculating it.
+However, it's presence can be seen through an equivalent formulation.
+Let $\Delta \mathbf f_i = f(\mathbf z_k) - f(\mathbf z_{k-i})$, $\Delta \mathbf z_i = \mathbf z_k - \mathbf z_{k-i}$,
+    \[ \mathcal F = [\Delta \mathbf f_1 \ldots \Delta \mathbf f_{m-1}] \]
+    and
+    \[ \mathcal Z = [\Delta \mathbf z_1 \ldots \Delta \mathbf z_{m-1}] \]
+    This means that $\mathbf\alpha$ has been replaced by $\mathbf\theta$ such that
+    \[ \mathbf \theta \; = \;\argmin_{\mathbf w} \left\|f(\mathbf z_{k}) - \mathcal  F \mathbf w\right\|_2\; = \;(\mathcal F^T \mathcal F)^{-1}\mathcal F^Tf(\mathbf z_k) \]
+    so 
+    \[ \mathbf z_{k+1} \;= \;\mathbf z_{k} - \mathcal Z \mathbf \theta\; = \; \mathbf z_{k} - \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T f(\mathbf z_k) \]
+    Which yields the approximated inverse Jacobian $G_k= \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T$
+        \end{enumerate}
+        In practice, it seemed that combining the first, third, and fourth of these variations leads to the most successful technique.
+However, should $n$ be large enough to provide a good approximation of the probability space, the second variation also saves a lot of time.
+
 \section{Binary Knockoff Performance}
@@
     \end{center}
-    \caption{Measures how much far the empirical cross-moment matrix $\frac{1}{n}X_L^TX_L$ is from the desired cross moment matrix $M_L$ for random binary knockoffs for various $n$ and $p$, averaged over $50$ simulations of $X$ from the Ising model.
-Where $W = \Sigma - \diag{\mathbf s} + \mathbf m \mathbf m^T$ is the top quadrant of $M_L$, deviation is calculated as \\ $(\|X^T\tilde X - W\|_{fro} + \|\tilde X^T \tilde X - M\|_{fro})/(\|W\|_{fro}+\|M\|_{fro})$}
+    \caption{Measures how much far the empirical cross-moment matrix $\frac{1}{n}X_{aug}^TX_{aug}$ is from the desired cross moment matrix $M_{aug}$ for random binary knockoffs for various $n$ and $p$, averaged over $50$ simulations of $X$ from the Ising model.
+Where $W = \Sigma - \diag{\mathbf s} + \mathbf m \mathbf m^T$ is the top quadrant of $M_{aug}$, deviation is calculated as \\ $(\|X^T\tilde X - W\|_{fro} + \|\tilde X^T \tilde X - M\|_{fro})/(\|W\|_{fro}+\|M\|_{fro})$}
     \end{figure}
@@
 The binary knockoffs only replicate the correlation structure of the original features in expectation, rather than exactly.
-Thus, for any given value $n$, the empirical correlation of the combined design matrix $X_L$ won't exactly be $\Sigma_L$, and the method won't work perfectly.
-However, even here, the actual FDR is only somewhat hire than the desired $.2$, and for larger values of $n$, the empirical correlation will get closer to $\Sigma_L$, and the FDR will be controlled better.
+Thus, for any given value $n$, the empirical correlation of the combined design matrix $X_{aug}$ won't exactly be $\Sigma_{aug}$, and the method won't work perfectly.
+However, even here, the actual FDR is only somewhat hire than the desired $.2$, and for larger values of $n$, the empirical correlation will get closer to $\Sigma_{aug}$, and the FDR will be controlled better.
 
@@
 Currently, they provide approximate FDR control for LASSO and seem like they also provide approximate FDR control for logistic regression.
-However, there are no theoretical guarantees for logistic regression, and the quality of FDR control for LASSO depends on how far the realized cross-moment matrix deviates from the desired $M_L$.
+However, there are no theoretical guarantees for logistic regression, and the quality of FDR control for LASSO depends on how far the realized cross-moment matrix deviates from the desired $M_{aug}$.
 In cases where $n$ is large, this should only be a small issue, but for smaller $n$ FDR control fails in a meaningful way.

### 2015-07-31 masters_paper/paper.tex
Lots more edits

@@
 \documentclass[11pt]{article}
-\usepackage[paper=letterpaper, margin=.5in]{geometry}
+\usepackage[paper=letterpaper, margin=.75in]{geometry}
 \linespread{1.5}
@@
 % placeins defines "\FloatBarrier", which stops tables from moving around
-\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins, caption} 
+\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins, caption,hyperref} 
 \usepackage[utf8]{inputenc}
@@
 
-\newcommand{\x}{\bf x}
+\newcommand{\x}{\mathbf x}
 
@@
 The knockoff filter method provides one method of variable selection for linear regression.
-It relies on generating 'knockoff' features, which replicate the correlation structure of the original variable; when the full path of LASSO regression is fit, the point at which a a null variable and its knockoff have nonzero coefficients will be exchangeable.
+It relies on generating `knockoff' features, which replicate the correlation structure of the original variable; when the full path of LASSO regression is fit, the point at which a null variable and its knockoff have nonzero coefficients will be exchangeable.
 However, for other GLMs, the method breaks down.
@@
 Variable selection is an essential problem to fitting a regression model.
-For linear regression, the knockoff filter \cite{2014arXiv1404.5609F} offers a method of exact FDR control.
+For linear regression, the knockoff filter\cite{2014arXiv1404.5609F} offers a method of exact FDR control.
 However, the method as originally formulated does not extend to other generalized linear models (GLMs) such as logistic regression.
@@
     \[\mathbf{y} = X\beta + \mathbf{z}\]
-    where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise.It is important to note that a intercept vector $\mathbf 1$ is not included in $X$.
-For the purpose of this paper, only the case where $2p\leq n$ will be considered, though the methods can be extended to $p\leq n$.
+    where $\mathbf{y}\in\R^n$ are the observed values of $y$, $X \in \R^{n\times p}$ is the matrix of predictor variables, $\beta\in\R^p$ are the unknown coefficients, and $\mathbf z$ is Gaussian noise.
+It is important to note that a intercept vector $\mathbf 1$ is not included in $X$.
+For the purpose of this paper, only the case where $2p\leq n$ will be considered, though the methods can be extended to $p\geq n$.
 I will to refer the $j$th column of $X$ as $X_j$, which is $n$ observations of the random variable $x_j$.
@@
     In other words, FDR is the portion of features which are thought to have an effect on $y$ but actually have no effect.
-FDR is controlled at level $q$ if FDR is less than $q$ irrespective of the coefficients $\beta$.
+FDR is controlled at level $q$ if FDR is less than or equal to $q$ irrespective of the coefficients $\beta$.
 \par
@@
     \[ \tilde X^T\tilde X=G \textrm{ and } \tilde X^T X = G - \diag{\mathbf s}\]
-    for some vector $\mathbf{s}\in \R^p$ such that, writing $\diag{G}$ as the vector made from the main diagonal of $G$, $\diag{G} - \mathbf s$ is small.
+    for some vector $\mathbf{s}\in \R^p$ such that, writing $\diag{G}$ as the vector made from the main diagonal of $G$, $\diag{G} - \mathbf s$ has small magnitude.
 Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$.
@@
 One feature entering the LASSO path before another is an indication that the first feature is a stronger predictor of the outcome than the second.
-Thus, since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to enter the LASSO path sooner than the knockoff features when the original is a valid predictor.
+Since, by construction, the knockoff features are weaker predictors than the originals, we would expect the original features to enter the LASSO path sooner than the knockoff features when the original is a valid predictor.
 On the other hand, since $\beta(\lambda)$ only depends on $X_{aug}$ through the sufficient statistics $G_{aug}:=X_{aug}^TX_{aug}$ and $X_{aug}^T\mathbf y$, for null predictors, the coefficients of the original features won't enter any sooner on average than knockoff feature.
@@
 \par
-   This observation leads to the final step. Let $Z_j$ be the point at which $X_j$ entered the path and $\tilde Z_j$ the point at which $\tilde X_j$ entered the path. Then, define $W_j$ as 
+   This observation leads to the final step. Let $Z_j$ be the point at which $X_j$ enters the path and $\tilde Z_j$ the point at which $\tilde X_j$ enters the path. Then, define $W_j$ as 
    \[ W_j = \begin{cases}   Z_j         &\mbox{if } Z_j>\tilde Z_j \\ 
@@
 Thus, we will make a selection $\hat S=\{j:W_j\geq T\}$ for some threshold $T$.
-When $X_j\not\in S$, $X_j$ and $\tilde X_j$ are both equally likely to enter the path first, so $W_j$ is equally as likely to be positive as negative for $j\in S$.
+When $j\not\in S$, $X_j$ and $\tilde X_j$ are both equally likely to enter the path first, so $W_j$ is equally as likely to be positive as negative for $j\in S$.
 Thus,
@@
     For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. The Gram matrix of $X_{aug}$ is
-    \[ G_{aug}= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right] \]
+    \[ G_{aug}:= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right] \]
     $A$, the Schur complement of $G$ in $G_{aug}$ can be calculated as 
@@
 This effect is particularly noticeable when the random variable $x_j$ is discrete but $\tilde X_j$ does not even consist of the same set of values.
-The end result is that for variable selection for other generalized linear models, which do not have the same sufficient statistics as linear regression which knockoffs are designed to hit, the original knockoffs will fail to control FDR.
+The end result is that for variable selection for other generalized linear models, which do not have the same sufficient statistics which knockoffs are designed to replicate, the original knockoffs will fail to control FDR.
 \par
@@
     Where $\Sigma$ is the covariance matrix of $X$.
-This is essentially relaxation of the original correlation condition, requiring the desired covariance to only hold in expectation.
+This is essentially a relaxation of the original correlation condition, requiring the desired covariance to only hold in expectation.
 For large values of $n$, the observed covariance will converge to its expectation, making this distinction minor.
@@
     \[ \beta(\lambda) = \argmin_\mathbf b \left\{l(\mathbf{b}\st X, \tilde X) + \lambda\|\mathbf{b}\|_1 \right\}\]
+
     The $Z_j$ and $\tilde Z_j$ are then, once again, the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before.
@@
     \end{center}
-        \caption{The mean cumulative portion of $W$ statistic which are positive by absolute rank of the value.
+    \caption{\linespread{1}\selectfont{} The mean cumulative portion of $W$ statistic which are positive by absolute rank of the value.
 The mean is over 1000 simulations.
@@
 
+    \FloatBarrier
+
     To evaluate whether this made a difference in practice, I tested the method using original style knockoffs in logistic regression.
@@
 This is particularly important for high values, which play the largest roll in selecting a threshold.
-In the top left corner, under the ``global null", where all features are null and normally distributed, that we almost achieve this.
+In the top left corner, under the ``global null", where all features are null and normally distributed, we almost achieve this.
 There is just a small, but consistent, deviation from $50\%$.
@@
 \par
-    \FloatBarrier
+    \begin{figure}[h]
+        \begin{center}
+            \includegraphics[width=14cm]{images/normalQQ}
+            \caption{\linespread{1}\selectfont{}Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
+        \end{center}
+    \end{figure}
+
     This gets much worse when some true predictors are included, as shown in the bottom row of the figure.
@@
 \par
-    \begin{figure}[h]
-        \begin{center}
-            \includegraphics[width=14cm]{images/normalQQ}
-            \caption{Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
-        \end{center}
-    \end{figure}
     Upon examination, the issue seems to be that the original style knockoff features have a very different empirical distribution than the features they are imitating.
@@
 Thus, instead, this section will develop methods of generating $\tilde X$ randomly such that the correlation condition holds in expectation.
-Mainly, $\tilde X\st X$ should be a random matrix such that
+Mainly, $\tilde X\st X$ should be a random binary matrix such that
         \[ \E(X^T \tilde X \st X) = X^TX - \diag{s} \quad \& \quad \E(\tilde X^T \tilde X \st X) = X^TX\]
@@
 The full class of random binary vector variables on $\{0,1\}^p$ can be specified as multinomial on the $2^p$ elements of the set.
-However, in practice, it is generally impractical to specify or estimate all $2^p-1$ necessary parameters, so it is useful to pick a method to match the first two moments, as is desirable for knockoffs.
-These are the mean vector $\E(\bf x)=m\in[0,1]^p$ and the cross-moment matrix $\E(\bf x\bf x^T)=M\in[0,1]^{p\times p}$.
+However, it is generally impractical to specify or estimate all $2^p-1$ necessary parameters, so it is useful to pick a method to match the first two moments, as is desirable for knockoffs.
+These are the mean vector $\E(\mathbf x)=\mathbf{m}\in[0,1]^p$ and the cross-moment matrix $\E(\mathbf x\mathbf x^T)=M\in[0,1]^{p\times p}$.
 Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag{M}$.
@@
     \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
-    for all $i\neq j$\cite{2011arXiv1111.0576S}.
+    for all $i\neq j$.\cite{2011arXiv1111.0576S}
 
@@
 Let the vector of empirical means of the columns of $X$ be $\mathbf m = \frac{1}{n}X^T\mathbf 1$ and the empirical covariance matrix of $X$ be $\Sigma = \frac{1}{n}X^TX - \mathbf{m}\mathbf{m}^T$.
-Much as before with the original knockoffs, both the equi-correlated and SDP method can be used to select  $\mathbf s$  such that $\diag{\Sigma}-\mathbf s$ is small and
+Much as before with the original knockoffs, both the equi-correlated and SDP method can be used to select  $\mathbf s$  such that $\diag{\Sigma}-\mathbf s$ has small magnitude and
         \[ \Sigma_{aug} = \left[\begin{array}{cc}  \Sigma & \Sigma - \diag{\mathbf s }\\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \succeq 0 \]
@@
         \[ M_{aug} = \Sigma_{aug} + \mathbf m_{aug}\mathbf m_{aug}^T\]
-        So it must be the case that  
-        \[\max\{0,\mathbf m_{aug,i} + \mathbf m_{aug,j} -1\} \leq \Sigma_{aug,ij} + \mathbf m_i \mathbf m_j \leq \min\{\mathbf m_{aug,i}, \mathbf m_{aug,j}\} \]
+        As mentioned above, for this cross-moment matrix to correspond to an actual random binary vector, it must be the case  
+        \[\max\{0,(m_{aug})_i + ( m_{aug})_j -1\} \leq (\Sigma_{aug})_{ij} +  (m_{aug})_i  (m_{aug})_j \leq \min\{( m_{aug})_i, ( m_{aug})_j\} \]
     However, since $\Sigma$ and $\mathbf m$ are from an actual random binary vector, the entries of $\Sigma$ satisfy this property by default.
 Only off diagonal entries which have had $\mathbf s$ subtracted from them could possibly violate this condition.
-Furthermore, for these entries, the goal is to make $\Sigma_{aug,ij}$ as close to $0$ as possible.
-Since the condition is always satisfied in a neighborhood of $M_{aug,ij} = m_{aug,j}m_{aug,i}$, this implies that this method should normally be satisfied.
-This is the principle advantage minimizing the $\textrm{cor}(x_i,\tilde x_i)$ rather than $\E[x_i\tilde x_i]$.
+Furthermore, for these entries, the goal is to make $(\Sigma_{aug})_{ij}$ as close to $0$ as possible.
+Since the condition is always satisfied in a neighborhood of $(M_{aug})_{ij} = (m_{aug})_j(m_{aug})_i$, this implies that this method should normally be satisfied.
+This is the principle advantage of minimizing the $\textrm{cov}(x_i,\tilde x_i)$ rather than $\E[x_i\tilde x_i]$.
 For the SDP method though, the restriction can be built directly into the optimization problem.
@@
                     subject to & $ 0 \preceq \diag{\mathbf s} \preceq 2\Sigma $ \\
-                               & $ \max(2m_i -1,0) \leq \Sigma_{ii}-\mathbf s_i + m_i^2 \leq m_i  \quad \forall i$
+                               & $ \max\{2m_i -1,0\} \leq \Sigma_{ii}-s_i + m_i^2 \leq m_i  \quad \forall i$
                 \end{tabular} 
@@
             the factor of $\diag{\sigma}^{-1}$ in the objective is to account for the $x_i$ not being normalized, as was the case with the original knockoffs. \par
-            Given the $\sigma_{aug}$ and corresponding $M_{aug}$, all that is needed is some process to generate $\tilde X$ that replicates the desired empirical cross moments.
+            Given the $\Sigma_{aug}$ and corresponding $M_{aug}$, all that is needed is some process to generate $\tilde X$ that replicates the desired empirical cross moments.
 
@@
         The simplest way to do this is based on a probability matrix $P\in (0,1)^{n\times p}$, where each $X_{ij}$ is drawn independently as $X_{ij}\sim\textrm{Bernoulli}(P_{ij})$.
-Such a $P$, to give as close as possible to the desired cross-moments, should satisfy
+Such a $P$ should satisfy
             \begin{center}
@@
             \end{center}
-            The first part of the objective corresponds to the deviation of $\E[X^T\tilde X]$ from $M-\diag{s}$ and the second part to the deviation of $\E[\tilde X^T \tilde X]$ from $M$, excluding the diagonal values which are matched by $ \mathbf 1^T P = \mathbf m $.
+            so that the cross-moments are as close as possible to $M_{aug}$.
+The first part of the objective corresponds to the deviation of $\E[X^T\tilde X]$ from $M-\diag{s}$ and the second part to the deviation of $\E[\tilde X^T \tilde X]$ from $M$ excluding the diagonal values which are matched by $ \mathbf 1^T P = \mathbf m $.
 This problem can be stated as a quadratic program via the introduction of slack variables:
@@
 In practice, this does seem to yield reasonable results, and is far quicker than other methods for large $p$.
-In particular, the following method seemed to work well:
+In particular, the following method seems to work well:
     \begin{enumerate}
-        \item Optimize initial $P_i$ as such that $X_j^T P_i = M_{ij}$ if $i\neq j$ and $X_i^T P_i = \sigma_i(1-s)$.   
+        \item Optimize initial $P_i$ such that $X_j^T P_i = M_{ij}$ if $i\neq j$ and $X_i^T P_i = M_{ii}-s_i$.   
             \begin{center}
                 \begin{tabular}{r l}
-                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_{ij})^2$ \\
+                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - ( M_{ii}-s_i)\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_{ij})^2$ \\
                     subject to   & $ \frac{1}{n}\mathbf 1^T P_i = M_{ii} $ \\
@@
                 \begin{tabular}{r l}
-                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_{ij})^2 + \sum_{i\neq j}(\frac{1}{n}\tilde X_j^T P_i - M_{ij})^2$  \\
+                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-s_i)\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_{ij})^2 + \sum_{i\neq j}(\frac{1}{n}\tilde X_j^T P_i - M_{ij})^2$  \\
                     subject to   & $ \frac{1}{n}\mathbf 1^T P_i = M_{ii} $ \\
@@
 An alternate approach to generating $\tilde X$ that will satisfy the relaxed correlation condition is to find a random binary vector $\mathbf{x}_{aug}\in\{0,1\}^{2p}$ that has the desired cross-moments $M_{aug}$.
-Let $\mathbf{x}_{aug,1}$ be the first $p$ entries of $\mathbf{x}_{aug}$ and $\mathbf{x}_{aug,2}$ be the remainder.
-Then, for a row $\xi$ of $X$, the corresponding row $\tilde \xi$ of $\tilde X$ would be drawn from $\mathbf{x}_{aug,2}\st \mathbf{x}_{aug,1}=\xi$.
-This necessitates a family of binary vectors which can match any cross-moment matrix and is easy to conditionally sample from.
+Let $(\mathbf{x}_{aug})_1$ be the first $p$ entries of $\mathbf{x}_{aug}$ and $(\mathbf{x}_{aug})_2$ be the remainder.
+Then, for a row $\xi$ of $X$, the corresponding row $\tilde \xi$ of $\tilde X$ would be drawn from $(\mathbf{x}_{aug})_2\st (\mathbf{x}_{aug})_1=\xi$.
+This method necessitates a family of binary vectors which can match any cross-moment matrix and is easy to conditionally sample from.
 
@@
 Since multivariate normal distributions are quite easy to sample from and are defined by their first two moments, an obvious choice is to use them to generate random binary vectors.
-This might be done by finding $\mu$ and $\Sigma$ such that for $\mathbf z\sim N_p(\mu,\Sigma)$, where $x_i$ is defined as $x_i = \mathrm{sign}(z_i)$, $\mathbf x$ has the desired first two moments.
+This might be done by defining $x_i = \mathrm{sign}(z_i)$ for $\mathbf z\sim N_p(\mu,\Sigma)$, where $\mu$ and $\Sigma$ are chosen such that $\mathbf x$ has the correct cross-moments.
 However, this is only guaranteed to be feasible in the bivariate case.
-For higher dimensions, this method can generally only provide random binary vectors with approximately the right moments, and the quality of the approximation quickly degenerates as $p$ increases\cite{2011arXiv1111.0576S}.
-Therefore, we do not explore this direction further.
+For higher dimensions, this method can generally only provide random binary vectors with approximately the right moments, and the quality of the approximation quickly degenerates as $p$ increases.\cite{2011arXiv1111.0576S} Therefore, we do not explore this direction further.
 
 \subsubsection{$\mu$-Conditionals family}
-There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals\cite{2011arXiv1111.0576S}.
-The basic idea is that the entries of a binary vector variable $\mathbf x$ can be sampled sequentially as
+There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals.\cite{2011arXiv1111.0576S} The basic idea is that the entries of a binary vector variable $\mathbf x$ can be sampled sequentially as
     \[ \p(x_{i}=1\st x_{1},...,x_{i-1}) = \mu\left(a_{ii}+\sum_{k=1}^{i-1}a_{ik}x_i\right) \]
@@
 
-    If $\mu$ is chosen such that it is a bijection and differentiable, there is a unique $A$ such that $\E(\mathbf x \mathbf x^T)=M$ when generated from this model.
-The natural choice for $\mu$ is the logistic link function $L(x)=\frac{1}{1-e^{-x}}$, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
+    If $\mu$ is chosen such that it is a bijection and differentiable, for any proper cross-moment matrix $M$, there is a corresponding $A$ such that $\E(\mathbf x \mathbf x^T \st A)=M$.
+The natural choice for $\mu$ is the logistic link function $L(x)=\frac{1}{1-e^{-x}}$, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients are log odds ratios:
     \[a_{ij} = \log\left(\frac{\p(x_j=x_k=1)\p(x_j=x_k=0)}{\p(x_j=0,x_k=1)\p(x_j=1,x_k=0)}\right) \]
-    when $i\neq j$. When $\bf x$ is generated from this model with $a_{ij}=0$, then $x_i$ and $x_j$ are conditionally independent.  \par
+    when $i\neq j$. When $\mathbf x$ is generated from this model with $a_{ij}=0$, then $x_i$ and $x_j$ are conditionally independent.  \par
 
     There is no closed form to calculate the entries in $A$ if $p>1$, but they can be derived numerically two ways.
-The first of these, for when one is attempting to replicate the empirical cross-moments from a data matrix $X$, is to fit $a_{1i}$ to $a_{ii}$ by successive binomial family GLM regressions.
+The first of these, for when one is attempting to replicate the empirical cross-moments from a data matrix $X$, is to fit each row $a_{1i}$ to $a_{ii}$ by successive binomial family GLM regressions.
 The link function is the given function $\mu$, and $X_i$ is regressed on $X_{1} \ldots X_{i-1}$ using maximum likelihood.
-$a_{ji}$ for $i\neq j$ will then be the coefficient on $X_j$ while $a_{ii}$ is the intercept of the regression.
+$a_{ji}$ for $i\neq j$ will then be the coefficient on $X_j$ in the model predicting $X_i$ while $a_{ii}$ is the intercept of the regression.
 \par
-    Otherwise, if one is just working with a desired cross-moment matrix $M$, the successive rows of $A$ can be fit via Newton-Raphson\cite{2011arXiv1111.0576S}.
-This is performed successively on each row of $A$.
+    Otherwise, if one is just working with a desired cross-moment matrix $M$, the successive rows of $A$ can be fit via Newton-Raphson.\cite{2011arXiv1111.0576S} This is performed successively on each row of $A$.
 If the first $i-1$ rows have been fit, then the upper left $(i-1)\times(i-1)$ sub matrix $A_{-i}$ of $A$ has already been filled.
@@
         \[ J\left( \mathbf a_i\right) = \sum_{\xi_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L\left(\mathbf \xi_{-i}^TA_{-i}\xi_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf \xi_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \xi_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \xi_{-i}^T & 1 \end{array}\right] \]
-            With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive. Instead, with K iid draws $\xi_{-i}^{(k)}\sim \mathbf x_{-i}$, the simulated values 
+            With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive.
+Instead, with K i.i.d.
+draws $\xi_{-i}^{(k)}\sim \mathbf x_{-i}$, the simulated values
             \[f\left(\mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L\left(\mathbf{a}_i^T\left[\begin{array}{c} \xi_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \xi_{-i}^{(k)} \\ 1 \end{array}\right] \]
@@
 In the simulation method in particular, if $K$ is not large enough, the equations can be very difficult or impossible to solve.
-When the initial problem can't be solved, besides increasing the simulation size or numerical accuracy, one can instead solve the relaxed problem  $f(\mathbf a)=\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
+When the initial problem can't be solved, besides increasing the simulation size or numerical precision, one can instead solve the relaxed problem  $f(\mathbf a)=\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
     \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T \]
@@
 The second half of the matrix still must be fit via Newton-Raphson.
-            \item One can avoid fitting the first half of the $A$ matrix entirely though by using the actual observed values in $X$ to simulate the first half of $\mathbf x_{aug}$, if performing Newton-Raphoson.
+            \item One can avoid fitting the first half of the $A$ matrix entirely though by using the actual observed values in $X$ to simulate values $[\xi_1\ldots \xi_p]\sim \mathbf x_{aug}$ if performing the Newton-Raphson method.
 This can be done by either bootstrapping rows from $X$ or using some number of copies of $X$ exactly.
@@
 Anderson mixing is one of several related quasi-Newtonian methods, which performs Newton-Raphson iteration $k$ with an approximation of the inverted inverted Jacobian labeled $G_k$.
-Thus, to solve for a variable $\mathbf z$ such that $f(\mathbf z)=\mathbf 0$, the best approximation for the root is updated as $\mathbf z_{k+1}=\mathbf z_k - G_k f(\mathbf{z}_k)$ at each iteration.
+Thus, to solve for a variable $\mathbf z$ such that $f(\mathbf z)=\mathbf 0$, the best approximation for the root is updated as $\mathbf{z}_{k+1}=\mathbf{z}_k - G_k f(\mathbf{z}_k)$ at each iteration.
 \par
@@
 These values are $f(\mathbf z_{k})$ through $f(\mathbf z_{k-m+1})$.
+This is analogous to estimating the slope using a secant line, except in a high dimension space.
 The best approximation for $f(\mathbf z)=0$ in the affine space which spans them is defined by
@@
         In practice, it seemed that combining the first, third, and fourth of these variations leads to the most successful technique.
-However, should $n$ be large enough to provide a good approximation of the probability space, the second variation also saves a lot of time.
+However, should $n$ be large enough to provide a good approximation of the probability space, the second variation also saves a lot of time.\footnote{An implementation of this method in python can be found on my github: \url{https://github.com/ajmaurer/Chicago-Course-Work/blob/master/masters_paper/knockoffGLM.py}}
 
 \section{Binary Knockoff Performance}
-    While the theory for original knockoffs and random binary variable generation suggests at how the binary knockoffs should perform, seeing their performance in practice still provides invaluable insight to their uses and limitations.
+    While the theory for original knockoffs and random binary variable generation suggests at how the binary knockoffs should behave, seeing their actual performance still provides invaluable insight to their uses and limitations.
 There are three major areas in which this is useful; how well do the binary knockoffs' empirical cross-moment matrices converge to their expectation, how well is FDR controlled for linear regression, and how well, if at all, is FDR controlled for other GLMs?
+\par
+    To answer each of these questions, I performed a number of different simulations of the random binary knockoffs.
+In each one, the rows of $X$ were drawn from the Ising model with a randomly generated $A$ matrix for the specified $p$.
+Where $z_{ij}\sim N(0,.2)$, $A_{ij}=z_{ij}$ if $i\neq j$, and $A_{ii}=z_{ii}-1$ otherwise.
+This leads to mild correlation among the columns of $X$.
+Then, for $i\in S$, $\beta_i\sim N(0,1)$, being otherwise set to $0$.
+For each observation for the linear regressions, $y\sim N(\mathbf{x}^T\beta,1)$, while for the logistic regressions, $y\sim \textrm{Bernoulli}(L(\mathbf{x}^T\beta))$.
+Given $X$, $\tilde X$ was generated based by the $\mu$-conditional method.
 
@@
     \FloatBarrier
-        In a certain sense, the question of the convergence of the binary knockoffs' empirical cross-moments is the least interesting of these questions.
+        In a certain sense, the question of the convergence of the binary knockoffs' empirical cross-moments is the least interesting of the given questions.
 Assuming the Ising model coefficient $A$ or the probability matrix $P$ was successfully fit, the cross-moment matrices' will converge to their expectation in probability for increasingly large values of $n$.
@@
     \end{center}
-    \caption{Measures how much far the empirical cross-moment matrix $\frac{1}{n}X_{aug}^TX_{aug}$ is from the desired cross moment matrix $M_{aug}$ for random binary knockoffs for various $n$ and $p$, averaged over $50$ simulations of $X$ from the Ising model.
-Where $W = \Sigma - \diag{\mathbf s} + \mathbf m \mathbf m^T$ is the top quadrant of $M_{aug}$, deviation is calculated as \\ $(\|X^T\tilde X - W\|_{fro} + \|\tilde X^T \tilde X - M\|_{fro})/(\|W\|_{fro}+\|M\|_{fro})$}
+    \caption{\linespread{1}\selectfont{}Measures how much far the empirical cross-moment matrix $\frac{1}{n}X_{aug}^TX_{aug}$ is from the desired cross moment matrix $M_{aug}$ for random binary knockoffs for various $n$ and $p$, averaged over $50$ simulations.
+Where $W = \Sigma - \diag{\mathbf s} + \mathbf m \mathbf m^T$ is the top quadrant of $M_{aug}$, deviation is calculated as $(\|X^T\tilde X - W\|_{fro} + \|\tilde X^T \tilde X - M\|_{fro})/(\|W\|_{fro}+\|M\|_{fro})$, where $\| \cdot \|_{fro}$ is the Frobenius norm.}
     \end{figure}
@@
 That it does so rather conservatively is a reflection of the relatively low number of total features compared to the sparsity.
-On the other hand, though mirroring the trend with the original knockoffs, the binary knockoffs fail to properly control FDR.
+On the other hand, though mirroring the trend with the original knockoffs, the binary knockoffs fail to exactly control FDR at $.2$.
 They select too many variables, which gives higher power at the expense of proper FDR control.
-As figure 5 shows, both methods make related, but no where near identical variable selections.
+As figure 5 shows, both methods make related, but non-identical, variable selections.
 \par
@@
     \end{center}
-    \caption{The mean FDR and power of binary and original knockoffs.
-Taken over $200$ simulations of linear regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ is generated from Ising model while $y$ is normally distributed about $X\beta$.}
+    \caption{\linespread{1}\selectfont{}The mean FDR and power of binary and original knockoffs.
+Taken over $200$ simulations of linear regression on $n=1000$, $p=50$, and targeting FDR $q=.2$ for varying levels of sparsity.}
     \end{figure}
@@
     \end{center}
-    \caption{The mean correlation between variables selected using binary and original knockoffs.
-Taken over $200$ simulations of linear regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ is generated from Ising model while $y$ is normally distributed about $X\beta$.}
+    \caption{\linespread{1}\selectfont{}Mean correlation between $z_i=I_{\hat S}(i)$ for binary knockoffs and for original knockoffs.
+Taken over $200$ simulations of linear regression on $n=1000$, $p=50$, and targeting FDR $q=.2$ for varying levels of sparsity.}
     \end{figure}
@@
 Thus, for any given value $n$, the empirical correlation of the combined design matrix $X_{aug}$ won't exactly be $\Sigma_{aug}$, and the method won't work perfectly.
-However, even here, the actual FDR is only somewhat hire than the desired $.2$, and for larger values of $n$, the empirical correlation will get closer to $\Sigma_{aug}$, and the FDR will be controlled better.
+However, even here, the actual FDR is only somewhat higher than the desired $.2$, and for larger values of $n$, the empirical correlation will get closer to $\Sigma_{aug}$, and the FDR will be controlled better.
 
@@
     In logistic regression, the binary knockoffs do seem to provide an improvement over the original knockoffs, as can be seen in figure 6.
-Though neither control FDR properly, using binary knockoffs result both in higher power and lower FDR.
+Though neither control FDR exactly at $.2$, using binary knockoffs result both in higher power and lower FDR.
 As well, while the original knockoffs do considerably worse in logistic regression than in linear regression, the binary knockoffs perform only a little worse, having similar FDR and somewhat lower power.
@@
     \end{center}
-    \caption{The mean FDR and power of binary and original knockoffs.
-Taken over $200$ simulations of logistic regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ is generated from Ising model while $y$ is Bernoulli with $p=L(X\beta)$.}
+    \caption{\linespread{1}\selectfont{}The mean FDR and power of binary and original knockoffs.
+Taken over $200$ simulations of logistic regression on $n=1000$, $p=50$, and targeting FDR $q=.2$ for varying levels of sparsity.}
     \end{figure}
@@
     \end{center}
-    \caption{The mean correlation between variables selected using binary and original knockoffs.
-Taken over $200$ simulations of logistic regression on $n=1000$, $p=50$, targeted FDR $q=.2$, and varying sparsity, or number of true predictors.
-$X$ is generated from Ising model while $y$ is Bernoulli with $p=L(X\beta)$.}
+    \caption{\linespread{1}\selectfont{}Mean correlation between $z_i=I_{\hat S}(i)$ for binary knockoffs and for original knockoffs.
+Taken over $200$ simulations of logistic regression on $n=1000$, $p=50$, and targeting FDR $q=.2$ for varying levels of sparsity.}
     \end{figure}
@@
 \section{Discussion}
-    Binary knockoffs, as a practical tool for statistical inference, are still a work in progress.
-Currently, they provide approximate FDR control for LASSO and seem like they also provide approximate FDR control for logistic regression.
-However, there are no theoretical guarantees for logistic regression, and the quality of FDR control for LASSO depends on how far the realized cross-moment matrix deviates from the desired $M_{aug}$.
-In cases where $n$ is large, this should only be a small issue, but for smaller $n$ FDR control fails in a meaningful way.
+    This paper has developed a method, along with several variations, of creating random knockoffs for binary data.
+These knockoffs serve as a replacement in the knockoff filter for the original style knockoffs; they serve the same purpose, but rather than meet the correlation condition exactly, they meet it in expectation.
+Since they are generated randomly so that they satisfy the correlation condition in expectation only, we see that in simulation they only provide approximate FDR control for linear regression.
+However, since they imitate the marginal distribution of the original $X_i$, they provide much better FDR control in logistic regression than the original style knockoff.
+\par
+    As a practical tool for statistical inference, they are still a work in progress.
+They do seem to provide approximate FDR control in logistic regression.
+However, there are no theoretical guarantees for this to be true, and the quality of FDR control for LASSO depends on how far the realized cross-moment matrix deviates from the desired $M_{aug}$.
+In cases where $n$ is large, this should only be a small issue, but for smaller $n$, FDR control fails in a meaningful way.
 As well, computational issues make binary knockoffs hard to generate for large $p$; as $p$ increases, the Ising model to generate the knockoffs gets rapidly more computationally intensive to fit, as do the methods based on quadratic programs.
 This leaves the method, as currently developed, only suitable for relatively high values of $n$ and low values of $p$.
-Considering FDR control for variable selection is most important for data sets with large $p$, this is not terribly satisfying.
 \par
@@
 A finally variable selection $S$ could be made based on theses, perhaps only selecting the variables selected by the majority of the $S_m$.
-This, or some other sort of average across multiple realizations of $\tilde X$, could likely be used to dampen the effect of variance in the generation of $\tilde X$.
-This would at least make binary knockoffs reliable for LASSO.
+This, or some other sort of average across multiple realizations of $\tilde X$, could likely be used to dampen the effect of variance in the generation of $\tilde X$ so as to make binary knockoffs more reliable for both LASSO and other GLMs.
 \par
@@
 Some of this can likely be mitigated by improvements in implementation, and the most time intensive issue, calculating $f(\mathbf a)$, can be easily parallelized.
-A less computationally intensive method to derive either $A$ or $P$ is necessary to make $p$ in the hundreds or thousands feasible.
+Still, a less computationally intensive method to derive either $A$ or $P$ is necessary to make $p$ in the hundreds or thousands feasible.
 A method based on optimizing each column of $P$ individually is thus appealing.
 So far, only approximate solutions have resulted from this method.
-However, in similar situations methods, such as the EM algorithm, have been developed which have strong guarantees for global optimization using such piecewise optimization.
+However, in similar situations, other methods, such as the EM algorithm, have been developed which have strong guarantees for global optimization using such piecewise optimization.
 If a similar development could be made here, or merely a method that hits the desired cross-moments more closely in practice, the result would be far more useful tool.
 \par
-    The final area of concern is how binary knockoffs will occur when the $X$ matrix is not fit well by the Ising model due to significant high order interactions.
+    The final area of concern is how binary knockoffs will perform when the $X$ matrix is not fit well by the Ising model due to significant high order interactions.
 Hopefully, having matched the first and second moments of $X$, binary knockoffs will continue to provide reasonable FDR control.
-An avenue to deal with this though would be to explicitly model these higher order interactions.
-One might introduce higher order moments into the fitting of $\mu$-conditionals.
-For instance, if $A$ is a three dimensional matrix whose entry $a_{ijk}$ would be added to the linear predictor when $x_i=x_j=x_k=1$.
-Just as the $\mu$-conditionals are based on a regression of $X_i$ on $X_{1:i-1}$ with no interactions, this would be akin to adding two way interactions.
+One avenue to ensure this would be to explicitly model these higher order interactions.
+We might introduce higher order moments into the fitting of $\mu$-conditionals.
+For instance, $A$ could be a three dimensional matrix whose entry $a_{ijk}$ would be added to the linear predictor when $x_i=x_j=x_k=1$.
+Just as the $\mu$-conditionals are based on a regression of $X_i$ on $X_{1:i-1}$ with no interactions, this would be akin to adding two-way interactions.
 This can be used to replicate arbitrary binary vectors to an arbitrary desired level of precision.
-However, in practice this has big limitations.
+However, this has big limitations.
 For one, adding additional levels of interactions will make the computation much more difficult very quickly.

### 2015-07-31 masters_paper/paper.tex
Final draft for department

@@
 Variable selection is an essential problem to fitting a regression model.
-For linear regression, the knockoff filter\cite{2014arXiv1404.5609F} offers a method of exact FDR control.
+For linear regression, the knockoff filter \cite{2014arXiv1404.5609F} offers a method of exact FDR control.
 However, the method as originally formulated does not extend to other generalized linear models (GLMs) such as logistic regression.
@@
 I will to refer the $j$th column of $X$ as $X_j$, which is $n$ observations of the random variable $x_j$.
-The random variable $x_j$ is in turn $j$th entry of the random vector valued variable $\mathbf x$.
+The random variable $x_j$ is in turn the $j$th entry of the random vector valued variable $\mathbf x$.
 Let $\beta$ be sparse, implying only a subset of the features $x_j$ have an effect on  $y$.
@@
     for some vector $\mathbf{s}\in \R^p$ such that, writing $\diag{G}$ as the vector made from the main diagonal of $G$, $\diag{G} - \mathbf s$ has small magnitude.
-Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$.
+Since $\tilde X_j$ and $X_j$ have relatively low correlation, as long as $\tilde X_j$ is created independently of $\mathbf y$, $\tilde X_j$ will have lower correlation with $\mathbf y$ than $X_j$ if $X_j$ is a true predictor.
 In the case where $X_j$ is a null predictor of $\mathbf y$, $\tilde X_j$ will thus also be a null predictor of $\mathbf y$.
@@
             \end{cases} \]
-            $W_j$ is positive when $Z_j$ entered the path first and negative when $\tilde X_j$ entered the path first.
-Large, positive values for $W_j$ will indicate that $X_j$ is a strong predictor, since $X_j$ entered the path early and before $\tilde X_j$.
+            $W_j$ is positive when $X_j$ enters the path first and negative when $\tilde X_j$ enters the path first.
+Large, positive values for $W_j$ will indicate that $X_j$ is a strong predictor, since $X_j$ enters the path early and before $\tilde X_j$.
 Thus, we will make a selection $\hat S=\{j:W_j\geq T\}$ for some threshold $T$.
@@
             The quantity on the left is, of course, the number of false discoveries made by choosing a particular threshold $T$. Also, it must be the case
-            \[\vert \{j:W_j\leq -T \; \& \; j\not\in S\}\vert \leq \vert\{j:W_j\leq -T\}\vert\]
+            \[\vert \{j:W_j\leq -T \; \& \; j\not\in S\}\vert \leq \vert\{j:W_j\leq -T\}\vert.\]
             This inequality should be pretty tight, since a true predictor should be a better predictor than its knockoff and unlikely to enter the path after it.
 Accordingly, the expected portion of false discoveries out of all discoveries is approximately
-            \[FDR \approx \E\left[\frac{\vert\{j:W_j\leq -T\}\vert}{\max\{\vert\{j:W_j\geq T\}\vert,1\}}\right]\] 
+            \[FDR \approx \E\left[\frac{\vert\{j:W_j\leq -T\}\vert}{\max\{\vert\{j:W_j\geq T\}\vert,1\}}\right].\] 
             So, when $T$ is selected as
@@
     For both methods, the first step is to normalize the matrix $X$ such that $X_j^TX_j=1$ for all $j$. The Gram matrix of $X_{aug}$ is
-    \[ G_{aug}:= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right] \]
-    $A$, the Schur complement of $G$ in $G_{aug}$ can be calculated as 
-    \[ A = 2\,\diag{\mathbf s} - \diag{\mathbf s}G^{-1}\diag{\mathbf s} \]
+    \[ G_{aug}:= [X\; \tilde X]^T[X\; \tilde X] = \left[ \begin{array}{cc} G & G - \diag{\mathbf s} \\ G - \diag{\mathbf s} & G \end{array}\right]. \]
+    $A$, the Schur complement of $G$ in $G_{aug}$, can be calculated as 
+    \[ A = 2\,\diag{\mathbf s} - \diag{\mathbf s}G^{-1}\diag{\mathbf s}. \]
     For $G_{aug}$ to exist, $G_{aug}$ must be positive semi-definite, which happens if and only if $A$ is positive semi-definite, which happens in turn if and only if 
-    \[ \diag{\mathbf s} \succeq 0  \; \textrm{ and } \; 2G - \diag{\mathbf s} \succeq 0 \]
+    \[ \diag{\mathbf s} \succeq 0  \; \textrm{ and } \; 2G - \diag{\mathbf s} \succeq 0. \]
     Given this is true, $A$ can be factored as $A=C^TC$.
 Combining this with a satisfactory $\mathbf s$ and an orthonormal matrix $\tilde U\in\R^{n\times p}$ such that $\tilde U^T X = 0$, a $\tilde X$ fulfilling the desired properties can be calculated as
-    \[ \tilde X = X(I-G^{-1}\,\diag{\mathbf s}) + \tilde UC\]
+    \[ \tilde X = X(I-G^{-1}\,\diag{\mathbf s}) + \tilde UC.\]
     As mentioned above, $\mathbf s$ should be chosen so as to make $\diag{G}-\mathbf s$ small.
@@
     \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=\Sigma \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = \Sigma - \diag{\mathbf s}\]
-    Where $\Sigma$ is the covariance matrix of $X$.
+    where $\Sigma$ is the covariance matrix of $X$.
 This is essentially a relaxation of the original correlation condition, requiring the desired covariance to only hold in expectation.
@@
     The $Z_j$ and $\tilde Z_j$ are then, once again, the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before.
-The problem with this is that if the given GLM doesn't have the same sufficient statistics as linear regression, there is no theoretical guarantee that the knockoff filter will control FDR.
-This is since, for null $X_j$, there is no theory to guarantee the $Z_j$ and $\tilde Z_j$ are exchangeable, and thus $W_j$ is not guaranteed to be symmetrically distributed around $0$.
+The is that if the given GLM doesn't have the same sufficient statistics as linear regression, there is no theoretical guarantee that the knockoff filter will control FDR.
+This is since, for null $X_j$, there is no guarantee the $Z_j$ and $\tilde Z_j$ are exchangeable, and thus $W_j$ is not guaranteed to be symmetrically distributed around $0$ for null features $X_j$.
 \par
@@
 For the method to work, for the null predictors, this should be a consistent $50\%$, indicative of the desired symmetric distribution.
-This is particularly important for high values, which play the largest roll in selecting a threshold.
+This is particularly important for high values of $|W_j|$, which play the largest roll in selecting a threshold.
 In the top left corner, under the ``global null", where all features are null and normally distributed, we almost achieve this.
@@
 
-    This gets much worse when some true predictors are included, as shown in the bottom row of the figure.
+    This gets much worse when some true predictors are included in the model, as shown in the bottom row of the figure.
 For both normal and binary predictors, the originals come in first far more often in the beginning.
@@
 \par
-    Upon examination, the issue seems to be that the original style knockoff features have a very different empirical distribution than the features they are imitating.
+    Upon examination, one issue seems to be that the original style knockoff features have a very different empirical distribution than the features they are imitating.
 This is obviously true with the binary features, but even with the normal features, the distributions differ substantially.
@@
 Mainly, $\tilde X\st X$ should be a random binary matrix such that
-        \[ \E(X^T \tilde X \st X) = X^TX - \diag{s} \quad \& \quad \E(\tilde X^T \tilde X \st X) = X^TX\]
+    \[ \E[X^T \tilde X \st X] = X^TX - \diag{\mathbf s} \quad \& \quad \E[\tilde X^T \tilde X \st X] = X^TX.\]
 
@@
 Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag{M}$.
-For an arbitrary symmetric $M$ to be a valid cross-moment matrix for some random binary vector variable, $M-mm^T\succeq0$, and
+For an arbitrary symmetric $M$ to be a valid cross-moment matrix for some random binary vector variable, the necessary and sufficient conditions are that $M-mm^T\succeq0$ and
     \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
-    for all $i\neq j$.\cite{2011arXiv1111.0576S}
+    for all $i\neq j$ \cite{2011arXiv1111.0576S}.
 
@@
 Much as before with the original knockoffs, both the equi-correlated and SDP method can be used to select  $\mathbf s$  such that $\diag{\Sigma}-\mathbf s$ has small magnitude and
-        \[ \Sigma_{aug} = \left[\begin{array}{cc}  \Sigma & \Sigma - \diag{\mathbf s }\\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \succeq 0 \]
+        \[ \Sigma_{aug} = \left[\begin{array}{cc}  \Sigma & \Sigma - \diag{\mathbf s }\\ \Sigma - \diag{\mathbf s} & \Sigma \end{array}\right] \succeq 0. \]
     While with the original knockoffs, the result was small values for $X_i^T\tilde X_i$, now the result is low covariance between $X_i$ and $\tilde X_i$.\par
@@
     then the desired cross-moment matrix of the joint distribution is 
-        \[ M_{aug} = \Sigma_{aug} + \mathbf m_{aug}\mathbf m_{aug}^T\]
+        \[ M_{aug} = \Sigma_{aug} + \mathbf m_{aug}\mathbf m_{aug}^T.\]
         As mentioned above, for this cross-moment matrix to correspond to an actual random binary vector, it must be the case  
-        \[\max\{0,(m_{aug})_i + ( m_{aug})_j -1\} \leq (\Sigma_{aug})_{ij} +  (m_{aug})_i  (m_{aug})_j \leq \min\{( m_{aug})_i, ( m_{aug})_j\} \]
+        \[\max\{0,(m_{aug})_i + ( m_{aug})_j -1\} \leq (\Sigma_{aug})_{ij} +  (m_{aug})_i  (m_{aug})_j \leq \min\{( m_{aug})_i, ( m_{aug})_j\}. \]
     However, since $\Sigma$ and $\mathbf m$ are from an actual random binary vector, the entries of $\Sigma$ satisfy this property by default.
@@
                 \begin{tabular}{r l}
-                    minimize     & $\|X^TP-(M-\diag{s})\|_{fro}^2 + \sum_{i\neq j}(P_i^T P_j - M_{ij})^2 $\\
+                    minimize     & $\|X^TP-(M-\diag{\mathbf{s}})\|_{fro}^2 + \sum_{i\neq j}(P_i^T P_j - M_{ij})^2 $\\
                     subject to   & $ \mathbf 1^T P = \mathbf m $ \\
@@
             so that the cross-moments are as close as possible to $M_{aug}$.
-The first part of the objective corresponds to the deviation of $\E[X^T\tilde X]$ from $M-\diag{s}$ and the second part to the deviation of $\E[\tilde X^T \tilde X]$ from $M$ excluding the diagonal values which are matched by $ \mathbf 1^T P = \mathbf m $.
+The first part of the objective corresponds to the deviation of $\E[X^T\tilde X]$ from $M-\diag{\mathbf{s}}$ and the second part to the deviation of $\E[\tilde X^T \tilde X]$ from $M$ excluding the diagonal values which are matched by $ \mathbf 1^T P = \mathbf m $.
 This problem can be stated as a quadratic program via the introduction of slack variables:
@@
                     minimize     & $\|W\|_{fro}^2 + \|V\|_{fro}^2 $ \\
-                    subject to   & $ -W \leq X^TP-(M-\diag{s})\leq W $ \\
+                    subject to   & $ -W \leq X^TP-(M-\diag{\mathbf{s}})\leq W $ \\
                                  & $ -V_{ij} \leq P_i^T P_j - M_{ij} \leq V_{ij} \quad \forall i\neq j$ \\
@@
 However, this is only guaranteed to be feasible in the bivariate case.
-For higher dimensions, this method can generally only provide random binary vectors with approximately the right moments, and the quality of the approximation quickly degenerates as $p$ increases.\cite{2011arXiv1111.0576S} Therefore, we do not explore this direction further.
+For higher dimensions, this method can generally only provide random binary vectors with approximately the right moments, and the quality of the approximation quickly degenerates as $p$ increases \cite{2011arXiv1111.0576S}.
+Therefore, we do not explore this direction further.
 
 \subsubsection{$\mu$-Conditionals family}
-There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals.\cite{2011arXiv1111.0576S} The basic idea is that the entries of a binary vector variable $\mathbf x$ can be sampled sequentially as
+There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals \cite{2011arXiv1111.0576S}.
+The basic idea is that the entries of a binary vector variable $\mathbf x$ can be sampled sequentially as
     \[ \p(x_{i}=1\st x_{1},...,x_{i-1}) = \mu\left(a_{ii}+\sum_{k=1}^{i-1}a_{ik}x_i\right) \]
@@
 \par
-    Otherwise, if one is just working with a desired cross-moment matrix $M$, the successive rows of $A$ can be fit via Newton-Raphson.\cite{2011arXiv1111.0576S} This is performed successively on each row of $A$.
+    Otherwise, if one is just working with a desired cross-moment matrix $M$, the successive rows of $A$ can be fit via Newton-Raphson \cite{2011arXiv1111.0576S}.
+This is performed successively on each row of $A$.
 If the first $i-1$ rows have been fit, then the upper left $(i-1)\times(i-1)$ sub matrix $A_{-i}$ of $A$ has already been filled.
@@
             where $J$ is the Jacobian of $f$. To do so, the Jacobian matrix is calculated as 
-        \[ J\left( \mathbf a_i\right) = \sum_{\xi_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L\left(\mathbf \xi_{-i}^TA_{-i}\xi_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf \xi_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \xi_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \xi_{-i}^T & 1 \end{array}\right] \]
+        \[ J\left( \mathbf a_i\right) = \sum_{\xi_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L\left(\mathbf \xi_{-i}^TA_{-i}\xi_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf \xi_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \xi_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \xi_{-i}^T & 1 \end{array}\right].
+\]
             With $2^{i-1}$ possible values for $\mathbf x_{-i}$, this can quickly become computationally expensive.
@@
 In the simulation method in particular, if $K$ is not large enough, the equations can be very difficult or impossible to solve.
-When the initial problem can't be solved, besides increasing the simulation size or numerical precision, one can instead solve the relaxed problem  $f(\mathbf a)=\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$
-    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T \]
+When the initial problem can't be solved, besides increasing the simulation size or numerical precision, one can instead solve the relaxed problem  $f(\mathbf a)=\mathbf m_i^*(\tau)$, where, for $\tau\in[0,1]$,
+    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T .\]
             When $\tau=0$, this yields the original problem, while when $\tau=1$, it is treats $x_i$ as independent of $\mathbf x_{-i}$. The latter will always have the solution
-            \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{M_{ii}}{1-M_{ii}}\right) \end{array} \right]^T \]
+            \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{M_{ii}}{1-M_{ii}}\right) \end{array} \right]^T.\]
             The hope is that for some $\tau$ close to $0$, convergence can be achieved, only causing a slight distortion from the desired cross-moments.
@@
 This is since it has $O(Kp^2)$ complexity compared to $O(Kp)$ complexity for $f(\mathbf a)$, which quickly becomes prohibitive.
-One might instead estimate the full Jacobian based on first differences, but in practice an even simpler method called Anderson mixing\cite{Fang07twoclasses}, which avoids additional computation of $f(\mathbf a)$, works well.
+One might instead estimate the full Jacobian based on first differences, but in practice an even simpler method called Anderson mixing \cite{Fang07twoclasses}, which avoids additional computation of $f(\mathbf a)$, works well.
 Anderson mixing is one of several related quasi-Newtonian methods, which performs Newton-Raphson iteration $k$ with an approximation of the inverted inverted Jacobian labeled $G_k$.
@@
     While the theory for original knockoffs and random binary variable generation suggests at how the binary knockoffs should behave, seeing their actual performance still provides invaluable insight to their uses and limitations.
-There are three major areas in which this is useful; how well do the binary knockoffs' empirical cross-moment matrices converge to their expectation, how well is FDR controlled for linear regression, and how well, if at all, is FDR controlled for other GLMs?
+There are three major areas in which this is useful: how well do the binary knockoffs' empirical cross-moment matrices converge to their expectation, how well is FDR controlled for linear regression, and how well, if at all, is FDR controlled for other GLMs?
 \par
@@
     When the knockoff filter is used in linear regression, we see that the original knockoffs are still superior to the binary knockoffs.
-As figure 4 shows, while they both fail to control the FDR when there are only a few true predictors\footnote{This issue is addressed in the original paper; a slightly more conservative filter called knockoff+ avoids this problem}, the original knockoffs properly control FDR for more valid predictors.
+As figure 4 shows, while they both fail to control the FDR when there are only a few true predictors\footnote{This issue is addressed in the original paper; a slightly more conservative filter called knockoff+ avoids this problem}, the original knockoffs properly control FDR when there is a larger number of valid predictors.
 That it does so rather conservatively is a reflection of the relatively low number of total features compared to the sparsity.
-On the other hand, though mirroring the trend with the original knockoffs, the binary knockoffs fail to exactly control FDR at $.2$.
+On the other hand, though mirroring the trend with the original knockoffs, the binary knockoffs fail to exactly control FDR at level $q=.2$.
 They select too many variables, which gives higher power at the expense of proper FDR control.
@@
 These knockoffs serve as a replacement in the knockoff filter for the original style knockoffs; they serve the same purpose, but rather than meet the correlation condition exactly, they meet it in expectation.
-Since they are generated randomly so that they satisfy the correlation condition in expectation only, we see that in simulation they only provide approximate FDR control for linear regression.
+Since they are generated randomly so that they satisfy the correlation condition in expectation only, we see that in simulations they only provide approximate FDR control for linear regression.
 However, since they imitate the marginal distribution of the original $X_i$, they provide much better FDR control in logistic regression than the original style knockoff.
